### 2021-02-26 tex/imp.tex
Hello, Tex!
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

%! Author = aybehrouz
%! Date = 2/26/21

% Preamble
\documentclass{article}

% Packages
%\usepackage{amsmath}
%\usepackage{amssymb}

% Document
\begin{document}
    \section{Consensus}\label{sec:consensus}

    \subsection{Estimating A User's Stake}\label{subsec:estimating-a-user's-stake}

    In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
    of stake the user has in the system.
    Conventionally in these systems, for estimating a user's stake, we use the amount of native system tokens the
    user is holding.
    Unfortunately, one problem with this approach is that a strong attacker may be able to obtain a considerable
    amount of system tokens, for example by borrowing from a DEFI application, and use this stake to attack the system.

    To mitigate this problem, for calculating a user's stake, instead of using the raw ALGO balance, we use the
    minimum of a \emph{trust value} that the system has calculated for the user and the user's ALGO balance:
    \[
        Stake_{user} = \min (Balance_{user},Trust_{user})
    \]
    For estimating the value of $Trust_{user}$ we use the exponential moving average of the user's ALGO balance.
    Therefore, in our system a user who held ALGOs and participated in the consensus for a long time is more trusted
    than a new user with a higher balance.
    An attacker who has obtained a large amount of ALGOs, also needs to hold them for a long period of time before
    being able to attack our system.

    For calculating the exponential moving average of a time series at the time step $t$, we can use the following
    recursive formula:
    \[
        M_{t} = (1 - \alpha) M_{t - 1} + \alpha X_{t} = M_{t - 1} + \alpha (X_{t} - M_{t - 1})
    \]
    Where:
    \begin{itemize}
        \item The coefficient $\alpha$ is a constant smoothing factor between 0 and 1 which represents the degree of
        weighting decrease, A higher $\alpha$ discounts older observations faster.
        \item $X_{t}$ is the value of the time series at the time step $t$.
        \item $M_{t}$ is the value of the EMA at the time step $t$.
    \end{itemize}

    Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
    $M_{t}$:
    \[
        M_{t} = (1 - \alpha)^{t - k}M_{k} + [1 - (1 - \alpha)^{t - k}]X
    \]
    Where:
    \[
        X = X_{k+1} = X_{k+2} = \dots = X_t
    \]
    When $|nx| \ll 1$ we can use the binomial approximation  $(1 + x)^n \approx 1 + nx$ to further simplify this
    formula:
    \[
        M_{t} = M_{k} + (t - k) \alpha (X - M_{k})
    \]

    For choosing the value of $\alpha$ we can consider the number of time steps that the trust value of a user needs
    for reaching a specified fraction of his account balance.
    We know that for large $n$ and $|x| < 1$ we have $(1 + x)^{n} \approx e^{nx}$, so by letting $k = 0$ and $t = n$
    we can write:
    \[
        \alpha =- \frac{\ln\left(1 - \frac{M_{n}}{X}\right)}{n}
    \]
    The value of $\alpha$ for a desired configuration can be calculated by this equation.
    For instance, we could calculate the $\alpha$ for a relatively good configuration in which $M_{n} = 0.8X$ and $n$
    equals to the number of time steps of 10 years.
\end{document}

### 2021-02-26 tex/imp.tex
Clarify the last equation
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
     for reaching a specified fraction of his account balance.
-    We know that for large $n$ and $|x| < 1$ we have $(1 + x)^{n} \approx e^{nx}$, so by letting $k = 0$ and $t = n$
+    We know that for large $n$ and $|x| < 1$ we have $(1 + x)^{n} \approx e^{nx}$, so by letting $M_{k} = 0$ and
+    $n = t - k$
     we can write:
     \[
-        \alpha =- \frac{\ln\left(1 - \frac{M_{n}}{X}\right)}{n}
+        \alpha =- \frac{\ln\left(1 - \frac{M_{n + k}}{X}\right)}{n}
     \]
     The value of $\alpha$ for a desired configuration can be calculated by this equation.
-    For instance, we could calculate the $\alpha$ for a relatively good configuration in which $M_{n} = 0.8X$ and $n$
+    For instance, we could calculate the $\alpha$ for a relatively good configuration in which $M_{n+k} = 0.8X$ and $n$
     equals to the number of time steps of 10 years.

### 2021-02-28 tex/con.tex
Add support for equation conversion from tex to md
The equations will be converted using url images in the gfm directory and by html tags in the md directory.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-02-26 tex/imp.tex
@@
 % Preamble
-\documentclass{article}
+\documentclass[11pt]{article}
 
 % Packages
+\usepackage{amssymb}
 %\usepackage{amsmath}
-%\usepackage{amssymb}
 
@@
     \]
-    For estimating the value of $Trust_{user}$ we use the exponential moving average of the user's ALGO balance.
+    For estimating the value of \(Trust_{user}\) we use the exponential moving average of the user's ALGO balance.
     Therefore, in our system a user who held ALGOs and participated in the consensus for a long time is more trusted
@@
 
-    For calculating the exponential moving average of a time series at the time step $t$, we can use the following
+    For calculating the exponential moving average of a time series at the time step \(t\), we can use the following
     recursive formula:
     \[
-        M_{t} = (1 - \alpha) M_{t - 1} + \alpha X_{t} = M_{t - 1} + \alpha (X_{t} - M_{t - 1})
+        M_t = (1 - \alpha) M_{t-1} + \alpha X_t = M_{t-1} + \alpha (X_t - M_{t-1})
     \]
@@
     \begin{itemize}
-        \item The coefficient $\alpha$ is a constant smoothing factor between 0 and 1 which represents the degree of
-        weighting decrease, A higher $\alpha$ discounts older observations faster.
-        \item $X_{t}$ is the value of the time series at the time step $t$.
-        \item $M_{t}$ is the value of the EMA at the time step $t$.
+        \item The coefficient \(\alpha\) is a constant smoothing factor between 0 and 1 which represents the degree of
+        weighting decrease, A higher \(\alpha\) discounts older observations faster.
+        \item \(X_t\) is the value of the time series at the time step \(t\).
+        \item \(M_t\) is the value of the EMA at the time step \(t\).
     \end{itemize}
@@
     Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
-    $M_{t}$:
+    \(M_t\):
     \[
-        M_{t} = (1 - \alpha)^{t - k}M_{k} + [1 - (1 - \alpha)^{t - k}]X
+        M_t = (1 - \alpha)^{t-k}M_k + [1 - (1 - \alpha)^{t - k}]X
     \]
@@
     \]
-    When $|nx| \ll 1$ we can use the binomial approximation  $(1 + x)^n \approx 1 + nx$ to further simplify this
+    When \(|nx| \ll 1\) we can use the binomial approximation  \((1 + x)^n \approx 1 + nx\) to further simplify this
     formula:
     \[
-        M_{t} = M_{k} + (t - k) \alpha (X - M_{k})
+        M_t = M_k + (t - k) \alpha (X - M_k)
     \]
 
-    For choosing the value of $\alpha$ we can consider the number of time steps that the trust value of a user needs
+    For choosing the value of \(\alpha\) we can consider the number of time steps that the trust value of a user needs
     for reaching a specified fraction of his account balance.
-    We know that for large $n$ and $|x| < 1$ we have $(1 + x)^{n} \approx e^{nx}$, so by letting $k = 0$ and $t = n$
+    We know that for large \(n\) and \(|x| < 1\) we have \((1 + x)^n \approx e^{nx}\), so by letting \(M_k = 0\) and
+    \(n = t - k\)
     we can write:
     \[
-        \alpha =- \frac{\ln\left(1 - \frac{M_{n}}{X}\right)}{n}
+        \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{X}\right)}{n}
     \]
-    The value of $\alpha$ for a desired configuration can be calculated by this equation.
-    For instance, we could calculate the $\alpha$ for a relatively good configuration in which $M_{n} = 0.8X$ and $n$
-    equals to the number of time steps of 10 years.
+    The value of \(\alpha\) for a desired configuration can be calculated by this equation.
+    For instance, we could calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8X\)
+    and \(n\) equals to the number of time steps of 10 years.
 \end{document}

### 2021-02-28 tex/specs.tex
Add support for equation conversion from tex to md
The equations will be converted using url images in the gfm directory and by html tags in the md directory.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

%! Author = aybehrouz
%! Date = 2/28/21

% Preamble
\documentclass[11pt]{article}

% Packages
%\usepackage{amsmath}

\title{The Algorand Virtual Machine Specification}
%\author{aybehrouz \thanks{funded by myself}}
%\date{2020}
% Document
\begin{document}
    \maketitle

    \section{Introduction}\label{sec:introduction}

    The Algorand Virtual Machine (AVM) is an abstract computing machine for executing Algorand's smart contracts.
    It is designed in a way that it could be efficiently implemented in either hardware or software.

    \section{The Structure of the Algorand Virtual Machine}\label{sec:the-structure-of-the-algorand-virtual-machine}

    ...

    \subsection{Data Types}\label{subsec:data-types}

    The Algorand Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
    and does
    not have to be done by the Algorand Virtual Machine itself.

    The instruction set of the Algorand Virtual Machine distinguishes its operand types using instructions intended to
    operate on values of specific types.
    For instance, \texttt{iadd} assumes that its operands are two 64-bit integers.

    \subsection{The \texttt{PC} Register}\label{subsec:the-pc-register}

    ...

    \subsection{Call Stack}\label{subsec:call-stack}

    A call stack contains the information that is needed for restoring the state of the invoker of a method.

    \subsection{Memory Unit}\label{subsec:memory-unit}

    The Algorand Virtual Machine has a byte addressable memory which is divided into separate segments.
    Every segment belongs to a smart contract that has a unique \texttt{applicationID}.
    The AVM always has a single working memory segment and memory locations outside its current working segment can
    not be accessed.
    The only instructions that can change the current working segment are \texttt{invoke\_external}, \texttt{athrow} and
    return instructions.

    Every smart contract has its own memory segment.
    Hence, there is no way for a smart contract to access another smart contract's memory.
    Interaction between smart contracts is done using \texttt{invoke\_external} instruction, and A smart contract can
    invoke methods of another smart contract by this instruction.

    A memory segment consists of five data areas:
    \begin{itemize}
        \item Code Area
        \item Constant Area
        \item Local Frame
        \item Operand Stack
        \item Heap
    \end{itemize}
    All data areas except operand stack, have their own address space which starts from 0.
    Operand stack is a last-in-first-out (LIFO) stack and is not addressable.
    Every memory access instruction operates on its specific data areas.

    \subsubsection{Code Area}

    The code area of a segment contains the byte-code of the smart contract which owns that segment.
    The AVM has no instructions for manipulating the code area.
    \textbf{Installing, removing and updating smart contracts need to be done externally.}

    \subsubsection{Constant Area}

    The constant area of a segment contains several kinds of constants, ranging from user defined constants to method
    address tables.
    A method address table stores method locations in the code area and their access type.
    The access type of a method can be either \texttt{public} or \texttt{private}.
    The AVM has no instructions for modifying the constant area.

    \emph{Only public methods can be invoked by \texttt{invoke\_external} instruction.}

    \subsubsection{Local Frame}

    A local frame is used to store methods parameters and local variables.
    A new frame is created each time a method is invoked.
    A frame is destroyed when its method invocation completes, whether the completion is normal or abrupt.

    \subsubsection{Operand Stack}

    Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too.
    AVM instructions take operands from the operand stack, operate on them, and push the result back onto the operand
    stack.
    An operand stack is destroyed when its owner method completes, whether that completion is normal or abrupt.

    \subsubsection{Heap}

    The heap of a segment is a persistent memory area which is divided into pages.
    Each page can be referenced by an index that starts from 0.
    Memory locations inside every page have a separate address space that starts from 0.
    In other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex,
        offset)}.
    Pages of a heap do not need to be equally sized.

    \emph{The reason behind this paged design is that the heap is usually persisted using a block device. A heap with
    a paged
    structure could expose the underlying block based nature of the persistence layer to the application layer. In
    this way, the compiler or the programmer could better optimize the code for the persistence layer.}

    \section{Instruction Set Summary}\label{subsec:instruction-set-summary}

    An Algorand Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
    performed,
    followed by zero or more operands supplying arguments or data that are used by the operation.
    \textbf{The number and size of the operands are determined solely by the opcode.}

    \subsection{Method Invocation}\label{subsec:method-invocation}

    The Algorand Virtual Machine has three types of method invocation:
    \begin{itemize}
        \item \texttt{invoke\_internal} invokes a method from the current running smart contract.
        \item \texttt{invoke\_external} invokes a \texttt{public} method from another smart contract.
        It will change the current memory segment to the segment of the invoked smart contract.
        \item \texttt{invoke\_native} invokes a method that is not hosted by the Algorand virtual machine.
        By this instruction, high performance native methods of a hosting machine could become available to AVM smart
        contracts.
    \end{itemize}

    \emph{In the future, we may need to add special instructions for invoking interface and virtual methods...}

    Each time a method is invoked a new local frame and operand stack is created.
    The Algorand Virtual Machine uses local frames to pass parameters on method invocation.
    On method invocation, any parameters are passed in consecutive local variables stored in the method's local frame
    starting from local variable 0.
    The invoker of a method writes the parameters in the local frame of the invoked method using \texttt{arg}
    instructions.

    \subsubsection{Exceptions}

    An exception is thrown programmatically using the \texttt{athrow} instruction.
    Exceptions can also be thrown by various Algorand Virtual Machine instructions if they detect an abnormal
    condition.
    Some exceptions are not catchable and will always abort the execution of the smart contract.

    \subsubsection{Method Invocation Completion}

    A method invocation completes normally if that invocation does not cause an exception to be thrown, either
    directly from the AVM or as a result of executing an explicit throw statement.
    If the invocation of the current method completes normally, then a value may be returned to the invoking method.
    This occurs when the invoked method executes one of the return instructions, the choice of which must be
    appropriate for the type of the value being returned (if any).
    Execution then continues normally in the invoking method's local frame with the returned value (if any)
    pushed onto the operand stack.

    A method invocation completes abruptly if an exceptions is thrown and is not caught by the current method.
    A method invocation that completes abruptly never returns a value to its invoker.

    When a method completes, whether normally or abruptly, the call stack is used to restore the state of the invoker,
    including its local frame and operand stack, with the \texttt{PC} register appropriately restored and incremented
    to skip past the method invocation instruction.
    If the invoker was another smart contract, i.e.\ the invocation was made by an \texttt{invoke\_external}
    instruction, the current memory segment will be changed to the invoker's segment.

    A thrown exception causes methods in the call stack to complete \textbf{abruptly} one by one, as long as the
    \texttt{PC} register is not pointing to a \texttt{catch} instruction.
    The \texttt{catch} instruction acts like a branch instruction that branches only if an exception is caught.
    \textbf{When an external method invocation completes abruptly, before changing the current segment, all changes
    made to the heap area will be rolled back.}

    \emph{By using the \texttt{athrow} instruction properly, a programmer can make any method act like an atomic
    operation.}

    \subsubsection{Authorizing Operations}

    In blockchain applications, we usually need to authorize certain operations.
    For example, for sending an asset from a user to another user, first we need to make sure that the sender has
    authorized this operation.
    The Algorand virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
    cryptographic instructions for validating signatures and cryptographic entities.
    By using these instructions and passing signatures as parameters to methods a programmer can implement the
    required logic for authorizing any operation.

    \emph{The Algorand virtual machine has no instructions for issuing cryptographic signatures.}

    In addition to signatures, a method can verify its invoker by using \texttt{get\_parent} instruction.
    This instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
    smart contract in the call stack.
    In other words, it gives the smart contract that has invoked the current smart contract. (if any)

    \subsubsection{Heap Allocation Instructions}

    ...

\end{document}

### 2021-03-02 tex/impl.tex
Complete TeX versions
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

%! Author = aybehrouz
%! Date = 2/28/21

% Preamble
\documentclass[11pt, A4]{article}

% Packages
%\usepackage{amsmath}

\usepackage{hyperref}

\title{The Algorand Virtual Machine Implementation}
\author{aybehrouz}
\date{January 2021}

% Document
\begin{document}
    \maketitle

    \section{Persistence}\label{sec:persistence}

    For implementing the persistence layer of the AVM, we assume that we have access to an updatable zero-knowledge
    elementary database (ZK-EDB) with the following properties:

    \begin{itemize}
        \item The ZK-EDB contains a mapping from a set of keys to a set of values.
        \item Every state of the database has a commitment \(C\).
        \item The ZK-EDB has a method \((D, p) = get(x)\), where \(x\) is a key and \(D\) is the associated value
        with \(x\) and \(p\) is a proof.
        \item A user can use \(C\) and \(p\) to verify that \(D\) is really associated with \(x\), and \(D\) is not
        altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB.
        \item Having \(p\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
        associated with \(x\) instead of \(D\).
    \end{itemize}

    We use a ZK-EDB for storing the AVM heap. We include the commitment of the current state of this DB in every
    block of the Algorand blockchain, so ZK-EDB servers need not be trusted servers.

    Every page of AVM heap will be stored with a key of the form: \texttt{applicationID|pageIndex} (the \texttt{|}
    operator concatenates two numbers). Nodes do not keep a full copy of the AVM heap and for validating block
    certificates or emulating the AVM ( i.e.~validating transactions) they need to connect to a ZK-EDB and retrieve
    the required pages of AVM heap. For better performance, nodes keep a cache of heap pages to
    reduce the amount of ZK-EDB access.

    We also use a ZK-EDB for storing the code area of each segment, and we include the commitment of this DB in every
    block. Every code area will be divided into blocks and every block will be stored in the DB with
    \texttt{applicationID|blockID} as its key. Like heap pages, nodes keep a cache of code area blocks.

    \emph{Unlike heap pages, the AVM is not aware of different blocks of the code area.}

    \section{Transactions}\label{sec:transactions}

    Algorand has four types of transaction:

    \begin{itemize}
        \item \texttt{avmCall} essentially is an \texttt{invoke\_external} instruction that invokes a method from an
        AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
        assets, including ALGOs, is done by these transactions.
        \item \texttt{installApp} installs an AVM smart contract and determines the update policy of the smart
        contract: if the contract is updatable or not, the accounts that can update or uninstall the contract, and so
        on.
        \item \texttt{unInstallApp} removes an AVM smart contract.
        \item \texttt{updateApp} updates an AVM smart contract.
    \end{itemize}

    All types of Algorand transactions contain an \texttt{invoke\_external} instruction which calls a special method
    from ALGO smart contract that transfers the proposed fee of the transaction in ALGOs from a sender account to the
    fee sink accounts.

    Every transaction is required to exactly specify what heap pages or code area blocks it will access. This enables
    validators to start retrieving the required memory blocks from available ZK-EDB servers as soon as they see a
    transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to access
    a memory block that is not included in its access lists, will be rejected. Users could
    use smart contract oracles to predict the list of memory blocks their transactions need. See
    Section~\ref{sec:smart-contract-oracle} for more details.

    \section{Blockchain}\label{sec:blockchain}

    Every block of the Algorand blockchain corresponds to a set of transactions. We store the commitment of this
    transaction set in every block, but we don't keep the set itself. To be able to detect replay attacks, we require
    every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
    and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
    must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
    a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound\ -\ issuance\
        \textgreater{}\ maxLifeTime} or if a signature of the same user with a bigger or equal nonce is already used
    (i.e
    .~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two nonces
    have an equal issuance, the nonce with the bigger sequence number will be considered bigger.

    To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. When the
    difference between \texttt{issuance} component of this nonce and the current round becomes bigger than the
    maximum allowed lifetime of a signature, this information can be safely deleted. \textbf{As a result, we will not
    have the problem of " unremovable empty accounts" like Ethereum.}

    The only information that Algorand nodes are required to store is \textbf{the most recent block} of the Algorand
    blockchain. Every block of the Algorand blockchain contains the following information:

    \begin{center}
        \begin{tabular}{||c||}
            \hline
            Block \\ [0.5ex]
            \hline\hline
            commitment to the ZK-EDB storing heap pages \\ [0.7ex]
            commitment to the ZK-EDB storing code areas \\ [0.7ex]
            commitment to the set of transactions       \\ [0.7ex]
            previous block hash                         \\ [0.7ex]
            random seed                                 \\ [0.7ex]
            \hline
        \end{tabular}
    \end{center}

    For confirming a new block, nodes that are not validators only need to verify the block certificate. For
    verifying a block certificate, a node needs to know the ALGO balances of validators, but it doesn't need to
    emulate the AVM execution.

    On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
    execution of the Algorand virtual machine. To do so, first they retrieve all heap pages and code area blocks they
    need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
    transactions included in the new block. This will modify some pages of the AVM memory, so they update the ZK-EDB
    commitments based on the modified pages and verify the commitments included in the new block. Validators also
    calculate and verify the commitment to the new block's transaction set.

    \emph{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDBs will receive the new
    block, and they will update their database by emulating the AVM execution.}

    \section{Incentive mechanism}\label{incentive-mechanism}

    \subsection{Transaction Fee}\label{subsec:transaction-fee}

    Every transaction in the Algorand blockchain starts with an \texttt{invoke\_external} instruction which calls a
    special method from ALGO smart contract. This method will transfer the proposed fee of the transaction in ALGOs
    from a sender account to the fee sink accounts. Algorand has two fee sink accounts: \texttt{execFeeSink} collects
    execution fees and \texttt{dbFeeSink} collects fees for ZK-EDBs. The Protocol decides how to distribute the
    transaction fee between these two fee sink accounts.

    When a block is added to the blockchain, the proposer of that block will receive a share of the block fees.
    Consequently, a block proposer is always incentivized to include more transactions in his block. However, if he
    puts too many transactions in his block and the validation of the block becomes too difficult, some validators
    may not be able to validate all transactions on time. If a validator can not validate a block in the required
    time, he will consider the block invalid. So, when a proposed block contains too many transactions, the network
    may reach consensus on another block, and the proposer of that block will not receive any fees. As a result, a
    proposer is incentivized to use network transaction capacity optimally.

    On the other hand, we believe that the proposer does not have enough incentives for optimizing the storage size
    of the transaction set. Therefore, we require that \textbf{the size of the transaction set of every block in
    bytes be lower than a certain threshold.}

    Validators need to spend resources for validating transactions. When a validator starts the emulation of the AVM
    to validate a transaction, solely from the code he can't predict the time the execution will finish. This will
    give an adversary an opportunity to attack the network by broadcasting transactions that never ends. Since,
    validators can not finish the execution of these transactions, the network will not be able to charge the
    attacker any fees, and he would be able to waste validators resources for free.

    To mitigate this problem, we require that every transaction specify a cap for all the resources it needs. This
    will include memory, network and processor related resources. Also, the protocol defines an execution cost for
    every AVM instruction reflecting the amount of resources its emulation needs. This will define a standard way for
    measuring the execution cost of any \texttt{avmCall} transaction. Every \texttt{avmCall} transaction is required
    to specify a maximum execution cost. If during emulation it reaches this maximum cost, the transaction will be
    considered failed and the network can receive the proposed fee of that transaction.

    Every \texttt{avmCall} transaction is required to provide the following information as an upper bound for the
    resources it needs:

    \begin{itemize}
        \item Execution cost
        \item A list of heap/code-area pages for reading
        \item A list of heap pages for writing
        \item A list of heap pages it will deallocate
        \item Number and size of heap pages it will allocate
    \end{itemize}

    If a transaction tries to violate any of these predefined limitations, for example, if it tries to read a memory
    location that is not included in its reading list, it will be considered failed and the network can receive the
    proposed fee of that transaction.

    \emph{A transaction always pays all of its proposed fee, no matter how much of its predefined resources were not
    used in the final emulation.}

    \subsection{ZK-EDB Servers}\label{subsec:zk-edb-servers}

    The incentive mechanism for ZK-EDB servers should have the following properties:

    \begin{itemize}
        \item It incentivizes storing all memory blocks, whether a heap page or a code area block, and not only those
        that are used more frequently.
        \item It incentivizes the ZK-EDB servers to actively provide the required memory blocks for validators.
        \item Making more accounts will not provide any advantages for a ZK-EDB server.
    \end{itemize}

    For our incentive mechanism, we require that every time a validator receives a memory block from a ZK-EDB, after
    validating the data, he give a receipt to the ZK-EDB. In this receipt the validator signs the following information:

    \begin{itemize}
        \item \texttt{ownerAddr} the ALGO address of the ZK-EDB\@.
        \item \texttt{receivedBlockID} the ID of the received memory block.
        \item \texttt{round} the current round number.
    \end{itemize}

    \emph{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}

    To incentivize ZK-EDB servers, every round a lottery will be held and a predefined amount of ALGOs from
    \texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
    between all the \emph{winning tickets} of the lottery.

    \emph{One ZK-EDB server could own multiple winning tickets in a round.}

    To run this lottery, In every round, based on the current block seed, a collection of \emph{valid} receipts will be
    selected randomly as the \emph{winning} receipts. A receipt is \emph{valid} in the round \(r\) if:

    \begin{itemize}
        \item The signer was a validator in the round \(r - 1\) and voted for the agreed-upon block.
        \item The data block in the receipt was needed for validating the \textbf{previous} block.
        \item The receipt round number is \(r - 1\).
        \item The signer did not sign a receipt for the same data block for two different ZK-EDBs in the previous round.
    \end{itemize}

    For selecting the winning receipts we could use a random generator:

    \begin{verbatim}
        if random(seed|validatorPK|receivedBlockID) < winProbability
            the receipt issued by validatorPK for receivedBlockID is a winner
    \end{verbatim}

    \begin{itemize}
        \item \texttt{random()} produces uniform random numbers between 0 and 1, using its input argument as a seed.
        \item \texttt{validatorPK} is the public key of the signer of the receipt.
        \item \texttt{receivedBlockID} is the ID of the memory block that the receipt was issued for.
        \item \texttt{winProbability} is the probability of winning in every round.
        \item \texttt{seed} is the current block seed.
        \item \texttt{|} is a concatenation operator.
    \end{itemize}

    \emph{The winners of the lottery were validators one round before the lottery round.}

    Also, based on the current block seed, a random memory block, whether a heap page or a code area block, is
    selected as the challenge of the round. A ZK-EDB that owns a winning receipt needs to broadcast a \emph{winning
    ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
    challenge. Solving a round challenge requires the content of the memory block which was selected as the round
    challenge. This will encourage ZK-EDBs to store all memory blocks.

    A possible choice for the challenge solution could be the cryptographic hash of the content of the challenge
    memory block combined with the ZK-EDB ALGO address: \texttt{hash(challenge.content|ownerAddr)}

    The winning tickets of the lottery of the round \(r\) need to be included in the block of the round \(r\),
    otherwise they will be considered expired. Validation and prize distribution for the winning tickets of the round
    \(r\) will be done in the round \(r + 1\). This way, \textbf{the content of the challenge memory block could be
    kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.

    \subsection{Memory Allocation and De-allocation}\label{subsec:memory-allocation-and-de-allocation}

    Every k round the protocol chooses a price per byte for the AVM memory. When a smart contract executes a heap
    allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ALGO
    address of the smart contract.

    To determine the price of AVM memory, Every k round, the protocol calculates \texttt{dbFee} and
    \texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees and
    \texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
    the protocol considers the total size of all memory pages that were accessed for either reading or writing during
    some time. These two values will be calculated for the last k rounds and the price per byte of the AVM memory
    will be a linear function of \texttt{dbFee/memTraffic}

    When a smart contract executes a heap de-allocation instruction, the protocol will refund the cost of
    de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
    calculates the refunded amount based on the average price the smart had paid for that allocated memory. This will
    prevent smart contracts from profit taking by trading memory with the protocol.

    \section{Concurrency}\label{sec:concurrency}

    Every block of the Algorand blockchain contains a list of transactions. This list is an ordered list and the
    effect of its contained transactions must be applied to the AVM state sequentially as they appear in the ordered
    list. The ordering of transactions in a block is solely chosen by the block proposer.

    \emph{Users should not have any assumption about the ordering of transactions in a block.}

    The fact that block transactions constitute a sequential list, does not mean they can not be validated
    concurrently. Many transactions are actually independent and the order of their execution does not matter. These
    transactions can be safely validated in parallel by validators.

    A transaction can change the AVM state by modifying either the code area or the AVM heap. In Algorand, all
    transactions declare the list of memory blocks they want to read or write. This will enable us to determine the
    independent sets of transactions which can be validated in parallel. To do so, we define the \emph{memory dependency
    graph} as follows:

    \begin{itemize}
        \item \(G\) is an undirected graph.
        \item Every vertex in \(G\) corresponds to a transaction.
        \item Vertices \(u\) and \(v\) are adjacent in \(G\) if and only if u has a memory block \(B\) in its writing
        list and \(v\) has \(B\) in either its writing list or reading list.
    \end{itemize}

    If we consider a proper vertex coloring of \(G\), every color class will give us an independent set of
    transactions that can be validated concurrently. To achieve the highest parallelization, we need to color \(G\)
    with minimum number of colors. Thus, the chromatic number of the memory dependency graph shows how good a
    transaction set could be run concurrently.

    Graph coloring is computationally NP-hard. However, in our use case we don't need to necessarily find an optimal
    solution. An approximate greedy algorithm will perform well enough in most circumstances. The block proposer is
    responsible for solving the graph coloring problem anda proposed block must determine the independent sets of
    transactions which can be run in parallel safely. Since with better parallelization a block can contain more
    transactions, a proposer is incentivized enough to find a good graph coloring.

    \section{Consensus}\label{sec:consensus}

    \subsection{Estimating A User's Stake}\label{subsec:estimating-a-user's-stake}

    In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
    of stake the user has in the system.
    Conventionally in these systems, for estimating a user's stake, we use the amount of native system tokens the
    user is holding.
    Unfortunately, one problem with this approach is that a strong attacker may be able to obtain a considerable
    amount of system tokens, for example by borrowing from a DEFI application, and use this stake to attack the system.

    To mitigate this problem, for calculating a user's stake, instead of using the raw ALGO balance, we use the
    minimum of a \emph{trust value} that the system has calculated for the user and the user's ALGO balance:
    \[
        Stake_{user} = \min (Balance_{user},Trust_{user})
    \]
    For estimating the value of \(Trust_{user}\) we use the exponential moving average of the user's ALGO balance.
    Therefore, in our system a user who held ALGOs and participated in the consensus for a long time is more trusted
    than a new user with a higher balance.
    An attacker who has obtained a large amount of ALGOs, also needs to hold them for a long period of time before
    being able to attack our system.

    For calculating the exponential moving average of a time series at the time step \(t\), we can use the following
    recursive formula:
    \[
        M_t = (1 - \alpha) M_{t-1} + \alpha X_t = M_{t-1} + \alpha (X_t - M_{t-1})
    \]
    Where:
    \begin{itemize}
        \item The coefficient \(\alpha\) is a constant smoothing factor between 0 and 1 which represents the degree of
        weighting decrease, A higher \(\alpha\) discounts older observations faster.
        \item \(X_t\) is the value of the time series at the time step \(t\).
        \item \(M_t\) is the value of the EMA at the time step \(t\).
    \end{itemize}

    Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
    \(M_t\):
    \[
        M_t = (1 - \alpha)^{t-k}M_k + [1 - (1 - \alpha)^{t - k}]X
    \]
    Where:
    \[
        X = X_{k+1} = X_{k+2} = \dots = X_t
    \]
    When \(|nx| \ll 1\) we can use the binomial approximation  \((1 + x)^n \approx 1 + nx\) to further simplify this
    formula:
    \[
        M_t = M_k + (t - k) \alpha (X - M_k)
    \]

    For choosing the value of \(\alpha\) we can consider the number of time steps that the trust value of a user needs
    for reaching a specified fraction of his account balance.
    We know that for large \(n\) and \(|x| < 1\) we have \((1 + x)^n \approx e^{nx}\), so by letting \(M_k = 0\) and
    \(n = t - k\)
    we can write:
    \[
        \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{X}\right)}{n}
    \]
    The value of \(\alpha\) for a desired configuration can be calculated by this equation.
    For instance, we could calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8X\)
    and \(n\) equals to the number of time steps of 10 years.

    \section{Smart Contract Oracle}\label{sec:smart-contract-oracle}

    A smart contract oracle is a full AVM emulator that keeps a full local copy of AVM memory and can emulate AVM
    execution without accessing aZK-EDB. Smart contract oracles can be used for reporting useful information about
    \texttt{avmCall} transactions such as accessed AVM heap pages or code area blocks, exact amount of execution cost,
    and soon.

\end{document}

### 2021-03-02 tex/specs.tex
Complete TeX versions
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
+\usepackage{hyperref}
+
 \title{The Algorand Virtual Machine Specification}
-%\author{aybehrouz \thanks{funded by myself}}
-%\date{2020}
+\author{aybehrouz}
+\date{January 2021}
+
 % Document
@@
 
-    The Algorand Virtual Machine (AVM) is an abstract computing machine for executing Algorand's smart contracts.
-    It is designed in a way that it could be efficiently implemented in either hardware or software.
+    The Algorand Virtual Machine (AVM) is an abstract computing machine for executing Algorand's smart contracts. It
+    is designed in a way that it could be efficiently implemented in either hardware or software.
 
@@
     The Algorand Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
-    and does
-    not have to be done by the Algorand Virtual Machine itself.
+    and does not have to be done by the Algorand Virtual Machine itself.
 
     The instruction set of the Algorand Virtual Machine distinguishes its operand types using instructions intended to
-    operate on values of specific types.
-    For instance, \texttt{iadd} assumes that its operands are two 64-bit integers.
+    operate on values of specific types. For instance, \texttt{iadd} assumes that its operands are two 64-bit integers.
 
@@
 
-    The Algorand Virtual Machine has a byte addressable memory which is divided into separate segments.
-    Every segment belongs to a smart contract that has a unique \texttt{applicationID}.
-    The AVM always has a single working memory segment and memory locations outside its current working segment can
-    not be accessed.
-    The only instructions that can change the current working segment are \texttt{invoke\_external}, \texttt{athrow} and
-    return instructions.
+    The Algorand Virtual Machine has a byte addressable memory which is divided into separate segments. Every segment
+    belongs to a smart contract that has a unique \texttt{applicationID}. The AVM always has a single working memory
+    segment and memory locations outside its current working segment can not be accessed. The only instructions which
+    can change the current working segment are \texttt{invoke\_external}, \texttt{athrow} and return instructions.
 
-    Every smart contract has its own memory segment.
-    Hence, there is no way for a smart contract to access another smart contract's memory.
-    Interaction between smart contracts is done using \texttt{invoke\_external} instruction, and A smart contract can
-    invoke methods of another smart contract by this instruction.
+    Every smart contract has its own memory segment. Hence, there is no way for a smart contract to access another
+    smart contract's memory. Interaction between smart contracts is done using \texttt{invoke\_external} instruction,
+    and A smart contract can invoke methods of another smart contract by this instruction.
 
@@
     \end{itemize}
-    All data areas except operand stack, have their own address space which starts from 0.
-    Operand stack is a last-in-first-out (LIFO) stack and is not addressable.
-    Every memory access instruction operates on its specific data areas.
+    All data areas except operand stack, have their own address space which starts from 0. Operand stack is a
+    last-in-first-out (LIFO) stack and is not addressable. Every memory access instruction operates on its specific
+    data areas.
 
@@
 
-    The code area of a segment contains the byte-code of the smart contract which owns that segment.
-    The AVM has no instructions for manipulating the code area.
-    \textbf{Installing, removing and updating smart contracts need to be done externally.}
+    The code area of a segment contains the byte-code of the smart contract which owns that segment. The AVM has no
+    instructions for manipulating the code area. \textbf{Installing, removing and updating smart contracts need to
+    be done externally.}
 
@@
     The constant area of a segment contains several kinds of constants, ranging from user defined constants to method
-    address tables.
-    A method address table stores method locations in the code area and their access type.
-    The access type of a method can be either \texttt{public} or \texttt{private}.
-    The AVM has no instructions for modifying the constant area.
+    address tables. A method address table stores method locations in the code area and their access type. The access
+    type of a method can be either \texttt{public} or \texttt{private}. The AVM has no instructions for modifying the
+    constant area.
 
@@
 
-    A local frame is used to store methods parameters and local variables.
-    A new frame is created each time a method is invoked.
-    A frame is destroyed when its method invocation completes, whether the completion is normal or abrupt.
+    A local frame is used to store methods parameters and local variables. A new frame is created each time a method
+    is invoked. A frame is destroyed when its method invocation completes, whether the completion is normal or abrupt.
 
@@
 
-    Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too.
-    AVM instructions take operands from the operand stack, operate on them, and push the result back onto the operand
-    stack.
-    An operand stack is destroyed when its owner method completes, whether that completion is normal or abrupt.
+    Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too. AVM
+    instructions take operands from the operand stack, operate on them, and push the result back onto the operand
+    stack. An operand stack is destroyed when its owner method completes, whether that completion is normal or abrupt.
 
@@
 
-    The heap of a segment is a persistent memory area which is divided into pages.
-    Each page can be referenced by an index that starts from 0.
-    Memory locations inside every page have a separate address space that starts from 0.
-    In other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex,
-        offset)}.
-    Pages of a heap do not need to be equally sized.
+    The heap of a segment is a persistent memory area which is divided into pages. Each page can be referenced by an
+    index that starts from 0. Memory locations inside every page have a separate address space that starts from 0. In
+    other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex, offset)
+    }. Pages of a heap do not need to be equally sized.
 
     \emph{The reason behind this paged design is that the heap is usually persisted using a block device. A heap with
-    a paged
-    structure could expose the underlying block based nature of the persistence layer to the application layer. In
-    this way, the compiler or the programmer could better optimize the code for the persistence layer.}
+    a paged structure could expose the underlying block based nature of the persistence layer to the application
+    layer. In this way, the compiler or the programmer could better optimize the code for the persistence layer.}
 
@@
     An Algorand Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
-    performed,
-    followed by zero or more operands supplying arguments or data that are used by the operation.
+    performed, followed by zero or more operands supplying arguments or data that are used by the operation.
     \textbf{The number and size of the operands are determined solely by the opcode.}
@@
         \item \texttt{invoke\_internal} invokes a method from the current running smart contract.
-        \item \texttt{invoke\_external} invokes a \texttt{public} method from another smart contract.
-        It will change the current memory segment to the segment of the invoked smart contract.
-        \item \texttt{invoke\_native} invokes a method that is not hosted by the Algorand virtual machine.
-        By this instruction, high performance native methods of a hosting machine could become available to AVM smart
-        contracts.
+        \item \texttt{invoke\_external} invokes a \texttt{public} method from another smart contract. It will change
+        the current memory segment to the segment of the invoked smart contract.
+        \item \texttt{invoke\_native} invokes a method that is not hosted by the Algorand virtual machine. By this
+        instruction, high performance native methods of a hosting machine could become available to AVM smart contracts.
     \end{itemize}
@@
 
-    Each time a method is invoked a new local frame and operand stack is created.
-    The Algorand Virtual Machine uses local frames to pass parameters on method invocation.
-    On method invocation, any parameters are passed in consecutive local variables stored in the method's local frame
-    starting from local variable 0.
-    The invoker of a method writes the parameters in the local frame of the invoked method using \texttt{arg}
-    instructions.
+    Each time a method is invoked a new local frame and operand stack is created. The Algorand Virtual Machine uses
+    local frames to pass parameters on method invocation. On method invocation, any parameters are passed in
+    consecutive local variables stored in the method's local frame starting from local variable 0. The invoker of a
+    method writes the parameters in the local frame of the invoked method using \texttt{arg} instructions.
 
@@
 
-    An exception is thrown programmatically using the \texttt{athrow} instruction.
-    Exceptions can also be thrown by various Algorand Virtual Machine instructions if they detect an abnormal
-    condition.
-    Some exceptions are not catchable and will always abort the execution of the smart contract.
+    An exception is thrown programmatically using the \texttt{athrow} instruction. Exceptions can also be thrown by
+    various Algorand Virtual Machine instructions if they detect an abnormal condition. Some exceptions are not
+    catchable and will always abort the execution of the smart contract.
 
@@
     A method invocation completes normally if that invocation does not cause an exception to be thrown, either
-    directly from the AVM or as a result of executing an explicit throw statement.
-    If the invocation of the current method completes normally, then a value may be returned to the invoking method.
-    This occurs when the invoked method executes one of the return instructions, the choice of which must be
-    appropriate for the type of the value being returned (if any).
-    Execution then continues normally in the invoking method's local frame with the returned value (if any)
-    pushed onto the operand stack.
+    directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
+    method completes normally, then a value may be returned to the invoking method. This occurs when the invoked
+    method executes one of the return instructions, the choice of which must be appropriate for the type of the value
+    being returned (if any). Execution then continues normally in the invoking method's local frame with the returned
+    value (if any) pushed onto the operand stack.
 
-    A method invocation completes abruptly if an exceptions is thrown and is not caught by the current method.
-    A method invocation that completes abruptly never returns a value to its invoker.
+    A method invocation completes abruptly if an exceptions is thrown and is not caught by the current method. A
+    method invocation that completes abruptly never returns a value to its invoker.
 
@@
     including its local frame and operand stack, with the \texttt{PC} register appropriately restored and incremented
-    to skip past the method invocation instruction.
-    If the invoker was another smart contract, i.e.\ the invocation was made by an \texttt{invoke\_external}
-    instruction, the current memory segment will be changed to the invoker's segment.
+    to skip past the method invocation instruction. If the invoker was another smart contract, i.e.\ the invocation
+    was made by an \texttt{invoke\_external} instruction, the current memory segment will be changed to the invoker's
+    segment.
 
     A thrown exception causes methods in the call stack to complete \textbf{abruptly} one by one, as long as the
-    \texttt{PC} register is not pointing to a \texttt{catch} instruction.
-    The \texttt{catch} instruction acts like a branch instruction that branches only if an exception is caught.
-    \textbf{When an external method invocation completes abruptly, before changing the current segment, all changes
-    made to the heap area will be rolled back.}
+    \texttt{PC} register is not pointing to a \texttt{catch} instruction. The \texttt{catch} instruction acts like a
+    branch instruction that branches only if an exception is caught. \textbf{When an external method invocation
+    completes abruptly, before changing the current segment, all changes made to the heap area will be rolled back.}
 
@@
 
-    In blockchain applications, we usually need to authorize certain operations.
-    For example, for sending an asset from a user to another user, first we need to make sure that the sender has
-    authorized this operation.
-    The Algorand virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
-    cryptographic instructions for validating signatures and cryptographic entities.
-    By using these instructions and passing signatures as parameters to methods a programmer can implement the
-    required logic for authorizing any operation.
+    In blockchain applications, we usually need to authorize certain operations. For example, for sending an asset
+    from a user to another user, first we need to make sure that the sender has authorized this operation. The
+    Algorand virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
+    cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
+    passing signatures as parameters to methods a programmer can implement the required logic for authorizing any
+    operation.
 
@@
 
-    In addition to signatures, a method can verify its invoker by using \texttt{get\_parent} instruction.
-    This instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
-    smart contract in the call stack.
-    In other words, it gives the smart contract that has invoked the current smart contract. (if any)
+    In addition to signatures, a method can verify its invoker by using \texttt{get\_parent} instruction. This
+    instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
+    smart contract in the call stack. In other words, it gives the smart contract that has invoked the current smart
+    contract. (if any)
 

### 2021-03-03 tex/impl.tex
Introduce `beta` for controlling the minimum trust value
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
     In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
-    of stake the user has in the system.
-    Conventionally in these systems, for estimating a user's stake, we use the amount of native system tokens the
-    user is holding.
-    Unfortunately, one problem with this approach is that a strong attacker may be able to obtain a considerable
-    amount of system tokens, for example by borrowing from a DEFI application, and use this stake to attack the system.
+    of stake the user has in the system. Conventionally in these systems, for estimating a user's stake, we use the
+    amount of native system tokens the user is holding. Unfortunately, one problem with this approach is that a
+    strong attacker may be able to obtain a considerable amount of system tokens, for example by borrowing from a
+    DEFI application, and use this stake to attack the system.
 
-    To mitigate this problem, for calculating a user's stake, instead of using the raw ALGO balance, we use the
-    minimum of a \emph{trust value} that the system has calculated for the user and the user's ALGO balance:
+    To mitigate this problem, for calculating a user's stake at the round \(n\), instead of using the raw ALGO
+    balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
+    ALGO balance:
     \[
-        Stake_{user} = \min (Balance_{user},Trust_{user})
+        Stake_{user,n} = \min (Balance_{user,n}, Trust_{user,n})
     \]
-    For estimating the value of \(Trust_{user}\) we use the exponential moving average of the user's ALGO balance.
-    Therefore, in our system a user who held ALGOs and participated in the consensus for a long time is more trusted
-    than a new user with a higher balance.
-    An attacker who has obtained a large amount of ALGOs, also needs to hold them for a long period of time before
-    being able to attack our system.
+    For estimating the value of \(Trust_{user,n}\) we use the following formula:
+    \[
+        Trust_{user,n} = \max (M_n, \beta Balance_{user,n})
+    \]
+    Where \(M_n\) is the exponential moving average of the user's ALGO balance at the round \(n\) and \(\beta\) is a
+    constant between \(0\) and \(1\) determining the initial trust value of new users.
+
+    In our system a user who held ALGOs and participated in the consensus for a long time is more trusted
+    than a new user with a higher balance. An attacker who has obtained a large amount of ALGOs, also needs to hold
+    them for a long period of time before being able to attack our system.
 
@@
     \begin{itemize}
-        \item The coefficient \(\alpha\) is a constant smoothing factor between 0 and 1 which represents the degree of
-        weighting decrease, A higher \(\alpha\) discounts older observations faster.
+        \item The coefficient \(\alpha\) is a constant smoothing factor between \(0\) and\(1\) which represents the
+        degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
         \item \(X_t\) is the value of the time series at the time step \(t\).
@@
     For choosing the value of \(\alpha\) we can consider the number of time steps that the trust value of a user needs
-    for reaching a specified fraction of his account balance.
-    We know that for large \(n\) and \(|x| < 1\) we have \((1 + x)^n \approx e^{nx}\), so by letting \(M_k = 0\) and
+    for reaching a specified fraction of his account balance. We know that for large \(n\) and \(|x| < 1\) we have \(
+    (1 + x)^n \approx e^{nx}\), so by letting \(M_k = 0\) and
     \(n = t - k\)
@@
     \]
-    The value of \(\alpha\) for a desired configuration can be calculated by this equation.
-    For instance, we could calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8X\)
-    and \(n\) equals to the number of time steps of 10 years.
+    The value of \(\alpha\) for a desired configuration can be calculated by this equation. For instance, we could
+    calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8X\) and \(n\) equals to the
+    number of time steps of 10 years.
 

### 2021-03-04 tex/avm.tex
Explain total stake calculation and update the makefile
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-03-03 tex/impl.tex
@@
 % Preamble
-\documentclass[11pt, A4]{article}
+\documentclass[11pt, A4]{report}
 
@@
 
-\title{The Algorand Virtual Machine Implementation}
+\title{The Algorand Virtual Machine}
 \author{aybehrouz}
@@
     \maketitle
+
+    \chapter{Specification}\label{ch:specification}
+
+    \section{Introduction}\label{sec:introduction}
+
+    The Algorand Virtual Machine (AVM) is an abstract computing machine for executing Algorand's smart contracts. It
+    is designed in a way that it could be efficiently implemented in either hardware or software.
+
+    \section{The Structure of the Algorand Virtual Machine}\label{sec:the-structure-of-the-algorand-virtual-machine}
+
+    ...
+
+    \subsection{Data Types}\label{subsec:data-types}
+
+    The Algorand Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
+    and does not have to be done by the Algorand Virtual Machine itself.
+
+    The instruction set of the Algorand Virtual Machine distinguishes its operand types using instructions intended to
+    operate on values of specific types. For instance, \texttt{iadd} assumes that its operands are two 64-bit integers.
+
+    \subsection{The \texttt{PC} Register}\label{subsec:the-pc-register}
+
+    ...
+
+    \subsection{Call Stack}\label{subsec:call-stack}
+
+    A call stack contains the information that is needed for restoring the state of the invoker of a method.
+
+    \subsection{Memory Unit}\label{subsec:memory-unit}
+
+    The Algorand Virtual Machine has a byte addressable memory which is divided into separate segments. Every segment
+    belongs to a smart contract that has a unique \texttt{applicationID}. The AVM always has a single working memory
+    segment and memory locations outside its current working segment can not be accessed. The only instructions which
+    can change the current working segment are \texttt{invoke\_external}, \texttt{athrow} and return instructions.
+
+    Every smart contract has its own memory segment. Hence, there is no way for a smart contract to access another
+    smart contract's memory. Interaction between smart contracts is done using \texttt{invoke\_external} instruction,
+    and A smart contract can invoke methods of another smart contract by this instruction.
+
+    A memory segment consists of five data areas:
+    \begin{itemize}
+        \item Code Area
+        \item Constant Area
+        \item Local Frame
+        \item Operand Stack
+        \item Heap
+    \end{itemize}
+    All data areas except operand stack, have their own address space which starts from 0. Operand stack is a
+    last-in-first-out (LIFO) stack and is not addressable. Every memory access instruction operates on its specific
+    data areas.
+
+    \subsubsection{Code Area}
+
+    The code area of a segment contains the byte-code of the smart contract which owns that segment. The AVM has no
+    instructions for manipulating the code area. \textbf{Installing, removing and updating smart contracts need to
+    be done externally.}
+
+    \subsubsection{Constant Area}
+
+    The constant area of a segment contains several kinds of constants, ranging from user defined constants to method
+    address tables. A method address table stores method locations in the code area and their access type. The access
+    type of a method can be either \texttt{public} or \texttt{private}. The AVM has no instructions for modifying the
+    constant area.
+
+    \emph{Only public methods can be invoked by \texttt{invoke\_external} instruction.}
+
+    \subsubsection{Local Frame}
+
+    A local frame is used to store methods parameters and local variables. A new frame is created each time a method
+    is invoked. A frame is destroyed when its method invocation completes, whether the completion is normal or abrupt.
+
+    \subsubsection{Operand Stack}
+
+    Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too. AVM
+    instructions take operands from the operand stack, operate on them, and push the result back onto the operand
+    stack. An operand stack is destroyed when its owner method completes, whether that completion is normal or abrupt.
+
+    \subsubsection{Heap}
+
+    The heap of a segment is a persistent memory area which is divided into pages. Each page can be referenced by an
+    index that starts from 0. Memory locations inside every page have a separate address space that starts from 0. In
+    other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex, offset)
+    }. Pages of a heap do not need to be equally sized.
+
+    \emph{The reason behind this paged design is that the heap is usually persisted using a block device. A heap with
+    a paged structure could expose the underlying block based nature of the persistence layer to the application
+    layer. In this way, the compiler or the programmer could better optimize the code for the persistence layer.}
+
+    \section{Instruction Set Summary}\label{subsec:instruction-set-summary}
+
+    An Algorand Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
+    performed, followed by zero or more operands supplying arguments or data that are used by the operation.
+    \textbf{The number and size of the operands are determined solely by the opcode.}
+
+    \subsection{Method Invocation}\label{subsec:method-invocation}
+
+    The Algorand Virtual Machine has three types of method invocation:
+    \begin{itemize}
+        \item \texttt{invoke\_internal} invokes a method from the current running smart contract.
+        \item \texttt{invoke\_external} invokes a \texttt{public} method from another smart contract. It will change
+        the current memory segment to the segment of the invoked smart contract.
+        \item \texttt{invoke\_native} invokes a method that is not hosted by the Algorand virtual machine. By this
+        instruction, high performance native methods of a hosting machine could become available to AVM smart contracts.
+    \end{itemize}
+
+    \emph{In the future, we may need to add special instructions for invoking interface and virtual methods...}
+
+    Each time a method is invoked a new local frame and operand stack is created. The Algorand Virtual Machine uses
+    local frames to pass parameters on method invocation. On method invocation, any parameters are passed in
+    consecutive local variables stored in the method's local frame starting from local variable 0. The invoker of a
+    method writes the parameters in the local frame of the invoked method using \texttt{arg} instructions.
+
+    \subsubsection{Exceptions}
+
+    An exception is thrown programmatically using the \texttt{athrow} instruction. Exceptions can also be thrown by
+    various Algorand Virtual Machine instructions if they detect an abnormal condition. Some exceptions are not
+    catchable and will always abort the execution of the smart contract.
+
+    \subsubsection{Method Invocation Completion}
+
+    A method invocation completes normally if that invocation does not cause an exception to be thrown, either
+    directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
+    method completes normally, then a value may be returned to the invoking method. This occurs when the invoked
+    method executes one of the return instructions, the choice of which must be appropriate for the type of the value
+    being returned (if any). Execution then continues normally in the invoking method's local frame with the returned
+    value (if any) pushed onto the operand stack.
+
+    A method invocation completes abruptly if an exceptions is thrown and is not caught by the current method. A
+    method invocation that completes abruptly never returns a value to its invoker.
+
+    When a method completes, whether normally or abruptly, the call stack is used to restore the state of the invoker,
+    including its local frame and operand stack, with the \texttt{PC} register appropriately restored and incremented
+    to skip past the method invocation instruction. If the invoker was another smart contract, i.e.\ the invocation
+    was made by an \texttt{invoke\_external} instruction, the current memory segment will be changed to the invoker's
+    segment.
+
+    A thrown exception causes methods in the call stack to complete \textbf{abruptly} one by one, as long as the
+    \texttt{PC} register is not pointing to a \texttt{catch} instruction. The \texttt{catch} instruction acts like a
+    branch instruction that branches only if an exception is caught. \textbf{When an external method invocation
+    completes abruptly, before changing the current segment, all changes made to the heap area will be rolled back.}
+
+    \emph{By using the \texttt{athrow} instruction properly, a programmer can make any method act like an atomic
+    operation.}
+
+    \subsubsection{Authorizing Operations}
+
+    In blockchain applications, we usually need to authorize certain operations. For example, for sending an asset
+    from a user to another user, first we need to make sure that the sender has authorized this operation. The
+    Algorand virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
+    cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
+    passing signatures as parameters to methods a programmer can implement the required logic for authorizing any
+    operation.
+
+    \emph{The Algorand virtual machine has no instructions for issuing cryptographic signatures.}
+
+    In addition to signatures, a method can verify its invoker by using \texttt{get\_parent} instruction. This
+    instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
+    smart contract in the call stack. In other words, it gives the smart contract that has invoked the current smart
+    contract. (if any)
+
+    \subsubsection{Heap Allocation Instructions}
+
+    ...
+
+    \chapter{Implementation}\label{ch:implementation}
 
@@
 
-    To mitigate this problem, for calculating a user's stake at the round \(n\), instead of using the raw ALGO
+    To mitigate this problem, for calculating a user's stake at the time step \(t\), instead of using the raw ALGO
     balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
@@
     \[
-        Stake_{user,n} = \min (Balance_{user,n}, Trust_{user,n})
+        Stake_{u,t} = \min (Balance_{u,t}, Trust_{u,t})
     \]
-    For estimating the value of \(Trust_{user,n}\) we use the following formula:
+    For estimating the value of \(Trust_{u,t}\) we use the following formula:
     \[
-        Trust_{user,n} = \max (M_n, \beta Balance_{user,n})
+        Trust_{u,t} = \max (M_{u,t}, \beta Balance_{u,t})
     \]
-    Where \(M_n\) is the exponential moving average of the user's ALGO balance at the round \(n\) and \(\beta\) is a
-    constant between \(0\) and \(1\) determining the initial trust value of new users.
+    Where \(M_{u,t}\) is the exponential moving average of the ALGO balance of the user \(u\) at the time step \(t\),
+    and \(\beta\) is a constant between \(0\) and \(1\) determining the minimum trust value of a user.
+
+    The agreement protocol, at the time step \(t\), will use \(\sum_{u}Stake_{u,t}\) to determine the required number of
+    votes for the confirmation of a block. At the start of the agreement protocol, \(\beta\) is initialized to \(0\).
+    If after confirming a block we have:
+    \[
+        \sum_{u}Stake_u < \gamma \sum_{u} Balance_u
+    \]
+    The protocol will increase the value of \(\beta\) to make sure that the total stake of the system goes high enough.
+    If after confirming a block we have \(\beta > 0\) and:
+    \[
+        \sum_{u}Stake_u > \lambda \sum_{u} Balance_u
+    \]
+    The protocol will decrease the value of \(\beta\) until it eventually reaches \(0\) again.
 
     In our system a user who held ALGOs and participated in the consensus for a long time is more trusted
-    than a new user with a higher balance. An attacker who has obtained a large amount of ALGOs, also needs to hold
-    them for a long period of time before being able to attack our system.
+    than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
+    amount of ALGOs, also needs to hold them for a long period of time before being able to attack our system.
 
@@
     A smart contract oracle is a full AVM emulator that keeps a full local copy of AVM memory and can emulate AVM
-    execution without accessing aZK-EDB. Smart contract oracles can be used for reporting useful information about
+    execution without accessing a ZK-EDB. Smart contract oracles can be used for reporting useful information about
     \texttt{avmCall} transactions such as accessed AVM heap pages or code area blocks, exact amount of execution cost,

### 2021-03-04 tex/avm.tex
Add proposal.md again
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-    \section{Instruction Set Summary}\label{subsec:instruction-set-summary}
+    \section{Instruction Set Summary}\label{sec:instruction-set-summary}
 
@@
     must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
-    a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound\ -\ issuance\
-        \textgreater{}\ maxLifeTime} or if a signature of the same user with a bigger or equal nonce is already used
-    (i.e
-    .~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two nonces
-    have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
+    a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance \textgreater{}\
+        maxLifeTime} or if a signature of the same user with a bigger or equal nonce is already used
+    (i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
+    nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
 
@@
     maximum allowed lifetime of a signature, this information can be safely deleted. \textbf{As a result, we will not
-    have the problem of " unremovable empty accounts" like Ethereum.}
+    have the problem of "un-removable empty accounts" like Ethereum.}
 
@@
 
-    \section{Incentive mechanism}\label{incentive-mechanism}
+    \section{Incentive mechanism}\label{sec:incentive-mechanism}
 
@@
         \item Every vertex in \(G\) corresponds to a transaction.
-        \item Vertices \(u\) and \(v\) are adjacent in \(G\) if and only if u has a memory block \(B\) in its writing
-        list and \(v\) has \(B\) in either its writing list or reading list.
+        \item Vertices \(u\) and \(v\) are adjacent in \(G\) if and only if \(u\) has a memory block \(B\) in its
+        writing list and \(v\) has \(B\) in either its writing list or reading list.
     \end{itemize}
@@
     If we consider a proper vertex coloring of \(G\), every color class will give us an independent set of
-    transactions that can be validated concurrently. To achieve the highest parallelization, we need to color \(G\)
+    transactions which can be validated concurrently. To achieve the highest parallelization, we need to color \(G\)
     with minimum number of colors. Thus, the chromatic number of the memory dependency graph shows how good a
@@
     \]
-    The protocol will increase the value of \(\beta\) to make sure that the total stake of the system goes high enough.
+    The protocol will increase the value of \(\beta\) to make sure that the total stake of the system is not too low.
+
     If after confirming a block we have \(\beta > 0\) and:

### 2021-03-05 tex/avm.tex
Change how we handle low total stake
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 \usepackage{hyperref}
+%\binoppenalty=99999
+%\relpenalty=99999
 
@@
     \[
-        Stake_{u,t} = \min (Balance_{u,t}, Trust_{u,t})
+        S_{u,t} = \min (B_{u,t}, Trust_{u,t})
     \]
-    For estimating the value of \(Trust_{u,t}\) we use the following formula:
-    \[
-        Trust_{u,t} = \max (M_{u,t}, \beta Balance_{u,t})
-    \]
-    Where \(M_{u,t}\) is the exponential moving average of the ALGO balance of the user \(u\) at the time step \(t\),
-    and \(\beta\) is a constant between \(0\) and \(1\) determining the minimum trust value of a user.
-
-    The agreement protocol, at the time step \(t\), will use \(\sum_{u}Stake_{u,t}\) to determine the required number of
-    votes for the confirmation of a block. At the start of the agreement protocol, \(\beta\) is initialized to \(0\).
-    If after confirming a block we have:
-    \[
-        \sum_{u}Stake_u < \gamma \sum_{u} Balance_u
-    \]
-    The protocol will increase the value of \(\beta\) to make sure that the total stake of the system is not too low.
-
-    If after confirming a block we have \(\beta > 0\) and:
-    \[
-        \sum_{u}Stake_u > \lambda \sum_{u} Balance_u
-    \]
-    The protocol will decrease the value of \(\beta\) until it eventually reaches \(0\) again.
+    The agreement protocol, at the time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
+    number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
+    exponential moving average of the ALGO balance of the user \(u\) at the time step \(t\).
 
@@
     than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
-    amount of ALGOs, also needs to hold them for a long period of time before being able to attack our system.
+    amount of ALGOs, also needs to hold them for a long period of time before being able to attack the system.
 
-    For calculating the exponential moving average of a time series at the time step \(t\), we can use the following
+    For calculating the exponential moving average of a user's balance at the time step \(t\), we can use the following
     recursive formula:
     \[
-        M_t = (1 - \alpha) M_{t-1} + \alpha X_t = M_{t-1} + \alpha (X_t - M_{t-1})
+        M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
     \]
@@
     \begin{itemize}
-        \item The coefficient \(\alpha\) is a constant smoothing factor between \(0\) and\(1\) which represents the
+        \item The coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
         degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
-        \item \(X_t\) is the value of the time series at the time step \(t\).
-        \item \(M_t\) is the value of the EMA at the time step \(t\).
+        \item \(B_{u,t}\) is the balance of the user \(u\) at the time step \(t\).
+        \item \(M_{u,t}\) is the EMA for the user \(u\) at the time step \(t\).
     \end{itemize}
@@
     Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
-    \(M_t\):
+    \(M_{u,t}\): (In the following equations the \(u\) subscript is dropped for simplicity)
     \[
-        M_t = (1 - \alpha)^{t-k}M_k + [1 - (1 - \alpha)^{t - k}]X
+        M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B
     \]
@@
     \[
-        X = X_{k+1} = X_{k+2} = \dots = X_t
+        B = B_{k+1} = B_{k+2} = \dots = B_{t}
     \]
-    When \(|nx| \ll 1\) we can use the binomial approximation  \((1 + x)^n \approx 1 + nx\) to further simplify this
-    formula:
+    We know that when \(|nx| \ll 1\) we can use the binomial approximation \({(1 + x)^n \approx 1 + nx}\). So, we can
+    further simplify this formula:
     \[
-        M_t = M_k + (t - k) \alpha (X - M_k)
+        M_{t} = M_{k} + (t - k) \alpha (B - M_{k})
     \]
@@
     For choosing the value of \(\alpha\) we can consider the number of time steps that the trust value of a user needs
-    for reaching a specified fraction of his account balance. We know that for large \(n\) and \(|x| < 1\) we have \(
-    (1 + x)^n \approx e^{nx}\), so by letting \(M_k = 0\) and
-    \(n = t - k\)
-    we can write:
+    for reaching a specified fraction of his account balance. We know that for large \(n\) and \(|x| < 1\) we have
+    \((1 + x)^n \approx e^{nx}\), so by letting \(M_{u,k} = 0\) and \(n = t - k\) we can write:
     \[
-        \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{X}\right)}{n}
+        \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{B}\right)}{n}
     \]
     The value of \(\alpha\) for a desired configuration can be calculated by this equation. For instance, we could
-    calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8X\) and \(n\) equals to the
+    calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8B\) and \(n\) equals to the
     number of time steps of 10 years.
+
+    In our system a newly created account will not have voting power for some time, no matter how high its
+    balance is. While this is a desirable property, in case a large proportion of total system tokens are
+    transferred to newly created accounts, it can result in too much voting power for older accounts. This may decrease
+    the degree of decentralization in our system.
+
+    However, this situation is easily detectable by comparing the total stake of the system with the total balance of
+    users. If after confirming a block the total stake of the system goes too low and we have:
+    \[
+        \sum_{u}S_{u,t} < \gamma \sum_{u}B_{u,t}
+    \]
+    The protocol will perform a \emph{time shift} in the system: the time step of the system
+    will be incremented for \(m\) steps while no blocks will be confirmed. This will increase the value of \(M_{u,t}\)
+    for new accounts with a non-zero balance, giving them more influence in the agreement protocol.
+
+    For calculating the value of \(m\) which determines the amount of time shift in the system, we should note that when
+    \(B_{u,t} = B_{u, t-1} = B_u\), we can derive a simple recursive rule for the stake of a user:
+    \[
+        S_{u,t} = (1 - \alpha) S_{u,t-1} + \alpha B_u
+    \]
+    Therefore, we have:
+    \[
+        \sum_{u}S_{u,t} = (1 - \alpha) \sum_{u}S_{u,t - 1} + \alpha \sum_{u}B_u
+    \]
+    This equation shows that when the balance of users is not changing over time the total stake of the system is the
+    exponential average of the total ALGOs of the system. Consequently, when we shift the time for \(m\) steps, we can
+    calculate the new total stake of the system from the following equation:
+
+    \[
+        \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
+    \]
+    Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
+    we can obtain \(m\) from the following formula, when \(\alpha\) is small:
+    \[
+        m = \frac{1}{\alpha} \ln \left(\frac{\gamma - 1}{\lambda - 1}\right)
+    \]
 

### 2021-03-07 tex/avm.tex
Fix minor readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
     transactions which can be validated concurrently. To achieve the highest parallelization, we need to color \(G\)
-    with minimum number of colors. Thus, the chromatic number of the memory dependency graph shows how good a
+    with minimum number of colors. Thus, the \emph{chromatic number} of the memory dependency graph shows how good a
     transaction set could be run concurrently.
@@
     Graph coloring is computationally NP-hard. However, in our use case we don't need to necessarily find an optimal
-    solution. An approximate greedy algorithm will perform well enough in most circumstances. The block proposer is
-    responsible for solving the graph coloring problem anda proposed block must determine the independent sets of
-    transactions which can be run in parallel safely. Since with better parallelization a block can contain more
+    solution. An approximate greedy algorithm will perform well enough in most circumstances. The block proposer is the
+    one who is responsible for solving the graph coloring problem and a proposed block must determine the independent
+    sets of transactions which can be run in parallel safely. Since with better parallelization a block can contain more
     transactions, a proposer is incentivized enough to find a good graph coloring.
@@
     \]
+    Where:
+    \begin{itemize}
+        \item \(S_{u,t}\) is the stake of the user \(u\) at the time step \(t\).
+        \item \(B_{u,t}\) is the ALGO balance of the user \(u\) at the time step \(t\).
+        \item \(Trust_{u,t}\) is an estimated trust value for the user \(u\) at the time step \(t\).
+    \end{itemize}
+
     The agreement protocol, at the time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
@@
     \]
-    Where:
-    \begin{itemize}
-        \item The coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
-        degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
-        \item \(B_{u,t}\) is the balance of the user \(u\) at the time step \(t\).
-        \item \(M_{u,t}\) is the EMA for the user \(u\) at the time step \(t\).
-    \end{itemize}
+    Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
+    degree of
+    weighting decrease, A higher \(\alpha\) discounts older observations faster.
 
@@
     Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
-    we can obtain \(m\) from the following formula, when \(\alpha\) is small:
+    we can obtain \(m\) from the following formula, assuming \(\alpha\) is small enough:
     \[
-        m = \frac{1}{\alpha} \ln \left(\frac{\gamma - 1}{\lambda - 1}\right)
+        m = \frac{1}{\alpha} \ln \left(\frac{1 - \gamma}{1 - \lambda}\right)
     \]

### 2021-03-13 tex/avm.tex
Update "Concurrency" section
This commit introduces the idea of memory chunks and concurrent counters, which greatly improves the concurrency of my design.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
     smart contract's memory. Interaction between smart contracts is done using \texttt{invoke\_external} instruction,
-    and A smart contract can invoke methods of another smart contract by this instruction.
+    and a smart contract can invoke methods of another smart contract by this instruction.
 
@@
     A local frame is used to store methods parameters and local variables. A new frame is created each time a method
-    is invoked. A frame is destroyed when its method invocation completes, whether the completion is normal or abrupt.
+    is invoked, and it is destroyed when its method invocation completes, whether the completion is normal or abrupt.
 
@@
 
-    The heap of a segment is a persistent memory area which is divided into pages. Each page can be referenced by an
-    index that starts from 0. Memory locations inside every page have a separate address space that starts from 0. In
-    other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex, offset)
-    }. Pages of a heap do not need to be equally sized.
+    The heap of a segment is a persistent memory area which is divided into pages. Memory locations inside every page
+    have a separate address space that starts from 0, and each page can be referenced by an index that starts from 0. In
+    other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex, offset)}.
+    Different pages of a heap do not need to be equally sized.
 
@@
         \item \texttt{invoke\_native} invokes a method that is not hosted by the Algorand virtual machine. By this
-        instruction, high performance native methods of a hosting machine could become available to AVM smart contracts.
+        instruction, high performance native methods of the hosting machine could become available to AVM smart
+        contracts.
     \end{itemize}
@@
     branch instruction that branches only if an exception is caught. \textbf{When an external method invocation
-    completes abruptly, before changing the current segment, all changes made to the heap area will be rolled back.}
+    completes abruptly, before changing the current segment, all changes made to the heap area by that method
+    call, including changes made to other segments, will be rolled back.} So, when a method call completes abruptly,
+    that method call essentially has no effect on the AVM state.
 
@@
 
-    \subsubsection{Authorizing Operations}
+    \subsection{Authorizing Operations}\label{subsec:authorizing-operations}
 
@@
     cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
-    passing signatures as parameters to methods a programmer can implement the required logic for authorizing any
+    passing signatures as parameters to methods, a programmer can implement the required logic for authorizing any
     operation.
@@
     instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
-    smart contract in the call stack. In other words, it gives the smart contract that has invoked the current smart
-    contract. (if any)
+    smart contract in the call stack. In other words, it returns the \texttt{applicationID} of the smart contract that
+    has invoked the current smart contract. (if any)
 
-    \subsubsection{Heap Allocation Instructions}
+    \subsection{Heap Allocation Instructions}\label{subsec:heap-allocation-instructions}
+
+    ...
+
+    \section{The AVM Standard Library}\label{sec:the-avm-standard-library}
 
@@
         \item Every state of the database has a commitment \(C\).
-        \item The ZK-EDB has a method \((D, p) = get(x)\), where \(x\) is a key and \(D\) is the associated value
-        with \(x\) and \(p\) is a proof.
+        \item The ZK-EDB has a method \((D, p) = get(x)\), where \(x\) is a key and \(D\) is the associated data
+        with \(x\), and \(p\) is a proof.
         \item A user can use \(C\) and \(p\) to verify that \(D\) is really associated with \(x\), and \(D\) is not
@@
 
-    \emph{Unlike heap pages, the AVM is not aware of different blocks of the code area.}
+    \emph{Unlike heap pages, the AVM is not aware of different blocks of code area.}
 
@@
         \item \texttt{installApp} installs an AVM smart contract and determines the update policy of the smart
-        contract: if the contract is updatable or not, the accounts that can update or uninstall the contract, and so
+        contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
         on.
@@
 
-    Every transaction is required to exactly specify what heap pages or code area blocks it will access. This enables
-    validators to start retrieving the required memory blocks from available ZK-EDB servers as soon as they see a
-    transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to access
-    a memory block that is not included in its access lists, will be rejected. Users could
-    use smart contract oracles to predict the list of memory blocks their transactions need. See
-    Section~\ref{sec:smart-contract-oracle} for more details.
+    Every transaction is required to exactly specify what heap locations or code area addresses it will access. This
+    enables validators to start retrieving the required memory blocks from available ZK-EDB servers as soon as they
+    see a transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to
+    access a memory location that is not included in its access lists, will be rejected. Users could use smart contract
+    oracles to predict the list of memory blocks their transactions need. See Section~\ref{sec:smart-contract-oracle}
+    for more details.
 
@@
         \item Execution cost
-        \item A list of heap/code-area pages for reading
-        \item A list of heap pages for writing
+        \item A list of heap/code-area locations for reading
+        \item A list of heap locations for writing
         \item A list of heap pages it will deallocate
@@
         \item It incentivizes storing all memory blocks, whether a heap page or a code area block, and not only those
-        that are used more frequently.
-        \item It incentivizes the ZK-EDB servers to actively provide the required memory blocks for validators.
+        which are used more frequently.
+        \item It incentivizes ZK-EDB servers to actively provide the required memory blocks for validators.
         \item Making more accounts will not provide any advantages for a ZK-EDB server.
@@
     \begin{itemize}
-        \item \texttt{ownerAddr} the ALGO address of the ZK-EDB\@.
-        \item \texttt{receivedBlockID} the ID of the received memory block.
-        \item \texttt{round} the current round number.
+        \item \texttt{ownerAddr}: the ALGO address of the ZK-EDB\@.
+        \item \texttt{receivedBlockID}: the ID of the received memory block.
+        \item \texttt{round}: the current round number.
     \end{itemize}
@@
 
-    To incentivize ZK-EDB servers, every round a lottery will be held and a predefined amount of ALGOs from
+    To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ALGOs from
     \texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
-    between all the \emph{winning tickets} of the lottery.
+    between all \emph{winning tickets} of the lottery.
 
@@
 
-    To run this lottery, In every round, based on the current block seed, a collection of \emph{valid} receipts will be
-    selected randomly as the \emph{winning} receipts. A receipt is \emph{valid} in the round \(r\) if:
+    To run this lottery, every round, based on the current block seed, a collection of \emph{valid} receipts will be
+    selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in the round \(r\) if:
 
@@
     \end{itemize}
-
     For selecting the winning receipts we could use a random generator:
-
     \begin{verbatim}
-        if random(seed|validatorPK|receivedBlockID) < winProbability
+IF random(seed|validatorPK|receivedBlockID) < winProbability THEN
     the receipt issued by validatorPK for receivedBlockID is a winner
     \end{verbatim}
-
     \begin{itemize}
@@
 
-    Every k round the protocol chooses a price per byte for the AVM memory. When a smart contract executes a heap
+    Every \(k\) round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
     allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ALGO
@@
 
-    To determine the price of AVM memory, Every k round, the protocol calculates \texttt{dbFee} and
-    \texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees and
+    To determine the price of AVM memory, Every \(k\) round, the protocol calculates \texttt{dbFee} and
+    \texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
     \texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
-    the protocol considers the total size of all memory pages that were accessed for either reading or writing during
-    some time. These two values will be calculated for the last k rounds and the price per byte of the AVM memory
-    will be a linear function of \texttt{dbFee/memTraffic}
+    the protocol considers the total size of all the memory pages that were accessed for either reading or writing
+    during a time period. These two values will be calculated for the last \(k\) rounds and the price per byte of
+    AVM memory will be a linear function of \texttt{dbFee/memTraffic}
 
@@
     de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
-    calculates the refunded amount based on the average price the smart had paid for that allocated memory. This will
-    prevent smart contracts from profit taking by trading memory with the protocol.
+    calculates the refunded amount based on the average price the smart contract had paid for that allocated memory.
+    This will prevent smart contracts from profit taking by trading memory with the protocol.
 
     \section{Concurrency}\label{sec:concurrency}
+
+    \subsection{Memory Dependency Graph}\label{subsec:memory-dependency-graph}
 
@@
     A transaction can change the AVM state by modifying either the code area or the AVM heap. In Algorand, all
-    transactions declare the list of memory blocks they want to read or write. This will enable us to determine the
+    transactions declare the list of memory locations they want to read or write. This will enable us to determine the
     independent sets of transactions which can be validated in parallel. To do so, we define the \emph{memory dependency
@@
         \item Every vertex in \(G\) corresponds to a transaction.
-        \item Vertices \(u\) and \(v\) are adjacent in \(G\) if and only if \(u\) has a memory block \(B\) in its
-        writing list and \(v\) has \(B\) in either its writing list or reading list.
+        \item Vertices \(u\) and \(v\) are adjacent in \(G\) if and only if \(u\) has a memory location \(L\) in its
+        writing list and \(v\) has \(L\) in either its writing list or reading list.
     \end{itemize}
@@
     transactions, a proposer is incentivized enough to find a good graph coloring.
+
+    \subsection{Concurrent Counters}\label{subsec:concurrent-counters}
+
+    We know that in Algorand every transaction needs to transfer its proposed fee to the \texttt{feeSink} account
+    first. This will essentially make every transaction a reader and a writer of the balance record of the
+    \texttt{feeSink} account. As a result, all transactions will be dependant and parallelism will be completely
+    impossible. Actually, any account that is highly active, for example the account of an exchange or a payment
+    processor, can become a concurrency bottleneck in our system which makes many transactions dependant.
+
+    This problem can be easily solved by using a concurrent counter (CC) for storing the balance of these kind of
+    accounts. A concurrent counter is a data structure which improves concurrency by using multiple memory locations for
+    storing a single counter. The value of the concurrent current is equal to the sum of the sub counters and it can
+    be incremented or decremented by incrementing/decrementing any of the sub counters. This way a concurrent
+    counter trades concurrency with memory usage.
+
+    A pseudocode for implementing a concurrent counter (CC) which returns an error when the value of the counter
+    becomes negative, follows:
+
+    \begin{samepage}
+        \begin{verbatim}
+INCREMENT(CC, value, seed)
+    i = seed MOD CC.size
+    ATOMIC_INCREMENT(CC.cell[i], value)
+
+DECREMENT(CC, value, seed, attempt)
+    IF attempt = CC.size THEN
+        restore CC by adding back the subtracted value to CC
+        RETURN ERROR
+    i = seed MOD CC.size
+    i = (i + attempt) MOD CC.size
+    IF CC.cell[i] >= value THEN
+        ATOMIC_DECREMENT(CC.cell[i], value)
+    ELSE
+        ATOMIC_SET(CC.cell[i], 0)
+        DECREMENT(CC, value - CC.cell[i], seed, attempt + 1)
+        \end{verbatim}
+    \end{samepage}
+
+    Concurrent counter data structure is a part of the AVM standard library, and any smart contract can use this data
+    structure for storing the balance of highly active accounts.
+
+    \subsection{Memory Chunks}\label{subsec:memory-chunks}
+
+    In order to further increase the concurrency level of Algorand, we can divide the AVM memory into \emph{chunks}.
+    Each memory chunk can be persisted using a different ZK-EDB, hence having a different commitment. Then, the
+    consensus on new values of different commitments can be done by different voting committees.
+
+    Any transaction needs to be validated only by voting committee of the memory chunk that it modifies. So, if we
+    choose chunks in a way that most transactions only modify memory locations of one chunk, the transactions of a
+    block can be divided between voting committees and be validated in parallel.
+
+    \emph{If a transaction modifies multiple chunks it will be validated by voting committees of all chunks it
+    modifies.}
+
+    Because the voting committees are selected by random sampling, by choosing large enough samples we can make sure
+    that having multiple voting committees will not change the security properties of our agreement protocol.
 
@@
     Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
-    degree of
-    weighting decrease, A higher \(\alpha\) discounts older observations faster.
+    degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
 
@@
     execution without accessing a ZK-EDB. Smart contract oracles can be used for reporting useful information about
-    \texttt{avmCall} transactions such as accessed AVM heap pages or code area blocks, exact amount of execution cost,
-    and soon.
+    \texttt{avmCall} transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
+    and so on.
 

### 2021-03-13 tex/avm.tex
Change a name
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-\title{The Algorand Virtual Machine}
+\title{The Argennon Virtual Machine}
 \author{aybehrouz}
@@
 
-    The Algorand Virtual Machine (AVM) is an abstract computing machine for executing Algorand's smart contracts. It
+    The Argennon Virtual Machine (AVM) is an abstract computing machine for executing Argennon's smart contracts. It
     is designed in a way that it could be efficiently implemented in either hardware or software.
 
-    \section{The Structure of the Algorand Virtual Machine}\label{sec:the-structure-of-the-algorand-virtual-machine}
+    \section{The Structure of the Argennon Virtual Machine}\label{sec:the-structure-of-the-algorand-virtual-machine}
 
@@
 
-    The Algorand Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
-    and does not have to be done by the Algorand Virtual Machine itself.
+    The Argennon Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
+    and does not have to be done by the Argennon Virtual Machine itself.
 
-    The instruction set of the Algorand Virtual Machine distinguishes its operand types using instructions intended to
+    The instruction set of the Argennon Virtual Machine distinguishes its operand types using instructions intended to
     operate on values of specific types. For instance, \texttt{iadd} assumes that its operands are two 64-bit integers.
@@
 
-    The Algorand Virtual Machine has a byte addressable memory which is divided into separate segments. Every segment
+    The Argennon Virtual Machine has a byte addressable memory which is divided into separate segments. Every segment
     belongs to a smart contract that has a unique \texttt{applicationID}. The AVM always has a single working memory
@@
 
-    An Algorand Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
+    An Argennon Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
     performed, followed by zero or more operands supplying arguments or data that are used by the operation.
@@
 
-    The Algorand Virtual Machine has three types of method invocation:
+    The Argennon Virtual Machine has three types of method invocation:
     \begin{itemize}
@@
         the current memory segment to the segment of the invoked smart contract.
-        \item \texttt{invoke\_native} invokes a method that is not hosted by the Algorand virtual machine. By this
+        \item \texttt{invoke\_native} invokes a method that is not hosted by the Argennon virtual machine. By this
         instruction, high performance native methods of the hosting machine could become available to AVM smart
@@
 
-    Each time a method is invoked a new local frame and operand stack is created. The Algorand Virtual Machine uses
+    Each time a method is invoked a new local frame and operand stack is created. The Argennon Virtual Machine uses
     local frames to pass parameters on method invocation. On method invocation, any parameters are passed in
@@
     An exception is thrown programmatically using the \texttt{athrow} instruction. Exceptions can also be thrown by
-    various Algorand Virtual Machine instructions if they detect an abnormal condition. Some exceptions are not
+    various Argennon Virtual Machine instructions if they detect an abnormal condition. Some exceptions are not
     catchable and will always abort the execution of the smart contract.
@@
     from a user to another user, first we need to make sure that the sender has authorized this operation. The
-    Algorand virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
+    Argennon virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
     cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
@@
 
-    \emph{The Algorand virtual machine has no instructions for issuing cryptographic signatures.}
+    \emph{The Argennon virtual machine has no instructions for issuing cryptographic signatures.}
 
@@
     We use a ZK-EDB for storing the AVM heap. We include the commitment of the current state of this DB in every
-    block of the Algorand blockchain, so ZK-EDB servers need not be trusted servers.
+    block of the Argennon blockchain, so ZK-EDB servers need not be trusted servers.
 
@@
 
-    Algorand has four types of transaction:
+    Argennon has four types of transaction:
 
@@
         AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
-        assets, including ALGOs, is done by these transactions.
+        assets, including ARGs, is done by these transactions.
         \item \texttt{installApp} installs an AVM smart contract and determines the update policy of the smart
@@
 
-    All types of Algorand transactions contain an \texttt{invoke\_external} instruction which calls a special method
-    from ALGO smart contract that transfers the proposed fee of the transaction in ALGOs from a sender account to the
+    All types of Argennon transactions contain an \texttt{invoke\_external} instruction which calls a special method
+    from ARG smart contract that transfers the proposed fee of the transaction in ARGs from a sender account to the
     fee sink accounts.
@@
 
-    Every block of the Algorand blockchain corresponds to a set of transactions. We store the commitment of this
+    Every block of the Argennon blockchain corresponds to a set of transactions. We store the commitment of this
     transaction set in every block, but we don't keep the set itself. To be able to detect replay attacks, we require
@@
 
-    The only information that Algorand nodes are required to store is \textbf{the most recent block} of the Algorand
-    blockchain. Every block of the Algorand blockchain contains the following information:
+    The only information that Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
+    blockchain. Every block of the Argennon blockchain contains the following information:
 
@@
     For confirming a new block, nodes that are not validators only need to verify the block certificate. For
-    verifying a block certificate, a node needs to know the ALGO balances of validators, but it doesn't need to
+    verifying a block certificate, a node needs to know the ARG balances of validators, but it doesn't need to
     emulate the AVM execution.
@@
     On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
-    execution of the Algorand virtual machine. To do so, first they retrieve all heap pages and code area blocks they
+    execution of the Argennon virtual machine. To do so, first they retrieve all heap pages and code area blocks they
     need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
@@
 
-    Every transaction in the Algorand blockchain starts with an \texttt{invoke\_external} instruction which calls a
-    special method from ALGO smart contract. This method will transfer the proposed fee of the transaction in ALGOs
-    from a sender account to the fee sink accounts. Algorand has two fee sink accounts: \texttt{execFeeSink} collects
+    Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
+    special method from ARG smart contract. This method will transfer the proposed fee of the transaction in ARGs
+    from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
     execution fees and \texttt{dbFeeSink} collects fees for ZK-EDBs. The Protocol decides how to distribute the
@@
     \begin{itemize}
-        \item \texttt{ownerAddr}: the ALGO address of the ZK-EDB\@.
+        \item \texttt{ownerAddr}: the ARG address of the ZK-EDB\@.
         \item \texttt{receivedBlockID}: the ID of the received memory block.
@@
 
-    To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ALGOs from
+    To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ARGs from
     \texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
@@
     A possible choice for the challenge solution could be the cryptographic hash of the content of the challenge
-    memory block combined with the ZK-EDB ALGO address: \texttt{hash(challenge.content|ownerAddr)}
+    memory block combined with the ZK-EDB ARG address: \texttt{hash(challenge.content|ownerAddr)}
 
@@
     Every \(k\) round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
-    allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ALGO
+    allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
     address of the smart contract.
@@
 
-    Every block of the Algorand blockchain contains a list of transactions. This list is an ordered list and the
+    Every block of the Argennon blockchain contains a list of transactions. This list is an ordered list and the
     effect of its contained transactions must be applied to the AVM state sequentially as they appear in the ordered
@@
 
-    A transaction can change the AVM state by modifying either the code area or the AVM heap. In Algorand, all
+    A transaction can change the AVM state by modifying either the code area or the AVM heap. In Argennon, all
     transactions declare the list of memory locations they want to read or write. This will enable us to determine the
@@
 
-    We know that in Algorand every transaction needs to transfer its proposed fee to the \texttt{feeSink} account
-    first. This will essentially make every transaction a reader and a writer of the balance record of the
-    \texttt{feeSink} account. As a result, all transactions will be dependant and parallelism will be completely
-    impossible. Actually, any account that is highly active, for example the account of an exchange or a payment
-    processor, can become a concurrency bottleneck in our system which makes many transactions dependant.
+    We know that in Argennon every transaction needs to transfer its proposed fee to the \texttt{feeSink} account
+    first. This essentially makes every transaction a reader and a writer of the memory location which stores the
+    balance record of the \texttt{feeSink} account. As a result, all transactions in Argennon will be dependant and
+    parallelism will be completely impossible. Actually, any account that is highly active, for example the account
+    of an exchange or a payment processor, could become a concurrency bottleneck, making all transactions that interact
+    with them dependant.
 
@@
     accounts. A concurrent counter is a data structure which improves concurrency by using multiple memory locations for
-    storing a single counter. The value of the concurrent current is equal to the sum of the sub counters and it can
-    be incremented or decremented by incrementing/decrementing any of the sub counters. This way a concurrent
+    storing a single counter. The value of the concurrent counter is equal to the sum of its sub counters and it can
+    be incremented or decremented by incrementing/decrementing any of the sub counters. This way, a concurrent
     counter trades concurrency with memory usage.
@@
     ELSE
+        remaining = value - CC.cell[i]
         ATOMIC_SET(CC.cell[i], 0)
-        DECREMENT(CC, value - CC.cell[i], seed, attempt + 1)
+        DECREMENT(CC, remaining, seed, attempt + 1)
         \end{verbatim}
     \end{samepage}
+
+    It should be noted that in a blockchain application we don't have concurrent threads and therefore we don't need
+    atomic functions. For usage in a smart contract, the atomic functions of this pseudocode can be implemented like
+    normal functions.
 
@@
 
-    In order to further increase the concurrency level of Algorand, we can divide the AVM memory into \emph{chunks}.
+    In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{chunks}.
     Each memory chunk can be persisted using a different ZK-EDB, hence having a different commitment. Then, the
@@
 
-    \emph{If a transaction modifies multiple chunks it will be validated by voting committees of all chunks it
+    \emph{If a transaction modifies multiple chunks it must be validated by the voting committees of all chunks it
     modifies.}
@@
 
-    To mitigate this problem, for calculating a user's stake at the time step \(t\), instead of using the raw ALGO
+    To mitigate this problem, for calculating a user's stake at the time step \(t\), instead of using the raw ARG
     balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
-    ALGO balance:
+    ARG balance:
     \[
@@
         \item \(S_{u,t}\) is the stake of the user \(u\) at the time step \(t\).
-        \item \(B_{u,t}\) is the ALGO balance of the user \(u\) at the time step \(t\).
+        \item \(B_{u,t}\) is the ARG balance of the user \(u\) at the time step \(t\).
         \item \(Trust_{u,t}\) is an estimated trust value for the user \(u\) at the time step \(t\).
@@
     number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
-    exponential moving average of the ALGO balance of the user \(u\) at the time step \(t\).
+    exponential moving average of the ARG balance of the user \(u\) at the time step \(t\).
 
-    In our system a user who held ALGOs and participated in the consensus for a long time is more trusted
+    In our system a user who held ARGs and participated in the consensus for a long time is more trusted
     than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
-    amount of ALGOs, also needs to hold them for a long period of time before being able to attack the system.
+    amount of ARGs, also needs to hold them for a long period of time before being able to attack the system.
 
@@
     This equation shows that when the balance of users is not changing over time the total stake of the system is the
-    exponential average of the total ALGOs of the system. Consequently, when we shift the time for \(m\) steps, we can
+    exponential average of the total ARGs of the system. Consequently, when we shift the time for \(m\) steps, we can
     calculate the new total stake of the system from the following equation:

### 2021-03-18 tex/avm.tex
Re-add "Concurrency" section
This commit introduces the idea of execution DAG.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 %\usepackage{amsmath}
-
+\usepackage{graphicx}
+\graphicspath{ {../img/} }
 \usepackage{hyperref}
+\usepackage{amsfonts}
+\usepackage[ruled,boxed]{algorithm2e}
+\usepackage{algpseudocode}
+\usepackage{listings}
+\usepackage{amssymb}
+\lstset{
+%  basicstyle=\ttfamily,
+    mathescape
+}
 %\binoppenalty=99999
@@
     effect of its contained transactions must be applied to the AVM state sequentially as they appear in the ordered
-    list. The ordering of transactions in a block is solely chosen by the block proposer.
+    list. This ordering is solely chosen by the block proposer, and users should not have any assumption about
+    the ordering of transactions in a block.
 
-    \emph{Users should not have any assumption about the ordering of transactions in a block.}
-
-    The fact that block transactions constitute a sequential list, does not mean they can not be validated
-    concurrently. Many transactions are actually independent and the order of their execution does not matter. These
-    transactions can be safely validated in parallel by validators.
+    The fact that block transactions constitute a sequential list, does not mean they can not be executed and applied
+    to the AVM state concurrently. Many transactions are actually independent and the order of their execution does not
+    matter. These transactions can be safely validated in parallel by validators.
 
@@
     transactions declare the list of memory locations they want to read or write. This will enable us to determine the
-    independent sets of transactions which can be validated in parallel. To do so, we define the \emph{memory dependency
-    graph} as follows:
+    independent sets of transactions which can be executed in parallel. To do so, we define the \emph{memory dependency
+    graph} \(G_d\) as follows:
 
     \begin{itemize}
-        \item \(G\) is an undirected graph.
-        \item Every vertex in \(G\) corresponds to a transaction.
-        \item Vertices \(u\) and \(v\) are adjacent in \(G\) if and only if \(u\) has a memory location \(L\) in its
+        \item \(G_d\) is an undirected graph.
+        \item Every vertex in \(G_d\) corresponds to a transaction and vice versa.
+        \item Vertices \(u\) and \(v\) are adjacent in \(G_d\) if and only if \(u\) has a memory location \(L\) in its
         writing list and \(v\) has \(L\) in either its writing list or reading list.
@@
 
-    If we consider a proper vertex coloring of \(G\), every color class will give us an independent set of
-    transactions which can be validated concurrently. To achieve the highest parallelization, we need to color \(G\)
-    with minimum number of colors. Thus, the \emph{chromatic number} of the memory dependency graph shows how good a
+    If we consider a proper vertex coloring of \(G_d\), every color class will give us an independent set of
+    transactions which can be executed concurrently. To achieve the highest parallelization, we need to color \(G_d\)
+    with minimum number of colors. The \emph{chromatic number} of the memory dependency graph thus shows how good a
     transaction set could be run concurrently.
@@
     Graph coloring is computationally NP-hard. However, in our use case we don't need to necessarily find an optimal
-    solution. An approximate greedy algorithm will perform well enough in most circumstances. The block proposer is the
-    one who is responsible for solving the graph coloring problem and a proposed block must determine the independent
-    sets of transactions which can be run in parallel safely. Since with better parallelization a block can contain more
-    transactions, a proposer is incentivized enough to find a good graph coloring.
+    solution. An approximate greedy algorithm will perform well enough in most circumstances.
+
+    After constructing the memory dependency graph of a transaction set, we can use it to construct the
+    \emph{execution DAG} of transactions. The execution DAG of a transaction set \(T\) is a directed acyclic
+    graph \(G_e\) which has the \emph{execution invariance} property:
+    \begin{itemize}
+        \item Every vertex in \(G_e\) corresponds to a transaction in \(T\) and vice versa.
+        \item Applying the transactions of \(T\) in any order that \emph{respects} \(G_e\) will result in
+        the same AVM state.
+        \begin{itemize}
+            \item An ordering of transactions of \(T\) respects \(G_e\) if for every directed edge \((u,v)\) in \(G_e\)
+            the transaction \(u\) comes before the transaction \(v\) in the ordering.
+        \end{itemize}
+    \end{itemize}
+
+    Having the execution DAG of a set of transactions, using Algorithm~\ref{alg:exec_dag}, we can apply the transaction
+    set to the AVM state concurrently, using multiple processor, while we can make sure that the resulted AVM state will
+    always be the same no matter how many processor we have used.
+
+    %\includegraphics[width=17cm]{../img/Alg1.png}
+    \begin{algorithm}
+        \DontPrintSemicolon
+        \SetKwData{Ready}{$R_e$}\SetKwData{V}{$v_f$}\SetKwData{Graph}{$G_e$}\SetKwData{Vertices}{$V$}\SetKwData
+        {Txns}{$T$}
+        \KwData{The execution dag $\Graph = (\Vertices,E)$ of a transaction set \Txns}
+        \KwResult{The state of the AVM after applying \Txns with any ordering respecting \Graph}
+        \BlankLine
+        \Ready $\gets$ the set of all vertices of \Vertices with in degree 0\;
+        \While{$\Vertices \neq \varnothing$}
+        {
+            wait until a new free processor is available\;
+            \If{the execution of a transaction was finished}
+            {
+                remove the vertex of the finished transaction \V from \Graph\;
+                \For{each vertex $u \in Adj[\V]$}
+                {
+                    \If{$u$ has zero in degree}
+                    {
+                        $\Ready \gets \Ready \cup u$\;
+                    }
+                }
+            }
+            \If{$\Ready \neq \varnothing$}
+            {
+                remove a vertex from \Ready and assign it to a processor\;
+            }
+        }
+        \caption{executing DAG transactions}\label{alg:exec_dag}
+    \end{algorithm}
+
+    By replacing every undirected edge of a memory dependency graph with a directed edge in such a way that the
+    resulted graph has no cycles, we will obtain a valid execution DAG. Thus, from a memory dependency graph different
+    execution DAGs can be constructed with different levels of parallelization ability.
+
+    If we assume that we have unlimited number of processors and all transactions take equal time for executing, it
+    can be shown that by providing a minimal graph coloring to Algorithm~\ref{alg:gen_dag} as input, the resulted
+    DAG will be optimal, in the sense that it results in the minimum overall execution time.
+
+    %\includegraphics[width=17cm]{../img/Alg2.png}
+    \begin{algorithm}
+        \DontPrintSemicolon
+        \SetKwData{Txns}{$T$}\SetKwData{Gd}{$G_d=(V_d,E_d)$}
+        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
+        \Input{The memory dependency graph \Gd of a transaction set \Txns\\A proper coloring of $G_d$}
+        \Output{An execution dag $G_e=(V_e,E_e)$ for the transaction set \Txns}
+        \BlankLine
+        $V_e \gets V_d$\;
+        $E_e \gets \varnothing$\;
+        define a total order on colors of $G_d$\;
+        \For{each edge $\{u,v\} \in E_d$}
+        {
+            \eIf{$color[u] < color[v]$}
+            {
+                $E_e \gets E_e \cup (u,v)$\;
+            }{
+                $E_e \gets E_e \cup (v,u)$\;
+            }
+        }
+        \caption{Constructing an execution DAG}\label{alg:gen_dag}
+    \end{algorithm}
+
+    The block proposer is responsible for proposing an efficient execution DAG alongside his proposed block which will
+    determine the ordering of block transactions and help validators to validate transactions in parallel. Since with
+    better parallelization a block can contain more transactions, a proposer is incentivized enough to find a good
+    execution DAG for transactions.
 
@@
 
-    We know that in Argennon every transaction needs to transfer its proposed fee to the \texttt{feeSink} account
-    first. This essentially makes every transaction a reader and a writer of the memory location which stores the
-    balance record of the \texttt{feeSink} account. As a result, all transactions in Argennon will be dependant and
+    We know that in Argennon every transaction needs to transfer its proposed fee to the \texttt{feeSink} accounts
+    first. This essentially makes every transaction a reader and a writer of the memory locations which store the
+    balance record of the \texttt{feeSink} accounts. As a result, all transactions in Argennon will be dependant and
     parallelism will be completely impossible. Actually, any account that is highly active, for example the account
-    of an exchange or a payment processor, could become a concurrency bottleneck, making all transactions that interact
-    with them dependant.
+    of an exchange or a payment processor, could become a concurrency bottleneck of the system, making all transactions
+    which interact with them dependant.
 
-    This problem can be easily solved by using a concurrent counter (CC) for storing the balance of these kind of
+    This problem can be easily solved by using a concurrent counter (CC) for storing the balance of this type of
     accounts. A concurrent counter is a data structure which improves concurrency by using multiple memory locations for
@@
 
-    \begin{samepage}
-        \begin{verbatim}
-INCREMENT(CC, value, seed)
-    i = seed MOD CC.size
-    ATOMIC_INCREMENT(CC.cell[i], value)
-
-DECREMENT(CC, value, seed, attempt)
-    IF attempt = CC.size THEN
-        restore CC by adding back the subtracted value to CC
-        RETURN ERROR
-    i = seed MOD CC.size
-    i = (i + attempt) MOD CC.size
-    IF CC.cell[i] >= value THEN
-        ATOMIC_DECREMENT(CC.cell[i], value)
-    ELSE
-        remaining = value - CC.cell[i]
-        ATOMIC_SET(CC.cell[i], 0)
-        DECREMENT(CC, remaining, seed, attempt + 1)
-        \end{verbatim}
-    \end{samepage}
+    %\includegraphics[width=17cm]{../img/Alg3.png}
+    \begin{algorithm}
+        \DontPrintSemicolon
+        \SetKwData{CC}{Counter}
+        \SetKwFunction{Inc}{Increment}\SetKwFunction{Dec}{Decrement}\SetKwFunction{AtomInc}{AtomicIncrement}
+        \SetKwFunction{AtomDec}{AtomicDecrement}\SetKwFunction{AtomSet}{AtomicSet}
+        \SetKwProg{Fn}{Function}{}{}
+        \BlankLine
+        \Fn{\Inc{\CC, value, seed}}
+        {
+            $i \gets seed \bmod \CC.size$\;
+           \AtomInc{$\CC.cell[i]$, value}\;
+        }
+        \BlankLine
+        \Fn{\Dec{\CC, value, seed, attempt}}
+        {
+            \If {attempt = \CC.size}
+            {
+                restore \CC by adding back the subtracted value\;
+                \KwRet{Error}\;
+            }
+            $i \gets seed \bmod \CC.size$\;
+            $i \gets (i + attempt) \bmod \CC.size$\;
+            \eIf {$\CC.cell[i] \geq value$}
+            {
+                \AtomDec{$\CC.cell[i]$, value}\;
+            }{
+                $r \gets value - \CC.cell[i]$\;
+                \AtomSet{$\CC.cell[i]$, $0$}\;
+                \Dec{\CC, r, seed, $attempt + 1$}\;
+            }
+        }
+        \caption{Concurrent counter}\label{alg:CC}
+    \end{algorithm}
 
@@
     In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{chunks}.
-    Each memory chunk can be persisted using a different ZK-EDB, hence having a different commitment. Then, the
-    consensus on new values of different commitments can be done by different voting committees.
+    Each memory chunk can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
+    consensus on new values of the commitment of any chunk can be achieved by different voting committees.
 
-    Any transaction needs to be validated only by voting committee of the memory chunk that it modifies. So, if we
-    choose chunks in a way that most transactions only modify memory locations of one chunk, the transactions of a
-    block can be divided between voting committees and be validated in parallel.
+    If a transaction does not modify a memory chunk and in the transaction ordering of the block it comes after
+    any transaction which modifies that chunk, then the execution of that transaction is not needed for calculating
+    the new commitment of the chunk. Consequently, the voting committee of the memory chunk can safely ignore such a
+    transaction. The execution DAG of transactions can be used for finding and pruning these transactions as
+    we see in Algorithm~\ref{alg:prune_dag}.
 
-    \emph{If a transaction modifies multiple chunks it must be validated by the voting committees of all chunks it
-    modifies.}
+    %\includegraphics[width=17cm]{../img/Alg4.png}
+    \begin{algorithm}
+        \DontPrintSemicolon
+        \SetKwData{V}{$v_0$}\SetKwData{Graph}{$G_e$}\SetKwData{Chunk}{$C_k$}\SetKwData{Txns}{$T$}
+        \SetKwFunction{RPrune}{RecursivePrune}
+        \SetKwProg{Fn}{Function}{}{}
+        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
+        \Input{An execution dag and a memory chunk \Chunk}
+        \Output{The pruned execution dag based on \Chunk}
+        \BlankLine
+        \For{each vertex \V with out degree $0$}
+        {
+            \RPrune{\V}\;
+        }
+        \BlankLine
+        \Fn{\RPrune{\V}}
+        {
+            \If{$Adj[\V] = \varnothing$ {\bf and} \V does not modify \Chunk}
+            {
+                remove \V from \Graph\;
+                \For{each $u$ such that $\V \in Adj[u]$}
+                {
+                    \RPrune{u}\;
+                }
+            }
+        }
+        \caption{Pruning the execution DAG}\label{alg:prune_dag}
+    \end{algorithm}
+
+    If we choose chunks in a way that most transactions only modify memory locations of one chunk,
+    likely the transactions of a block are divided between voting committees and are validated in parallel.
 
     Because the voting committees are selected by random sampling, by choosing large enough samples we can make sure
-    that having multiple voting committees will not change the security properties of our agreement protocol.
+    that having multiple voting committees will not change the security properties of the Argennon agreement protocol.
 
     \section{Consensus}\label{sec:consensus}
+
+    The consensus protocol of Argennon is similar to Algorand with a few minor improvements.
 

### 2021-03-18 tex/avm.tex
Make images a bit smaller
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Remove htm directory from git
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-    %\includegraphics[width=17cm]{../img/Alg1.png}
+    %\includegraphics[width=17cm]{../img/Alg1s.png}
     \begin{algorithm}
@@
 
-    %\includegraphics[width=17cm]{../img/Alg2.png}
+    %\includegraphics[width=17cm]{../img/Alg2s.png}
     \begin{algorithm}
@@
 
-    This problem can be easily solved by using a concurrent counter (CC) for storing the balance of this type of
+    This problem can be easily solved by using a concurrent counter for storing the balance of this type of
     accounts. A concurrent counter is a data structure which improves concurrency by using multiple memory locations for
@@
 
-    A pseudocode for implementing a concurrent counter (CC) which returns an error when the value of the counter
+    A pseudocode for implementing a concurrent counter which returns an error when the value of the counter
     becomes negative, follows:
 
-    %\includegraphics[width=17cm]{../img/Alg3.png}
+    %\includegraphics[width=17cm]{../img/Alg3s.png}
     \begin{algorithm}
@@
 
-    %\includegraphics[width=17cm]{../img/Alg4.png}
+    %\includegraphics[width=17cm]{../img/Alg4s.png}
     \begin{algorithm}

### 2021-03-18 tex/avm.tex
Remove htm directory from git and fix readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 % Preamble
-\documentclass[11pt, A4]{report}
+\documentclass[a4paper]{report}
 
@@
 %\usepackage{amsmath}
+%\usepackage[a4paper]{geometry}
 \usepackage{graphicx}
-\graphicspath{ {../img/} }
 \usepackage{hyperref}
@@
     \maketitle
+    \tableofcontents
 
@@
     must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
-    a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance \textgreater{}\
-        maxLifeTime} or if a signature of the same user with a bigger or equal nonce is already used
+    a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
+    if a signature of the same user with a bigger or equal nonce is already used
     (i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
@@
     from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
-    execution fees and \texttt{dbFeeSink} collects fees for ZK-EDBs. The Protocol decides how to distribute the
+    execution fees and \texttt{dbFeeSink} collects fees for ZK-EDB servers. The Protocol decides how to distribute the
     transaction fee between these two fee sink accounts.
@@
         \item Vertices \(u\) and \(v\) are adjacent in \(G_d\) if and only if \(u\) has a memory location \(L\) in its
-        writing list and \(v\) has \(L\) in either its writing list or reading list.
+        writing list and \(v\) has \(L\) in either its writing list or its reading list.
     \end{itemize}
@@
     transactions which can be executed concurrently. To achieve the highest parallelization, we need to color \(G_d\)
-    with minimum number of colors. The \emph{chromatic number} of the memory dependency graph thus shows how good a
+    with minimum number of colors. Thus, the \emph{chromatic number} of the memory dependency graph shows how good a
     transaction set could be run concurrently.
@@
 
-    After constructing the memory dependency graph of a transaction set, we can use it to construct the
-    \emph{execution DAG} of transactions. The execution DAG of a transaction set \(T\) is a directed acyclic
-    graph \(G_e\) which has the \emph{execution invariance} property:
+    After constructing the memory dependency graph, we can use it to construct the
+    \emph{execution DAG} of transactions. The execution DAG of transaction set \(T\) is a directed acyclic
+    graph \(G_e = (V_e,E_e)\) which has the \emph{execution invariance} property:
     \begin{itemize}
-        \item Every vertex in \(G_e\) corresponds to a transaction in \(T\) and vice versa.
-        \item Applying the transactions of \(T\) in any order that \emph{respects} \(G_e\) will result in
+        \item Every vertex in \(V_e\) corresponds to a transaction in \(T\) and vice versa.
+        \item Executing the transactions of \(T\) in any order that \emph{respects} \(G_e\) will result in
         the same AVM state.
         \begin{itemize}
-            \item An ordering of transactions of \(T\) respects \(G_e\) if for every directed edge \((u,v)\) in \(G_e\)
+            \item An ordering of transactions of \(T\) respects \(G_e\) if for every directed edge \((u,v) \in E_e\)
             the transaction \(u\) comes before the transaction \(v\) in the ordering.
@@
     Having the execution DAG of a set of transactions, using Algorithm~\ref{alg:exec_dag}, we can apply the transaction
-    set to the AVM state concurrently, using multiple processor, while we can make sure that the resulted AVM state will
+    set to the AVM state concurrently, using multiple processor, while we can be sure that the resulted AVM state will
     always be the same no matter how many processor we have used.
 
-    %\includegraphics[width=17cm]{../img/Alg1s.png}
+    %##\includegraphics[width=17cm]{../img/Alg1s.png}
     \begin{algorithm}
@@
         {Txns}{$T$}
-        \KwData{The execution dag $\Graph = (\Vertices,E)$ of a transaction set \Txns}
+        \KwData{The execution dag $\Graph = (\Vertices,E)$ of transaction set \Txns}
         \KwResult{The state of the AVM after applying \Txns with any ordering respecting \Graph}
@@
 
-    %\includegraphics[width=17cm]{../img/Alg2s.png}
+    %##\includegraphics[width=17cm]{../img/Alg2s.png}
     \begin{algorithm}
@@
         \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{The memory dependency graph \Gd of a transaction set \Txns\\A proper coloring of $G_d$}
+        \Input{The memory dependency graph \Gd of transaction set \Txns\\A proper coloring of $G_d$}
         \Output{An execution dag $G_e=(V_e,E_e)$ for the transaction set \Txns}
@@
 
-    The block proposer is responsible for proposing an efficient execution DAG alongside his proposed block which will
-    determine the ordering of block transactions and help validators to validate transactions in parallel. Since with
-    better parallelization a block can contain more transactions, a proposer is incentivized enough to find a good
-    execution DAG for transactions.
+    The block proposer is responsible for proposing an efficient execution DAG alongside his proposed block. This
+    execution DAG will determine the ordering of block transactions and help validators to validate transactions in
+    parallel. Since with better parallelization a block can contain more transactions, a proposer is incentivized enough
+    to find a good execution DAG for transactions.
 
@@
     parallelism will be completely impossible. Actually, any account that is highly active, for example the account
-    of an exchange or a payment processor, could become a concurrency bottleneck of the system, making all transactions
-    which interact with them dependant.
+    of an exchange or a payment processor, could become a concurrency bottleneck in our system which makes all
+    transactions interacting with them dependant.
 
-    This problem can be easily solved by using a concurrent counter for storing the balance of this type of
+    This problem can be easily solved by using a concurrent counter for storing the balance record of this type of
     accounts. A concurrent counter is a data structure which improves concurrency by using multiple memory locations for
@@
 
-    A pseudocode for implementing a concurrent counter which returns an error when the value of the counter
-    becomes negative, follows:
+    Algorithm~\ref{alg:CC} implements a concurrent counter which returns an error when the value of the counter
+    becomes negative.
 
-    %\includegraphics[width=17cm]{../img/Alg3s.png}
+    %##\includegraphics[width=17cm]{../img/Alg3s.png}
     \begin{algorithm}
@@
     Concurrent counter data structure is a part of the AVM standard library, and any smart contract can use this data
-    structure for storing the balance of highly active accounts.
+    structure for storing the balance record of highly active accounts.
 
@@
     Each memory chunk can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
-    consensus on new values of the commitment of any chunk can be achieved by different voting committees.
+    consensus on new values of the commitment of any chunk can be achieved by a different voting committee.
 
@@
     any transaction which modifies that chunk, then the execution of that transaction is not needed for calculating
-    the new commitment of the chunk. Consequently, the voting committee of the memory chunk can safely ignore such a
+    the new commitment of the chunk. Consequently, the voting committee of that memory chunk can safely ignore such a
     transaction. The execution DAG of transactions can be used for finding and pruning these transactions as
@@
 
-    %\includegraphics[width=17cm]{../img/Alg4s.png}
+    %##\includegraphics[width=17cm]{../img/Alg4s.png}
     \begin{algorithm}
@@
         \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{An execution dag and a memory chunk \Chunk}
+        \Input{An execution dag \Graph and a memory chunk \Chunk}
         \Output{The pruned execution dag based on \Chunk}
@@
                 remove \V from \Graph\;
-                \For{each $u$ such that $\V \in Adj[u]$}
+                \For{each $u$ such that edge $(u,\V)$ was in \Graph}
                 {
@@
         }
-        \caption{Pruning the execution DAG}\label{alg:prune_dag}
+        \caption{Pruning an execution DAG}\label{alg:prune_dag}
     \end{algorithm}
@@
     If we choose chunks in a way that most transactions only modify memory locations of one chunk,
-    likely the transactions of a block are divided between voting committees and are validated in parallel.
+    likely many transactions of a block only need to be validated by one voting committee and can be validated in
+    parallel by different committees.
 
@@
 
-    To mitigate this problem, for calculating a user's stake at the time step \(t\), instead of using the raw ARG
+    To mitigate this problem, for calculating a user's stake at time step \(t\), instead of using the raw ARG
     balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
@@
     \begin{itemize}
-        \item \(S_{u,t}\) is the stake of the user \(u\) at the time step \(t\).
-        \item \(B_{u,t}\) is the ARG balance of the user \(u\) at the time step \(t\).
-        \item \(Trust_{u,t}\) is an estimated trust value for the user \(u\) at the time step \(t\).
+        \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
+        \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
+        \item \(Trust_{u,t}\) is an estimated trust value for user \(u\) at time step \(t\).
     \end{itemize}
 
-    The agreement protocol, at the time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
+    The agreement protocol, at time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
     number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
-    exponential moving average of the ARG balance of the user \(u\) at the time step \(t\).
+    exponential moving average of the ARG balance of user \(u\) at time step \(t\).
 
@@
 
-    For calculating the exponential moving average of a user's balance at the time step \(t\), we can use the following
+    For calculating the exponential moving average of a user's balance at time step \(t\), we can use the following
     recursive formula:

### 2021-03-22 tex/avm.tex
Add memory spooling
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 \usepackage{amssymb}
+\usepackage{cleveref}
 \lstset{
@@
 
-    \subsection{ZK-EDB Servers}\label{subsec:zk-edb-servers}
+    \subsection{Incentives for ZK-EDB Servers}\label{subsec:zk-edb-servers}
 
@@
 
-    \subsection{Memory Allocation and De-allocation}\label{subsec:memory-allocation-and-de-allocation}
+    \subsection{Memory Allocation and De-allocation Fee}\label{subsec:memory-allocation-and-de-allocation}
 
@@
     effect of its contained transactions must be applied to the AVM state sequentially as they appear in the ordered
-    list. This ordering is solely chosen by the block proposer, and users should not have any assumption about
+    list. This ordering is solely chosen by the block proposer, and users should not have any assumptions about
     the ordering of transactions in a block.
@@
 
+    \subsection{Memory Spooling}\label{subsec:spooling}
+
+    When two transactions are dependant and they are connected with an edge \((u,v)\) in the execution DAG,
+    the transaction \(u\) needs to be run before the transaction \(v\). However, if \(v\) does not read any
+    memory locations that \(u\) modifies, we can run \(u\) and \(v\) in parallel. We just need to make sure
+    \(u\) does not see any changes \(v\) is making in AVM memory. This can be done by appropriate versioning
+    of the memory locations which is shared between \(u\) and \(v\). We call this method \emph{memory spooling}.
+    After enabling memory spooling between two transactions the edge connecting them can be safely removed from the
+    execution DAG\@.
+
     \subsection{Concurrent Counters}\label{subsec:concurrent-counters}
@@
         \SetKwFunction{Inc}{Increment}\SetKwFunction{Dec}{Decrement}\SetKwFunction{AtomInc}{AtomicIncrement}
-        \SetKwFunction{AtomDec}{AtomicDecrement}\SetKwFunction{AtomSet}{AtomicSet}
+        \SetKwFunction{AtomDec}{AtomicDecrement}\SetKwFunction{AtomSet}{AtomicSet}\SetKwFunction{Get}{GetValue}
+        \SetKwFunction{Acquire}{Lock.Acquire}\SetKwFunction{Release}{Lock.Release}
         \SetKwProg{Fn}{Function}{}{}
+        \Fn{\Get{\CC}}
+        {
+            $s \gets 0$\;
+            \Acquire{}\;
+            \For{$i \gets 0$ \KwTo $\CC.size - 1$}
+            {
+                $s \gets s + \CC.cell[i]$\;
+            }
+            \Release{}\;
+            \KwRet{s}\;
+        }
         \BlankLine

### 2021-03-24 tex/avm.tex
Add note boxes
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 \usepackage{amssymb}
-\usepackage{cleveref}
+\usepackage{tcolorbox}
 \lstset{
@@
 %\relpenalty=99999
+\newcommand{\note}[1] {
+    \begin{tcolorbox}[colframe=white,colback=white]
+        \emph{#1}
+        %\emph{#1}\medskip
+    \end{tcolorbox}
+}
 
@@
 
-    \emph{Only public methods can be invoked by \texttt{invoke\_external} instruction.}
+    \note{Only public methods can be invoked by \texttt{invoke\_external} instruction.}
 
@@
 
-    \emph{The reason behind this paged design is that the heap is usually persisted using a block device. A heap with
+    \note{The reason behind this paged design is that the heap is usually persisted using a block device. A heap with
     a paged structure could expose the underlying block based nature of the persistence layer to the application
@@
 
-    \emph{In the future, we may need to add special instructions for invoking interface and virtual methods...}
+    \note{In the future, we may need to add special instructions for invoking interface and virtual methods...}
 
@@
 
-    \emph{By using the \texttt{athrow} instruction properly, a programmer can make any method act like an atomic
+    \note{By using the \texttt{athrow} instruction properly, a programmer can make any method act like an atomic
     operation.}
@@
 
-    \emph{The Argennon virtual machine has no instructions for issuing cryptographic signatures.}
+    \note{The Argennon virtual machine has no instructions for issuing cryptographic signatures.}
 
@@
 
-    \emph{Unlike heap pages, the AVM is not aware of different blocks of code area.}
+    \note{Unlike heap pages, the AVM is not aware of different blocks of code area.}
 
@@
 
-    \emph{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDBs will receive the new
+    \note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDBs will receive the new
     block, and they will update their database by emulating the AVM execution.}
@@
 
-    \emph{A transaction always pays all of its proposed fee, no matter how much of its predefined resources were not
+    \note{A transaction always pays all of its proposed fee, no matter how much of its predefined resources were not
     used in the final emulation.}
@@
 
-    \emph{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}
+    \note{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}
 
@@
 
-    \emph{One ZK-EDB server could own multiple winning tickets in a round.}
+    \note{One ZK-EDB server could own multiple winning tickets in a round.}
 
     To run this lottery, every round, based on the current block seed, a collection of \emph{valid} receipts will be
-    selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in the round \(r\) if:
+    selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in round \(r\) if:
 
@@
 
-    \emph{The winners of the lottery were validators one round before the lottery round.}
+    \note{The winners of the lottery were validators one round before the lottery round.}
 
@@
 
-    The winning tickets of the lottery of the round \(r\) need to be included in the block of the round \(r\),
-    otherwise they will be considered expired. Validation and prize distribution for the winning tickets of the round
+    The winning tickets of the lottery of round \(r\) need to be included in the block of the round \(r\),
+    otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
     \(r\) will be done in the round \(r + 1\). This way, \textbf{the content of the challenge memory block could be

### 2021-03-29 tex/avm.tex
Update README.md
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
         \item A user can use \(C\) and \(p\) to verify that \(D\) is really associated with \(x\), and \(D\) is not
-        altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB.
+        altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB\@.
         \item Having \(p\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
@@
 
-    \note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDBs will receive the new
+    \note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDB servers will receive the new
     block, and they will update their database by emulating the AVM execution.}
+
+    \section{Consensus}\label{sec:consensus}
+
+    \note{It will be a randomized proof of stake which is under development....}
+
+    \subsection{Estimating A User's Stake}\label{subsec:estimating-a-user's-stake}
+
+    In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
+    of stake the user has in the system. Conventionally in these systems, for estimating a user's stake, we use the
+    amount of native system tokens the user is holding. Unfortunately, one problem with this approach is that a
+    strong attacker may be able to obtain a considerable amount of system tokens, for example by borrowing from a
+    DEFI application, and use this stake to attack the system.
+
+    To mitigate this problem, for calculating a user's stake at time step \(t\), instead of using the raw ARG
+    balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
+    ARG balance:
+    \[
+        S_{u,t} = \min (B_{u,t}, Trust_{u,t})
+    \]
+    Where:
+    \begin{itemize}
+        \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
+        \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
+        \item \(Trust_{u,t}\) is an estimated trust value for user \(u\) at time step \(t\).
+    \end{itemize}
+
+    The agreement protocol, at time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
+    number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
+    exponential moving average of the ARG balance of user \(u\) at time step \(t\).
+
+    In our system a user who held ARGs and participated in the consensus for a long time is more trusted
+    than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
+    amount of ARGs, also needs to hold them for a long period of time before being able to attack the system.
+
+    For calculating the exponential moving average of a user's balance at time step \(t\), we can use the following
+    recursive formula:
+    \[
+        M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
+    \]
+    Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
+    degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
+
+    Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
+    \(M_{u,t}\): (In the following equations the \(u\) subscript is dropped for simplicity)
+    \[
+        M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B
+    \]
+    Where:
+    \[
+        B = B_{k+1} = B_{k+2} = \dots = B_{t}
+    \]
+    We know that when \(|nx| \ll 1\) we can use the binomial approximation \({(1 + x)^n \approx 1 + nx}\). So, we can
+    further simplify this formula:
+    \[
+        M_{t} = M_{k} + (t - k) \alpha (B - M_{k})
+    \]
+
+    For choosing the value of \(\alpha\) we can consider the number of time steps that the trust value of a user needs
+    for reaching a specified fraction of his account balance. We know that for large \(n\) and \(|x| < 1\) we have
+    \((1 + x)^n \approx e^{nx}\), so by letting \(M_{u,k} = 0\) and \(n = t - k\) we can write:
+    \[
+        \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{B}\right)}{n}
+    \]
+    The value of \(\alpha\) for a desired configuration can be calculated by this equation. For instance, we could
+    calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8B\) and \(n\) equals to the
+    number of time steps of 10 years.
+
+    In our system a newly created account will not have voting power for some time, no matter how high its
+    balance is. While this is a desirable property, in case a large proportion of total system tokens are
+    transferred to newly created accounts, it can result in too much voting power for older accounts. This may decrease
+    the degree of decentralization in our system.
+
+    However, this situation is easily detectable by comparing the total stake of the system with the total balance of
+    users. If after confirming a block the total stake of the system goes too low and we have:
+    \[
+        \sum_{u}S_{u,t} < \gamma \sum_{u}B_{u,t}
+    \]
+    The protocol will perform a \emph{time shift} in the system: the time step of the system
+    will be incremented for \(m\) steps while no blocks will be confirmed. This will increase the value of \(M_{u,t}\)
+    for new accounts with a non-zero balance, giving them more influence in the agreement protocol.
+
+    For calculating the value of \(m\) which determines the amount of time shift in the system, we should note that when
+    \(B_{u,t} = B_{u, t-1} = B_u\), we can derive a simple recursive rule for the stake of a user:
+    \[
+        S_{u,t} = (1 - \alpha) S_{u,t-1} + \alpha B_u
+    \]
+    Therefore, we have:
+    \[
+        \sum_{u}S_{u,t} = (1 - \alpha) \sum_{u}S_{u,t - 1} + \alpha \sum_{u}B_u
+    \]
+    This equation shows that when the balance of users is not changing over time the total stake of the system is the
+    exponential average of the total ARGs of the system. Consequently, when we shift the time for \(m\) steps, we can
+    calculate the new total stake of the system from the following equation:
+
+    \[
+        \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
+    \]
+    Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
+    we can obtain \(m\) from the following formula, assuming \(\alpha\) is small enough:
+    \[
+        m = \frac{1}{\alpha} \ln \left(\frac{1 - \gamma}{1 - \lambda}\right)
+    \]
 
@@
 
-    \section{Consensus}\label{sec:consensus}
-
-    The consensus protocol of Argennon is similar to Algorand with a few minor improvements.
-
-    \subsection{Estimating A User's Stake}\label{subsec:estimating-a-user's-stake}
-
-    In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
-    of stake the user has in the system. Conventionally in these systems, for estimating a user's stake, we use the
-    amount of native system tokens the user is holding. Unfortunately, one problem with this approach is that a
-    strong attacker may be able to obtain a considerable amount of system tokens, for example by borrowing from a
-    DEFI application, and use this stake to attack the system.
-
-    To mitigate this problem, for calculating a user's stake at time step \(t\), instead of using the raw ARG
-    balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
-    ARG balance:
-    \[
-        S_{u,t} = \min (B_{u,t}, Trust_{u,t})
-    \]
-    Where:
-    \begin{itemize}
-        \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
-        \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
-        \item \(Trust_{u,t}\) is an estimated trust value for user \(u\) at time step \(t\).
-    \end{itemize}
-
-    The agreement protocol, at time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
-    number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
-    exponential moving average of the ARG balance of user \(u\) at time step \(t\).
-
-    In our system a user who held ARGs and participated in the consensus for a long time is more trusted
-    than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
-    amount of ARGs, also needs to hold them for a long period of time before being able to attack the system.
-
-    For calculating the exponential moving average of a user's balance at time step \(t\), we can use the following
-    recursive formula:
-    \[
-        M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
-    \]
-    Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
-    degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
-
-    Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
-    \(M_{u,t}\): (In the following equations the \(u\) subscript is dropped for simplicity)
-    \[
-        M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B
-    \]
-    Where:
-    \[
-        B = B_{k+1} = B_{k+2} = \dots = B_{t}
-    \]
-    We know that when \(|nx| \ll 1\) we can use the binomial approximation \({(1 + x)^n \approx 1 + nx}\). So, we can
-    further simplify this formula:
-    \[
-        M_{t} = M_{k} + (t - k) \alpha (B - M_{k})
-    \]
-
-    For choosing the value of \(\alpha\) we can consider the number of time steps that the trust value of a user needs
-    for reaching a specified fraction of his account balance. We know that for large \(n\) and \(|x| < 1\) we have
-    \((1 + x)^n \approx e^{nx}\), so by letting \(M_{u,k} = 0\) and \(n = t - k\) we can write:
-    \[
-        \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{B}\right)}{n}
-    \]
-    The value of \(\alpha\) for a desired configuration can be calculated by this equation. For instance, we could
-    calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8B\) and \(n\) equals to the
-    number of time steps of 10 years.
-
-    In our system a newly created account will not have voting power for some time, no matter how high its
-    balance is. While this is a desirable property, in case a large proportion of total system tokens are
-    transferred to newly created accounts, it can result in too much voting power for older accounts. This may decrease
-    the degree of decentralization in our system.
-
-    However, this situation is easily detectable by comparing the total stake of the system with the total balance of
-    users. If after confirming a block the total stake of the system goes too low and we have:
-    \[
-        \sum_{u}S_{u,t} < \gamma \sum_{u}B_{u,t}
-    \]
-    The protocol will perform a \emph{time shift} in the system: the time step of the system
-    will be incremented for \(m\) steps while no blocks will be confirmed. This will increase the value of \(M_{u,t}\)
-    for new accounts with a non-zero balance, giving them more influence in the agreement protocol.
-
-    For calculating the value of \(m\) which determines the amount of time shift in the system, we should note that when
-    \(B_{u,t} = B_{u, t-1} = B_u\), we can derive a simple recursive rule for the stake of a user:
-    \[
-        S_{u,t} = (1 - \alpha) S_{u,t-1} + \alpha B_u
-    \]
-    Therefore, we have:
-    \[
-        \sum_{u}S_{u,t} = (1 - \alpha) \sum_{u}S_{u,t - 1} + \alpha \sum_{u}B_u
-    \]
-    This equation shows that when the balance of users is not changing over time the total stake of the system is the
-    exponential average of the total ARGs of the system. Consequently, when we shift the time for \(m\) steps, we can
-    calculate the new total stake of the system from the following equation:
-
-    \[
-        \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
-    \]
-    Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
-    we can obtain \(m\) from the following formula, assuming \(\alpha\) is small enough:
-    \[
-        m = \frac{1}{\alpha} \ln \left(\frac{1 - \gamma}{1 - \lambda}\right)
-    \]
-
     \section{Smart Contract Oracle}\label{sec:smart-contract-oracle}

### 2021-04-21 tex/avm.tex
Redesign chapter one
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
+%! suppress = TooLargeSection
 %! Author = aybehrouz
@@
 % Preamble
-\documentclass[a4paper]{report}
+\documentclass[11pt, a4paper]{report}
 
@@
 
-    \chapter{Specification}\label{ch:specification}
+    \chapter{The Argennon Virtual Machine}\label{ch:AVM}
 
@@
 
-    The Argennon Virtual Machine (AVM) is an abstract computing machine for executing Argennon's smart contracts. It
-    is designed in a way that it could be efficiently implemented in either hardware or software.
+    The Argennon Virtual Machine (AVM) is an abstract computing machine for executing Argennon's smart contracts. The
+    Argennon Virtual Machine knows nothing of the Argennon blockchain, only of cleanup sessions and
+    execution sessions.
 
-    \section{The Structure of the Argennon Virtual Machine}\label{sec:the-structure-of-the-algorand-virtual-machine}
+    Execution sessions are separate sessions of executing smart contact's code by the Argennon Virtual Machine.
+    These sessions usually are correspondent to the concept of a transaction in a blockchain. (See
+    Section~\ref{subsec:execution-sessions})
 
-    ...
+    A cleanup session is the time slot in which the Argennon Virtual Machine performs cleanup and re-optimization of
+    its internal data structures in a "Stop the World" manner. Cleanup sessions usually are done as the final phase of
+    the block confirmation process in a blockchain. (See Section ...)
 
-    \subsection{Data Types}\label{subsec:data-types}
+    \section{Data Types}\label{sec:data-types}
 
@@
 
-    The instruction set of the Argennon Virtual Machine distinguishes its operand types using instructions intended to
-    operate on values of specific types. For instance, \texttt{iadd} assumes that its operands are two 64-bit integers.
+    The Argennon Virtual Machine operates on two kinds of types: primitive types and reference types. There are,
+    correspondingly, two kinds of values that can be stored in memory locations, passed as arguments, returned by
+    methods, and operated upon: primitive values and reference values.
 
-    \subsection{The \texttt{PC} Register}\label{subsec:the-pc-register}
+    Values of primitive or reference types need not be tagged or otherwise be inspectable to determine their types
+    at run time, or to be distinguished from values of other types. Instead, the instruction set of the Argennon
+    Virtual Machine distinguishes its operand types using instructions intended to operate on values of specific
+    types. For instance, \texttt{iadd64} assumes that its operands are two 64-bit signed integers or \texttt{rstore}
+    assumes that its operand is a reference type.
 
-    ...
+    A reference value may be the special \texttt{null} reference, a reference to no location, which will be denoted here
+    by \texttt{null}. The default value of a reference type is \texttt{null}.
 
-    \subsection{Call Stack}\label{subsec:call-stack}
+    The Argennon virtual machine does not have a fixed word size. Any instruction of the Argennon Virtual Machine has
+    its specific word size, and the addressable memory areas are byte addressable.
 
-    A call stack contains the information that is needed for restoring the state of the invoker of a method.
+    \section{Arithmetics}\label{sec:arithmetics}
 
-    \subsection{Memory Unit}\label{subsec:memory-unit}
+    All arithmetic operations in the Argennon Virtual machine are safe and there are strict overflow checks and
+    blah blah blah
 
-    The Argennon Virtual Machine has a byte addressable memory which is divided into separate segments. Every segment
-    belongs to a smart contract that has a unique \texttt{applicationID}. The AVM always has a single working memory
-    segment and memory locations outside its current working segment can not be accessed. The only instructions which
-    can change the current working segment are \texttt{invoke\_external}, \texttt{athrow} and return instructions.
+    \section{Architecture}\label{sec:arch}
 
-    Every smart contract has its own memory segment. Hence, there is no way for a smart contract to access another
-    smart contract's memory. Interaction between smart contracts is done using \texttt{invoke\_external} instruction,
-    and a smart contract can invoke methods of another smart contract by this instruction.
+    \subsection{The \texttt{pc} Register}\label{subsec:the-pc-register}
 
-    A memory segment consists of five data areas:
+    The Argennon Virtual Machine always has a single thread of execution and exactly has one \texttt{pc} register.
+    When the Argennon Virtual Machine is executing a method, if that method is not native, the \texttt{pc} register
+    contains the address of the AVM instruction currently being executed. If the method currently being executed
+    is native, the value of the Argennon Virtual Machine's \texttt{pc} register is undefined.
+
+    \subsection{Call Stack Queue}\label{subsec:call-stack}
+
+    Every AVM execution session has a queue of call stacks. A call stack contains all the information that is needed for
+    restoring the state, and continuing the execution after the method invocation completes.
+    This information is represented by a \texttt{CallInfo} struct:
+    \begin{verbatim}
+CallInfo {
+    applicationID,
+    methodID,
+    pc,
+    canModify,
+    localFrame,
+    operandStack
+}
+    \end{verbatim}
+    The \texttt{applicationID} field is the unique identifier that the AVM assigns to every smart contract,
+    \texttt{methodID} is the unique identifier of a method's bytecode and \texttt{canModify} is a flag
+    indicating whether a method is allowed to execute state modifying instructions or not.
+
+    Every method invocation has a corresponding \texttt{CallInfo} struct and when a method is invoked its
+    \texttt{CallInfo} struct is pushed onto the call stack. Hence, the top of the call stack always contains
+    the \texttt{CallInfo} struct of the current method. When a method invocation completes its call
+    information is popped from the call stack.
+
+    An AVM execution session will continue as long as there is a non-empty call stack in the call stack queue. When the
+    execution of the current call stack finishes the execution of the next call stack in the call stack queue starts.
+    It's possible that an AVM method invocation creates new call stacks. These newly created call stacks will be a part
+    of the current execution session and are added at the end of the call stack queue. A newly created call stack always
+    contains a single \texttt{CallInfo} struct. See Section~\ref{subsec:method-invocation} for more details.
+
+    \subsection{Run-Time Data Areas}\label{subsec:data-areas}
+
+    The Argennon Virtual Machine defines various run-time data areas that are used during an execution session
+    (i.e.\ transaction), Some of these data areas are persistent and are stored on the blockchain, Other data areas
+    are per execution session or per method invocation. These data areas are created when their context starts and
+    destroyed when their context ends.
+
+    The Argennon Virtual Machine has five run-time data areas:
     \begin{itemize}
-        \item Code Area
+        \item Method Area
         \item Constant Area
@@
     \end{itemize}
-    All data areas except operand stack, have their own address space which starts from 0. Operand stack is a
-    last-in-first-out (LIFO) stack and is not addressable. Every memory access instruction operates on its specific
+    All data areas except operand stack, have their own address space. Operand stack is a
+    last-in-first-out (LIFO) stack and is not addressable. Every AVM memory access instruction operates on its specific
     data areas.
 
-    \subsubsection{Code Area}
+    \subsubsection{Method Area}
 
-    The code area of a segment contains the byte-code of the smart contract which owns that segment. The AVM has no
-    instructions for manipulating the code area. \textbf{Installing, removing and updating smart contracts need to
-    be done externally.}
+    The method area contains the byte-code of any method which can be called by a non-native method invocation
+    instruction. In the Argennon Virtual Machine every method has a unique identifier:
+    \texttt{methodID}, and the method area is a map from method identifiers to their byte codes.
+
+    Instructions that modify the code area can only be run in privileged mode which means they can only be executed
+    by the \emph{root} smart contract. As a result, the access of smart contracts to the code area is essentially
+    read-only.
 
@@
 
-    The constant area of a segment contains several kinds of constants, ranging from user defined constants to method
-    address tables. A method address table stores method locations in the code area and their access type. The access
-    type of a method can be either \texttt{public} or \texttt{private}. The AVM has no instructions for modifying the
-    constant area.
+    Every smart contract has a single constant area which is stored in the AVM code area as a special type of
+    non-executable method. The constant area of a smart contract contains several kinds of constants, ranging from user
+    defined constants to method address tables. A method address table stores the list of method identifiers of a smart
+    contract and their access type. The access type of a method can be either \texttt{external} or \texttt{internal}.
+    Only external methods can be invoked by \texttt{invoke\_external} instruction.
 
-    \note{Only public methods can be invoked by \texttt{invoke\_external} instruction.}
+    \note{Instructions that modify the constant area can only be run in privileged mode.}
 
@@
 
-    Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too. AVM
-    instructions take operands from the operand stack, operate on them, and push the result back onto the operand
-    stack. An operand stack is destroyed when its owner method completes, whether that completion is normal or abrupt.
+    Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too. The
+    Argennon Virtual Machine is a stack machine and its instructions take operands from the operand stack, operate on
+    them, and push the result back onto the operand stack. An operand stack is destroyed when its owner method
+    completes, whether that completion is normal or abrupt.
 
@@
 
-    The heap of a segment is a persistent memory area which is divided into pages. Memory locations inside every page
-    have a separate address space that starts from 0, and each page can be referenced by an index that starts from 0. In
-    other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex, offset)}.
-    Different pages of a heap do not need to be equally sized.
+    The heap of the Argennon Virtual Machine is a persistent memory area which stores \emph{memory chunks}. A Memory
+    chunk is a byte addressable piece of memory which has a separate address space starting from 0. Each chunk has a
+    fixed size and different chunks need not be equally sized. Chunks are stored in the heap area and every chunk is
+    assigned a unique identifier: \texttt{chunkID}. As a result, the address of every memory location inside
+    the heap area can be considered as a pair: \texttt{(chunkID, offset)}.
 
-    \note{The reason behind this paged design is that the heap is usually persisted using a block device. A heap with
-    a paged structure could expose the underlying block based nature of the persistence layer to the application
-    layer. In this way, the compiler or the programmer could better optimize the code for the persistence layer.}
+    A smart contract can read any chunk stored in the heap by having its \texttt{chunkID}. However, it can only modify
+    those chunks that was created by itself. In other words, every memory chunk in the AVM heap has an owner. Only
+    the owner can modify a chunk while any one can read that chunk. In addition, for modifying a chunk the \texttt{canModify}
+    flag of the current method's \texttt{callInfo} struct must be true.
 
-    \section{Instruction Set Summary}\label{sec:instruction-set-summary}
+    \note{The reason behind this type of access control design is the fact that smart contract code is usually
+    immutable. That means if a smart contract does not implement a getter mechanism for some parts of its internal
+    data, this functionality can never be added later. Although the internal data is publicly available, there will
+    be no way for other smart contracts to use this data on-chain. This design eliminates the need for implementing
+    \emph{trivial} getters.}
+
+    The AVM Heap is able to save snapshots of its state and later restore them. This will enable the Argennon
+    Virtual Machine to have state reversion capability. The snapshot management of the heap area is done by
+    the following functions:
+    \begin{itemize}
+        \item \texttt{Save()} saves a snapshot of the current state of the heap and pushes it onto the
+        snapshot stack.
+        \item \texttt{Restore()} pops the snapshot stored on the top of the snapshot stack and restores it. The
+        current state of the heap will be lost.
+        \item \texttt{Discard()} pops the snapshot stored on the top of the snapshot stack and discards it. The current
+        state of the heap will be preserved.
+    \end{itemize}
+
+    \note{These functions are internal functions of the AVM. They are not instructions, and can not be called by smart
+    contracts.}
+
+    \section{Instruction Set Overview}\label{sec:instruction-set-overview}
 
@@
     performed, followed by zero or more operands supplying arguments or data that are used by the operation.
-    \textbf{The number and size of the operands are determined solely by the opcode.}
+    The number and size of the operands are determined solely by the opcode.
 
@@
 
-    The Argennon Virtual Machine has three types of method invocation:
+    The Argennon Virtual Machine has four types of method invocation instruction:
     \begin{itemize}
-        \item \texttt{invoke\_internal} invokes a method from the current running smart contract.
-        \item \texttt{invoke\_external} invokes a \texttt{public} method from another smart contract. It will change
-        the current memory segment to the segment of the invoked smart contract.
-        \item \texttt{invoke\_native} invokes a method that is not hosted by the Argennon virtual machine. By this
+        \item \texttt{invoke\_internal} invokes a method without changing the context of method execution. In other
+        words, \texttt{applicationID} and \texttt{canModify} fields of the invoked method in the
+        call stack will be the same as the invoker.
+        A smart contract can invoke any method by \texttt{invoke\_internal} instruction, even if that method is an
+        internal method of another smart contract. Since the invoked method will always be executed in the context of
+        the invoking smart contract, a smart contract will not be able to modify another smart contract state by
+        this instruction. This instruction facilitates the usage of libraries and code reuse.
+        \item \texttt{invoke\_external} invokes a method and changes the context of method execution to another
+        smart contract. Only external methods of another smart contract can be called by this instruction. For changing
+        the context of method execution this instruction sets the \texttt{applicationID} field of the invoked method
+        to the \texttt{applicationID} of the called smart contract. Then it searches the call stack and if
+        the \texttt{applicationID} of the called smart contract is found in the call stack it sets \texttt{canModify}
+        to \texttt{false}, otherwise \texttt{canModify} is set to \texttt{true}. As a result, if the
+        called smart contract is already called, the invoked method will not be able to modify the heap area.
+        \item \texttt{invoke\_native} invokes a method that is not hosted by the Argennon Virtual Machine. By this
         instruction, high performance native methods of the hosting machine could become available to AVM smart
-        contracts.
+        contracts. This instruction will not modify the AVM state.
+        \item \texttt{spawn} spawns a new method invocation. Spawning a method invocation is the invocation of a method
+        with changing the context of method execution and creating a new call stack. The created call stack is added at
+        the end of the call stack queue and the \texttt{canModify} field of the \texttt{CallInfo} struct will always
+        be set to \texttt{true}. This instruction does not modify the \texttt{pc} register and the execution of the
+        caller will continue after this instruction. As a result, \texttt{spawn} can not return
+        any value to the caller and the return value of the called method (if any) will be added
+        to the transaction output. If the invoked method completes abruptly it will cause all the call stack queue to
+        complete abruptly.
     \end{itemize}
 
-    \note{In the future, we may need to add special instructions for invoking interface and virtual methods...}
+    \note{The AVM does not natively support polymorphism and virtual methods. A compiler could easily
+    generates appropriate code for implementing these features.}
 
-    Each time a method is invoked a new local frame and operand stack is created. The Argennon Virtual Machine uses
-    local frames to pass parameters on method invocation. On method invocation, any parameters are passed in
-    consecutive local variables stored in the method's local frame starting from local variable 0. The invoker of a
-    method writes the parameters in the local frame of the invoked method using \texttt{arg} instructions.
+    Each time a method is invoked the \texttt{Save()} function of the AVM heap is called, then a new local frame
+    and operand stack is created. The Argennon Virtual Machine uses local frames to pass parameters on
+    method invocation. On method invocation, any parameters are passed in consecutive local variables stored in the
+    method's local frame starting from address 0. The invoker of a method writes the parameters in the local frame
+    of the invoked method using \texttt{arg} instructions.
 
-    \subsubsection{Exceptions}
+    \note{For AVM smart contracts \emph{reentrancy} can only happen in read-only mode and is totally safe.
+    At the same time call-back patterns can be easily implemented by using \texttt{spawn} instruction.}
+
+    \subsection{Exceptions}\label{subsec:exceptions}
 
@@
     various Argennon Virtual Machine instructions if they detect an abnormal condition. Some exceptions are not
-    catchable and will always abort the execution of the smart contract.
-
-    \subsubsection{Method Invocation Completion}
-
-    A method invocation completes normally if that invocation does not cause an exception to be thrown, either
-    directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
-    method completes normally, then a value may be returned to the invoking method. This occurs when the invoked
-    method executes one of the return instructions, the choice of which must be appropriate for the type of the value
-    being returned (if any). Execution then continues normally in the invoking method's local frame with the returned
-    value (if any) pushed onto the operand stack.
-
-    A method invocation completes abruptly if an exceptions is thrown and is not caught by the current method. A
-    method invocation that completes abruptly never returns a value to its invoker.
-
-    When a method completes, whether normally or abruptly, the call stack is used to restore the state of the invoker,
-    including its local frame and operand stack, with the \texttt{PC} register appropriately restored and incremented
-    to skip past the method invocation instruction. If the invoker was another smart contract, i.e.\ the invocation
-    was made by an \texttt{invoke\_external} instruction, the current memory segment will be changed to the invoker's
-    segment.
-
-    A thrown exception causes methods in the call stack to complete \textbf{abruptly} one by one, as long as the
-    \texttt{PC} register is not pointing to a \texttt{catch} instruction. The \texttt{catch} instruction acts like a
-    branch instruction that branches only if an exception is caught. \textbf{When an external method invocation
-    completes abruptly, before changing the current segment, all changes made to the heap area by that method
-    call, including changes made to other segments, will be rolled back.} So, when a method call completes abruptly,
-    that method call essentially has no effect on the AVM state.
+    catchable and will always abort the method invocation.
 
@@
 
-    \subsection{Authorizing Operations}\label{subsec:authorizing-operations}
+    \subsection{Method Invocation Completion}\label{subsec:method-invocation-completion}
+
+    A method invocation completes normally if that invocation does not cause an exception to be thrown, either
+    directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
+    method completes normally and the invocation was made by an \texttt{invoke} instruction, then a value may be
+    returned to the invoking method. This occurs when the invoked method executes one of the return instructions, the
+    choice of which must be appropriate for the type of the value being returned (if any). Execution then continues
+    normally in the invoking method's local frame with the returned
+    value (if any) pushed onto the operand stack. If the method was invoked by a \texttt{spawn} instruction,
+    the execution of the next call stack (if any) begins and the returned value
+    is appended at the end of the transaction output. When a method invocation completes normally
+    the \texttt{Heap.Discard()} function is called as a part of the return instruction.
+
+    A method invocation completes abruptly if an exception is thrown and is not caught by the current method. A
+    method invocation that completes abruptly never returns a value.
+
+    When a method completes, whether normally or abruptly, the call stack is used to restore the state of the invoker,
+    including its local frame and operand stack, with the \texttt{pc} register appropriately restored and incremented
+    to skip past the method invocation instruction. If the current call stack is empty the next call stack in the call
+    stack queue will be used for continuing the execution, and when there are no call stacks left in the queue the
+    current execution session ends.
+
+    A thrown exception causes methods in the call stack to complete \emph{abruptly} one by one, as long as the
+    \texttt{pc} register is not pointing to a \texttt{catch} instruction. The \texttt{catch} instruction acts like a
+    branch instruction that branches only if an exception is caught. When an exception is thrown
+    Algorithm~\ref{alg:Abrupt} is used to restore the state of the AVM heap.
+
+    \begin{algorithm}
+        \DontPrintSemicolon
+        \SetKwData{Heap}{Heap}\SetKwData{Stack}{CallStack}
+        \SetKwFunction{Discard}{Discard}\SetKwFunction{Pop}{pop}\SetKwFunction{Restore}{Restore}
+        \SetKwProg{Fn}{Function}{}{}
+
+        \While{\Stack is not empty}
+        {
+            \Stack.\Pop{}\;
+            \eIf{\Stack is not empty \textbf{and} $\Stack[head].pc = catch$}
+            {
+                \Heap.\Restore{}\;
+            }{
+                \Heap.\Discard{}\;
+            }
+        }
+        \Heap.\Restore{}\;
+        \caption{Abrupt Method Completion}\label{alg:Abrupt}
+    \end{algorithm}
+
+    Closer investigation of Algorithm~\ref{alg:Abrupt} shows that when an invocation which was made by
+    a \texttt{spawn} instruction completes abruptly, the heap snapshot that was saved before the start of the
+    execution of the first call stack of the execution session is restored. This means
+    that all changes made to the AVM heap during the execution session will be restored, but the changes made by
+    the root smart contract including the transferring of the transaction fee will not be restored.
+    See Section~\ref{subsec:execution-sessions} for more details.
+
+    \subsection{Heap Allocation and De-Allocation}\label{subsec:heap-allocation-instructions}
+
+    \section{Authorizing Operations}\label{sec:authorizing-operations}
 
@@
     cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
-    passing signatures as parameters to methods, a programmer can implement the required logic for authorizing any
-    operation.
+    passing cryptographic signatures as parameters to methods, a programmer can implement the required logic
+    for authorizing any operation.
 
@@
     instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
-    smart contract in the call stack. In other words, it returns the \texttt{applicationID} of the smart contract that
-    has invoked the current smart contract. (if any)
+    smart contract in the call stack or if there is none, the smart contract that created the current call stack.
+    In other words, it returns the \texttt{applicationID} of the smart contract that has invoked or spawned
+    the current smart contract.
 
-    \subsection{Heap Allocation Instructions}\label{subsec:heap-allocation-instructions}
+    \section{Execution Sessions}\label{subsec:execution-sessions}
+
+    The execution of smart contracts code in the AVM is done in execution sessions. Every execution session in the
+    Argennon Virtual Machine constitutes of a single method invocation from the \emph{root} smart contract and a list
+    of resource allocation requests. These allocation requests always include a request for allocating some amount
+    of execution time and a request for accessing some memory locations for reading or writing. If after the start of
+    execution an execution session tries to violate its allocated resources, an uncatchable exception will
+    be thrown by the AVM\@.
+
+    \note{The AVM does not have a \texttt{calldata} memory area. The arguments of the root smart contract's method will
+    be copied from the transaction to the local frame of the invoked method, exactly like a normal method invocation.}
+
+    The root smart contract with the \texttt{applicationID} of zero, is a special smart contract in the Argennon
+    Virtual Machine. Its code is immutable and a part of the AVM implementation and always is run in privileged mode.
+    Every execution session starts with a call to a method of the root smart contract. This method always transfers
+    some amount of fee from an account ot the \texttt{feeSink} account and then it performs the requested operation.
+
+    If the requested operation is a method invocation the root smart contract will call the \texttt{Heap.Save()}
+    function once and then it will invoke the requested method using \texttt{spawn} instruction. This means that the
+    invoked method will have a separate call stack and if it throws an exception it will not cause a rollback in
+    changes the is made by the root smart contract to the state.
+
+    \note{Argennon transactions do not have a sender, and the payer of a transaction fee could be any
+    account who has provided the required digital signature.}
+
+    \section{Cleanup Sessions}\label{sec:cleanup-sessions}
 
@@
 
-    \chapter{Implementation}\label{ch:implementation}
+    \chapter{The Argennon Blockchain}\label{ch:argennon-blockchain}
 

### 2021-04-25 tex/avm.tex
Introduce the Argon language
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 \documentclass[11pt, a4paper]{report}
+\usepackage[a4paper]{geometry}
 
@@
 %\usepackage{amsmath}
-%\usepackage[a4paper]{geometry}
 \usepackage{graphicx}
@@
 \usepackage{tcolorbox}
+\usepackage{amsmath}
+
 \lstset{
-%  basicstyle=\ttfamily,
+    language=Java,
+    morekeywords={contract, initialize, constructor, external, virtual, map, account, signature, overrides, is, string},
+    basicstyle=\small\sffamily,
+    columns=fullflexible,
+%   keywordstyle=\bfseries,
+    breaklines=true,
+    showstringspaces=false,
     mathescape
@@
     The Argennon Virtual Machine (AVM) is an abstract computing machine for executing Argennon's smart contracts. The
-    Argennon Virtual Machine knows nothing of the Argennon blockchain, only of cleanup sessions and
-    execution sessions.
+    Argennon Virtual Machine knows nothing of the Argennon blockchain, only of Argennon's identifier tries and
+    the concept of execution sessions.
 
     Execution sessions are separate sessions of executing smart contact's code by the Argennon Virtual Machine.
-    These sessions usually are correspondent to the concept of a transaction in a blockchain. (See
-    Section~\ref{subsec:execution-sessions})
-
-    A cleanup session is the time slot in which the Argennon Virtual Machine performs cleanup and re-optimization of
-    its internal data structures in a "Stop the World" manner. Cleanup sessions usually are done as the final phase of
-    the block confirmation process in a blockchain. (See Section ...)
+    These sessions are usually correspondent to the concept of a transaction in a blockchain. (See
+    Section~\ref{sec:execution-sessions})
 
@@
 
-    The Argennon Virtual Machine operates on two kinds of types: primitive types and reference types. There are,
+    The Argennon Virtual Machine operates on two kinds of types: primitive types and identifier types. There are,
     correspondingly, two kinds of values that can be stored in memory locations, passed as arguments, returned by
-    methods, and operated upon: primitive values and reference values.
-
-    Values of primitive or reference types need not be tagged or otherwise be inspectable to determine their types
+    methods, and operated upon: primitive values and identifier values.
+
+    Values of primitive or identifier types need not be tagged or otherwise be inspectable to determine their types
     at run time, or to be distinguished from values of other types. Instead, the instruction set of the Argennon
     Virtual Machine distinguishes its operand types using instructions intended to operate on values of specific
-    types. For instance, \texttt{iadd64} assumes that its operands are two 64-bit signed integers or \texttt{rstore}
-    assumes that its operand is a reference type.
-
-    A reference value may be the special \texttt{null} reference, a reference to no location, which will be denoted here
-    by \texttt{null}. The default value of a reference type is \texttt{null}.
+    types. For instance, \texttt{iadd64} assumes that its operands are two 64-bit signed integers or \texttt{delete}
+    assumes that its operand is an identifier type.
+
+    An identifier value in the Argennon Virtual Machine is a variable length array of bytes. Some instructions
+    that work with identifiers are able to determine the length of their identifier operands, while some
+    other instructions, for performance reasons, require the length to be specified.
 
@@
 
+    \section{Identifiers}\label{sec:identifiers}
+
+    In the Argennon Virtual Machine four distinct identifier values exist: \texttt{applicationID}, \texttt{accountID},
+    \texttt{methodID} and \texttt{chunkID}.
+
+    All these identifiers are \emph{prefix codes}, and hence can be represented by
+    \emph{prefix trees} \footnote{Also knows as tries.}. The AVM has three primitive prefix trees:
+    \emph{applications, accounts} and \emph{local}. Any identifier in Argennon is a
+    prefix code built by using one or more of these prefix trees:
+    \begin{itemize}
+        \item \texttt{applicationID} is a prefix code built by \emph{applications} prefix tree.
+        \item \texttt{accountID} is a prefix code built by \emph{accounts} prefix tree.
+        \item \texttt{methodID} is a composite prefix code built by concatenating an \texttt{applicationID} to
+        a prefix code made by \emph{local} prefix tree.(\texttt{|} is the concatenating operator.)
+
+        \texttt{methodID = (applicationID|<local prefix code>)}
+        \item \texttt{chunkID} is a composite prefix code built by concatenating an \texttt{applicationID} to
+        an \texttt{accountID} to a prefix code made by \emph{local} prefix tree
+
+        \texttt{chunkID = (applicationID|accountID|<local prefix code>)}
+    \end{itemize}
+
+    All Argennon prefix trees have an equal branching factor \(\beta\). Therefore, we can represent an Argennon
+    prefix tree as a sequence of fractional numbers in base \(\beta\):
+    \[
+        <A^{(1)},A^{(2)},A^{(3)},\dots>
+    \]
+    Where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)}<A^{(i+1)}\). A typical choice for \(\beta\)
+    could be \(2^8\).
+
+    One important property of prefix identifiers is that while they have variable and unlimited length, they are
+    uniquely extractable from any data. Assume that we have a string of digits which are in base $\beta$, we
+    know that $k$ first digits belong to an Argennon's identifier, but we don't know the value of $k$.
+    Algorithm~\ref{alg:prefix_id} can extract the prefixed identifier uniquely. We can apply this algorithm
+    multiple times to a sequence to extract a composite identifier, for example \texttt{chunkID}, from a sequence.
+
+    %##\includegraphics[width=17cm]{../img/Alg1s.png}
+    \begin{algorithm}
+        \DontPrintSemicolon
+        \SetKwData{Ready}{$R_e$}\SetKwData{V}{$v_f$}\SetKwData{Graph}{$G_e$}\SetKwData{Vertices}{$V$}\SetKwData
+        {Txns}{$T$}
+        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
+        \Input{A sequence of $n$ digits in base $\beta$: $d_{1}d_{2}\dots d_{n}$ \newline
+            A prefix tree: $<A^{(1)},A^{(2)},A^{(3)},\dots>$}
+        \Output{Valid identifier prefix of the sequence.}
+        \BlankLine
+        \For{$i = 1$ \KwTo $n$}
+        {
+            \If{$(0.d_{1}d_{2}\dots d_{i})_\beta < A^{(i)}$}
+            {
+                \KwRet{$d_{1}d_{2}\dots d_{i}$}\;
+            }
+        }
+        \KwRet{NIL}\;
+        \caption{Finding a prefixed identifier}\label{alg:prefix_id}
+    \end{algorithm}
+
+    In Argennon the shorter prefix codes are assigned to more active accounts and smart contracts which tend to own more
+    data objects in the system. The prefix trees are designed by analyzing empirical data to make sure the number
+    of leaves in each level is chosen appropriately.
+
     \section{Arithmetics}\label{sec:arithmetics}
 
-    All arithmetic operations in the Argennon Virtual machine are safe and there are strict overflow checks and
-    blah blah blah
+    The Argennon Virtual Machine supports signed integer and signed floating point operations. The Argennon Virtual
+    Machine does not support any type of unsigned arithmetics. All arithmetic operations in the Argennon Virtual Machine
+    are checked and any type of overflow or underflow will cause a catchable exception to be thrown.
 
@@
     All data areas except operand stack, have their own address space. Operand stack is a
-    last-in-first-out (LIFO) stack and is not addressable. Every AVM memory access instruction operates on its specific
+    last-in-first-out (LIFO) stack and is not addressable. Every AVM instruction operates on its specific
     data areas.
@@
     non-executable method. The constant area of a smart contract contains several kinds of constants, ranging from user
-    defined constants to method address tables. A method address table stores the list of method identifiers of a smart
-    contract and their access type. The access type of a method can be either \texttt{external} or \texttt{internal}.
+    defined constants to method address tables. A method address table stores the list of methods of a smart
+    contract and their access type. The access type of method can be either \texttt{external} or \texttt{internal}.
     Only external methods can be invoked by \texttt{invoke\_external} instruction.
@@
     those chunks that was created by itself. In other words, every memory chunk in the AVM heap has an owner. Only
-    the owner can modify a chunk while any one can read that chunk. In addition, for modifying a chunk the \texttt{canModify}
-    flag of the current method's \texttt{callInfo} struct must be true.
+    the owner can modify a chunk while anyone can read that chunk. In addition, for modifying a chunk the \texttt{canModify}
+    flag of the current method's \texttt{callInfo} struct must be true. The owner of a chunk can be easily determined
+    by the \texttt{applicationID} part of the chunk identifier.
 
@@
         \item \texttt{Discard()} pops the snapshot stored on the top of the snapshot stack and discards it. The current
-        state of the heap will be preserved.
+        state of the heap will not change.
     \end{itemize}
@@
     \begin{itemize}
-        \item \texttt{invoke\_internal} invokes a method without changing the context of method execution. In other
+        \item \texttt{invoke\_internal}: invokes a method without changing the context of method execution. In other
         words, \texttt{applicationID} and \texttt{canModify} fields of the invoked method in the
@@
         the invoking smart contract, a smart contract will not be able to modify another smart contract state by
-        this instruction. This instruction facilitates the usage of libraries and code reuse.
-        \item \texttt{invoke\_external} invokes a method and changes the context of method execution to another
+        this instruction. This instruction facilitates code reuse and the usage of libraries.
+        \item \texttt{invoke\_external}: invokes a method and changes the context of method execution to another
         smart contract. Only external methods of another smart contract can be called by this instruction. For changing
@@
         to the \texttt{applicationID} of the called smart contract. Then it searches the call stack and if
-        the \texttt{applicationID} of the called smart contract is found in the call stack it sets \texttt{canModify}
-        to \texttt{false}, otherwise \texttt{canModify} is set to \texttt{true}. As a result, if the
-        called smart contract is already called, the invoked method will not be able to modify the heap area.
-        \item \texttt{invoke\_native} invokes a method that is not hosted by the Argennon Virtual Machine. By this
+        the \texttt{applicationID} of the called smart contract is found in the current call stack it sets
+        \texttt{canModify} to \texttt{false}, otherwise \texttt{canModify} is set to \texttt{true}.
+        As a result, if the called smart contract is already called, the invoked method will not be
+        able to modify the heap area.
+        \item \texttt{invoke\_native}: invokes a method that is not hosted by the Argennon Virtual Machine. By this
         instruction, high performance native methods of the hosting machine could become available to AVM smart
         contracts. This instruction will not modify the AVM state.
-        \item \texttt{spawn} spawns a new method invocation. Spawning a method invocation is the invocation of a method
-        with changing the context of method execution and creating a new call stack. The created call stack is added at
-        the end of the call stack queue and the \texttt{canModify} field of the \texttt{CallInfo} struct will always
+        \item \texttt{spawn}: spawns a new method invocation. Spawning a method invocation is the invocation of a method
+        with changing the context of method execution and creating a new call stack. Only external methods can
+        be spawned. The created call stack is added at the end of the call stack queue and
+        the \texttt{canModify} field of the pushed \texttt{CallInfo} struct will always
         be set to \texttt{true}. This instruction does not modify the \texttt{pc} register and the execution of the
@@
         any value to the caller and the return value of the called method (if any) will be added
-        to the transaction output. If the invoked method completes abruptly it will cause all the call stack queue to
+        to the transaction output. If the invoked method completes abruptly, it will cause all the call stack queue to
         complete abruptly.
@@
     the execution of the next call stack (if any) begins and the returned value
-    is appended at the end of the transaction output. When a method invocation completes normally
+    is appended at the end of the transaction output. When a method invocation completes normally,
     the \texttt{Heap.Discard()} function is called as a part of the return instruction.
@@
             \Stack.\Pop{}\;
-            \eIf{\Stack is not empty \textbf{and} $\Stack[head].pc = catch$}
+            \eIf{\Stack is not empty \textbf{and} $\Stack[head].pc \rightarrow catch$}
             {
@@
         \Heap.\Restore{}\;
-        \caption{Abrupt Method Completion}\label{alg:Abrupt}
+        \caption{Abrupt method completion}\label{alg:Abrupt}
     \end{algorithm}
 
-    Closer investigation of Algorithm~\ref{alg:Abrupt} shows that when an invocation which was made by
+    A closer investigation of Algorithm~\ref{alg:Abrupt} shows that when an invocation which was made by
     a \texttt{spawn} instruction completes abruptly, the heap snapshot that was saved before the start of the
-    execution of the first call stack of the execution session is restored. This means
+    execution of the first call stack of the execution session is restored. Since the root
+    smart contract calls \texttt{Heap.Save()} function before spawning the requested method, this means
     that all changes made to the AVM heap during the execution session will be restored, but the changes made by
-    the root smart contract including the transferring of the transaction fee will not be restored.
-    See Section~\ref{subsec:execution-sessions} for more details.
-
-    \subsection{Heap Allocation and De-Allocation}\label{subsec:heap-allocation-instructions}
+    the root smart contract including the transferring of the transaction fee will not be restored. See
+    Section~\ref{sec:execution-sessions} for more details.
+
+    \subsection{Heap Allocation and Deallocation}\label{subsec:heap-allocation-instructions}
+
+    Heap allocation and deallocation in the Argennon Virtual Machine is done by the following instructions:
+
+    \begin{itemize}
+        \item \texttt{alloc}: creates a new memory chunk of the specifies size in the heap. the identifier
+        of the new chunk will be the concatenation of the \texttt{applicationID} of the creator of the chunk and
+        the \texttt{id} operand. The specified identifier must be concatenation of
+        an \texttt{accountID} and a prefix code
+        generated by the \emph{local} prefix tree: \texttt{id = accountID|localID}. If the identifier is not valid or
+        if it already exists in the heap a catchable exception will be thrown.
+
+        \item \texttt{delete}: deletes the chunk that its identifier is the concatenation of the current smart
+        contract \texttt{applicationID} to the specified \texttt{id}. If the identifier is not valid or
+        if the chunk was not found a catchable exception will be thrown.
+    \end{itemize}
+
+    When a smart contract allocates a new memory chunk, the identifier of the new chunk is not generated by
+    the AVM, instead the smart contract can choose an identifier itself. This is a very important feature of
+    the AVM heap which allows smart contracts to use the AVM heap as a dictionary (map) data structure.
+    Because the \texttt{chunkID} is a prefix code, any smart contract has its own identifier space and a smart contract
+    can easily generate unique identifiers for its chunks.
+
+    The AVM requires the chosen identifier to be a prefix code made by concatenating a code compatible with
+    the \emph{accounts} prefix tree to a code compatible with the \emph{local} prefix tree. This requirement will
+    enable AVM to detect invalid identifiers while smart contracts can easily associate data with account identifiers.
+    At the same time, A smart contract is able to create maps with costume keys by appending local identifiers
+    to the \texttt{accountID} zero, which does not correspond to a valid account in the Argennon.
+
+    \section{Execution Sessions}\label{sec:execution-sessions}
+
+    The execution of smart contracts code in the AVM is done in execution sessions. Every execution session in the
+    Argennon Virtual Machine consists of a single method invocation from the \emph{root} smart contract and a list
+    of resource allocation requests. These allocation requests always include a request for allocating some amount
+    of execution time and a request for accessing some memory locations for reading or writing. If after the start
+    an execution session tries to violate its allocated resources, an \textbf{uncatchable} exception will
+    be thrown by the AVM\@.
+
+    \note{The AVM does not have a \texttt{calldata} memory area. The arguments of the root smart contract method will
+    be copied from the transaction to the local frame of the invoked method, exactly like a normal method invocation.}
+
+    The root smart contract with the \texttt{applicationID} of zero, is a special smart contract in the Argennon
+    Virtual Machine. Its code is immutable and is a part of the AVM implementation. The root smart contract is always
+    run in the privileged mode, and every execution session starts with a call to a method of the root smart contract.
+    This method always transfers some amount of fee from an account ot the \texttt{feeSink} account and then it
+    performs the requested operation.
+
+    \note{Argennon transactions do not have a sender, and the payer of the transaction fee could be any
+    account who has provided the required digital signature.}
+
+    If the requested operation is a method invocation, the root smart contract will call the \texttt{Heap.Save()}
+    function once, and then it will invoke the requested method using \texttt{spawn} instruction. This means that the
+    invoked method will have a separate call stack and if it completes abruptly the changes made by the root
+    smart contract to the state will not be reverted.
 
@@
     from a user to another user, first we need to make sure that the sender has authorized this operation. The
-    Argennon virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
+    Argennon virtual machine has no built-in mechanism for authorizing operations, but it provides a rich set of
     cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
@@
 
-    \section{Execution Sessions}\label{subsec:execution-sessions}
-
-    The execution of smart contracts code in the AVM is done in execution sessions. Every execution session in the
-    Argennon Virtual Machine constitutes of a single method invocation from the \emph{root} smart contract and a list
-    of resource allocation requests. These allocation requests always include a request for allocating some amount
-    of execution time and a request for accessing some memory locations for reading or writing. If after the start of
-    execution an execution session tries to violate its allocated resources, an uncatchable exception will
-    be thrown by the AVM\@.
-
-    \note{The AVM does not have a \texttt{calldata} memory area. The arguments of the root smart contract's method will
-    be copied from the transaction to the local frame of the invoked method, exactly like a normal method invocation.}
-
-    The root smart contract with the \texttt{applicationID} of zero, is a special smart contract in the Argennon
-    Virtual Machine. Its code is immutable and a part of the AVM implementation and always is run in privileged mode.
-    Every execution session starts with a call to a method of the root smart contract. This method always transfers
-    some amount of fee from an account ot the \texttt{feeSink} account and then it performs the requested operation.
-
-    If the requested operation is a method invocation the root smart contract will call the \texttt{Heap.Save()}
-    function once and then it will invoke the requested method using \texttt{spawn} instruction. This means that the
-    invoked method will have a separate call stack and if it throws an exception it will not cause a rollback in
-    changes the is made by the root smart contract to the state.
-
-    \note{Argennon transactions do not have a sender, and the payer of a transaction fee could be any
-    account who has provided the required digital signature.}
-
-    \section{Cleanup Sessions}\label{sec:cleanup-sessions}
+    \section{The AVM Standard Library}\label{sec:the-avm-standard-library}
 
@@
 
-    \section{The AVM Standard Library}\label{sec:the-avm-standard-library}
-
-    ...
+    \chapter{The Argon Language}\label{ch:argon-lang}
+
+    \section{Introduction}\label{sec:introduction2}
+
+    The Argon programming language is a class-based, object-oriented language designed for writing Argennon's smart
+    contracts. The Argon programming language is inspired by Solidity and is similar to Java, with a number of aspects
+    of them omitted and a few ideas from other languages included. Argon is designed to be fully compatible with
+    the Argennon Virtual Machine and be able to use all advanced features of the Argennon blockchain.
+
+    Argon Programs are organized as sets of packages. Each package has its own set of names for types, which helps
+    to prevent name conflicts. Every package can contain an arbitrary number of classes and a single contract. Every
+    Argon package corresponds to at max one AVM smart contract.
+
+    \section{Features Overview}\label{sec:features-overview}
+
+    \subsection{Static Classes}\label{subsec:static-classes}
+
+    In Argon a static class is an uninstantiable class which all its members and methods are static. Every static
+    class can define a special method \texttt{initialize} which can only be called in the \texttt{initialize}
+    method of the single contract of the package the static class is defined in.
+
+    A Static class is like a normal class which can only have one instance. So like normal classes, static classes can
+    be a super class or a subclass of other static classes and can implement interfaces.
+
+    \subsection{Contracts}\label{subsec:contract}
+
+    Every package of an Argon program can have one or zero contracts. A contact is a special static class which is
+    allowed to define methods with \texttt{external} visibility. When an Argon package is deployed as an AVM smart
+    contract, these external methods will define the contract's public interface.
+
+    \subsection{Access Level Modifiers}\label{subsec:access-level-modifiers}
+
+    Access level modifiers determine whether other classes can use a particular field or invoke a particular method or
+    if a method can be invoked externally by other smart contracts.
+
+    \begin{center}
+        \begin{tabular}{lllllr}
+            \hline
+                    & Class & Package & Subclass & Program & World \\
+            \hline
+            private   & yes   & no      & no       & no      & no    \\
+            protected & yes   & no      & yes      & no      & no    \\
+            package   & yes   & yes     & yes      & no      & no    \\
+            public    & yes   & yes     & yes      & yes     & no    \\
+            external  & no    & no      & no       & no      & yes   \\
+            \hline
+        \end{tabular}\label{tab:table}
+    \end{center}
+
+    \begin{lstlisting}[frame=TB, float, title=Argon contracts are static classes,label={lst:code1}]
+// A contract is a static class and all its methods are automatically static.
+contract MirrorToken {
+    private SimpleToken token;
+    private SimpleToken reflection;
+
+    // `initialize` is a special static method that is called by the AVM after the code of a contract
+    // is stored in the AVM code area. It can not be called after that.
+    initialize(double supply1, double supply2) {
+    // `new` does not create a new smart contract. It just makes an ordinary object.
+        token = new SimpleToken(supply1);
+        reflection = new SimpleToken(supply2);
+    }
+
+    external void transfer(account sender, account recipient,
+        double amount, signature sig) {
+        // Checks if the signature is issued by the sender and verifies the calling of this function
+        // with the current parameters.
+        sig.verifyCallBy(sender);
+        token.transfer(sender, recipient, amount);
+        reflection.transfer(recipient, sender, Math.sqrt(amount));
+    }
+
+    external double balanceOf(account user) { return token.balanceOf(user); }
+
+    external double balanceOfReflection(account user) { return reflection.balanceOf(user); }
+}
+
+public static class Math {
+    public double sqrt(double x) {
+        return // calculate square root of x
+    }
+}
+
+package class SimpleToken {
+    private map(account -> double) balances;
+
+    // The visibility of a member without an access modifier will be the same as its defining
+    // class so `SimpleToken` constructor has package visibility.
+    constructor(double initialSupply) {
+        // initializes the object
+    }
+
+    void transfer(account sender, account recipient, double amount) {
+        if (balances[sender] < amount) throw("Not enough balance.");
+        // implements the required logic...
+    }
+    // implements other methods...
+}
+    \end{lstlisting}
+
+    \subsection{Shadowing}\label{subsec:shadowing}
+
+    If a declaration of a type (such as a member variable or a parameter name) in a particular scope (such as an
+    inner block or a method definition) has the same name as another declaration in the enclosing scope, it will
+    result in a compiler error. In other words, the Argon programming language does not allow shadowing.
 

### 2021-04-26 tex/A.tex
Change project structure
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-03-29 tex/avm.tex
@@
 % Preamble
-\documentclass[a4paper]{report}
+\documentclass[11pt, a4paper]{report}
+\usepackage[a4paper]{geometry}
 
@@
 
-\title{The Argennon Virtual Machine}
+\title{Argennon: A Scalable Smart Contract Platform}
 \author{aybehrouz}
@@
 
-% Document
 \begin{document}
-    \maketitle
+    %\maketitle
     \tableofcontents
 
-    \chapter{Specification}\label{ch:specification}
+    \chapter{The Argennon Virtual Machine}\label{ch:AVM}
+    \input{sections/avm}
 
-    \section{Introduction}\label{sec:introduction}
+    \chapter{The Argon Language}\label{ch:argon-lang}
+    \input{sections/argon}
 
-    The Argennon Virtual Machine (AVM) is an abstract computing machine for executing Argennon's smart contracts. It
-    is designed in a way that it could be efficiently implemented in either hardware or software.
-
-    \section{The Structure of the Argennon Virtual Machine}\label{sec:the-structure-of-the-algorand-virtual-machine}
-
-    ...
-
-    \subsection{Data Types}\label{subsec:data-types}
-
-    The Argennon Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
-    and does not have to be done by the Argennon Virtual Machine itself.
-
-    The instruction set of the Argennon Virtual Machine distinguishes its operand types using instructions intended to
-    operate on values of specific types. For instance, \texttt{iadd} assumes that its operands are two 64-bit integers.
-
-    \subsection{The \texttt{PC} Register}\label{subsec:the-pc-register}
-
-    ...
-
-    \subsection{Call Stack}\label{subsec:call-stack}
-
-    A call stack contains the information that is needed for restoring the state of the invoker of a method.
-
-    \subsection{Memory Unit}\label{subsec:memory-unit}
-
-    The Argennon Virtual Machine has a byte addressable memory which is divided into separate segments. Every segment
-    belongs to a smart contract that has a unique \texttt{applicationID}. The AVM always has a single working memory
-    segment and memory locations outside its current working segment can not be accessed. The only instructions which
-    can change the current working segment are \texttt{invoke\_external}, \texttt{athrow} and return instructions.
-
-    Every smart contract has its own memory segment. Hence, there is no way for a smart contract to access another
-    smart contract's memory. Interaction between smart contracts is done using \texttt{invoke\_external} instruction,
-    and a smart contract can invoke methods of another smart contract by this instruction.
-
-    A memory segment consists of five data areas:
-    \begin{itemize}
-        \item Code Area
-        \item Constant Area
-        \item Local Frame
-        \item Operand Stack
-        \item Heap
-    \end{itemize}
-    All data areas except operand stack, have their own address space which starts from 0. Operand stack is a
-    last-in-first-out (LIFO) stack and is not addressable. Every memory access instruction operates on its specific
-    data areas.
-
-    \subsubsection{Code Area}
-
-    The code area of a segment contains the byte-code of the smart contract which owns that segment. The AVM has no
-    instructions for manipulating the code area. \textbf{Installing, removing and updating smart contracts need to
-    be done externally.}
-
-    \subsubsection{Constant Area}
-
-    The constant area of a segment contains several kinds of constants, ranging from user defined constants to method
-    address tables. A method address table stores method locations in the code area and their access type. The access
-    type of a method can be either \texttt{public} or \texttt{private}. The AVM has no instructions for modifying the
-    constant area.
-
-    \note{Only public methods can be invoked by \texttt{invoke\_external} instruction.}
-
-    \subsubsection{Local Frame}
-
-    A local frame is used to store methods parameters and local variables. A new frame is created each time a method
-    is invoked, and it is destroyed when its method invocation completes, whether the completion is normal or abrupt.
-
-    \subsubsection{Operand Stack}
-
-    Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too. AVM
-    instructions take operands from the operand stack, operate on them, and push the result back onto the operand
-    stack. An operand stack is destroyed when its owner method completes, whether that completion is normal or abrupt.
-
-    \subsubsection{Heap}
-
-    The heap of a segment is a persistent memory area which is divided into pages. Memory locations inside every page
-    have a separate address space that starts from 0, and each page can be referenced by an index that starts from 0. In
-    other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex, offset)}.
-    Different pages of a heap do not need to be equally sized.
-
-    \note{The reason behind this paged design is that the heap is usually persisted using a block device. A heap with
-    a paged structure could expose the underlying block based nature of the persistence layer to the application
-    layer. In this way, the compiler or the programmer could better optimize the code for the persistence layer.}
-
-    \section{Instruction Set Summary}\label{sec:instruction-set-summary}
-
-    An Argennon Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
-    performed, followed by zero or more operands supplying arguments or data that are used by the operation.
-    \textbf{The number and size of the operands are determined solely by the opcode.}
-
-    \subsection{Method Invocation}\label{subsec:method-invocation}
-
-    The Argennon Virtual Machine has three types of method invocation:
-    \begin{itemize}
-        \item \texttt{invoke\_internal} invokes a method from the current running smart contract.
-        \item \texttt{invoke\_external} invokes a \texttt{public} method from another smart contract. It will change
-        the current memory segment to the segment of the invoked smart contract.
-        \item \texttt{invoke\_native} invokes a method that is not hosted by the Argennon virtual machine. By this
-        instruction, high performance native methods of the hosting machine could become available to AVM smart
-        contracts.
-    \end{itemize}
-
-    \note{In the future, we may need to add special instructions for invoking interface and virtual methods...}
-
-    Each time a method is invoked a new local frame and operand stack is created. The Argennon Virtual Machine uses
-    local frames to pass parameters on method invocation. On method invocation, any parameters are passed in
-    consecutive local variables stored in the method's local frame starting from local variable 0. The invoker of a
-    method writes the parameters in the local frame of the invoked method using \texttt{arg} instructions.
-
-    \subsubsection{Exceptions}
-
-    An exception is thrown programmatically using the \texttt{athrow} instruction. Exceptions can also be thrown by
-    various Argennon Virtual Machine instructions if they detect an abnormal condition. Some exceptions are not
-    catchable and will always abort the execution of the smart contract.
-
-    \subsubsection{Method Invocation Completion}
-
-    A method invocation completes normally if that invocation does not cause an exception to be thrown, either
-    directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
-    method completes normally, then a value may be returned to the invoking method. This occurs when the invoked
-    method executes one of the return instructions, the choice of which must be appropriate for the type of the value
-    being returned (if any). Execution then continues normally in the invoking method's local frame with the returned
-    value (if any) pushed onto the operand stack.
-
-    A method invocation completes abruptly if an exceptions is thrown and is not caught by the current method. A
-    method invocation that completes abruptly never returns a value to its invoker.
-
-    When a method completes, whether normally or abruptly, the call stack is used to restore the state of the invoker,
-    including its local frame and operand stack, with the \texttt{PC} register appropriately restored and incremented
-    to skip past the method invocation instruction. If the invoker was another smart contract, i.e.\ the invocation
-    was made by an \texttt{invoke\_external} instruction, the current memory segment will be changed to the invoker's
-    segment.
-
-    A thrown exception causes methods in the call stack to complete \textbf{abruptly} one by one, as long as the
-    \texttt{PC} register is not pointing to a \texttt{catch} instruction. The \texttt{catch} instruction acts like a
-    branch instruction that branches only if an exception is caught. \textbf{When an external method invocation
-    completes abruptly, before changing the current segment, all changes made to the heap area by that method
-    call, including changes made to other segments, will be rolled back.} So, when a method call completes abruptly,
-    that method call essentially has no effect on the AVM state.
-
-    \note{By using the \texttt{athrow} instruction properly, a programmer can make any method act like an atomic
-    operation.}
-
-    \subsection{Authorizing Operations}\label{subsec:authorizing-operations}
-
-    In blockchain applications, we usually need to authorize certain operations. For example, for sending an asset
-    from a user to another user, first we need to make sure that the sender has authorized this operation. The
-    Argennon virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
-    cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
-    passing signatures as parameters to methods, a programmer can implement the required logic for authorizing any
-    operation.
-
-    \note{The Argennon virtual machine has no instructions for issuing cryptographic signatures.}
-
-    In addition to signatures, a method can verify its invoker by using \texttt{get\_parent} instruction. This
-    instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
-    smart contract in the call stack. In other words, it returns the \texttt{applicationID} of the smart contract that
-    has invoked the current smart contract. (if any)
-
-    \subsection{Heap Allocation Instructions}\label{subsec:heap-allocation-instructions}
-
-    ...
-
-    \section{The AVM Standard Library}\label{sec:the-avm-standard-library}
-
-    ...
-
-    \chapter{Implementation}\label{ch:implementation}
-
-    \section{Persistence}\label{sec:persistence}
-
-    For implementing the persistence layer of the AVM, we assume that we have access to an updatable zero-knowledge
-    elementary database (ZK-EDB) with the following properties:
-
-    \begin{itemize}
-        \item The ZK-EDB contains a mapping from a set of keys to a set of values.
-        \item Every state of the database has a commitment \(C\).
-        \item The ZK-EDB has a method \((D, p) = get(x)\), where \(x\) is a key and \(D\) is the associated data
-        with \(x\), and \(p\) is a proof.
-        \item A user can use \(C\) and \(p\) to verify that \(D\) is really associated with \(x\), and \(D\) is not
-        altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB\@.
-        \item Having \(p\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
-        associated with \(x\) instead of \(D\).
-    \end{itemize}
-
-    We use a ZK-EDB for storing the AVM heap. We include the commitment of the current state of this DB in every
-    block of the Argennon blockchain, so ZK-EDB servers need not be trusted servers.
-
-    Every page of AVM heap will be stored with a key of the form: \texttt{applicationID|pageIndex} (the \texttt{|}
-    operator concatenates two numbers). Nodes do not keep a full copy of the AVM heap and for validating block
-    certificates or emulating the AVM ( i.e.~validating transactions) they need to connect to a ZK-EDB and retrieve
-    the required pages of AVM heap. For better performance, nodes keep a cache of heap pages to
-    reduce the amount of ZK-EDB access.
-
-    We also use a ZK-EDB for storing the code area of each segment, and we include the commitment of this DB in every
-    block. Every code area will be divided into blocks and every block will be stored in the DB with
-    \texttt{applicationID|blockID} as its key. Like heap pages, nodes keep a cache of code area blocks.
-
-    \note{Unlike heap pages, the AVM is not aware of different blocks of code area.}
-
-    \section{Transactions}\label{sec:transactions}
-
-    Argennon has four types of transaction:
-
-    \begin{itemize}
-        \item \texttt{avmCall} essentially is an \texttt{invoke\_external} instruction that invokes a method from an
-        AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
-        assets, including ARGs, is done by these transactions.
-        \item \texttt{installApp} installs an AVM smart contract and determines the update policy of the smart
-        contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
-        on.
-        \item \texttt{unInstallApp} removes an AVM smart contract.
-        \item \texttt{updateApp} updates an AVM smart contract.
-    \end{itemize}
-
-    All types of Argennon transactions contain an \texttt{invoke\_external} instruction which calls a special method
-    from ARG smart contract that transfers the proposed fee of the transaction in ARGs from a sender account to the
-    fee sink accounts.
-
-    Every transaction is required to exactly specify what heap locations or code area addresses it will access. This
-    enables validators to start retrieving the required memory blocks from available ZK-EDB servers as soon as they
-    see a transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to
-    access a memory location that is not included in its access lists, will be rejected. Users could use smart contract
-    oracles to predict the list of memory blocks their transactions need. See Section~\ref{sec:smart-contract-oracle}
-    for more details.
-
-    \section{Blockchain}\label{sec:blockchain}
-
-    Every block of the Argennon blockchain corresponds to a set of transactions. We store the commitment of this
-    transaction set in every block, but we don't keep the set itself. To be able to detect replay attacks, we require
-    every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
-    and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
-    must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
-    a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
-    if a signature of the same user with a bigger or equal nonce is already used
-    (i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
-    nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
-
-    To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. When the
-    difference between \texttt{issuance} component of this nonce and the current round becomes bigger than the
-    maximum allowed lifetime of a signature, this information can be safely deleted. \textbf{As a result, we will not
-    have the problem of "un-removable empty accounts" like Ethereum.}
-
-    The only information that Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
-    blockchain. Every block of the Argennon blockchain contains the following information:
-
-    \begin{center}
-        \begin{tabular}{||c||}
-            \hline
-            Block \\ [0.5ex]
-            \hline\hline
-            commitment to the ZK-EDB storing heap pages \\ [0.7ex]
-            commitment to the ZK-EDB storing code areas \\ [0.7ex]
-            commitment to the set of transactions       \\ [0.7ex]
-            previous block hash                         \\ [0.7ex]
-            random seed                                 \\ [0.7ex]
-            \hline
-        \end{tabular}
-    \end{center}
-
-    For confirming a new block, nodes that are not validators only need to verify the block certificate. For
-    verifying a block certificate, a node needs to know the ARG balances of validators, but it doesn't need to
-    emulate the AVM execution.
-
-    On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
-    execution of the Argennon virtual machine. To do so, first they retrieve all heap pages and code area blocks they
-    need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
-    transactions included in the new block. This will modify some pages of the AVM memory, so they update the ZK-EDB
-    commitments based on the modified pages and verify the commitments included in the new block. Validators also
-    calculate and verify the commitment to the new block's transaction set.
-
-    \note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDB servers will receive the new
-    block, and they will update their database by emulating the AVM execution.}
-
-    \section{Consensus}\label{sec:consensus}
-
-    \note{It will be a randomized proof of stake which is under development....}
-
-    \subsection{Estimating A User's Stake}\label{subsec:estimating-a-user's-stake}
-
-    In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
-    of stake the user has in the system. Conventionally in these systems, for estimating a user's stake, we use the
-    amount of native system tokens the user is holding. Unfortunately, one problem with this approach is that a
-    strong attacker may be able to obtain a considerable amount of system tokens, for example by borrowing from a
-    DEFI application, and use this stake to attack the system.
-
-    To mitigate this problem, for calculating a user's stake at time step \(t\), instead of using the raw ARG
-    balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
-    ARG balance:
-    \[
-        S_{u,t} = \min (B_{u,t}, Trust_{u,t})
-    \]
-    Where:
-    \begin{itemize}
-        \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
-        \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
-        \item \(Trust_{u,t}\) is an estimated trust value for user \(u\) at time step \(t\).
-    \end{itemize}
-
-    The agreement protocol, at time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
-    number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
-    exponential moving average of the ARG balance of user \(u\) at time step \(t\).
-
-    In our system a user who held ARGs and participated in the consensus for a long time is more trusted
-    than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
-    amount of ARGs, also needs to hold them for a long period of time before being able to attack the system.
-
-    For calculating the exponential moving average of a user's balance at time step \(t\), we can use the following
-    recursive formula:
-    \[
-        M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
-    \]
-    Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
-    degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
-
-    Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
-    \(M_{u,t}\): (In the following equations the \(u\) subscript is dropped for simplicity)
-    \[
-        M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B
-    \]
-    Where:
-    \[
-        B = B_{k+1} = B_{k+2} = \dots = B_{t}
-    \]
-    We know that when \(|nx| \ll 1\) we can use the binomial approximation \({(1 + x)^n \approx 1 + nx}\). So, we can
-    further simplify this formula:
-    \[
-        M_{t} = M_{k} + (t - k) \alpha (B - M_{k})
-    \]
-
-    For choosing the value of \(\alpha\) we can consider the number of time steps that the trust value of a user needs
-    for reaching a specified fraction of his account balance. We know that for large \(n\) and \(|x| < 1\) we have
-    \((1 + x)^n \approx e^{nx}\), so by letting \(M_{u,k} = 0\) and \(n = t - k\) we can write:
-    \[
-        \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{B}\right)}{n}
-    \]
-    The value of \(\alpha\) for a desired configuration can be calculated by this equation. For instance, we could
-    calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8B\) and \(n\) equals to the
-    number of time steps of 10 years.
-
-    In our system a newly created account will not have voting power for some time, no matter how high its
-    balance is. While this is a desirable property, in case a large proportion of total system tokens are
-    transferred to newly created accounts, it can result in too much voting power for older accounts. This may decrease
-    the degree of decentralization in our system.
-
-    However, this situation is easily detectable by comparing the total stake of the system with the total balance of
-    users. If after confirming a block the total stake of the system goes too low and we have:
-    \[
-        \sum_{u}S_{u,t} < \gamma \sum_{u}B_{u,t}
-    \]
-    The protocol will perform a \emph{time shift} in the system: the time step of the system
-    will be incremented for \(m\) steps while no blocks will be confirmed. This will increase the value of \(M_{u,t}\)
-    for new accounts with a non-zero balance, giving them more influence in the agreement protocol.
-
-    For calculating the value of \(m\) which determines the amount of time shift in the system, we should note that when
-    \(B_{u,t} = B_{u, t-1} = B_u\), we can derive a simple recursive rule for the stake of a user:
-    \[
-        S_{u,t} = (1 - \alpha) S_{u,t-1} + \alpha B_u
-    \]
-    Therefore, we have:
-    \[
-        \sum_{u}S_{u,t} = (1 - \alpha) \sum_{u}S_{u,t - 1} + \alpha \sum_{u}B_u
-    \]
-    This equation shows that when the balance of users is not changing over time the total stake of the system is the
-    exponential average of the total ARGs of the system. Consequently, when we shift the time for \(m\) steps, we can
-    calculate the new total stake of the system from the following equation:
-
-    \[
-        \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
-    \]
-    Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
-    we can obtain \(m\) from the following formula, assuming \(\alpha\) is small enough:
-    \[
-        m = \frac{1}{\alpha} \ln \left(\frac{1 - \gamma}{1 - \lambda}\right)
-    \]
-
-    \section{Incentive mechanism}\label{sec:incentive-mechanism}
-
-    \subsection{Transaction Fee}\label{subsec:transaction-fee}
-
-    Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
-    special method from ARG smart contract. This method will transfer the proposed fee of the transaction in ARGs
-    from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
-    execution fees and \texttt{dbFeeSink} collects fees for ZK-EDB servers. The Protocol decides how to distribute the
-    transaction fee between these two fee sink accounts.
-
-    When a block is added to the blockchain, the proposer of that block will receive a share of the block fees.
-    Consequently, a block proposer is always incentivized to include more transactions in his block. However, if he
-    puts too many transactions in his block and the validation of the block becomes too difficult, some validators
-    may not be able to validate all transactions on time. If a validator can not validate a block in the required
-    time, he will consider the block invalid. So, when a proposed block contains too many transactions, the network
-    may reach consensus on another block, and the proposer of that block will not receive any fees. As a result, a
-    proposer is incentivized to use network transaction capacity optimally.
-
-    On the other hand, we believe that the proposer does not have enough incentives for optimizing the storage size
-    of the transaction set. Therefore, we require that \textbf{the size of the transaction set of every block in
-    bytes be lower than a certain threshold.}
-
-    Validators need to spend resources for validating transactions. When a validator starts the emulation of the AVM
-    to validate a transaction, solely from the code he can't predict the time the execution will finish. This will
-    give an adversary an opportunity to attack the network by broadcasting transactions that never ends. Since,
-    validators can not finish the execution of these transactions, the network will not be able to charge the
-    attacker any fees, and he would be able to waste validators resources for free.
-
-    To mitigate this problem, we require that every transaction specify a cap for all the resources it needs. This
-    will include memory, network and processor related resources. Also, the protocol defines an execution cost for
-    every AVM instruction reflecting the amount of resources its emulation needs. This will define a standard way for
-    measuring the execution cost of any \texttt{avmCall} transaction. Every \texttt{avmCall} transaction is required
-    to specify a maximum execution cost. If during emulation it reaches this maximum cost, the transaction will be
-    considered failed and the network can receive the proposed fee of that transaction.
-
-    Every \texttt{avmCall} transaction is required to provide the following information as an upper bound for the
-    resources it needs:
-
-    \begin{itemize}
-        \item Execution cost
-        \item A list of heap/code-area locations for reading
-        \item A list of heap locations for writing
-        \item A list of heap pages it will deallocate
-        \item Number and size of heap pages it will allocate
-    \end{itemize}
-
-    If a transaction tries to violate any of these predefined limitations, for example, if it tries to read a memory
-    location that is not included in its reading list, it will be considered failed and the network can receive the
-    proposed fee of that transaction.
-
-    \note{A transaction always pays all of its proposed fee, no matter how much of its predefined resources were not
-    used in the final emulation.}
-
-    \subsection{Incentives for ZK-EDB Servers}\label{subsec:zk-edb-servers}
-
-    The incentive mechanism for ZK-EDB servers should have the following properties:
-
-    \begin{itemize}
-        \item It incentivizes storing all memory blocks, whether a heap page or a code area block, and not only those
-        which are used more frequently.
-        \item It incentivizes ZK-EDB servers to actively provide the required memory blocks for validators.
-        \item Making more accounts will not provide any advantages for a ZK-EDB server.
-    \end{itemize}
-
-    For our incentive mechanism, we require that every time a validator receives a memory block from a ZK-EDB, after
-    validating the data, he give a receipt to the ZK-EDB. In this receipt the validator signs the following information:
-
-    \begin{itemize}
-        \item \texttt{ownerAddr}: the ARG address of the ZK-EDB\@.
-        \item \texttt{receivedBlockID}: the ID of the received memory block.
-        \item \texttt{round}: the current round number.
-    \end{itemize}
-
-    \note{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}
-
-    To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ARGs from
-    \texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
-    between all \emph{winning tickets} of the lottery.
-
-    \note{One ZK-EDB server could own multiple winning tickets in a round.}
-
-    To run this lottery, every round, based on the current block seed, a collection of \emph{valid} receipts will be
-    selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in round \(r\) if:
-
-    \begin{itemize}
-        \item The signer was a validator in the round \(r - 1\) and voted for the agreed-upon block.
-        \item The data block in the receipt was needed for validating the \textbf{previous} block.
-        \item The receipt round number is \(r - 1\).
-        \item The signer did not sign a receipt for the same data block for two different ZK-EDBs in the previous round.
-    \end{itemize}
-    For selecting the winning receipts we could use a random generator:
-    \begin{verbatim}
-IF random(seed|validatorPK|receivedBlockID) < winProbability THEN
-    the receipt issued by validatorPK for receivedBlockID is a winner
-    \end{verbatim}
-    \begin{itemize}
-        \item \texttt{random()} produces uniform random numbers between 0 and 1, using its input argument as a seed.
-        \item \texttt{validatorPK} is the public key of the signer of the receipt.
-        \item \texttt{receivedBlockID} is the ID of the memory block that the receipt was issued for.
-        \item \texttt{winProbability} is the probability of winning in every round.
-        \item \texttt{seed} is the current block seed.
-        \item \texttt{|} is a concatenation operator.
-    \end{itemize}
-
-    \note{The winners of the lottery were validators one round before the lottery round.}
-
-    Also, based on the current block seed, a random memory block, whether a heap page or a code area block, is
-    selected as the challenge of the round. A ZK-EDB that owns a winning receipt needs to broadcast a \emph{winning
-    ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
-    challenge. Solving a round challenge requires the content of the memory block which was selected as the round
-    challenge. This will encourage ZK-EDBs to store all memory blocks.
-
-    A possible choice for the challenge solution could be the cryptographic hash of the content of the challenge
-    memory block combined with the ZK-EDB ARG address: \texttt{hash(challenge.content|ownerAddr)}
-
-    The winning tickets of the lottery of round \(r\) need to be included in the block of the round \(r\),
-    otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
-    \(r\) will be done in the round \(r + 1\). This way, \textbf{the content of the challenge memory block could be
-    kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.
-
-    \subsection{Memory Allocation and De-allocation Fee}\label{subsec:memory-allocation-and-de-allocation}
-
-    Every \(k\) round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
-    allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
-    address of the smart contract.
-
-    To determine the price of AVM memory, Every \(k\) round, the protocol calculates \texttt{dbFee} and
-    \texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
-    \texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
-    the protocol considers the total size of all the memory pages that were accessed for either reading or writing
-    during a time period. These two values will be calculated for the last \(k\) rounds and the price per byte of
-    AVM memory will be a linear function of \texttt{dbFee/memTraffic}
-
-    When a smart contract executes a heap de-allocation instruction, the protocol will refund the cost of
-    de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
-    calculates the refunded amount based on the average price the smart contract had paid for that allocated memory.
-    This will prevent smart contracts from profit taking by trading memory with the protocol.
-
-    \section{Concurrency}\label{sec:concurrency}
-
-    \subsection{Memory Dependency Graph}\label{subsec:memory-dependency-graph}
-
-    Every block of the Argennon blockchain contains a list of transactions. This list is an ordered list and the
-    effect of its contained transactions must be applied to the AVM state sequentially as they appear in the ordered
-    list. This ordering is solely chosen by the block proposer, and users should not have any assumptions about
-    the ordering of transactions in a block.
-
-    The fact that block transactions constitute a sequential list, does not mean they can not be executed and applied
-    to the AVM state concurrently. Many transactions are actually independent and the order of their execution does not
-    matter. These transactions can be safely validated in parallel by validators.
-
-    A transaction can change the AVM state by modifying either the code area or the AVM heap. In Argennon, all
-    transactions declare the list of memory locations they want to read or write. This will enable us to determine the
-    independent sets of transactions which can be executed in parallel. To do so, we define the \emph{memory dependency
-    graph} \(G_d\) as follows:
-
-    \begin{itemize}
-        \item \(G_d\) is an undirected graph.
-        \item Every vertex in \(G_d\) corresponds to a transaction and vice versa.
-        \item Vertices \(u\) and \(v\) are adjacent in \(G_d\) if and only if \(u\) has a memory location \(L\) in its
-        writing list and \(v\) has \(L\) in either its writing list or its reading list.
-    \end{itemize}
-
-    If we consider a proper vertex coloring of \(G_d\), every color class will give us an independent set of
-    transactions which can be executed concurrently. To achieve the highest parallelization, we need to color \(G_d\)
-    with minimum number of colors. Thus, the \emph{chromatic number} of the memory dependency graph shows how good a
-    transaction set could be run concurrently.
-
-    Graph coloring is computationally NP-hard. However, in our use case we don't need to necessarily find an optimal
-    solution. An approximate greedy algorithm will perform well enough in most circumstances.
-
-    After constructing the memory dependency graph, we can use it to construct the
-    \emph{execution DAG} of transactions. The execution DAG of transaction set \(T\) is a directed acyclic
-    graph \(G_e = (V_e,E_e)\) which has the \emph{execution invariance} property:
-    \begin{itemize}
-        \item Every vertex in \(V_e\) corresponds to a transaction in \(T\) and vice versa.
-        \item Executing the transactions of \(T\) in any order that \emph{respects} \(G_e\) will result in
-        the same AVM state.
-        \begin{itemize}
-            \item An ordering of transactions of \(T\) respects \(G_e\) if for every directed edge \((u,v) \in E_e\)
-            the transaction \(u\) comes before the transaction \(v\) in the ordering.
-        \end{itemize}
-    \end{itemize}
-
-    Having the execution DAG of a set of transactions, using Algorithm~\ref{alg:exec_dag}, we can apply the transaction
-    set to the AVM state concurrently, using multiple processor, while we can be sure that the resulted AVM state will
-    always be the same no matter how many processor we have used.
-
-    %##\includegraphics[width=17cm]{../img/Alg1s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Ready}{$R_e$}\SetKwData{V}{$v_f$}\SetKwData{Graph}{$G_e$}\SetKwData{Vertices}{$V$}\SetKwData
-        {Txns}{$T$}
-        \KwData{The execution dag $\Graph = (\Vertices,E)$ of transaction set \Txns}
-        \KwResult{The state of the AVM after applying \Txns with any ordering respecting \Graph}
-        \BlankLine
-        \Ready $\gets$ the set of all vertices of \Vertices with in degree 0\;
-        \While{$\Vertices \neq \varnothing$}
-        {
-            wait until a new free processor is available\;
-            \If{the execution of a transaction was finished}
-            {
-                remove the vertex of the finished transaction \V from \Graph\;
-                \For{each vertex $u \in Adj[\V]$}
-                {
-                    \If{$u$ has zero in degree}
-                    {
-                        $\Ready \gets \Ready \cup u$\;
-                    }
-                }
-            }
-            \If{$\Ready \neq \varnothing$}
-            {
-                remove a vertex from \Ready and assign it to a processor\;
-            }
-        }
-        \caption{executing DAG transactions}\label{alg:exec_dag}
-    \end{algorithm}
-
-    By replacing every undirected edge of a memory dependency graph with a directed edge in such a way that the
-    resulted graph has no cycles, we will obtain a valid execution DAG. Thus, from a memory dependency graph different
-    execution DAGs can be constructed with different levels of parallelization ability.
-
-    If we assume that we have unlimited number of processors and all transactions take equal time for executing, it
-    can be shown that by providing a minimal graph coloring to Algorithm~\ref{alg:gen_dag} as input, the resulted
-    DAG will be optimal, in the sense that it results in the minimum overall execution time.
-
-    %##\includegraphics[width=17cm]{../img/Alg2s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Txns}{$T$}\SetKwData{Gd}{$G_d=(V_d,E_d)$}
-        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{The memory dependency graph \Gd of transaction set \Txns\\A proper coloring of $G_d$}
-        \Output{An execution dag $G_e=(V_e,E_e)$ for the transaction set \Txns}
-        \BlankLine
-        $V_e \gets V_d$\;
-        $E_e \gets \varnothing$\;
-        define a total order on colors of $G_d$\;
-        \For{each edge $\{u,v\} \in E_d$}
-        {
-            \eIf{$color[u] < color[v]$}
-            {
-                $E_e \gets E_e \cup (u,v)$\;
-            }{
-                $E_e \gets E_e \cup (v,u)$\;
-            }
-        }
-        \caption{Constructing an execution DAG}\label{alg:gen_dag}
-    \end{algorithm}
-
-    The block proposer is responsible for proposing an efficient execution DAG alongside his proposed block. This
-    execution DAG will determine the ordering of block transactions and help validators to validate transactions in
-    parallel. Since with better parallelization a block can contain more transactions, a proposer is incentivized enough
-    to find a good execution DAG for transactions.
-
-    \subsection{Memory Spooling}\label{subsec:spooling}
-
-    When two transactions are dependant and they are connected with an edge \((u,v)\) in the execution DAG,
-    the transaction \(u\) needs to be run before the transaction \(v\). However, if \(v\) does not read any
-    memory locations that \(u\) modifies, we can run \(u\) and \(v\) in parallel. We just need to make sure
-    \(u\) does not see any changes \(v\) is making in AVM memory. This can be done by appropriate versioning
-    of the memory locations which is shared between \(u\) and \(v\). We call this method \emph{memory spooling}.
-    After enabling memory spooling between two transactions the edge connecting them can be safely removed from the
-    execution DAG\@.
-
-    \subsection{Concurrent Counters}\label{subsec:concurrent-counters}
-
-    We know that in Argennon every transaction needs to transfer its proposed fee to the \texttt{feeSink} accounts
-    first. This essentially makes every transaction a reader and a writer of the memory locations which store the
-    balance record of the \texttt{feeSink} accounts. As a result, all transactions in Argennon will be dependant and
-    parallelism will be completely impossible. Actually, any account that is highly active, for example the account
-    of an exchange or a payment processor, could become a concurrency bottleneck in our system which makes all
-    transactions interacting with them dependant.
-
-    This problem can be easily solved by using a concurrent counter for storing the balance record of this type of
-    accounts. A concurrent counter is a data structure which improves concurrency by using multiple memory locations for
-    storing a single counter. The value of the concurrent counter is equal to the sum of its sub counters and it can
-    be incremented or decremented by incrementing/decrementing any of the sub counters. This way, a concurrent
-    counter trades concurrency with memory usage.
-
-    Algorithm~\ref{alg:CC} implements a concurrent counter which returns an error when the value of the counter
-    becomes negative.
-
-    %##\includegraphics[width=17cm]{../img/Alg3s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{CC}{Counter}
-        \SetKwFunction{Inc}{Increment}\SetKwFunction{Dec}{Decrement}\SetKwFunction{AtomInc}{AtomicIncrement}
-        \SetKwFunction{AtomDec}{AtomicDecrement}\SetKwFunction{AtomSet}{AtomicSet}\SetKwFunction{Get}{GetValue}
-        \SetKwFunction{Acquire}{Lock.Acquire}\SetKwFunction{Release}{Lock.Release}
-        \SetKwProg{Fn}{Function}{}{}
-        \Fn{\Get{\CC}}
-        {
-            $s \gets 0$\;
-            \Acquire{}\;
-            \For{$i \gets 0$ \KwTo $\CC.size - 1$}
-            {
-                $s \gets s + \CC.cell[i]$\;
-            }
-            \Release{}\;
-            \KwRet{s}\;
-        }
-        \BlankLine
-        \Fn{\Inc{\CC, value, seed}}
-        {
-            $i \gets seed \bmod \CC.size$\;
-            \AtomInc{$\CC.cell[i]$, value}\;
-        }
-        \BlankLine
-        \Fn{\Dec{\CC, value, seed, attempt}}
-        {
-            \If {attempt = \CC.size}
-            {
-                restore \CC by adding back the subtracted value\;
-                \KwRet{Error}\;
-            }
-            $i \gets seed \bmod \CC.size$\;
-            $i \gets (i + attempt) \bmod \CC.size$\;
-            \eIf {$\CC.cell[i] \geq value$}
-            {
-                \AtomDec{$\CC.cell[i]$, value}\;
-            }{
-                $r \gets value - \CC.cell[i]$\;
-                \AtomSet{$\CC.cell[i]$, $0$}\;
-                \Dec{\CC, r, seed, $attempt + 1$}\;
-            }
-        }
-        \caption{Concurrent counter}\label{alg:CC}
-    \end{algorithm}
-
-    It should be noted that in a blockchain application we don't have concurrent threads and therefore we don't need
-    atomic functions. For usage in a smart contract, the atomic functions of this pseudocode can be implemented like
-    normal functions.
-
-    Concurrent counter data structure is a part of the AVM standard library, and any smart contract can use this data
-    structure for storing the balance record of highly active accounts.
-
-    \subsection{Memory Chunks}\label{subsec:memory-chunks}
-
-    In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{chunks}.
-    Each memory chunk can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
-    consensus on new values of the commitment of any chunk can be achieved by a different voting committee.
-
-    If a transaction does not modify a memory chunk and in the transaction ordering of the block it comes after
-    any transaction which modifies that chunk, then the execution of that transaction is not needed for calculating
-    the new commitment of the chunk. Consequently, the voting committee of that memory chunk can safely ignore such a
-    transaction. The execution DAG of transactions can be used for finding and pruning these transactions as
-    we see in Algorithm~\ref{alg:prune_dag}.
-
-    %##\includegraphics[width=17cm]{../img/Alg4s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{V}{$v_0$}\SetKwData{Graph}{$G_e$}\SetKwData{Chunk}{$C_k$}\SetKwData{Txns}{$T$}
-        \SetKwFunction{RPrune}{RecursivePrune}
-        \SetKwProg{Fn}{Function}{}{}
-        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{An execution dag \Graph and a memory chunk \Chunk}
-        \Output{The pruned execution dag based on \Chunk}
-        \BlankLine
-        \For{each vertex \V with out degree $0$}
-        {
-            \RPrune{\V}\;
-        }
-        \BlankLine
-        \Fn{\RPrune{\V}}
-        {
-            \If{$Adj[\V] = \varnothing$ {\bf and} \V does not modify \Chunk}
-            {
-                remove \V from \Graph\;
-                \For{each $u$ such that edge $(u,\V)$ was in \Graph}
-                {
-                    \RPrune{u}\;
-                }
-            }
-        }
-        \caption{Pruning an execution DAG}\label{alg:prune_dag}
-    \end{algorithm}
-
-    If we choose chunks in a way that most transactions only modify memory locations of one chunk,
-    likely many transactions of a block only need to be validated by one voting committee and can be validated in
-    parallel by different committees.
-
-    Because the voting committees are selected by random sampling, by choosing large enough samples we can make sure
-    that having multiple voting committees will not change the security properties of the Argennon agreement protocol.
-
-    \section{Smart Contract Oracle}\label{sec:smart-contract-oracle}
-
-    A smart contract oracle is a full AVM emulator that keeps a full local copy of AVM memory and can emulate AVM
-    execution without accessing a ZK-EDB. Smart contract oracles can be used for reporting useful information about
-    \texttt{avmCall} transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
-    and so on.
+    \chapter{The Argennon Blockchain}\label{ch:argennon-blockchain}
+    \input{sections/blockchain}
 

### 2021-04-26 tex/sections/argon.tex
Change project structure
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-03-29 tex/avm.tex
@@
 %! Author = aybehrouz
-%! Date = 2/28/21
 
-% Preamble
-\documentclass[a4paper]{report}
+\section{Introduction}\label{sec:introduction2}
 
-% Packages
-%\usepackage{amsmath}
-%\usepackage[a4paper]{geometry}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{amsfonts}
-\usepackage[ruled,boxed]{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{tcolorbox}
-\lstset{
-%  basicstyle=\ttfamily,
-    mathescape
-}
-%\binoppenalty=99999
-%\relpenalty=99999
-\newcommand{\note}[1] {
-    \begin{tcolorbox}[colframe=white,colback=white]
-        \emph{#1}
-        %\emph{#1}\medskip
-    \end{tcolorbox}
+The Argon programming language is a class-based, object-oriented language designed for writing Argennon's smart
+contracts. The Argon programming language is inspired by Solidity and is similar to Java, with a number of aspects
+of them omitted and a few ideas from other languages included. Argon is designed to be fully compatible with
+the Argennon Virtual Machine and be able to use all advanced features of the Argennon blockchain.
+
+Argon Programs are organized as sets of packages. Each package has its own set of names for types, which helps
+to prevent name conflicts. Every package can contain an arbitrary number of classes and a single contract. Every
+Argon package corresponds to at max one AVM smart contract.
+
+\section{Features Overview}\label{sec:features-overview}
+
+\subsection{Static Classes}\label{subsec:static-classes}
+
+In Argon a static class is an uninstantiable class which all its members and methods are static. Every static
+class can define a special method \texttt{initialize} which can only be called in the \texttt{initialize}
+method of the single contract of the package the static class is defined in.
+
+A Static class is like a normal class which can only have one instance. So like normal classes, static classes can
+be a super class or a subclass of other static classes and can implement interfaces.
+
+\subsection{Contracts}\label{subsec:contract}
+
+Every package of an Argon program can have one or zero contracts. A contact is a special static class which is
+allowed to define methods with \texttt{external} visibility. When an Argon package is deployed as an AVM smart
+contract, these external methods will define the contract's public interface.
+
+\subsection{Access Level Modifiers}\label{subsec:access-level-modifiers}
+
+Access level modifiers determine whether other classes can use a particular field or invoke a particular method or
+if a method can be invoked externally by other smart contracts.
+
+\begin{center}
+    \begin{tabular}{llllll}
+        \hline
+        & Class & Package & Subclass & Program & World \\
+        \hline
+        private   & yes   & no      & no       & no      & no    \\
+        protected & yes   & no      & yes      & no      & no    \\
+        package   & yes   & yes     & yes      & no      & no    \\
+        public    & yes   & yes     & yes      & yes     & no    \\
+        external  & no    & no      & no       & no      & yes   \\
+        \hline
+    \end{tabular}\label{tab:table}
+\end{center}
+
+\begin{lstlisting}[frame=TB, float, title=Argon contracts are static classes,label={lst:code1}]
+// A contract is a static class and all its methods are automatically static.
+contract MirrorToken {
+    private SimpleToken token;
+    private SimpleToken reflection;
+
+    // `initialize` is a special static method that is called by the AVM after the code of a contract
+    // is stored in the AVM code area. It can not be called after that.
+    initialize(double supply1, double supply2) {
+        // `new` does not create a new smart contract. It just makes an ordinary object.
+        token = new SimpleToken(supply1);
+        reflection = new SimpleToken(supply2);
     }
 
-\title{The Argennon Virtual Machine}
-\author{aybehrouz}
-\date{January 2021}
+    external void transfer(account sender, account recipient,
+        double amount, signature sig) {
+        // Checks if the signature is issued by the sender and verifies the calling of this function
+        // with the current parameters.
+        sig.verifyCallBy(sender);
+        token.transfer(sender, recipient, amount);
+        reflection.transfer(recipient, sender, Math.sqrt(amount));
+    }
 
-% Document
-\begin{document}
-    \maketitle
-    \tableofcontents
+    external double balanceOf(account user) { return token.balanceOf(user); }
 
-    \chapter{Specification}\label{ch:specification}
+    external double balanceOfReflection(account user) { return reflection.balanceOf(user); }
+}
 
-    \section{Introduction}\label{sec:introduction}
-
-    The Argennon Virtual Machine (AVM) is an abstract computing machine for executing Argennon's smart contracts. It
-    is designed in a way that it could be efficiently implemented in either hardware or software.
-
-    \section{The Structure of the Argennon Virtual Machine}\label{sec:the-structure-of-the-algorand-virtual-machine}
-
-    ...
-
-    \subsection{Data Types}\label{subsec:data-types}
-
-    The Argennon Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
-    and does not have to be done by the Argennon Virtual Machine itself.
-
-    The instruction set of the Argennon Virtual Machine distinguishes its operand types using instructions intended to
-    operate on values of specific types. For instance, \texttt{iadd} assumes that its operands are two 64-bit integers.
-
-    \subsection{The \texttt{PC} Register}\label{subsec:the-pc-register}
-
-    ...
-
-    \subsection{Call Stack}\label{subsec:call-stack}
-
-    A call stack contains the information that is needed for restoring the state of the invoker of a method.
-
-    \subsection{Memory Unit}\label{subsec:memory-unit}
-
-    The Argennon Virtual Machine has a byte addressable memory which is divided into separate segments. Every segment
-    belongs to a smart contract that has a unique \texttt{applicationID}. The AVM always has a single working memory
-    segment and memory locations outside its current working segment can not be accessed. The only instructions which
-    can change the current working segment are \texttt{invoke\_external}, \texttt{athrow} and return instructions.
-
-    Every smart contract has its own memory segment. Hence, there is no way for a smart contract to access another
-    smart contract's memory. Interaction between smart contracts is done using \texttt{invoke\_external} instruction,
-    and a smart contract can invoke methods of another smart contract by this instruction.
-
-    A memory segment consists of five data areas:
-    \begin{itemize}
-        \item Code Area
-        \item Constant Area
-        \item Local Frame
-        \item Operand Stack
-        \item Heap
-    \end{itemize}
-    All data areas except operand stack, have their own address space which starts from 0. Operand stack is a
-    last-in-first-out (LIFO) stack and is not addressable. Every memory access instruction operates on its specific
-    data areas.
-
-    \subsubsection{Code Area}
-
-    The code area of a segment contains the byte-code of the smart contract which owns that segment. The AVM has no
-    instructions for manipulating the code area. \textbf{Installing, removing and updating smart contracts need to
-    be done externally.}
-
-    \subsubsection{Constant Area}
-
-    The constant area of a segment contains several kinds of constants, ranging from user defined constants to method
-    address tables. A method address table stores method locations in the code area and their access type. The access
-    type of a method can be either \texttt{public} or \texttt{private}. The AVM has no instructions for modifying the
-    constant area.
-
-    \note{Only public methods can be invoked by \texttt{invoke\_external} instruction.}
-
-    \subsubsection{Local Frame}
-
-    A local frame is used to store methods parameters and local variables. A new frame is created each time a method
-    is invoked, and it is destroyed when its method invocation completes, whether the completion is normal or abrupt.
-
-    \subsubsection{Operand Stack}
-
-    Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too. AVM
-    instructions take operands from the operand stack, operate on them, and push the result back onto the operand
-    stack. An operand stack is destroyed when its owner method completes, whether that completion is normal or abrupt.
-
-    \subsubsection{Heap}
-
-    The heap of a segment is a persistent memory area which is divided into pages. Memory locations inside every page
-    have a separate address space that starts from 0, and each page can be referenced by an index that starts from 0. In
-    other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex, offset)}.
-    Different pages of a heap do not need to be equally sized.
-
-    \note{The reason behind this paged design is that the heap is usually persisted using a block device. A heap with
-    a paged structure could expose the underlying block based nature of the persistence layer to the application
-    layer. In this way, the compiler or the programmer could better optimize the code for the persistence layer.}
-
-    \section{Instruction Set Summary}\label{sec:instruction-set-summary}
-
-    An Argennon Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
-    performed, followed by zero or more operands supplying arguments or data that are used by the operation.
-    \textbf{The number and size of the operands are determined solely by the opcode.}
-
-    \subsection{Method Invocation}\label{subsec:method-invocation}
-
-    The Argennon Virtual Machine has three types of method invocation:
-    \begin{itemize}
-        \item \texttt{invoke\_internal} invokes a method from the current running smart contract.
-        \item \texttt{invoke\_external} invokes a \texttt{public} method from another smart contract. It will change
-        the current memory segment to the segment of the invoked smart contract.
-        \item \texttt{invoke\_native} invokes a method that is not hosted by the Argennon virtual machine. By this
-        instruction, high performance native methods of the hosting machine could become available to AVM smart
-        contracts.
-    \end{itemize}
-
-    \note{In the future, we may need to add special instructions for invoking interface and virtual methods...}
-
-    Each time a method is invoked a new local frame and operand stack is created. The Argennon Virtual Machine uses
-    local frames to pass parameters on method invocation. On method invocation, any parameters are passed in
-    consecutive local variables stored in the method's local frame starting from local variable 0. The invoker of a
-    method writes the parameters in the local frame of the invoked method using \texttt{arg} instructions.
-
-    \subsubsection{Exceptions}
-
-    An exception is thrown programmatically using the \texttt{athrow} instruction. Exceptions can also be thrown by
-    various Argennon Virtual Machine instructions if they detect an abnormal condition. Some exceptions are not
-    catchable and will always abort the execution of the smart contract.
-
-    \subsubsection{Method Invocation Completion}
-
-    A method invocation completes normally if that invocation does not cause an exception to be thrown, either
-    directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
-    method completes normally, then a value may be returned to the invoking method. This occurs when the invoked
-    method executes one of the return instructions, the choice of which must be appropriate for the type of the value
-    being returned (if any). Execution then continues normally in the invoking method's local frame with the returned
-    value (if any) pushed onto the operand stack.
-
-    A method invocation completes abruptly if an exceptions is thrown and is not caught by the current method. A
-    method invocation that completes abruptly never returns a value to its invoker.
-
-    When a method completes, whether normally or abruptly, the call stack is used to restore the state of the invoker,
-    including its local frame and operand stack, with the \texttt{PC} register appropriately restored and incremented
-    to skip past the method invocation instruction. If the invoker was another smart contract, i.e.\ the invocation
-    was made by an \texttt{invoke\_external} instruction, the current memory segment will be changed to the invoker's
-    segment.
-
-    A thrown exception causes methods in the call stack to complete \textbf{abruptly} one by one, as long as the
-    \texttt{PC} register is not pointing to a \texttt{catch} instruction. The \texttt{catch} instruction acts like a
-    branch instruction that branches only if an exception is caught. \textbf{When an external method invocation
-    completes abruptly, before changing the current segment, all changes made to the heap area by that method
-    call, including changes made to other segments, will be rolled back.} So, when a method call completes abruptly,
-    that method call essentially has no effect on the AVM state.
-
-    \note{By using the \texttt{athrow} instruction properly, a programmer can make any method act like an atomic
-    operation.}
-
-    \subsection{Authorizing Operations}\label{subsec:authorizing-operations}
-
-    In blockchain applications, we usually need to authorize certain operations. For example, for sending an asset
-    from a user to another user, first we need to make sure that the sender has authorized this operation. The
-    Argennon virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
-    cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
-    passing signatures as parameters to methods, a programmer can implement the required logic for authorizing any
-    operation.
-
-    \note{The Argennon virtual machine has no instructions for issuing cryptographic signatures.}
-
-    In addition to signatures, a method can verify its invoker by using \texttt{get\_parent} instruction. This
-    instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
-    smart contract in the call stack. In other words, it returns the \texttt{applicationID} of the smart contract that
-    has invoked the current smart contract. (if any)
-
-    \subsection{Heap Allocation Instructions}\label{subsec:heap-allocation-instructions}
-
-    ...
-
-    \section{The AVM Standard Library}\label{sec:the-avm-standard-library}
-
-    ...
-
-    \chapter{Implementation}\label{ch:implementation}
-
-    \section{Persistence}\label{sec:persistence}
-
-    For implementing the persistence layer of the AVM, we assume that we have access to an updatable zero-knowledge
-    elementary database (ZK-EDB) with the following properties:
-
-    \begin{itemize}
-        \item The ZK-EDB contains a mapping from a set of keys to a set of values.
-        \item Every state of the database has a commitment \(C\).
-        \item The ZK-EDB has a method \((D, p) = get(x)\), where \(x\) is a key and \(D\) is the associated data
-        with \(x\), and \(p\) is a proof.
-        \item A user can use \(C\) and \(p\) to verify that \(D\) is really associated with \(x\), and \(D\) is not
-        altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB\@.
-        \item Having \(p\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
-        associated with \(x\) instead of \(D\).
-    \end{itemize}
-
-    We use a ZK-EDB for storing the AVM heap. We include the commitment of the current state of this DB in every
-    block of the Argennon blockchain, so ZK-EDB servers need not be trusted servers.
-
-    Every page of AVM heap will be stored with a key of the form: \texttt{applicationID|pageIndex} (the \texttt{|}
-    operator concatenates two numbers). Nodes do not keep a full copy of the AVM heap and for validating block
-    certificates or emulating the AVM ( i.e.~validating transactions) they need to connect to a ZK-EDB and retrieve
-    the required pages of AVM heap. For better performance, nodes keep a cache of heap pages to
-    reduce the amount of ZK-EDB access.
-
-    We also use a ZK-EDB for storing the code area of each segment, and we include the commitment of this DB in every
-    block. Every code area will be divided into blocks and every block will be stored in the DB with
-    \texttt{applicationID|blockID} as its key. Like heap pages, nodes keep a cache of code area blocks.
-
-    \note{Unlike heap pages, the AVM is not aware of different blocks of code area.}
-
-    \section{Transactions}\label{sec:transactions}
-
-    Argennon has four types of transaction:
-
-    \begin{itemize}
-        \item \texttt{avmCall} essentially is an \texttt{invoke\_external} instruction that invokes a method from an
-        AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
-        assets, including ARGs, is done by these transactions.
-        \item \texttt{installApp} installs an AVM smart contract and determines the update policy of the smart
-        contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
-        on.
-        \item \texttt{unInstallApp} removes an AVM smart contract.
-        \item \texttt{updateApp} updates an AVM smart contract.
-    \end{itemize}
-
-    All types of Argennon transactions contain an \texttt{invoke\_external} instruction which calls a special method
-    from ARG smart contract that transfers the proposed fee of the transaction in ARGs from a sender account to the
-    fee sink accounts.
-
-    Every transaction is required to exactly specify what heap locations or code area addresses it will access. This
-    enables validators to start retrieving the required memory blocks from available ZK-EDB servers as soon as they
-    see a transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to
-    access a memory location that is not included in its access lists, will be rejected. Users could use smart contract
-    oracles to predict the list of memory blocks their transactions need. See Section~\ref{sec:smart-contract-oracle}
-    for more details.
-
-    \section{Blockchain}\label{sec:blockchain}
-
-    Every block of the Argennon blockchain corresponds to a set of transactions. We store the commitment of this
-    transaction set in every block, but we don't keep the set itself. To be able to detect replay attacks, we require
-    every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
-    and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
-    must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
-    a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
-    if a signature of the same user with a bigger or equal nonce is already used
-    (i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
-    nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
-
-    To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. When the
-    difference between \texttt{issuance} component of this nonce and the current round becomes bigger than the
-    maximum allowed lifetime of a signature, this information can be safely deleted. \textbf{As a result, we will not
-    have the problem of "un-removable empty accounts" like Ethereum.}
-
-    The only information that Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
-    blockchain. Every block of the Argennon blockchain contains the following information:
-
-    \begin{center}
-        \begin{tabular}{||c||}
-            \hline
-            Block \\ [0.5ex]
-            \hline\hline
-            commitment to the ZK-EDB storing heap pages \\ [0.7ex]
-            commitment to the ZK-EDB storing code areas \\ [0.7ex]
-            commitment to the set of transactions       \\ [0.7ex]
-            previous block hash                         \\ [0.7ex]
-            random seed                                 \\ [0.7ex]
-            \hline
-        \end{tabular}
-    \end{center}
-
-    For confirming a new block, nodes that are not validators only need to verify the block certificate. For
-    verifying a block certificate, a node needs to know the ARG balances of validators, but it doesn't need to
-    emulate the AVM execution.
-
-    On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
-    execution of the Argennon virtual machine. To do so, first they retrieve all heap pages and code area blocks they
-    need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
-    transactions included in the new block. This will modify some pages of the AVM memory, so they update the ZK-EDB
-    commitments based on the modified pages and verify the commitments included in the new block. Validators also
-    calculate and verify the commitment to the new block's transaction set.
-
-    \note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDB servers will receive the new
-    block, and they will update their database by emulating the AVM execution.}
-
-    \section{Consensus}\label{sec:consensus}
-
-    \note{It will be a randomized proof of stake which is under development....}
-
-    \subsection{Estimating A User's Stake}\label{subsec:estimating-a-user's-stake}
-
-    In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
-    of stake the user has in the system. Conventionally in these systems, for estimating a user's stake, we use the
-    amount of native system tokens the user is holding. Unfortunately, one problem with this approach is that a
-    strong attacker may be able to obtain a considerable amount of system tokens, for example by borrowing from a
-    DEFI application, and use this stake to attack the system.
-
-    To mitigate this problem, for calculating a user's stake at time step \(t\), instead of using the raw ARG
-    balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
-    ARG balance:
-    \[
-        S_{u,t} = \min (B_{u,t}, Trust_{u,t})
-    \]
-    Where:
-    \begin{itemize}
-        \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
-        \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
-        \item \(Trust_{u,t}\) is an estimated trust value for user \(u\) at time step \(t\).
-    \end{itemize}
-
-    The agreement protocol, at time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
-    number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
-    exponential moving average of the ARG balance of user \(u\) at time step \(t\).
-
-    In our system a user who held ARGs and participated in the consensus for a long time is more trusted
-    than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
-    amount of ARGs, also needs to hold them for a long period of time before being able to attack the system.
-
-    For calculating the exponential moving average of a user's balance at time step \(t\), we can use the following
-    recursive formula:
-    \[
-        M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
-    \]
-    Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
-    degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
-
-    Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
-    \(M_{u,t}\): (In the following equations the \(u\) subscript is dropped for simplicity)
-    \[
-        M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B
-    \]
-    Where:
-    \[
-        B = B_{k+1} = B_{k+2} = \dots = B_{t}
-    \]
-    We know that when \(|nx| \ll 1\) we can use the binomial approximation \({(1 + x)^n \approx 1 + nx}\). So, we can
-    further simplify this formula:
-    \[
-        M_{t} = M_{k} + (t - k) \alpha (B - M_{k})
-    \]
-
-    For choosing the value of \(\alpha\) we can consider the number of time steps that the trust value of a user needs
-    for reaching a specified fraction of his account balance. We know that for large \(n\) and \(|x| < 1\) we have
-    \((1 + x)^n \approx e^{nx}\), so by letting \(M_{u,k} = 0\) and \(n = t - k\) we can write:
-    \[
-        \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{B}\right)}{n}
-    \]
-    The value of \(\alpha\) for a desired configuration can be calculated by this equation. For instance, we could
-    calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8B\) and \(n\) equals to the
-    number of time steps of 10 years.
-
-    In our system a newly created account will not have voting power for some time, no matter how high its
-    balance is. While this is a desirable property, in case a large proportion of total system tokens are
-    transferred to newly created accounts, it can result in too much voting power for older accounts. This may decrease
-    the degree of decentralization in our system.
-
-    However, this situation is easily detectable by comparing the total stake of the system with the total balance of
-    users. If after confirming a block the total stake of the system goes too low and we have:
-    \[
-        \sum_{u}S_{u,t} < \gamma \sum_{u}B_{u,t}
-    \]
-    The protocol will perform a \emph{time shift} in the system: the time step of the system
-    will be incremented for \(m\) steps while no blocks will be confirmed. This will increase the value of \(M_{u,t}\)
-    for new accounts with a non-zero balance, giving them more influence in the agreement protocol.
-
-    For calculating the value of \(m\) which determines the amount of time shift in the system, we should note that when
-    \(B_{u,t} = B_{u, t-1} = B_u\), we can derive a simple recursive rule for the stake of a user:
-    \[
-        S_{u,t} = (1 - \alpha) S_{u,t-1} + \alpha B_u
-    \]
-    Therefore, we have:
-    \[
-        \sum_{u}S_{u,t} = (1 - \alpha) \sum_{u}S_{u,t - 1} + \alpha \sum_{u}B_u
-    \]
-    This equation shows that when the balance of users is not changing over time the total stake of the system is the
-    exponential average of the total ARGs of the system. Consequently, when we shift the time for \(m\) steps, we can
-    calculate the new total stake of the system from the following equation:
-
-    \[
-        \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
-    \]
-    Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
-    we can obtain \(m\) from the following formula, assuming \(\alpha\) is small enough:
-    \[
-        m = \frac{1}{\alpha} \ln \left(\frac{1 - \gamma}{1 - \lambda}\right)
-    \]
-
-    \section{Incentive mechanism}\label{sec:incentive-mechanism}
-
-    \subsection{Transaction Fee}\label{subsec:transaction-fee}
-
-    Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
-    special method from ARG smart contract. This method will transfer the proposed fee of the transaction in ARGs
-    from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
-    execution fees and \texttt{dbFeeSink} collects fees for ZK-EDB servers. The Protocol decides how to distribute the
-    transaction fee between these two fee sink accounts.
-
-    When a block is added to the blockchain, the proposer of that block will receive a share of the block fees.
-    Consequently, a block proposer is always incentivized to include more transactions in his block. However, if he
-    puts too many transactions in his block and the validation of the block becomes too difficult, some validators
-    may not be able to validate all transactions on time. If a validator can not validate a block in the required
-    time, he will consider the block invalid. So, when a proposed block contains too many transactions, the network
-    may reach consensus on another block, and the proposer of that block will not receive any fees. As a result, a
-    proposer is incentivized to use network transaction capacity optimally.
-
-    On the other hand, we believe that the proposer does not have enough incentives for optimizing the storage size
-    of the transaction set. Therefore, we require that \textbf{the size of the transaction set of every block in
-    bytes be lower than a certain threshold.}
-
-    Validators need to spend resources for validating transactions. When a validator starts the emulation of the AVM
-    to validate a transaction, solely from the code he can't predict the time the execution will finish. This will
-    give an adversary an opportunity to attack the network by broadcasting transactions that never ends. Since,
-    validators can not finish the execution of these transactions, the network will not be able to charge the
-    attacker any fees, and he would be able to waste validators resources for free.
-
-    To mitigate this problem, we require that every transaction specify a cap for all the resources it needs. This
-    will include memory, network and processor related resources. Also, the protocol defines an execution cost for
-    every AVM instruction reflecting the amount of resources its emulation needs. This will define a standard way for
-    measuring the execution cost of any \texttt{avmCall} transaction. Every \texttt{avmCall} transaction is required
-    to specify a maximum execution cost. If during emulation it reaches this maximum cost, the transaction will be
-    considered failed and the network can receive the proposed fee of that transaction.
-
-    Every \texttt{avmCall} transaction is required to provide the following information as an upper bound for the
-    resources it needs:
-
-    \begin{itemize}
-        \item Execution cost
-        \item A list of heap/code-area locations for reading
-        \item A list of heap locations for writing
-        \item A list of heap pages it will deallocate
-        \item Number and size of heap pages it will allocate
-    \end{itemize}
-
-    If a transaction tries to violate any of these predefined limitations, for example, if it tries to read a memory
-    location that is not included in its reading list, it will be considered failed and the network can receive the
-    proposed fee of that transaction.
-
-    \note{A transaction always pays all of its proposed fee, no matter how much of its predefined resources were not
-    used in the final emulation.}
-
-    \subsection{Incentives for ZK-EDB Servers}\label{subsec:zk-edb-servers}
-
-    The incentive mechanism for ZK-EDB servers should have the following properties:
-
-    \begin{itemize}
-        \item It incentivizes storing all memory blocks, whether a heap page or a code area block, and not only those
-        which are used more frequently.
-        \item It incentivizes ZK-EDB servers to actively provide the required memory blocks for validators.
-        \item Making more accounts will not provide any advantages for a ZK-EDB server.
-    \end{itemize}
-
-    For our incentive mechanism, we require that every time a validator receives a memory block from a ZK-EDB, after
-    validating the data, he give a receipt to the ZK-EDB. In this receipt the validator signs the following information:
-
-    \begin{itemize}
-        \item \texttt{ownerAddr}: the ARG address of the ZK-EDB\@.
-        \item \texttt{receivedBlockID}: the ID of the received memory block.
-        \item \texttt{round}: the current round number.
-    \end{itemize}
-
-    \note{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}
-
-    To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ARGs from
-    \texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
-    between all \emph{winning tickets} of the lottery.
-
-    \note{One ZK-EDB server could own multiple winning tickets in a round.}
-
-    To run this lottery, every round, based on the current block seed, a collection of \emph{valid} receipts will be
-    selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in round \(r\) if:
-
-    \begin{itemize}
-        \item The signer was a validator in the round \(r - 1\) and voted for the agreed-upon block.
-        \item The data block in the receipt was needed for validating the \textbf{previous} block.
-        \item The receipt round number is \(r - 1\).
-        \item The signer did not sign a receipt for the same data block for two different ZK-EDBs in the previous round.
-    \end{itemize}
-    For selecting the winning receipts we could use a random generator:
-    \begin{verbatim}
-IF random(seed|validatorPK|receivedBlockID) < winProbability THEN
-    the receipt issued by validatorPK for receivedBlockID is a winner
-    \end{verbatim}
-    \begin{itemize}
-        \item \texttt{random()} produces uniform random numbers between 0 and 1, using its input argument as a seed.
-        \item \texttt{validatorPK} is the public key of the signer of the receipt.
-        \item \texttt{receivedBlockID} is the ID of the memory block that the receipt was issued for.
-        \item \texttt{winProbability} is the probability of winning in every round.
-        \item \texttt{seed} is the current block seed.
-        \item \texttt{|} is a concatenation operator.
-    \end{itemize}
-
-    \note{The winners of the lottery were validators one round before the lottery round.}
-
-    Also, based on the current block seed, a random memory block, whether a heap page or a code area block, is
-    selected as the challenge of the round. A ZK-EDB that owns a winning receipt needs to broadcast a \emph{winning
-    ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
-    challenge. Solving a round challenge requires the content of the memory block which was selected as the round
-    challenge. This will encourage ZK-EDBs to store all memory blocks.
-
-    A possible choice for the challenge solution could be the cryptographic hash of the content of the challenge
-    memory block combined with the ZK-EDB ARG address: \texttt{hash(challenge.content|ownerAddr)}
-
-    The winning tickets of the lottery of round \(r\) need to be included in the block of the round \(r\),
-    otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
-    \(r\) will be done in the round \(r + 1\). This way, \textbf{the content of the challenge memory block could be
-    kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.
-
-    \subsection{Memory Allocation and De-allocation Fee}\label{subsec:memory-allocation-and-de-allocation}
-
-    Every \(k\) round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
-    allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
-    address of the smart contract.
-
-    To determine the price of AVM memory, Every \(k\) round, the protocol calculates \texttt{dbFee} and
-    \texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
-    \texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
-    the protocol considers the total size of all the memory pages that were accessed for either reading or writing
-    during a time period. These two values will be calculated for the last \(k\) rounds and the price per byte of
-    AVM memory will be a linear function of \texttt{dbFee/memTraffic}
-
-    When a smart contract executes a heap de-allocation instruction, the protocol will refund the cost of
-    de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
-    calculates the refunded amount based on the average price the smart contract had paid for that allocated memory.
-    This will prevent smart contracts from profit taking by trading memory with the protocol.
-
-    \section{Concurrency}\label{sec:concurrency}
-
-    \subsection{Memory Dependency Graph}\label{subsec:memory-dependency-graph}
-
-    Every block of the Argennon blockchain contains a list of transactions. This list is an ordered list and the
-    effect of its contained transactions must be applied to the AVM state sequentially as they appear in the ordered
-    list. This ordering is solely chosen by the block proposer, and users should not have any assumptions about
-    the ordering of transactions in a block.
-
-    The fact that block transactions constitute a sequential list, does not mean they can not be executed and applied
-    to the AVM state concurrently. Many transactions are actually independent and the order of their execution does not
-    matter. These transactions can be safely validated in parallel by validators.
-
-    A transaction can change the AVM state by modifying either the code area or the AVM heap. In Argennon, all
-    transactions declare the list of memory locations they want to read or write. This will enable us to determine the
-    independent sets of transactions which can be executed in parallel. To do so, we define the \emph{memory dependency
-    graph} \(G_d\) as follows:
-
-    \begin{itemize}
-        \item \(G_d\) is an undirected graph.
-        \item Every vertex in \(G_d\) corresponds to a transaction and vice versa.
-        \item Vertices \(u\) and \(v\) are adjacent in \(G_d\) if and only if \(u\) has a memory location \(L\) in its
-        writing list and \(v\) has \(L\) in either its writing list or its reading list.
-    \end{itemize}
-
-    If we consider a proper vertex coloring of \(G_d\), every color class will give us an independent set of
-    transactions which can be executed concurrently. To achieve the highest parallelization, we need to color \(G_d\)
-    with minimum number of colors. Thus, the \emph{chromatic number} of the memory dependency graph shows how good a
-    transaction set could be run concurrently.
-
-    Graph coloring is computationally NP-hard. However, in our use case we don't need to necessarily find an optimal
-    solution. An approximate greedy algorithm will perform well enough in most circumstances.
-
-    After constructing the memory dependency graph, we can use it to construct the
-    \emph{execution DAG} of transactions. The execution DAG of transaction set \(T\) is a directed acyclic
-    graph \(G_e = (V_e,E_e)\) which has the \emph{execution invariance} property:
-    \begin{itemize}
-        \item Every vertex in \(V_e\) corresponds to a transaction in \(T\) and vice versa.
-        \item Executing the transactions of \(T\) in any order that \emph{respects} \(G_e\) will result in
-        the same AVM state.
-        \begin{itemize}
-            \item An ordering of transactions of \(T\) respects \(G_e\) if for every directed edge \((u,v) \in E_e\)
-            the transaction \(u\) comes before the transaction \(v\) in the ordering.
-        \end{itemize}
-    \end{itemize}
-
-    Having the execution DAG of a set of transactions, using Algorithm~\ref{alg:exec_dag}, we can apply the transaction
-    set to the AVM state concurrently, using multiple processor, while we can be sure that the resulted AVM state will
-    always be the same no matter how many processor we have used.
-
-    %##\includegraphics[width=17cm]{../img/Alg1s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Ready}{$R_e$}\SetKwData{V}{$v_f$}\SetKwData{Graph}{$G_e$}\SetKwData{Vertices}{$V$}\SetKwData
-        {Txns}{$T$}
-        \KwData{The execution dag $\Graph = (\Vertices,E)$ of transaction set \Txns}
-        \KwResult{The state of the AVM after applying \Txns with any ordering respecting \Graph}
-        \BlankLine
-        \Ready $\gets$ the set of all vertices of \Vertices with in degree 0\;
-        \While{$\Vertices \neq \varnothing$}
-        {
-            wait until a new free processor is available\;
-            \If{the execution of a transaction was finished}
-            {
-                remove the vertex of the finished transaction \V from \Graph\;
-                \For{each vertex $u \in Adj[\V]$}
-                {
-                    \If{$u$ has zero in degree}
-                    {
-                        $\Ready \gets \Ready \cup u$\;
+public static class Math {
+    public double sqrt(double x) {
+        return // calculate square root of x
     }
 }
+
+package class SimpleToken {
+    private map(account -> double) balances;
+
+    // The visibility of a member without an access modifier will be the same as its defining
+    // class so `SimpleToken` constructor has package visibility.
+    constructor(double initialSupply) {
+        // initializes the object
     }
-            \If{$\Ready \neq \varnothing$}
-            {
-                remove a vertex from \Ready and assign it to a processor\;
+
+    void transfer(account sender, account recipient, double amount) {
+        if (balances[sender] < amount) throw("Not enough balance.");
+        // implements the required logic...
     }
+    // implements other methods...
 }
-        \caption{executing DAG transactions}\label{alg:exec_dag}
-    \end{algorithm}
+\end{lstlisting}
 
-    By replacing every undirected edge of a memory dependency graph with a directed edge in such a way that the
-    resulted graph has no cycles, we will obtain a valid execution DAG. Thus, from a memory dependency graph different
-    execution DAGs can be constructed with different levels of parallelization ability.
+\subsection{Shadowing}\label{subsec:shadowing}
 
-    If we assume that we have unlimited number of processors and all transactions take equal time for executing, it
-    can be shown that by providing a minimal graph coloring to Algorithm~\ref{alg:gen_dag} as input, the resulted
-    DAG will be optimal, in the sense that it results in the minimum overall execution time.
-
-    %##\includegraphics[width=17cm]{../img/Alg2s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Txns}{$T$}\SetKwData{Gd}{$G_d=(V_d,E_d)$}
-        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{The memory dependency graph \Gd of transaction set \Txns\\A proper coloring of $G_d$}
-        \Output{An execution dag $G_e=(V_e,E_e)$ for the transaction set \Txns}
-        \BlankLine
-        $V_e \gets V_d$\;
-        $E_e \gets \varnothing$\;
-        define a total order on colors of $G_d$\;
-        \For{each edge $\{u,v\} \in E_d$}
-        {
-            \eIf{$color[u] < color[v]$}
-            {
-                $E_e \gets E_e \cup (u,v)$\;
-            }{
-                $E_e \gets E_e \cup (v,u)$\;
-            }
-        }
-        \caption{Constructing an execution DAG}\label{alg:gen_dag}
-    \end{algorithm}
-
-    The block proposer is responsible for proposing an efficient execution DAG alongside his proposed block. This
-    execution DAG will determine the ordering of block transactions and help validators to validate transactions in
-    parallel. Since with better parallelization a block can contain more transactions, a proposer is incentivized enough
-    to find a good execution DAG for transactions.
-
-    \subsection{Memory Spooling}\label{subsec:spooling}
-
-    When two transactions are dependant and they are connected with an edge \((u,v)\) in the execution DAG,
-    the transaction \(u\) needs to be run before the transaction \(v\). However, if \(v\) does not read any
-    memory locations that \(u\) modifies, we can run \(u\) and \(v\) in parallel. We just need to make sure
-    \(u\) does not see any changes \(v\) is making in AVM memory. This can be done by appropriate versioning
-    of the memory locations which is shared between \(u\) and \(v\). We call this method \emph{memory spooling}.
-    After enabling memory spooling between two transactions the edge connecting them can be safely removed from the
-    execution DAG\@.
-
-    \subsection{Concurrent Counters}\label{subsec:concurrent-counters}
-
-    We know that in Argennon every transaction needs to transfer its proposed fee to the \texttt{feeSink} accounts
-    first. This essentially makes every transaction a reader and a writer of the memory locations which store the
-    balance record of the \texttt{feeSink} accounts. As a result, all transactions in Argennon will be dependant and
-    parallelism will be completely impossible. Actually, any account that is highly active, for example the account
-    of an exchange or a payment processor, could become a concurrency bottleneck in our system which makes all
-    transactions interacting with them dependant.
-
-    This problem can be easily solved by using a concurrent counter for storing the balance record of this type of
-    accounts. A concurrent counter is a data structure which improves concurrency by using multiple memory locations for
-    storing a single counter. The value of the concurrent counter is equal to the sum of its sub counters and it can
-    be incremented or decremented by incrementing/decrementing any of the sub counters. This way, a concurrent
-    counter trades concurrency with memory usage.
-
-    Algorithm~\ref{alg:CC} implements a concurrent counter which returns an error when the value of the counter
-    becomes negative.
-
-    %##\includegraphics[width=17cm]{../img/Alg3s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{CC}{Counter}
-        \SetKwFunction{Inc}{Increment}\SetKwFunction{Dec}{Decrement}\SetKwFunction{AtomInc}{AtomicIncrement}
-        \SetKwFunction{AtomDec}{AtomicDecrement}\SetKwFunction{AtomSet}{AtomicSet}\SetKwFunction{Get}{GetValue}
-        \SetKwFunction{Acquire}{Lock.Acquire}\SetKwFunction{Release}{Lock.Release}
-        \SetKwProg{Fn}{Function}{}{}
-        \Fn{\Get{\CC}}
-        {
-            $s \gets 0$\;
-            \Acquire{}\;
-            \For{$i \gets 0$ \KwTo $\CC.size - 1$}
-            {
-                $s \gets s + \CC.cell[i]$\;
-            }
-            \Release{}\;
-            \KwRet{s}\;
-        }
-        \BlankLine
-        \Fn{\Inc{\CC, value, seed}}
-        {
-            $i \gets seed \bmod \CC.size$\;
-            \AtomInc{$\CC.cell[i]$, value}\;
-        }
-        \BlankLine
-        \Fn{\Dec{\CC, value, seed, attempt}}
-        {
-            \If {attempt = \CC.size}
-            {
-                restore \CC by adding back the subtracted value\;
-                \KwRet{Error}\;
-            }
-            $i \gets seed \bmod \CC.size$\;
-            $i \gets (i + attempt) \bmod \CC.size$\;
-            \eIf {$\CC.cell[i] \geq value$}
-            {
-                \AtomDec{$\CC.cell[i]$, value}\;
-            }{
-                $r \gets value - \CC.cell[i]$\;
-                \AtomSet{$\CC.cell[i]$, $0$}\;
-                \Dec{\CC, r, seed, $attempt + 1$}\;
-            }
-        }
-        \caption{Concurrent counter}\label{alg:CC}
-    \end{algorithm}
-
-    It should be noted that in a blockchain application we don't have concurrent threads and therefore we don't need
-    atomic functions. For usage in a smart contract, the atomic functions of this pseudocode can be implemented like
-    normal functions.
-
-    Concurrent counter data structure is a part of the AVM standard library, and any smart contract can use this data
-    structure for storing the balance record of highly active accounts.
-
-    \subsection{Memory Chunks}\label{subsec:memory-chunks}
-
-    In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{chunks}.
-    Each memory chunk can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
-    consensus on new values of the commitment of any chunk can be achieved by a different voting committee.
-
-    If a transaction does not modify a memory chunk and in the transaction ordering of the block it comes after
-    any transaction which modifies that chunk, then the execution of that transaction is not needed for calculating
-    the new commitment of the chunk. Consequently, the voting committee of that memory chunk can safely ignore such a
-    transaction. The execution DAG of transactions can be used for finding and pruning these transactions as
-    we see in Algorithm~\ref{alg:prune_dag}.
-
-    %##\includegraphics[width=17cm]{../img/Alg4s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{V}{$v_0$}\SetKwData{Graph}{$G_e$}\SetKwData{Chunk}{$C_k$}\SetKwData{Txns}{$T$}
-        \SetKwFunction{RPrune}{RecursivePrune}
-        \SetKwProg{Fn}{Function}{}{}
-        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{An execution dag \Graph and a memory chunk \Chunk}
-        \Output{The pruned execution dag based on \Chunk}
-        \BlankLine
-        \For{each vertex \V with out degree $0$}
-        {
-            \RPrune{\V}\;
-        }
-        \BlankLine
-        \Fn{\RPrune{\V}}
-        {
-            \If{$Adj[\V] = \varnothing$ {\bf and} \V does not modify \Chunk}
-            {
-                remove \V from \Graph\;
-                \For{each $u$ such that edge $(u,\V)$ was in \Graph}
-                {
-                    \RPrune{u}\;
-                }
-            }
-        }
-        \caption{Pruning an execution DAG}\label{alg:prune_dag}
-    \end{algorithm}
-
-    If we choose chunks in a way that most transactions only modify memory locations of one chunk,
-    likely many transactions of a block only need to be validated by one voting committee and can be validated in
-    parallel by different committees.
-
-    Because the voting committees are selected by random sampling, by choosing large enough samples we can make sure
-    that having multiple voting committees will not change the security properties of the Argennon agreement protocol.
-
-    \section{Smart Contract Oracle}\label{sec:smart-contract-oracle}
-
-    A smart contract oracle is a full AVM emulator that keeps a full local copy of AVM memory and can emulate AVM
-    execution without accessing a ZK-EDB. Smart contract oracles can be used for reporting useful information about
-    \texttt{avmCall} transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
-    and so on.
-
-\end{document}
+If a declaration of a type (such as a member variable or a parameter name) in a particular scope (such as an
+inner block or a method definition) has the same name as another declaration in the enclosing scope, it will
+result in a compiler error. In other words, the Argon programming language does not allow shadowing.

### 2021-04-26 tex/sections/avm.tex
Change project structure
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-04-25 tex/avm.tex
@@
-%! suppress = TooLargeSection
 %! Author = aybehrouz
-%! Date = 2/28/21
-
-% Preamble
-\documentclass[11pt, a4paper]{report}
-\usepackage[a4paper]{geometry}
-
-% Packages
-%\usepackage{amsmath}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{amsfonts}
-\usepackage[ruled,boxed]{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{tcolorbox}
-\usepackage{amsmath}
-
-\lstset{
-    language=Java,
-    morekeywords={contract, initialize, constructor, external, virtual, map, account, signature, overrides, is, string},
-    basicstyle=\small\sffamily,
-    columns=fullflexible,
-%   keywordstyle=\bfseries,
-    breaklines=true,
-    showstringspaces=false,
-    mathescape
-}
-%\binoppenalty=99999
-%\relpenalty=99999
-\newcommand{\note}[1] {
-    \begin{tcolorbox}[colframe=white,colback=white]
-        \emph{#1}
-        %\emph{#1}\medskip
-    \end{tcolorbox}
-}
-
-\title{The Argennon Virtual Machine}
-\author{aybehrouz}
-\date{January 2021}
-
-% Document
-\begin{document}
-    \maketitle
-    \tableofcontents
-
-    \chapter{The Argennon Virtual Machine}\label{ch:AVM}
 
@@
 All these identifiers are \emph{prefix codes}, and hence can be represented by
-    \emph{prefix trees} \footnote{Also knows as tries.}. The AVM has three primitive prefix trees:
+\emph{prefix trees}\footnote{Also called a trie.}. The AVM has three primitive prefix trees:
 \emph{applications, accounts} and \emph{local}. Any identifier in Argennon is a
@@
 \section{The AVM Standard Library}\label{sec:the-avm-standard-library}
-
-    ...
-
-    \chapter{The Argon Language}\label{ch:argon-lang}
-
-    \section{Introduction}\label{sec:introduction2}
-
-    The Argon programming language is a class-based, object-oriented language designed for writing Argennon's smart
-    contracts. The Argon programming language is inspired by Solidity and is similar to Java, with a number of aspects
-    of them omitted and a few ideas from other languages included. Argon is designed to be fully compatible with
-    the Argennon Virtual Machine and be able to use all advanced features of the Argennon blockchain.
-
-    Argon Programs are organized as sets of packages. Each package has its own set of names for types, which helps
-    to prevent name conflicts. Every package can contain an arbitrary number of classes and a single contract. Every
-    Argon package corresponds to at max one AVM smart contract.
-
-    \section{Features Overview}\label{sec:features-overview}
-
-    \subsection{Static Classes}\label{subsec:static-classes}
-
-    In Argon a static class is an uninstantiable class which all its members and methods are static. Every static
-    class can define a special method \texttt{initialize} which can only be called in the \texttt{initialize}
-    method of the single contract of the package the static class is defined in.
-
-    A Static class is like a normal class which can only have one instance. So like normal classes, static classes can
-    be a super class or a subclass of other static classes and can implement interfaces.
-
-    \subsection{Contracts}\label{subsec:contract}
-
-    Every package of an Argon program can have one or zero contracts. A contact is a special static class which is
-    allowed to define methods with \texttt{external} visibility. When an Argon package is deployed as an AVM smart
-    contract, these external methods will define the contract's public interface.
-
-    \subsection{Access Level Modifiers}\label{subsec:access-level-modifiers}
-
-    Access level modifiers determine whether other classes can use a particular field or invoke a particular method or
-    if a method can be invoked externally by other smart contracts.
-
-    \begin{center}
-        \begin{tabular}{lllllr}
-            \hline
-                    & Class & Package & Subclass & Program & World \\
-            \hline
-            private   & yes   & no      & no       & no      & no    \\
-            protected & yes   & no      & yes      & no      & no    \\
-            package   & yes   & yes     & yes      & no      & no    \\
-            public    & yes   & yes     & yes      & yes     & no    \\
-            external  & no    & no      & no       & no      & yes   \\
-            \hline
-        \end{tabular}\label{tab:table}
-    \end{center}
-
-    \begin{lstlisting}[frame=TB, float, title=Argon contracts are static classes,label={lst:code1}]
-// A contract is a static class and all its methods are automatically static.
-contract MirrorToken {
-    private SimpleToken token;
-    private SimpleToken reflection;
-
-    // `initialize` is a special static method that is called by the AVM after the code of a contract
-    // is stored in the AVM code area. It can not be called after that.
-    initialize(double supply1, double supply2) {
-    // `new` does not create a new smart contract. It just makes an ordinary object.
-        token = new SimpleToken(supply1);
-        reflection = new SimpleToken(supply2);
-    }
-
-    external void transfer(account sender, account recipient,
-        double amount, signature sig) {
-        // Checks if the signature is issued by the sender and verifies the calling of this function
-        // with the current parameters.
-        sig.verifyCallBy(sender);
-        token.transfer(sender, recipient, amount);
-        reflection.transfer(recipient, sender, Math.sqrt(amount));
-    }
-
-    external double balanceOf(account user) { return token.balanceOf(user); }
-
-    external double balanceOfReflection(account user) { return reflection.balanceOf(user); }
-}
-
-public static class Math {
-    public double sqrt(double x) {
-        return // calculate square root of x
-    }
-}
-
-package class SimpleToken {
-    private map(account -> double) balances;
-
-    // The visibility of a member without an access modifier will be the same as its defining
-    // class so `SimpleToken` constructor has package visibility.
-    constructor(double initialSupply) {
-        // initializes the object
-    }
-
-    void transfer(account sender, account recipient, double amount) {
-        if (balances[sender] < amount) throw("Not enough balance.");
-        // implements the required logic...
-    }
-    // implements other methods...
-}
-    \end{lstlisting}
-
-    \subsection{Shadowing}\label{subsec:shadowing}
-
-    If a declaration of a type (such as a member variable or a parameter name) in a particular scope (such as an
-    inner block or a method definition) has the same name as another declaration in the enclosing scope, it will
-    result in a compiler error. In other words, the Argon programming language does not allow shadowing.
-
-    \chapter{The Argennon Blockchain}\label{ch:argennon-blockchain}
-
-    \section{Persistence}\label{sec:persistence}
-
-    For implementing the persistence layer of the AVM, we assume that we have access to an updatable zero-knowledge
-    elementary database (ZK-EDB) with the following properties:
-
-    \begin{itemize}
-        \item The ZK-EDB contains a mapping from a set of keys to a set of values.
-        \item Every state of the database has a commitment \(C\).
-        \item The ZK-EDB has a method \((D, p) = get(x)\), where \(x\) is a key and \(D\) is the associated data
-        with \(x\), and \(p\) is a proof.
-        \item A user can use \(C\) and \(p\) to verify that \(D\) is really associated with \(x\), and \(D\) is not
-        altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB\@.
-        \item Having \(p\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
-        associated with \(x\) instead of \(D\).
-    \end{itemize}
-
-    We use a ZK-EDB for storing the AVM heap. We include the commitment of the current state of this DB in every
-    block of the Argennon blockchain, so ZK-EDB servers need not be trusted servers.
-
-    Every page of AVM heap will be stored with a key of the form: \texttt{applicationID|pageIndex} (the \texttt{|}
-    operator concatenates two numbers). Nodes do not keep a full copy of the AVM heap and for validating block
-    certificates or emulating the AVM ( i.e.~validating transactions) they need to connect to a ZK-EDB and retrieve
-    the required pages of AVM heap. For better performance, nodes keep a cache of heap pages to
-    reduce the amount of ZK-EDB access.
-
-    We also use a ZK-EDB for storing the code area of each segment, and we include the commitment of this DB in every
-    block. Every code area will be divided into blocks and every block will be stored in the DB with
-    \texttt{applicationID|blockID} as its key. Like heap pages, nodes keep a cache of code area blocks.
-
-    \note{Unlike heap pages, the AVM is not aware of different blocks of code area.}
-
-    \section{Transactions}\label{sec:transactions}
-
-    Argennon has four types of transaction:
-
-    \begin{itemize}
-        \item \texttt{avmCall} essentially is an \texttt{invoke\_external} instruction that invokes a method from an
-        AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
-        assets, including ARGs, is done by these transactions.
-        \item \texttt{installApp} installs an AVM smart contract and determines the update policy of the smart
-        contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
-        on.
-        \item \texttt{unInstallApp} removes an AVM smart contract.
-        \item \texttt{updateApp} updates an AVM smart contract.
-    \end{itemize}
-
-    All types of Argennon transactions contain an \texttt{invoke\_external} instruction which calls a special method
-    from ARG smart contract that transfers the proposed fee of the transaction in ARGs from a sender account to the
-    fee sink accounts.
-
-    Every transaction is required to exactly specify what heap locations or code area addresses it will access. This
-    enables validators to start retrieving the required memory blocks from available ZK-EDB servers as soon as they
-    see a transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to
-    access a memory location that is not included in its access lists, will be rejected. Users could use smart contract
-    oracles to predict the list of memory blocks their transactions need. See Section~\ref{sec:smart-contract-oracle}
-    for more details.
-
-    \section{Blockchain}\label{sec:blockchain}
-
-    Every block of the Argennon blockchain corresponds to a set of transactions. We store the commitment of this
-    transaction set in every block, but we don't keep the set itself. To be able to detect replay attacks, we require
-    every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
-    and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
-    must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
-    a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
-    if a signature of the same user with a bigger or equal nonce is already used
-    (i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
-    nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
-
-    To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. When the
-    difference between \texttt{issuance} component of this nonce and the current round becomes bigger than the
-    maximum allowed lifetime of a signature, this information can be safely deleted. \textbf{As a result, we will not
-    have the problem of "un-removable empty accounts" like Ethereum.}
-
-    The only information that Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
-    blockchain. Every block of the Argennon blockchain contains the following information:
-
-    \begin{center}
-        \begin{tabular}{||c||}
-            \hline
-            Block \\ [0.5ex]
-            \hline\hline
-            commitment to the ZK-EDB storing heap pages \\ [0.7ex]
-            commitment to the ZK-EDB storing code areas \\ [0.7ex]
-            commitment to the set of transactions       \\ [0.7ex]
-            previous block hash                         \\ [0.7ex]
-            random seed                                 \\ [0.7ex]
-            \hline
-        \end{tabular}
-    \end{center}
-
-    For confirming a new block, nodes that are not validators only need to verify the block certificate. For
-    verifying a block certificate, a node needs to know the ARG balances of validators, but it doesn't need to
-    emulate the AVM execution.
-
-    On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
-    execution of the Argennon virtual machine. To do so, first they retrieve all heap pages and code area blocks they
-    need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
-    transactions included in the new block. This will modify some pages of the AVM memory, so they update the ZK-EDB
-    commitments based on the modified pages and verify the commitments included in the new block. Validators also
-    calculate and verify the commitment to the new block's transaction set.
-
-    \note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDB servers will receive the new
-    block, and they will update their database by emulating the AVM execution.}
-
-    \section{Consensus}\label{sec:consensus}
-
-    \note{It will be a randomized proof of stake which is under development....}
-
-    \subsection{Estimating A User's Stake}\label{subsec:estimating-a-user's-stake}
-
-    In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
-    of stake the user has in the system. Conventionally in these systems, for estimating a user's stake, we use the
-    amount of native system tokens the user is holding. Unfortunately, one problem with this approach is that a
-    strong attacker may be able to obtain a considerable amount of system tokens, for example by borrowing from a
-    DEFI application, and use this stake to attack the system.
-
-    To mitigate this problem, for calculating a user's stake at time step \(t\), instead of using the raw ARG
-    balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
-    ARG balance:
-    \[
-        S_{u,t} = \min (B_{u,t}, Trust_{u,t})
-    \]
-    Where:
-    \begin{itemize}
-        \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
-        \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
-        \item \(Trust_{u,t}\) is an estimated trust value for user \(u\) at time step \(t\).
-    \end{itemize}
-
-    The agreement protocol, at time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
-    number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
-    exponential moving average of the ARG balance of user \(u\) at time step \(t\).
-
-    In our system a user who held ARGs and participated in the consensus for a long time is more trusted
-    than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
-    amount of ARGs, also needs to hold them for a long period of time before being able to attack the system.
-
-    For calculating the exponential moving average of a user's balance at time step \(t\), we can use the following
-    recursive formula:
-    \[
-        M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
-    \]
-    Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
-    degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
-
-    Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
-    \(M_{u,t}\): (In the following equations the \(u\) subscript is dropped for simplicity)
-    \[
-        M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B
-    \]
-    Where:
-    \[
-        B = B_{k+1} = B_{k+2} = \dots = B_{t}
-    \]
-    We know that when \(|nx| \ll 1\) we can use the binomial approximation \({(1 + x)^n \approx 1 + nx}\). So, we can
-    further simplify this formula:
-    \[
-        M_{t} = M_{k} + (t - k) \alpha (B - M_{k})
-    \]
-
-    For choosing the value of \(\alpha\) we can consider the number of time steps that the trust value of a user needs
-    for reaching a specified fraction of his account balance. We know that for large \(n\) and \(|x| < 1\) we have
-    \((1 + x)^n \approx e^{nx}\), so by letting \(M_{u,k} = 0\) and \(n = t - k\) we can write:
-    \[
-        \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{B}\right)}{n}
-    \]
-    The value of \(\alpha\) for a desired configuration can be calculated by this equation. For instance, we could
-    calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8B\) and \(n\) equals to the
-    number of time steps of 10 years.
-
-    In our system a newly created account will not have voting power for some time, no matter how high its
-    balance is. While this is a desirable property, in case a large proportion of total system tokens are
-    transferred to newly created accounts, it can result in too much voting power for older accounts. This may decrease
-    the degree of decentralization in our system.
-
-    However, this situation is easily detectable by comparing the total stake of the system with the total balance of
-    users. If after confirming a block the total stake of the system goes too low and we have:
-    \[
-        \sum_{u}S_{u,t} < \gamma \sum_{u}B_{u,t}
-    \]
-    The protocol will perform a \emph{time shift} in the system: the time step of the system
-    will be incremented for \(m\) steps while no blocks will be confirmed. This will increase the value of \(M_{u,t}\)
-    for new accounts with a non-zero balance, giving them more influence in the agreement protocol.
-
-    For calculating the value of \(m\) which determines the amount of time shift in the system, we should note that when
-    \(B_{u,t} = B_{u, t-1} = B_u\), we can derive a simple recursive rule for the stake of a user:
-    \[
-        S_{u,t} = (1 - \alpha) S_{u,t-1} + \alpha B_u
-    \]
-    Therefore, we have:
-    \[
-        \sum_{u}S_{u,t} = (1 - \alpha) \sum_{u}S_{u,t - 1} + \alpha \sum_{u}B_u
-    \]
-    This equation shows that when the balance of users is not changing over time the total stake of the system is the
-    exponential average of the total ARGs of the system. Consequently, when we shift the time for \(m\) steps, we can
-    calculate the new total stake of the system from the following equation:
-
-    \[
-        \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
-    \]
-    Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
-    we can obtain \(m\) from the following formula, assuming \(\alpha\) is small enough:
-    \[
-        m = \frac{1}{\alpha} \ln \left(\frac{1 - \gamma}{1 - \lambda}\right)
-    \]
-
-    \section{Incentive mechanism}\label{sec:incentive-mechanism}
-
-    \subsection{Transaction Fee}\label{subsec:transaction-fee}
-
-    Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
-    special method from ARG smart contract. This method will transfer the proposed fee of the transaction in ARGs
-    from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
-    execution fees and \texttt{dbFeeSink} collects fees for ZK-EDB servers. The Protocol decides how to distribute the
-    transaction fee between these two fee sink accounts.
-
-    When a block is added to the blockchain, the proposer of that block will receive a share of the block fees.
-    Consequently, a block proposer is always incentivized to include more transactions in his block. However, if he
-    puts too many transactions in his block and the validation of the block becomes too difficult, some validators
-    may not be able to validate all transactions on time. If a validator can not validate a block in the required
-    time, he will consider the block invalid. So, when a proposed block contains too many transactions, the network
-    may reach consensus on another block, and the proposer of that block will not receive any fees. As a result, a
-    proposer is incentivized to use network transaction capacity optimally.
-
-    On the other hand, we believe that the proposer does not have enough incentives for optimizing the storage size
-    of the transaction set. Therefore, we require that \textbf{the size of the transaction set of every block in
-    bytes be lower than a certain threshold.}
-
-    Validators need to spend resources for validating transactions. When a validator starts the emulation of the AVM
-    to validate a transaction, solely from the code he can't predict the time the execution will finish. This will
-    give an adversary an opportunity to attack the network by broadcasting transactions that never ends. Since,
-    validators can not finish the execution of these transactions, the network will not be able to charge the
-    attacker any fees, and he would be able to waste validators resources for free.
-
-    To mitigate this problem, we require that every transaction specify a cap for all the resources it needs. This
-    will include memory, network and processor related resources. Also, the protocol defines an execution cost for
-    every AVM instruction reflecting the amount of resources its emulation needs. This will define a standard way for
-    measuring the execution cost of any \texttt{avmCall} transaction. Every \texttt{avmCall} transaction is required
-    to specify a maximum execution cost. If during emulation it reaches this maximum cost, the transaction will be
-    considered failed and the network can receive the proposed fee of that transaction.
-
-    Every \texttt{avmCall} transaction is required to provide the following information as an upper bound for the
-    resources it needs:
-
-    \begin{itemize}
-        \item Execution cost
-        \item A list of heap/code-area locations for reading
-        \item A list of heap locations for writing
-        \item A list of heap pages it will deallocate
-        \item Number and size of heap pages it will allocate
-    \end{itemize}
-
-    If a transaction tries to violate any of these predefined limitations, for example, if it tries to read a memory
-    location that is not included in its reading list, it will be considered failed and the network can receive the
-    proposed fee of that transaction.
-
-    \note{A transaction always pays all of its proposed fee, no matter how much of its predefined resources were not
-    used in the final emulation.}
-
-    \subsection{Incentives for ZK-EDB Servers}\label{subsec:zk-edb-servers}
-
-    The incentive mechanism for ZK-EDB servers should have the following properties:
-
-    \begin{itemize}
-        \item It incentivizes storing all memory blocks, whether a heap page or a code area block, and not only those
-        which are used more frequently.
-        \item It incentivizes ZK-EDB servers to actively provide the required memory blocks for validators.
-        \item Making more accounts will not provide any advantages for a ZK-EDB server.
-    \end{itemize}
-
-    For our incentive mechanism, we require that every time a validator receives a memory block from a ZK-EDB, after
-    validating the data, he give a receipt to the ZK-EDB. In this receipt the validator signs the following information:
-
-    \begin{itemize}
-        \item \texttt{ownerAddr}: the ARG address of the ZK-EDB\@.
-        \item \texttt{receivedBlockID}: the ID of the received memory block.
-        \item \texttt{round}: the current round number.
-    \end{itemize}
-
-    \note{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}
-
-    To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ARGs from
-    \texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
-    between all \emph{winning tickets} of the lottery.
-
-    \note{One ZK-EDB server could own multiple winning tickets in a round.}
-
-    To run this lottery, every round, based on the current block seed, a collection of \emph{valid} receipts will be
-    selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in round \(r\) if:
-
-    \begin{itemize}
-        \item The signer was a validator in the round \(r - 1\) and voted for the agreed-upon block.
-        \item The data block in the receipt was needed for validating the \textbf{previous} block.
-        \item The receipt round number is \(r - 1\).
-        \item The signer did not sign a receipt for the same data block for two different ZK-EDBs in the previous round.
-    \end{itemize}
-    For selecting the winning receipts we could use a random generator:
-    \begin{verbatim}
-IF random(seed|validatorPK|receivedBlockID) < winProbability THEN
-    the receipt issued by validatorPK for receivedBlockID is a winner
-    \end{verbatim}
-    \begin{itemize}
-        \item \texttt{random()} produces uniform random numbers between 0 and 1, using its input argument as a seed.
-        \item \texttt{validatorPK} is the public key of the signer of the receipt.
-        \item \texttt{receivedBlockID} is the ID of the memory block that the receipt was issued for.
-        \item \texttt{winProbability} is the probability of winning in every round.
-        \item \texttt{seed} is the current block seed.
-        \item \texttt{|} is a concatenation operator.
-    \end{itemize}
-
-    \note{The winners of the lottery were validators one round before the lottery round.}
-
-    Also, based on the current block seed, a random memory block, whether a heap page or a code area block, is
-    selected as the challenge of the round. A ZK-EDB that owns a winning receipt needs to broadcast a \emph{winning
-    ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
-    challenge. Solving a round challenge requires the content of the memory block which was selected as the round
-    challenge. This will encourage ZK-EDBs to store all memory blocks.
-
-    A possible choice for the challenge solution could be the cryptographic hash of the content of the challenge
-    memory block combined with the ZK-EDB ARG address: \texttt{hash(challenge.content|ownerAddr)}
-
-    The winning tickets of the lottery of round \(r\) need to be included in the block of the round \(r\),
-    otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
-    \(r\) will be done in the round \(r + 1\). This way, \textbf{the content of the challenge memory block could be
-    kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.
-
-    \subsection{Memory Allocation and De-allocation Fee}\label{subsec:memory-allocation-and-de-allocation}
-
-    Every \(k\) round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
-    allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
-    address of the smart contract.
-
-    To determine the price of AVM memory, Every \(k\) round, the protocol calculates \texttt{dbFee} and
-    \texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
-    \texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
-    the protocol considers the total size of all the memory pages that were accessed for either reading or writing
-    during a time period. These two values will be calculated for the last \(k\) rounds and the price per byte of
-    AVM memory will be a linear function of \texttt{dbFee/memTraffic}
-
-    When a smart contract executes a heap de-allocation instruction, the protocol will refund the cost of
-    de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
-    calculates the refunded amount based on the average price the smart contract had paid for that allocated memory.
-    This will prevent smart contracts from profit taking by trading memory with the protocol.
-
-    \section{Concurrency}\label{sec:concurrency}
-
-    \subsection{Memory Dependency Graph}\label{subsec:memory-dependency-graph}
-
-    Every block of the Argennon blockchain contains a list of transactions. This list is an ordered list and the
-    effect of its contained transactions must be applied to the AVM state sequentially as they appear in the ordered
-    list. This ordering is solely chosen by the block proposer, and users should not have any assumptions about
-    the ordering of transactions in a block.
-
-    The fact that block transactions constitute a sequential list, does not mean they can not be executed and applied
-    to the AVM state concurrently. Many transactions are actually independent and the order of their execution does not
-    matter. These transactions can be safely validated in parallel by validators.
-
-    A transaction can change the AVM state by modifying either the code area or the AVM heap. In Argennon, all
-    transactions declare the list of memory locations they want to read or write. This will enable us to determine the
-    independent sets of transactions which can be executed in parallel. To do so, we define the \emph{memory dependency
-    graph} \(G_d\) as follows:
-
-    \begin{itemize}
-        \item \(G_d\) is an undirected graph.
-        \item Every vertex in \(G_d\) corresponds to a transaction and vice versa.
-        \item Vertices \(u\) and \(v\) are adjacent in \(G_d\) if and only if \(u\) has a memory location \(L\) in its
-        writing list and \(v\) has \(L\) in either its writing list or its reading list.
-    \end{itemize}
-
-    If we consider a proper vertex coloring of \(G_d\), every color class will give us an independent set of
-    transactions which can be executed concurrently. To achieve the highest parallelization, we need to color \(G_d\)
-    with minimum number of colors. Thus, the \emph{chromatic number} of the memory dependency graph shows how good a
-    transaction set could be run concurrently.
-
-    Graph coloring is computationally NP-hard. However, in our use case we don't need to necessarily find an optimal
-    solution. An approximate greedy algorithm will perform well enough in most circumstances.
-
-    After constructing the memory dependency graph, we can use it to construct the
-    \emph{execution DAG} of transactions. The execution DAG of transaction set \(T\) is a directed acyclic
-    graph \(G_e = (V_e,E_e)\) which has the \emph{execution invariance} property:
-    \begin{itemize}
-        \item Every vertex in \(V_e\) corresponds to a transaction in \(T\) and vice versa.
-        \item Executing the transactions of \(T\) in any order that \emph{respects} \(G_e\) will result in
-        the same AVM state.
-        \begin{itemize}
-            \item An ordering of transactions of \(T\) respects \(G_e\) if for every directed edge \((u,v) \in E_e\)
-            the transaction \(u\) comes before the transaction \(v\) in the ordering.
-        \end{itemize}
-    \end{itemize}
-
-    Having the execution DAG of a set of transactions, using Algorithm~\ref{alg:exec_dag}, we can apply the transaction
-    set to the AVM state concurrently, using multiple processor, while we can be sure that the resulted AVM state will
-    always be the same no matter how many processor we have used.
-
-    %##\includegraphics[width=17cm]{../img/Alg1s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Ready}{$R_e$}\SetKwData{V}{$v_f$}\SetKwData{Graph}{$G_e$}\SetKwData{Vertices}{$V$}\SetKwData
-        {Txns}{$T$}
-        \KwData{The execution dag $\Graph = (\Vertices,E)$ of transaction set \Txns}
-        \KwResult{The state of the AVM after applying \Txns with any ordering respecting \Graph}
-        \BlankLine
-        \Ready $\gets$ the set of all vertices of \Vertices with in degree 0\;
-        \While{$\Vertices \neq \varnothing$}
-        {
-            wait until a new free processor is available\;
-            \If{the execution of a transaction was finished}
-            {
-                remove the vertex of the finished transaction \V from \Graph\;
-                \For{each vertex $u \in Adj[\V]$}
-                {
-                    \If{$u$ has zero in degree}
-                    {
-                        $\Ready \gets \Ready \cup u$\;
-                    }
-                }
-            }
-            \If{$\Ready \neq \varnothing$}
-            {
-                remove a vertex from \Ready and assign it to a processor\;
-            }
-        }
-        \caption{executing DAG transactions}\label{alg:exec_dag}
-    \end{algorithm}
-
-    By replacing every undirected edge of a memory dependency graph with a directed edge in such a way that the
-    resulted graph has no cycles, we will obtain a valid execution DAG. Thus, from a memory dependency graph different
-    execution DAGs can be constructed with different levels of parallelization ability.
-
-    If we assume that we have unlimited number of processors and all transactions take equal time for executing, it
-    can be shown that by providing a minimal graph coloring to Algorithm~\ref{alg:gen_dag} as input, the resulted
-    DAG will be optimal, in the sense that it results in the minimum overall execution time.
-
-    %##\includegraphics[width=17cm]{../img/Alg2s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Txns}{$T$}\SetKwData{Gd}{$G_d=(V_d,E_d)$}
-        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{The memory dependency graph \Gd of transaction set \Txns\\A proper coloring of $G_d$}
-        \Output{An execution dag $G_e=(V_e,E_e)$ for the transaction set \Txns}
-        \BlankLine
-        $V_e \gets V_d$\;
-        $E_e \gets \varnothing$\;
-        define a total order on colors of $G_d$\;
-        \For{each edge $\{u,v\} \in E_d$}
-        {
-            \eIf{$color[u] < color[v]$}
-            {
-                $E_e \gets E_e \cup (u,v)$\;
-            }{
-                $E_e \gets E_e \cup (v,u)$\;
-            }
-        }
-        \caption{Constructing an execution DAG}\label{alg:gen_dag}
-    \end{algorithm}
-
-    The block proposer is responsible for proposing an efficient execution DAG alongside his proposed block. This
-    execution DAG will determine the ordering of block transactions and help validators to validate transactions in
-    parallel. Since with better parallelization a block can contain more transactions, a proposer is incentivized enough
-    to find a good execution DAG for transactions.
-
-    \subsection{Memory Spooling}\label{subsec:spooling}
-
-    When two transactions are dependant and they are connected with an edge \((u,v)\) in the execution DAG,
-    the transaction \(u\) needs to be run before the transaction \(v\). However, if \(v\) does not read any
-    memory locations that \(u\) modifies, we can run \(u\) and \(v\) in parallel. We just need to make sure
-    \(u\) does not see any changes \(v\) is making in AVM memory. This can be done by appropriate versioning
-    of the memory locations which is shared between \(u\) and \(v\). We call this method \emph{memory spooling}.
-    After enabling memory spooling between two transactions the edge connecting them can be safely removed from the
-    execution DAG\@.
-
-    \subsection{Concurrent Counters}\label{subsec:concurrent-counters}
-
-    We know that in Argennon every transaction needs to transfer its proposed fee to the \texttt{feeSink} accounts
-    first. This essentially makes every transaction a reader and a writer of the memory locations which store the
-    balance record of the \texttt{feeSink} accounts. As a result, all transactions in Argennon will be dependant and
-    parallelism will be completely impossible. Actually, any account that is highly active, for example the account
-    of an exchange or a payment processor, could become a concurrency bottleneck in our system which makes all
-    transactions interacting with them dependant.
-
-    This problem can be easily solved by using a concurrent counter for storing the balance record of this type of
-    accounts. A concurrent counter is a data structure which improves concurrency by using multiple memory locations for
-    storing a single counter. The value of the concurrent counter is equal to the sum of its sub counters and it can
-    be incremented or decremented by incrementing/decrementing any of the sub counters. This way, a concurrent
-    counter trades concurrency with memory usage.
-
-    Algorithm~\ref{alg:CC} implements a concurrent counter which returns an error when the value of the counter
-    becomes negative.
-
-    %##\includegraphics[width=17cm]{../img/Alg3s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{CC}{Counter}
-        \SetKwFunction{Inc}{Increment}\SetKwFunction{Dec}{Decrement}\SetKwFunction{AtomInc}{AtomicIncrement}
-        \SetKwFunction{AtomDec}{AtomicDecrement}\SetKwFunction{AtomSet}{AtomicSet}\SetKwFunction{Get}{GetValue}
-        \SetKwFunction{Acquire}{Lock.Acquire}\SetKwFunction{Release}{Lock.Release}
-        \SetKwProg{Fn}{Function}{}{}
-        \Fn{\Get{\CC}}
-        {
-            $s \gets 0$\;
-            \Acquire{}\;
-            \For{$i \gets 0$ \KwTo $\CC.size - 1$}
-            {
-                $s \gets s + \CC.cell[i]$\;
-            }
-            \Release{}\;
-            \KwRet{s}\;
-        }
-        \BlankLine
-        \Fn{\Inc{\CC, value, seed}}
-        {
-            $i \gets seed \bmod \CC.size$\;
-            \AtomInc{$\CC.cell[i]$, value}\;
-        }
-        \BlankLine
-        \Fn{\Dec{\CC, value, seed, attempt}}
-        {
-            \If {attempt = \CC.size}
-            {
-                restore \CC by adding back the subtracted value\;
-                \KwRet{Error}\;
-            }
-            $i \gets seed \bmod \CC.size$\;
-            $i \gets (i + attempt) \bmod \CC.size$\;
-            \eIf {$\CC.cell[i] \geq value$}
-            {
-                \AtomDec{$\CC.cell[i]$, value}\;
-            }{
-                $r \gets value - \CC.cell[i]$\;
-                \AtomSet{$\CC.cell[i]$, $0$}\;
-                \Dec{\CC, r, seed, $attempt + 1$}\;
-            }
-        }
-        \caption{Concurrent counter}\label{alg:CC}
-    \end{algorithm}
-
-    It should be noted that in a blockchain application we don't have concurrent threads and therefore we don't need
-    atomic functions. For usage in a smart contract, the atomic functions of this pseudocode can be implemented like
-    normal functions.
-
-    Concurrent counter data structure is a part of the AVM standard library, and any smart contract can use this data
-    structure for storing the balance record of highly active accounts.
-
-    \subsection{Memory Chunks}\label{subsec:memory-chunks}
-
-    In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{chunks}.
-    Each memory chunk can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
-    consensus on new values of the commitment of any chunk can be achieved by a different voting committee.
-
-    If a transaction does not modify a memory chunk and in the transaction ordering of the block it comes after
-    any transaction which modifies that chunk, then the execution of that transaction is not needed for calculating
-    the new commitment of the chunk. Consequently, the voting committee of that memory chunk can safely ignore such a
-    transaction. The execution DAG of transactions can be used for finding and pruning these transactions as
-    we see in Algorithm~\ref{alg:prune_dag}.
-
-    %##\includegraphics[width=17cm]{../img/Alg4s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{V}{$v_0$}\SetKwData{Graph}{$G_e$}\SetKwData{Chunk}{$C_k$}\SetKwData{Txns}{$T$}
-        \SetKwFunction{RPrune}{RecursivePrune}
-        \SetKwProg{Fn}{Function}{}{}
-        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{An execution dag \Graph and a memory chunk \Chunk}
-        \Output{The pruned execution dag based on \Chunk}
-        \BlankLine
-        \For{each vertex \V with out degree $0$}
-        {
-            \RPrune{\V}\;
-        }
-        \BlankLine
-        \Fn{\RPrune{\V}}
-        {
-            \If{$Adj[\V] = \varnothing$ {\bf and} \V does not modify \Chunk}
-            {
-                remove \V from \Graph\;
-                \For{each $u$ such that edge $(u,\V)$ was in \Graph}
-                {
-                    \RPrune{u}\;
-                }
-            }
-        }
-        \caption{Pruning an execution DAG}\label{alg:prune_dag}
-    \end{algorithm}
-
-    If we choose chunks in a way that most transactions only modify memory locations of one chunk,
-    likely many transactions of a block only need to be validated by one voting committee and can be validated in
-    parallel by different committees.
-
-    Because the voting committees are selected by random sampling, by choosing large enough samples we can make sure
-    that having multiple voting committees will not change the security properties of the Argennon agreement protocol.
-
-    \section{Smart Contract Oracle}\label{sec:smart-contract-oracle}
-
-    A smart contract oracle is a full AVM emulator that keeps a full local copy of AVM memory and can emulate AVM
-    execution without accessing a ZK-EDB. Smart contract oracles can be used for reporting useful information about
-    \texttt{avmCall} transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
-    and so on.
-
-\end{document}

### 2021-04-26 tex/sections/blockchain.tex
Change project structure
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-03-29 tex/avm.tex
@@
 %! Author = aybehrouz
-%! Date = 2/28/21
-
-% Preamble
-\documentclass[a4paper]{report}
-
-% Packages
-%\usepackage{amsmath}
-%\usepackage[a4paper]{geometry}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{amsfonts}
-\usepackage[ruled,boxed]{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{tcolorbox}
-\lstset{
-%  basicstyle=\ttfamily,
-    mathescape
-}
-%\binoppenalty=99999
-%\relpenalty=99999
-\newcommand{\note}[1] {
-    \begin{tcolorbox}[colframe=white,colback=white]
-        \emph{#1}
-        %\emph{#1}\medskip
-    \end{tcolorbox}
-}
-
-\title{The Argennon Virtual Machine}
-\author{aybehrouz}
-\date{January 2021}
-
-% Document
-\begin{document}
-    \maketitle
-    \tableofcontents
-
-    \chapter{Specification}\label{ch:specification}
-
-    \section{Introduction}\label{sec:introduction}
-
-    The Argennon Virtual Machine (AVM) is an abstract computing machine for executing Argennon's smart contracts. It
-    is designed in a way that it could be efficiently implemented in either hardware or software.
-
-    \section{The Structure of the Argennon Virtual Machine}\label{sec:the-structure-of-the-algorand-virtual-machine}
-
-    ...
-
-    \subsection{Data Types}\label{subsec:data-types}
-
-    The Argennon Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
-    and does not have to be done by the Argennon Virtual Machine itself.
-
-    The instruction set of the Argennon Virtual Machine distinguishes its operand types using instructions intended to
-    operate on values of specific types. For instance, \texttt{iadd} assumes that its operands are two 64-bit integers.
-
-    \subsection{The \texttt{PC} Register}\label{subsec:the-pc-register}
-
-    ...
-
-    \subsection{Call Stack}\label{subsec:call-stack}
-
-    A call stack contains the information that is needed for restoring the state of the invoker of a method.
-
-    \subsection{Memory Unit}\label{subsec:memory-unit}
-
-    The Argennon Virtual Machine has a byte addressable memory which is divided into separate segments. Every segment
-    belongs to a smart contract that has a unique \texttt{applicationID}. The AVM always has a single working memory
-    segment and memory locations outside its current working segment can not be accessed. The only instructions which
-    can change the current working segment are \texttt{invoke\_external}, \texttt{athrow} and return instructions.
-
-    Every smart contract has its own memory segment. Hence, there is no way for a smart contract to access another
-    smart contract's memory. Interaction between smart contracts is done using \texttt{invoke\_external} instruction,
-    and a smart contract can invoke methods of another smart contract by this instruction.
-
-    A memory segment consists of five data areas:
-    \begin{itemize}
-        \item Code Area
-        \item Constant Area
-        \item Local Frame
-        \item Operand Stack
-        \item Heap
-    \end{itemize}
-    All data areas except operand stack, have their own address space which starts from 0. Operand stack is a
-    last-in-first-out (LIFO) stack and is not addressable. Every memory access instruction operates on its specific
-    data areas.
-
-    \subsubsection{Code Area}
-
-    The code area of a segment contains the byte-code of the smart contract which owns that segment. The AVM has no
-    instructions for manipulating the code area. \textbf{Installing, removing and updating smart contracts need to
-    be done externally.}
-
-    \subsubsection{Constant Area}
-
-    The constant area of a segment contains several kinds of constants, ranging from user defined constants to method
-    address tables. A method address table stores method locations in the code area and their access type. The access
-    type of a method can be either \texttt{public} or \texttt{private}. The AVM has no instructions for modifying the
-    constant area.
-
-    \note{Only public methods can be invoked by \texttt{invoke\_external} instruction.}
-
-    \subsubsection{Local Frame}
-
-    A local frame is used to store methods parameters and local variables. A new frame is created each time a method
-    is invoked, and it is destroyed when its method invocation completes, whether the completion is normal or abrupt.
-
-    \subsubsection{Operand Stack}
-
-    Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too. AVM
-    instructions take operands from the operand stack, operate on them, and push the result back onto the operand
-    stack. An operand stack is destroyed when its owner method completes, whether that completion is normal or abrupt.
-
-    \subsubsection{Heap}
-
-    The heap of a segment is a persistent memory area which is divided into pages. Memory locations inside every page
-    have a separate address space that starts from 0, and each page can be referenced by an index that starts from 0. In
-    other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex, offset)}.
-    Different pages of a heap do not need to be equally sized.
-
-    \note{The reason behind this paged design is that the heap is usually persisted using a block device. A heap with
-    a paged structure could expose the underlying block based nature of the persistence layer to the application
-    layer. In this way, the compiler or the programmer could better optimize the code for the persistence layer.}
-
-    \section{Instruction Set Summary}\label{sec:instruction-set-summary}
-
-    An Argennon Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
-    performed, followed by zero or more operands supplying arguments or data that are used by the operation.
-    \textbf{The number and size of the operands are determined solely by the opcode.}
-
-    \subsection{Method Invocation}\label{subsec:method-invocation}
-
-    The Argennon Virtual Machine has three types of method invocation:
-    \begin{itemize}
-        \item \texttt{invoke\_internal} invokes a method from the current running smart contract.
-        \item \texttt{invoke\_external} invokes a \texttt{public} method from another smart contract. It will change
-        the current memory segment to the segment of the invoked smart contract.
-        \item \texttt{invoke\_native} invokes a method that is not hosted by the Argennon virtual machine. By this
-        instruction, high performance native methods of the hosting machine could become available to AVM smart
-        contracts.
-    \end{itemize}
-
-    \note{In the future, we may need to add special instructions for invoking interface and virtual methods...}
-
-    Each time a method is invoked a new local frame and operand stack is created. The Argennon Virtual Machine uses
-    local frames to pass parameters on method invocation. On method invocation, any parameters are passed in
-    consecutive local variables stored in the method's local frame starting from local variable 0. The invoker of a
-    method writes the parameters in the local frame of the invoked method using \texttt{arg} instructions.
-
-    \subsubsection{Exceptions}
-
-    An exception is thrown programmatically using the \texttt{athrow} instruction. Exceptions can also be thrown by
-    various Argennon Virtual Machine instructions if they detect an abnormal condition. Some exceptions are not
-    catchable and will always abort the execution of the smart contract.
-
-    \subsubsection{Method Invocation Completion}
-
-    A method invocation completes normally if that invocation does not cause an exception to be thrown, either
-    directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
-    method completes normally, then a value may be returned to the invoking method. This occurs when the invoked
-    method executes one of the return instructions, the choice of which must be appropriate for the type of the value
-    being returned (if any). Execution then continues normally in the invoking method's local frame with the returned
-    value (if any) pushed onto the operand stack.
-
-    A method invocation completes abruptly if an exceptions is thrown and is not caught by the current method. A
-    method invocation that completes abruptly never returns a value to its invoker.
-
-    When a method completes, whether normally or abruptly, the call stack is used to restore the state of the invoker,
-    including its local frame and operand stack, with the \texttt{PC} register appropriately restored and incremented
-    to skip past the method invocation instruction. If the invoker was another smart contract, i.e.\ the invocation
-    was made by an \texttt{invoke\_external} instruction, the current memory segment will be changed to the invoker's
-    segment.
-
-    A thrown exception causes methods in the call stack to complete \textbf{abruptly} one by one, as long as the
-    \texttt{PC} register is not pointing to a \texttt{catch} instruction. The \texttt{catch} instruction acts like a
-    branch instruction that branches only if an exception is caught. \textbf{When an external method invocation
-    completes abruptly, before changing the current segment, all changes made to the heap area by that method
-    call, including changes made to other segments, will be rolled back.} So, when a method call completes abruptly,
-    that method call essentially has no effect on the AVM state.
-
-    \note{By using the \texttt{athrow} instruction properly, a programmer can make any method act like an atomic
-    operation.}
-
-    \subsection{Authorizing Operations}\label{subsec:authorizing-operations}
-
-    In blockchain applications, we usually need to authorize certain operations. For example, for sending an asset
-    from a user to another user, first we need to make sure that the sender has authorized this operation. The
-    Argennon virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
-    cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
-    passing signatures as parameters to methods, a programmer can implement the required logic for authorizing any
-    operation.
-
-    \note{The Argennon virtual machine has no instructions for issuing cryptographic signatures.}
-
-    In addition to signatures, a method can verify its invoker by using \texttt{get\_parent} instruction. This
-    instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
-    smart contract in the call stack. In other words, it returns the \texttt{applicationID} of the smart contract that
-    has invoked the current smart contract. (if any)
-
-    \subsection{Heap Allocation Instructions}\label{subsec:heap-allocation-instructions}
-
-    ...
-
-    \section{The AVM Standard Library}\label{sec:the-avm-standard-library}
-
-    ...
-
-    \chapter{Implementation}\label{ch:implementation}
 
 \section{Persistence}\label{sec:persistence}
-
-    For implementing the persistence layer of the AVM, we assume that we have access to an updatable zero-knowledge
-    elementary database (ZK-EDB) with the following properties:
-
-    \begin{itemize}
-        \item The ZK-EDB contains a mapping from a set of keys to a set of values.
-        \item Every state of the database has a commitment \(C\).
-        \item The ZK-EDB has a method \((D, p) = get(x)\), where \(x\) is a key and \(D\) is the associated data
-        with \(x\), and \(p\) is a proof.
-        \item A user can use \(C\) and \(p\) to verify that \(D\) is really associated with \(x\), and \(D\) is not
-        altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB\@.
-        \item Having \(p\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
-        associated with \(x\) instead of \(D\).
-    \end{itemize}
-
-    We use a ZK-EDB for storing the AVM heap. We include the commitment of the current state of this DB in every
-    block of the Argennon blockchain, so ZK-EDB servers need not be trusted servers.
-
-    Every page of AVM heap will be stored with a key of the form: \texttt{applicationID|pageIndex} (the \texttt{|}
-    operator concatenates two numbers). Nodes do not keep a full copy of the AVM heap and for validating block
-    certificates or emulating the AVM ( i.e.~validating transactions) they need to connect to a ZK-EDB and retrieve
-    the required pages of AVM heap. For better performance, nodes keep a cache of heap pages to
-    reduce the amount of ZK-EDB access.
-
-    We also use a ZK-EDB for storing the code area of each segment, and we include the commitment of this DB in every
-    block. Every code area will be divided into blocks and every block will be stored in the DB with
-    \texttt{applicationID|blockID} as its key. Like heap pages, nodes keep a cache of code area blocks.
-
-    \note{Unlike heap pages, the AVM is not aware of different blocks of code area.}
+\input{sections/persistence}
 
@@
 \section{Consensus}\label{sec:consensus}
-
-    \note{It will be a randomized proof of stake which is under development....}
-
-    \subsection{Estimating A User's Stake}\label{subsec:estimating-a-user's-stake}
-
-    In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
-    of stake the user has in the system. Conventionally in these systems, for estimating a user's stake, we use the
-    amount of native system tokens the user is holding. Unfortunately, one problem with this approach is that a
-    strong attacker may be able to obtain a considerable amount of system tokens, for example by borrowing from a
-    DEFI application, and use this stake to attack the system.
-
-    To mitigate this problem, for calculating a user's stake at time step \(t\), instead of using the raw ARG
-    balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
-    ARG balance:
-    \[
-        S_{u,t} = \min (B_{u,t}, Trust_{u,t})
-    \]
-    Where:
-    \begin{itemize}
-        \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
-        \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
-        \item \(Trust_{u,t}\) is an estimated trust value for user \(u\) at time step \(t\).
-    \end{itemize}
-
-    The agreement protocol, at time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
-    number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
-    exponential moving average of the ARG balance of user \(u\) at time step \(t\).
-
-    In our system a user who held ARGs and participated in the consensus for a long time is more trusted
-    than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
-    amount of ARGs, also needs to hold them for a long period of time before being able to attack the system.
-
-    For calculating the exponential moving average of a user's balance at time step \(t\), we can use the following
-    recursive formula:
-    \[
-        M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
-    \]
-    Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
-    degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
-
-    Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
-    \(M_{u,t}\): (In the following equations the \(u\) subscript is dropped for simplicity)
-    \[
-        M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B
-    \]
-    Where:
-    \[
-        B = B_{k+1} = B_{k+2} = \dots = B_{t}
-    \]
-    We know that when \(|nx| \ll 1\) we can use the binomial approximation \({(1 + x)^n \approx 1 + nx}\). So, we can
-    further simplify this formula:
-    \[
-        M_{t} = M_{k} + (t - k) \alpha (B - M_{k})
-    \]
-
-    For choosing the value of \(\alpha\) we can consider the number of time steps that the trust value of a user needs
-    for reaching a specified fraction of his account balance. We know that for large \(n\) and \(|x| < 1\) we have
-    \((1 + x)^n \approx e^{nx}\), so by letting \(M_{u,k} = 0\) and \(n = t - k\) we can write:
-    \[
-        \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{B}\right)}{n}
-    \]
-    The value of \(\alpha\) for a desired configuration can be calculated by this equation. For instance, we could
-    calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8B\) and \(n\) equals to the
-    number of time steps of 10 years.
-
-    In our system a newly created account will not have voting power for some time, no matter how high its
-    balance is. While this is a desirable property, in case a large proportion of total system tokens are
-    transferred to newly created accounts, it can result in too much voting power for older accounts. This may decrease
-    the degree of decentralization in our system.
-
-    However, this situation is easily detectable by comparing the total stake of the system with the total balance of
-    users. If after confirming a block the total stake of the system goes too low and we have:
-    \[
-        \sum_{u}S_{u,t} < \gamma \sum_{u}B_{u,t}
-    \]
-    The protocol will perform a \emph{time shift} in the system: the time step of the system
-    will be incremented for \(m\) steps while no blocks will be confirmed. This will increase the value of \(M_{u,t}\)
-    for new accounts with a non-zero balance, giving them more influence in the agreement protocol.
-
-    For calculating the value of \(m\) which determines the amount of time shift in the system, we should note that when
-    \(B_{u,t} = B_{u, t-1} = B_u\), we can derive a simple recursive rule for the stake of a user:
-    \[
-        S_{u,t} = (1 - \alpha) S_{u,t-1} + \alpha B_u
-    \]
-    Therefore, we have:
-    \[
-        \sum_{u}S_{u,t} = (1 - \alpha) \sum_{u}S_{u,t - 1} + \alpha \sum_{u}B_u
-    \]
-    This equation shows that when the balance of users is not changing over time the total stake of the system is the
-    exponential average of the total ARGs of the system. Consequently, when we shift the time for \(m\) steps, we can
-    calculate the new total stake of the system from the following equation:
-
-    \[
-        \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
-    \]
-    Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
-    we can obtain \(m\) from the following formula, assuming \(\alpha\) is small enough:
-    \[
-        m = \frac{1}{\alpha} \ln \left(\frac{1 - \gamma}{1 - \lambda}\right)
-    \]
+\input{sections/consensus}
 
@@
 \section{Concurrency}\label{sec:concurrency}
-
-    \subsection{Memory Dependency Graph}\label{subsec:memory-dependency-graph}
-
-    Every block of the Argennon blockchain contains a list of transactions. This list is an ordered list and the
-    effect of its contained transactions must be applied to the AVM state sequentially as they appear in the ordered
-    list. This ordering is solely chosen by the block proposer, and users should not have any assumptions about
-    the ordering of transactions in a block.
-
-    The fact that block transactions constitute a sequential list, does not mean they can not be executed and applied
-    to the AVM state concurrently. Many transactions are actually independent and the order of their execution does not
-    matter. These transactions can be safely validated in parallel by validators.
-
-    A transaction can change the AVM state by modifying either the code area or the AVM heap. In Argennon, all
-    transactions declare the list of memory locations they want to read or write. This will enable us to determine the
-    independent sets of transactions which can be executed in parallel. To do so, we define the \emph{memory dependency
-    graph} \(G_d\) as follows:
-
-    \begin{itemize}
-        \item \(G_d\) is an undirected graph.
-        \item Every vertex in \(G_d\) corresponds to a transaction and vice versa.
-        \item Vertices \(u\) and \(v\) are adjacent in \(G_d\) if and only if \(u\) has a memory location \(L\) in its
-        writing list and \(v\) has \(L\) in either its writing list or its reading list.
-    \end{itemize}
-
-    If we consider a proper vertex coloring of \(G_d\), every color class will give us an independent set of
-    transactions which can be executed concurrently. To achieve the highest parallelization, we need to color \(G_d\)
-    with minimum number of colors. Thus, the \emph{chromatic number} of the memory dependency graph shows how good a
-    transaction set could be run concurrently.
-
-    Graph coloring is computationally NP-hard. However, in our use case we don't need to necessarily find an optimal
-    solution. An approximate greedy algorithm will perform well enough in most circumstances.
-
-    After constructing the memory dependency graph, we can use it to construct the
-    \emph{execution DAG} of transactions. The execution DAG of transaction set \(T\) is a directed acyclic
-    graph \(G_e = (V_e,E_e)\) which has the \emph{execution invariance} property:
-    \begin{itemize}
-        \item Every vertex in \(V_e\) corresponds to a transaction in \(T\) and vice versa.
-        \item Executing the transactions of \(T\) in any order that \emph{respects} \(G_e\) will result in
-        the same AVM state.
-        \begin{itemize}
-            \item An ordering of transactions of \(T\) respects \(G_e\) if for every directed edge \((u,v) \in E_e\)
-            the transaction \(u\) comes before the transaction \(v\) in the ordering.
-        \end{itemize}
-    \end{itemize}
-
-    Having the execution DAG of a set of transactions, using Algorithm~\ref{alg:exec_dag}, we can apply the transaction
-    set to the AVM state concurrently, using multiple processor, while we can be sure that the resulted AVM state will
-    always be the same no matter how many processor we have used.
-
-    %##\includegraphics[width=17cm]{../img/Alg1s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Ready}{$R_e$}\SetKwData{V}{$v_f$}\SetKwData{Graph}{$G_e$}\SetKwData{Vertices}{$V$}\SetKwData
-        {Txns}{$T$}
-        \KwData{The execution dag $\Graph = (\Vertices,E)$ of transaction set \Txns}
-        \KwResult{The state of the AVM after applying \Txns with any ordering respecting \Graph}
-        \BlankLine
-        \Ready $\gets$ the set of all vertices of \Vertices with in degree 0\;
-        \While{$\Vertices \neq \varnothing$}
-        {
-            wait until a new free processor is available\;
-            \If{the execution of a transaction was finished}
-            {
-                remove the vertex of the finished transaction \V from \Graph\;
-                \For{each vertex $u \in Adj[\V]$}
-                {
-                    \If{$u$ has zero in degree}
-                    {
-                        $\Ready \gets \Ready \cup u$\;
-                    }
-                }
-            }
-            \If{$\Ready \neq \varnothing$}
-            {
-                remove a vertex from \Ready and assign it to a processor\;
-            }
-        }
-        \caption{executing DAG transactions}\label{alg:exec_dag}
-    \end{algorithm}
-
-    By replacing every undirected edge of a memory dependency graph with a directed edge in such a way that the
-    resulted graph has no cycles, we will obtain a valid execution DAG. Thus, from a memory dependency graph different
-    execution DAGs can be constructed with different levels of parallelization ability.
-
-    If we assume that we have unlimited number of processors and all transactions take equal time for executing, it
-    can be shown that by providing a minimal graph coloring to Algorithm~\ref{alg:gen_dag} as input, the resulted
-    DAG will be optimal, in the sense that it results in the minimum overall execution time.
-
-    %##\includegraphics[width=17cm]{../img/Alg2s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Txns}{$T$}\SetKwData{Gd}{$G_d=(V_d,E_d)$}
-        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{The memory dependency graph \Gd of transaction set \Txns\\A proper coloring of $G_d$}
-        \Output{An execution dag $G_e=(V_e,E_e)$ for the transaction set \Txns}
-        \BlankLine
-        $V_e \gets V_d$\;
-        $E_e \gets \varnothing$\;
-        define a total order on colors of $G_d$\;
-        \For{each edge $\{u,v\} \in E_d$}
-        {
-            \eIf{$color[u] < color[v]$}
-            {
-                $E_e \gets E_e \cup (u,v)$\;
-            }{
-                $E_e \gets E_e \cup (v,u)$\;
-            }
-        }
-        \caption{Constructing an execution DAG}\label{alg:gen_dag}
-    \end{algorithm}
-
-    The block proposer is responsible for proposing an efficient execution DAG alongside his proposed block. This
-    execution DAG will determine the ordering of block transactions and help validators to validate transactions in
-    parallel. Since with better parallelization a block can contain more transactions, a proposer is incentivized enough
-    to find a good execution DAG for transactions.
-
-    \subsection{Memory Spooling}\label{subsec:spooling}
-
-    When two transactions are dependant and they are connected with an edge \((u,v)\) in the execution DAG,
-    the transaction \(u\) needs to be run before the transaction \(v\). However, if \(v\) does not read any
-    memory locations that \(u\) modifies, we can run \(u\) and \(v\) in parallel. We just need to make sure
-    \(u\) does not see any changes \(v\) is making in AVM memory. This can be done by appropriate versioning
-    of the memory locations which is shared between \(u\) and \(v\). We call this method \emph{memory spooling}.
-    After enabling memory spooling between two transactions the edge connecting them can be safely removed from the
-    execution DAG\@.
-
-    \subsection{Concurrent Counters}\label{subsec:concurrent-counters}
-
-    We know that in Argennon every transaction needs to transfer its proposed fee to the \texttt{feeSink} accounts
-    first. This essentially makes every transaction a reader and a writer of the memory locations which store the
-    balance record of the \texttt{feeSink} accounts. As a result, all transactions in Argennon will be dependant and
-    parallelism will be completely impossible. Actually, any account that is highly active, for example the account
-    of an exchange or a payment processor, could become a concurrency bottleneck in our system which makes all
-    transactions interacting with them dependant.
-
-    This problem can be easily solved by using a concurrent counter for storing the balance record of this type of
-    accounts. A concurrent counter is a data structure which improves concurrency by using multiple memory locations for
-    storing a single counter. The value of the concurrent counter is equal to the sum of its sub counters and it can
-    be incremented or decremented by incrementing/decrementing any of the sub counters. This way, a concurrent
-    counter trades concurrency with memory usage.
-
-    Algorithm~\ref{alg:CC} implements a concurrent counter which returns an error when the value of the counter
-    becomes negative.
-
-    %##\includegraphics[width=17cm]{../img/Alg3s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{CC}{Counter}
-        \SetKwFunction{Inc}{Increment}\SetKwFunction{Dec}{Decrement}\SetKwFunction{AtomInc}{AtomicIncrement}
-        \SetKwFunction{AtomDec}{AtomicDecrement}\SetKwFunction{AtomSet}{AtomicSet}\SetKwFunction{Get}{GetValue}
-        \SetKwFunction{Acquire}{Lock.Acquire}\SetKwFunction{Release}{Lock.Release}
-        \SetKwProg{Fn}{Function}{}{}
-        \Fn{\Get{\CC}}
-        {
-            $s \gets 0$\;
-            \Acquire{}\;
-            \For{$i \gets 0$ \KwTo $\CC.size - 1$}
-            {
-                $s \gets s + \CC.cell[i]$\;
-            }
-            \Release{}\;
-            \KwRet{s}\;
-        }
-        \BlankLine
-        \Fn{\Inc{\CC, value, seed}}
-        {
-            $i \gets seed \bmod \CC.size$\;
-            \AtomInc{$\CC.cell[i]$, value}\;
-        }
-        \BlankLine
-        \Fn{\Dec{\CC, value, seed, attempt}}
-        {
-            \If {attempt = \CC.size}
-            {
-                restore \CC by adding back the subtracted value\;
-                \KwRet{Error}\;
-            }
-            $i \gets seed \bmod \CC.size$\;
-            $i \gets (i + attempt) \bmod \CC.size$\;
-            \eIf {$\CC.cell[i] \geq value$}
-            {
-                \AtomDec{$\CC.cell[i]$, value}\;
-            }{
-                $r \gets value - \CC.cell[i]$\;
-                \AtomSet{$\CC.cell[i]$, $0$}\;
-                \Dec{\CC, r, seed, $attempt + 1$}\;
-            }
-        }
-        \caption{Concurrent counter}\label{alg:CC}
-    \end{algorithm}
-
-    It should be noted that in a blockchain application we don't have concurrent threads and therefore we don't need
-    atomic functions. For usage in a smart contract, the atomic functions of this pseudocode can be implemented like
-    normal functions.
-
-    Concurrent counter data structure is a part of the AVM standard library, and any smart contract can use this data
-    structure for storing the balance record of highly active accounts.
-
-    \subsection{Memory Chunks}\label{subsec:memory-chunks}
-
-    In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{chunks}.
-    Each memory chunk can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
-    consensus on new values of the commitment of any chunk can be achieved by a different voting committee.
-
-    If a transaction does not modify a memory chunk and in the transaction ordering of the block it comes after
-    any transaction which modifies that chunk, then the execution of that transaction is not needed for calculating
-    the new commitment of the chunk. Consequently, the voting committee of that memory chunk can safely ignore such a
-    transaction. The execution DAG of transactions can be used for finding and pruning these transactions as
-    we see in Algorithm~\ref{alg:prune_dag}.
-
-    %##\includegraphics[width=17cm]{../img/Alg4s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{V}{$v_0$}\SetKwData{Graph}{$G_e$}\SetKwData{Chunk}{$C_k$}\SetKwData{Txns}{$T$}
-        \SetKwFunction{RPrune}{RecursivePrune}
-        \SetKwProg{Fn}{Function}{}{}
-        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{An execution dag \Graph and a memory chunk \Chunk}
-        \Output{The pruned execution dag based on \Chunk}
-        \BlankLine
-        \For{each vertex \V with out degree $0$}
-        {
-            \RPrune{\V}\;
-        }
-        \BlankLine
-        \Fn{\RPrune{\V}}
-        {
-            \If{$Adj[\V] = \varnothing$ {\bf and} \V does not modify \Chunk}
-            {
-                remove \V from \Graph\;
-                \For{each $u$ such that edge $(u,\V)$ was in \Graph}
-                {
-                    \RPrune{u}\;
-                }
-            }
-        }
-        \caption{Pruning an execution DAG}\label{alg:prune_dag}
-    \end{algorithm}
-
-    If we choose chunks in a way that most transactions only modify memory locations of one chunk,
-    likely many transactions of a block only need to be validated by one voting committee and can be validated in
-    parallel by different committees.
-
-    Because the voting committees are selected by random sampling, by choosing large enough samples we can make sure
-    that having multiple voting committees will not change the security properties of the Argennon agreement protocol.
+\input{sections/concurrency}
 
@@
 and so on.
-
-\end{document}

### 2021-04-26 tex/sections/concurrency.tex
Change project structure
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-03-29 tex/avm.tex
@@
 %! Author = aybehrouz
-%! Date = 2/28/21
-
-% Preamble
-\documentclass[a4paper]{report}
-
-% Packages
-%\usepackage{amsmath}
-%\usepackage[a4paper]{geometry}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{amsfonts}
-\usepackage[ruled,boxed]{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{tcolorbox}
-\lstset{
-%  basicstyle=\ttfamily,
-    mathescape
-}
-%\binoppenalty=99999
-%\relpenalty=99999
-\newcommand{\note}[1] {
-    \begin{tcolorbox}[colframe=white,colback=white]
-        \emph{#1}
-        %\emph{#1}\medskip
-    \end{tcolorbox}
-}
-
-\title{The Argennon Virtual Machine}
-\author{aybehrouz}
-\date{January 2021}
-
-% Document
-\begin{document}
-    \maketitle
-    \tableofcontents
-
-    \chapter{Specification}\label{ch:specification}
-
-    \section{Introduction}\label{sec:introduction}
-
-    The Argennon Virtual Machine (AVM) is an abstract computing machine for executing Argennon's smart contracts. It
-    is designed in a way that it could be efficiently implemented in either hardware or software.
-
-    \section{The Structure of the Argennon Virtual Machine}\label{sec:the-structure-of-the-algorand-virtual-machine}
-
-    ...
-
-    \subsection{Data Types}\label{subsec:data-types}
-
-    The Argennon Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
-    and does not have to be done by the Argennon Virtual Machine itself.
-
-    The instruction set of the Argennon Virtual Machine distinguishes its operand types using instructions intended to
-    operate on values of specific types. For instance, \texttt{iadd} assumes that its operands are two 64-bit integers.
-
-    \subsection{The \texttt{PC} Register}\label{subsec:the-pc-register}
-
-    ...
-
-    \subsection{Call Stack}\label{subsec:call-stack}
-
-    A call stack contains the information that is needed for restoring the state of the invoker of a method.
-
-    \subsection{Memory Unit}\label{subsec:memory-unit}
-
-    The Argennon Virtual Machine has a byte addressable memory which is divided into separate segments. Every segment
-    belongs to a smart contract that has a unique \texttt{applicationID}. The AVM always has a single working memory
-    segment and memory locations outside its current working segment can not be accessed. The only instructions which
-    can change the current working segment are \texttt{invoke\_external}, \texttt{athrow} and return instructions.
-
-    Every smart contract has its own memory segment. Hence, there is no way for a smart contract to access another
-    smart contract's memory. Interaction between smart contracts is done using \texttt{invoke\_external} instruction,
-    and a smart contract can invoke methods of another smart contract by this instruction.
-
-    A memory segment consists of five data areas:
-    \begin{itemize}
-        \item Code Area
-        \item Constant Area
-        \item Local Frame
-        \item Operand Stack
-        \item Heap
-    \end{itemize}
-    All data areas except operand stack, have their own address space which starts from 0. Operand stack is a
-    last-in-first-out (LIFO) stack and is not addressable. Every memory access instruction operates on its specific
-    data areas.
-
-    \subsubsection{Code Area}
-
-    The code area of a segment contains the byte-code of the smart contract which owns that segment. The AVM has no
-    instructions for manipulating the code area. \textbf{Installing, removing and updating smart contracts need to
-    be done externally.}
-
-    \subsubsection{Constant Area}
-
-    The constant area of a segment contains several kinds of constants, ranging from user defined constants to method
-    address tables. A method address table stores method locations in the code area and their access type. The access
-    type of a method can be either \texttt{public} or \texttt{private}. The AVM has no instructions for modifying the
-    constant area.
-
-    \note{Only public methods can be invoked by \texttt{invoke\_external} instruction.}
-
-    \subsubsection{Local Frame}
-
-    A local frame is used to store methods parameters and local variables. A new frame is created each time a method
-    is invoked, and it is destroyed when its method invocation completes, whether the completion is normal or abrupt.
-
-    \subsubsection{Operand Stack}
-
-    Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too. AVM
-    instructions take operands from the operand stack, operate on them, and push the result back onto the operand
-    stack. An operand stack is destroyed when its owner method completes, whether that completion is normal or abrupt.
-
-    \subsubsection{Heap}
-
-    The heap of a segment is a persistent memory area which is divided into pages. Memory locations inside every page
-    have a separate address space that starts from 0, and each page can be referenced by an index that starts from 0. In
-    other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex, offset)}.
-    Different pages of a heap do not need to be equally sized.
-
-    \note{The reason behind this paged design is that the heap is usually persisted using a block device. A heap with
-    a paged structure could expose the underlying block based nature of the persistence layer to the application
-    layer. In this way, the compiler or the programmer could better optimize the code for the persistence layer.}
-
-    \section{Instruction Set Summary}\label{sec:instruction-set-summary}
-
-    An Argennon Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
-    performed, followed by zero or more operands supplying arguments or data that are used by the operation.
-    \textbf{The number and size of the operands are determined solely by the opcode.}
-
-    \subsection{Method Invocation}\label{subsec:method-invocation}
-
-    The Argennon Virtual Machine has three types of method invocation:
-    \begin{itemize}
-        \item \texttt{invoke\_internal} invokes a method from the current running smart contract.
-        \item \texttt{invoke\_external} invokes a \texttt{public} method from another smart contract. It will change
-        the current memory segment to the segment of the invoked smart contract.
-        \item \texttt{invoke\_native} invokes a method that is not hosted by the Argennon virtual machine. By this
-        instruction, high performance native methods of the hosting machine could become available to AVM smart
-        contracts.
-    \end{itemize}
-
-    \note{In the future, we may need to add special instructions for invoking interface and virtual methods...}
-
-    Each time a method is invoked a new local frame and operand stack is created. The Argennon Virtual Machine uses
-    local frames to pass parameters on method invocation. On method invocation, any parameters are passed in
-    consecutive local variables stored in the method's local frame starting from local variable 0. The invoker of a
-    method writes the parameters in the local frame of the invoked method using \texttt{arg} instructions.
-
-    \subsubsection{Exceptions}
-
-    An exception is thrown programmatically using the \texttt{athrow} instruction. Exceptions can also be thrown by
-    various Argennon Virtual Machine instructions if they detect an abnormal condition. Some exceptions are not
-    catchable and will always abort the execution of the smart contract.
-
-    \subsubsection{Method Invocation Completion}
-
-    A method invocation completes normally if that invocation does not cause an exception to be thrown, either
-    directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
-    method completes normally, then a value may be returned to the invoking method. This occurs when the invoked
-    method executes one of the return instructions, the choice of which must be appropriate for the type of the value
-    being returned (if any). Execution then continues normally in the invoking method's local frame with the returned
-    value (if any) pushed onto the operand stack.
-
-    A method invocation completes abruptly if an exceptions is thrown and is not caught by the current method. A
-    method invocation that completes abruptly never returns a value to its invoker.
-
-    When a method completes, whether normally or abruptly, the call stack is used to restore the state of the invoker,
-    including its local frame and operand stack, with the \texttt{PC} register appropriately restored and incremented
-    to skip past the method invocation instruction. If the invoker was another smart contract, i.e.\ the invocation
-    was made by an \texttt{invoke\_external} instruction, the current memory segment will be changed to the invoker's
-    segment.
-
-    A thrown exception causes methods in the call stack to complete \textbf{abruptly} one by one, as long as the
-    \texttt{PC} register is not pointing to a \texttt{catch} instruction. The \texttt{catch} instruction acts like a
-    branch instruction that branches only if an exception is caught. \textbf{When an external method invocation
-    completes abruptly, before changing the current segment, all changes made to the heap area by that method
-    call, including changes made to other segments, will be rolled back.} So, when a method call completes abruptly,
-    that method call essentially has no effect on the AVM state.
-
-    \note{By using the \texttt{athrow} instruction properly, a programmer can make any method act like an atomic
-    operation.}
-
-    \subsection{Authorizing Operations}\label{subsec:authorizing-operations}
-
-    In blockchain applications, we usually need to authorize certain operations. For example, for sending an asset
-    from a user to another user, first we need to make sure that the sender has authorized this operation. The
-    Argennon virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
-    cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
-    passing signatures as parameters to methods, a programmer can implement the required logic for authorizing any
-    operation.
-
-    \note{The Argennon virtual machine has no instructions for issuing cryptographic signatures.}
-
-    In addition to signatures, a method can verify its invoker by using \texttt{get\_parent} instruction. This
-    instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
-    smart contract in the call stack. In other words, it returns the \texttt{applicationID} of the smart contract that
-    has invoked the current smart contract. (if any)
-
-    \subsection{Heap Allocation Instructions}\label{subsec:heap-allocation-instructions}
-
-    ...
-
-    \section{The AVM Standard Library}\label{sec:the-avm-standard-library}
-
-    ...
-
-    \chapter{Implementation}\label{ch:implementation}
-
-    \section{Persistence}\label{sec:persistence}
-
-    For implementing the persistence layer of the AVM, we assume that we have access to an updatable zero-knowledge
-    elementary database (ZK-EDB) with the following properties:
-
-    \begin{itemize}
-        \item The ZK-EDB contains a mapping from a set of keys to a set of values.
-        \item Every state of the database has a commitment \(C\).
-        \item The ZK-EDB has a method \((D, p) = get(x)\), where \(x\) is a key and \(D\) is the associated data
-        with \(x\), and \(p\) is a proof.
-        \item A user can use \(C\) and \(p\) to verify that \(D\) is really associated with \(x\), and \(D\) is not
-        altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB\@.
-        \item Having \(p\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
-        associated with \(x\) instead of \(D\).
-    \end{itemize}
-
-    We use a ZK-EDB for storing the AVM heap. We include the commitment of the current state of this DB in every
-    block of the Argennon blockchain, so ZK-EDB servers need not be trusted servers.
-
-    Every page of AVM heap will be stored with a key of the form: \texttt{applicationID|pageIndex} (the \texttt{|}
-    operator concatenates two numbers). Nodes do not keep a full copy of the AVM heap and for validating block
-    certificates or emulating the AVM ( i.e.~validating transactions) they need to connect to a ZK-EDB and retrieve
-    the required pages of AVM heap. For better performance, nodes keep a cache of heap pages to
-    reduce the amount of ZK-EDB access.
-
-    We also use a ZK-EDB for storing the code area of each segment, and we include the commitment of this DB in every
-    block. Every code area will be divided into blocks and every block will be stored in the DB with
-    \texttt{applicationID|blockID} as its key. Like heap pages, nodes keep a cache of code area blocks.
-
-    \note{Unlike heap pages, the AVM is not aware of different blocks of code area.}
-
-    \section{Transactions}\label{sec:transactions}
-
-    Argennon has four types of transaction:
-
-    \begin{itemize}
-        \item \texttt{avmCall} essentially is an \texttt{invoke\_external} instruction that invokes a method from an
-        AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
-        assets, including ARGs, is done by these transactions.
-        \item \texttt{installApp} installs an AVM smart contract and determines the update policy of the smart
-        contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
-        on.
-        \item \texttt{unInstallApp} removes an AVM smart contract.
-        \item \texttt{updateApp} updates an AVM smart contract.
-    \end{itemize}
-
-    All types of Argennon transactions contain an \texttt{invoke\_external} instruction which calls a special method
-    from ARG smart contract that transfers the proposed fee of the transaction in ARGs from a sender account to the
-    fee sink accounts.
-
-    Every transaction is required to exactly specify what heap locations or code area addresses it will access. This
-    enables validators to start retrieving the required memory blocks from available ZK-EDB servers as soon as they
-    see a transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to
-    access a memory location that is not included in its access lists, will be rejected. Users could use smart contract
-    oracles to predict the list of memory blocks their transactions need. See Section~\ref{sec:smart-contract-oracle}
-    for more details.
-
-    \section{Blockchain}\label{sec:blockchain}
-
-    Every block of the Argennon blockchain corresponds to a set of transactions. We store the commitment of this
-    transaction set in every block, but we don't keep the set itself. To be able to detect replay attacks, we require
-    every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
-    and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
-    must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
-    a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
-    if a signature of the same user with a bigger or equal nonce is already used
-    (i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
-    nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
-
-    To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. When the
-    difference between \texttt{issuance} component of this nonce and the current round becomes bigger than the
-    maximum allowed lifetime of a signature, this information can be safely deleted. \textbf{As a result, we will not
-    have the problem of "un-removable empty accounts" like Ethereum.}
-
-    The only information that Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
-    blockchain. Every block of the Argennon blockchain contains the following information:
-
-    \begin{center}
-        \begin{tabular}{||c||}
-            \hline
-            Block \\ [0.5ex]
-            \hline\hline
-            commitment to the ZK-EDB storing heap pages \\ [0.7ex]
-            commitment to the ZK-EDB storing code areas \\ [0.7ex]
-            commitment to the set of transactions       \\ [0.7ex]
-            previous block hash                         \\ [0.7ex]
-            random seed                                 \\ [0.7ex]
-            \hline
-        \end{tabular}
-    \end{center}
-
-    For confirming a new block, nodes that are not validators only need to verify the block certificate. For
-    verifying a block certificate, a node needs to know the ARG balances of validators, but it doesn't need to
-    emulate the AVM execution.
-
-    On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
-    execution of the Argennon virtual machine. To do so, first they retrieve all heap pages and code area blocks they
-    need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
-    transactions included in the new block. This will modify some pages of the AVM memory, so they update the ZK-EDB
-    commitments based on the modified pages and verify the commitments included in the new block. Validators also
-    calculate and verify the commitment to the new block's transaction set.
-
-    \note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDB servers will receive the new
-    block, and they will update their database by emulating the AVM execution.}
-
-    \section{Consensus}\label{sec:consensus}
-
-    \note{It will be a randomized proof of stake which is under development....}
-
-    \subsection{Estimating A User's Stake}\label{subsec:estimating-a-user's-stake}
-
-    In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
-    of stake the user has in the system. Conventionally in these systems, for estimating a user's stake, we use the
-    amount of native system tokens the user is holding. Unfortunately, one problem with this approach is that a
-    strong attacker may be able to obtain a considerable amount of system tokens, for example by borrowing from a
-    DEFI application, and use this stake to attack the system.
-
-    To mitigate this problem, for calculating a user's stake at time step \(t\), instead of using the raw ARG
-    balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
-    ARG balance:
-    \[
-        S_{u,t} = \min (B_{u,t}, Trust_{u,t})
-    \]
-    Where:
-    \begin{itemize}
-        \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
-        \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
-        \item \(Trust_{u,t}\) is an estimated trust value for user \(u\) at time step \(t\).
-    \end{itemize}
-
-    The agreement protocol, at time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
-    number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
-    exponential moving average of the ARG balance of user \(u\) at time step \(t\).
-
-    In our system a user who held ARGs and participated in the consensus for a long time is more trusted
-    than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
-    amount of ARGs, also needs to hold them for a long period of time before being able to attack the system.
-
-    For calculating the exponential moving average of a user's balance at time step \(t\), we can use the following
-    recursive formula:
-    \[
-        M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
-    \]
-    Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
-    degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
-
-    Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
-    \(M_{u,t}\): (In the following equations the \(u\) subscript is dropped for simplicity)
-    \[
-        M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B
-    \]
-    Where:
-    \[
-        B = B_{k+1} = B_{k+2} = \dots = B_{t}
-    \]
-    We know that when \(|nx| \ll 1\) we can use the binomial approximation \({(1 + x)^n \approx 1 + nx}\). So, we can
-    further simplify this formula:
-    \[
-        M_{t} = M_{k} + (t - k) \alpha (B - M_{k})
-    \]
-
-    For choosing the value of \(\alpha\) we can consider the number of time steps that the trust value of a user needs
-    for reaching a specified fraction of his account balance. We know that for large \(n\) and \(|x| < 1\) we have
-    \((1 + x)^n \approx e^{nx}\), so by letting \(M_{u,k} = 0\) and \(n = t - k\) we can write:
-    \[
-        \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{B}\right)}{n}
-    \]
-    The value of \(\alpha\) for a desired configuration can be calculated by this equation. For instance, we could
-    calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8B\) and \(n\) equals to the
-    number of time steps of 10 years.
-
-    In our system a newly created account will not have voting power for some time, no matter how high its
-    balance is. While this is a desirable property, in case a large proportion of total system tokens are
-    transferred to newly created accounts, it can result in too much voting power for older accounts. This may decrease
-    the degree of decentralization in our system.
-
-    However, this situation is easily detectable by comparing the total stake of the system with the total balance of
-    users. If after confirming a block the total stake of the system goes too low and we have:
-    \[
-        \sum_{u}S_{u,t} < \gamma \sum_{u}B_{u,t}
-    \]
-    The protocol will perform a \emph{time shift} in the system: the time step of the system
-    will be incremented for \(m\) steps while no blocks will be confirmed. This will increase the value of \(M_{u,t}\)
-    for new accounts with a non-zero balance, giving them more influence in the agreement protocol.
-
-    For calculating the value of \(m\) which determines the amount of time shift in the system, we should note that when
-    \(B_{u,t} = B_{u, t-1} = B_u\), we can derive a simple recursive rule for the stake of a user:
-    \[
-        S_{u,t} = (1 - \alpha) S_{u,t-1} + \alpha B_u
-    \]
-    Therefore, we have:
-    \[
-        \sum_{u}S_{u,t} = (1 - \alpha) \sum_{u}S_{u,t - 1} + \alpha \sum_{u}B_u
-    \]
-    This equation shows that when the balance of users is not changing over time the total stake of the system is the
-    exponential average of the total ARGs of the system. Consequently, when we shift the time for \(m\) steps, we can
-    calculate the new total stake of the system from the following equation:
-
-    \[
-        \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
-    \]
-    Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
-    we can obtain \(m\) from the following formula, assuming \(\alpha\) is small enough:
-    \[
-        m = \frac{1}{\alpha} \ln \left(\frac{1 - \gamma}{1 - \lambda}\right)
-    \]
-
-    \section{Incentive mechanism}\label{sec:incentive-mechanism}
-
-    \subsection{Transaction Fee}\label{subsec:transaction-fee}
-
-    Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
-    special method from ARG smart contract. This method will transfer the proposed fee of the transaction in ARGs
-    from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
-    execution fees and \texttt{dbFeeSink} collects fees for ZK-EDB servers. The Protocol decides how to distribute the
-    transaction fee between these two fee sink accounts.
-
-    When a block is added to the blockchain, the proposer of that block will receive a share of the block fees.
-    Consequently, a block proposer is always incentivized to include more transactions in his block. However, if he
-    puts too many transactions in his block and the validation of the block becomes too difficult, some validators
-    may not be able to validate all transactions on time. If a validator can not validate a block in the required
-    time, he will consider the block invalid. So, when a proposed block contains too many transactions, the network
-    may reach consensus on another block, and the proposer of that block will not receive any fees. As a result, a
-    proposer is incentivized to use network transaction capacity optimally.
-
-    On the other hand, we believe that the proposer does not have enough incentives for optimizing the storage size
-    of the transaction set. Therefore, we require that \textbf{the size of the transaction set of every block in
-    bytes be lower than a certain threshold.}
-
-    Validators need to spend resources for validating transactions. When a validator starts the emulation of the AVM
-    to validate a transaction, solely from the code he can't predict the time the execution will finish. This will
-    give an adversary an opportunity to attack the network by broadcasting transactions that never ends. Since,
-    validators can not finish the execution of these transactions, the network will not be able to charge the
-    attacker any fees, and he would be able to waste validators resources for free.
-
-    To mitigate this problem, we require that every transaction specify a cap for all the resources it needs. This
-    will include memory, network and processor related resources. Also, the protocol defines an execution cost for
-    every AVM instruction reflecting the amount of resources its emulation needs. This will define a standard way for
-    measuring the execution cost of any \texttt{avmCall} transaction. Every \texttt{avmCall} transaction is required
-    to specify a maximum execution cost. If during emulation it reaches this maximum cost, the transaction will be
-    considered failed and the network can receive the proposed fee of that transaction.
-
-    Every \texttt{avmCall} transaction is required to provide the following information as an upper bound for the
-    resources it needs:
-
-    \begin{itemize}
-        \item Execution cost
-        \item A list of heap/code-area locations for reading
-        \item A list of heap locations for writing
-        \item A list of heap pages it will deallocate
-        \item Number and size of heap pages it will allocate
-    \end{itemize}
-
-    If a transaction tries to violate any of these predefined limitations, for example, if it tries to read a memory
-    location that is not included in its reading list, it will be considered failed and the network can receive the
-    proposed fee of that transaction.
-
-    \note{A transaction always pays all of its proposed fee, no matter how much of its predefined resources were not
-    used in the final emulation.}
-
-    \subsection{Incentives for ZK-EDB Servers}\label{subsec:zk-edb-servers}
-
-    The incentive mechanism for ZK-EDB servers should have the following properties:
-
-    \begin{itemize}
-        \item It incentivizes storing all memory blocks, whether a heap page or a code area block, and not only those
-        which are used more frequently.
-        \item It incentivizes ZK-EDB servers to actively provide the required memory blocks for validators.
-        \item Making more accounts will not provide any advantages for a ZK-EDB server.
-    \end{itemize}
-
-    For our incentive mechanism, we require that every time a validator receives a memory block from a ZK-EDB, after
-    validating the data, he give a receipt to the ZK-EDB. In this receipt the validator signs the following information:
-
-    \begin{itemize}
-        \item \texttt{ownerAddr}: the ARG address of the ZK-EDB\@.
-        \item \texttt{receivedBlockID}: the ID of the received memory block.
-        \item \texttt{round}: the current round number.
-    \end{itemize}
-
-    \note{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}
-
-    To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ARGs from
-    \texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
-    between all \emph{winning tickets} of the lottery.
-
-    \note{One ZK-EDB server could own multiple winning tickets in a round.}
-
-    To run this lottery, every round, based on the current block seed, a collection of \emph{valid} receipts will be
-    selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in round \(r\) if:
-
-    \begin{itemize}
-        \item The signer was a validator in the round \(r - 1\) and voted for the agreed-upon block.
-        \item The data block in the receipt was needed for validating the \textbf{previous} block.
-        \item The receipt round number is \(r - 1\).
-        \item The signer did not sign a receipt for the same data block for two different ZK-EDBs in the previous round.
-    \end{itemize}
-    For selecting the winning receipts we could use a random generator:
-    \begin{verbatim}
-IF random(seed|validatorPK|receivedBlockID) < winProbability THEN
-    the receipt issued by validatorPK for receivedBlockID is a winner
-    \end{verbatim}
-    \begin{itemize}
-        \item \texttt{random()} produces uniform random numbers between 0 and 1, using its input argument as a seed.
-        \item \texttt{validatorPK} is the public key of the signer of the receipt.
-        \item \texttt{receivedBlockID} is the ID of the memory block that the receipt was issued for.
-        \item \texttt{winProbability} is the probability of winning in every round.
-        \item \texttt{seed} is the current block seed.
-        \item \texttt{|} is a concatenation operator.
-    \end{itemize}
-
-    \note{The winners of the lottery were validators one round before the lottery round.}
-
-    Also, based on the current block seed, a random memory block, whether a heap page or a code area block, is
-    selected as the challenge of the round. A ZK-EDB that owns a winning receipt needs to broadcast a \emph{winning
-    ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
-    challenge. Solving a round challenge requires the content of the memory block which was selected as the round
-    challenge. This will encourage ZK-EDBs to store all memory blocks.
-
-    A possible choice for the challenge solution could be the cryptographic hash of the content of the challenge
-    memory block combined with the ZK-EDB ARG address: \texttt{hash(challenge.content|ownerAddr)}
-
-    The winning tickets of the lottery of round \(r\) need to be included in the block of the round \(r\),
-    otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
-    \(r\) will be done in the round \(r + 1\). This way, \textbf{the content of the challenge memory block could be
-    kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.
-
-    \subsection{Memory Allocation and De-allocation Fee}\label{subsec:memory-allocation-and-de-allocation}
-
-    Every \(k\) round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
-    allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
-    address of the smart contract.
-
-    To determine the price of AVM memory, Every \(k\) round, the protocol calculates \texttt{dbFee} and
-    \texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
-    \texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
-    the protocol considers the total size of all the memory pages that were accessed for either reading or writing
-    during a time period. These two values will be calculated for the last \(k\) rounds and the price per byte of
-    AVM memory will be a linear function of \texttt{dbFee/memTraffic}
-
-    When a smart contract executes a heap de-allocation instruction, the protocol will refund the cost of
-    de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
-    calculates the refunded amount based on the average price the smart contract had paid for that allocated memory.
-    This will prevent smart contracts from profit taking by trading memory with the protocol.
-
-    \section{Concurrency}\label{sec:concurrency}
 
@@
 that having multiple voting committees will not change the security properties of the Argennon agreement protocol.
-
-    \section{Smart Contract Oracle}\label{sec:smart-contract-oracle}
-
-    A smart contract oracle is a full AVM emulator that keeps a full local copy of AVM memory and can emulate AVM
-    execution without accessing a ZK-EDB. Smart contract oracles can be used for reporting useful information about
-    \texttt{avmCall} transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
-    and so on.
-
-\end{document}

### 2021-04-26 tex/sections/consensus.tex
Change project structure
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-03-29 tex/avm.tex
@@
 %! Author = aybehrouz
-%! Date = 2/28/21
-
-% Preamble
-\documentclass[a4paper]{report}
-
-% Packages
-%\usepackage{amsmath}
-%\usepackage[a4paper]{geometry}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{amsfonts}
-\usepackage[ruled,boxed]{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{tcolorbox}
-\lstset{
-%  basicstyle=\ttfamily,
-    mathescape
-}
-%\binoppenalty=99999
-%\relpenalty=99999
-\newcommand{\note}[1] {
-    \begin{tcolorbox}[colframe=white,colback=white]
-        \emph{#1}
-        %\emph{#1}\medskip
-    \end{tcolorbox}
-}
-
-\title{The Argennon Virtual Machine}
-\author{aybehrouz}
-\date{January 2021}
-
-% Document
-\begin{document}
-    \maketitle
-    \tableofcontents
-
-    \chapter{Specification}\label{ch:specification}
-
-    \section{Introduction}\label{sec:introduction}
-
-    The Argennon Virtual Machine (AVM) is an abstract computing machine for executing Argennon's smart contracts. It
-    is designed in a way that it could be efficiently implemented in either hardware or software.
-
-    \section{The Structure of the Argennon Virtual Machine}\label{sec:the-structure-of-the-algorand-virtual-machine}
-
-    ...
-
-    \subsection{Data Types}\label{subsec:data-types}
-
-    The Argennon Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
-    and does not have to be done by the Argennon Virtual Machine itself.
-
-    The instruction set of the Argennon Virtual Machine distinguishes its operand types using instructions intended to
-    operate on values of specific types. For instance, \texttt{iadd} assumes that its operands are two 64-bit integers.
-
-    \subsection{The \texttt{PC} Register}\label{subsec:the-pc-register}
-
-    ...
-
-    \subsection{Call Stack}\label{subsec:call-stack}
-
-    A call stack contains the information that is needed for restoring the state of the invoker of a method.
-
-    \subsection{Memory Unit}\label{subsec:memory-unit}
-
-    The Argennon Virtual Machine has a byte addressable memory which is divided into separate segments. Every segment
-    belongs to a smart contract that has a unique \texttt{applicationID}. The AVM always has a single working memory
-    segment and memory locations outside its current working segment can not be accessed. The only instructions which
-    can change the current working segment are \texttt{invoke\_external}, \texttt{athrow} and return instructions.
-
-    Every smart contract has its own memory segment. Hence, there is no way for a smart contract to access another
-    smart contract's memory. Interaction between smart contracts is done using \texttt{invoke\_external} instruction,
-    and a smart contract can invoke methods of another smart contract by this instruction.
-
-    A memory segment consists of five data areas:
-    \begin{itemize}
-        \item Code Area
-        \item Constant Area
-        \item Local Frame
-        \item Operand Stack
-        \item Heap
-    \end{itemize}
-    All data areas except operand stack, have their own address space which starts from 0. Operand stack is a
-    last-in-first-out (LIFO) stack and is not addressable. Every memory access instruction operates on its specific
-    data areas.
-
-    \subsubsection{Code Area}
-
-    The code area of a segment contains the byte-code of the smart contract which owns that segment. The AVM has no
-    instructions for manipulating the code area. \textbf{Installing, removing and updating smart contracts need to
-    be done externally.}
-
-    \subsubsection{Constant Area}
-
-    The constant area of a segment contains several kinds of constants, ranging from user defined constants to method
-    address tables. A method address table stores method locations in the code area and their access type. The access
-    type of a method can be either \texttt{public} or \texttt{private}. The AVM has no instructions for modifying the
-    constant area.
-
-    \note{Only public methods can be invoked by \texttt{invoke\_external} instruction.}
-
-    \subsubsection{Local Frame}
-
-    A local frame is used to store methods parameters and local variables. A new frame is created each time a method
-    is invoked, and it is destroyed when its method invocation completes, whether the completion is normal or abrupt.
-
-    \subsubsection{Operand Stack}
-
-    Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too. AVM
-    instructions take operands from the operand stack, operate on them, and push the result back onto the operand
-    stack. An operand stack is destroyed when its owner method completes, whether that completion is normal or abrupt.
-
-    \subsubsection{Heap}
-
-    The heap of a segment is a persistent memory area which is divided into pages. Memory locations inside every page
-    have a separate address space that starts from 0, and each page can be referenced by an index that starts from 0. In
-    other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex, offset)}.
-    Different pages of a heap do not need to be equally sized.
-
-    \note{The reason behind this paged design is that the heap is usually persisted using a block device. A heap with
-    a paged structure could expose the underlying block based nature of the persistence layer to the application
-    layer. In this way, the compiler or the programmer could better optimize the code for the persistence layer.}
-
-    \section{Instruction Set Summary}\label{sec:instruction-set-summary}
-
-    An Argennon Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
-    performed, followed by zero or more operands supplying arguments or data that are used by the operation.
-    \textbf{The number and size of the operands are determined solely by the opcode.}
-
-    \subsection{Method Invocation}\label{subsec:method-invocation}
-
-    The Argennon Virtual Machine has three types of method invocation:
-    \begin{itemize}
-        \item \texttt{invoke\_internal} invokes a method from the current running smart contract.
-        \item \texttt{invoke\_external} invokes a \texttt{public} method from another smart contract. It will change
-        the current memory segment to the segment of the invoked smart contract.
-        \item \texttt{invoke\_native} invokes a method that is not hosted by the Argennon virtual machine. By this
-        instruction, high performance native methods of the hosting machine could become available to AVM smart
-        contracts.
-    \end{itemize}
-
-    \note{In the future, we may need to add special instructions for invoking interface and virtual methods...}
-
-    Each time a method is invoked a new local frame and operand stack is created. The Argennon Virtual Machine uses
-    local frames to pass parameters on method invocation. On method invocation, any parameters are passed in
-    consecutive local variables stored in the method's local frame starting from local variable 0. The invoker of a
-    method writes the parameters in the local frame of the invoked method using \texttt{arg} instructions.
-
-    \subsubsection{Exceptions}
-
-    An exception is thrown programmatically using the \texttt{athrow} instruction. Exceptions can also be thrown by
-    various Argennon Virtual Machine instructions if they detect an abnormal condition. Some exceptions are not
-    catchable and will always abort the execution of the smart contract.
-
-    \subsubsection{Method Invocation Completion}
-
-    A method invocation completes normally if that invocation does not cause an exception to be thrown, either
-    directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
-    method completes normally, then a value may be returned to the invoking method. This occurs when the invoked
-    method executes one of the return instructions, the choice of which must be appropriate for the type of the value
-    being returned (if any). Execution then continues normally in the invoking method's local frame with the returned
-    value (if any) pushed onto the operand stack.
-
-    A method invocation completes abruptly if an exceptions is thrown and is not caught by the current method. A
-    method invocation that completes abruptly never returns a value to its invoker.
-
-    When a method completes, whether normally or abruptly, the call stack is used to restore the state of the invoker,
-    including its local frame and operand stack, with the \texttt{PC} register appropriately restored and incremented
-    to skip past the method invocation instruction. If the invoker was another smart contract, i.e.\ the invocation
-    was made by an \texttt{invoke\_external} instruction, the current memory segment will be changed to the invoker's
-    segment.
-
-    A thrown exception causes methods in the call stack to complete \textbf{abruptly} one by one, as long as the
-    \texttt{PC} register is not pointing to a \texttt{catch} instruction. The \texttt{catch} instruction acts like a
-    branch instruction that branches only if an exception is caught. \textbf{When an external method invocation
-    completes abruptly, before changing the current segment, all changes made to the heap area by that method
-    call, including changes made to other segments, will be rolled back.} So, when a method call completes abruptly,
-    that method call essentially has no effect on the AVM state.
-
-    \note{By using the \texttt{athrow} instruction properly, a programmer can make any method act like an atomic
-    operation.}
-
-    \subsection{Authorizing Operations}\label{subsec:authorizing-operations}
-
-    In blockchain applications, we usually need to authorize certain operations. For example, for sending an asset
-    from a user to another user, first we need to make sure that the sender has authorized this operation. The
-    Argennon virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
-    cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
-    passing signatures as parameters to methods, a programmer can implement the required logic for authorizing any
-    operation.
-
-    \note{The Argennon virtual machine has no instructions for issuing cryptographic signatures.}
-
-    In addition to signatures, a method can verify its invoker by using \texttt{get\_parent} instruction. This
-    instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
-    smart contract in the call stack. In other words, it returns the \texttt{applicationID} of the smart contract that
-    has invoked the current smart contract. (if any)
-
-    \subsection{Heap Allocation Instructions}\label{subsec:heap-allocation-instructions}
-
-    ...
-
-    \section{The AVM Standard Library}\label{sec:the-avm-standard-library}
-
-    ...
-
-    \chapter{Implementation}\label{ch:implementation}
-
-    \section{Persistence}\label{sec:persistence}
-
-    For implementing the persistence layer of the AVM, we assume that we have access to an updatable zero-knowledge
-    elementary database (ZK-EDB) with the following properties:
-
-    \begin{itemize}
-        \item The ZK-EDB contains a mapping from a set of keys to a set of values.
-        \item Every state of the database has a commitment \(C\).
-        \item The ZK-EDB has a method \((D, p) = get(x)\), where \(x\) is a key and \(D\) is the associated data
-        with \(x\), and \(p\) is a proof.
-        \item A user can use \(C\) and \(p\) to verify that \(D\) is really associated with \(x\), and \(D\) is not
-        altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB\@.
-        \item Having \(p\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
-        associated with \(x\) instead of \(D\).
-    \end{itemize}
-
-    We use a ZK-EDB for storing the AVM heap. We include the commitment of the current state of this DB in every
-    block of the Argennon blockchain, so ZK-EDB servers need not be trusted servers.
-
-    Every page of AVM heap will be stored with a key of the form: \texttt{applicationID|pageIndex} (the \texttt{|}
-    operator concatenates two numbers). Nodes do not keep a full copy of the AVM heap and for validating block
-    certificates or emulating the AVM ( i.e.~validating transactions) they need to connect to a ZK-EDB and retrieve
-    the required pages of AVM heap. For better performance, nodes keep a cache of heap pages to
-    reduce the amount of ZK-EDB access.
-
-    We also use a ZK-EDB for storing the code area of each segment, and we include the commitment of this DB in every
-    block. Every code area will be divided into blocks and every block will be stored in the DB with
-    \texttt{applicationID|blockID} as its key. Like heap pages, nodes keep a cache of code area blocks.
-
-    \note{Unlike heap pages, the AVM is not aware of different blocks of code area.}
-
-    \section{Transactions}\label{sec:transactions}
-
-    Argennon has four types of transaction:
-
-    \begin{itemize}
-        \item \texttt{avmCall} essentially is an \texttt{invoke\_external} instruction that invokes a method from an
-        AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
-        assets, including ARGs, is done by these transactions.
-        \item \texttt{installApp} installs an AVM smart contract and determines the update policy of the smart
-        contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
-        on.
-        \item \texttt{unInstallApp} removes an AVM smart contract.
-        \item \texttt{updateApp} updates an AVM smart contract.
-    \end{itemize}
-
-    All types of Argennon transactions contain an \texttt{invoke\_external} instruction which calls a special method
-    from ARG smart contract that transfers the proposed fee of the transaction in ARGs from a sender account to the
-    fee sink accounts.
-
-    Every transaction is required to exactly specify what heap locations or code area addresses it will access. This
-    enables validators to start retrieving the required memory blocks from available ZK-EDB servers as soon as they
-    see a transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to
-    access a memory location that is not included in its access lists, will be rejected. Users could use smart contract
-    oracles to predict the list of memory blocks their transactions need. See Section~\ref{sec:smart-contract-oracle}
-    for more details.
-
-    \section{Blockchain}\label{sec:blockchain}
-
-    Every block of the Argennon blockchain corresponds to a set of transactions. We store the commitment of this
-    transaction set in every block, but we don't keep the set itself. To be able to detect replay attacks, we require
-    every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
-    and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
-    must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
-    a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
-    if a signature of the same user with a bigger or equal nonce is already used
-    (i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
-    nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
-
-    To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. When the
-    difference between \texttt{issuance} component of this nonce and the current round becomes bigger than the
-    maximum allowed lifetime of a signature, this information can be safely deleted. \textbf{As a result, we will not
-    have the problem of "un-removable empty accounts" like Ethereum.}
-
-    The only information that Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
-    blockchain. Every block of the Argennon blockchain contains the following information:
-
-    \begin{center}
-        \begin{tabular}{||c||}
-            \hline
-            Block \\ [0.5ex]
-            \hline\hline
-            commitment to the ZK-EDB storing heap pages \\ [0.7ex]
-            commitment to the ZK-EDB storing code areas \\ [0.7ex]
-            commitment to the set of transactions       \\ [0.7ex]
-            previous block hash                         \\ [0.7ex]
-            random seed                                 \\ [0.7ex]
-            \hline
-        \end{tabular}
-    \end{center}
-
-    For confirming a new block, nodes that are not validators only need to verify the block certificate. For
-    verifying a block certificate, a node needs to know the ARG balances of validators, but it doesn't need to
-    emulate the AVM execution.
-
-    On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
-    execution of the Argennon virtual machine. To do so, first they retrieve all heap pages and code area blocks they
-    need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
-    transactions included in the new block. This will modify some pages of the AVM memory, so they update the ZK-EDB
-    commitments based on the modified pages and verify the commitments included in the new block. Validators also
-    calculate and verify the commitment to the new block's transaction set.
-
-    \note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDB servers will receive the new
-    block, and they will update their database by emulating the AVM execution.}
-
-    \section{Consensus}\label{sec:consensus}
 
@@
 \]
-
-    \section{Incentive mechanism}\label{sec:incentive-mechanism}
-
-    \subsection{Transaction Fee}\label{subsec:transaction-fee}
-
-    Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
-    special method from ARG smart contract. This method will transfer the proposed fee of the transaction in ARGs
-    from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
-    execution fees and \texttt{dbFeeSink} collects fees for ZK-EDB servers. The Protocol decides how to distribute the
-    transaction fee between these two fee sink accounts.
-
-    When a block is added to the blockchain, the proposer of that block will receive a share of the block fees.
-    Consequently, a block proposer is always incentivized to include more transactions in his block. However, if he
-    puts too many transactions in his block and the validation of the block becomes too difficult, some validators
-    may not be able to validate all transactions on time. If a validator can not validate a block in the required
-    time, he will consider the block invalid. So, when a proposed block contains too many transactions, the network
-    may reach consensus on another block, and the proposer of that block will not receive any fees. As a result, a
-    proposer is incentivized to use network transaction capacity optimally.
-
-    On the other hand, we believe that the proposer does not have enough incentives for optimizing the storage size
-    of the transaction set. Therefore, we require that \textbf{the size of the transaction set of every block in
-    bytes be lower than a certain threshold.}
-
-    Validators need to spend resources for validating transactions. When a validator starts the emulation of the AVM
-    to validate a transaction, solely from the code he can't predict the time the execution will finish. This will
-    give an adversary an opportunity to attack the network by broadcasting transactions that never ends. Since,
-    validators can not finish the execution of these transactions, the network will not be able to charge the
-    attacker any fees, and he would be able to waste validators resources for free.
-
-    To mitigate this problem, we require that every transaction specify a cap for all the resources it needs. This
-    will include memory, network and processor related resources. Also, the protocol defines an execution cost for
-    every AVM instruction reflecting the amount of resources its emulation needs. This will define a standard way for
-    measuring the execution cost of any \texttt{avmCall} transaction. Every \texttt{avmCall} transaction is required
-    to specify a maximum execution cost. If during emulation it reaches this maximum cost, the transaction will be
-    considered failed and the network can receive the proposed fee of that transaction.
-
-    Every \texttt{avmCall} transaction is required to provide the following information as an upper bound for the
-    resources it needs:
-
-    \begin{itemize}
-        \item Execution cost
-        \item A list of heap/code-area locations for reading
-        \item A list of heap locations for writing
-        \item A list of heap pages it will deallocate
-        \item Number and size of heap pages it will allocate
-    \end{itemize}
-
-    If a transaction tries to violate any of these predefined limitations, for example, if it tries to read a memory
-    location that is not included in its reading list, it will be considered failed and the network can receive the
-    proposed fee of that transaction.
-
-    \note{A transaction always pays all of its proposed fee, no matter how much of its predefined resources were not
-    used in the final emulation.}
-
-    \subsection{Incentives for ZK-EDB Servers}\label{subsec:zk-edb-servers}
-
-    The incentive mechanism for ZK-EDB servers should have the following properties:
-
-    \begin{itemize}
-        \item It incentivizes storing all memory blocks, whether a heap page or a code area block, and not only those
-        which are used more frequently.
-        \item It incentivizes ZK-EDB servers to actively provide the required memory blocks for validators.
-        \item Making more accounts will not provide any advantages for a ZK-EDB server.
-    \end{itemize}
-
-    For our incentive mechanism, we require that every time a validator receives a memory block from a ZK-EDB, after
-    validating the data, he give a receipt to the ZK-EDB. In this receipt the validator signs the following information:
-
-    \begin{itemize}
-        \item \texttt{ownerAddr}: the ARG address of the ZK-EDB\@.
-        \item \texttt{receivedBlockID}: the ID of the received memory block.
-        \item \texttt{round}: the current round number.
-    \end{itemize}
-
-    \note{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}
-
-    To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ARGs from
-    \texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
-    between all \emph{winning tickets} of the lottery.
-
-    \note{One ZK-EDB server could own multiple winning tickets in a round.}
-
-    To run this lottery, every round, based on the current block seed, a collection of \emph{valid} receipts will be
-    selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in round \(r\) if:
-
-    \begin{itemize}
-        \item The signer was a validator in the round \(r - 1\) and voted for the agreed-upon block.
-        \item The data block in the receipt was needed for validating the \textbf{previous} block.
-        \item The receipt round number is \(r - 1\).
-        \item The signer did not sign a receipt for the same data block for two different ZK-EDBs in the previous round.
-    \end{itemize}
-    For selecting the winning receipts we could use a random generator:
-    \begin{verbatim}
-IF random(seed|validatorPK|receivedBlockID) < winProbability THEN
-    the receipt issued by validatorPK for receivedBlockID is a winner
-    \end{verbatim}
-    \begin{itemize}
-        \item \texttt{random()} produces uniform random numbers between 0 and 1, using its input argument as a seed.
-        \item \texttt{validatorPK} is the public key of the signer of the receipt.
-        \item \texttt{receivedBlockID} is the ID of the memory block that the receipt was issued for.
-        \item \texttt{winProbability} is the probability of winning in every round.
-        \item \texttt{seed} is the current block seed.
-        \item \texttt{|} is a concatenation operator.
-    \end{itemize}
-
-    \note{The winners of the lottery were validators one round before the lottery round.}
-
-    Also, based on the current block seed, a random memory block, whether a heap page or a code area block, is
-    selected as the challenge of the round. A ZK-EDB that owns a winning receipt needs to broadcast a \emph{winning
-    ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
-    challenge. Solving a round challenge requires the content of the memory block which was selected as the round
-    challenge. This will encourage ZK-EDBs to store all memory blocks.
-
-    A possible choice for the challenge solution could be the cryptographic hash of the content of the challenge
-    memory block combined with the ZK-EDB ARG address: \texttt{hash(challenge.content|ownerAddr)}
-
-    The winning tickets of the lottery of round \(r\) need to be included in the block of the round \(r\),
-    otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
-    \(r\) will be done in the round \(r + 1\). This way, \textbf{the content of the challenge memory block could be
-    kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.
-
-    \subsection{Memory Allocation and De-allocation Fee}\label{subsec:memory-allocation-and-de-allocation}
-
-    Every \(k\) round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
-    allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
-    address of the smart contract.
-
-    To determine the price of AVM memory, Every \(k\) round, the protocol calculates \texttt{dbFee} and
-    \texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
-    \texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
-    the protocol considers the total size of all the memory pages that were accessed for either reading or writing
-    during a time period. These two values will be calculated for the last \(k\) rounds and the price per byte of
-    AVM memory will be a linear function of \texttt{dbFee/memTraffic}
-
-    When a smart contract executes a heap de-allocation instruction, the protocol will refund the cost of
-    de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
-    calculates the refunded amount based on the average price the smart contract had paid for that allocated memory.
-    This will prevent smart contracts from profit taking by trading memory with the protocol.
-
-    \section{Concurrency}\label{sec:concurrency}
-
-    \subsection{Memory Dependency Graph}\label{subsec:memory-dependency-graph}
-
-    Every block of the Argennon blockchain contains a list of transactions. This list is an ordered list and the
-    effect of its contained transactions must be applied to the AVM state sequentially as they appear in the ordered
-    list. This ordering is solely chosen by the block proposer, and users should not have any assumptions about
-    the ordering of transactions in a block.
-
-    The fact that block transactions constitute a sequential list, does not mean they can not be executed and applied
-    to the AVM state concurrently. Many transactions are actually independent and the order of their execution does not
-    matter. These transactions can be safely validated in parallel by validators.
-
-    A transaction can change the AVM state by modifying either the code area or the AVM heap. In Argennon, all
-    transactions declare the list of memory locations they want to read or write. This will enable us to determine the
-    independent sets of transactions which can be executed in parallel. To do so, we define the \emph{memory dependency
-    graph} \(G_d\) as follows:
-
-    \begin{itemize}
-        \item \(G_d\) is an undirected graph.
-        \item Every vertex in \(G_d\) corresponds to a transaction and vice versa.
-        \item Vertices \(u\) and \(v\) are adjacent in \(G_d\) if and only if \(u\) has a memory location \(L\) in its
-        writing list and \(v\) has \(L\) in either its writing list or its reading list.
-    \end{itemize}
-
-    If we consider a proper vertex coloring of \(G_d\), every color class will give us an independent set of
-    transactions which can be executed concurrently. To achieve the highest parallelization, we need to color \(G_d\)
-    with minimum number of colors. Thus, the \emph{chromatic number} of the memory dependency graph shows how good a
-    transaction set could be run concurrently.
-
-    Graph coloring is computationally NP-hard. However, in our use case we don't need to necessarily find an optimal
-    solution. An approximate greedy algorithm will perform well enough in most circumstances.
-
-    After constructing the memory dependency graph, we can use it to construct the
-    \emph{execution DAG} of transactions. The execution DAG of transaction set \(T\) is a directed acyclic
-    graph \(G_e = (V_e,E_e)\) which has the \emph{execution invariance} property:
-    \begin{itemize}
-        \item Every vertex in \(V_e\) corresponds to a transaction in \(T\) and vice versa.
-        \item Executing the transactions of \(T\) in any order that \emph{respects} \(G_e\) will result in
-        the same AVM state.
-        \begin{itemize}
-            \item An ordering of transactions of \(T\) respects \(G_e\) if for every directed edge \((u,v) \in E_e\)
-            the transaction \(u\) comes before the transaction \(v\) in the ordering.
-        \end{itemize}
-    \end{itemize}
-
-    Having the execution DAG of a set of transactions, using Algorithm~\ref{alg:exec_dag}, we can apply the transaction
-    set to the AVM state concurrently, using multiple processor, while we can be sure that the resulted AVM state will
-    always be the same no matter how many processor we have used.
-
-    %##\includegraphics[width=17cm]{../img/Alg1s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Ready}{$R_e$}\SetKwData{V}{$v_f$}\SetKwData{Graph}{$G_e$}\SetKwData{Vertices}{$V$}\SetKwData
-        {Txns}{$T$}
-        \KwData{The execution dag $\Graph = (\Vertices,E)$ of transaction set \Txns}
-        \KwResult{The state of the AVM after applying \Txns with any ordering respecting \Graph}
-        \BlankLine
-        \Ready $\gets$ the set of all vertices of \Vertices with in degree 0\;
-        \While{$\Vertices \neq \varnothing$}
-        {
-            wait until a new free processor is available\;
-            \If{the execution of a transaction was finished}
-            {
-                remove the vertex of the finished transaction \V from \Graph\;
-                \For{each vertex $u \in Adj[\V]$}
-                {
-                    \If{$u$ has zero in degree}
-                    {
-                        $\Ready \gets \Ready \cup u$\;
-                    }
-                }
-            }
-            \If{$\Ready \neq \varnothing$}
-            {
-                remove a vertex from \Ready and assign it to a processor\;
-            }
-        }
-        \caption{executing DAG transactions}\label{alg:exec_dag}
-    \end{algorithm}
-
-    By replacing every undirected edge of a memory dependency graph with a directed edge in such a way that the
-    resulted graph has no cycles, we will obtain a valid execution DAG. Thus, from a memory dependency graph different
-    execution DAGs can be constructed with different levels of parallelization ability.
-
-    If we assume that we have unlimited number of processors and all transactions take equal time for executing, it
-    can be shown that by providing a minimal graph coloring to Algorithm~\ref{alg:gen_dag} as input, the resulted
-    DAG will be optimal, in the sense that it results in the minimum overall execution time.
-
-    %##\includegraphics[width=17cm]{../img/Alg2s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Txns}{$T$}\SetKwData{Gd}{$G_d=(V_d,E_d)$}
-        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{The memory dependency graph \Gd of transaction set \Txns\\A proper coloring of $G_d$}
-        \Output{An execution dag $G_e=(V_e,E_e)$ for the transaction set \Txns}
-        \BlankLine
-        $V_e \gets V_d$\;
-        $E_e \gets \varnothing$\;
-        define a total order on colors of $G_d$\;
-        \For{each edge $\{u,v\} \in E_d$}
-        {
-            \eIf{$color[u] < color[v]$}
-            {
-                $E_e \gets E_e \cup (u,v)$\;
-            }{
-                $E_e \gets E_e \cup (v,u)$\;
-            }
-        }
-        \caption{Constructing an execution DAG}\label{alg:gen_dag}
-    \end{algorithm}
-
-    The block proposer is responsible for proposing an efficient execution DAG alongside his proposed block. This
-    execution DAG will determine the ordering of block transactions and help validators to validate transactions in
-    parallel. Since with better parallelization a block can contain more transactions, a proposer is incentivized enough
-    to find a good execution DAG for transactions.
-
-    \subsection{Memory Spooling}\label{subsec:spooling}
-
-    When two transactions are dependant and they are connected with an edge \((u,v)\) in the execution DAG,
-    the transaction \(u\) needs to be run before the transaction \(v\). However, if \(v\) does not read any
-    memory locations that \(u\) modifies, we can run \(u\) and \(v\) in parallel. We just need to make sure
-    \(u\) does not see any changes \(v\) is making in AVM memory. This can be done by appropriate versioning
-    of the memory locations which is shared between \(u\) and \(v\). We call this method \emph{memory spooling}.
-    After enabling memory spooling between two transactions the edge connecting them can be safely removed from the
-    execution DAG\@.
-
-    \subsection{Concurrent Counters}\label{subsec:concurrent-counters}
-
-    We know that in Argennon every transaction needs to transfer its proposed fee to the \texttt{feeSink} accounts
-    first. This essentially makes every transaction a reader and a writer of the memory locations which store the
-    balance record of the \texttt{feeSink} accounts. As a result, all transactions in Argennon will be dependant and
-    parallelism will be completely impossible. Actually, any account that is highly active, for example the account
-    of an exchange or a payment processor, could become a concurrency bottleneck in our system which makes all
-    transactions interacting with them dependant.
-
-    This problem can be easily solved by using a concurrent counter for storing the balance record of this type of
-    accounts. A concurrent counter is a data structure which improves concurrency by using multiple memory locations for
-    storing a single counter. The value of the concurrent counter is equal to the sum of its sub counters and it can
-    be incremented or decremented by incrementing/decrementing any of the sub counters. This way, a concurrent
-    counter trades concurrency with memory usage.
-
-    Algorithm~\ref{alg:CC} implements a concurrent counter which returns an error when the value of the counter
-    becomes negative.
-
-    %##\includegraphics[width=17cm]{../img/Alg3s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{CC}{Counter}
-        \SetKwFunction{Inc}{Increment}\SetKwFunction{Dec}{Decrement}\SetKwFunction{AtomInc}{AtomicIncrement}
-        \SetKwFunction{AtomDec}{AtomicDecrement}\SetKwFunction{AtomSet}{AtomicSet}\SetKwFunction{Get}{GetValue}
-        \SetKwFunction{Acquire}{Lock.Acquire}\SetKwFunction{Release}{Lock.Release}
-        \SetKwProg{Fn}{Function}{}{}
-        \Fn{\Get{\CC}}
-        {
-            $s \gets 0$\;
-            \Acquire{}\;
-            \For{$i \gets 0$ \KwTo $\CC.size - 1$}
-            {
-                $s \gets s + \CC.cell[i]$\;
-            }
-            \Release{}\;
-            \KwRet{s}\;
-        }
-        \BlankLine
-        \Fn{\Inc{\CC, value, seed}}
-        {
-            $i \gets seed \bmod \CC.size$\;
-            \AtomInc{$\CC.cell[i]$, value}\;
-        }
-        \BlankLine
-        \Fn{\Dec{\CC, value, seed, attempt}}
-        {
-            \If {attempt = \CC.size}
-            {
-                restore \CC by adding back the subtracted value\;
-                \KwRet{Error}\;
-            }
-            $i \gets seed \bmod \CC.size$\;
-            $i \gets (i + attempt) \bmod \CC.size$\;
-            \eIf {$\CC.cell[i] \geq value$}
-            {
-                \AtomDec{$\CC.cell[i]$, value}\;
-            }{
-                $r \gets value - \CC.cell[i]$\;
-                \AtomSet{$\CC.cell[i]$, $0$}\;
-                \Dec{\CC, r, seed, $attempt + 1$}\;
-            }
-        }
-        \caption{Concurrent counter}\label{alg:CC}
-    \end{algorithm}
-
-    It should be noted that in a blockchain application we don't have concurrent threads and therefore we don't need
-    atomic functions. For usage in a smart contract, the atomic functions of this pseudocode can be implemented like
-    normal functions.
-
-    Concurrent counter data structure is a part of the AVM standard library, and any smart contract can use this data
-    structure for storing the balance record of highly active accounts.
-
-    \subsection{Memory Chunks}\label{subsec:memory-chunks}
-
-    In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{chunks}.
-    Each memory chunk can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
-    consensus on new values of the commitment of any chunk can be achieved by a different voting committee.
-
-    If a transaction does not modify a memory chunk and in the transaction ordering of the block it comes after
-    any transaction which modifies that chunk, then the execution of that transaction is not needed for calculating
-    the new commitment of the chunk. Consequently, the voting committee of that memory chunk can safely ignore such a
-    transaction. The execution DAG of transactions can be used for finding and pruning these transactions as
-    we see in Algorithm~\ref{alg:prune_dag}.
-
-    %##\includegraphics[width=17cm]{../img/Alg4s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{V}{$v_0$}\SetKwData{Graph}{$G_e$}\SetKwData{Chunk}{$C_k$}\SetKwData{Txns}{$T$}
-        \SetKwFunction{RPrune}{RecursivePrune}
-        \SetKwProg{Fn}{Function}{}{}
-        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{An execution dag \Graph and a memory chunk \Chunk}
-        \Output{The pruned execution dag based on \Chunk}
-        \BlankLine
-        \For{each vertex \V with out degree $0$}
-        {
-            \RPrune{\V}\;
-        }
-        \BlankLine
-        \Fn{\RPrune{\V}}
-        {
-            \If{$Adj[\V] = \varnothing$ {\bf and} \V does not modify \Chunk}
-            {
-                remove \V from \Graph\;
-                \For{each $u$ such that edge $(u,\V)$ was in \Graph}
-                {
-                    \RPrune{u}\;
-                }
-            }
-        }
-        \caption{Pruning an execution DAG}\label{alg:prune_dag}
-    \end{algorithm}
-
-    If we choose chunks in a way that most transactions only modify memory locations of one chunk,
-    likely many transactions of a block only need to be validated by one voting committee and can be validated in
-    parallel by different committees.
-
-    Because the voting committees are selected by random sampling, by choosing large enough samples we can make sure
-    that having multiple voting committees will not change the security properties of the Argennon agreement protocol.
-
-    \section{Smart Contract Oracle}\label{sec:smart-contract-oracle}
-
-    A smart contract oracle is a full AVM emulator that keeps a full local copy of AVM memory and can emulate AVM
-    execution without accessing a ZK-EDB. Smart contract oracles can be used for reporting useful information about
-    \texttt{avmCall} transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
-    and so on.
-
-\end{document}

### 2021-04-26 tex/sections/persistence.tex
Change project structure
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-03-29 tex/avm.tex
@@
 %! Author = aybehrouz
-%! Date = 2/28/21
-
-% Preamble
-\documentclass[a4paper]{report}
-
-% Packages
-%\usepackage{amsmath}
-%\usepackage[a4paper]{geometry}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{amsfonts}
-\usepackage[ruled,boxed]{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{tcolorbox}
-\lstset{
-%  basicstyle=\ttfamily,
-    mathescape
-}
-%\binoppenalty=99999
-%\relpenalty=99999
-\newcommand{\note}[1] {
-    \begin{tcolorbox}[colframe=white,colback=white]
-        \emph{#1}
-        %\emph{#1}\medskip
-    \end{tcolorbox}
-}
-
-\title{The Argennon Virtual Machine}
-\author{aybehrouz}
-\date{January 2021}
-
-% Document
-\begin{document}
-    \maketitle
-    \tableofcontents
-
-    \chapter{Specification}\label{ch:specification}
-
-    \section{Introduction}\label{sec:introduction}
-
-    The Argennon Virtual Machine (AVM) is an abstract computing machine for executing Argennon's smart contracts. It
-    is designed in a way that it could be efficiently implemented in either hardware or software.
-
-    \section{The Structure of the Argennon Virtual Machine}\label{sec:the-structure-of-the-algorand-virtual-machine}
-
-    ...
-
-    \subsection{Data Types}\label{subsec:data-types}
-
-    The Argennon Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
-    and does not have to be done by the Argennon Virtual Machine itself.
-
-    The instruction set of the Argennon Virtual Machine distinguishes its operand types using instructions intended to
-    operate on values of specific types. For instance, \texttt{iadd} assumes that its operands are two 64-bit integers.
-
-    \subsection{The \texttt{PC} Register}\label{subsec:the-pc-register}
-
-    ...
-
-    \subsection{Call Stack}\label{subsec:call-stack}
-
-    A call stack contains the information that is needed for restoring the state of the invoker of a method.
-
-    \subsection{Memory Unit}\label{subsec:memory-unit}
-
-    The Argennon Virtual Machine has a byte addressable memory which is divided into separate segments. Every segment
-    belongs to a smart contract that has a unique \texttt{applicationID}. The AVM always has a single working memory
-    segment and memory locations outside its current working segment can not be accessed. The only instructions which
-    can change the current working segment are \texttt{invoke\_external}, \texttt{athrow} and return instructions.
-
-    Every smart contract has its own memory segment. Hence, there is no way for a smart contract to access another
-    smart contract's memory. Interaction between smart contracts is done using \texttt{invoke\_external} instruction,
-    and a smart contract can invoke methods of another smart contract by this instruction.
-
-    A memory segment consists of five data areas:
-    \begin{itemize}
-        \item Code Area
-        \item Constant Area
-        \item Local Frame
-        \item Operand Stack
-        \item Heap
-    \end{itemize}
-    All data areas except operand stack, have their own address space which starts from 0. Operand stack is a
-    last-in-first-out (LIFO) stack and is not addressable. Every memory access instruction operates on its specific
-    data areas.
-
-    \subsubsection{Code Area}
-
-    The code area of a segment contains the byte-code of the smart contract which owns that segment. The AVM has no
-    instructions for manipulating the code area. \textbf{Installing, removing and updating smart contracts need to
-    be done externally.}
-
-    \subsubsection{Constant Area}
-
-    The constant area of a segment contains several kinds of constants, ranging from user defined constants to method
-    address tables. A method address table stores method locations in the code area and their access type. The access
-    type of a method can be either \texttt{public} or \texttt{private}. The AVM has no instructions for modifying the
-    constant area.
-
-    \note{Only public methods can be invoked by \texttt{invoke\_external} instruction.}
-
-    \subsubsection{Local Frame}
-
-    A local frame is used to store methods parameters and local variables. A new frame is created each time a method
-    is invoked, and it is destroyed when its method invocation completes, whether the completion is normal or abrupt.
-
-    \subsubsection{Operand Stack}
-
-    Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too. AVM
-    instructions take operands from the operand stack, operate on them, and push the result back onto the operand
-    stack. An operand stack is destroyed when its owner method completes, whether that completion is normal or abrupt.
-
-    \subsubsection{Heap}
-
-    The heap of a segment is a persistent memory area which is divided into pages. Memory locations inside every page
-    have a separate address space that starts from 0, and each page can be referenced by an index that starts from 0. In
-    other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex, offset)}.
-    Different pages of a heap do not need to be equally sized.
-
-    \note{The reason behind this paged design is that the heap is usually persisted using a block device. A heap with
-    a paged structure could expose the underlying block based nature of the persistence layer to the application
-    layer. In this way, the compiler or the programmer could better optimize the code for the persistence layer.}
-
-    \section{Instruction Set Summary}\label{sec:instruction-set-summary}
-
-    An Argennon Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
-    performed, followed by zero or more operands supplying arguments or data that are used by the operation.
-    \textbf{The number and size of the operands are determined solely by the opcode.}
-
-    \subsection{Method Invocation}\label{subsec:method-invocation}
-
-    The Argennon Virtual Machine has three types of method invocation:
-    \begin{itemize}
-        \item \texttt{invoke\_internal} invokes a method from the current running smart contract.
-        \item \texttt{invoke\_external} invokes a \texttt{public} method from another smart contract. It will change
-        the current memory segment to the segment of the invoked smart contract.
-        \item \texttt{invoke\_native} invokes a method that is not hosted by the Argennon virtual machine. By this
-        instruction, high performance native methods of the hosting machine could become available to AVM smart
-        contracts.
-    \end{itemize}
-
-    \note{In the future, we may need to add special instructions for invoking interface and virtual methods...}
-
-    Each time a method is invoked a new local frame and operand stack is created. The Argennon Virtual Machine uses
-    local frames to pass parameters on method invocation. On method invocation, any parameters are passed in
-    consecutive local variables stored in the method's local frame starting from local variable 0. The invoker of a
-    method writes the parameters in the local frame of the invoked method using \texttt{arg} instructions.
-
-    \subsubsection{Exceptions}
-
-    An exception is thrown programmatically using the \texttt{athrow} instruction. Exceptions can also be thrown by
-    various Argennon Virtual Machine instructions if they detect an abnormal condition. Some exceptions are not
-    catchable and will always abort the execution of the smart contract.
-
-    \subsubsection{Method Invocation Completion}
-
-    A method invocation completes normally if that invocation does not cause an exception to be thrown, either
-    directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
-    method completes normally, then a value may be returned to the invoking method. This occurs when the invoked
-    method executes one of the return instructions, the choice of which must be appropriate for the type of the value
-    being returned (if any). Execution then continues normally in the invoking method's local frame with the returned
-    value (if any) pushed onto the operand stack.
-
-    A method invocation completes abruptly if an exceptions is thrown and is not caught by the current method. A
-    method invocation that completes abruptly never returns a value to its invoker.
-
-    When a method completes, whether normally or abruptly, the call stack is used to restore the state of the invoker,
-    including its local frame and operand stack, with the \texttt{PC} register appropriately restored and incremented
-    to skip past the method invocation instruction. If the invoker was another smart contract, i.e.\ the invocation
-    was made by an \texttt{invoke\_external} instruction, the current memory segment will be changed to the invoker's
-    segment.
-
-    A thrown exception causes methods in the call stack to complete \textbf{abruptly} one by one, as long as the
-    \texttt{PC} register is not pointing to a \texttt{catch} instruction. The \texttt{catch} instruction acts like a
-    branch instruction that branches only if an exception is caught. \textbf{When an external method invocation
-    completes abruptly, before changing the current segment, all changes made to the heap area by that method
-    call, including changes made to other segments, will be rolled back.} So, when a method call completes abruptly,
-    that method call essentially has no effect on the AVM state.
-
-    \note{By using the \texttt{athrow} instruction properly, a programmer can make any method act like an atomic
-    operation.}
-
-    \subsection{Authorizing Operations}\label{subsec:authorizing-operations}
-
-    In blockchain applications, we usually need to authorize certain operations. For example, for sending an asset
-    from a user to another user, first we need to make sure that the sender has authorized this operation. The
-    Argennon virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
-    cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
-    passing signatures as parameters to methods, a programmer can implement the required logic for authorizing any
-    operation.
-
-    \note{The Argennon virtual machine has no instructions for issuing cryptographic signatures.}
-
-    In addition to signatures, a method can verify its invoker by using \texttt{get\_parent} instruction. This
-    instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
-    smart contract in the call stack. In other words, it returns the \texttt{applicationID} of the smart contract that
-    has invoked the current smart contract. (if any)
-
-    \subsection{Heap Allocation Instructions}\label{subsec:heap-allocation-instructions}
-
-    ...
-
-    \section{The AVM Standard Library}\label{sec:the-avm-standard-library}
-
-    ...
-
-    \chapter{Implementation}\label{ch:implementation}
-
-    \section{Persistence}\label{sec:persistence}
 
@@
 \note{Unlike heap pages, the AVM is not aware of different blocks of code area.}
-
-    \section{Transactions}\label{sec:transactions}
-
-    Argennon has four types of transaction:
-
-    \begin{itemize}
-        \item \texttt{avmCall} essentially is an \texttt{invoke\_external} instruction that invokes a method from an
-        AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
-        assets, including ARGs, is done by these transactions.
-        \item \texttt{installApp} installs an AVM smart contract and determines the update policy of the smart
-        contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
-        on.
-        \item \texttt{unInstallApp} removes an AVM smart contract.
-        \item \texttt{updateApp} updates an AVM smart contract.
-    \end{itemize}
-
-    All types of Argennon transactions contain an \texttt{invoke\_external} instruction which calls a special method
-    from ARG smart contract that transfers the proposed fee of the transaction in ARGs from a sender account to the
-    fee sink accounts.
-
-    Every transaction is required to exactly specify what heap locations or code area addresses it will access. This
-    enables validators to start retrieving the required memory blocks from available ZK-EDB servers as soon as they
-    see a transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to
-    access a memory location that is not included in its access lists, will be rejected. Users could use smart contract
-    oracles to predict the list of memory blocks their transactions need. See Section~\ref{sec:smart-contract-oracle}
-    for more details.
-
-    \section{Blockchain}\label{sec:blockchain}
-
-    Every block of the Argennon blockchain corresponds to a set of transactions. We store the commitment of this
-    transaction set in every block, but we don't keep the set itself. To be able to detect replay attacks, we require
-    every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
-    and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
-    must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
-    a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
-    if a signature of the same user with a bigger or equal nonce is already used
-    (i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
-    nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
-
-    To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. When the
-    difference between \texttt{issuance} component of this nonce and the current round becomes bigger than the
-    maximum allowed lifetime of a signature, this information can be safely deleted. \textbf{As a result, we will not
-    have the problem of "un-removable empty accounts" like Ethereum.}
-
-    The only information that Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
-    blockchain. Every block of the Argennon blockchain contains the following information:
-
-    \begin{center}
-        \begin{tabular}{||c||}
-            \hline
-            Block \\ [0.5ex]
-            \hline\hline
-            commitment to the ZK-EDB storing heap pages \\ [0.7ex]
-            commitment to the ZK-EDB storing code areas \\ [0.7ex]
-            commitment to the set of transactions       \\ [0.7ex]
-            previous block hash                         \\ [0.7ex]
-            random seed                                 \\ [0.7ex]
-            \hline
-        \end{tabular}
-    \end{center}
-
-    For confirming a new block, nodes that are not validators only need to verify the block certificate. For
-    verifying a block certificate, a node needs to know the ARG balances of validators, but it doesn't need to
-    emulate the AVM execution.
-
-    On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
-    execution of the Argennon virtual machine. To do so, first they retrieve all heap pages and code area blocks they
-    need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
-    transactions included in the new block. This will modify some pages of the AVM memory, so they update the ZK-EDB
-    commitments based on the modified pages and verify the commitments included in the new block. Validators also
-    calculate and verify the commitment to the new block's transaction set.
-
-    \note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDB servers will receive the new
-    block, and they will update their database by emulating the AVM execution.}
-
-    \section{Consensus}\label{sec:consensus}
-
-    \note{It will be a randomized proof of stake which is under development....}
-
-    \subsection{Estimating A User's Stake}\label{subsec:estimating-a-user's-stake}
-
-    In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
-    of stake the user has in the system. Conventionally in these systems, for estimating a user's stake, we use the
-    amount of native system tokens the user is holding. Unfortunately, one problem with this approach is that a
-    strong attacker may be able to obtain a considerable amount of system tokens, for example by borrowing from a
-    DEFI application, and use this stake to attack the system.
-
-    To mitigate this problem, for calculating a user's stake at time step \(t\), instead of using the raw ARG
-    balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
-    ARG balance:
-    \[
-        S_{u,t} = \min (B_{u,t}, Trust_{u,t})
-    \]
-    Where:
-    \begin{itemize}
-        \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
-        \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
-        \item \(Trust_{u,t}\) is an estimated trust value for user \(u\) at time step \(t\).
-    \end{itemize}
-
-    The agreement protocol, at time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
-    number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
-    exponential moving average of the ARG balance of user \(u\) at time step \(t\).
-
-    In our system a user who held ARGs and participated in the consensus for a long time is more trusted
-    than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
-    amount of ARGs, also needs to hold them for a long period of time before being able to attack the system.
-
-    For calculating the exponential moving average of a user's balance at time step \(t\), we can use the following
-    recursive formula:
-    \[
-        M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
-    \]
-    Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
-    degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
-
-    Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
-    \(M_{u,t}\): (In the following equations the \(u\) subscript is dropped for simplicity)
-    \[
-        M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B
-    \]
-    Where:
-    \[
-        B = B_{k+1} = B_{k+2} = \dots = B_{t}
-    \]
-    We know that when \(|nx| \ll 1\) we can use the binomial approximation \({(1 + x)^n \approx 1 + nx}\). So, we can
-    further simplify this formula:
-    \[
-        M_{t} = M_{k} + (t - k) \alpha (B - M_{k})
-    \]
-
-    For choosing the value of \(\alpha\) we can consider the number of time steps that the trust value of a user needs
-    for reaching a specified fraction of his account balance. We know that for large \(n\) and \(|x| < 1\) we have
-    \((1 + x)^n \approx e^{nx}\), so by letting \(M_{u,k} = 0\) and \(n = t - k\) we can write:
-    \[
-        \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{B}\right)}{n}
-    \]
-    The value of \(\alpha\) for a desired configuration can be calculated by this equation. For instance, we could
-    calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8B\) and \(n\) equals to the
-    number of time steps of 10 years.
-
-    In our system a newly created account will not have voting power for some time, no matter how high its
-    balance is. While this is a desirable property, in case a large proportion of total system tokens are
-    transferred to newly created accounts, it can result in too much voting power for older accounts. This may decrease
-    the degree of decentralization in our system.
-
-    However, this situation is easily detectable by comparing the total stake of the system with the total balance of
-    users. If after confirming a block the total stake of the system goes too low and we have:
-    \[
-        \sum_{u}S_{u,t} < \gamma \sum_{u}B_{u,t}
-    \]
-    The protocol will perform a \emph{time shift} in the system: the time step of the system
-    will be incremented for \(m\) steps while no blocks will be confirmed. This will increase the value of \(M_{u,t}\)
-    for new accounts with a non-zero balance, giving them more influence in the agreement protocol.
-
-    For calculating the value of \(m\) which determines the amount of time shift in the system, we should note that when
-    \(B_{u,t} = B_{u, t-1} = B_u\), we can derive a simple recursive rule for the stake of a user:
-    \[
-        S_{u,t} = (1 - \alpha) S_{u,t-1} + \alpha B_u
-    \]
-    Therefore, we have:
-    \[
-        \sum_{u}S_{u,t} = (1 - \alpha) \sum_{u}S_{u,t - 1} + \alpha \sum_{u}B_u
-    \]
-    This equation shows that when the balance of users is not changing over time the total stake of the system is the
-    exponential average of the total ARGs of the system. Consequently, when we shift the time for \(m\) steps, we can
-    calculate the new total stake of the system from the following equation:
-
-    \[
-        \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
-    \]
-    Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
-    we can obtain \(m\) from the following formula, assuming \(\alpha\) is small enough:
-    \[
-        m = \frac{1}{\alpha} \ln \left(\frac{1 - \gamma}{1 - \lambda}\right)
-    \]
-
-    \section{Incentive mechanism}\label{sec:incentive-mechanism}
-
-    \subsection{Transaction Fee}\label{subsec:transaction-fee}
-
-    Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
-    special method from ARG smart contract. This method will transfer the proposed fee of the transaction in ARGs
-    from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
-    execution fees and \texttt{dbFeeSink} collects fees for ZK-EDB servers. The Protocol decides how to distribute the
-    transaction fee between these two fee sink accounts.
-
-    When a block is added to the blockchain, the proposer of that block will receive a share of the block fees.
-    Consequently, a block proposer is always incentivized to include more transactions in his block. However, if he
-    puts too many transactions in his block and the validation of the block becomes too difficult, some validators
-    may not be able to validate all transactions on time. If a validator can not validate a block in the required
-    time, he will consider the block invalid. So, when a proposed block contains too many transactions, the network
-    may reach consensus on another block, and the proposer of that block will not receive any fees. As a result, a
-    proposer is incentivized to use network transaction capacity optimally.
-
-    On the other hand, we believe that the proposer does not have enough incentives for optimizing the storage size
-    of the transaction set. Therefore, we require that \textbf{the size of the transaction set of every block in
-    bytes be lower than a certain threshold.}
-
-    Validators need to spend resources for validating transactions. When a validator starts the emulation of the AVM
-    to validate a transaction, solely from the code he can't predict the time the execution will finish. This will
-    give an adversary an opportunity to attack the network by broadcasting transactions that never ends. Since,
-    validators can not finish the execution of these transactions, the network will not be able to charge the
-    attacker any fees, and he would be able to waste validators resources for free.
-
-    To mitigate this problem, we require that every transaction specify a cap for all the resources it needs. This
-    will include memory, network and processor related resources. Also, the protocol defines an execution cost for
-    every AVM instruction reflecting the amount of resources its emulation needs. This will define a standard way for
-    measuring the execution cost of any \texttt{avmCall} transaction. Every \texttt{avmCall} transaction is required
-    to specify a maximum execution cost. If during emulation it reaches this maximum cost, the transaction will be
-    considered failed and the network can receive the proposed fee of that transaction.
-
-    Every \texttt{avmCall} transaction is required to provide the following information as an upper bound for the
-    resources it needs:
-
-    \begin{itemize}
-        \item Execution cost
-        \item A list of heap/code-area locations for reading
-        \item A list of heap locations for writing
-        \item A list of heap pages it will deallocate
-        \item Number and size of heap pages it will allocate
-    \end{itemize}
-
-    If a transaction tries to violate any of these predefined limitations, for example, if it tries to read a memory
-    location that is not included in its reading list, it will be considered failed and the network can receive the
-    proposed fee of that transaction.
-
-    \note{A transaction always pays all of its proposed fee, no matter how much of its predefined resources were not
-    used in the final emulation.}
-
-    \subsection{Incentives for ZK-EDB Servers}\label{subsec:zk-edb-servers}
-
-    The incentive mechanism for ZK-EDB servers should have the following properties:
-
-    \begin{itemize}
-        \item It incentivizes storing all memory blocks, whether a heap page or a code area block, and not only those
-        which are used more frequently.
-        \item It incentivizes ZK-EDB servers to actively provide the required memory blocks for validators.
-        \item Making more accounts will not provide any advantages for a ZK-EDB server.
-    \end{itemize}
-
-    For our incentive mechanism, we require that every time a validator receives a memory block from a ZK-EDB, after
-    validating the data, he give a receipt to the ZK-EDB. In this receipt the validator signs the following information:
-
-    \begin{itemize}
-        \item \texttt{ownerAddr}: the ARG address of the ZK-EDB\@.
-        \item \texttt{receivedBlockID}: the ID of the received memory block.
-        \item \texttt{round}: the current round number.
-    \end{itemize}
-
-    \note{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}
-
-    To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ARGs from
-    \texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
-    between all \emph{winning tickets} of the lottery.
-
-    \note{One ZK-EDB server could own multiple winning tickets in a round.}
-
-    To run this lottery, every round, based on the current block seed, a collection of \emph{valid} receipts will be
-    selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in round \(r\) if:
-
-    \begin{itemize}
-        \item The signer was a validator in the round \(r - 1\) and voted for the agreed-upon block.
-        \item The data block in the receipt was needed for validating the \textbf{previous} block.
-        \item The receipt round number is \(r - 1\).
-        \item The signer did not sign a receipt for the same data block for two different ZK-EDBs in the previous round.
-    \end{itemize}
-    For selecting the winning receipts we could use a random generator:
-    \begin{verbatim}
-IF random(seed|validatorPK|receivedBlockID) < winProbability THEN
-    the receipt issued by validatorPK for receivedBlockID is a winner
-    \end{verbatim}
-    \begin{itemize}
-        \item \texttt{random()} produces uniform random numbers between 0 and 1, using its input argument as a seed.
-        \item \texttt{validatorPK} is the public key of the signer of the receipt.
-        \item \texttt{receivedBlockID} is the ID of the memory block that the receipt was issued for.
-        \item \texttt{winProbability} is the probability of winning in every round.
-        \item \texttt{seed} is the current block seed.
-        \item \texttt{|} is a concatenation operator.
-    \end{itemize}
-
-    \note{The winners of the lottery were validators one round before the lottery round.}
-
-    Also, based on the current block seed, a random memory block, whether a heap page or a code area block, is
-    selected as the challenge of the round. A ZK-EDB that owns a winning receipt needs to broadcast a \emph{winning
-    ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
-    challenge. Solving a round challenge requires the content of the memory block which was selected as the round
-    challenge. This will encourage ZK-EDBs to store all memory blocks.
-
-    A possible choice for the challenge solution could be the cryptographic hash of the content of the challenge
-    memory block combined with the ZK-EDB ARG address: \texttt{hash(challenge.content|ownerAddr)}
-
-    The winning tickets of the lottery of round \(r\) need to be included in the block of the round \(r\),
-    otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
-    \(r\) will be done in the round \(r + 1\). This way, \textbf{the content of the challenge memory block could be
-    kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.
-
-    \subsection{Memory Allocation and De-allocation Fee}\label{subsec:memory-allocation-and-de-allocation}
-
-    Every \(k\) round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
-    allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
-    address of the smart contract.
-
-    To determine the price of AVM memory, Every \(k\) round, the protocol calculates \texttt{dbFee} and
-    \texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
-    \texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
-    the protocol considers the total size of all the memory pages that were accessed for either reading or writing
-    during a time period. These two values will be calculated for the last \(k\) rounds and the price per byte of
-    AVM memory will be a linear function of \texttt{dbFee/memTraffic}
-
-    When a smart contract executes a heap de-allocation instruction, the protocol will refund the cost of
-    de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
-    calculates the refunded amount based on the average price the smart contract had paid for that allocated memory.
-    This will prevent smart contracts from profit taking by trading memory with the protocol.
-
-    \section{Concurrency}\label{sec:concurrency}
-
-    \subsection{Memory Dependency Graph}\label{subsec:memory-dependency-graph}
-
-    Every block of the Argennon blockchain contains a list of transactions. This list is an ordered list and the
-    effect of its contained transactions must be applied to the AVM state sequentially as they appear in the ordered
-    list. This ordering is solely chosen by the block proposer, and users should not have any assumptions about
-    the ordering of transactions in a block.
-
-    The fact that block transactions constitute a sequential list, does not mean they can not be executed and applied
-    to the AVM state concurrently. Many transactions are actually independent and the order of their execution does not
-    matter. These transactions can be safely validated in parallel by validators.
-
-    A transaction can change the AVM state by modifying either the code area or the AVM heap. In Argennon, all
-    transactions declare the list of memory locations they want to read or write. This will enable us to determine the
-    independent sets of transactions which can be executed in parallel. To do so, we define the \emph{memory dependency
-    graph} \(G_d\) as follows:
-
-    \begin{itemize}
-        \item \(G_d\) is an undirected graph.
-        \item Every vertex in \(G_d\) corresponds to a transaction and vice versa.
-        \item Vertices \(u\) and \(v\) are adjacent in \(G_d\) if and only if \(u\) has a memory location \(L\) in its
-        writing list and \(v\) has \(L\) in either its writing list or its reading list.
-    \end{itemize}
-
-    If we consider a proper vertex coloring of \(G_d\), every color class will give us an independent set of
-    transactions which can be executed concurrently. To achieve the highest parallelization, we need to color \(G_d\)
-    with minimum number of colors. Thus, the \emph{chromatic number} of the memory dependency graph shows how good a
-    transaction set could be run concurrently.
-
-    Graph coloring is computationally NP-hard. However, in our use case we don't need to necessarily find an optimal
-    solution. An approximate greedy algorithm will perform well enough in most circumstances.
-
-    After constructing the memory dependency graph, we can use it to construct the
-    \emph{execution DAG} of transactions. The execution DAG of transaction set \(T\) is a directed acyclic
-    graph \(G_e = (V_e,E_e)\) which has the \emph{execution invariance} property:
-    \begin{itemize}
-        \item Every vertex in \(V_e\) corresponds to a transaction in \(T\) and vice versa.
-        \item Executing the transactions of \(T\) in any order that \emph{respects} \(G_e\) will result in
-        the same AVM state.
-        \begin{itemize}
-            \item An ordering of transactions of \(T\) respects \(G_e\) if for every directed edge \((u,v) \in E_e\)
-            the transaction \(u\) comes before the transaction \(v\) in the ordering.
-        \end{itemize}
-    \end{itemize}
-
-    Having the execution DAG of a set of transactions, using Algorithm~\ref{alg:exec_dag}, we can apply the transaction
-    set to the AVM state concurrently, using multiple processor, while we can be sure that the resulted AVM state will
-    always be the same no matter how many processor we have used.
-
-    %##\includegraphics[width=17cm]{../img/Alg1s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Ready}{$R_e$}\SetKwData{V}{$v_f$}\SetKwData{Graph}{$G_e$}\SetKwData{Vertices}{$V$}\SetKwData
-        {Txns}{$T$}
-        \KwData{The execution dag $\Graph = (\Vertices,E)$ of transaction set \Txns}
-        \KwResult{The state of the AVM after applying \Txns with any ordering respecting \Graph}
-        \BlankLine
-        \Ready $\gets$ the set of all vertices of \Vertices with in degree 0\;
-        \While{$\Vertices \neq \varnothing$}
-        {
-            wait until a new free processor is available\;
-            \If{the execution of a transaction was finished}
-            {
-                remove the vertex of the finished transaction \V from \Graph\;
-                \For{each vertex $u \in Adj[\V]$}
-                {
-                    \If{$u$ has zero in degree}
-                    {
-                        $\Ready \gets \Ready \cup u$\;
-                    }
-                }
-            }
-            \If{$\Ready \neq \varnothing$}
-            {
-                remove a vertex from \Ready and assign it to a processor\;
-            }
-        }
-        \caption{executing DAG transactions}\label{alg:exec_dag}
-    \end{algorithm}
-
-    By replacing every undirected edge of a memory dependency graph with a directed edge in such a way that the
-    resulted graph has no cycles, we will obtain a valid execution DAG. Thus, from a memory dependency graph different
-    execution DAGs can be constructed with different levels of parallelization ability.
-
-    If we assume that we have unlimited number of processors and all transactions take equal time for executing, it
-    can be shown that by providing a minimal graph coloring to Algorithm~\ref{alg:gen_dag} as input, the resulted
-    DAG will be optimal, in the sense that it results in the minimum overall execution time.
-
-    %##\includegraphics[width=17cm]{../img/Alg2s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Txns}{$T$}\SetKwData{Gd}{$G_d=(V_d,E_d)$}
-        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{The memory dependency graph \Gd of transaction set \Txns\\A proper coloring of $G_d$}
-        \Output{An execution dag $G_e=(V_e,E_e)$ for the transaction set \Txns}
-        \BlankLine
-        $V_e \gets V_d$\;
-        $E_e \gets \varnothing$\;
-        define a total order on colors of $G_d$\;
-        \For{each edge $\{u,v\} \in E_d$}
-        {
-            \eIf{$color[u] < color[v]$}
-            {
-                $E_e \gets E_e \cup (u,v)$\;
-            }{
-                $E_e \gets E_e \cup (v,u)$\;
-            }
-        }
-        \caption{Constructing an execution DAG}\label{alg:gen_dag}
-    \end{algorithm}
-
-    The block proposer is responsible for proposing an efficient execution DAG alongside his proposed block. This
-    execution DAG will determine the ordering of block transactions and help validators to validate transactions in
-    parallel. Since with better parallelization a block can contain more transactions, a proposer is incentivized enough
-    to find a good execution DAG for transactions.
-
-    \subsection{Memory Spooling}\label{subsec:spooling}
-
-    When two transactions are dependant and they are connected with an edge \((u,v)\) in the execution DAG,
-    the transaction \(u\) needs to be run before the transaction \(v\). However, if \(v\) does not read any
-    memory locations that \(u\) modifies, we can run \(u\) and \(v\) in parallel. We just need to make sure
-    \(u\) does not see any changes \(v\) is making in AVM memory. This can be done by appropriate versioning
-    of the memory locations which is shared between \(u\) and \(v\). We call this method \emph{memory spooling}.
-    After enabling memory spooling between two transactions the edge connecting them can be safely removed from the
-    execution DAG\@.
-
-    \subsection{Concurrent Counters}\label{subsec:concurrent-counters}
-
-    We know that in Argennon every transaction needs to transfer its proposed fee to the \texttt{feeSink} accounts
-    first. This essentially makes every transaction a reader and a writer of the memory locations which store the
-    balance record of the \texttt{feeSink} accounts. As a result, all transactions in Argennon will be dependant and
-    parallelism will be completely impossible. Actually, any account that is highly active, for example the account
-    of an exchange or a payment processor, could become a concurrency bottleneck in our system which makes all
-    transactions interacting with them dependant.
-
-    This problem can be easily solved by using a concurrent counter for storing the balance record of this type of
-    accounts. A concurrent counter is a data structure which improves concurrency by using multiple memory locations for
-    storing a single counter. The value of the concurrent counter is equal to the sum of its sub counters and it can
-    be incremented or decremented by incrementing/decrementing any of the sub counters. This way, a concurrent
-    counter trades concurrency with memory usage.
-
-    Algorithm~\ref{alg:CC} implements a concurrent counter which returns an error when the value of the counter
-    becomes negative.
-
-    %##\includegraphics[width=17cm]{../img/Alg3s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{CC}{Counter}
-        \SetKwFunction{Inc}{Increment}\SetKwFunction{Dec}{Decrement}\SetKwFunction{AtomInc}{AtomicIncrement}
-        \SetKwFunction{AtomDec}{AtomicDecrement}\SetKwFunction{AtomSet}{AtomicSet}\SetKwFunction{Get}{GetValue}
-        \SetKwFunction{Acquire}{Lock.Acquire}\SetKwFunction{Release}{Lock.Release}
-        \SetKwProg{Fn}{Function}{}{}
-        \Fn{\Get{\CC}}
-        {
-            $s \gets 0$\;
-            \Acquire{}\;
-            \For{$i \gets 0$ \KwTo $\CC.size - 1$}
-            {
-                $s \gets s + \CC.cell[i]$\;
-            }
-            \Release{}\;
-            \KwRet{s}\;
-        }
-        \BlankLine
-        \Fn{\Inc{\CC, value, seed}}
-        {
-            $i \gets seed \bmod \CC.size$\;
-            \AtomInc{$\CC.cell[i]$, value}\;
-        }
-        \BlankLine
-        \Fn{\Dec{\CC, value, seed, attempt}}
-        {
-            \If {attempt = \CC.size}
-            {
-                restore \CC by adding back the subtracted value\;
-                \KwRet{Error}\;
-            }
-            $i \gets seed \bmod \CC.size$\;
-            $i \gets (i + attempt) \bmod \CC.size$\;
-            \eIf {$\CC.cell[i] \geq value$}
-            {
-                \AtomDec{$\CC.cell[i]$, value}\;
-            }{
-                $r \gets value - \CC.cell[i]$\;
-                \AtomSet{$\CC.cell[i]$, $0$}\;
-                \Dec{\CC, r, seed, $attempt + 1$}\;
-            }
-        }
-        \caption{Concurrent counter}\label{alg:CC}
-    \end{algorithm}
-
-    It should be noted that in a blockchain application we don't have concurrent threads and therefore we don't need
-    atomic functions. For usage in a smart contract, the atomic functions of this pseudocode can be implemented like
-    normal functions.
-
-    Concurrent counter data structure is a part of the AVM standard library, and any smart contract can use this data
-    structure for storing the balance record of highly active accounts.
-
-    \subsection{Memory Chunks}\label{subsec:memory-chunks}
-
-    In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{chunks}.
-    Each memory chunk can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
-    consensus on new values of the commitment of any chunk can be achieved by a different voting committee.
-
-    If a transaction does not modify a memory chunk and in the transaction ordering of the block it comes after
-    any transaction which modifies that chunk, then the execution of that transaction is not needed for calculating
-    the new commitment of the chunk. Consequently, the voting committee of that memory chunk can safely ignore such a
-    transaction. The execution DAG of transactions can be used for finding and pruning these transactions as
-    we see in Algorithm~\ref{alg:prune_dag}.
-
-    %##\includegraphics[width=17cm]{../img/Alg4s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{V}{$v_0$}\SetKwData{Graph}{$G_e$}\SetKwData{Chunk}{$C_k$}\SetKwData{Txns}{$T$}
-        \SetKwFunction{RPrune}{RecursivePrune}
-        \SetKwProg{Fn}{Function}{}{}
-        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{An execution dag \Graph and a memory chunk \Chunk}
-        \Output{The pruned execution dag based on \Chunk}
-        \BlankLine
-        \For{each vertex \V with out degree $0$}
-        {
-            \RPrune{\V}\;
-        }
-        \BlankLine
-        \Fn{\RPrune{\V}}
-        {
-            \If{$Adj[\V] = \varnothing$ {\bf and} \V does not modify \Chunk}
-            {
-                remove \V from \Graph\;
-                \For{each $u$ such that edge $(u,\V)$ was in \Graph}
-                {
-                    \RPrune{u}\;
-                }
-            }
-        }
-        \caption{Pruning an execution DAG}\label{alg:prune_dag}
-    \end{algorithm}
-
-    If we choose chunks in a way that most transactions only modify memory locations of one chunk,
-    likely many transactions of a block only need to be validated by one voting committee and can be validated in
-    parallel by different committees.
-
-    Because the voting committees are selected by random sampling, by choosing large enough samples we can make sure
-    that having multiple voting committees will not change the security properties of the Argennon agreement protocol.
-
-    \section{Smart Contract Oracle}\label{sec:smart-contract-oracle}
-
-    A smart contract oracle is a full AVM emulator that keeps a full local copy of AVM memory and can emulate AVM
-    execution without accessing a ZK-EDB. Smart contract oracles can be used for reporting useful information about
-    \texttt{avmCall} transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
-    and so on.
-
-\end{document}

### 2021-04-27 tex/sections/argon.tex
Update persistence section
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
         public    & yes   & yes     & yes      & yes     & no    \\
-        external  & no    & no      & no       & no      & yes   \\
+        external  & no    & no      & no       & yes     & yes   \\
         \hline

### 2021-04-27 tex/sections/blockchain.tex
Update persistence section
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
         \hline
-        Block \\ [0.5ex]
+        \textbf{Block} \\ [0.5ex]
         \hline\hline
-        commitment to the ZK-EDB storing heap pages \\ [0.7ex]
-        commitment to the ZK-EDB storing code areas \\ [0.7ex]
+        commitment to the staking database \\ [0.7ex]
+        commitment to the method database \\ [0.7ex]
+        commitment to the heap database \\ [0.7ex]
         commitment to the set of transactions       \\ [0.7ex]

### 2021-04-27 tex/sections/persistence.tex
Update persistence section
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-For implementing the persistence layer of the AVM, we assume that we have access to an updatable zero-knowledge
-elementary database (ZK-EDB) with the following properties:
+The Argennon Virtual Machine has two persistent memory areas: \emph{method area}, and \emph{heap}. Method area stores
+method byte codes \footnote{also it stores constant area blocks.}, and heap stores memory chunks. Both of these
+data elements, methods and chunks, can be considered as continuous pieces of byte addressable memory
+which we shall call \emph{objects} throughout this chapter.
 
+In the AVM persistence layer, similar objects are clustered together and constitute a bigger data element which we call a
+\emph{page}.\footnote{we avoided calling them a cluster, because usually a cluster refers to a \emph{set}. AVM object
+clusters are not sets. They are ordered lists, like a page containing an ordered list of words or sentences.}
+A page is an ordered list of an arbitrary number of objects, which their order reflects the order they were added to
+the page:
+\[
+    P = (O_1,O_2,\dots,O_n),\quad i < j \; \Leftrightarrow \; \textrm{$O_i$ was added before $O_j$}
+\]
+
+A page of the AVM storage should contain objects that have very similar access pattern. We expect that when a page
+is needed for validating a block, almost all of its objects are needed. We also prefer that the objects of the page
+are needed for the same access type. In other words, the objects of a page are chosen in a way that
+for validating a block, we usually need to either read all of them or modify\footnote{and probably read.} all of them.
+
+Pages of the AVM storage space are persisted using updatable zero-knowledge elementary databases (ZK-EDB). Argennon
+has three zero-knowledge databases. \emph{Staking} database which stores all the data that is associated with
+the Argennon agreement protocol. \emph{Method} database which stores the AVM method area, and \emph{Heap} database which
+stores the AVM heap. The commitment of these three ZK-EDBs are included in every block of the Argennon blockchain.
+
+We consider the following properties for a ZK-EDB:
 \begin{itemize}
@@
     \item Every state of the database has a commitment \(C\).
-    \item The ZK-EDB has a method \((D, p) = get(x)\), where \(x\) is a key and \(D\) is the associated data
-    with \(x\), and \(p\) is a proof.
-    \item A user can use \(C\) and \(p\) to verify that \(D\) is really associated with \(x\), and \(D\) is not
+    \item The ZK-EDB has a method \((D, \pi) = get(x)\), where \(x\) is a key and \(D\) is the associated data
+    with \(x\), and \(\pi\) is a proof.
+    \item A user can use \(C\) and \(\pi\) to verify that \(D\) is really associated with \(x\), and \(D\) is not
     altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB\@.
-    \item Having \(p\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
+    \item Having \(\pi\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
     associated with \(x\) instead of \(D\).
@@
 
-We use a ZK-EDB for storing the AVM heap. We include the commitment of the current state of this DB in every
-block of the Argennon blockchain, so ZK-EDB servers need not be trusted servers.
+Every page of the AVM storage is stored with an index as its key: \texttt{pageIndex}. The \texttt{pageIndex} is
+required to be smaller than a certain value determined by the
+protocol.\footnote{this facilitates the usage of ZK-EDBs that use vector commitments.}
+Therefore, the AVM clustering algorithm tries to reuse indices and keep the number of used indices as low as
+possible.
 
-Every page of AVM heap will be stored with a key of the form: \texttt{applicationID|pageIndex} (the \texttt{|}
-operator concatenates two numbers). Nodes do not keep a full copy of the AVM heap and for validating block
-certificates or emulating the AVM ( i.e.~validating transactions) they need to connect to a ZK-EDB and retrieve
-the required pages of AVM heap. For better performance, nodes keep a cache of heap pages to
-reduce the amount of ZK-EDB access.
+In Argennon \emph{light} nodes do not keep a full copy of the AVM storage and for emulating the Argennon Virtual Machine
+( i.e.~validating transactions) they need to connect to a ZK-EDB and retrieve the required pages.
+Because light node need to be able to validate block certificates, they usually cache a large part of the staking
+database, and keep their cache updated to make sure they can keep themselves in sync with the blockchain.
 
-We also use a ZK-EDB for storing the code area of each segment, and we include the commitment of this DB in every
-block. Every code area will be divided into blocks and every block will be stored in the DB with
-\texttt{applicationID|blockID} as its key. Like heap pages, nodes keep a cache of code area blocks.
+The commitments of the AVM databases are dependant to the clustering of data objects. So, the clustering algorithm of
+the AVM storage has to be a part of the Argennon agreement protocol. Clustering of objects is one of the
+tasks that a block proposer performs in Argennon.
 
-\note{Unlike heap pages, the AVM is not aware of different blocks of code area.}
+Every block of the Argennon blockchain contains a set of \emph{clustering directives}. These directives
+can only modify pages that were used for validating the block, and can
+include directives for moving an object from one page to another or directives specifying which pages will contain
+the newly created objects. These directives will always be run by nodes at the end of block validation.
+
+A block proposer usually obtains clustering directives from a third party source who enjoys a considerable amount of
+computational power. This federated approach does not affect Argennon security, because the integrity of a
+database can not be altered by clustering directives. Those
+directives can only affect the performance of the argennon blockchain, and directives of a single block can
+not change the performance much.

### 2021-04-27 tex/sections/blockchain.tex
Update the rest of the document
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 \begin{itemize}
-    \item \texttt{avmCall} essentially is an \texttt{invoke\_external} instruction that invokes a method from an
+    \item \texttt{avmCall}: invokes a method from an
     AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
     assets, including ARGs, is done by these transactions.
-    \item \texttt{installApp} installs an AVM smart contract and determines the update policy of the smart
+    \item \texttt{installApp}: installs an AVM smart contract and determines the update policy of the smart
     contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
     on.
-    \item \texttt{unInstallApp} removes an AVM smart contract.
-    \item \texttt{updateApp} updates an AVM smart contract.
+    \item \texttt{unInstallApp}: removes an AVM smart contract.
+    \item \texttt{updateApp}: updates an AVM smart contract.
 \end{itemize}
 
-All types of Argennon transactions contain an \texttt{invoke\_external} instruction which calls a special method
-from ARG smart contract that transfers the proposed fee of the transaction in ARGs from a sender account to the
-fee sink accounts.
+All types of Argennon transactions contain a single \texttt{invoke\_external} instruction which calls a method
+from the root smart contract that transfers the proposed fee of the transaction in ARGs from a sender account to the
+fee sink accounts and then performs the requested operation.
 
 Every transaction is required to exactly specify what heap locations or code area addresses it will access. This
-enables validators to start retrieving the required memory blocks from available ZK-EDB servers as soon as they
+enables validators to start retrieving the required storage pages from available ZK-EDB servers as soon as they
 see a transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to
 access a memory location that is not included in its access lists, will be rejected. Users could use smart contract
-oracles to predict the list of memory blocks their transactions need. See Section~\ref{sec:smart-contract-oracle}
+oracles to predict the list of memory locations their transactions need. See Section~\ref{sec:smart-contract-oracle}
 for more details.
@@
 
-The only information that Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
+The only information that all Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
 blockchain. Every block of the Argennon blockchain contains the following information:
@@
 For confirming a new block, nodes that are not validators only need to verify the block certificate. For
-verifying a block certificate, a node needs to know the ARG balances of validators, but it doesn't need to
+verifying a block certificate, a node needs to know the staking information of validators, but it doesn't need to
 emulate the AVM execution.
@@
 On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
-execution of the Argennon virtual machine. To do so, first they retrieve all heap pages and code area blocks they
+execution of the Argennon Virtual Machine. To do so, first they retrieve all storage pages they
 need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
-transactions included in the new block. This will modify some pages of the AVM memory, so they update the ZK-EDB
+transactions included in the new block. This will modify some pages of the AVM storage, so they update the ZK-EDB
 commitments based on the modified pages and verify the commitments included in the new block. Validators also
@@
 Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
-special method from ARG smart contract. This method will transfer the proposed fee of the transaction in ARGs
+special method from the root smart contract. This method will transfer the proposed fee of the transaction in ARGs
 from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects

### 2021-04-28 tex/sections/blockchain.tex
Do minor modification
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
         commitment to the set of transactions   \\ [0.7ex]
+        clustering directives                   \\ [0.7ex]
         previous block hash                     \\ [0.7ex]

### 2021-04-28 tex/sections/concurrency.tex
Do minor modification
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
     }
-    \caption{executing DAG transactions}\label{alg:exec_dag}
+    \caption{Executing DAG transactions}\label{alg:exec_dag}
 \end{algorithm}
@@
 
-\subsection{Memory Chunks}\label{subsec:memory-chunks}
+\subsection{Memory Shards}\label{subsec:memory-shards}
 
-In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{chunks}.
-Each memory chunk can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
-consensus on new values of the commitment of any chunk can be achieved by a different voting committee.
+In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{shards}.
+Each memory shard can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
+consensus on new values of the commitment of any shard can be achieved by a different voting committee.
 
-If a transaction does not modify a memory chunk and in the transaction ordering of the block it comes after
-any transaction which modifies that chunk, then the execution of that transaction is not needed for calculating
-the new commitment of the chunk. Consequently, the voting committee of that memory chunk can safely ignore such a
+If a transaction does not modify a memory shard and in the transaction ordering of the block it comes after
+any transaction which modifies that shard, then the execution of that transaction is not needed for calculating
+the new commitment of the shard. Consequently, the voting committee of that memory shard can safely ignore such a
 transaction. The execution DAG of transactions can be used for finding and pruning these transactions as
@@
     \DontPrintSemicolon
-    \SetKwData{V}{$v_0$}\SetKwData{Graph}{$G_e$}\SetKwData{Chunk}{$C_k$}\SetKwData{Txns}{$T$}
+    \SetKwData{V}{$v_0$}\SetKwData{Graph}{$G_e$}\SetKwData{Shard}{$Sh_k$}\SetKwData{Txns}{$T$}
     \SetKwFunction{RPrune}{RecursivePrune}
@@
     \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-    \Input{An execution dag \Graph and a memory chunk \Chunk}
-    \Output{The pruned execution dag based on \Chunk}
+    \Input{An execution dag \Graph and a memory shard \Shard}
+    \Output{The pruned execution dag based on \Shard}
     \BlankLine
@@
     {
-        \If{$Adj[\V] = \varnothing$ {\bf and} \V does not modify \Chunk}
+        \If{$Adj[\V] = \varnothing$ {\bf and} \V does not modify \Shard}
         {
@@
 
-If we choose chunks in a way that most transactions only modify memory locations of one chunk,
+If we choose shards in a way that most transactions only modify memory locations of one shard,
 likely many transactions of a block only need to be validated by one voting committee and can be validated in

### 2021-04-28 tex/sections/persistence.tex
Update the rest of the document
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Do minor modification
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The Argennon Virtual Machine has two persistent memory areas: \emph{method area}, and \emph{heap}. Method area stores
-method byte codes \footnote{also it stores constant area blocks.}, and heap stores memory chunks. Both of these
+method byte codes\footnote{also it stores constant area blocks.}, and heap stores memory chunks. Both of these
 data elements, methods and chunks, can be considered as continuous pieces of byte addressable memory
@@
 In the AVM persistence layer, similar objects are clustered together and constitute a bigger data element which we call a
-\emph{page}.\footnote{we avoided calling them a cluster, because usually a cluster refers to a \emph{set}. AVM object
+\emph{page}.\footnote{we avoided calling them clusters, because usually a cluster refers to a \emph{set}. AVM object
 clusters are not sets. They are ordered lists, like a page containing an ordered list of words or sentences.}
@@
 A page of the AVM storage should contain objects that have very similar access pattern. We expect that when a page
-is needed for validating a block, almost all of its objects are needed. We also prefer that the objects of the page
+is needed for validating a block, almost all of its objects are needed. We also prefer that the objects
 are needed for the same access type. In other words, the objects of a page are chosen in a way that
@@
 Pages of the AVM storage space are persisted using updatable zero-knowledge elementary databases (ZK-EDB). Argennon
-has three zero-knowledge databases. \emph{Staking} database which stores all the data that is associated with
-the Argennon agreement protocol. \emph{Method} database which stores the AVM method area, and \emph{Heap} database which
-stores the AVM heap. The commitment of these three ZK-EDBs are included in every block of the Argennon blockchain.
+has three zero-knowledge databases. \emph{staking} database, which stores all the data that is associated with
+the Argennon agreement protocol. \emph{method} database, which stores the AVM method area, and \emph{heap} database,
+which stores the AVM heap. The commitment of these three ZK-EDBs are included in every block of the Argennon blockchain.
 
@@
 
-The commitments of the AVM databases are dependant to the clustering of data objects. So, the clustering algorithm of
-the AVM storage has to be a part of the Argennon agreement protocol. Clustering of objects is one of the
-tasks that a block proposer performs in Argennon.
+The commitments of the AVM databases are dependant to the shape of clusters of data objects. So,
+the clustering algorithm of the AVM storage has to be a part of the Argennon agreement protocol. Clustering of
+objects is one of the tasks that a block proposer performs in Argennon.
 
@@
 A block proposer usually obtains clustering directives from a third party source who enjoys a considerable amount of
-computational power. This federated approach does not affect Argennon security, because the integrity of a
-database can not be altered by clustering directives. Those
-directives can only affect the performance of the argennon blockchain, and directives of a single block can
+computational power.\footnote{we can say the AVM clustering algorithm is essentially off-chain.} This federated approach
+does not affect Argennon security, because the integrity of a database can not be altered by clustering directives.
+Those directives can only affect the performance of the argennon blockchain, and directives of a single block can
 not change the performance much.

### 2021-04-28 tex/sections/avm.tex
Update persistence section
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Fix a typo
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The Argennon virtual machine does not have a fixed word size. Any instruction of the Argennon Virtual Machine has
+The Argennon virtual machine does not have a fixed word size. Any instruction of the AVM has
 its specific word size, and the addressable memory areas are byte addressable.
@@
 All these identifiers are \emph{prefix codes}, and hence can be represented by
-\emph{prefix trees}\footnote{Also called a trie.}. The AVM has three primitive prefix trees:
+\emph{prefix trees}\footnote{Also called a trie.}. The Argennon has three primitive prefix trees:
 \emph{applications, accounts} and \emph{local}. Any identifier in Argennon is a
@@
 \[
-    <A^{(1)},A^{(2)},A^{(3)},\dots>
+    (A^{(1)},A^{(2)},A^{(3)},\dots)
 \]
@@
 One important property of prefix identifiers is that while they have variable and unlimited length, they are
-uniquely extractable from any data. Assume that we have a string of digits which are in base $\beta$, we
+uniquely extractable from any sequence. Assume that we have a string of digits in base $\beta$, we
 know that $k$ first digits belong to an Argennon's identifier, but we don't know the value of $k$.
-Algorithm~\ref{alg:prefix_id} can extract the prefixed identifier uniquely. We can apply this algorithm
-multiple times to a sequence to extract a composite identifier, for example \texttt{chunkID}, from a sequence.
+Algorithm~\ref{alg:prefix_id} can be used to extract the prefixed identifier uniquely. Also, we can apply this algorithm
+multiple times to extract a composite identifier, for example \texttt{chunkID}, from a sequence.
 
@@
 The heap of the Argennon Virtual Machine is a persistent memory area which stores \emph{memory chunks}. A Memory
-chunk is a byte addressable piece of memory which has a separate address space starting from 0. Each chunk has a
-fixed size and different chunks need not be equally sized. Chunks are stored in the heap area and every chunk is
+chunk is a byte addressable piece of continuous memory which has a separate address space starting from 0. Each chunk
+has a fixed size and different chunks need not be equally sized. Chunks are stored in the heap area and every chunk is
 assigned a unique identifier: \texttt{chunkID}. As a result, the address of every memory location inside
@@
 \note{For AVM smart contracts \emph{reentrancy} can only happen in read-only mode and is totally safe.
-At the same time call-back patterns can be easily implemented by using \texttt{spawn} instruction.}
+At the same time, call-back patterns can be easily implemented by using \texttt{spawn} instruction.}
 

### 2021-06-03 tex/sections/avm.tex
Add the pronunciation note
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Fix a few typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The Argennon Virtual Machine (AVM) is an abstract computing machine for executing Argennon's smart contracts. The
+The Argennon\footnote{the classical pronunciation should be used:\textipa{/ar"gen.non/}} Virtual Machine (AVM) is an
+abstract computing machine for executing Argennon's smart contracts. The
 Argennon Virtual Machine knows nothing of the Argennon blockchain, only of Argennon's identifier tries and
@@
 \begin{itemize}
-    \item \texttt{alloc}: creates a new memory chunk of the specifies size in the heap. the identifier
+    \item \texttt{alloc}: creates a new memory chunk of the specified size in the heap. the identifier
     of the new chunk will be the concatenation of the \texttt{applicationID} of the creator of the chunk and
-    the \texttt{id} operand. The specified identifier must be concatenation of
+    the \texttt{id} operand of the instruction. The \texttt{id} operand must be the concatenation of
     an \texttt{accountID} and a prefix code
@@
     \item \texttt{delete}: deletes the chunk that its identifier is the concatenation of the current smart
-    contract \texttt{applicationID} to the specified \texttt{id}. If the identifier is not valid or
+    contract \texttt{applicationID} and the specified \texttt{id} operand. If the identifier is not valid or
     if the chunk was not found a catchable exception will be thrown.
@@
 the AVM, instead the smart contract can choose an identifier itself. This is a very important feature of
-the AVM heap which allows smart contracts to use the AVM heap as a dictionary (map) data structure.
-Because the \texttt{chunkID} is a prefix code, any smart contract has its own identifier space and a smart contract
+the AVM heap, which allows smart contracts to use the AVM heap as a dictionary (map) data structure.
+Since the \texttt{chunkID} is a prefix code, any smart contract has its own identifier space and a smart contract
 can easily generate unique identifiers for its chunks.
@@
 The AVM requires the chosen identifier to be a prefix code made by concatenating a code compatible with
-the \emph{accounts} prefix tree to a code compatible with the \emph{local} prefix tree. This requirement will
-enable AVM to detect invalid identifiers while smart contracts can easily associate data with account identifiers.
+the \emph{accounts} prefix tree to a code compatible with the \emph{local} prefix tree. This requirement
+enables AVM to detect invalid identifiers, while smart contracts can easily associate data with account identifiers.
 At the same time, A smart contract is able to create maps with costume keys by appending local identifiers
-to the \texttt{accountID} zero, which does not correspond to a valid account in the Argennon.
+to the zero \texttt{accountID}. Zero \texttt{accountID} does not correspond to a valid account in the Argennon
+blockchain.
 

### 2021-06-06 tex/sections/avm.tex
Describe the AVM standard library
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 \section{The AVM Standard Library}\label{sec:the-avm-standard-library}
+
+As we explained in Section~\ref{subsec:method-invocation} by using \texttt{invoke\_internal} instruction, a smart
+contract can invoke methods of another smart contract in its own context. This is how AVM smart contracts
+invoke methods of the AVM standard library. Methods of the AVM standard library are stored
+with the \texttt{applicationID} zero in the AVM method area. This implies that the AVM standard library is actually
+a part of the root smart contract.
+
+In Argennon, the root smart contract is an updatable smart contract, which can be updated by the Argennon consensus
+protocol. This means that bugs or security vulnerabilities in the AVM standard library could be quickly patched and
+smart contracts can use the AVM standard library safely. Many important and useful functionalities,
+such as fungible and non-fungible assets, access control mechanisms,
+and general purpose DAOs are implemented in the Argennon standard library.
+
+All Argennon standards, for instance ARC standard series, which defines standards regarding transferable assets,
+are defined based on how a contract should use the AVM standard library. As a result, Argennon standards are
+different from conventional blockchain standards. Argennon standards define some type of standard logic and
+behaviour for a smart contract, not only a set of method signatures. This enables users to expect certain type
+of behaviour from a contract which complies with an Argennon standard.

### 2021-06-08 tex/sections/avm.tex
Fix a few typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-In the Argennon Virtual Machine four distinct identifier values exist: \texttt{applicationID}, \texttt{accountID},
+In the Argennon Virtual Machine four distinct identifier types exist: \texttt{applicationID}, \texttt{accountID},
 \texttt{methodID} and \texttt{chunkID}.
@@
 As we explained in Section~\ref{subsec:method-invocation} by using \texttt{invoke\_internal} instruction, a smart
-contract can invoke methods of another smart contract in its own context. This is how AVM smart contracts
+contract can invoke methods of another smart contract in its own context. AVM smart contracts use this instruction to
 invoke methods of the AVM standard library. Methods of the AVM standard library are stored

### 2021-06-08 tex/sections/persistence.tex
Fix a few typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 method byte codes\footnote{also it stores constant area blocks.}, and heap stores memory chunks. Both of these
-data elements, methods and chunks, can be considered as continuous pieces of byte addressable memory
-which we shall call \emph{objects} throughout this chapter.
+data elements, methods and chunks, can be considered as continuous pieces of byte addressable memory. Throughout this
+chapter, we shall call these data elements \emph{objects}.
 
@@
 include directives for moving an object from one page to another or directives specifying which pages will contain
-the newly created objects. These directives will always be run by nodes at the end of block validation.
+the newly created objects. These directives are always executed by nodes at the end of block validation.
 
-A block proposer usually obtains clustering directives from a third party source who enjoys a considerable amount of
+A block proposer usually obtains clustering directives from a third party source that enjoys a considerable amount of
 computational power.\footnote{we can say the AVM clustering algorithm is essentially off-chain.} This federated approach

### 2021-06-09 tex/sections/avm.tex
Fix a few typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The execution of smart contracts code in the AVM is done in execution sessions. Every execution session in the
-Argennon Virtual Machine consists of a single method invocation from the \emph{root} smart contract and a list
+The execution of smart contracts code by the Argennon Virtual Machine is performed in execution sessions. Every
+execution session consists of a single method invocation from the \emph{root} smart contract and a list
 of resource allocation requests. These allocation requests always include a request for allocating some amount
-of execution time and a request for accessing some memory locations for reading or writing. If after the start
+of execution time and a request for accessing some memory locations for reading or writing. If after starting,
 an execution session tries to violate its allocated resources, an \textbf{uncatchable} exception will
@@
 The root smart contract with the \texttt{applicationID} of zero, is a special smart contract in the Argennon
-Virtual Machine. Its code is immutable and is a part of the AVM implementation. The root smart contract is always
-run in the privileged mode, and every execution session starts with a call to a method of the root smart contract.
-This method always transfers some amount of fee from an account ot the \texttt{feeSink} account and then it
+Virtual Machine. Its code is mutable and is a part of the Argennon protocol. The root smart contract is always
+run in the privileged mode, and every Argennon transaction starts with a call to a method of the root smart contract.
+This method always transfers some amount of fee from an account to the \texttt{feeSink} accounts and then it
 performs the requested operation.
 
+If the requested operation is a method invocation, the root smart contract will call the \texttt{Heap.Save()}
+function, and then it will invoke the requested method using \texttt{spawn} instruction. This implies that the
+invoked method will always have a separate call stack, and if it completes abruptly the changes made by the root
+smart contract to the AVM state will not be reverted.
+
 \note{Argennon transactions do not have a sender, and the payer of the transaction fee could be any
-account who has provided the required digital signature.}
-
-If the requested operation is a method invocation, the root smart contract will call the \texttt{Heap.Save()}
-function once, and then it will invoke the requested method using \texttt{spawn} instruction. This means that the
-invoked method will have a separate call stack and if it completes abruptly the changes made by the root
-smart contract to the state will not be reverted.
+account who has provided the required digital signature for fee payment.}
 

### 2021-06-20 tex/sections/avm.tex
Add the Argennon consensus protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The Argennon\footnote{the classical pronunciation should be used:\textipa{/ar"gen.non/}} Virtual Machine (AVM) is an
+The Argennon\footnote{the classical pronunciation should be used:\textipa{/Ar"gen.non/}} Virtual Machine (AVM) is an
 abstract computing machine for executing Argennon's smart contracts. The
@@
 All these identifiers are \emph{prefix codes}, and hence can be represented by
-\emph{prefix trees}\footnote{Also called a trie.}. The Argennon has three primitive prefix trees:
+\emph{prefix trees}\footnote{Also called tries.}. The Argennon has three primitive prefix trees:
 \emph{applications, accounts} and \emph{local}. Any identifier in Argennon is a
@@
     \item \texttt{methodID} is a composite prefix code built by concatenating an \texttt{applicationID} to
-    a prefix code made by \emph{local} prefix tree.(\texttt{|} is the concatenating operator.)
+    a prefix code made by \emph{local} prefix tree: (\texttt{|} is the concatenating operator)
 
@@
     \item \texttt{chunkID} is a composite prefix code built by concatenating an \texttt{applicationID} to
-    an \texttt{accountID} to a prefix code made by \emph{local} prefix tree
+    an \texttt{accountID} to a prefix code made by \emph{local} prefix tree:
 
@@
 All Argennon prefix trees have an equal branching factor \(\beta\). Therefore, we can represent an Argennon
-prefix tree as a sequence of fractional numbers in base \(\beta\):
+prefix tree as a sequence of fractional numbers\footnote{It's possible to have \(a_i=0\). For
+instance, \(A^{(4)}=(0.2000)_{10}\) is a valid prefix tree.} in base \(\beta\):
 \[
@@
 of resource allocation requests. These allocation requests always include a request for allocating some amount
-of execution time and a request for accessing some memory locations for reading or writing. If after starting,
+of execution cost and a request for accessing some memory locations for reading or writing. If after starting,
 an execution session tries to violate its allocated resources, an \textbf{uncatchable} exception will
@@
 \note{Argennon transactions do not have a sender, and the payer of the transaction fee could be any
-account who has provided the required digital signature for fee payment.}
+account who has provided the required digital signature for the fee payment.}
 

### 2021-06-20 tex/sections/blockchain.tex
Add the Argennon consensus protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-04-27 tex/sections/blockchain.tex
@@
 
-\section{Persistence}\label{sec:persistence}
-\input{sections/persistence}
+\section{Applications}\label{sec:applications}
+
+An Argennon application or a smart contract is a collection of method bytecodes and heap chunks that are stored in
+the AVM storage, identified by a unique application identifier. An application identifier, \texttt{applicationID}, is
+a unique prefix code generated by the \emph{applications} prefix tree. (See Section~\ref{sec:identifiers}.)
+
+An application identifier can
+be considered as the address of an application and has the following standard symbolic representation:
+\begin{verbatim}
+<application-id> ::= "["<hex-prefix-code>"]"
+<hex-prefix-code> ::= "0x"<hex-num>"."<hex-prefix-code> | "0x"<hex-num>
+\end{verbatim}
+where \texttt{<hex-num>} is a hexadecimal number between $0$ and $255$, using lower case
+letters \texttt{[a-f]} for showing digits greater than $9$.
+
+For example \texttt{[0x24.0xff.0xda]}, \texttt{[0x0]} and \texttt{[0x3.0xa0.0x0.0x0]}, are valid application addresses.
+
+Argennon has two special smart contracts: the \emph{root} smart contract\footnote{also called the root app} and
+the \emph{ARG} smart contract\footnote{also called the Argennon smart contract or the Argennon app.}.
+The root smart contract, with \texttt{applicationID=[0x0]}, is a privileged smart contract responsible for
+method invocation and installation/uninstallation of other smart contracts.
+On the other hand, the ARG smart contract, with \texttt{applicationID=[0x1]}, controls the ARG token, the main
+currency of the Argennon blockchain, and also manages a database of public keys and handles signature verification.
+
+Both of these smart contracts are mutable smart contracts and can be updated by the Argennon governance protocol.
+
+\section{Accounts}\label{sec:accounts}
+
+Argennon accounts are entities defined inside the ARG smart contract.
+Every Argennon account is uniquely identified by a prefix code generated using \emph{accounts} prefix tree. An account
+identifier can be considered as the address of an account and has the following standard symbolic representation:
+\begin{verbatim}
+<account-id> ::= "["<decimal-prefix-code>"]"
+<decimal-prefix-code> ::= <dec-num>"."<decimal-prefix-code> | <dec-num>
+\end{verbatim}
+where \texttt{<dec-num>} is a normal decimal number between $0$ and $255$.
+
+For example \texttt{[21.255.37]}, \texttt{[0]} and \texttt{[1.0.0.0.0]}, are valid standard symbolic representations
+of account addresses.
+
+A new account can be created by invoking \texttt{createAccount} method from the Argennon smart contract. For creating
+a new account two public keys need to be provided by the caller and registered in the Argennon smart contract.
+One public key will be used for issuing digital signatures, and the other one will be used for voting. The
+provided public keys need to meet certain cryptographic requirements,\footnote{Argennon uses Prove
+Knowledge of the Secret Key (KOSK) scheme.} and can not be already registered in the system.
+
+If the owner of the new account is an application, the \texttt{applicationID} of the owner will be registered in the
+ARG smart contract and no public keys are needed. An application can own an arbitrary number of accounts.
 
@@
 
-Argennon has four types of transaction:
+Every Argennon transaction is a single \texttt{invoke\_external} instruction which calls a method from the Argennon
+root smart contract. This method always transfers the proposed fee of the transaction in ARGs from a
+sender account to the fee sink accounts and then performs the requested operation. The Argennon root smart contract
+has four public methods:
 
@@
     \item \texttt{unInstallApp}: removes an AVM smart contract.
-    \item \texttt{updateApp}: updates an AVM smart contract.
+    \item \texttt{updateApp}: updates an AVM smart contract (if allowed).
 \end{itemize}
 
-All types of Argennon transactions contain a single \texttt{invoke\_external} instruction which calls a method
-from the root smart contract that transfers the proposed fee of the transaction in ARGs from a sender account to the
-fee sink accounts and then performs the requested operation.
+Argennon transactions do not have a sender. The authorization of the requested operation is always done by checking the
+digital signatures that are provided as a part of the argument list to the root smart contract method.
 
-Every transaction is required to exactly specify what heap locations or code area addresses it will access. This
-enables validators to start retrieving the required storage pages from available ZK-EDB servers as soon as they
-see a transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to
-access a memory location that is not included in its access lists, will be rejected. Users could use smart contract
-oracles to predict the list of memory locations their transactions need. See Section~\ref{sec:smart-contract-oracle}
-for more details.
+Every Argennon transaction is required to specify a cap for all the resources it needs. This
+will include memory, network and processor related resources. The protocol defines an execution cost for
+every AVM instruction, reflecting the amount of resources its emulation needs, and every transaction is required
+to specify a maximum execution cost. If during emulation, a transaction reaches this maximum cost, it will be
+considered failed and the network can receive the proposed fee of that transaction.
 
-\section{Blockchain}\label{sec:blockchain}
+Also, Argennon transactions are required to specify what heap or code area addresses they will access. This will
+enable validators to parallelize transaction validation as we will see in Section~\ref{sec:concurrency}. A transaction
+that tries to access a memory location that is not in its access lists, will be rejected.
+Users could use off-chain \emph{smart contract oracles} to predict the list of memory locations their transactions need.
 
-Every block of the Argennon blockchain corresponds to a set of transactions. We store the commitment of this
-transaction set in every block, but we don't keep the set itself. To be able to detect replay attacks, we require
-every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
-and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
-must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
-a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
-if a signature of the same user with a bigger or equal nonce is already used
-(i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
-nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
+A smart contract oracle is a full AVM emulator that keeps a full local copy of the AVM memory and can emulate AVM
+execution without accessing a ZK-EDB server. Smart contract oracles can be used for reporting useful information about
+Argennon transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
+and so on.
 
-To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. When the
-difference between \texttt{issuance} component of this nonce and the current round becomes bigger than the
-maximum allowed lifetime of a signature, this information can be safely deleted. \textbf{As a result, we will not
-have the problem of "un-removable empty accounts" like Ethereum.}
+Every Argennon transaction is required to provide the following information as an upper bound for the
+resources it needs:
 
-The only information that all Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
-blockchain. Every block of the Argennon blockchain contains the following information:
+\begin{itemize}
+    \item Maximum execution cost
+    \item A list of heap/code-area locations for reading
+    \item A list of heap locations for writing
+    \item A list of heap chunks it will deallocate
+    \item A list of methods it will delete (if any)
+    \item Number and size of heap chunks it will allocate
+    \item Number and size of method bytecodes it will allocate (if any)
+\end{itemize}
+
+If a transaction tries to violate any of these predefined limitations, it will be considered failed, and the network
+can receive the proposed fee of that transaction.
+
+\note{An Argennon transaction always pays all of its proposed fee, no matter how much of its predefined resources
+were not used in the final emulation. This will incentivize users to exactly report their
+transactions resource usage.}
+
+\section{Blocks}\label{sec:blocks}
+
+The Argennon blockchain is a sequence of blocks. Every block represents an ordered list of transactions, intended to be
+executed by the Argennon Virtual Machine. The first block of the blockchain, the \emph{genesis} block, is a spacial
+block that fully describes the initial state of the AVM. Every block of the Argennon blockchain thus corresponds to a
+unique AVM state which can be calculated deterministically from the genesis block.
+
+A block of the Argennon blockchain contains the following information:
 
@@
         \hline\hline
-        commitment to the staking database \\ [0.7ex]
-        commitment to the method database \\ [0.7ex]
-        commitment to the heap database \\ [0.7ex]
-        commitment to the set of transactions       \\ [0.7ex]
-        previous block hash                         \\ [0.7ex]
-        random seed                                 \\ [0.7ex]
+        commitment to the staking database            \\ [1.2ex]
+        commitment to the method database             \\ [1.2ex]
+        commitment to the heap database               \\ [1.2ex]
+        commitment to the set of transactions         \\ [1.2ex]
+        last block certificate issued by         \\
+        the validators committee                      \\ [1.2ex]
+        clustering directives                         \\ [1.2ex]
+        random seed                                   \\ [1.2ex]
+        previous block hash                           \\ [1.2ex]
         \hline
@@
 
-For confirming a new block, nodes that are not validators only need to verify the block certificate. For
-verifying a block certificate, a node needs to know the staking information of validators, but it doesn't need to
-emulate the AVM execution.
+\subsection{Block Validation}\label{subsec:block-validation}
 
-On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
-execution of the Argennon Virtual Machine. To do so, first they retrieve all storage pages they
-need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
-transactions included in the new block. This will modify some pages of the AVM storage, so they update the ZK-EDB
-commitments based on the modified pages and verify the commitments included in the new block. Validators also
-calculate and verify the commitment to the new block's transaction set.
+Having the previous AVM state, the transaction list and the clustering directives of a block, a node can calculate
+commitments to the staking, method and heap databases of the current block by emulating the AVM execution. If the
+node can obtain the previous block information from a trusted source, it does not need to have a trusted local
+copy of the AVM state,
+and it can reliably retrieve the required storage pages from a ZK-EDB server. We call this type of block verification
+\emph{conditional} block validation. This validation is conditional because the validity of the current block is
+conditioned on the validity of the previous block.
 
-\note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDB servers will receive the new
-block, and they will update their database by emulating the AVM execution.}
+Interestingly, conditional block validation of multiple blocks can be done in parallel. If a node has enough bandwidth
+and computational resources, it can conditionally verify any number of blocks from a previously created blockchain
+simultaneously and in parallel. As we will see in Section~\ref{subsec:validators-committee}, this property plays an
+important role in the Argennon consensus protocol.
+
+To some extent, conditional validation of a single block could be parallelized as well. Many transactions
+in a block are actually independent and the order of their execution does not
+matter. These transactions can be safely validated in parallel. Section~\ref{sec:concurrency} further
+develops this concept.
+
+\subsection{Block Certificate}\label{subsec:block-certificate}
+
+An Argennon block certificate is an aggregate signature of some predefined subset of accounts. This predefined subset
+is called the certificate committee and their signature ensures that the certified block is conditionally
+valid given the validity of some previous block.
+
+Argennon uses BLS aggregate signatures to represent block certificates. To better understand block certificates and
+the Argennon consensus protocol, we need to briefly review the BLS signature scheme and its aggregation mechanism.
+
+The BLS signature scheme operates in a prime order group and supports simple threshold signature generation,
+threshold key generation, and signature aggregation. To review, the scheme uses the following ingredients:
+
+\newcommand{\G}{\mathbb{G}}
+\newcommand{\Z}{\mathbb{Z}}
+\newcommand{\adv}{{\cal A}}
+\newcommand{\bdv}{{\cal B}}
+\newcommand{\deq}{\mathrel{\mathop:}=}
+\newcommand{\SK}{\mathit{sk}}
+\newcommand{\PK}{\mathit{pk}}
+\newcommand{\C}{\mathit{cert}}
+\newcommand{\APK}{\mathit{apk}}
+\newcommand{\DPK}{\mathit{\Delta pk}}
+\newcommand{\MM}{\mathcal{M}}
+\newcommand{\xwedge}{\, \operatorname{\text{$\wedge$}}\, }
+\newcommand{\abs}[1]{\lvert #1 \rvert}
+\newcommand{\Hm}{H_0}
+\newcommand{\Hpk}{H_1}
+\newcommand{\qHpk}{Q_{\Hpk}}
+\newcommand{\qHm}{Q_{\Hm}}
+\newcommand{\qsig}{Q_{\text{sig}}}
+
+\begin{itemize}
+    \item An efficiently computable \emph{non-degenerate} pairing $e:\G_0 \times \G_1 \to \G_T$
+    in groups $\G_0$, $\G_1$ and $\G_T$ of prime order $q$. We let $g_0$ and $g_1$ be generators
+    of $\G_0$ and $\G_1$ respectively.
+    \item A hash function $H_0: \mathcal{M} \rightarrow \mathbb{G}_0$, where $\mathcal{M}$ is the message space.
+    The hash function will be treated as a random oracle.
+\end{itemize}
+
+The BLS signature scheme is defined as follows:
+
+\begin{itemize}
+    \item $\textbf{KeyGen}()$: choose a random $\alpha$ from $\Z_q$ and set $h \gets g_1^\alpha \in \G_1$.
+    output $\PK \deq (h)$ and $\SK \deq (\alpha)$.
+    \item $\textbf{Sign}(\SK, m)$: output $\sigma \gets \Hm(m)^\alpha \in \G_0$.
+    The signature $\sigma$ is a \emph{single} group element.
+    \item $\textbf{Verify}(\PK,m,\sigma)$: if $e(g_1, \sigma) = e\big(\PK,\ \Hm(m)\big)$  then output "accept",
+    otherwise output "reject".
+\end{itemize}
+
+Given triples $(\PK_i,\ m_i,\ \sigma_i)$ for $i=1,\ldots,n$,
+anyone can aggregate the signatures $\sigma_1,\ldots,\sigma_n \in \G_0$
+into a short convincing aggregate signature $\sigma$ by computing
+\begin{equation}
+    \label{eq:agg}
+    \sigma \gets \sigma_1 \cdots \sigma_n \in \G_0.
+\end{equation}
+Verifying an aggregate signature $\sigma \in \G_0$ is done by checking that
+\begin{equation}
+    \label{eq:aggdiff}
+    e(g_1, \sigma) = e\big(\PK_1,\ \Hm(m_1)\big) \cdots e\big(\PK_n,\ \Hm(m_n)\big).
+\end{equation}
+When all the messages are the same ($m = m_1 = \ldots = m_n$), the verification relation~\eqref{eq:aggdiff} reduces to
+a simpler test that requires only two pairings:
+\begin{equation}
+    \label{eq:aggsame}
+    e(g_1, \sigma) = e\Big(\PK_1 \cdots \PK_n,\ \Hm(m)\Big).
+\end{equation}
+We call $\APK=\PK_1 \cdots \PK_n$ the aggregate public key.
+
+To defend against \emph{rogue public key} attacks, Argennon uses Prove Knowledge of the Secret Key (KOSK) scheme. As we
+explained in Section~\ref{sec:accounts}, when an account is created its public keys need to be registered in
+the ARG smart contract. Therefore, the KOSK scheme can be easily implemented in Argennon.
+
+Because it is not usually possible to collect the signatures of all members of a certificate committee, an Argennon
+block certificate essentially is an Accountable-Subgroup Multi-signature (ASM). Argennon uses a simple ASM scheme
+based on BLS aggregate signatures.
+
+Argennon block certificates constitute an ordered sequence based on the order of blocks they certify. If we show
+the certificate of committee $C$ for the $i$-th block\footnote{note that the $i$-th block certificate is not
+necessarily the certificate of the $i$-th block.} with $\C_i$, and the set of signers
+with $S_i$, then the block certificate $\C_i$ can be considered as a tuple:
+\begin{equation}
+    \C_i=(\sigma_i,\ C-S_{i})\label{eq:cert},
+\end{equation}
+where $\sigma_i$ is the aggregate signature issued by $S_i$.
+
+The aggregate public key of the certificate can
+be calculated from:
+\begin{equation}
+    \APK_i=\APK_C\APK_{C-S_i}^{-1}\label{eq:aggCertPK}\ ,
+\end{equation}
+where $\APK_{A}$ shows the aggregate public key of all accounts in $A$.
+
+Alternately we can use $\APK_{i-1}$ to calculate the aggregate public key:
+\begin{equation}
+    \APK_i=\APK_{i-1}\APK_{S_i-S_{i-1}}\APK_{S_{i-1}-S_i}^{-1}\ .\label{eq:aggPK-2}
+\end{equation}
+
+When an Argennon account is created, both its $\PK$ and $\PK^{-1}$ is registered in the ARG smart contract, so the
+inverse of any aggregate public key can be easily computed.\footnote{since the group operator of a cyclic
+group is commutative, we have $(ab)^{-1}=a^{-1}b^{-1}$.}
 
@@
 \section{Incentive mechanism}\label{sec:incentive-mechanism}
+
+\note{TODO: update this section!}
 
@@
 
-Every \texttt{avmCall} transaction is required to provide the following information as an upper bound for the
-resources it needs:
-
-\begin{itemize}
-    \item Execution cost
-    \item A list of heap/code-area locations for reading
-    \item A list of heap locations for writing
-    \item A list of heap pages it will deallocate
-    \item Number and size of heap pages it will allocate
-\end{itemize}
-
-If a transaction tries to violate any of these predefined limitations, for example, if it tries to read a memory
-location that is not included in its reading list, it will be considered failed and the network can receive the
-proposed fee of that transaction.
-
-\note{A transaction always pays all of its proposed fee, no matter how much of its predefined resources were not
-used in the final emulation.}
-
 \subsection{Incentives for ZK-EDB Servers}\label{subsec:zk-edb-servers}
@@
 To run this lottery, every round, based on the current block seed, a collection of \emph{valid} receipts will be
-selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in round \(r\) if:
+selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in round $r$ if:
 
 \begin{itemize}
-    \item The signer was a validator in the round \(r - 1\) and voted for the agreed-upon block.
+    \item The signer was a validator in the round $r - 1$ and voted for the agreed-upon block.
     \item The data block in the receipt was needed for validating the \textbf{previous} block.
-    \item The receipt round number is \(r - 1\).
+    \item The receipt round number is $r - 1$.
     \item The signer did not sign a receipt for the same data block for two different ZK-EDBs in the previous round.
@@
 
-The winning tickets of the lottery of round \(r\) need to be included in the block of the round \(r\),
+The winning tickets of the lottery of round $r$ need to be included in the block of the round $r$,
 otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
-\(r\) will be done in the round \(r + 1\). This way, \textbf{the content of the challenge memory block could be
+$r$ will be done in the round $r + 1$. This way, \textbf{the content of the challenge memory block could be
 kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.
@@
 
-Every \(k\) round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
+Every $k$ round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
 allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
@@
 
-To determine the price of AVM memory, Every \(k\) round, the protocol calculates \texttt{dbFee} and
+To determine the price of AVM memory, Every $k$ round, the protocol calculates \texttt{dbFee} and
 \texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
@@
 the protocol considers the total size of all the memory pages that were accessed for either reading or writing
-during a time period. These two values will be calculated for the last \(k\) rounds and the price per byte of
+during a time period. These two values will be calculated for the last $k$ rounds and the price per byte of
 AVM memory will be a linear function of \texttt{dbFee/memTraffic}
@@
 
+\section{Governance}\label{sec:governance}
+
+The Argennon Decentralized Autonomous Governance system (ADAGs)
+\note{not yet written...}
+
 \section{Concurrency}\label{sec:concurrency}
 \input{sections/concurrency}
-
-\section{Smart Contract Oracle}\label{sec:smart-contract-oracle}
-
-A smart contract oracle is a full AVM emulator that keeps a full local copy of AVM memory and can emulate AVM
-execution without accessing a ZK-EDB. Smart contract oracles can be used for reporting useful information about
-\texttt{avmCall} transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
-and so on.

### 2021-06-20 tex/sections/concurrency.tex
Add the Argennon consensus protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 structure for storing the balance record of highly active accounts.
-
-\subsection{Memory Shards}\label{subsec:memory-shards}
-
-In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{shards}.
-Each memory shard can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
-consensus on new values of the commitment of any shard can be achieved by a different voting committee.
-
-If a transaction does not modify a memory shard and in the transaction ordering of the block it comes after
-any transaction which modifies that shard, then the execution of that transaction is not needed for calculating
-the new commitment of the shard. Consequently, the voting committee of that memory shard can safely ignore such a
-transaction. The execution DAG of transactions can be used for finding and pruning these transactions as
-we see in Algorithm~\ref{alg:prune_dag}.
-
-%##\includegraphics[width=17cm]{../img/Alg4s.png}
-\begin{algorithm}
-    \DontPrintSemicolon
-    \SetKwData{V}{$v_0$}\SetKwData{Graph}{$G_e$}\SetKwData{Shard}{$Sh_k$}\SetKwData{Txns}{$T$}
-    \SetKwFunction{RPrune}{RecursivePrune}
-    \SetKwProg{Fn}{Function}{}{}
-    \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-    \Input{An execution dag \Graph and a memory shard \Shard}
-    \Output{The pruned execution dag based on \Shard}
-    \BlankLine
-    \For{each vertex \V with out degree $0$}
-    {
-        \RPrune{\V}\;
-    }
-    \BlankLine
-    \Fn{\RPrune{\V}}
-    {
-        \If{$Adj[\V] = \varnothing$ {\bf and} \V does not modify \Shard}
-        {
-            remove \V from \Graph\;
-            \For{each $u$ such that edge $(u,\V)$ was in \Graph}
-            {
-                \RPrune{u}\;
-            }
-        }
-    }
-    \caption{Pruning an execution DAG}\label{alg:prune_dag}
-\end{algorithm}
-
-If we choose shards in a way that most transactions only modify memory locations of one shard,
-likely many transactions of a block only need to be validated by one voting committee and can be validated in
-parallel by different committees.
-
-Because the voting committees are selected by random sampling, by choosing large enough samples we can make sure
-that having multiple voting committees will not change the security properties of the Argennon agreement protocol.

### 2021-06-20 tex/sections/consensus.tex
Update persistence section
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Add the Argennon consensus protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-\note{It will be a randomized proof of stake which is under development....}
+Argennon has two primary certificate committees: the committee of \emph{delegates} and the committee
+of \emph{validators}. A block is considered final only after it gets a certificate from both of these committees.
 
-\subsection{Estimating A User's Stake}\label{subsec:estimating-a-user's-stake}
+The delegates committee generates a certificate for every block of the Argennon blockchain, and the
+validators committee generates a certificate every $m$ blocks. The validators committee will
+certify a block only if it has already been certified by the committee of delegates.
+
+In addition to primary committees, Argennon has several community driven committees. Certificates of these
+committees are not required for block validation, but they could be used by the members of the validators
+committee to better decide about the validity of a block.
+
+When an anomaly is detected in the consensus mechanism, the \emph{recovery} protocol is initiated by nodes. The
+recovery protocol is designed to be resilient to many types of attacks in order to be able to restore the normal
+system functionality.
+
+\subsection{The Committee of Delegates}\label{subsec:the-committee-of-delegates}
+
+The committee of delegates is a small committee of trusted delegates, elected by Argennon users either through the
+Argennon Decentralized Autonomous Governance system (ADAGs\footnote{pronounced \textipa{/eI-dagz/}.}) or by emergency
+voting during the recovery protocol. At the start of the Argennon mainnet, this committee will have
+five members, and later its size could be changed by the ADAGs in a procedure described
+in Section~\ref{sec:governance}.
+
+The committee of delegates is responsible for creating new blocks of the Argennon blockchain, and it issues a
+certificate for every block of the Argennon blockchain. A certificate needs to be signed
+by \textbf{all} of the committee members in order to be considered valid.
+
+Usually, the delegates are large organizations, and they have enough computational resources to generate blocks
+very fast. However, a block is not completely final if it does not have the certificate of the validators committee.
+A certified block by the delegates will not be accepted by the network, if the last block certified by
+the validators committee is behind more than a certain number of blocks.
+
+The committee of delegates may use any type of agreement protocol to reach consensus on the
+next block. Usually a very simple and fast protocol can do the job: one of the members
+is randomly chosen as the proposer, and other members vote "yes" or "no" on the proposed block.
+
+If one of the delegates lose its network connectivity, no new blocks can be generated. For this reason,
+the delegates should invest on different types of communication infrastructure, to make sure they never lose
+connectivity to each other and to the Argennon network.
+
+\subsection{The Committee of Validators}\label{subsec:validators-committee}
+
+The Argennon protocol calculates a stake value for every account, which is an estimate of a user's stake in the
+system. This value is measured in ARGs. The committee of validators is composed of
+\textbf{all} accounts whose stake value is higher than \texttt{minValidatorsStake} threshold. This
+threshold is determined by the ADAGs, but it can never be higher
+than $100$ ARGs.
+
+Every member of the validators committee has a status which can be either \texttt{online} or \texttt{offline}.
+This status is stored in the ARG smart contract, and it can be changed through a method invocation
+from this smart contract. When an account sets its status to \texttt{offline}, it receives a small reward, and
+it can not change it back to \texttt{online} for \texttt{statusCoolDown} number of blocks.
+
+\note{When an account changes its status, the change has no effect until the block containing the status
+change transaction gets certified by the committee of validators.}
+
+A block certificate issued by some members of the validators committee is considered valid, if according to
+the staking database of the previous \textbf{certified block by the validators committee}, we have:
+\begin{itemize}
+    \item The total stake of \texttt{online} accounts is higher than \texttt{minOnlineStake} percent of the
+    total stake of all accounts. This threshold can be changed by the ADAGs, but it can never be lower
+    than 70 percent.
+    \item All signers of the certificate have \texttt{online} status.
+    \item The sum of stake values of the certificate signers is higher than 80 percent of the total stake
+    of accounts that have \texttt{online} status.
+\end{itemize}
+
+If the total online stake goes lower than \texttt{minOnlineStake} threshold, nodes initiate the recovery
+consensus protocol.
+
+As we mentioned before, every $m$ blocks of the Argennon blockchain needs a certificate from the committee
+of validators. This means that assuming the last certified block by the validators is valid, for issuing the block
+certificate, the committee needs to conditionally validate $m$ blocks. We call these $m$ blocks the
+validation set. To reduce resource usage of validators' nodes, the protocol requires honest nodes to only
+conditionally validate a subset of the validation set. Validators that have a stake lower than
+average should only validate one block in expectation.
+
+More precisely, to decide about signing the block certificate, every honest member of the validators
+committee iterates over blocks of the validation set, and selects a block with
+probability $p_u$. In this way, a member of the validators committee will select $p_{u}m$ blocks in expectation.
+Then, the honest member conditionally validates the selected blocks. If the member has selected more than one block,
+he may validate the selected blocks in parallel.
+If all checked blocks are valid, the validator issues an "accept"
+signature, otherwise he issues a "reject" signature.
+
+One recommended formula for calculating the probability of selecting a block for validation by user $u$ is:
+\begin{equation}
+    p_u=\max\left(\frac{s_u}{m\mu},\ \frac{1}{m}\right),\label{eq:block-selection}
+\end{equation}
+where $\mu$ is the average stake of committee members and $s_u$ is the stake of user $u$.
+
+If the committee of validators, at any point issues a "reject" certificate for a block, the network will initiate the
+recovery protocol.
+
+\subsubsection{Signature Aggregation}
+
+In Argennon, signature aggregation is mostly performed by ZK-EDB servers. To distribute the aggregation workload
+between different servers, The committee of validators is divided into pre-determined groups, and each ZK-EDB
+server is responsible for signature aggregation of one group. To make sure that there is enough redundancy, the number
+of groups should be less than the number of ZK-EDB servers and each group should be assigned to
+multiple ZK-EDB servers.
+
+Any member of a group knows all the servers that are responsible for signature aggregation of his group. When a member
+signs a block certificate, he sends his signature to \textbf{all} the severs that aggregate the signatures of his group.
+These servers aggregate the signatures they receive, and then send the aggregated signature to the delegates.
+Furthermore, the delegates aggregate these signatures to produce the final block certificate,
+and then include it in the next block.
+
+The role of the delegates in the signature aggregation is limited. The important part of the work is done by ZK-EDB
+servers. As long as there are enough honest ZK-EDB servers, the network will be able to perform signature aggregation
+even if the delegates are malicious. (See Section~\ref{subsec:recovery})
+
+\subsection{Estimating A User's Stake}\label{subsec:user's-stake}
 
 In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
-of stake the user has in the system. Conventionally in these systems, for estimating a user's stake, we use the
-amount of native system tokens the user is holding. Unfortunately, one problem with this approach is that a
-strong attacker may be able to obtain a considerable amount of system tokens, for example by borrowing from a
-DEFI application, and use this stake to attack the system.
+of stake the user has in the system. Conventionally in these systems, a user's stake is considered to be equal with the
+amount of native system tokens, the user has "staked" in the system. A user stakes his tokens by locking them in
+his account or a staking account for some period of time, and during this time he will not be able to transfer
+his tokens.
 
-To mitigate this problem, for calculating a user's stake at time step \(t\), instead of using the raw ARG
-balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
-ARG balance:
+Unfortunately, there is a subtle problem with this approach. It is not clear that in a real world economic system
+how much of the main currency of the system can be locked and kept out of the circulation indefinitely. It seems that
+this amount for currencies like US dollar, is quite low comparing to the total market cap of the currency.
+This means that for a real world currency this type of staking mechanisms will result in putting the
+fate of the system in the hands of the owners of a small fraction of the total supply of a currency.
+
+To mitigate this problem, Argennon uses a hybrid approach for estimating the stake of a user.
+Every \texttt{stakingDuration} blocks, which is called a \emph{staking period}, Argennon calculates
+a \emph{trust value} for each user. The user's stake
+at time step \(t\), is estimated based on the user's trust value and his ARG balance:
 \[
-    S_{u,t} = \min (B_{u,t}, Trust_{u,t})
+    S_{u,t} = \min (B_{u,t}, Trust_{u,k}),
 \]
-Where:
+where:
 \begin{itemize}
@@
     \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
-    \item \(Trust_{u,t}\) is an estimated trust value for user \(u\) at time step \(t\).
+    \item \(Trust_{u,k}\) is an estimated trust value for user \(u\) at staking period \(k\).
 \end{itemize}
 
-The agreement protocol, at time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
-number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
-exponential moving average of the ARG balance of user \(u\) at time step \(t\).
+Argennon users can lock their ARG tokens in their account for any period of time. During this time a user
+will not be able to transfer his tokens and there is no way for cancelling a lock.
+The trust value of a user is calculated based on the amount of his locked tokens and the
+Exponential Moving Average (EMA) of his ARG balance:
+\[
+    Trust_{u,k} = \max (L_{u,k}, M_{u,t_k}),
+\]
+where
+\begin{itemize}
+    \item $L_{u,k}$ is the amount of locked tokens of user $u$, whose release time is \textbf{after the end} of
+    the staking period $k+1$.
+    \item $M_{u,t_k}$ is the Exponential Moving Average (EMA) of the ARG balance of user \(u\) at time step \(t_k\).
+    $t_k$ is the start time of the staking period $k$.
+\end{itemize}
 
-In our system a user who held ARGs and participated in the consensus for a long time is more trusted
+In Argennon a user who held ARGs and participated in the consensus for a long time is more trusted
 than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
@@
 
-For calculating the exponential moving average of a user's balance at time step \(t\), we can use the following
+For calculating the EMA of a user's balance at time step \(t\), we can use the following
 recursive formula:
@@
 \]
-Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
-degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
+where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\), which represents the
+degree of weighting decrease. A higher \(\alpha\) discounts older observations faster.
 
@@
 \]
-Where:
+where:
 \[
@@
 
-In our system a newly created account will not have voting power for some time, no matter how high its
-balance is. While this is a desirable property, in case a large proportion of total system tokens are
-transferred to newly created accounts, it can result in too much voting power for older accounts. This may decrease
-the degree of decentralization in our system.
+\subsection{The Recovery Protocol}\label{subsec:recovery}
 
-However, this situation is easily detectable by comparing the total stake of the system with the total balance of
-users. If after confirming a block the total stake of the system goes too low and we have:
-\[
-    \sum_{u}S_{u,t} < \gamma \sum_{u}B_{u,t}
-\]
-The protocol will perform a \emph{time shift} in the system: the time step of the system
-will be incremented for \(m\) steps while no blocks will be confirmed. This will increase the value of \(M_{u,t}\)
-for new accounts with a non-zero balance, giving them more influence in the agreement protocol.
-
-For calculating the value of \(m\) which determines the amount of time shift in the system, we should note that when
-\(B_{u,t} = B_{u, t-1} = B_u\), we can derive a simple recursive rule for the stake of a user:
-\[
-    S_{u,t} = (1 - \alpha) S_{u,t-1} + \alpha B_u
-\]
-Therefore, we have:
-\[
-    \sum_{u}S_{u,t} = (1 - \alpha) \sum_{u}S_{u,t - 1} + \alpha \sum_{u}B_u
-\]
-This equation shows that when the balance of users is not changing over time the total stake of the system is the
-exponential average of the total ARGs of the system. Consequently, when we shift the time for \(m\) steps, we can
-calculate the new total stake of the system from the following equation:
-
-\[
-    \sum_{u}S_{u,t+m} = (1 - \alpha)^{m}\sum_{u}S_{u,t} + [1 - (1 - \alpha)^{m}]\sum_{u}B_u
-\]
-Hence, if we want to increase the total stake of the system from \(\gamma \sum_{u}B_u\) to \(\lambda \sum_{u}B_u\),
-we can obtain \(m\) from the following formula, assuming \(\alpha\) is small enough:
-\[
-    m = \frac{1}{\alpha} \ln \left(\frac{1 - \gamma}{1 - \lambda}\right)
-\]
+\note{not yet written...}

### 2021-06-20 tex/sections/draft.tex
Add the Argennon consensus protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-03-29 tex/avm.tex
@@
 %! Author = aybehrouz
-%! Date = 2/28/21
 
-% Preamble
-\documentclass[a4paper]{report}
+In Argennon \emph{light} nodes do not keep a full copy of the AVM storage and for emulating the Argennon Virtual Machine
+( i.e.~validating transactions) they need to connect to a ZK-EDB and retrieve the required pages.
+Because light node need to be able to validate block certificates, they usually cache a large part of the staking
+database, and keep their cache updated to make sure they can keep themselves in sync with the blockchain.
 
-% Packages
-%\usepackage{amsmath}
-%\usepackage[a4paper]{geometry}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{amsfonts}
-\usepackage[ruled,boxed]{algorithm2e}
-\usepackage{algpseudocode}
-\usepackage{listings}
-\usepackage{amssymb}
-\usepackage{tcolorbox}
-\lstset{
-%  basicstyle=\ttfamily,
-    mathescape
-}
-%\binoppenalty=99999
-%\relpenalty=99999
-\newcommand{\note}[1] {
-    \begin{tcolorbox}[colframe=white,colback=white]
-        \emph{#1}
-        %\emph{#1}\medskip
-    \end{tcolorbox}
-}
+On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
+execution of the Argennon Virtual Machine. To do so, first they retrieve all storage pages they
+need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
+transactions included in the new block. This will modify some pages of the AVM storage, so they update the ZK-EDB
+commitments based on the modified pages and verify the commitments included in the new block. Validators also
+calculate and verify the commitment to the new block's transaction set.
 
-\title{The Argennon Virtual Machine}
-\author{aybehrouz}
-\date{January 2021}
+\note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDB servers will receive the new
+block, and they will update their database by emulating the AVM execution.}
 
-% Document
-\begin{document}
-    \maketitle
-    \tableofcontents
-
-    \chapter{Specification}\label{ch:specification}
-
-    \section{Introduction}\label{sec:introduction}
-
-    The Argennon Virtual Machine (AVM) is an abstract computing machine for executing Argennon's smart contracts. It
-    is designed in a way that it could be efficiently implemented in either hardware or software.
-
-    \section{The Structure of the Argennon Virtual Machine}\label{sec:the-structure-of-the-algorand-virtual-machine}
-
-    ...
-
-    \subsection{Data Types}\label{subsec:data-types}
-
-    The Argennon Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
-    and does not have to be done by the Argennon Virtual Machine itself.
-
-    The instruction set of the Argennon Virtual Machine distinguishes its operand types using instructions intended to
-    operate on values of specific types. For instance, \texttt{iadd} assumes that its operands are two 64-bit integers.
-
-    \subsection{The \texttt{PC} Register}\label{subsec:the-pc-register}
-
-    ...
-
-    \subsection{Call Stack}\label{subsec:call-stack}
-
-    A call stack contains the information that is needed for restoring the state of the invoker of a method.
-
-    \subsection{Memory Unit}\label{subsec:memory-unit}
-
-    The Argennon Virtual Machine has a byte addressable memory which is divided into separate segments. Every segment
-    belongs to a smart contract that has a unique \texttt{applicationID}. The AVM always has a single working memory
-    segment and memory locations outside its current working segment can not be accessed. The only instructions which
-    can change the current working segment are \texttt{invoke\_external}, \texttt{athrow} and return instructions.
-
-    Every smart contract has its own memory segment. Hence, there is no way for a smart contract to access another
-    smart contract's memory. Interaction between smart contracts is done using \texttt{invoke\_external} instruction,
-    and a smart contract can invoke methods of another smart contract by this instruction.
-
-    A memory segment consists of five data areas:
-    \begin{itemize}
-        \item Code Area
-        \item Constant Area
-        \item Local Frame
-        \item Operand Stack
-        \item Heap
-    \end{itemize}
-    All data areas except operand stack, have their own address space which starts from 0. Operand stack is a
-    last-in-first-out (LIFO) stack and is not addressable. Every memory access instruction operates on its specific
-    data areas.
-
-    \subsubsection{Code Area}
-
-    The code area of a segment contains the byte-code of the smart contract which owns that segment. The AVM has no
-    instructions for manipulating the code area. \textbf{Installing, removing and updating smart contracts need to
-    be done externally.}
-
-    \subsubsection{Constant Area}
-
-    The constant area of a segment contains several kinds of constants, ranging from user defined constants to method
-    address tables. A method address table stores method locations in the code area and their access type. The access
-    type of a method can be either \texttt{public} or \texttt{private}. The AVM has no instructions for modifying the
-    constant area.
-
-    \note{Only public methods can be invoked by \texttt{invoke\_external} instruction.}
-
-    \subsubsection{Local Frame}
-
-    A local frame is used to store methods parameters and local variables. A new frame is created each time a method
-    is invoked, and it is destroyed when its method invocation completes, whether the completion is normal or abrupt.
-
-    \subsubsection{Operand Stack}
-
-    Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too. AVM
-    instructions take operands from the operand stack, operate on them, and push the result back onto the operand
-    stack. An operand stack is destroyed when its owner method completes, whether that completion is normal or abrupt.
-
-    \subsubsection{Heap}
-
-    The heap of a segment is a persistent memory area which is divided into pages. Memory locations inside every page
-    have a separate address space that starts from 0, and each page can be referenced by an index that starts from 0. In
-    other words, the address of every memory location inside a heap is a pair of indices: \texttt{(pageIndex, offset)}.
-    Different pages of a heap do not need to be equally sized.
-
-    \note{The reason behind this paged design is that the heap is usually persisted using a block device. A heap with
-    a paged structure could expose the underlying block based nature of the persistence layer to the application
-    layer. In this way, the compiler or the programmer could better optimize the code for the persistence layer.}
-
-    \section{Instruction Set Summary}\label{sec:instruction-set-summary}
-
-    An Argennon Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
-    performed, followed by zero or more operands supplying arguments or data that are used by the operation.
-    \textbf{The number and size of the operands are determined solely by the opcode.}
-
-    \subsection{Method Invocation}\label{subsec:method-invocation}
-
-    The Argennon Virtual Machine has three types of method invocation:
-    \begin{itemize}
-        \item \texttt{invoke\_internal} invokes a method from the current running smart contract.
-        \item \texttt{invoke\_external} invokes a \texttt{public} method from another smart contract. It will change
-        the current memory segment to the segment of the invoked smart contract.
-        \item \texttt{invoke\_native} invokes a method that is not hosted by the Argennon virtual machine. By this
-        instruction, high performance native methods of the hosting machine could become available to AVM smart
-        contracts.
-    \end{itemize}
-
-    \note{In the future, we may need to add special instructions for invoking interface and virtual methods...}
-
-    Each time a method is invoked a new local frame and operand stack is created. The Argennon Virtual Machine uses
-    local frames to pass parameters on method invocation. On method invocation, any parameters are passed in
-    consecutive local variables stored in the method's local frame starting from local variable 0. The invoker of a
-    method writes the parameters in the local frame of the invoked method using \texttt{arg} instructions.
-
-    \subsubsection{Exceptions}
-
-    An exception is thrown programmatically using the \texttt{athrow} instruction. Exceptions can also be thrown by
-    various Argennon Virtual Machine instructions if they detect an abnormal condition. Some exceptions are not
-    catchable and will always abort the execution of the smart contract.
-
-    \subsubsection{Method Invocation Completion}
-
-    A method invocation completes normally if that invocation does not cause an exception to be thrown, either
-    directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
-    method completes normally, then a value may be returned to the invoking method. This occurs when the invoked
-    method executes one of the return instructions, the choice of which must be appropriate for the type of the value
-    being returned (if any). Execution then continues normally in the invoking method's local frame with the returned
-    value (if any) pushed onto the operand stack.
-
-    A method invocation completes abruptly if an exceptions is thrown and is not caught by the current method. A
-    method invocation that completes abruptly never returns a value to its invoker.
-
-    When a method completes, whether normally or abruptly, the call stack is used to restore the state of the invoker,
-    including its local frame and operand stack, with the \texttt{PC} register appropriately restored and incremented
-    to skip past the method invocation instruction. If the invoker was another smart contract, i.e.\ the invocation
-    was made by an \texttt{invoke\_external} instruction, the current memory segment will be changed to the invoker's
-    segment.
-
-    A thrown exception causes methods in the call stack to complete \textbf{abruptly} one by one, as long as the
-    \texttt{PC} register is not pointing to a \texttt{catch} instruction. The \texttt{catch} instruction acts like a
-    branch instruction that branches only if an exception is caught. \textbf{When an external method invocation
-    completes abruptly, before changing the current segment, all changes made to the heap area by that method
-    call, including changes made to other segments, will be rolled back.} So, when a method call completes abruptly,
-    that method call essentially has no effect on the AVM state.
-
-    \note{By using the \texttt{athrow} instruction properly, a programmer can make any method act like an atomic
-    operation.}
-
-    \subsection{Authorizing Operations}\label{subsec:authorizing-operations}
-
-    In blockchain applications, we usually need to authorize certain operations. For example, for sending an asset
-    from a user to another user, first we need to make sure that the sender has authorized this operation. The
-    Argennon virtual machine has no built in mechanism for authorizing operations, but it provides a rich set of
-    cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
-    passing signatures as parameters to methods, a programmer can implement the required logic for authorizing any
-    operation.
-
-    \note{The Argennon virtual machine has no instructions for issuing cryptographic signatures.}
-
-    In addition to signatures, a method can verify its invoker by using \texttt{get\_parent} instruction. This
-    instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
-    smart contract in the call stack. In other words, it returns the \texttt{applicationID} of the smart contract that
-    has invoked the current smart contract. (if any)
-
-    \subsection{Heap Allocation Instructions}\label{subsec:heap-allocation-instructions}
-
-    ...
-
-    \section{The AVM Standard Library}\label{sec:the-avm-standard-library}
-
-    ...
-
-    \chapter{Implementation}\label{ch:implementation}
-
-    \section{Persistence}\label{sec:persistence}
-
-    For implementing the persistence layer of the AVM, we assume that we have access to an updatable zero-knowledge
-    elementary database (ZK-EDB) with the following properties:
-
-    \begin{itemize}
-        \item The ZK-EDB contains a mapping from a set of keys to a set of values.
-        \item Every state of the database has a commitment \(C\).
-        \item The ZK-EDB has a method \((D, p) = get(x)\), where \(x\) is a key and \(D\) is the associated data
-        with \(x\), and \(p\) is a proof.
-        \item A user can use \(C\) and \(p\) to verify that \(D\) is really associated with \(x\), and \(D\) is not
-        altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB\@.
-        \item Having \(p\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
-        associated with \(x\) instead of \(D\).
-    \end{itemize}
-
-    We use a ZK-EDB for storing the AVM heap. We include the commitment of the current state of this DB in every
-    block of the Argennon blockchain, so ZK-EDB servers need not be trusted servers.
-
-    Every page of AVM heap will be stored with a key of the form: \texttt{applicationID|pageIndex} (the \texttt{|}
-    operator concatenates two numbers). Nodes do not keep a full copy of the AVM heap and for validating block
-    certificates or emulating the AVM ( i.e.~validating transactions) they need to connect to a ZK-EDB and retrieve
-    the required pages of AVM heap. For better performance, nodes keep a cache of heap pages to
-    reduce the amount of ZK-EDB access.
-
-    We also use a ZK-EDB for storing the code area of each segment, and we include the commitment of this DB in every
-    block. Every code area will be divided into blocks and every block will be stored in the DB with
-    \texttt{applicationID|blockID} as its key. Like heap pages, nodes keep a cache of code area blocks.
-
-    \note{Unlike heap pages, the AVM is not aware of different blocks of code area.}
-
-    \section{Transactions}\label{sec:transactions}
-
-    Argennon has four types of transaction:
-
-    \begin{itemize}
-        \item \texttt{avmCall} essentially is an \texttt{invoke\_external} instruction that invokes a method from an
-        AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
-        assets, including ARGs, is done by these transactions.
-        \item \texttt{installApp} installs an AVM smart contract and determines the update policy of the smart
-        contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
-        on.
-        \item \texttt{unInstallApp} removes an AVM smart contract.
-        \item \texttt{updateApp} updates an AVM smart contract.
-    \end{itemize}
-
-    All types of Argennon transactions contain an \texttt{invoke\_external} instruction which calls a special method
-    from ARG smart contract that transfers the proposed fee of the transaction in ARGs from a sender account to the
-    fee sink accounts.
-
-    Every transaction is required to exactly specify what heap locations or code area addresses it will access. This
-    enables validators to start retrieving the required memory blocks from available ZK-EDB servers as soon as they
-    see a transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to
-    access a memory location that is not included in its access lists, will be rejected. Users could use smart contract
-    oracles to predict the list of memory blocks their transactions need. See Section~\ref{sec:smart-contract-oracle}
-    for more details.
-
-    \section{Blockchain}\label{sec:blockchain}
-
-    Every block of the Argennon blockchain corresponds to a set of transactions. We store the commitment of this
+The only information that all Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
+blockchain.
+We store the commitment of this
 transaction set in every block, but we don't keep the set itself. To be able to detect replay attacks, we require
@@
 have the problem of "un-removable empty accounts" like Ethereum.}
-
-    The only information that Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
-    blockchain. Every block of the Argennon blockchain contains the following information:
-
-    \begin{center}
-        \begin{tabular}{||c||}
-            \hline
-            Block \\ [0.5ex]
-            \hline\hline
-            commitment to the ZK-EDB storing heap pages \\ [0.7ex]
-            commitment to the ZK-EDB storing code areas \\ [0.7ex]
-            commitment to the set of transactions       \\ [0.7ex]
-            previous block hash                         \\ [0.7ex]
-            random seed                                 \\ [0.7ex]
-            \hline
-        \end{tabular}
-    \end{center}
-
-    For confirming a new block, nodes that are not validators only need to verify the block certificate. For
-    verifying a block certificate, a node needs to know the ARG balances of validators, but it doesn't need to
-    emulate the AVM execution.
-
-    On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
-    execution of the Argennon virtual machine. To do so, first they retrieve all heap pages and code area blocks they
-    need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
-    transactions included in the new block. This will modify some pages of the AVM memory, so they update the ZK-EDB
-    commitments based on the modified pages and verify the commitments included in the new block. Validators also
-    calculate and verify the commitment to the new block's transaction set.
-
-    \note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDB servers will receive the new
-    block, and they will update their database by emulating the AVM execution.}
-
-    \section{Consensus}\label{sec:consensus}
-
-    \note{It will be a randomized proof of stake which is under development....}
-
-    \subsection{Estimating A User's Stake}\label{subsec:estimating-a-user's-stake}
-
-    In a proof of stake system the influence of a user in the consensus protocol should be proportional to the amount
-    of stake the user has in the system. Conventionally in these systems, for estimating a user's stake, we use the
-    amount of native system tokens the user is holding. Unfortunately, one problem with this approach is that a
-    strong attacker may be able to obtain a considerable amount of system tokens, for example by borrowing from a
-    DEFI application, and use this stake to attack the system.
-
-    To mitigate this problem, for calculating a user's stake at time step \(t\), instead of using the raw ARG
-    balance, we use the minimum of a \emph{trust value} the system has calculated for the user and the user's
-    ARG balance:
-    \[
-        S_{u,t} = \min (B_{u,t}, Trust_{u,t})
-    \]
-    Where:
-    \begin{itemize}
-        \item \(S_{u,t}\) is the stake of user \(u\) at time step \(t\).
-        \item \(B_{u,t}\) is the ARG balance of user \(u\) at time step \(t\).
-        \item \(Trust_{u,t}\) is an estimated trust value for user \(u\) at time step \(t\).
-    \end{itemize}
-
-    The agreement protocol, at time step \(t\), will use \(\sum_{u}S_{u,t}\) to determine the required
-    number of votes for the confirmation of a block, and we let \(Trust_{u,t} = M_{u,t}\), where \(M_{u,t}\) is the
-    exponential moving average of the ARG balance of user \(u\) at time step \(t\).
-
-    In our system a user who held ARGs and participated in the consensus for a long time is more trusted
-    than a user with a higher balance whose balance has increased recently. An attacker who has obtained a large
-    amount of ARGs, also needs to hold them for a long period of time before being able to attack the system.
-
-    For calculating the exponential moving average of a user's balance at time step \(t\), we can use the following
-    recursive formula:
-    \[
-        M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
-    \]
-    Where the coefficient \(\alpha\) is a constant smoothing factor between \(0\) and \(1\) which represents the
-    degree of weighting decrease, A higher \(\alpha\) discounts older observations faster.
-
-    Usually an account balance will not change in every time step, and we can use older values of EMA for calculating
-    \(M_{u,t}\): (In the following equations the \(u\) subscript is dropped for simplicity)
-    \[
-        M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B
-    \]
-    Where:
-    \[
-        B = B_{k+1} = B_{k+2} = \dots = B_{t}
-    \]
-    We know that when \(|nx| \ll 1\) we can use the binomial approximation \({(1 + x)^n \approx 1 + nx}\). So, we can
-    further simplify this formula:
-    \[
-        M_{t} = M_{k} + (t - k) \alpha (B - M_{k})
-    \]
-
-    For choosing the value of \(\alpha\) we can consider the number of time steps that the trust value of a user needs
-    for reaching a specified fraction of his account balance. We know that for large \(n\) and \(|x| < 1\) we have
-    \((1 + x)^n \approx e^{nx}\), so by letting \(M_{u,k} = 0\) and \(n = t - k\) we can write:
-    \[
-        \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{B}\right)}{n}
-    \]
-    The value of \(\alpha\) for a desired configuration can be calculated by this equation. For instance, we could
-    calculate the \(\alpha\) for a relatively good configuration in which \(M_{n+k} = 0.8B\) and \(n\) equals to the
-    number of time steps of 10 years.
 
@@
 
-    \section{Incentive mechanism}\label{sec:incentive-mechanism}
+An Argennon block certificate is not a part of the block data. This means that the Argennon blockchain does not have a
+unique certificate sequence. Any node may have its own valid certificate sequence based on the signatures
+it has collected over time.
 
-    \subsection{Transaction Fee}\label{subsec:transaction-fee}
+\subsection{Memory Shards}\label{subsec:memory-shards}
 
-    Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
-    special method from ARG smart contract. This method will transfer the proposed fee of the transaction in ARGs
-    from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
-    execution fees and \texttt{dbFeeSink} collects fees for ZK-EDB servers. The Protocol decides how to distribute the
-    transaction fee between these two fee sink accounts.
+In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{shards}.
+Each memory shard can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
+consensus on new values of the commitment of any shard can be achieved by a different voting committee.
 
-    When a block is added to the blockchain, the proposer of that block will receive a share of the block fees.
-    Consequently, a block proposer is always incentivized to include more transactions in his block. However, if he
-    puts too many transactions in his block and the validation of the block becomes too difficult, some validators
-    may not be able to validate all transactions on time. If a validator can not validate a block in the required
-    time, he will consider the block invalid. So, when a proposed block contains too many transactions, the network
-    may reach consensus on another block, and the proposer of that block will not receive any fees. As a result, a
-    proposer is incentivized to use network transaction capacity optimally.
-
-    On the other hand, we believe that the proposer does not have enough incentives for optimizing the storage size
-    of the transaction set. Therefore, we require that \textbf{the size of the transaction set of every block in
-    bytes be lower than a certain threshold.}
-
-    Validators need to spend resources for validating transactions. When a validator starts the emulation of the AVM
-    to validate a transaction, solely from the code he can't predict the time the execution will finish. This will
-    give an adversary an opportunity to attack the network by broadcasting transactions that never ends. Since,
-    validators can not finish the execution of these transactions, the network will not be able to charge the
-    attacker any fees, and he would be able to waste validators resources for free.
-
-    To mitigate this problem, we require that every transaction specify a cap for all the resources it needs. This
-    will include memory, network and processor related resources. Also, the protocol defines an execution cost for
-    every AVM instruction reflecting the amount of resources its emulation needs. This will define a standard way for
-    measuring the execution cost of any \texttt{avmCall} transaction. Every \texttt{avmCall} transaction is required
-    to specify a maximum execution cost. If during emulation it reaches this maximum cost, the transaction will be
-    considered failed and the network can receive the proposed fee of that transaction.
-
-    Every \texttt{avmCall} transaction is required to provide the following information as an upper bound for the
-    resources it needs:
-
-    \begin{itemize}
-        \item Execution cost
-        \item A list of heap/code-area locations for reading
-        \item A list of heap locations for writing
-        \item A list of heap pages it will deallocate
-        \item Number and size of heap pages it will allocate
-    \end{itemize}
-
-    If a transaction tries to violate any of these predefined limitations, for example, if it tries to read a memory
-    location that is not included in its reading list, it will be considered failed and the network can receive the
-    proposed fee of that transaction.
-
-    \note{A transaction always pays all of its proposed fee, no matter how much of its predefined resources were not
-    used in the final emulation.}
-
-    \subsection{Incentives for ZK-EDB Servers}\label{subsec:zk-edb-servers}
-
-    The incentive mechanism for ZK-EDB servers should have the following properties:
-
-    \begin{itemize}
-        \item It incentivizes storing all memory blocks, whether a heap page or a code area block, and not only those
-        which are used more frequently.
-        \item It incentivizes ZK-EDB servers to actively provide the required memory blocks for validators.
-        \item Making more accounts will not provide any advantages for a ZK-EDB server.
-    \end{itemize}
-
-    For our incentive mechanism, we require that every time a validator receives a memory block from a ZK-EDB, after
-    validating the data, he give a receipt to the ZK-EDB. In this receipt the validator signs the following information:
-
-    \begin{itemize}
-        \item \texttt{ownerAddr}: the ARG address of the ZK-EDB\@.
-        \item \texttt{receivedBlockID}: the ID of the received memory block.
-        \item \texttt{round}: the current round number.
-    \end{itemize}
-
-    \note{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}
-
-    To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ARGs from
-    \texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
-    between all \emph{winning tickets} of the lottery.
-
-    \note{One ZK-EDB server could own multiple winning tickets in a round.}
-
-    To run this lottery, every round, based on the current block seed, a collection of \emph{valid} receipts will be
-    selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in round \(r\) if:
-
-    \begin{itemize}
-        \item The signer was a validator in the round \(r - 1\) and voted for the agreed-upon block.
-        \item The data block in the receipt was needed for validating the \textbf{previous} block.
-        \item The receipt round number is \(r - 1\).
-        \item The signer did not sign a receipt for the same data block for two different ZK-EDBs in the previous round.
-    \end{itemize}
-    For selecting the winning receipts we could use a random generator:
-    \begin{verbatim}
-IF random(seed|validatorPK|receivedBlockID) < winProbability THEN
-    the receipt issued by validatorPK for receivedBlockID is a winner
-    \end{verbatim}
-    \begin{itemize}
-        \item \texttt{random()} produces uniform random numbers between 0 and 1, using its input argument as a seed.
-        \item \texttt{validatorPK} is the public key of the signer of the receipt.
-        \item \texttt{receivedBlockID} is the ID of the memory block that the receipt was issued for.
-        \item \texttt{winProbability} is the probability of winning in every round.
-        \item \texttt{seed} is the current block seed.
-        \item \texttt{|} is a concatenation operator.
-    \end{itemize}
-
-    \note{The winners of the lottery were validators one round before the lottery round.}
-
-    Also, based on the current block seed, a random memory block, whether a heap page or a code area block, is
-    selected as the challenge of the round. A ZK-EDB that owns a winning receipt needs to broadcast a \emph{winning
-    ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
-    challenge. Solving a round challenge requires the content of the memory block which was selected as the round
-    challenge. This will encourage ZK-EDBs to store all memory blocks.
-
-    A possible choice for the challenge solution could be the cryptographic hash of the content of the challenge
-    memory block combined with the ZK-EDB ARG address: \texttt{hash(challenge.content|ownerAddr)}
-
-    The winning tickets of the lottery of round \(r\) need to be included in the block of the round \(r\),
-    otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
-    \(r\) will be done in the round \(r + 1\). This way, \textbf{the content of the challenge memory block could be
-    kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.
-
-    \subsection{Memory Allocation and De-allocation Fee}\label{subsec:memory-allocation-and-de-allocation}
-
-    Every \(k\) round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
-    allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
-    address of the smart contract.
-
-    To determine the price of AVM memory, Every \(k\) round, the protocol calculates \texttt{dbFee} and
-    \texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
-    \texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
-    the protocol considers the total size of all the memory pages that were accessed for either reading or writing
-    during a time period. These two values will be calculated for the last \(k\) rounds and the price per byte of
-    AVM memory will be a linear function of \texttt{dbFee/memTraffic}
-
-    When a smart contract executes a heap de-allocation instruction, the protocol will refund the cost of
-    de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
-    calculates the refunded amount based on the average price the smart contract had paid for that allocated memory.
-    This will prevent smart contracts from profit taking by trading memory with the protocol.
-
-    \section{Concurrency}\label{sec:concurrency}
-
-    \subsection{Memory Dependency Graph}\label{subsec:memory-dependency-graph}
-
-    Every block of the Argennon blockchain contains a list of transactions. This list is an ordered list and the
-    effect of its contained transactions must be applied to the AVM state sequentially as they appear in the ordered
-    list. This ordering is solely chosen by the block proposer, and users should not have any assumptions about
-    the ordering of transactions in a block.
-
-    The fact that block transactions constitute a sequential list, does not mean they can not be executed and applied
-    to the AVM state concurrently. Many transactions are actually independent and the order of their execution does not
-    matter. These transactions can be safely validated in parallel by validators.
-
-    A transaction can change the AVM state by modifying either the code area or the AVM heap. In Argennon, all
-    transactions declare the list of memory locations they want to read or write. This will enable us to determine the
-    independent sets of transactions which can be executed in parallel. To do so, we define the \emph{memory dependency
-    graph} \(G_d\) as follows:
-
-    \begin{itemize}
-        \item \(G_d\) is an undirected graph.
-        \item Every vertex in \(G_d\) corresponds to a transaction and vice versa.
-        \item Vertices \(u\) and \(v\) are adjacent in \(G_d\) if and only if \(u\) has a memory location \(L\) in its
-        writing list and \(v\) has \(L\) in either its writing list or its reading list.
-    \end{itemize}
-
-    If we consider a proper vertex coloring of \(G_d\), every color class will give us an independent set of
-    transactions which can be executed concurrently. To achieve the highest parallelization, we need to color \(G_d\)
-    with minimum number of colors. Thus, the \emph{chromatic number} of the memory dependency graph shows how good a
-    transaction set could be run concurrently.
-
-    Graph coloring is computationally NP-hard. However, in our use case we don't need to necessarily find an optimal
-    solution. An approximate greedy algorithm will perform well enough in most circumstances.
-
-    After constructing the memory dependency graph, we can use it to construct the
-    \emph{execution DAG} of transactions. The execution DAG of transaction set \(T\) is a directed acyclic
-    graph \(G_e = (V_e,E_e)\) which has the \emph{execution invariance} property:
-    \begin{itemize}
-        \item Every vertex in \(V_e\) corresponds to a transaction in \(T\) and vice versa.
-        \item Executing the transactions of \(T\) in any order that \emph{respects} \(G_e\) will result in
-        the same AVM state.
-        \begin{itemize}
-            \item An ordering of transactions of \(T\) respects \(G_e\) if for every directed edge \((u,v) \in E_e\)
-            the transaction \(u\) comes before the transaction \(v\) in the ordering.
-        \end{itemize}
-    \end{itemize}
-
-    Having the execution DAG of a set of transactions, using Algorithm~\ref{alg:exec_dag}, we can apply the transaction
-    set to the AVM state concurrently, using multiple processor, while we can be sure that the resulted AVM state will
-    always be the same no matter how many processor we have used.
-
-    %##\includegraphics[width=17cm]{../img/Alg1s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Ready}{$R_e$}\SetKwData{V}{$v_f$}\SetKwData{Graph}{$G_e$}\SetKwData{Vertices}{$V$}\SetKwData
-        {Txns}{$T$}
-        \KwData{The execution dag $\Graph = (\Vertices,E)$ of transaction set \Txns}
-        \KwResult{The state of the AVM after applying \Txns with any ordering respecting \Graph}
-        \BlankLine
-        \Ready $\gets$ the set of all vertices of \Vertices with in degree 0\;
-        \While{$\Vertices \neq \varnothing$}
-        {
-            wait until a new free processor is available\;
-            \If{the execution of a transaction was finished}
-            {
-                remove the vertex of the finished transaction \V from \Graph\;
-                \For{each vertex $u \in Adj[\V]$}
-                {
-                    \If{$u$ has zero in degree}
-                    {
-                        $\Ready \gets \Ready \cup u$\;
-                    }
-                }
-            }
-            \If{$\Ready \neq \varnothing$}
-            {
-                remove a vertex from \Ready and assign it to a processor\;
-            }
-        }
-        \caption{executing DAG transactions}\label{alg:exec_dag}
-    \end{algorithm}
-
-    By replacing every undirected edge of a memory dependency graph with a directed edge in such a way that the
-    resulted graph has no cycles, we will obtain a valid execution DAG. Thus, from a memory dependency graph different
-    execution DAGs can be constructed with different levels of parallelization ability.
-
-    If we assume that we have unlimited number of processors and all transactions take equal time for executing, it
-    can be shown that by providing a minimal graph coloring to Algorithm~\ref{alg:gen_dag} as input, the resulted
-    DAG will be optimal, in the sense that it results in the minimum overall execution time.
-
-    %##\includegraphics[width=17cm]{../img/Alg2s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{Txns}{$T$}\SetKwData{Gd}{$G_d=(V_d,E_d)$}
-        \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{The memory dependency graph \Gd of transaction set \Txns\\A proper coloring of $G_d$}
-        \Output{An execution dag $G_e=(V_e,E_e)$ for the transaction set \Txns}
-        \BlankLine
-        $V_e \gets V_d$\;
-        $E_e \gets \varnothing$\;
-        define a total order on colors of $G_d$\;
-        \For{each edge $\{u,v\} \in E_d$}
-        {
-            \eIf{$color[u] < color[v]$}
-            {
-                $E_e \gets E_e \cup (u,v)$\;
-            }{
-                $E_e \gets E_e \cup (v,u)$\;
-            }
-        }
-        \caption{Constructing an execution DAG}\label{alg:gen_dag}
-    \end{algorithm}
-
-    The block proposer is responsible for proposing an efficient execution DAG alongside his proposed block. This
-    execution DAG will determine the ordering of block transactions and help validators to validate transactions in
-    parallel. Since with better parallelization a block can contain more transactions, a proposer is incentivized enough
-    to find a good execution DAG for transactions.
-
-    \subsection{Memory Spooling}\label{subsec:spooling}
-
-    When two transactions are dependant and they are connected with an edge \((u,v)\) in the execution DAG,
-    the transaction \(u\) needs to be run before the transaction \(v\). However, if \(v\) does not read any
-    memory locations that \(u\) modifies, we can run \(u\) and \(v\) in parallel. We just need to make sure
-    \(u\) does not see any changes \(v\) is making in AVM memory. This can be done by appropriate versioning
-    of the memory locations which is shared between \(u\) and \(v\). We call this method \emph{memory spooling}.
-    After enabling memory spooling between two transactions the edge connecting them can be safely removed from the
-    execution DAG\@.
-
-    \subsection{Concurrent Counters}\label{subsec:concurrent-counters}
-
-    We know that in Argennon every transaction needs to transfer its proposed fee to the \texttt{feeSink} accounts
-    first. This essentially makes every transaction a reader and a writer of the memory locations which store the
-    balance record of the \texttt{feeSink} accounts. As a result, all transactions in Argennon will be dependant and
-    parallelism will be completely impossible. Actually, any account that is highly active, for example the account
-    of an exchange or a payment processor, could become a concurrency bottleneck in our system which makes all
-    transactions interacting with them dependant.
-
-    This problem can be easily solved by using a concurrent counter for storing the balance record of this type of
-    accounts. A concurrent counter is a data structure which improves concurrency by using multiple memory locations for
-    storing a single counter. The value of the concurrent counter is equal to the sum of its sub counters and it can
-    be incremented or decremented by incrementing/decrementing any of the sub counters. This way, a concurrent
-    counter trades concurrency with memory usage.
-
-    Algorithm~\ref{alg:CC} implements a concurrent counter which returns an error when the value of the counter
-    becomes negative.
-
-    %##\includegraphics[width=17cm]{../img/Alg3s.png}
-    \begin{algorithm}
-        \DontPrintSemicolon
-        \SetKwData{CC}{Counter}
-        \SetKwFunction{Inc}{Increment}\SetKwFunction{Dec}{Decrement}\SetKwFunction{AtomInc}{AtomicIncrement}
-        \SetKwFunction{AtomDec}{AtomicDecrement}\SetKwFunction{AtomSet}{AtomicSet}\SetKwFunction{Get}{GetValue}
-        \SetKwFunction{Acquire}{Lock.Acquire}\SetKwFunction{Release}{Lock.Release}
-        \SetKwProg{Fn}{Function}{}{}
-        \Fn{\Get{\CC}}
-        {
-            $s \gets 0$\;
-            \Acquire{}\;
-            \For{$i \gets 0$ \KwTo $\CC.size - 1$}
-            {
-                $s \gets s + \CC.cell[i]$\;
-            }
-            \Release{}\;
-            \KwRet{s}\;
-        }
-        \BlankLine
-        \Fn{\Inc{\CC, value, seed}}
-        {
-            $i \gets seed \bmod \CC.size$\;
-            \AtomInc{$\CC.cell[i]$, value}\;
-        }
-        \BlankLine
-        \Fn{\Dec{\CC, value, seed, attempt}}
-        {
-            \If {attempt = \CC.size}
-            {
-                restore \CC by adding back the subtracted value\;
-                \KwRet{Error}\;
-            }
-            $i \gets seed \bmod \CC.size$\;
-            $i \gets (i + attempt) \bmod \CC.size$\;
-            \eIf {$\CC.cell[i] \geq value$}
-            {
-                \AtomDec{$\CC.cell[i]$, value}\;
-            }{
-                $r \gets value - \CC.cell[i]$\;
-                \AtomSet{$\CC.cell[i]$, $0$}\;
-                \Dec{\CC, r, seed, $attempt + 1$}\;
-            }
-        }
-        \caption{Concurrent counter}\label{alg:CC}
-    \end{algorithm}
-
-    It should be noted that in a blockchain application we don't have concurrent threads and therefore we don't need
-    atomic functions. For usage in a smart contract, the atomic functions of this pseudocode can be implemented like
-    normal functions.
-
-    Concurrent counter data structure is a part of the AVM standard library, and any smart contract can use this data
-    structure for storing the balance record of highly active accounts.
-
-    \subsection{Memory Chunks}\label{subsec:memory-chunks}
-
-    In order to further increase the concurrency level of Argennon, we can divide the AVM memory into \emph{chunks}.
-    Each memory chunk can be persisted using a different ZK-EDB, hence having its own commitment. Then, the
-    consensus on new values of the commitment of any chunk can be achieved by a different voting committee.
-
-    If a transaction does not modify a memory chunk and in the transaction ordering of the block it comes after
-    any transaction which modifies that chunk, then the execution of that transaction is not needed for calculating
-    the new commitment of the chunk. Consequently, the voting committee of that memory chunk can safely ignore such a
+If a transaction does not modify a memory shard and in the transaction ordering of the block it comes after
+any transaction which modifies that shard, then the execution of that transaction is not needed for calculating
+the new commitment of the shard. Consequently, the voting committee of that memory shard can safely ignore such a
 transaction. The execution DAG of transactions can be used for finding and pruning these transactions as
@@
     \DontPrintSemicolon
-        \SetKwData{V}{$v_0$}\SetKwData{Graph}{$G_e$}\SetKwData{Chunk}{$C_k$}\SetKwData{Txns}{$T$}
+    \SetKwData{V}{$v_0$}\SetKwData{Graph}{$G_e$}\SetKwData{Shard}{$Sh_k$}\SetKwData{Txns}{$T$}
     \SetKwFunction{RPrune}{RecursivePrune}
@@
     \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
-        \Input{An execution dag \Graph and a memory chunk \Chunk}
-        \Output{The pruned execution dag based on \Chunk}
+    \Input{An execution dag \Graph and a memory shard \Shard}
+    \Output{The pruned execution dag based on \Shard}
     \BlankLine
@@
     {
-            \If{$Adj[\V] = \varnothing$ {\bf and} \V does not modify \Chunk}
+        \If{$Adj[\V] = \varnothing$ {\bf and} \V does not modify \Shard}
         {
@@
 
-    If we choose chunks in a way that most transactions only modify memory locations of one chunk,
+If we choose shards in a way that most transactions only modify memory locations of one shard,
 likely many transactions of a block only need to be validated by one voting committee and can be validated in

### 2021-06-20 tex/sections/infra.tex
Add the Argennon consensus protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-06-08 tex/sections/persistence.tex
@@
 
+\section{Persistence Layer}\label{sec:persistence-layer}
+
 The Argennon Virtual Machine has two persistent memory areas: \emph{method area}, and \emph{heap}. Method area stores
-method byte codes\footnote{also it stores constant area blocks.}, and heap stores memory chunks. Both of these
-data elements, methods and chunks, can be considered as continuous pieces of byte addressable memory. Throughout this
+bytecodes of methods\footnote{also it stores constant area blocks.}, and heap stores memory chunks. Both of these
+data elements, bytecodes and chunks, can be considered as continuous pieces of byte addressable memory. Throughout this
 chapter, we shall call these data elements \emph{objects}.
 
-In the AVM persistence layer, similar objects are clustered together and constitute a bigger data element which we call a
-\emph{page}.\footnote{we avoided calling them clusters, because usually a cluster refers to a \emph{set}. AVM object
+\subsection{Storage Pages}\label{subsec:pages}
+
+In the AVM persistence layer, similar objects are clustered together and constitute a bigger data element which we call
+a \emph{page}.\footnote{we avoided calling them clusters, because usually a cluster refers to a \emph{set}. AVM object
 clusters are not sets. They are ordered lists, like a page containing an ordered list of words or sentences.}
@@
 A page of the AVM storage should contain objects that have very similar access pattern. We expect that when a page
-is needed for validating a block, almost all of its objects are needed. We also prefer that the objects
-are needed for the same access type. In other words, the objects of a page are chosen in a way that
+is needed for validating a block, almost all of its objects are needed for either reading or writing. We also prefer
+that the objects are needed for the same access type. In other words, the objects of a page are chosen in a way that
 for validating a block, we usually need to either read all of them or modify\footnote{and probably read.} all of them.
 
-Pages of the AVM storage space are persisted using updatable zero-knowledge elementary databases (ZK-EDB). Argennon
-has three zero-knowledge databases. \emph{staking} database, which stores all the data that is associated with
-the Argennon agreement protocol. \emph{method} database, which stores the AVM method area, and \emph{heap} database,
+\subsection{Zero-knowledge Databases}\label{subsec:zk-edb}
+
+Pages of the AVM storage are persisted using updatable zero-knowledge elementary databases (ZK-EDB). Argennon
+has three zero-knowledge databases: \emph{staking} database, which stores all the data that is associated with
+the Argennon consensus protocol. \emph{method} database, which stores the AVM method area, and \emph{heap} database,
 which stores the AVM heap. The commitment of these three ZK-EDBs are included in every block of the Argennon blockchain.
@@
     \item Every state of the database has a commitment \(C\).
-    \item The ZK-EDB has a method \((D, \pi) = get(x)\), where \(x\) is a key and \(D\) is the associated data
+    \item The ZK-EDB has a method \((D, \pi) = \text{get}(x)\), where \(x\) is a key and \(D\) is the associated data
     with \(x\), and \(\pi\) is a proof.
@@
 
-Every page of the AVM storage is stored with an index as its key: \texttt{pageIndex}. The \texttt{pageIndex} is
-required to be smaller than a certain value determined by the
-protocol.\footnote{this facilitates the usage of ZK-EDBs that use vector commitments.}
-Therefore, the AVM clustering algorithm tries to reuse indices and keep the number of used indices as low as
-possible.
+Pages of the AVM storage are stored in the ZK-EDBs, with an index: \texttt{pageIndex} as their key.
+The \texttt{pageIndex} is required to be smaller than a certain value, determined by the
+protocol, to facilitate the usage of ZK-EDBs that are based on vector commitments.
+For this reason, the AVM clustering algorithm always tries to reuse indices and keep the number of used indices
+as low as possible.
 
-In Argennon \emph{light} nodes do not keep a full copy of the AVM storage and for emulating the Argennon Virtual Machine
-( i.e.~validating transactions) they need to connect to a ZK-EDB and retrieve the required pages.
-Because light node need to be able to validate block certificates, they usually cache a large part of the staking
-database, and keep their cache updated to make sure they can keep themselves in sync with the blockchain.
-
-The commitments of the AVM databases are dependant to the shape of clusters of data objects. So,
-the clustering algorithm of the AVM storage has to be a part of the Argennon agreement protocol. Clustering of
-objects is one of the tasks that a block proposer performs in Argennon.
+The commitments of the AVM ZK-EDBs are affected by the way the clustering of data objects is done. So,
+the clustering algorithm has to be a part of the Argennon consensus protocol.
 
@@
 
-A block proposer usually obtains clustering directives from a third party source that enjoys a considerable amount of
-computational power.\footnote{we can say the AVM clustering algorithm is essentially off-chain.} This federated approach
-does not affect Argennon security, because the integrity of a database can not be altered by clustering directives.
+A block proposer could obtain clustering directives from any third party source\footnote{we can say the AVM clustering
+algorithm is essentially off-chain.}. This will not
+affect Argennon security, since the integrity of a database can not be altered by clustering directives.
 Those directives can only affect the performance of the argennon blockchain, and directives of a single block can
-not change the performance much.
+not affect the performance considerably.
+
+\subsubsection[Implementation]{Implementation Based on Vector Commitments}\label{subsubsec:impl-zk-edbs}
+
+\note{not yet written...}
+
+\subsection{Object Clustering Algorithm}\label{subsec:clustering}
+
+\note{not yet written...}
+
+\section{Networking Layer}\label{sec:networking-layer}
+
+Unlike conventional blockchains, Argennon does not use a p2p network architecture. Instead, it uses a
+client-server topology, based on a permission-less list of ZK-EDB servers. ZK-EDB servers are a
+crucial part of the Argennon ecosystem, and they form the backbone of the Argennon networking layer.
+
+\note{not yet written...}
+
+\subsection{Censorship Resilient Mode}\label{subsec:cens-res-mode}
+
+\note{not yet written...}

### 2021-06-21 tex/sections/blockchain.tex
Add the forgotten parts
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
+\note{Explicit key registration enables Argennon to decouple cryptography from blockchain design. In this way,
+    if the used cryptographic algorithms turn out to be insecure for some reason, for example because
+    of the introduction of quantum computers, the cryptographic algorithms could be easily upgraded.}
+
 \section{Transactions}\label{sec:transactions}
@@
 
+\subsection{Authorization}\label{subsec:txn-auth}
+
 Argennon transactions do not have a sender. The authorization of the requested operation is always done by checking the
 digital signatures that are provided as a part of the argument list to the root smart contract method.
+
+While every block of the Argennon blockchain stores the commitment of the transaction list, Argennon does not enforce
+storage of the transaction history. To be able to detect replay attacks, we require
+every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
+and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
+must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
+a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
+if a signature of the same user with a bigger or equal nonce is already used
+(i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
+nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
+
+To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. When the
+difference between \texttt{issuance} component of this nonce and the current round becomes bigger than the
+maximum allowed lifetime of a signature, this information can be safely deleted.\footnote{in some conventional
+blockchains, the nonce data can never be deleted, even if the account has zero balance and is not used anymore.}
+
+\subsection{Resource Declaration}\label{subsec:resource-declaration}
 
@@
 
+\section{Concurrency}\label{sec:concurrency}
+\input{sections/concurrency}
+
 \section{Governance}\label{sec:governance}
-
 The Argennon Decentralized Autonomous Governance system (ADAGs)
 \note{not yet written...}
-
-\section{Concurrency}\label{sec:concurrency}
-\input{sections/concurrency}

### 2021-06-21 tex/sections/consensus.tex
Add the forgotten parts
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-More precisely, to decide about signing the block certificate, every honest member of the validators
-committee iterates over blocks of the validation set, and selects a block with
+More precisely, to decide about signing the block certificate, a honest member of the validators
+committee iterates over every block of the validation set, and selects a block for validation with
 probability $p_u$. In this way, a member of the validators committee will select $p_{u}m$ blocks in expectation.
@@
 where $\mu$ is the average stake of committee members and $s_u$ is the stake of user $u$.
+
+The value of $m$ is determined by the ADAGs. but it can never be set higher than $100$. This way, it is guaranteed
+that on average, any block of the Argennon blockchain is validated by at least $0.005$ of the total ARG supply.
 
@@
 \[
-    Trust_{u,k} = \max (L_{u,k}, M_{u,t_k}),
+    Trust_{u,k} = L_{u,k} + M_{u,t_k}\ ,
 \]
@@
 \subsection{The Recovery Protocol}\label{subsec:recovery}
+\note{not yet written...}
 
+\subsection{Analysis}\label{subsec:consensus-math}
 \note{not yet written...}

### 2021-06-21 tex/A.tex
Update persistence section
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Update the rest of the document
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Add the pronunciation note
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Add the Argennon consensus protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Refactor project file structure
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-% Preamble
 \documentclass[11pt, a4paper]{report}
@@
 % Packages
-%\usepackage{amsmath}
+\usepackage{tipa}
 \usepackage{graphicx}
@@
 
+    \chapter{Infrastructure}\label{ch:infrastructure}
+
+    \section{Persistence Layer}\label{sec:persistence-layer}
+    \input{sections/persistence}
+
+    \section{Networking Layer}\label{sec:networking-layer}
+    Unlike conventional blockchains, Argennon does not use a p2p network architecture. Instead, it uses a
+    client-server topology, based on a permission-less list of ZK-EDB servers. ZK-EDB servers are a
+    crucial part of the Argennon ecosystem, and they form the backbone of the Argennon networking layer.
+    \note{not yet written...}
+
+    \subsection{Censorship Resilient Mode}\label{subsec:cens-res-mode}
+    \note{not yet written...}
+
     \chapter{The Argennon Blockchain}\label{ch:argennon-blockchain}
-    \input{sections/blockchain}
+    \input{sections/elements}
+
+    \section{Blocks}\label{sec:blocks}
+    \input{sections/blocks}
+
+    \section{Consensus}\label{sec:consensus}
+    \input{sections/consensus}
+
+    \section{Incentive mechanism}\label{sec:incentive-mechanism}
+    \input{sections/incentive}
+
+    \section{Concurrency}\label{sec:concurrency}
+    \input{sections/concurrency}
+
+    \section{Governance}\label{sec:governance}
+    The Argennon Decentralized Autonomous Governance system (ADAGs)
+    \note{not yet written...}
 

### 2021-06-21 tex/sections/blocks.tex
Refactor project file structure
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-06-20 tex/sections/blockchain.tex
@@
 %! Author = aybehrouz
-
-\section{Applications}\label{sec:applications}
-
-An Argennon application or a smart contract is a collection of method bytecodes and heap chunks that are stored in
-the AVM storage, identified by a unique application identifier. An application identifier, \texttt{applicationID}, is
-a unique prefix code generated by the \emph{applications} prefix tree. (See Section~\ref{sec:identifiers}.)
-
-An application identifier can
-be considered as the address of an application and has the following standard symbolic representation:
-\begin{verbatim}
-<application-id> ::= "["<hex-prefix-code>"]"
-<hex-prefix-code> ::= "0x"<hex-num>"."<hex-prefix-code> | "0x"<hex-num>
-\end{verbatim}
-where \texttt{<hex-num>} is a hexadecimal number between $0$ and $255$, using lower case
-letters \texttt{[a-f]} for showing digits greater than $9$.
-
-For example \texttt{[0x24.0xff.0xda]}, \texttt{[0x0]} and \texttt{[0x3.0xa0.0x0.0x0]}, are valid application addresses.
-
-Argennon has two special smart contracts: the \emph{root} smart contract\footnote{also called the root app} and
-the \emph{ARG} smart contract\footnote{also called the Argennon smart contract or the Argennon app.}.
-The root smart contract, with \texttt{applicationID=[0x0]}, is a privileged smart contract responsible for
-method invocation and installation/uninstallation of other smart contracts.
-On the other hand, the ARG smart contract, with \texttt{applicationID=[0x1]}, controls the ARG token, the main
-currency of the Argennon blockchain, and also manages a database of public keys and handles signature verification.
-
-Both of these smart contracts are mutable smart contracts and can be updated by the Argennon governance protocol.
-
-\section{Accounts}\label{sec:accounts}
-
-Argennon accounts are entities defined inside the ARG smart contract.
-Every Argennon account is uniquely identified by a prefix code generated using \emph{accounts} prefix tree. An account
-identifier can be considered as the address of an account and has the following standard symbolic representation:
-\begin{verbatim}
-<account-id> ::= "["<decimal-prefix-code>"]"
-<decimal-prefix-code> ::= <dec-num>"."<decimal-prefix-code> | <dec-num>
-\end{verbatim}
-where \texttt{<dec-num>} is a normal decimal number between $0$ and $255$.
-
-For example \texttt{[21.255.37]}, \texttt{[0]} and \texttt{[1.0.0.0.0]}, are valid standard symbolic representations
-of account addresses.
-
-A new account can be created by invoking \texttt{createAccount} method from the Argennon smart contract. For creating
-a new account two public keys need to be provided by the caller and registered in the Argennon smart contract.
-One public key will be used for issuing digital signatures, and the other one will be used for voting. The
-provided public keys need to meet certain cryptographic requirements,\footnote{Argennon uses Prove
-Knowledge of the Secret Key (KOSK) scheme.} and can not be already registered in the system.
-
-If the owner of the new account is an application, the \texttt{applicationID} of the owner will be registered in the
-ARG smart contract and no public keys are needed. An application can own an arbitrary number of accounts.
-
-\section{Transactions}\label{sec:transactions}
-
-Every Argennon transaction is a single \texttt{invoke\_external} instruction which calls a method from the Argennon
-root smart contract. This method always transfers the proposed fee of the transaction in ARGs from a
-sender account to the fee sink accounts and then performs the requested operation. The Argennon root smart contract
-has four public methods:
-
-\begin{itemize}
-    \item \texttt{avmCall}: invokes a method from an
-    AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
-    assets, including ARGs, is done by these transactions.
-    \item \texttt{installApp}: installs an AVM smart contract and determines the update policy of the smart
-    contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
-    on.
-    \item \texttt{unInstallApp}: removes an AVM smart contract.
-    \item \texttt{updateApp}: updates an AVM smart contract (if allowed).
-\end{itemize}
-
-Argennon transactions do not have a sender. The authorization of the requested operation is always done by checking the
-digital signatures that are provided as a part of the argument list to the root smart contract method.
-
-Every Argennon transaction is required to specify a cap for all the resources it needs. This
-will include memory, network and processor related resources. The protocol defines an execution cost for
-every AVM instruction, reflecting the amount of resources its emulation needs, and every transaction is required
-to specify a maximum execution cost. If during emulation, a transaction reaches this maximum cost, it will be
-considered failed and the network can receive the proposed fee of that transaction.
-
-Also, Argennon transactions are required to specify what heap or code area addresses they will access. This will
-enable validators to parallelize transaction validation as we will see in Section~\ref{sec:concurrency}. A transaction
-that tries to access a memory location that is not in its access lists, will be rejected.
-Users could use off-chain \emph{smart contract oracles} to predict the list of memory locations their transactions need.
-
-A smart contract oracle is a full AVM emulator that keeps a full local copy of the AVM memory and can emulate AVM
-execution without accessing a ZK-EDB server. Smart contract oracles can be used for reporting useful information about
-Argennon transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
-and so on.
-
-Every Argennon transaction is required to provide the following information as an upper bound for the
-resources it needs:
-
-\begin{itemize}
-    \item Maximum execution cost
-    \item A list of heap/code-area locations for reading
-    \item A list of heap locations for writing
-    \item A list of heap chunks it will deallocate
-    \item A list of methods it will delete (if any)
-    \item Number and size of heap chunks it will allocate
-    \item Number and size of method bytecodes it will allocate (if any)
-\end{itemize}
-
-If a transaction tries to violate any of these predefined limitations, it will be considered failed, and the network
-can receive the proposed fee of that transaction.
-
-\note{An Argennon transaction always pays all of its proposed fee, no matter how much of its predefined resources
-were not used in the final emulation. This will incentivize users to exactly report their
-transactions resource usage.}
-
-\section{Blocks}\label{sec:blocks}
 
@@
 group is commutative, we have $(ab)^{-1}=a^{-1}b^{-1}$.}
-
-\section{Consensus}\label{sec:consensus}
-\input{sections/consensus}
-
-\section{Incentive mechanism}\label{sec:incentive-mechanism}
-
-\note{TODO: update this section!}
-
-\subsection{Transaction Fee}\label{subsec:transaction-fee}
-
-Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
-special method from the root smart contract. This method will transfer the proposed fee of the transaction in ARGs
-from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
-execution fees and \texttt{dbFeeSink} collects fees for ZK-EDB servers. The Protocol decides how to distribute the
-transaction fee between these two fee sink accounts.
-
-When a block is added to the blockchain, the proposer of that block will receive a share of the block fees.
-Consequently, a block proposer is always incentivized to include more transactions in his block. However, if he
-puts too many transactions in his block and the validation of the block becomes too difficult, some validators
-may not be able to validate all transactions on time. If a validator can not validate a block in the required
-time, he will consider the block invalid. So, when a proposed block contains too many transactions, the network
-may reach consensus on another block, and the proposer of that block will not receive any fees. As a result, a
-proposer is incentivized to use network transaction capacity optimally.
-
-On the other hand, we believe that the proposer does not have enough incentives for optimizing the storage size
-of the transaction set. Therefore, we require that \textbf{the size of the transaction set of every block in
-bytes be lower than a certain threshold.}
-
-Validators need to spend resources for validating transactions. When a validator starts the emulation of the AVM
-to validate a transaction, solely from the code he can't predict the time the execution will finish. This will
-give an adversary an opportunity to attack the network by broadcasting transactions that never ends. Since,
-validators can not finish the execution of these transactions, the network will not be able to charge the
-attacker any fees, and he would be able to waste validators resources for free.
-
-To mitigate this problem, we require that every transaction specify a cap for all the resources it needs. This
-will include memory, network and processor related resources. Also, the protocol defines an execution cost for
-every AVM instruction reflecting the amount of resources its emulation needs. This will define a standard way for
-measuring the execution cost of any \texttt{avmCall} transaction. Every \texttt{avmCall} transaction is required
-to specify a maximum execution cost. If during emulation it reaches this maximum cost, the transaction will be
-considered failed and the network can receive the proposed fee of that transaction.
-
-\subsection{Incentives for ZK-EDB Servers}\label{subsec:zk-edb-servers}
-
-The incentive mechanism for ZK-EDB servers should have the following properties:
-
-\begin{itemize}
-    \item It incentivizes storing all memory blocks, whether a heap page or a code area block, and not only those
-    which are used more frequently.
-    \item It incentivizes ZK-EDB servers to actively provide the required memory blocks for validators.
-    \item Making more accounts will not provide any advantages for a ZK-EDB server.
-\end{itemize}
-
-For our incentive mechanism, we require that every time a validator receives a memory block from a ZK-EDB, after
-validating the data, he give a receipt to the ZK-EDB. In this receipt the validator signs the following information:
-
-\begin{itemize}
-    \item \texttt{ownerAddr}: the ARG address of the ZK-EDB\@.
-    \item \texttt{receivedBlockID}: the ID of the received memory block.
-    \item \texttt{round}: the current round number.
-\end{itemize}
-
-\note{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}
-
-To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ARGs from
-\texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
-between all \emph{winning tickets} of the lottery.
-
-\note{One ZK-EDB server could own multiple winning tickets in a round.}
-
-To run this lottery, every round, based on the current block seed, a collection of \emph{valid} receipts will be
-selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in round $r$ if:
-
-\begin{itemize}
-    \item The signer was a validator in the round $r - 1$ and voted for the agreed-upon block.
-    \item The data block in the receipt was needed for validating the \textbf{previous} block.
-    \item The receipt round number is $r - 1$.
-    \item The signer did not sign a receipt for the same data block for two different ZK-EDBs in the previous round.
-\end{itemize}
-For selecting the winning receipts we could use a random generator:
-\begin{verbatim}
-IF random(seed|validatorPK|receivedBlockID) < winProbability THEN
-    the receipt issued by validatorPK for receivedBlockID is a winner
-\end{verbatim}
-\begin{itemize}
-    \item \texttt{random()} produces uniform random numbers between 0 and 1, using its input argument as a seed.
-    \item \texttt{validatorPK} is the public key of the signer of the receipt.
-    \item \texttt{receivedBlockID} is the ID of the memory block that the receipt was issued for.
-    \item \texttt{winProbability} is the probability of winning in every round.
-    \item \texttt{seed} is the current block seed.
-    \item \texttt{|} is a concatenation operator.
-\end{itemize}
-
-\note{The winners of the lottery were validators one round before the lottery round.}
-
-Also, based on the current block seed, a random memory block, whether a heap page or a code area block, is
-selected as the challenge of the round. A ZK-EDB that owns a winning receipt needs to broadcast a \emph{winning
-ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
-challenge. Solving a round challenge requires the content of the memory block which was selected as the round
-challenge. This will encourage ZK-EDBs to store all memory blocks.
-
-A possible choice for the challenge solution could be the cryptographic hash of the content of the challenge
-memory block combined with the ZK-EDB ARG address: \texttt{hash(challenge.content|ownerAddr)}
-
-The winning tickets of the lottery of round $r$ need to be included in the block of the round $r$,
-otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
-$r$ will be done in the round $r + 1$. This way, \textbf{the content of the challenge memory block could be
-kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.
-
-\subsection{Memory Allocation and De-allocation Fee}\label{subsec:memory-allocation-and-de-allocation}
-
-Every $k$ round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
-allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
-address of the smart contract.
-
-To determine the price of AVM memory, Every $k$ round, the protocol calculates \texttt{dbFee} and
-\texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
-\texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
-the protocol considers the total size of all the memory pages that were accessed for either reading or writing
-during a time period. These two values will be calculated for the last $k$ rounds and the price per byte of
-AVM memory will be a linear function of \texttt{dbFee/memTraffic}
-
-When a smart contract executes a heap de-allocation instruction, the protocol will refund the cost of
-de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
-calculates the refunded amount based on the average price the smart contract had paid for that allocated memory.
-This will prevent smart contracts from profit taking by trading memory with the protocol.
-
-\section{Governance}\label{sec:governance}
-
-The Argennon Decentralized Autonomous Governance system (ADAGs)
-\note{not yet written...}
-
-\section{Concurrency}\label{sec:concurrency}
-\input{sections/concurrency}

### 2021-06-21 tex/sections/elements.tex
Refactor project file structure
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-06-21 tex/sections/blockchain.tex
@@
 transactions resource usage.}
-
-\section{Blocks}\label{sec:blocks}
-
-The Argennon blockchain is a sequence of blocks. Every block represents an ordered list of transactions, intended to be
-executed by the Argennon Virtual Machine. The first block of the blockchain, the \emph{genesis} block, is a spacial
-block that fully describes the initial state of the AVM. Every block of the Argennon blockchain thus corresponds to a
-unique AVM state which can be calculated deterministically from the genesis block.
-
-A block of the Argennon blockchain contains the following information:
-
-\begin{center}
-    \begin{tabular}{||c||}
-        \hline
-        \textbf{Block} \\ [0.5ex]
-        \hline\hline
-        commitment to the staking database            \\ [1.2ex]
-        commitment to the method database             \\ [1.2ex]
-        commitment to the heap database               \\ [1.2ex]
-        commitment to the set of transactions         \\ [1.2ex]
-        last block certificate issued by         \\
-        the validators committee                      \\ [1.2ex]
-        clustering directives                         \\ [1.2ex]
-        random seed                                   \\ [1.2ex]
-        previous block hash                           \\ [1.2ex]
-        \hline
-    \end{tabular}
-\end{center}
-
-\subsection{Block Validation}\label{subsec:block-validation}
-
-Having the previous AVM state, the transaction list and the clustering directives of a block, a node can calculate
-commitments to the staking, method and heap databases of the current block by emulating the AVM execution. If the
-node can obtain the previous block information from a trusted source, it does not need to have a trusted local
-copy of the AVM state,
-and it can reliably retrieve the required storage pages from a ZK-EDB server. We call this type of block verification
-\emph{conditional} block validation. This validation is conditional because the validity of the current block is
-conditioned on the validity of the previous block.
-
-Interestingly, conditional block validation of multiple blocks can be done in parallel. If a node has enough bandwidth
-and computational resources, it can conditionally verify any number of blocks from a previously created blockchain
-simultaneously and in parallel. As we will see in Section~\ref{subsec:validators-committee}, this property plays an
-important role in the Argennon consensus protocol.
-
-To some extent, conditional validation of a single block could be parallelized as well. Many transactions
-in a block are actually independent and the order of their execution does not
-matter. These transactions can be safely validated in parallel. Section~\ref{sec:concurrency} further
-develops this concept.
-
-\subsection{Block Certificate}\label{subsec:block-certificate}
-
-An Argennon block certificate is an aggregate signature of some predefined subset of accounts. This predefined subset
-is called the certificate committee and their signature ensures that the certified block is conditionally
-valid given the validity of some previous block.
-
-Argennon uses BLS aggregate signatures to represent block certificates. To better understand block certificates and
-the Argennon consensus protocol, we need to briefly review the BLS signature scheme and its aggregation mechanism.
-
-The BLS signature scheme operates in a prime order group and supports simple threshold signature generation,
-threshold key generation, and signature aggregation. To review, the scheme uses the following ingredients:
-
-\newcommand{\G}{\mathbb{G}}
-\newcommand{\Z}{\mathbb{Z}}
-\newcommand{\adv}{{\cal A}}
-\newcommand{\bdv}{{\cal B}}
-\newcommand{\deq}{\mathrel{\mathop:}=}
-\newcommand{\SK}{\mathit{sk}}
-\newcommand{\PK}{\mathit{pk}}
-\newcommand{\C}{\mathit{cert}}
-\newcommand{\APK}{\mathit{apk}}
-\newcommand{\DPK}{\mathit{\Delta pk}}
-\newcommand{\MM}{\mathcal{M}}
-\newcommand{\xwedge}{\, \operatorname{\text{$\wedge$}}\, }
-\newcommand{\abs}[1]{\lvert #1 \rvert}
-\newcommand{\Hm}{H_0}
-\newcommand{\Hpk}{H_1}
-\newcommand{\qHpk}{Q_{\Hpk}}
-\newcommand{\qHm}{Q_{\Hm}}
-\newcommand{\qsig}{Q_{\text{sig}}}
-
-\begin{itemize}
-    \item An efficiently computable \emph{non-degenerate} pairing $e:\G_0 \times \G_1 \to \G_T$
-    in groups $\G_0$, $\G_1$ and $\G_T$ of prime order $q$. We let $g_0$ and $g_1$ be generators
-    of $\G_0$ and $\G_1$ respectively.
-    \item A hash function $H_0: \mathcal{M} \rightarrow \mathbb{G}_0$, where $\mathcal{M}$ is the message space.
-    The hash function will be treated as a random oracle.
-\end{itemize}
-
-The BLS signature scheme is defined as follows:
-
-\begin{itemize}
-    \item $\textbf{KeyGen}()$: choose a random $\alpha$ from $\Z_q$ and set $h \gets g_1^\alpha \in \G_1$.
-    output $\PK \deq (h)$ and $\SK \deq (\alpha)$.
-    \item $\textbf{Sign}(\SK, m)$: output $\sigma \gets \Hm(m)^\alpha \in \G_0$.
-    The signature $\sigma$ is a \emph{single} group element.
-    \item $\textbf{Verify}(\PK,m,\sigma)$: if $e(g_1, \sigma) = e\big(\PK,\ \Hm(m)\big)$  then output "accept",
-    otherwise output "reject".
-\end{itemize}
-
-Given triples $(\PK_i,\ m_i,\ \sigma_i)$ for $i=1,\ldots,n$,
-anyone can aggregate the signatures $\sigma_1,\ldots,\sigma_n \in \G_0$
-into a short convincing aggregate signature $\sigma$ by computing
-\begin{equation}
-    \label{eq:agg}
-    \sigma \gets \sigma_1 \cdots \sigma_n \in \G_0.
-\end{equation}
-Verifying an aggregate signature $\sigma \in \G_0$ is done by checking that
-\begin{equation}
-    \label{eq:aggdiff}
-    e(g_1, \sigma) = e\big(\PK_1,\ \Hm(m_1)\big) \cdots e\big(\PK_n,\ \Hm(m_n)\big).
-\end{equation}
-When all the messages are the same ($m = m_1 = \ldots = m_n$), the verification relation~\eqref{eq:aggdiff} reduces to
-a simpler test that requires only two pairings:
-\begin{equation}
-    \label{eq:aggsame}
-    e(g_1, \sigma) = e\Big(\PK_1 \cdots \PK_n,\ \Hm(m)\Big).
-\end{equation}
-We call $\APK=\PK_1 \cdots \PK_n$ the aggregate public key.
-
-To defend against \emph{rogue public key} attacks, Argennon uses Prove Knowledge of the Secret Key (KOSK) scheme. As we
-explained in Section~\ref{sec:accounts}, when an account is created its public keys need to be registered in
-the ARG smart contract. Therefore, the KOSK scheme can be easily implemented in Argennon.
-
-Because it is not usually possible to collect the signatures of all members of a certificate committee, an Argennon
-block certificate essentially is an Accountable-Subgroup Multi-signature (ASM). Argennon uses a simple ASM scheme
-based on BLS aggregate signatures.
-
-Argennon block certificates constitute an ordered sequence based on the order of blocks they certify. If we show
-the certificate of committee $C$ for the $i$-th block\footnote{note that the $i$-th block certificate is not
-necessarily the certificate of the $i$-th block.} with $\C_i$, and the set of signers
-with $S_i$, then the block certificate $\C_i$ can be considered as a tuple:
-\begin{equation}
-    \C_i=(\sigma_i,\ C-S_{i})\label{eq:cert},
-\end{equation}
-where $\sigma_i$ is the aggregate signature issued by $S_i$.
-
-The aggregate public key of the certificate can
-be calculated from:
-\begin{equation}
-    \APK_i=\APK_C\APK_{C-S_i}^{-1}\label{eq:aggCertPK}\ ,
-\end{equation}
-where $\APK_{A}$ shows the aggregate public key of all accounts in $A$.
-
-Alternately we can use $\APK_{i-1}$ to calculate the aggregate public key:
-\begin{equation}
-    \APK_i=\APK_{i-1}\APK_{S_i-S_{i-1}}\APK_{S_{i-1}-S_i}^{-1}\ .\label{eq:aggPK-2}
-\end{equation}
-
-When an Argennon account is created, both its $\PK$ and $\PK^{-1}$ is registered in the ARG smart contract, so the
-inverse of any aggregate public key can be easily computed.\footnote{since the group operator of a cyclic
-group is commutative, we have $(ab)^{-1}=a^{-1}b^{-1}$.}
-
-\section{Consensus}\label{sec:consensus}
-\input{sections/consensus}
-
-\section{Incentive mechanism}\label{sec:incentive-mechanism}
-
-\note{TODO: update this section!}
-
-\subsection{Transaction Fee}\label{subsec:transaction-fee}
-
-Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
-special method from the root smart contract. This method will transfer the proposed fee of the transaction in ARGs
-from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
-execution fees and \texttt{dbFeeSink} collects fees for ZK-EDB servers. The Protocol decides how to distribute the
-transaction fee between these two fee sink accounts.
-
-When a block is added to the blockchain, the proposer of that block will receive a share of the block fees.
-Consequently, a block proposer is always incentivized to include more transactions in his block. However, if he
-puts too many transactions in his block and the validation of the block becomes too difficult, some validators
-may not be able to validate all transactions on time. If a validator can not validate a block in the required
-time, he will consider the block invalid. So, when a proposed block contains too many transactions, the network
-may reach consensus on another block, and the proposer of that block will not receive any fees. As a result, a
-proposer is incentivized to use network transaction capacity optimally.
-
-On the other hand, we believe that the proposer does not have enough incentives for optimizing the storage size
-of the transaction set. Therefore, we require that \textbf{the size of the transaction set of every block in
-bytes be lower than a certain threshold.}
-
-Validators need to spend resources for validating transactions. When a validator starts the emulation of the AVM
-to validate a transaction, solely from the code he can't predict the time the execution will finish. This will
-give an adversary an opportunity to attack the network by broadcasting transactions that never ends. Since,
-validators can not finish the execution of these transactions, the network will not be able to charge the
-attacker any fees, and he would be able to waste validators resources for free.
-
-To mitigate this problem, we require that every transaction specify a cap for all the resources it needs. This
-will include memory, network and processor related resources. Also, the protocol defines an execution cost for
-every AVM instruction reflecting the amount of resources its emulation needs. This will define a standard way for
-measuring the execution cost of any \texttt{avmCall} transaction. Every \texttt{avmCall} transaction is required
-to specify a maximum execution cost. If during emulation it reaches this maximum cost, the transaction will be
-considered failed and the network can receive the proposed fee of that transaction.
-
-\subsection{Incentives for ZK-EDB Servers}\label{subsec:zk-edb-servers}
-
-The incentive mechanism for ZK-EDB servers should have the following properties:
-
-\begin{itemize}
-    \item It incentivizes storing all memory blocks, whether a heap page or a code area block, and not only those
-    which are used more frequently.
-    \item It incentivizes ZK-EDB servers to actively provide the required memory blocks for validators.
-    \item Making more accounts will not provide any advantages for a ZK-EDB server.
-\end{itemize}
-
-For our incentive mechanism, we require that every time a validator receives a memory block from a ZK-EDB, after
-validating the data, he give a receipt to the ZK-EDB. In this receipt the validator signs the following information:
-
-\begin{itemize}
-    \item \texttt{ownerAddr}: the ARG address of the ZK-EDB\@.
-    \item \texttt{receivedBlockID}: the ID of the received memory block.
-    \item \texttt{round}: the current round number.
-\end{itemize}
-
-\note{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}
-
-To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ARGs from
-\texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
-between all \emph{winning tickets} of the lottery.
-
-\note{One ZK-EDB server could own multiple winning tickets in a round.}
-
-To run this lottery, every round, based on the current block seed, a collection of \emph{valid} receipts will be
-selected randomly as the \emph{winning receipts} of the round. A receipt is \emph{valid} in round $r$ if:
-
-\begin{itemize}
-    \item The signer was a validator in the round $r - 1$ and voted for the agreed-upon block.
-    \item The data block in the receipt was needed for validating the \textbf{previous} block.
-    \item The receipt round number is $r - 1$.
-    \item The signer did not sign a receipt for the same data block for two different ZK-EDBs in the previous round.
-\end{itemize}
-For selecting the winning receipts we could use a random generator:
-\begin{verbatim}
-IF random(seed|validatorPK|receivedBlockID) < winProbability THEN
-    the receipt issued by validatorPK for receivedBlockID is a winner
-\end{verbatim}
-\begin{itemize}
-    \item \texttt{random()} produces uniform random numbers between 0 and 1, using its input argument as a seed.
-    \item \texttt{validatorPK} is the public key of the signer of the receipt.
-    \item \texttt{receivedBlockID} is the ID of the memory block that the receipt was issued for.
-    \item \texttt{winProbability} is the probability of winning in every round.
-    \item \texttt{seed} is the current block seed.
-    \item \texttt{|} is a concatenation operator.
-\end{itemize}
-
-\note{The winners of the lottery were validators one round before the lottery round.}
-
-Also, based on the current block seed, a random memory block, whether a heap page or a code area block, is
-selected as the challenge of the round. A ZK-EDB that owns a winning receipt needs to broadcast a \emph{winning
-ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
-challenge. Solving a round challenge requires the content of the memory block which was selected as the round
-challenge. This will encourage ZK-EDBs to store all memory blocks.
-
-A possible choice for the challenge solution could be the cryptographic hash of the content of the challenge
-memory block combined with the ZK-EDB ARG address: \texttt{hash(challenge.content|ownerAddr)}
-
-The winning tickets of the lottery of round $r$ need to be included in the block of the round $r$,
-otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
-$r$ will be done in the round $r + 1$. This way, \textbf{the content of the challenge memory block could be
-kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.
-
-\subsection{Memory Allocation and De-allocation Fee}\label{subsec:memory-allocation-and-de-allocation}
-
-Every $k$ round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
-allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
-address of the smart contract.
-
-To determine the price of AVM memory, Every $k$ round, the protocol calculates \texttt{dbFee} and
-\texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
-\texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
-the protocol considers the total size of all the memory pages that were accessed for either reading or writing
-during a time period. These two values will be calculated for the last $k$ rounds and the price per byte of
-AVM memory will be a linear function of \texttt{dbFee/memTraffic}
-
-When a smart contract executes a heap de-allocation instruction, the protocol will refund the cost of
-de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
-calculates the refunded amount based on the average price the smart contract had paid for that allocated memory.
-This will prevent smart contracts from profit taking by trading memory with the protocol.
-
-\section{Concurrency}\label{sec:concurrency}
-\input{sections/concurrency}
-
-\section{Governance}\label{sec:governance}
-The Argennon Decentralized Autonomous Governance system (ADAGs)
-\note{not yet written...}

### 2021-06-21 tex/sections/incentive.tex
Refactor project file structure
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-04-28 tex/sections/blockchain.tex
@@
 
-\section{Persistence}\label{sec:persistence}
-\input{sections/persistence}
-
-\section{Transactions}\label{sec:transactions}
-
-Argennon has four types of transaction:
-
-\begin{itemize}
-    \item \texttt{avmCall}: invokes a method from an
-    AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
-    assets, including ARGs, is done by these transactions.
-    \item \texttt{installApp}: installs an AVM smart contract and determines the update policy of the smart
-    contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
-    on.
-    \item \texttt{unInstallApp}: removes an AVM smart contract.
-    \item \texttt{updateApp}: updates an AVM smart contract.
-\end{itemize}
-
-All types of Argennon transactions contain a single \texttt{invoke\_external} instruction which calls a method
-from the root smart contract that transfers the proposed fee of the transaction in ARGs from a sender account to the
-fee sink accounts and then performs the requested operation.
-
-Every transaction is required to exactly specify what heap locations or code area addresses it will access. This
-enables validators to start retrieving the required storage pages from available ZK-EDB servers as soon as they
-see a transaction, and they won't need to wait for receiving the new proposed block. A transaction that tries to
-access a memory location that is not included in its access lists, will be rejected. Users could use smart contract
-oracles to predict the list of memory locations their transactions need. See Section~\ref{sec:smart-contract-oracle}
-for more details.
-
-\section{Blockchain}\label{sec:blockchain}
-
-Every block of the Argennon blockchain corresponds to a set of transactions. We store the commitment of this
-transaction set in every block, but we don't keep the set itself. To be able to detect replay attacks, we require
-every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
-and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
-must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
-a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
-if a signature of the same user with a bigger or equal nonce is already used
-(i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
-nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
-
-To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. When the
-difference between \texttt{issuance} component of this nonce and the current round becomes bigger than the
-maximum allowed lifetime of a signature, this information can be safely deleted. \textbf{As a result, we will not
-have the problem of "un-removable empty accounts" like Ethereum.}
-
-The only information that all Argennon nodes are required to store is \textbf{the most recent block} of the Argennon
-blockchain. Every block of the Argennon blockchain contains the following information:
-
-\begin{center}
-    \begin{tabular}{||c||}
-        \hline
-        \textbf{Block} \\ [0.5ex]
-        \hline\hline
-        commitment to the staking database      \\ [0.7ex]
-        commitment to the method database       \\ [0.7ex]
-        commitment to the heap database         \\ [0.7ex]
-        commitment to the set of transactions   \\ [0.7ex]
-        clustering directives                   \\ [0.7ex]
-        previous block hash                     \\ [0.7ex]
-        random seed                             \\ [0.7ex]
-        \hline
-    \end{tabular}
-\end{center}
-
-For confirming a new block, nodes that are not validators only need to verify the block certificate. For
-verifying a block certificate, a node needs to know the staking information of validators, but it doesn't need to
-emulate the AVM execution.
-
-On the other hand, nodes that are chosen to be validators, for validating a new block, need to emulate the
-execution of the Argennon Virtual Machine. To do so, first they retrieve all storage pages they
-need from available ZK-EDBs. Then, they emulate the execution of AVM instructions and validate all the
-transactions included in the new block. This will modify some pages of the AVM storage, so they update the ZK-EDB
-commitments based on the modified pages and verify the commitments included in the new block. Validators also
-calculate and verify the commitment to the new block's transaction set.
-
-\note{Validators do not need to write the modified pages back to ZK-EDB servers. ZK-EDB servers will receive the new
-block, and they will update their database by emulating the AVM execution.}
-
-\section{Consensus}\label{sec:consensus}
-\input{sections/consensus}
-
-\section{Incentive mechanism}\label{sec:incentive-mechanism}
+\note{TODO: update this section!}
 
@@
 This will prevent smart contracts from profit taking by trading memory with the protocol.
-
-\section{Concurrency}\label{sec:concurrency}
-\input{sections/concurrency}
-
-\section{Smart Contract Oracle}\label{sec:smart-contract-oracle}
-
-A smart contract oracle is a full AVM emulator that keeps a full local copy of AVM memory and can emulate AVM
-execution without accessing a ZK-EDB. Smart contract oracles can be used for reporting useful information about
-\texttt{avmCall} transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
-and so on.

### 2021-06-21 tex/sections/persistence.tex
Refactor project file structure
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The Argennon Virtual Machine has two persistent memory areas: \emph{method area}, and \emph{heap}. Method area stores
-method byte codes\footnote{also it stores constant area blocks.}, and heap stores memory chunks. Both of these
-data elements, methods and chunks, can be considered as continuous pieces of byte addressable memory. Throughout this
+bytecodes of methods\footnote{also it stores constant area blocks.}, and heap stores memory chunks. Both of these
+data elements, bytecodes and chunks, can be considered as continuous pieces of byte addressable memory. Throughout this
 chapter, we shall call these data elements \emph{objects}.
 
-In the AVM persistence layer, similar objects are clustered together and constitute a bigger data element which we call a
-\emph{page}.\footnote{we avoided calling them clusters, because usually a cluster refers to a \emph{set}. AVM object
+\subsection{Storage Pages}\label{subsec:pages}
+
+In the AVM persistence layer, similar objects are clustered together and constitute a bigger data element which we call
+a \emph{page}.\footnote{we avoided calling them clusters, because usually a cluster refers to a \emph{set}. AVM object
 clusters are not sets. They are ordered lists, like a page containing an ordered list of words or sentences.}
@@
 
-A block proposer usually obtains clustering directives from a third party source that enjoys a considerable amount of
-computational power.\footnote{we can say the AVM clustering algorithm is essentially off-chain.} This federated approach
-does not affect Argennon security, because the integrity of a database can not be altered by clustering directives.
+A block proposer could obtain clustering directives from any third party source\footnote{we can say the AVM clustering
+algorithm is essentially off-chain.}. This will not
+affect Argennon security, since the integrity of a database can not be altered by clustering directives.
 Those directives can only affect the performance of the argennon blockchain, and directives of a single block can
-not change the performance much.
+not affect the performance considerably.
+
+\subsubsection[Implementation]{Implementation Based on Vector Commitments}\label{subsubsec:impl-zk-edbs}
+
+\note{not yet written...}
+
+\subsection{Object Clustering Algorithm}\label{subsec:clustering}
+
+\note{not yet written...}

### 2021-06-22 tex/sections/consensus.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The delegates committee generates a certificate for every block of the Argennon blockchain, and the
-validators committee generates a certificate every $m$ blocks. The validators committee will
+The committee of delegates generates a certificate for every block of the Argennon blockchain, and the
+committee of validators generates a certificate every $m$ blocks. The validators' committee will
 certify a block only if it has already been certified by the committee of delegates.
@@
 In addition to primary committees, Argennon has several community driven committees. Certificates of these
-committees are not required for block validation, but they could be used by the members of the validators
-committee to better decide about the validity of a block.
+committees are not required for block finality, but they could be used by the members of the
+validators' committee to better decide about the validity of a block.
 
@@
 recovery protocol is designed to be resilient to many types of attacks in order to be able to restore the normal
-system functionality.
+functionality of the system.
 
@@
 Usually, the delegates are large organizations, and they have enough computational resources to generate blocks
-very fast. However, a block is not completely final if it does not have the certificate of the validators committee.
+very fast. However, a block is not completely final if it does not have the certificate of the validators' committee.
 A certified block by the delegates will not be accepted by the network, if the last block certified by
-the validators committee is behind more than a certain number of blocks.
+the validators is behind it more than a certain number of blocks.
 
@@
 A block certificate issued by some members of the validators committee is considered valid, if according to
-the staking database of the previous \textbf{certified block by the validators committee}, we have:
+the staking database of the previous \textbf{certified block by the validators' committee}, we have:
 \begin{itemize}
@@
 
-More precisely, to decide about signing the block certificate, a honest member of the validators
+More precisely, to decide about signing the block certificate, a honest member of the validators'
 committee iterates over every block of the validation set, and selects a block for validation with
-probability $p_u$. In this way, a member of the validators committee will select $p_{u}m$ blocks in expectation.
+probability $p_u$. In this way, a member of the validators' committee will select $p_{u}m$ blocks in expectation.
 Then, the honest member conditionally validates the selected blocks. If the member has selected more than one block,
@@
 
-One recommended formula for calculating the probability of selecting a block for validation by user $u$ is:
+One recommended formula for calculating the probability of selecting a block by user $u$ for validation is:
 \begin{equation}
@@
 
-The value of $m$ is determined by the ADAGs. but it can never be set higher than $100$. This way, it is guaranteed
+The value of $m$ is determined by the ADAGs. but it can never be higher than $100$. This way, it is guaranteed
 that on average, any block of the Argennon blockchain is validated by at least $0.005$ of the total ARG supply.
@@
 signs a block certificate, he sends his signature to \textbf{all} the severs that aggregate the signatures of his group.
-These servers aggregate the signatures they receive, and then send the aggregated signature to the delegates.
-Furthermore, the delegates aggregate these signatures to produce the final block certificate,
+These servers aggregate the signatures they receive and then send the aggregated signature to the delegates.
+Furthermore, the delegates aggregate these signatures to produce the final block certificate
 and then include it in the next block.
@@
 at time step \(t\), is estimated based on the user's trust value and his ARG balance:
-\[
-    S_{u,t} = \min (B_{u,t}, Trust_{u,k}),
-\]
+\begin{equation}
+    S_{u,t} = \min (B_{u,t}, Trust_{u,k})\ ,\label{eq:stake}
+\end{equation}
 where:
@@
 Exponential Moving Average (EMA) of his ARG balance:
-\[
-    Trust_{u,k} = L_{u,k} + M_{u,t_k}\ ,
-\]
+\begin{equation}
+    Trust_{u,k} = L_{u,k} + M_{u,t_k}\ ,\label{eq:trust}
+\end{equation}
 where
@@
 \[
-    M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})
+    M_{u,t} = (1 - \alpha) M_{u,t-1} + \alpha B_{u,t} = M_{u,t-1} + \alpha (B_{u,t} - M_{u,t-1})\ ,
 \]
@@
 \[
-    M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B
+    M_{t} = (1 - \alpha)^{t-k}M_{k} + [1 - (1 - \alpha)^{t - k}]B\ ,
 \]
@@
 \[
-    B = B_{k+1} = B_{k+2} = \dots = B_{t}
+    B = B_{k+1} = B_{k+2} = \dots = B_{t}\ .
 \]
@@
 \[
-    M_{t} = M_{k} + (t - k) \alpha (B - M_{k})
+    M_{t} = M_{k} + (t - k) \alpha (B - M_{k})\ .
 \]
@@
 \((1 + x)^n \approx e^{nx}\), so by letting \(M_{u,k} = 0\) and \(n = t - k\) we can write:
-\[
-    \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{B}\right)}{n}
-\]
+\begin{equation}
+    \alpha =- \frac{\ln\left(1 - \frac{M_{n+k}}{B}\right)}{n}\ .\label{eq:alpha}
+\end{equation}
 The value of \(\alpha\) for a desired configuration can be calculated by this equation. For instance, we could

### 2021-06-22 tex/sections/elements.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 Argennon accounts are entities defined inside the ARG smart contract.
-Every Argennon account is uniquely identified by a prefix code generated using \emph{accounts} prefix tree. An account
+Every Argennon account is uniquely identified by a prefix code generated using \emph{accounts} prefix
+tree. (See Section~\ref{sec:identifiers}) An account
 identifier can be considered as the address of an account and has the following standard symbolic representation:
@@
 
-To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. When the
-difference between \texttt{issuance} component of this nonce and the current round becomes bigger than the
-maximum allowed lifetime of a signature, this information can be safely deleted.\footnote{in some conventional
-blockchains, the nonce data can never be deleted, even if the account has zero balance and is not used anymore.}
+To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. This information
+is stored in the ARG smart contract and when the difference between \texttt{issuance} component of the nonce and
+the current round becomes bigger than the maximum allowed lifetime of a signature, it can be safely
+deleted.\footnote{in some conventional
+blockchains, the nonce data can never be deleted, even if the account has zero balance and is no longer used.}
 

### 2021-06-22 tex/sections/persistence.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 In the AVM persistence layer, similar objects are clustered together and constitute a bigger data element which we call
-a \emph{page}.\footnote{we avoided calling them clusters, because usually a cluster refers to a \emph{set}. AVM object
+a \emph{page}.\footnote{we avoid calling them clusters, because usually a cluster refers to a \emph{set}. AVM object
 clusters are not sets. They are ordered lists, like a page containing an ordered list of words or sentences.}
@@
 \[
-    P = (O_1,O_2,\dots,O_n),\quad i < j \; \Leftrightarrow \; \textrm{$O_i$ was added before $O_j$}
+    P = (O_1,O_2,\dots,O_n),\quad i < j \; \Leftrightarrow \; \textrm{$O_i$ was added before $O_j$}\ .
 \]
@@
     with \(x\), and \(\pi\) is a proof.
-    \item A user can use \(C\) and \(\pi\) to verify that \(D\) is really associated with \(x\), and \(D\) is not
+    \item A user having \(C\) and \(\pi\) can verify that \(D\) is really associated with \(x\), and \(D\) is not
     altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB\@.
@@
 
-The commitments of the AVM ZK-EDBs are affected by the way the clustering of data objects is done. So,
-the clustering algorithm has to be a part of the Argennon consensus protocol.
+The commitments of the AVM ZK-EDBs are affected by the way data objects are clustered. Therefore,
+the Argennon clustering algorithm has to be a part of the consensus protocol.
 
@@
 affect Argennon security, since the integrity of a database can not be altered by clustering directives.
-Those directives can only affect the performance of the argennon blockchain, and directives of a single block can
+Those directives can only affect the performance of the Argennon network, and directives of a single block can
 not affect the performance considerably.

### 2021-06-22 tex/sections/draft.tex
Remove outdated section
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 %! Author = aybehrouz
+
+Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
+special method from the root smart contract. This method will transfer the proposed fee of the transaction in ARGs
+from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
+execution fees and \texttt{dbFeeSink} collects fees for ZK-EDB servers. The Protocol decides how to distribute the
+transaction fee between these two fee sink accounts.
+
+When a block is added to the blockchain, the proposer of that block will receive a share of the block fees.
+Consequently, a block proposer is always incentivized to include more transactions in his block. However, if he
+puts too many transactions in his block and the validation of the block becomes too difficult, some validators
+may not be able to validate all transactions on time. If a validator can not validate a block in the required
+time, he will consider the block invalid. So, when a proposed block contains too many transactions, the network
+may reach consensus on another block, and the proposer of that block will not receive any fees. As a result, a
+proposer is incentivized to use network transaction capacity optimally.
+
+On the other hand, we believe that the proposer does not have enough incentives for optimizing the storage size
+of the transaction set. Therefore, we require that \textbf{the size of the transaction set of every block in
+bytes be lower than a certain threshold.}
+
+Validators need to spend resources for validating transactions. When a validator starts the emulation of the AVM
+to validate a transaction, solely from the code he can't predict the time the execution will finish. This will
+give an adversary an opportunity to attack the network by broadcasting transactions that never ends. Since,
+validators can not finish the execution of these transactions, the network will not be able to charge the
+attacker any fees, and he would be able to waste validators resources for free.
+
+To mitigate this problem, we require that every transaction specify a cap for all the resources it needs. This
+will include memory, network and processor related resources. Also, the protocol defines an execution cost for
+every AVM instruction reflecting the amount of resources its emulation needs. This will define a standard way for
+measuring the execution cost of any \texttt{avmCall} transaction. Every \texttt{avmCall} transaction is required
+to specify a maximum execution cost. If during emulation it reaches this maximum cost, the transaction will be
+considered failed and the network can receive the proposed fee of that transaction.
 

### 2021-06-22 tex/sections/incentive.tex
Remove outdated section
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 \subsection{Transaction Fee}\label{subsec:transaction-fee}
-
-Every transaction in the Argennon blockchain starts with an \texttt{invoke\_external} instruction which calls a
-special method from the root smart contract. This method will transfer the proposed fee of the transaction in ARGs
-from a sender account to the fee sink accounts. Argennon has two fee sink accounts: \texttt{execFeeSink} collects
-execution fees and \texttt{dbFeeSink} collects fees for ZK-EDB servers. The Protocol decides how to distribute the
-transaction fee between these two fee sink accounts.
-
-When a block is added to the blockchain, the proposer of that block will receive a share of the block fees.
-Consequently, a block proposer is always incentivized to include more transactions in his block. However, if he
-puts too many transactions in his block and the validation of the block becomes too difficult, some validators
-may not be able to validate all transactions on time. If a validator can not validate a block in the required
-time, he will consider the block invalid. So, when a proposed block contains too many transactions, the network
-may reach consensus on another block, and the proposer of that block will not receive any fees. As a result, a
-proposer is incentivized to use network transaction capacity optimally.
-
-On the other hand, we believe that the proposer does not have enough incentives for optimizing the storage size
-of the transaction set. Therefore, we require that \textbf{the size of the transaction set of every block in
-bytes be lower than a certain threshold.}
-
-Validators need to spend resources for validating transactions. When a validator starts the emulation of the AVM
-to validate a transaction, solely from the code he can't predict the time the execution will finish. This will
-give an adversary an opportunity to attack the network by broadcasting transactions that never ends. Since,
-validators can not finish the execution of these transactions, the network will not be able to charge the
-attacker any fees, and he would be able to waste validators resources for free.
-
-To mitigate this problem, we require that every transaction specify a cap for all the resources it needs. This
-will include memory, network and processor related resources. Also, the protocol defines an execution cost for
-every AVM instruction reflecting the amount of resources its emulation needs. This will define a standard way for
-measuring the execution cost of any \texttt{avmCall} transaction. Every \texttt{avmCall} transaction is required
-to specify a maximum execution cost. If during emulation it reaches this maximum cost, the transaction will be
-considered failed and the network can receive the proposed fee of that transaction.
 

### 2021-06-23 tex/A.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Change the structure of chapters
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-    \chapter{Infrastructure}\label{ch:infrastructure}
-
-    \section{Persistence Layer}\label{sec:persistence-layer}
+    \chapter{Persistence Layer}\label{ch:persistance}
     \input{sections/persistence}
 
-    \section{Networking Layer}\label{sec:networking-layer}
-    Unlike conventional blockchains, Argennon does not use a p2p network architecture. Instead, it uses a
+    \chapter{Networking Layer}\label{ch:networking}
+    Unlike conventional blockchains, Argennon does not use a P2P network architecture. Instead, it uses a
     client-server topology, based on a permission-less list of ZK-EDB servers. ZK-EDB servers are a
@@
 
-    \subsection{Censorship Resilient Mode}\label{subsec:cens-res-mode}
+    \section{Censorship Resilient Mode}\label{sec:cens-res-mode}
     \note{not yet written...}
@@
 
-    \section{Concurrency}\label{sec:concurrency}
+    \section{Concurrent Transaction Validation}\label{sec:concurrency}
     \input{sections/concurrency}
 
-    \section{Governance}\label{sec:governance}
+    \chapter{Governance}\label{ch:governance}
     The Argennon Decentralized Autonomous Governance system (ADAGs)
     \note{not yet written...}
-
 \end{document}

### 2021-06-23 tex/sections/persistence.tex
Change the structure of chapters
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 chapter, we shall call these data elements \emph{objects}.
-
-\subsection{Storage Pages}\label{subsec:pages}
 
@@
 
-\subsection{Zero-knowledge Databases}\label{subsec:zk-edb}
+\section{Zero-knowledge Databases}\label{sec:zk-edb}
 
@@
 
-\subsubsection[Implementation]{Implementation Based on Vector Commitments}\label{subsubsec:impl-zk-edbs}
+\subsection[Implementation]{Implementation Based on Vector Commitments}\label{subsec:impl-zk-edbs}
 
@@
 
-\subsection{Object Clustering Algorithm}\label{subsec:clustering}
+\section{Object Clustering Algorithm}\label{sec:clustering}
 

### 2021-06-24 tex/sections/consensus.tex
Comment out a buggy part in the consensus
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 consensus protocol.
-
-As we mentioned before, every $m$ blocks of the Argennon blockchain needs a certificate from the committee
-of validators. This means that assuming the last certified block by the validators is valid, for issuing the block
-certificate, the committee needs to conditionally validate $m$ blocks. We call these $m$ blocks the
-validation set. To reduce resource usage of validators' nodes, the protocol requires honest nodes to only
-conditionally validate a subset of the validation set. Validators that have a stake lower than
-average should only validate one block in expectation.
-
-More precisely, to decide about signing the block certificate, a honest member of the validators'
-committee iterates over every block of the validation set, and selects a block for validation with
-probability $p_u$. In this way, a member of the validators' committee will select $p_{u}m$ blocks in expectation.
-Then, the honest member conditionally validates the selected blocks. If the member has selected more than one block,
-he may validate the selected blocks in parallel.
-If all checked blocks are valid, the validator issues an "accept"
-signature, otherwise he issues a "reject" signature.
+%
+%As we mentioned before, every $m$ blocks of the Argennon blockchain needs a certificate from the committee
+%of validators. This means that assuming the last certified block by the validators is valid, for issuing the block
+%certificate, the committee needs to conditionally validate $m$ blocks. We call these $m$ blocks the
+%validation set. To reduce resource usage of validators' nodes, the protocol requires honest nodes to only
+%conditionally validate a subset of the validation set. Validators that have a stake lower than
+%average should only validate one block in expectation.
+%
+%More precisely, to decide about signing the block certificate, a honest member of the validators'
+%committee iterates over every block of the validation set, and selects a block for validation with
+%probability $p_u$. In this way, a member of the validators' committee will select $p_{u}m$ blocks in expectation.
+%Then, the honest member conditionally validates the selected blocks. If the member has selected more than one block,
+%he may validate the selected blocks in parallel.
+%If all checked blocks are valid, the validator issues an "accept"
+%signature, otherwise he issues a "reject" signature.
 

### 2021-06-24 tex/sections/elements.tex
Comment out a buggy part in the consensus
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 Every Argennon transaction is required to specify a cap for all the resources it needs. This
-will include memory, network and processor related resources. The protocol defines an execution cost for
+includes memory, network and processor related resources. The protocol defines an execution cost for
 every AVM instruction, reflecting the amount of resources its emulation needs, and every transaction is required
@@
 
+\subsection{Transaction Fee}\label{subsec:fee}
+
+Every Argennon transaction is required to pay two types of fees: execution fee and storage fee. A transaction pays
+its fees by providing digital signatures of one or more accounts authorizing the transfer of the amount of fee in
+ARG from one or more accounts to the fee sink accounts.
 \note{An Argennon transaction always pays all of its proposed fee, no matter how much of its predefined resources
-were not used in the final emulation. This will incentivize users to exactly report their
-transactions resource usage.}
+were not used in the final emulation. This will incentivize users to report the resource usage of their
+transactions more accurately.}

### 2021-06-24 tex/sections/consensus.tex
Initially fix the consensus problem
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-06-22 tex/sections/consensus.tex
@@
 
-Argennon has two primary certificate committees: the committee of \emph{delegates} and the committee
-of \emph{validators}. A block is considered final only after it gets a certificate from both of these committees.
+There is two primary types of certificate committees in Argennon: the committee of \emph{delegates} and the committee
+of \emph{validators}. Argennon has one committee of delegates and $m$ committees of validators.
 
-The committee of delegates generates a certificate for every block of the Argennon blockchain, and the
-committee of validators generates a certificate every $m$ blocks. The validators' committee will
-certify a block only if it has already been certified by the committee of delegates.
+The committee of delegates generates a certificate for every block of the Argennon blockchain, and each
+committee of validators generates a certificate every $m$ blocks. A validators' committee will
+certify a block only if it has already been certified by the committee of delegates and every committee
+generates a certificate for a different block. A block is considered
+final only after it gets a certificate from both of these committees.
 
@@
 five members, and later its size could be changed by the ADAGs in a procedure described
-in Section~\ref{sec:governance}.
+in Section~\ref{ch:governance}.
 
@@
 The Argennon protocol calculates a stake value for every account, which is an estimate of a user's stake in the
-system. This value is measured in ARGs. The committee of validators is composed of
-\textbf{all} accounts whose stake value is higher than \texttt{minValidatorsStake} threshold. This
-threshold is determined by the ADAGs, but it can never be higher
-than $100$ ARGs.
+system. This value is measured in ARGs. Every $n$ block randomly $m$ committees are selected from
+accounts whose stake value is higher than \texttt{minValidatorsStake} threshold. This
+threshold is determined by the ADAGs, but it can never be higher than $100$ ARGs.
+These committees are chosen in a
+way that the total stake of members of each committee is approximately equal and every account is a member of at
+least one committee.
 
-Every member of the validators committee has a status which can be either \texttt{online} or \texttt{offline}.
+Every member of a validators committee has a status which can be either \texttt{online} or \texttt{offline}.
 This status is stored in the ARG smart contract, and it can be changed through a method invocation
@@
 
-A block certificate issued by some members of the validators committee is considered valid, if according to
-the staking database of the previous \textbf{certified block by the validators' committee}, we have:
+A block certificate issued by some members of a validators committee is considered valid, if according to
+the staking database of the previous \textbf{block certified by the same committee}, we have:
 \begin{itemize}
-    \item The total stake of \texttt{online} accounts is higher than \texttt{minOnlineStake} percent of the
-    total stake of all accounts. This threshold can be changed by the ADAGs, but it can never be lower
+    \item The total stake of \texttt{online} members is higher than \texttt{minOnlineStake} percent of the
+    total stake of all members. This threshold can be changed by the ADAGs, but it can never be lower
     than 70 percent.
@@
     \item The sum of stake values of the certificate signers is higher than 80 percent of the total stake
-    of accounts that have \texttt{online} status.
+    of committee members that have \texttt{online} status.
 \end{itemize}
 
-If the total online stake goes lower than \texttt{minOnlineStake} threshold, nodes initiate the recovery
+If the total online stake of a committee goes lower than \texttt{minOnlineStake} threshold, nodes initiate the recovery
 consensus protocol.
 
-As we mentioned before, every $m$ blocks of the Argennon blockchain needs a certificate from the committee
-of validators. This means that assuming the last certified block by the validators is valid, for issuing the block
-certificate, the committee needs to conditionally validate $m$ blocks. We call these $m$ blocks the
-validation set. To reduce resource usage of validators' nodes, the protocol requires honest nodes to only
-conditionally validate a subset of the validation set. Validators that have a stake lower than
-average should only validate one block in expectation.
-
-More precisely, to decide about signing the block certificate, a honest member of the validators'
-committee iterates over every block of the validation set, and selects a block for validation with
-probability $p_u$. In this way, a member of the validators' committee will select $p_{u}m$ blocks in expectation.
-Then, the honest member conditionally validates the selected blocks. If the member has selected more than one block,
-he may validate the selected blocks in parallel.
-If all checked blocks are valid, the validator issues an "accept"
+As we mentioned before, every $m$ blocks of the Argennon blockchain needs a certificate from a specific committee
+of validators. To decide about signing the block certificate, an honest member of the validators'
+committee checks the conditional validity of the committee block, and if the block is valid he issues an "accept"
 signature, otherwise he issues a "reject" signature.
 
-One recommended formula for calculating the probability of selecting a block by user $u$ for validation is:
-\begin{equation}
-    p_u=\max\left(\frac{s_u}{m\mu},\ \frac{1}{m}\right),\label{eq:block-selection}
-\end{equation}
-where $\mu$ is the average stake of committee members and $s_u$ is the stake of user $u$.
+The value of $m$ is determined by the ADAGs. but it can never be higher than $25$. This way, it is guaranteed
+that on average, any block of the Argennon blockchain is validated by at least $0.02$ of the total ARG supply.
 
-The value of $m$ is determined by the ADAGs. but it can never be higher than $100$. This way, it is guaranteed
-that on average, any block of the Argennon blockchain is validated by at least $0.005$ of the total ARG supply.
-
-If the committee of validators, at any point issues a "reject" certificate for a block, the network will initiate the
+If a committee of validators, at any point issues a "reject" certificate for a block, the network will initiate the
 recovery protocol.

### 2021-06-24 tex/sections/blocks.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Change the structure of chapters
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
         commitment to the set of transactions         \\ [1.2ex]
-        last block certificate issued by         \\
-        the validators committee                      \\ [1.2ex]
+        a consecutive list of block certificates issued     \\
+        by validators' committee (if any)           \\ [1.2ex]
         clustering directives                         \\ [1.2ex]
@@
     \label{eq:agg}
-    \sigma \gets \sigma_1 \cdots \sigma_n \in \G_0.
+    \sigma \gets \sigma_1 \cdots \sigma_n \in \G_0\ .
 \end{equation}
@@
     \label{eq:aggdiff}
-    e(g_1, \sigma) = e\big(\PK_1,\ \Hm(m_1)\big) \cdots e\big(\PK_n,\ \Hm(m_n)\big).
+    e(g_1, \sigma) = e\big(\PK_1,\ \Hm(m_1)\big) \cdots e\big(\PK_n,\ \Hm(m_n)\big)\ .
 \end{equation}
@@
     \label{eq:aggsame}
-    e(g_1, \sigma) = e\Big(\PK_1 \cdots \PK_n,\ \Hm(m)\Big).
+    e(g_1, \sigma) = e\Big(\PK_1 \cdots \PK_n,\ \Hm(m)\Big)\ .
 \end{equation}
@@
 \begin{equation}
-    \C_i=(\sigma_i,\ C-S_{i})\label{eq:cert},
+    \C_i=(\sigma_i,\ C-S_{i})\label{eq:cert}\ ,
 \end{equation}

### 2021-06-24 tex/sections/consensus.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 These committees are chosen in a
-way that the total stake of members of each committee is approximately equal and every account is a member of at
-least one committee.
+way that the total stake of members of each committee is approximately equal and every account is a member of
+\textbf{at least} one committee.
 
-Every member of a validators committee has a status which can be either \texttt{online} or \texttt{offline}.
-This status is stored in the ARG smart contract, and it can be changed through a method invocation
+Every member of a validators' committee has a status which can be either \texttt{online} or \texttt{offline}.
+This status is stored in the ARG smart contract and is a part of the staking database. A validator can change
+his status through a method invocation
 from this smart contract. When an account sets its status to \texttt{offline}, it receives a small reward, and
@@
 
-\note{When an account changes its status, the change has no effect until the block containing the status
-change transaction gets certified by the committee of validators.}
+\note{When a validator changes his status, the change has no effect until the block containing the status
+change transaction gets certified by his committee.}
 
-A block certificate issued by some members of a validators committee is considered valid, if according to
-the staking database of the previous \textbf{block certified by the same committee}, we have:
+A block certificate issued by some members of a validators' committee is considered valid, if according to
+the staking database of the previous block \textbf{certified by the same committee}, we have:
 \begin{itemize}
@@
 
-As we mentioned before, every $m$ blocks of the Argennon blockchain needs a certificate from a specific committee
-of validators. To decide about signing the block certificate, an honest member of the validators'
-committee checks the conditional validity of the committee block, and if the block is valid he issues an "accept"
-signature, otherwise he issues a "reject" signature.
+As we mentioned before, every block of the Argennon blockchain needs a certificate from a specific committee of
+validators, which repeats every $m$ blocks. To decide about signing the block certificate, an honest member of a
+validators' committee checks the conditional validity of the committee block, and if the block is valid he issues
+an "accept" signature, otherwise he issues a "reject" signature.
+
+block certificates issued by validators' committees are included in the Argennon blockchain. A block may contain
+multiple certificates, provided that those certificates belong to consecutive blocks.
 
@@
 In Argennon, signature aggregation is mostly performed by ZK-EDB servers. To distribute the aggregation workload
-between different servers, The committee of validators is divided into pre-determined groups, and each ZK-EDB
-server is responsible for signature aggregation of one group. To make sure that there is enough redundancy, the number
-of groups should be less than the number of ZK-EDB servers and each group should be assigned to
+between different servers, Every committee of validators is divided into pre-determined groups, and each ZK-EDB
+server is responsible for signature aggregation of one group. To make sure that there is enough redundancy, the
+total number of groups should be less than the number of ZK-EDB servers and each group should be assigned to
 multiple ZK-EDB servers.

### 2021-06-25 tex/sections/consensus.tex
Add the reserve delegates concept
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-There is two primary types of certificate committees in Argennon: the committee of \emph{delegates} and the committee
+There are two primary types of certificate committees in Argennon: the committee of \emph{delegates} and the committee
 of \emph{validators}. Argennon has one committee of delegates and $m$ committees of validators.
@@
 committee of validators generates a certificate every $m$ blocks. A validators' committee will
-certify a block only if it has already been certified by the committee of delegates and every committee
-generates a certificate for a different block. A block is considered
-final only after it gets a certificate from both of these committees.
+certify a block only if it has already been certified by the committee of delegates. Every committee of validators has
+an index between $0$ and $m - 1$, and it issues a certificate for block number $n$, if $n\mod m$ equals
+the committee index.
+
+A block is considered final only if it has a valid certificate from both the delegates and the
+validators. A valid certificate from validators could be either a valid certificate from the committee of
+validators with the appropriate index or a certificate from all the validators.
 
@@
 five members, and later its size could be changed by the ADAGs in a procedure described
-in Section~\ref{ch:governance}.
+in Chapter~\ref{ch:governance}.
 
@@
 
+Besides the main delegates' committee, a reserve committee of delegates consists of three members is elected by users
+through the ADAGs. In case the main committee fails to generate new blocks or behaves maliciously, the task of
+block generation will be assigned to the reserve committee until a new main delegates' committee is elected
+through the ADAGs.
+
 Usually, the delegates are large organizations, and they have enough computational resources to generate blocks
-very fast. However, a block is not completely final if it does not have the certificate of the validators' committee.
+very fast. However, a block is not completely final if it does not have the certificate of the validators.
 A certified block by the delegates will not be accepted by the network, if the last block certified by
@@
 
-\subsection{The Committee of Validators}\label{subsec:validators-committee}
+\subsection{Validators}\label{subsec:validators-committee}
 
 The Argennon protocol calculates a stake value for every account, which is an estimate of a user's stake in the
-system. This value is measured in ARGs. Every $n$ block randomly $m$ committees are selected from
-accounts whose stake value is higher than \texttt{minValidatorsStake} threshold. This
+system, and is measured in ARGs. Any account whose stake value s higher than
+\texttt{minValidatorsStake} threshold is considered a validator.
+The \texttt{minValidatorsStake}
 threshold is determined by the ADAGs, but it can never be higher than $100$ ARGs.
-These committees are chosen in a
-way that the total stake of members of each committee is approximately equal and every account is a member of
-\textbf{at least} one committee.
+
+Every $n$ blocks randomly $m$ committees are selected from
+validators, in a way that the total stake of members of each committee is approximately equal and every
+account is a member of \textbf{at least} one committee.
 
@@
 his status through a method invocation
-from this smart contract. When an account sets its status to \texttt{offline}, it receives a small reward, and
+from the ARG smart contract. When an account sets its status to \texttt{offline}, it receives a small reward, and
 it can not change it back to \texttt{online} for \texttt{statusCoolDown} number of blocks.
@@
 \begin{itemize}
-    \item The total stake of \texttt{online} members is higher than \texttt{minOnlineStake} percent of the
-    total stake of all members. This threshold can be changed by the ADAGs, but it can never be lower
+    \item The total stake of \texttt{online} members of the committee is higher than \texttt{minOnlineStake} percent
+    of the total stake of all members. This threshold can be changed by the ADAGs, but it can never be lower
     than 70 percent.
@@
 
-If the total online stake of a committee goes lower than \texttt{minOnlineStake} threshold, nodes initiate the recovery
-consensus protocol.
+If the total online stake of a committee goes lower than \texttt{minOnlineStake} threshold, a partial reshuffling
+of committee members will be performed by the protocol in order to move some online members to the committee without
+enough online stake. If the reshuffling can not solve the problem, nodes will initiate the recovery consensus
+protocol.
 
 As we mentioned before, every block of the Argennon blockchain needs a certificate from a specific committee of
-validators, which repeats every $m$ blocks. To decide about signing the block certificate, an honest member of a
+validators. To decide about signing the block certificate, an honest member of a
 validators' committee checks the conditional validity of the committee block, and if the block is valid he issues
@@
 
-block certificates issued by validators' committees are included in the Argennon blockchain. A block may contain
-multiple certificates, provided that those certificates belong to consecutive blocks.
+block certificates issued by validators' committees are included in the blocks of the Argennon blockchain.
+A block can contain multiple certificates, provided that those certificates belong to consecutive blocks.
 
@@
 
-If a committee of validators, at any point issues a "reject" certificate for a block, the network will initiate the
-recovery protocol.
+If a committee of validators, at any point issues a "reject" certificate or does not issue a certificate for its
+block, after \texttt{certificateTimeOut} period of time,
+nodes will initiate the recovery protocol.
 
@@
 \note{not yet written...}
-
+%
+%Validators stop issuing certificates for normal blocks and they will only issue certificates for \emph{statuts} blocks.
+%A status block is a special block which can only contain status change transactions. Status blocks need to be
+%certified by all the validators.
+%all online validators will check the
+%validity of that block. if they do not issue a certificate for that black either
 \subsection{Analysis}\label{subsec:consensus-math}

### 2021-06-28 tex/sections/consensus.tex
Add changes needed by the recovery protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-There are two primary types of certificate committees in Argennon: the committee of \emph{delegates} and the committee
+The credibility of a block of the Argennon blockchain is determined based on the certificates it has received
+from different sets of users, knows as committees. There are two primary types of certificate committees in
+Argennon: the committee of \emph{delegates} and the committee
 of \emph{validators}. Argennon has one committee of delegates and $m$ committees of validators.
 
-The committee of delegates generates a certificate for every block of the Argennon blockchain, and each
-committee of validators generates a certificate every $m$ blocks. A validators' committee will
+The committee of delegates issues a certificate for every block of the Argennon blockchain, and each
+committee of validators issues a certificate every $m$ blocks. A validators' committee will
 certify a block only if it has already been certified by the committee of delegates. Every committee of validators has
-an index between $0$ and $m - 1$, and it issues a certificate for block number $n$, if $n\mod m$ equals
+an index between $0$ and $m - 1$, and it issues a certificate for block number $n$, if $n$ modulo $m$ equals
 the committee index.
 
-A block is considered final only if it has a valid certificate from both the delegates and the
-validators. A valid certificate from validators could be either a valid certificate from the committee of
-validators with the appropriate index or a certificate from all the validators.
+Every block of the Argennon blockchain needs a certificate from both the committee of delegates and
+the committee of validators. \textbf{A block is considered final after its next block receives both of
+its certificates.}
 
 In addition to primary committees, Argennon has several community driven committees. Certificates of these
-committees are not required for block finality, but they could be used by the members of the
+committees are not required for block finality, but they could be used by members of the
 validators' committee to better decide about the validity of a block.
@@
 
-The committee of delegates is a small committee of trusted delegates, elected by Argennon users either through the
-Argennon Decentralized Autonomous Governance system (ADAGs\footnote{pronounced \textipa{/eI-dagz/}.}) or by emergency
-voting during the recovery protocol. At the start of the Argennon mainnet, this committee will have
+The committee of delegates is a small committee of trusted delegates, elected by Argennon users through the
+Argennon Decentralized Autonomous Governance system (ADAGs\footnote{pronounced \textipa{/eI-dagz/}.}).
+At the start of the Argennon mainnet, this committee will have
 five members, and later its size could be changed by the ADAGs in a procedure described
-in Chapter~\ref{ch:governance}.
+in Section~\ref{sec:adags}.
 
@@
 Besides the main delegates' committee, a reserve committee of delegates consists of three members is elected by users
-through the ADAGs. In case the main committee fails to generate new blocks or behaves maliciously, the task of
+either through the ADAGs or by \emph{emergency agreement} during the recovery protocol. In case the main committee
+fails to generate new blocks or behaves maliciously, the task of
 block generation will be assigned to the reserve committee until a new main delegates' committee is elected
@@
 The Argennon protocol calculates a stake value for every account, which is an estimate of a user's stake in the
-system, and is measured in ARGs. Any account whose stake value s higher than
-\texttt{minValidatorsStake} threshold is considered a validator.
+system, and is measured in ARGs. Any account whose stake value is higher than
+\texttt{minValidatorsStake} threshold is considered a \emph{validator}.
 The \texttt{minValidatorsStake}
-threshold is determined by the ADAGs, but it can never be higher than $100$ ARGs.
+threshold is determined by the ADAGs, but it can never be higher than $500$ ARGs.
 
-Every $n$ blocks randomly $m$ committees are selected from
-validators, in a way that the total stake of members of each committee is approximately equal and every
+Every \texttt{committeeLifeTime} blocks, randomly $m$ committees are selected from
+validators, in a way that the total stake of members of each committee is approximately equal, and every
 account is a member of \textbf{at least} one committee.
 
-Every member of a validators' committee has a status which can be either \texttt{online} or \texttt{offline}.
+Every validator has a status which can be either \texttt{online} or \texttt{offline}.
 This status is stored in the ARG smart contract and is a part of the staking database. A validator can change
@@
 A block certificate issued by some members of a validators' committee is considered valid, if according to
-the staking database of the previous block \textbf{certified by the same committee}, we have:
+the staking database of the previous block \textbf{certified by the same committee}, we have:\footnote{If we calculate
+the stake values based on the previous block a malicious committee can affect the validators of the next block.}
 \begin{itemize}
     \item The total stake of \texttt{online} members of the committee is higher than \texttt{minOnlineStake} percent
-    of the total stake of all members. This threshold can be changed by the ADAGs, but it can never be lower
-    than 70 percent.
+    of the total stake of the committee. This threshold can be changed by the ADAGs, but it can never be lower
+    than $2/3$.
     \item All signers of the certificate have \texttt{online} status.
-    \item The sum of stake values of the certificate signers is higher than 80 percent of the total stake
-    of committee members that have \texttt{online} status.
+    \item The sum of stake values of certificate signers is higher than 80 percent of the total stake
+    of the committee members that have \texttt{online} status.
 \end{itemize}
 
-If the total online stake of a committee goes lower than \texttt{minOnlineStake} threshold, a partial reshuffling
-of committee members will be performed by the protocol in order to move some online members to the committee without
-enough online stake. If the reshuffling can not solve the problem, nodes will initiate the recovery consensus
-protocol.
+The delegates can generate blocks very fast. Therefore, the Argennon blockchain always has an
+unvalidated part which contains blocks that have a certificate from the committee of delegates but have not received
+a certificate from the validators yet.
 
-As we mentioned before, every block of the Argennon blockchain needs a certificate from a specific committee of
-validators. To decide about signing the block certificate, an honest member of a
-validators' committee checks the conditional validity of the committee block, and if the block is valid he issues
-an "accept" signature, otherwise he issues a "reject" signature.
+As we mentioned before, the block with height $n$ needs a certificate from the committee of
+validators with index $n$ modulo $m$. To decide about signing a block certificate, a validator
+checks the conditional validity\footnote{See Section~\ref{subsec:block-validation}} of the block, and
+if the block is valid he issues
+an "accept" signature. If the block is invalid, he initiates the recovery protocol. The validator will broadcast the
+certificate \textbf{only after} he sees the certificate of the validators of the previous block.
 
-block certificates issued by validators' committees are included in the blocks of the Argennon blockchain.
-A block can contain multiple certificates, provided that those certificates belong to consecutive blocks.
+Some validators may also require seeing a certificate from
+some community driven committee for the previous block, before they broadcast an "accept" certificate for a block.
+
+So in Argennon, the block validation by committees is performed in parallel, and validators
+do not wait for seeing the validators' certificate of the previous block to start transaction validation. On the
+other hand, the block certificates are published and broadcast sequentially. This ensures that an invalid
+fork made by malicious delegates will not receive any certificates from validators.
 
@@
 
-If a committee of validators, at any point issues a "reject" certificate or does not issue a certificate for its
-block, after \texttt{certificateTimeOut} period of time,
-nodes will initiate the recovery protocol.
+block certificates issued by committees of validators are included in the blocks of the Argennon blockchain.
+A block can contain multiple certificates, provided that
+those certificates belong to consecutive blocks.
 
-\subsubsection{Signature Aggregation}
+\subsection{Status Blocks}\label{subsec:status-blocks}
+
+If according to the staking database of block $n$, the total online stake of the committee with index $n$ modulo $m$ is
+lower than \texttt{minOnlineStake} threshold, the block $n + m$ can never be certified by validators.
+
+To prevent blockchain from halting in such situations, the protocol performs a predefined partial
+reshuffling of committee members.
+In this reshuffling which is based on the block random seed, some online members from other committees will
+be moved to the committee without enough online stake to make it active again.
+
+If the reshuffling can not solve the problem due to low total online stake, the protocol requires the next block
+of the blockchain to be a special \emph{status} block. A status block is a special block which can only contain
+status change transactions. The status block need to be certified by the delegates and by 2/3 of the total stake of
+the validators. The \texttt{online}/\texttt{offline} status of validators will not be considered in the validity of
+the status block certificate.\footnote{Theoretically at the status block, the total online stake of the system could
+be very low, even zero. Therefore, the status block should not be certified only by online stake.} After applying
+the transactions of the status block, the total online stake of validators
+must go higher than \texttt{minOnlineStake} threshold.
+
+\subsection{Signature Aggregation}\label{subsec:sig-agg}
 
@@
 servers. As long as there are enough honest ZK-EDB servers, the network will be able to perform signature aggregation
-even if the delegates are malicious. (See Section~\ref{subsec:recovery})
+even if the delegates are malicious.
+
+\subsection{The Recovery Protocol}\label{subsec:recovery}
+
+The recovery protocol is a resilient protocol intended for recovering the Argennon blockchain from critical situations.
+In the terminology of the CAP theorem, the recovery protocol is designed to choose consistency over availability
+and is not a protocol supposed to be executed occasionally. Ideally this protocol should never be used
+during the lifetime of the Argennon blockchain.
 
@@
 
-\subsection{The Recovery Protocol}\label{subsec:recovery}
-\note{not yet written...}
-%
-%Validators stop issuing certificates for normal blocks and they will only issue certificates for \emph{statuts} blocks.
-%A status block is a special block which can only contain status change transactions. Status blocks need to be
-%certified by all the validators.
-%all online validators will check the
-%validity of that block. if they do not issue a certificate for that black either
 \subsection{Analysis}\label{subsec:consensus-math}

### 2021-06-28 tex/sections/draft.tex
Add changes needed by the recovery protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 that having multiple voting committees will not change the security properties of the Argennon agreement protocol.
+
+If a block receives a "reject" certificate from its committee or does not receive a certificate
+after \texttt{certificateTimeOut} period of time, it will be conditionally validated by \textbf{all} validators. If
+it does not receive an "accept" certificate, the validators initiate the recovery protocol.
+
+The recovery protocol can recover the functionality
+of the Argennon blockchain as long as more than 2/3 of the total stake of the system is controlled by honest users
+and any network partition resolves after a finite amount of time. The recovery protocol uses two main emergency
+procedures to recover the : emergency forking and emergency agreement protocol.
+
+\subsubsection{Emergency Forking}
+
+The reserve committee of delegates can fork the Argennon blockchain, if it gets a valid fork request from
+the validators. A valid fork request at
+block $b$ is a signature issued by at least $0.8$ of the total \texttt{online} stake
+of the validators. The stake values are calculated based on the staking database at block $b - 1$.
+
+For forking at block $b$, the reserve committee of delegates
+makes a special fork block which only contains a valid fork request at block $b$, and its parent block is block $b$.
+The fork block needs a valid certificate from all $m$ validator committees in order to be valid. A valid fork block
+will force the fork, in the sense that any validator receiving it considers the forked chain as the new valid chain.
+
+A validator will issue
+an "accept" certificate for a fork block at position $b + 1$ if
+\begin{itemize}
+    \item the fork block is signed by the reserve committee of delegates.
+    \item the fork block contains a valid fork request at block $b$ and block $b$ is its parent.
+    \item the validator has not issued a certificate for any block $\geq b + 2$.
+    \item the validator has not seen a certificate of validators for any block $\geq b + 2$.
+\end{itemize}
+
+Emergency forking may discard at max one block with a valid validator's certificate
+from the blockchain, in the worst case scenario.
+
+\subsubsection{Emergency Agreement Protocol}
+
+Emergency agreement protocol is a protocol for deciding between a set of proposals when no committee of
+delegates is available. For initiating the protocol, a validator signs a message containing
+the hash of a block which will the stake values during voting rounds will be calculated based on its
+staking database. A validator starts the protocol if he
+receives a message containing a valid block signed by more than 2/3 of the total stake of the validators.
+if the block is too old the validator will not enter the agreement protocol.
+
+Emergency agreement is done by voting in rounds. Each round lasts for approximately $\lambda$ units of time.
+Users cast two type of votes: \emph{$i$-votes}, which are votes that are valid only in round $i$,
+and \emph{final-votes}, which are votes that are valid in any round. A user executes the
+following procedure in round $r$ of the voting phase:
+\begin{itemize}
+    \item if the user has not yet final-voted any value or if $r\mod k = 1$ he $r$-votes a desired proposal
+    \item if he sees more than 2/3 round $r$-votes for a proposal $p$ he final-votes $p$
+    \item if he sees more than 2/3 final-votes for a proposal he goes to the confirmation phase for $p$
+    \item if $clock > r \cdot \lambda$ user goes to the round $r + 1$
+\end{itemize}
+
+the confirmation phase
+\begin{itemize}
+    \item user confirm-votes $p$
+    \item if he sees more than 2/3 confirm votes for $p$ he selects $p$ and ends the agreement protocol
+    \item if $clock > k \cdot \lambda$ user goes to the round $k + 1$ of voting
+\end{itemize}
+
+\subsubsection{Initiating the Recovery Protocol}
+
+A validator will try to initiate the recovery protocol in two situations: when he does not receive any blocks for
+\texttt{blockTimeOut} amount of time, or when he sees an evidence which proves the delegates are malicious.
+
+When a validator is not receiving any blocks, first he switches his network to censorship resilient mode. If still he
+does not receive any new blocks, he will start validating all the blocks in his blockchain that does not have
+a validators' certificate. After he successfully validates the last block, if still he does not receive any blocks,
+he will sign and broadcast an \textbf{emergency fork request} message at the last block he has
+verified.
+
+If the reserve committee of delegates is already active or If a validator sees a fork request signed
+by 2/3 of the total stake of the validators, but does not receive the fork
+block after a certain amount of time he will sign and broadcast a request for
+emergency agreement on a new reserve committee. the validator puts the last valid block of his
+blockchain in the request.
+
+If a validator sees an evidence that proves the delegates are behaving maliciously, he signs and broadcasts a
+fork request at approprate block. if the reserve committee is generating blocks he signs a agreement for choosing
+a new reserve committee of delegates.
+
+A validator finds a block that conditionally is not valid. in this case validators start validating previous blocks of the
+invalid block that does not have a validators' certificate. then the validator sign a fork request at the last valid
+block before the invalid block.
+the block depends on the specific evidence
+recieving a malformed block signed by the delegates for example a block which does not contain all the required feilds
+
+two different blocks signed by the delegates at the same height (same block number)
+a block signed by the delegates which is not conditionlly valid and its previous block has a valid certificate from the validators commitee
+
+time out in receiving validators certificate
+
+time out in receiving a new block
+
+finds an invalid block in his blockchain
+
+Online stake goes too low
+
+The delegates stop generating blocks
+
+\paragraph{The delegates fork the blockchain}
+%
+%Validators stop issuing certificates for normal blocks and they will only issue certificates for \emph{statuts} blocks.
+%A status block is a special block which can only contain status change transactions. Status blocks need to be
+%certified by all the validators.
+%all online validators will check the
+%validity of that block. if they do not issue a certificate for that black either

### 2021-06-30 tex/A.tex
Add changes needed by the recovery protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Add the recovery protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
     \chapter{Networking Layer}\label{ch:networking}
+    \section{Normal Mode}\label{sec:normal-mode}
     Unlike conventional blockchains, Argennon does not use a P2P network architecture. Instead, it uses a
@@
     \chapter{Governance}\label{ch:governance}
+    \section{ADAGs}\label{sec:adags}
     The Argennon Decentralized Autonomous Governance system (ADAGs)

### 2021-06-30 tex/sections/consensus.tex
Add the recovery protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 Every block of the Argennon blockchain needs a certificate from both the committee of delegates and
-the committee of validators. \textbf{A block is considered final after its next block receives both of
-its certificates.}
+the committee of validators. \textbf{A block is considered final after it receives both of
+its certificates.} In Argennon as long as more than half of the total stake of validators is controlled by honest users,
+the probability of discarding a final block is near zero even if all the delegates are malicious.
 
@@
 
-When an anomaly is detected in the consensus mechanism, the \emph{recovery} protocol is initiated by nodes. The
+When an anomaly is detected in the consensus mechanism, the \emph{recovery} protocol is initiated by validators. The
 recovery protocol is designed to be resilient to many types of attacks in order to be able to restore the normal
@@
 Usually, the delegates are large organizations, and they have enough computational resources to generate blocks
-very fast. However, a block is not completely final if it does not have the certificate of the validators.
+very fast. However, a block is not completely final if it does not have the certificate of its validators.
 A certified block by the delegates will not be accepted by the network, if the last block certified by
@@
 
-Every \texttt{committeeLifeTime} blocks, randomly $m$ committees are selected from
+Every \texttt{committeeLifeTime} number of blocks, randomly $m$ committees are selected from
 validators, in a way that the total stake of members of each committee is approximately equal, and every
@@
 the staking database of the previous block \textbf{certified by the same committee}, we have:\footnote{If we calculate
-the stake values based on the previous block a malicious committee can affect the validators of the next block.}
+the stake values based on the previous block a malicious committee can select the validators of the next block.}
 \begin{itemize}
@@
     \item All signers of the certificate have \texttt{online} status.
-    \item The sum of stake values of certificate signers is higher than 80 percent of the total stake
+    \item The sum of stake values of certificate signers is higher than $3/4$ of the total stake
     of the committee members that have \texttt{online} status.
@@
 The delegates can generate blocks very fast. Therefore, the Argennon blockchain always has an
-unvalidated part which contains blocks that have a certificate from the committee of delegates but have not received
+unvalidated part, which contains blocks that have a certificate from the committee of delegates but have not received
 a certificate from the validators yet.
@@
 certificate \textbf{only after} he sees the certificate of the validators of the previous block.
-
 Some validators may also require seeing a certificate from
-some community driven committee for the previous block, before they broadcast an "accept" certificate for a block.
+some community driven committee. An honest validator never signs a certificate for two different blocks with the
+same height.
 
@@
 do not wait for seeing the validators' certificate of the previous block to start transaction validation. On the
-other hand, the block certificates are published and broadcast sequentially. This ensures that an invalid
+other hand, the block certificates are published and broadcast sequentially. A validator does not publish its
+certificate if the certificate of the previous block has not been published yet. This ensures that an invalid
 fork made by malicious delegates will not receive any certificates from validators.
@@
 If the reshuffling can not solve the problem due to low total online stake, the protocol requires the next block
-of the blockchain to be a special \emph{status} block. A status block is a special block which can only contain
+of the blockchain to be a special \emph{status block}. A status block is a special block which can only contain
 status change transactions. The status block need to be certified by the delegates and by 2/3 of the total stake of
-the validators. The \texttt{online}/\texttt{offline} status of validators will not be considered in the validity of
+validators. The \texttt{online}/\texttt{offline} status of validators will not be considered in the validity of
 the status block certificate.\footnote{Theoretically at the status block, the total online stake of the system could
@@
 
-The recovery protocol is a resilient protocol intended for recovering the Argennon blockchain from critical situations.
+The recovery protocol is a resilient protocol designed for recovering the Argennon blockchain from critical situations.
 In the terminology of the CAP theorem, the recovery protocol is designed to choose consistency over availability
@@
 
-\subsection{Estimating A User's Stake}\label{subsec:user's-stake}
+The recovery protocol can recover the functionality
+of the Argennon blockchain as long as more than 2/3 of the total stake of the system is controlled by honest users
+and any network partition resolves after a finite amount of time. The recovery protocol uses two main emergency
+procedures to recover the functionality of the Argennon blockchain: \emph{emergency forking} and \emph{emergency
+agreement} protocol.
+
+\subsubsection{Emergency Forking}
+
+The reserve committee of delegates is able to fork the Argennon blockchain, if it receives a valid fork request
+from the validators.
+This fork needs to be confirmed by the validators and can not discard any blocks that has been already certified
+by validators.
+A valid fork request is an unexpired request signed by more than half of the
+total \texttt{online} stake of the validators.
+
+For forking at block $b$, the reserve committee of delegates
+makes a special \emph{fork block} which only contains a valid fork request, and its parent is the block $b$.
+The height of the fork block is
+$b + 1$ so it needs a valid certificate from the committee of validator with the index $b+1$ modulo $m$.
+
+Validators always choose a valid fork block over a block of the main chain. However, as we mentioned before,
+a validator never signs a certificate for two different blocks with the same height. Consequently, if a validator
+has already signed the block $b$ of the main chain, he will not sign the fork block and vice versa.
+
+When the fork block is broadcast, it is possible that the validators of the committee with index $b+1$ modulo $m$
+get divided between the fork block and the block $b+1$ of the main chain, in a way that no block gets enough validators.
+This will cause the blockchain to halt. To prevent this, the protocol allows the reserve delegates to revoke a
+fork block. After a fork block
+is revoked, the validators who voted for it are allowed to vote for another block with the same
+height.
+
+To revoke a fork block with height $b+1$, the delegates need to seal the fork by adding a special
+\emph{seal block} after the fork block.\footnote{The parent of the seal block must be the fork block.} The seal block
+has the height $b+2$ and needs to be certified by the validators' committee with the
+index $b+2$ modulo $m$. The fork block is considered revoked only after the seal block is certified by the validators.
+
+A member of the appropriate committee will sign the certificate of a seal block only if he has not already signed
+a block after the fork block (i.e., a block with the same parent as the seal block).
+
+The seal block is not a normal block and validators who signed a certificate for a seal block are allowed
+to sign a certificate for a block with the same height if the parent of that block is not the fork
+block. In other words, the signers of the seal block are not allowed to sign a certificate for
+any blocks on the sealed chain, but are allowed to sign certificates for other chains.
+
+As long as more than half of the total online stake of every committee of validators is controlled
+by honest users, a malicious committee of delegates can not use the emergency forking procedure to discard blocks
+that have a validators' certificate.
+
+Moreover, an honest committee of delegates will always try to perform the emergency forking in such a way that
+valid blocks do not get discarded, including blocks that are not certified by the validators yet.
+
+\subsubsection{Emergency Agreement}
+
+The emergency agreement protocol is a resilient protocol for deciding between a set of proposals when no committee of
+delegates can be trusted. For initiating the protocol, a validator signs a message containing the subject
+of the agreement and a start time.
+
+A validator enters the agreement protocol if he
+receives a request that is signed by more than half of the total stake
+of the validators and its start time has not passed. The validator calculates
+the stake values based on the staking database of the last final block in his blockchain without
+considering the \texttt{online}/\texttt{offline} status of validators.
+
+The agreement protocol consists of two phases: the \emph{voting phase}, which selects a single proposal
+and the \emph{confirmation phase} which confirms the selected proposal. The voting phase is done in rounds.
+Each round lasts for approximately $\lambda$ units of time, and after $k$ rounds, the current session ends and a new
+agreement session starts. All votes and messages are tagged in a way that the messages of one session can
+not be used in another session.
+
+Users cast three type of votes: \emph{$i$-votes}, which are votes that are valid only in
+round $i$, \emph{final-votes}, which are votes that are valid in any round, and \emph{c-votes}, which are votes
+used only in the confirmation phase.
+
+A user executes the following procedure
+in round $r$ of the voting phase:
+\paragraph{Voting Phase:}
+\begin{itemize}
+    \item if the user has not yet final-voted any value, he $r$-votes a single desired proposal.
+    \item if he sees more than 2/3 $r$-votes for a proposal $p$, he final-votes $p$.
+    \item if he sees more than 2/3 final-votes for a proposal, he goes to the confirmation phase for $p$.
+    \item if $clock > r \cdot \lambda$ and $r < k$ user goes to the round $r + 1$ and if $r = k$ user starts a
+    new agreement session.
+\end{itemize}
+\paragraph{Confirmation Phase for Proposal $p$:}
+\begin{itemize}
+    \item user c-votes $p$.
+    \item if he sees more than 2/3 c-votes for $p$ he selects $p$ and ends the agreement protocol.
+    \item if $clock > k \cdot \lambda$ user starts a new agreement session.
+\end{itemize}
+
+We assume that users have clocks with the same speed, and $\lambda \gg \epsilon$, where $\epsilon$ is the maximum
+clock difference between users. We also assume that more than 2/3 of the total stake of the system is controlled
+by honest users, and network partitions are resolved after a finite amount of time. With these assumptions it can be
+shown that the emergency recovery protocol has the following important properties:
+\begin{itemize}
+    \item no two users will end the agreement protocol with two different proposals as the result of the agreement.
+    \item if honest users can agree upon some proposal value, the agreement protocol will converge to that value
+    after a finite number of sessions.
+\end{itemize}
+
+When the emergency agreement protocol is used for electing a new reserve committee to fork the blockchain, the
+confirmation phase could be skipped. In that case, the confirmation of the fork block by the committee of validators
+acts like the confirmation phase.
+
+\subsubsection{Initiating the Recovery Protocol}
+
+When a validator does not receive any blocks for \texttt{blockTimeOut} amount of time, or when he sees an
+evidence which proves the delegates are malicious, he initiates the recovery protocol.
+
+To do so, first the validator activates the censorship resilient mode of his networking module, then he checks the
+validity of the blocks in his blockchain that do not have a validators' certificate and determines the
+last valid block of his blockchain.
+
+In the next step, he will sign and broadcast an \textbf{emergency fork request} message, alongside some useful metadata
+such as the last valid block of his blockchain and the evidence of delegates' misbehaviour.\footnote{this metadata is
+not a part of the fork request.}
+
+If the reserve committee of delegates is already active, or if a validator sees a valid fork request signed
+by more than half of the total online stake of the validators, but does not receive the fork
+block after a certain amount of time, he will sign and broadcast a request for
+\textbf{emergency agreement} on a new reserve committee. The agreement on new delegates usually needs user
+interaction and is not a fully automatic process.
+
+The evidence which proves a committee of delegates is malicious is an invalid block that is signed by at least
+one delegate:
+\begin{itemize}
+    \item a block that is not conditionally valid
+    \item two different blocks at the same height
+    \item a block that has an invalid format
+\end{itemize}
+
+\subsection{Estimating Stake Values}\label{subsec:user's-stake}
 

### 2021-06-30 tex/sections/draft.tex
Add the recovery protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-06-22 tex/sections/draft.tex
@@
 that having multiple voting committees will not change the security properties of the Argennon agreement protocol.
+
+If a block receives a "reject" certificate from its committee or does not receive a certificate
+after \texttt{certificateTimeOut} period of time, it will be conditionally validated by \textbf{all} validators. If
+it does not receive an "accept" certificate, the validators initiate the recovery protocol.
+
+A valid fork block
+A validator issues
+an "accept" certificate for a fork block at position $b + 1$ (which forks the blockchain at block $b$), if:
+\begin{itemize}
+    \item the fork block is signed by the reserve committee of delegates and block $b$ is its parent.
+    \item the block $b$ has a valid certificate from validators.
+    \item the fork block contains a valid fork request according to the staking database of block $b$.
+    \item the validator has not signed a certificate for block $b + 1$.
+\end{itemize}
+
+will force the fork, in the sense that any validator after receiving it considers the forked chain as the new
+valid chain.
+Emergency forking may discard at max one block with a valid validator's certificate
+from the blockchain, in the worst case scenario.

### 2021-06-30 tex/sections/incentive.tex
Comment out a buggy part in the consensus
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Add incentives
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-\note{TODO: update this section!}
+\subsection{Certificate Rewards}\label{subsec:rewards}
 
-\subsection{Transaction Fee}\label{subsec:transaction-fee}
+When the certificate of a committee of validators is included in a block, the signers of that certificate and the
+delegates will be rewarded. Every validator will be rewarded equally. The delegates will be rewarded
+proportional to the total stake of the signers of the certificate and their reward increases if the
+certificate has more signers.
+
+Rewards will not be distributed instantly, instead they will be distributed at the end of the staking period.
+This will facilitate efficient implementations which avoid frequent updates in the AVM storage.
+Rewards of every staking period depends on the amount of fees that are collected during that period.
+
+\subsection{Stashing}\label{subsec:stashing}
+
+If an account behaves maliciously, and that behaviour could not have happened due to a mistake, by providing a proof
+in a block, the account will be disabled forever in the ARG smart contract. Disabling an account in the
+ARG smart contract will prevent that account from signing any valid signature in the future.
+
+Only behaviours will be stashed that can not happen due to a mistake or an attack. These behaviours include:
+\begin{itemize}
+    \item Signing a certificate for a block that is not conditionally valid.
+    \item Signing a certificate for two different blocks at the same height if none of them
+    is a fork block or a seal block.\footnote{signing
+    a fork block and a normal block at the same height usually is a malicious behaviour. However, it will not be
+    stashed because there are circumstances that a honest user could mistakenly do that.}
+\end{itemize}
 

### 2021-06-30 tex/sections/persistence.tex
Add the reserve delegates concept
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Add the recovery protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Add incentives
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 chapter, we shall call these data elements \emph{objects}.
+
+\section{Storage Pages}\label{sec:storage-pages}
 
@@
 
-\subsection[Implementation]{Implementation Based on Vector Commitments}\label{subsec:impl-zk-edbs}
+\subsection{Vector Commitments}\label{subsec:impl-zk-edbs}
+
+Informally, vector commitments allow committing to an ordered sequence of $q$ values
+(i.e.\ a vector), rather than to single messages. This is done in a way such that it is later possible
+to open the commitment with respect to specific positions (e.g., to prove that $m_i$ is the $i$-th committed
+message). More precisely, vector commitments are required to satisfy what is called position binding.
+Position binding states that an adversary should not be able to open a commitment to two different
+values at the same position. While this property, by itself, would be trivial to realize using standard
+commitment schemes, what makes vector commitments interesting is that they are concise, i.e.,
+the size of the commitment string as well as the size of each opening \textbf{is independent of the
+vector length}.
 

### 2021-07-01 tex/sections/incentive.tex
Fix a typo
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-\subsection{Stashing}\label{subsec:stashing}
+\subsection{Penalties}\label{subsec:penalties}
 
@@
 
-Only behaviours will be stashed that can not happen due to a mistake or an attack. These behaviours include:
+Only behaviours will be punished that can not happen due to a mistake or an attack. These behaviours include:
 \begin{itemize}
@@
     a fork block and a normal block at the same height usually is a malicious behaviour. However, it will not be
-    stashed because there are circumstances that a honest user could mistakenly do that.}
+    penalized because there are circumstances that a honest user could mistakenly do that.}
 \end{itemize}

### 2021-07-01 tex/sections/incentive.tex
Update names in Section ZK-EDB incentives
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
     \item Signing a certificate for two different blocks at the same height if none of them
-    is a fork block or a seal block.\footnote{signing
+    is a fork block or a seal block.\footnote{Signing
     a fork block and a normal block at the same height usually is a malicious behaviour. However, it will not be
@@
 \begin{itemize}
-    \item It incentivizes storing all memory blocks, whether a heap page or a code area block, and not only those
-    which are used more frequently.
-    \item It incentivizes ZK-EDB servers to actively provide the required memory blocks for validators.
-    \item Making more accounts will not provide any advantages for a ZK-EDB server.
+    \item It incentivizes storing all storage pages and not only those pages that are used more frequently.
+    \item It incentivizes ZK-EDB servers to actively provide the required storage pages for validators.
+    \item Making more accounts will not provide any advantage for a ZK-EDB server.
 \end{itemize}
 
-For our incentive mechanism, we require that every time a validator receives a memory block from a ZK-EDB, after
-validating the data, he give a receipt to the ZK-EDB. In this receipt the validator signs the following information:
+For our incentive mechanism, we require that every time a validator receives a storage page from a ZK-EDB, after
+validating the data, he give a receipt to the ZK-EDB server. In this receipt the validator signs the
+following information:
 
 \begin{itemize}
-    \item \texttt{ownerAddr}: the ARG address of the ZK-EDB\@.
-    \item \texttt{receivedBlockID}: the ID of the received memory block.
-    \item \texttt{round}: the current round number.
+    \item \texttt{ownerAddr}: the account address of the ZK-EDB server.
+    \item \texttt{receivedPageID}: the ID of the received page.
+    \item \texttt{round}: the current block number.
 \end{itemize}
 
-\note{In a round, an honest validator never gives a receipt for an identical memory block to two different ZK-EDBs.}
+\note{In a round, an honest validator never gives a receipt for an identical page to two different ZK-EDB servers.}
 
-To incentivize ZK-EDB servers, a lottery will be held every round and a predefined amount of ARGs from
+To incentivize ZK-EDB servers, a lottery will be held every round,\footnote{A round is the time interval between
+two consecutive blocks.} and a predefined amount of ARGs from
 \texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
@@
 \begin{itemize}
-    \item The signer was a validator in the round $r - 1$ and voted for the agreed-upon block.
-    \item The data block in the receipt was needed for validating the \textbf{previous} block.
+    \item The signer was a member of the validators' committee of the block $r - 1$ and signed the block certificate.
+    \item The page in the receipt was needed for validating the \textbf{previous} block.
     \item The receipt round number is $r - 1$.
-    \item The signer did not sign a receipt for the same data block for two different ZK-EDBs in the previous round.
+    \item The signer did not sign a receipt for the same storage page for two different ZK-EDB servers in
+    the previous round.
 \end{itemize}
@@
 \begin{verbatim}
-IF random(seed|validatorPK|receivedBlockID) < winProbability THEN
-    the receipt issued by validatorPK for receivedBlockID is a winner
+IF random(seed|validatorPK|receivedPageID) < winProbability THEN
+    the receipt issued by validatorPK for receivedPageID is a winner
 \end{verbatim}
@@
     \item \texttt{validatorPK} is the public key of the signer of the receipt.
-    \item \texttt{receivedBlockID} is the ID of the memory block that the receipt was issued for.
+    \item \texttt{receivedPageID} is the ID of the storage page that the receipt was issued for.
     \item \texttt{winProbability} is the probability of winning in every round.
@@
 
-\note{The winners of the lottery were validators one round before the lottery round.}
-
-Also, based on the current block seed, a random memory block, whether a heap page or a code area block, is
-selected as the challenge of the round. A ZK-EDB that owns a winning receipt needs to broadcast a \emph{winning
+Also, based on the current block seed, a random storage page is
+selected as the challenge of the round. A ZK-EDB server that owns a winning receipt needs to broadcast a \emph{winning
 ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
-challenge. Solving a round challenge requires the content of the memory block which was selected as the round
-challenge. This will encourage ZK-EDBs to store all memory blocks.
+challenge. Solving a round challenge requires the content of the storage page which was selected as the round
+challenge. This will encourage ZK-EDB servers to store all storage pages.
 
 A possible choice for the challenge solution could be the cryptographic hash of the content of the challenge
-memory block combined with the ZK-EDB ARG address: \texttt{hash(challenge.content|ownerAddr)}
+page combined with the server account address:
+
+\texttt{hash(challenge.content|ownerAddr)}
 
@@
 otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
-$r$ will be done in the round $r + 1$. This way, \textbf{the content of the challenge memory block could be
+$r$ will be done in the round $r + 1$. This way, \textbf{the content of the challenge page could be
 kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.

### 2021-07-01 tex/sections/draft.tex
Remove outdated section
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 from the blockchain, in the worst case scenario.
+
+\subsection{Memory Allocation and De-allocation Fee}\label{subsec:memory-allocation-and-de-allocation}
+
+Every $k$ round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
+allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
+address of the smart contract.
+
+To determine the price of AVM memory, Every $k$ round, the protocol calculates \texttt{dbFee} and
+\texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
+\texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
+the protocol considers the total size of all the memory pages that were accessed for either reading or writing
+during a time period. These two values will be calculated for the last $k$ rounds and the price per byte of
+AVM memory will be a linear function of \texttt{dbFee/memTraffic}
+
+When a smart contract executes a heap de-allocation instruction, the protocol will refund the cost of
+de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
+calculates the refunded amount based on the average price the smart contract had paid for that allocated memory.
+This will prevent smart contracts from profit taking by trading memory with the protocol.

### 2021-07-01 tex/sections/incentive.tex
Remove outdated section
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.
-
-\subsection{Memory Allocation and De-allocation Fee}\label{subsec:memory-allocation-and-de-allocation}
-
-Every $k$ round the protocol chooses a price per byte for AVM memory. When a smart contract executes a heap
-allocation instruction, the protocol will automatically deduce the cost of the allocated memory from the ARG
-address of the smart contract.
-
-To determine the price of AVM memory, Every $k$ round, the protocol calculates \texttt{dbFee} and
-\texttt{memTraffic} values. \texttt{dbFee} is the aggregate amount of collected database fees, and
-\texttt{memTraffic} is the total memory traffic of the system. For calculating the memory traffic of the system
-the protocol considers the total size of all the memory pages that were accessed for either reading or writing
-during a time period. These two values will be calculated for the last $k$ rounds and the price per byte of
-AVM memory will be a linear function of \texttt{dbFee/memTraffic}
-
-When a smart contract executes a heap de-allocation instruction, the protocol will refund the cost of
-de-allocated memory to the smart contract. Here, the current price of AVM memory does not matter and the protocol
-calculates the refunded amount based on the average price the smart contract had paid for that allocated memory.
-This will prevent smart contracts from profit taking by trading memory with the protocol.

### 2021-07-03 tex/sections/blocks.tex
Fix some readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 commitments to the staking, method and heap databases of the current block by emulating the AVM execution. If the
-node can obtain the previous block information from a trusted source, it does not need to have a trusted local
+node can obtain the previous block commitments from a trusted source, it does not need to have a trusted local
 copy of the AVM state,
 and it can reliably retrieve the required storage pages from a ZK-EDB server. We call this type of block verification
-\emph{conditional} block validation. This validation is conditional because the validity of the current block is
+\emph{conditional} block validation, since the validity of the current block is
 conditioned on the validity of the previous block.
@@
 Argennon block certificates constitute an ordered sequence based on the order of blocks they certify. If we show
-the certificate of committee $C$ for the $i$-th block\footnote{note that the $i$-th block certificate is not
-necessarily the certificate of the $i$-th block.} with $\C_i$, and the set of signers
+the $i$-th certificate\footnote{note that the $i$-th certificate is not
+necessarily the certificate of the $i$-th block.} of committee $C$ with $\C_i$, and the set of signers
 with $S_i$, then the block certificate $\C_i$ can be considered as a tuple:
@@
 
-Alternately we can use $\APK_{i-1}$ to calculate the aggregate public key:
+Alternately, we can use $\APK_{i-1}$ to calculate the aggregate public key:
 \begin{equation}

### 2021-07-03 tex/sections/consensus.tex
Add incentives
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Fix some readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The credibility of a block of the Argennon blockchain is determined based on the certificates it has received
-from different sets of users, knows as committees. There are two primary types of certificate committees in
+The credibility of a block of the Argennon blockchain is determined by the certificates it has received
+from different sets of users, knows as committees. There are two primary type of certificate committees in
 Argennon: the committee of \emph{delegates} and the committee
-of \emph{validators}. Argennon has one committee of delegates and $m$ committees of validators.
+of \emph{validators}. Argennon has \emph{one} committee of delegates and $m$ committees of validators.
 
@@
 Every \texttt{committeeLifeTime} number of blocks, randomly $m$ committees are selected from
-validators, in a way that the total stake of members of each committee is approximately equal, and every
+validators, in a way that the total stakes of committees are almost equal, and every
 account is a member of \textbf{at least} one committee.
@@
 \begin{itemize}
-    \item The total stake of \texttt{online} members of the committee is higher than \texttt{minOnlineStake} percent
+    \item The total stake of \texttt{online} members of the committee is higher than \texttt{minOnlineStake} fraction
     of the total stake of the committee. This threshold can be changed by the ADAGs, but it can never be lower
@@
     \item All signers of the certificate have \texttt{online} status.
-    \item The sum of stake values of certificate signers is higher than $3/4$ of the total stake
+    \item The sum of stake values of the certificate signers is higher than $3/4$ of the total stake
     of the committee members that have \texttt{online} status.
@@
 The delegates can generate blocks very fast. Therefore, the Argennon blockchain always has an
-unvalidated part, which contains blocks that have a certificate from the committee of delegates but have not received
-a certificate from the validators yet.
+unvalidated part, which contains blocks that have a certificate from the committee of delegates but have not
+yet received a certificate from the validators.
 
@@
 So in Argennon, the block validation by committees is performed in parallel, and validators
-do not wait for seeing the validators' certificate of the previous block to start transaction validation. On the
+do not wait for seeing the certificate of the previous block validators to start transaction validation. On the
 other hand, the block certificates are published and broadcast sequentially. A validator does not publish its
@@
 the status block certificate.\footnote{Theoretically at the status block, the total online stake of the system could
-be very low, even zero. Therefore, the status block should not be certified only by online stake.} After applying
-the transactions of the status block, the total online stake of validators
+be very low. Therefore, the status block should not be certified only by online stake.} After applying
+the transactions of the status block, the total online stake of all committees of validators
 must go higher than \texttt{minOnlineStake} threshold.
@@
 The recovery protocol is a resilient protocol designed for recovering the Argennon blockchain from critical situations.
-In the terminology of the CAP theorem, the recovery protocol is designed to choose consistency over availability
+In the terminology of the CAP theorem, the recovery protocol is designed to choose consistency over availability,
 and is not a protocol supposed to be executed occasionally. Ideally this protocol should never be used
@@
 from the validators.
-This fork needs to be confirmed by the validators and can not discard any blocks that has been already certified
+This fork needs to be confirmed by validators and can not discard any blocks that has been already certified
 by validators.
@@
 by honest users, a malicious committee of delegates can not use the emergency forking procedure to discard blocks
-that have a validators' certificate.
+that have a certificate from validators.
 
@@
 and the \emph{confirmation phase} which confirms the selected proposal. The voting phase is done in rounds.
-Each round lasts for approximately $\lambda$ units of time, and after $k$ rounds, the current session ends and a new
-agreement session starts. All votes and messages are tagged in a way that the messages of one session can
+Each round lasts for approximately $\lambda$ units of time, and after $k$ rounds, the current agreement session ends and a
+new session starts. All votes and messages are tagged in a way that the messages of one session can
 not be used in another session.
@@
 When the emergency agreement protocol is used for electing a new reserve committee to fork the blockchain, the
-confirmation phase could be skipped. In that case, the confirmation of the fork block by the committee of validators
-acts like the confirmation phase.
+confirmation phase could be skipped. In that case, the confirmation of the fork block by the appropriate committee
+of validators acts like the confirmation phase.
 
@@
 To do so, first the validator activates the censorship resilient mode of his networking module, then he checks the
-validity of the blocks in his blockchain that do not have a validators' certificate and determines the
+validity of the blocks that do not have a validators' certificate and determines the
 last valid block of his blockchain.
@@
 of stake the user has in the system. Conventionally in these systems, a user's stake is considered to be equal with the
-amount of native system tokens, the user has "staked" in the system. A user stakes his tokens by locking them in
-his account or a staking account for some period of time, and during this time he will not be able to transfer
+amount of native system tokens, he has "staked" in the system. A user stakes his tokens by locking them in
+his account or a separate staking account for some period of time. During this time, he will not be able to transfer
 his tokens.
@@
 This means that for a real world currency this type of staking mechanisms will result in putting the
-fate of the system in the hands of the owners of a small fraction of the total supply of a currency.
+fate of the system in the hands of the owners of a small fraction of the total supply.
 
@@
 Every \texttt{stakingDuration} blocks, which is called a \emph{staking period}, Argennon calculates
-a \emph{trust value} for each user. The user's stake
+a \emph{trust value} for each user.
+
+The user's stake
 at time step \(t\), is estimated based on the user's trust value and his ARG balance:

### 2021-07-03 tex/sections/elements.tex
Add incentives
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Fix some readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-\note{Explicit key registration enables Argennon to decouple cryptography from blockchain design. In this way,
-    if the used cryptographic algorithms turn out to be insecure for some reason, for example because
-    of the introduction of quantum computers, the cryptographic algorithms could be easily upgraded.}
+\note{Explicit key registration enables Argennon to decouple cryptography from the blockchain design. In this way,
+    if the cryptographic algorithms used become insecure for some reason, for example because
+    of the introduction of quantum computers, they could be easily upgraded.}
 
@@
     \item \texttt{avmCall}: invokes a method from an
-    AVM smart contract. Users interact with AVM smart contracts using these transactions. Transferring all
-    assets, including ARGs, is done by these transactions.
+    AVM smart contract. Users interact with AVM smart contracts using this method of the root smart contract.
+    Transferring all assets, including ARG, is done by these transactions.
     \item \texttt{installApp}: installs an AVM smart contract and determines the update policy of the smart
@@
 enable validators to parallelize transaction validation as we will see in Section~\ref{sec:concurrency}. A transaction
-that tries to access a memory location that is not in its access lists, will be rejected.
+that tries to access a memory location that is not in its access list, will be rejected.
 Users could use off-chain \emph{smart contract oracles} to predict the list of memory locations their transactions need.
 
-A smart contract oracle is a full AVM emulator that keeps a full local copy of the AVM memory and can emulate AVM
+A smart contract oracle is a full AVM emulator that keeps a full local copy of the AVM storage and can emulate AVM
 execution without accessing a ZK-EDB server. Smart contract oracles can be used for reporting useful information about
@@
 Every Argennon transaction is required to pay two types of fees: execution fee and storage fee. A transaction pays
-its fees by providing digital signatures of one or more accounts authorizing the transfer of the amount of fee in
+its fees by providing digital signatures of one or more accounts, authorizing the transfer of the amount of fee in
 ARG from one or more accounts to the fee sink accounts.
+
 \note{An Argennon transaction always pays all of its proposed fee, no matter how much of its predefined resources

### 2021-07-04 tex/sections/consensus.tex
Fix an issue in the emergency forking protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 Every block of the Argennon blockchain needs a certificate from both the committee of delegates and
-the committee of validators. \textbf{A block is considered final after it receives both of
-its certificates.} In Argennon as long as more than half of the total stake of validators is controlled by honest users,
+the committee of validators. A block is considered final after its \textbf{next} block receives \textbf{both} of
+its certificates. In Argennon as long as more than half of the total stake of validators is controlled by honest users,
 the probability of discarding a final block is near zero even if all the delegates are malicious.
@@
 
-Validators always choose a valid fork block over a block of the main chain. However, as we mentioned before,
+For signing a fork block, a validator ensures that the block is signed by the reserve committee and contains
+a valid fork request. The parent of the fork block does not necessarily need a validators' certificate. If the
+parent does not have a certificate, the validator checks the certificate of the block before the parent and
+the conditional validity of the parent instead. This enables the reserve committee
+to recover the liveliness of
+the blockchain in a situation where a malicious committee has generated multiple blocks at the same height.
+
+A validator always choose a valid fork block over a block of the main chain. However, as we mentioned before,
 a validator never signs a certificate for two different blocks with the same height. Consequently, if a validator
@@
 
-A member of the appropriate committee will sign the certificate of a seal block only if he has not already signed
-a block after the fork block (i.e., a block with the same parent as the seal block).
-
 The seal block is not a normal block and validators who signed a certificate for a seal block are allowed
-to sign a certificate for a block with the same height if the parent of that block is not the fork
-block. In other words, the signers of the seal block are not allowed to sign a certificate for
-any blocks on the sealed chain, but are allowed to sign certificates for other chains.
+to sign a certificate for a block with the same height and vice versa. However, it should be noted that
+generating a block with the same parent as the seal block is considered a malicious behaviour of the reserve
+committee and validators will not sign such a block.
 
@@
     \item a block that is not conditionally valid
-    \item two different blocks at the same height
+    \item two different blocks with the same parent
     \item a block that has an invalid format

### 2021-07-05 tex/sections/avm.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Update `README.md`
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-    \texttt{methodID = (applicationID|<local prefix code>)}
+    \texttt{methodID = (applicationID|<local-prefix-code>)}
     \item \texttt{chunkID} is a composite prefix code built by concatenating an \texttt{applicationID} to
@@
 
-    \texttt{chunkID = (applicationID|accountID|<local prefix code>)}
+    \texttt{chunkID = (applicationID|accountID|<local-prefix-code>)}
 \end{itemize}
@@
 prefix tree as a sequence of fractional numbers\footnote{It's possible to have \(a_i=0\). For
-instance, \(A^{(4)}=(0.2000)_{10}\) is a valid prefix tree.} in base \(\beta\):
+exmaple \(A^{(4)}=(0.2000)_{10}\) is correct.} in base \(\beta\):
 \[
-    (A^{(1)},A^{(2)},A^{(3)},\dots)
+    (A^{(1)},A^{(2)},A^{(3)},\dots)\ ,
 \]
-Where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)}<A^{(i+1)}\). A typical choice for \(\beta\)
+where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)}<A^{(i+1)}\). A typical choice for \(\beta\)
 could be \(2^8\).
@@
 protocol. This means that bugs or security vulnerabilities in the AVM standard library could be quickly patched and
-smart contracts can use the AVM standard library safely. Many important and useful functionalities,
+smart contracts could benefit from bugfixes and improvements of the AVM standard library even if they are
+non-updatable. Many important and useful functionalities,
 such as fungible and non-fungible assets, access control mechanisms,

### 2021-07-11 tex/sections/avm.tex
Make the AVM RESTful
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 abstract computing machine for executing Argennon's smart contracts. The
-Argennon Virtual Machine knows nothing of the Argennon blockchain, only of Argennon's identifier tries and
-the concept of execution sessions.
+Argennon Virtual Machine knows nothing of the Argennon blockchain, transactions or any concept of gas usage, only of
+Argennon's identifier tries and the concept of execution sessions.
 
-Execution sessions are separate sessions of executing smart contact's code by the Argennon Virtual Machine.
-These sessions are usually correspondent to the concept of a transaction in a blockchain. (See
-Section~\ref{sec:execution-sessions})
+Execution sessions are separate sessions of executing smart contract's code by the Argennon Virtual Machine.
+Every execution session starts with a single method invocation, but does not necessarily end when
+that method invocation completes. During an execution session additional method invocations could be scheduled,
+which will be executed after the initial method completion.
+
+An execution session may complete normally or abruptly. When an execution session completes abruptly, it will
+not have any effects on the AVM state.
 
@@
 Every AVM execution session has a queue of call stacks. A call stack contains all the information that is needed for
-restoring the state, and continuing the execution after the method invocation completes.
+restoring the AVM state, and continuing the execution after the method invocation completes.
 This information is represented by a \texttt{CallInfo} struct:
@@
 CallInfo {
-    applicationID,
     methodID,
     pc,
-    canModify,
     localFrame,
     operandStack
-}
+}.
 \end{verbatim}
-The \texttt{applicationID} field is the unique identifier that the AVM assigns to every smart contract,
-\texttt{methodID} is the unique identifier of a method's bytecode and \texttt{canModify} is a flag
-indicating whether a method is allowed to execute state modifying instructions or not.
+
+The \texttt{methodID} field is the unique identifier of a method's bytecode that
+is a composite prefix code built by concatenating an \texttt{applicationID} to
+a local identifier.
 
@@
 \texttt{CallInfo} struct is pushed onto the call stack. Hence, the top of the call stack always contains
-the \texttt{CallInfo} struct of the current method. When a method invocation completes its call
+the \texttt{CallInfo} struct of the current method. When a method invocation completes, its call
 information is popped from the call stack.
@@
 An AVM execution session will continue as long as there is a non-empty call stack in the call stack queue. When the
-execution of the current call stack finishes the execution of the next call stack in the call stack queue starts.
-It's possible that an AVM method invocation creates new call stacks. These newly created call stacks will be a part
+execution of the current call stack finishes, the execution of the next call stack in the call stack queue starts.
+It is possible that an AVM method invocation creates new call stacks. These newly created call stacks will be a part
 of the current execution session and are added at the end of the call stack queue. A newly created call stack always
@@
 
-Instructions that modify the code area can only be run in privileged mode which means they can only be executed
+Every smart contract is required to have a special \texttt{dispatcher} method. The \texttt{dispatcher} method is the
+only method of a smart contract that can be called by other smart contracts and has a
+predefined \texttt{methodID}: \texttt{(ApplicationID|0)}.
+
+\note{The Argennon Virtual Machine does not have internal or external methods. All methods are considered internal
+except the \texttt{dispatcher} method, which is the only external method of any smart contract.
+This design enables the AVM to execute smart contracts with \emph{RESTful} APIs more efficiently.}
+
+Instructions that modify the code area can only be run in privileged mode, which means they can only be executed
 by the \emph{root} smart contract. As a result, the access of smart contracts to the code area is essentially
@@
 Every smart contract has a single constant area which is stored in the AVM code area as a special type of
-non-executable method. The constant area of a smart contract contains several kinds of constants, ranging from user
-defined constants to method address tables. A method address table stores the list of methods of a smart
-contract and their access type. The access type of method can be either \texttt{external} or \texttt{internal}.
-Only external methods can be invoked by \texttt{invoke\_external} instruction.
+non-executable method. The constant area of a smart contract contains user
+defined constants.
 
@@
 those chunks that was created by itself. In other words, every memory chunk in the AVM heap has an owner. Only
-the owner can modify a chunk while anyone can read that chunk. In addition, for modifying a chunk the \texttt{canModify}
-flag of the current method's \texttt{callInfo} struct must be true. The owner of a chunk can be easily determined
+the owner can modify a chunk while anyone can read that chunk. The owner of a chunk can be easily determined
 by the \texttt{applicationID} part of the chunk identifier.
@@
 immutable. That means if a smart contract does not implement a getter mechanism for some parts of its internal
-data, this functionality can never be added later. Although the internal data is publicly available, there will
+data, this functionality can never be added later, and despite the internal data is publicly available, there will
 be no way for other smart contracts to use this data on-chain. This design eliminates the need for implementing
@@
     \item \texttt{invoke\_internal}: invokes a method without changing the context of method execution. In other
-    words, \texttt{applicationID} and \texttt{canModify} fields of the invoked method in the
-    call stack will be the same as the invoker.
+    words, the invoked method can only modify the state of the current smart contract.
     A smart contract can invoke any method by \texttt{invoke\_internal} instruction, even if that method is an
@@
     this instruction. This instruction facilitates code reuse and the usage of libraries.
-    \item \texttt{invoke\_external}: invokes a method and changes the context of method execution to another
-    smart contract. Only external methods of another smart contract can be called by this instruction. For changing
-    the context of method execution this instruction sets the \texttt{applicationID} field of the invoked method
-    to the \texttt{applicationID} of the called smart contract. Then it searches the call stack and if
-    the \texttt{applicationID} of the called smart contract is found in the current call stack it sets
-    \texttt{canModify} to \texttt{false}, otherwise \texttt{canModify} is set to \texttt{true}.
-    As a result, if the called smart contract is already called, the invoked method will not be
-    able to modify the heap area.
+    If the invoked method completes abruptly, that will cause the invoker method to complete abruptly too.
+    \item \texttt{invoke\_dispatcher}: invokes the \texttt{dispatcher} method of a smart contract and changes the
+    context of method execution to that
+    smart contract. If the invoked method completes abruptly, that will cause the invoker method to complete
+    abruptly too.
+    \item \texttt{i\_invoke\_dispatcher, i\_invoke\_internal}: independently invokes a method. These instructions
+    are similar to their \texttt{invoke} counterparts, with the difference that when
+    the invoked method completes abruptly, the invoker method can continue execution. When a method invoked by
+    an \texttt{i\_invoke} instruction completes abruptly, all changes made
+    by the invoked method to the AVM state will be reverted and the instruction will act like
+    a jump to the specified error handling code.
     \item \texttt{invoke\_native}: invokes a method that is not hosted by the Argennon Virtual Machine. By this
-    instruction, high performance native methods of the hosting machine could become available to AVM smart
-    contracts. This instruction will not modify the AVM state.
-    \item \texttt{spawn}: spawns a new method invocation. Spawning a method invocation is the invocation of a method
-    with changing the context of method execution and creating a new call stack. Only external methods can
-    be spawned. The created call stack is added at the end of the call stack queue and
-    the \texttt{canModify} field of the pushed \texttt{CallInfo} struct will always
-    be set to \texttt{true}. This instruction does not modify the \texttt{pc} register and the execution of the
-    caller will continue after this instruction. As a result, \texttt{spawn} can not return
-    any value to the caller and the return value of the called method (if any) will be added
-    to the transaction output. If the invoked method completes abruptly, it will cause all the call stack queue to
-    complete abruptly.
+    instruction, high performance native methods of the hosting machine, such as GPU accelerated operations,
+    could become available to AVM smart contracts. This instruction will not modify the AVM state.
+    \item \texttt{invoke\_later}: invokes the \texttt{dispatcher} method of another smart contract by creating a new
+    call stack. The created call stack is added at the end of the call stack queue, and
+    the context of method execution will change after the execution of that call stack starts. This instruction does
+    not modify the \texttt{pc} register and the execution of the
+    invoker will continue after this instruction. As a result, \texttt{invoke\_later} can not return
+    any values to the caller and the return value of the called method (if any) will be discarded. If the invoked
+    method completes abruptly, it will cause \textbf{all the call stack queue to
+    complete abruptly.}
 \end{itemize}
+
+\note{The \texttt{invoke\_later} instruction can be used to call a smart contract while its reentracy lock is active.
+Because the invocation uses a separate call stack, it will not cause the lock to throw an exception.}
+
+Each time a method is invoked a new local frame
+and operand stack is created. If the method was invoked by an \texttt{i\_invoke} instruction,
+the \texttt{Save()} function of the AVM heap is called before the execution of the method starts.
+
+The Argennon Virtual Machine uses local frames to pass parameters on
+method invocation. On method invocation, any parameters are passed in consecutive local variables stored in the
+method's local frame starting from address 0. The invoker of a method writes the parameters in the local frame
+of the invoked method using \texttt{arg} instructions.
 
@@
 
-Each time a method is invoked the \texttt{Save()} function of the AVM heap is called, then a new local frame
-and operand stack is created. The Argennon Virtual Machine uses local frames to pass parameters on
-method invocation. On method invocation, any parameters are passed in consecutive local variables stored in the
-method's local frame starting from address 0. The invoker of a method writes the parameters in the local frame
-of the invoked method using \texttt{arg} instructions.
-
-\note{For AVM smart contracts \emph{reentrancy} can only happen in read-only mode and is totally safe.
-At the same time, call-back patterns can be easily implemented by using \texttt{spawn} instruction.}
-
 \subsection{Exceptions}\label{subsec:exceptions}
@@
 An exception is thrown programmatically using the \texttt{athrow} instruction. Exceptions can also be thrown by
-various Argennon Virtual Machine instructions if they detect an abnormal condition. Some exceptions are not
-catchable and will always abort the method invocation.
+various Argennon Virtual Machine instructions if they detect an abnormal condition.
 
-\note{By using the \texttt{athrow} instruction properly, a programmer can make any method act like an atomic
-operation.}
+\note{Any AVM instruction may throw an exception. This happens when the execution cost of a transaction has exceeded
+the declared execution cost of the transaction.}
 
@@
 directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
-method completes normally and the invocation was made by an \texttt{invoke} instruction, then a value may be
+method completes normally and the invocation was not made by an \texttt{invoke\_later} instruction, then a value may be
 returned to the invoking method. This occurs when the invoked method executes one of the return instructions, the
@@
 normally in the invoking method's local frame with the returned
-value (if any) pushed onto the operand stack. If the method was invoked by a \texttt{spawn} instruction,
+value (if any) pushed onto the operand stack. If the method was invoked by an \texttt{invoke\_later} instruction,
 the execution of the next call stack (if any) begins and the returned value
-is appended at the end of the transaction output. When a method invocation completes normally,
+is discarded. When a method invocation that was made by
+an \texttt{i\_invoke} instruction completes normally,
 the \texttt{Heap.Discard()} function is called as a part of the return instruction.
 
-A method invocation completes abruptly if an exception is thrown and is not caught by the current method. A
+A method invocation completes abruptly when an exception is thrown by one its instructions.
+The \texttt{invoke\_internal} and \texttt{invoke\_dispatcher} instructions
+always throw an exception when their invoked method completes abruptly. A
 method invocation that completes abruptly never returns a value.
 
-When a method completes, whether normally or abruptly, the call stack is used to restore the state of the invoker,
+When a method completes normally, the call stack is used to restore the state of the invoker,
 including its local frame and operand stack, with the \texttt{pc} register appropriately restored and incremented
@@
 
-A thrown exception causes methods in the call stack to complete \emph{abruptly} one by one, as long as the
-\texttt{pc} register is not pointing to a \texttt{catch} instruction. The \texttt{catch} instruction acts like a
-branch instruction that branches only if an exception is caught. When an exception is thrown
-Algorithm~\ref{alg:Abrupt} is used to restore the state of the AVM heap.
+When a method completes abruptly, the call stack is used to restore the \texttt{pc} register of the invoker. If the pc
+register points to an \texttt{invoke} instruction, the current method completes abruptly as well. Otherwise, if
+the \texttt{pc} register is pointing to an \texttt{i\_invoke} instruction, the \texttt{Heap.Restore()} function is
+called to restore the state of the AVM heap. Then, the
+local frame and operand stack of the invoker is restored, and execution continues after branching to the error handling
+code specified by the instruction operand.
 
-\begin{algorithm}
-    \DontPrintSemicolon
-    \SetKwData{Heap}{Heap}\SetKwData{Stack}{CallStack}
-    \SetKwFunction{Discard}{Discard}\SetKwFunction{Pop}{pop}\SetKwFunction{Restore}{Restore}
-    \SetKwProg{Fn}{Function}{}{}
+When an invocation which was made by an \texttt{invoke\_later} instruction completes abruptly, the
+heap snapshot that was saved before the start of the
+execution of the first call stack is restored. This means
+that all changes made by the execution session to the AVM heap will be completely reverted.
 
-    \While{\Stack is not empty}
-    {
-        \Stack.\Pop{}\;
-        \eIf{\Stack is not empty \textbf{and} $\Stack[head].pc \rightarrow catch$}
-        {
-            \Heap.\Restore{}\;
-        }{
-            \Heap.\Discard{}\;
-        }
-    }
-    \Heap.\Restore{}\;
-    \caption{Abrupt method completion}\label{alg:Abrupt}
-\end{algorithm}
+\subsection{Reentrancy Protection}\label{subsec:reentrancy}
 
-A closer investigation of Algorithm~\ref{alg:Abrupt} shows that when an invocation which was made by
-a \texttt{spawn} instruction completes abruptly, the heap snapshot that was saved before the start of the
-execution of the first call stack of the execution session is restored. Since the root
-smart contract calls \texttt{Heap.Save()} function before spawning the requested method, this means
-that all changes made to the AVM heap during the execution session will be restored, but the changes made by
-the root smart contract including the transferring of the transaction fee will not be restored. See
-Section~\ref{sec:execution-sessions} for more details.
+The Argennon Virtual Machine provides low level reentrancy protection by defining a dedicated
+instruction: \texttt{enter}. When a smart contract executes the \texttt{enter} instruction, it
+can not execute that instruction again as long
+as the method executing that instruction completes (abruptly or normally). If a smart contract tries to execute
+the \texttt{enter} instruction, while the method executing that instruction for the first time has not yet completed,
+an exception will be thrown.
+
+Smart contracts can completely prevent reentrancy by executing \texttt{enter} instruction as the first instruction
+of their \texttt{dispatcher} method.
 
@@
 blockchain.
-
-\section{Execution Sessions}\label{sec:execution-sessions}
-
-The execution of smart contracts code by the Argennon Virtual Machine is performed in execution sessions. Every
-execution session consists of a single method invocation from the \emph{root} smart contract and a list
-of resource allocation requests. These allocation requests always include a request for allocating some amount
-of execution cost and a request for accessing some memory locations for reading or writing. If after starting,
-an execution session tries to violate its allocated resources, an \textbf{uncatchable} exception will
-be thrown by the AVM\@.
-
-\note{The AVM does not have a \texttt{calldata} memory area. The arguments of the root smart contract method will
-be copied from the transaction to the local frame of the invoked method, exactly like a normal method invocation.}
-
-The root smart contract with the \texttt{applicationID} of zero, is a special smart contract in the Argennon
-Virtual Machine. Its code is mutable and is a part of the Argennon protocol. The root smart contract is always
-run in the privileged mode, and every Argennon transaction starts with a call to a method of the root smart contract.
-This method always transfers some amount of fee from an account to the \texttt{feeSink} accounts and then it
-performs the requested operation.
-
-If the requested operation is a method invocation, the root smart contract will call the \texttt{Heap.Save()}
-function, and then it will invoke the requested method using \texttt{spawn} instruction. This implies that the
-invoked method will always have a separate call stack, and if it completes abruptly the changes made by the root
-smart contract to the AVM state will not be reverted.
-
-\note{Argennon transactions do not have a sender, and the payer of the transaction fee could be any
-account who has provided the required digital signature for the fee payment.}
 

### 2021-07-11 tex/sections/draft.tex
Make the AVM RESTful
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 This will prevent smart contracts from profit taking by trading memory with the protocol.
+
+If the pc register points to an \texttt{invoke} instruction, the current method completes abruptly too.
+including its local frame and operand stack, with the \texttt{pc} register appropriately restored and incremented
+to skip past the method invocation instruction. If the current call stack is empty the next call stack in the call
+stack queue will be used for continuing the execution, and when there are no call stacks left in the queue the
+current execution session ends.
+
+When an exception is thrown
+Algorithm~\ref{alg:Abrupt} is used to restore the state of the AVM heap.
+
+\begin{algorithm}
+    \DontPrintSemicolon
+    \SetKwData{Heap}{Heap}\SetKwData{Stack}{CallStack}
+    \SetKwFunction{Discard}{Discard}\SetKwFunction{Pop}{pop}\SetKwFunction{Restore}{Restore}
+    \SetKwProg{Fn}{Function}{}{}
+
+    \While{\Stack is not empty}
+    {
+        \Stack.\Pop{}\;
+        \eIf{\Stack is not empty \textbf{and} $\Stack[head].pc \rightarrow catch$}
+        {
+            \Heap.\Restore{}\;
+        }{
+            \Heap.\Discard{}\;
+        }
+    }
+    \Heap.\Restore{}\;
+    \caption{Abrupt method completion}\label{alg:Abrupt}
+\end{algorithm}
+
+A closer investigation of Algorithm~\ref{alg:Abrupt} shows that when an invocation which was made by
+a \texttt{spawn} instruction completes abruptly, the heap snapshot that was saved before the start of the
+execution of the first call stack of the execution session is restored. Since the root
+smart contract calls \texttt{Heap.Save()} function before spawning the requested method, this means
+that all changes made to the AVM heap during the execution session will be restored, but the changes made by
+the root smart contract including the transferring of the transaction fee will not be restored. See
+Section~\ref{sec:execution-sessions} for more details.

### 2021-07-11 tex/sections/elements.tex
Make the AVM RESTful
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-Argennon has two special smart contracts: the \emph{root} smart contract\footnote{also called the root app} and
-the \emph{ARG} smart contract\footnote{also called the Argennon smart contract or the Argennon app.}.
-The root smart contract, with \texttt{applicationID=[0x0]}, is a privileged smart contract responsible for
-method invocation and installation/uninstallation of other smart contracts.
-On the other hand, the ARG smart contract, with \texttt{applicationID=[0x1]}, controls the ARG token, the main
-currency of the Argennon blockchain, and also manages a database of public keys and handles signature verification.
+Argennon has two special smart contracts: the \emph{root smart contract}, also called the \emph{root application}, and
+the \emph{ARG smart contract}, which is also called the Argennon smart contract or the \emph{ARG application}.
 
-Both of these smart contracts are mutable smart contracts and can be updated by the Argennon governance protocol.
+\subsection{The Root Application}\label{subsec:the-root-app}
+
+The root application or the root smart contract, with \texttt{applicationID=[0x0]}, is a privileged smart contract
+responsible for installation/uninstallation of other smart contracts. The Argennon root smart contract
+performs three main operations:
+
+\begin{itemize}
+    \item Installation of AVM smart contracts and determining the update policy of a smart
+    contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
+    on.
+    \item Removing AVM smart contracts.
+    \item Updating an AVM smart contract (if allowed).
+\end{itemize}
+
+The root smart contract is a mutable smart contract and can be updated by the Argennon governance system.
+(See Section~\ref{sec:adags})
+
+\subsection{The ARG Application}\label{subsec:the-arg-app}
+
+The ARG application or the ARG smart contract,
+with \texttt{applicationID=[0x1]}, controls the ARG token, the main
+currency of the Argennon blockchain. This smart contract also manages a database of public keys and
+handles signature verification.
+
+The ARG smart contract is a mutable smart contract and can be updated by the Argennon governance system.
 
@@
 
-Every Argennon transaction is a single \texttt{invoke\_external} instruction which calls a method from the Argennon
-root smart contract. This method always transfers the proposed fee of the transaction in ARGs from a
-sender account to the fee sink accounts and then performs the requested operation. The Argennon root smart contract
-has four public methods:
+Every Argennon transaction consists of two \texttt{i\_invoke\_dispatcher} instructions, the first instruction always
+transfers the proposed fee of the transaction in ARG from a sender account to the fee sink accounts, and
+the second performs the requested operation.
 
-\begin{itemize}
-    \item \texttt{avmCall}: invokes a method from an
-    AVM smart contract. Users interact with AVM smart contracts using this method of the root smart contract.
-    Transferring all assets, including ARG, is done by these transactions.
-    \item \texttt{installApp}: installs an AVM smart contract and determines the update policy of the smart
-    contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
-    on.
-    \item \texttt{unInstallApp}: removes an AVM smart contract.
-    \item \texttt{updateApp}: updates an AVM smart contract (if allowed).
-\end{itemize}
+Users interact with AVM smart contracts using
+the second \texttt{i\_invoke\_dispatcher} instruction of a transaction,
+Transferring all assets, including ARG, is done by that instruction.
 
@@
 Every Argennon transaction is required to specify a cap for all the resources it needs. This
-includes memory, network and processor related resources. The protocol defines an execution cost for
-every AVM instruction, reflecting the amount of resources its emulation needs, and every transaction is required
-to specify a maximum execution cost. If during emulation, a transaction reaches this maximum cost, it will be
-considered failed and the network can receive the proposed fee of that transaction.
+includes memory, network and processor related resources. When a transaction reaches one of its
+pre-declared resource caps, executing any AVM instruction which uses that resource, will result in an AVM exception.
+
+As we know, every Argennon transaction consists of two \texttt{i\_invoke\_dipatcher} instructions. The first
+instruction is always considered as a \emph{free} instruction and resources spent during its execution
+session will not be counted. As a result, only the second instruction could fail due to exceeding resource limits.
+
+The Argennon protocol defines an execution cost for
+every AVM instruction, reflecting the amount of resources its emulation needs. Every
+transaction is required to specify two maximum execution costs: \texttt{maxInternalCost}
+and \texttt{maxExternalCost}. The \emph{external} execution cost of a transaction is the \textbf{overall} cost of its
+\texttt{invoke\_dispatcher} and \texttt{invoke\_later} instructions,\footnote{By overall cost, we mean the execution
+cost needed for reaching the next instruction.} and the remaining execution cost will be considered as
+the \emph{internal} cost. If a transaction reaches one of its maximum costs, executing any instruction which has
+that type of cost, will throw an AVM exception.
+
+\note{When a transaction reaches its \texttt{maxExternalCost}, it can still keep executing
+its own code, while it can not call other smart contracts.
+This way the execution cost of a smart contract is completely decoupled from the
+smart contract it calls, and a malicious contract can not make its invoker fail unconditionally
+by using infinite loops.}
 
@@
 \begin{itemize}
-    \item Maximum execution cost
+    \item Maximum internal execution cost
+    \item Maximum external execution cost
     \item A list of heap/code-area locations for reading
@@
 
-Every Argennon transaction is required to pay two types of fees: execution fee and storage fee. A transaction pays
+Every Argennon transaction is required to pay two types of fees: execution fee, which is paid for executing the
+transaction, and storage fee, which is paid for the amount of storage the transaction allocates.
+
+A transaction pays
 its fees by providing digital signatures of one or more accounts, authorizing the transfer of the amount of fee in
-ARG from one or more accounts to the fee sink accounts.
+ARG from one or more accounts to the fee sink accounts, and the fee is transferred by the first
+\texttt{i\_invoke\_dispatcher} instruction of the transaction.
 

### 2021-07-12 tex/sections/elements.tex
Use HTTP for smart contracts
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 Every Argennon transaction consists of two \texttt{i\_invoke\_dispatcher} instructions, the first instruction always
-transfers the proposed fee of the transaction in ARG from a sender account to the fee sink accounts, and
-the second performs the requested operation.
+transfers the proposed fee of the transaction in ARGs from a sender account to the fee sink accounts, and
+the second performs the requested operation. If the first instruction fails, the transaction will not be added to
+the Argennon blockchain.
 
 Users interact with AVM smart contracts using
-the second \texttt{i\_invoke\_dispatcher} instruction of a transaction,
+the second \texttt{i\_invoke\_dispatcher} instruction of a transaction.
 Transferring all assets, including ARG, is done by that instruction.
 
-\subsection{Authorization}\label{subsec:txn-auth}
+The \texttt{i\_invoke\_dispatcher} instruction invokes the \texttt{dispatcher} method of a smart contract. Every AVM
+smart contract has a \texttt{dispatcher} method, and the Argennon protocol requires
+the \texttt{dispatcher} method to accept an HTTP request as its argument and
+return an HTTP response.
 
-Argennon transactions do not have a sender. The authorization of the requested operation is always done by checking the
-digital signatures that are provided as a part of the argument list to the root smart contract method.
+As a result, every Argennon transaction contains two HTTP requests. In addition to these HTTP requests, the
+transaction is required to contain a resource declaration object, specifying the maximum amount of resources
+it needs for execution.
 
-While every block of the Argennon blockchain stores the commitment of the transaction list, Argennon does not enforce
-storage of the transaction history. To be able to detect replay attacks, we require
-every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
-and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
-must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
-a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
-if a signature of the same user with a bigger or equal nonce is already used
-(i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
-nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
-
-To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. This information
-is stored in the ARG smart contract and when the difference between \texttt{issuance} component of the nonce and
-the current round becomes bigger than the maximum allowed lifetime of a signature, it can be safely
-deleted.\footnote{in some conventional
-blockchains, the nonce data can never be deleted, even if the account has zero balance and is no longer used.}
+\note{Argennon smart contracts use HTTP as the application protocol and they are advised to have a RESTful API design.}
 
@@
 
+\subsection{Authorization}\label{subsec:txn-auth}
+
+Argennon transactions do not have a sender. The authorization of the requested operation is always done by checking the
+digital signatures that are provided as a part of the HTTP request to the \texttt{dispatcher} method.
+
+While every block of the Argennon blockchain stores the commitment of the transaction list, Argennon does not enforce
+storage of the transaction history. To be able to detect replay attacks, we require
+every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
+and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
+must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
+a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
+if a signature of the same user with a bigger or equal nonce is already used
+(i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
+nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
+
+To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. This information
+is stored in the ARG smart contract and when the difference between \texttt{issuance} component of the nonce and
+the current round becomes bigger than the maximum allowed lifetime of a signature, it can be safely
+deleted.\footnote{in some conventional
+blockchains, the nonce data can never be deleted, even if the account has zero balance and is no longer used.}
+
 \subsection{Transaction Fee}\label{subsec:fee}

### 2021-07-12 tex/sections/argon.tex
Remove Argon chapter
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 %! Author = aybehrouz
+
+\lstset{
+    language=Java,
+    morekeywords={main, initialize, constructor, external, virtual, map, account, signature, overrides, is, string},
+    basicstyle=\small\sffamily,
+    columns=fullflexible,
+%   keywordstyle=\bfseries,
+    breaklines=true,
+    showstringspaces=false,
+    mathescape
+}
 
@@
 \begin{lstlisting}[frame=TB, float, title=Argon contracts are static classes,label={lst:code1}]
-// A contract is a static class and all its methods are automatically static.
-contract MirrorToken {
+// .
+public class MirrorToken {
     private SimpleToken token;
@@
     // `initialize` is a special static method that is called by the AVM after the code of a contract
-    // is stored in the AVM code area. It can not be called after that.
-    initialize(double supply1, double supply2) {
+    // is stored in the AVM code area.
+    public static void initialize(double supply1, double supply2) {
         // `new` does not create a new smart contract. It just makes an ordinary object.
@@
         reflection = new SimpleToken(supply2);
+    }
+    // `main` is the only method of the application (i.e. smart contract) that can be called by other
+    // applications. Every application should have exactly one main method defined in a class.
+    // Alternativly the keyword `dispatcher` could be used instead of `main`.
+    public static RestResponse main(RestRequest arg) {
+        if (arg.pathMatches("\balances\{user}")) {
+            account sender = arg.getParameter<account>("user");
+            if (arg.operationIsPUT()) {
+                verify(arg.toMessage(), sender, arg.getParameter<signature>("sig"));
+                token.transfer(sender, recipient, amount);
+                reflection.transfer(recipient, sender, Math.sqrt(amount));
+            } else if (arg.operationIsGET()) {
+            } else {
+            }
+        }
+        arg.operationIsGET();
+        arg.operationIsPUT();
+        arg.getParameter<address>("user")
     }
@@
     external double balanceOfReflection(account user) { return reflection.balanceOf(user); }
-}
-
-public static class Math {
-    public double sqrt(double x) {
-        return // calculate square root of x
-    }
 }

### 2021-07-12 tex/sections/argon.tex
Re-add Argon chapter
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2021-04-26 tex/sections/argon.tex
@@
 %! Author = aybehrouz
+
+\lstset{
+    language=Java,
+    morekeywords={main, initialize, constructor, external, virtual, Map, Account, signature, overrides, is, string},
+    basicstyle=\small\sffamily,
+    columns=fullflexible,
+%   keywordstyle=\bfseries,
+    breaklines=true,
+    showstringspaces=false,
+    mathescape
+}
 
@@
 
-Argon Programs are organized as sets of packages. Each package has its own set of names for types, which helps
-to prevent name conflicts. Every package can contain an arbitrary number of classes and a single contract. Every
-Argon package corresponds to at max one AVM smart contract.
+Argon applications (i.e.\ smart contracts) are organized as sets of packages. Each package has its own set of names
+for types, which helps to prevent name conflicts. Every package can contain an arbitrary number of classes.
+Every Argon application is required
+to have exactly one \texttt{main} method and one \texttt{initialize} method. The \texttt{main} method is the
+only method of an Argon application which can be called by other smart contracts.
+
+The \texttt{main} method is required to have a single parameter named \texttt{request}. The type of this parameter
+should be \texttt{RestRequest} or \texttt{HttpRequest}. The return value of the \texttt{main} function needs to be a
+\texttt{RestResponse} or \texttt{HttpResponse}.
 
 \section{Features Overview}\label{sec:features-overview}
-
-\subsection{Static Classes}\label{subsec:static-classes}
-
-In Argon a static class is an uninstantiable class which all its members and methods are static. Every static
-class can define a special method \texttt{initialize} which can only be called in the \texttt{initialize}
-method of the single contract of the package the static class is defined in.
-
-A Static class is like a normal class which can only have one instance. So like normal classes, static classes can
-be a super class or a subclass of other static classes and can implement interfaces.
-
-\subsection{Contracts}\label{subsec:contract}
-
-Every package of an Argon program can have one or zero contracts. A contact is a special static class which is
-allowed to define methods with \texttt{external} visibility. When an Argon package is deployed as an AVM smart
-contract, these external methods will define the contract's public interface.
 
@@
 \begin{center}
-    \begin{tabular}{llllll}
+    \begin{tabular}{lllll}
         \hline
-        & Class & Package & Subclass & Program & World \\
+        & Class & Package & Subclass & Program \\
         \hline
-        private   & yes   & no      & no       & no      & no    \\
-        protected & yes   & no      & yes      & no      & no    \\
-        package   & yes   & yes     & yes      & no      & no    \\
-        public    & yes   & yes     & yes      & yes     & no    \\
-        external  & no    & no      & no       & no      & yes   \\
+        private   & yes   & no      & no       & no  \\
+        protected & yes   & no      & yes      & no  \\
+        package   & yes   & yes     & yes      & no  \\
+        public    & yes   & yes     & yes      & yes \\
         \hline
@@
 
-\begin{lstlisting}[frame=TB, float, title=Argon contracts are static classes,label={lst:code1}]
-// A contract is a static class and all its methods are automatically static.
-contract MirrorToken {
-    private SimpleToken token;
-    private SimpleToken reflection;
+\begin{lstlisting}[frame=TB, float, title=A simple Argon application,label={lst:code1}]
+public class MirrorToken {
+    private static SimpleToken token;
+    private static SimpleToken reflection;
 
     // `initialize` is a special static method that is called by the AVM after the code of a contract
-    // is stored in the AVM code area. It can not be called after that.
-    initialize(double supply1, double supply2) {
+    // is stored in the AVM code area.
+    public static void initialize(double supply1, double supply2) {
         // `new` does not create a new smart contract. It just makes an ordinary object.
@@
     }
-
-    external void transfer(account sender, account recipient,
-        double amount, signature sig) {
-        // Checks if the signature is issued by the sender and verifies the calling of this function
-        // with the current parameters.
-        sig.verifyCallBy(sender);
+    // `main` is the only method of the application (i.e. smart contract) that can be called
+    // by other applications. Every application should have exactly one main method defined
+    // in some class. Alternatively, the keyword `dispatcher` could be used instead of `main`.
+    public static RestResponse main(RestRequest request) {
+        RestResponse response = new RestResponse();
+        if (request.pathMatches("/balances/{user}")) {
+            Account sender = request.getParameter<Account>("user");
+            if (request.operationIsPUT()) {
+                sender.authorize(request.toMessage(), request.getParameter<byte[]>("sig"));
+                Account recipient = request.getParameter<Account>("to");
+                double amount = request.getParameter<double>("amount");
                 token.transfer(sender, recipient, amount);
                 reflection.transfer(recipient, sender, Math.sqrt(amount));
+                return response.setStatus(Http.Status.OK);
+            } else if (request.operationIsGET()) {
+                response.append<double>("balance", token.balanceOf(sender));
+                response.append<double>("reflection", reflection.balanceOf(user));
+                return response.setStatus(Http.Status.OK);
+            } else {
+                return response.setStatus(Http.Status.MethodNotAllowed);
             }
-
-    external double balanceOf(account user) { return token.balanceOf(user); }
-
-    external double balanceOfReflection(account user) { return reflection.balanceOf(user); }
         }
-
-public static class Math {
-    public double sqrt(double x) {
-        return // calculate square root of x
     }
@@
 package class SimpleToken {
-    private map(account -> double) balances;
+    private Map(Account -> double) balances;
 
-    // The visibility of a member without an access modifier will be the same as its defining
-    // class so `SimpleToken` constructor has package visibility.
+    // The visibility of a member without an access modifier will be the package level.
     constructor(double initialSupply) {
@@
 
-    void transfer(account sender, account recipient, double amount) {
+    void transfer(Account sender, Account recipient, double amount) {
         if (balances[sender] < amount) throw("Not enough balance.");

### 2021-07-13 tex/sections/avm.tex
Add the virtual signature concept
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
-passing cryptographic signatures as parameters to methods, a programmer can implement the required logic
-for authorizing any operation.
+passing cryptographic signatures as parameters to methods, a programmer, having user's public keys, can implement
+the required logic for authorizing any operation.
+
+\note{Authorization by explicit signatures eliminates the need for approval methods or call back patterns.}
+
+In addition to cryptographic signatures, the Argennon Virtual Machine provides instructions that enable
+smart contracts to issue \emph{virtual signatures}. A smart contract can virtually sign a message for a target
+smart contract by \texttt{v\_sign} instruction. When a smart contract
+virtually signs a message for a target, that message is stored in the internal memory of the AVM, associated with
+its signer and the target. The application that is the target of the signature can verify the message
+by \texttt{v\_verify} instruction. After verifying the signature, it will be removed from the AVM's internal memory
+to make sure a signature will not be used multiple times. The virtual signatures stored in the AVM's internal memory
+last for one execution session.
 
 \note{The Argennon virtual machine has no instructions for issuing cryptographic signatures.}
-
-In addition to signatures, a method can verify its invoker by using \texttt{get\_parent} instruction. This
-instruction gets the \texttt{applicationID} of the smart contract that is one level deeper than the current
-smart contract in the call stack or if there is none, the smart contract that created the current call stack.
-In other words, it returns the \texttt{applicationID} of the smart contract that has invoked or spawned
-the current smart contract.
 

### 2021-07-13 tex/sections/elements.tex
Add the virtual signature concept
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 \begin{verbatim}
-<application-id> ::= "["<hex-prefix-code>"]"
-<hex-prefix-code> ::= "0x"<hex-num>"."<hex-prefix-code> | "0x"<hex-num>
+<application-id> ::= <decimal-prefix-code>
+<decimal-prefix-code> ::= <dec-num>"."<decimal-prefix-code> | <dec-num>
 \end{verbatim}
-where \texttt{<hex-num>} is a hexadecimal number between $0$ and $255$, using lower case
-letters \texttt{[a-f]} for showing digits greater than $9$.
+where \texttt{<dec-num>} is a normal decimal number between $0$ and $255$.
 
-For example \texttt{[0x24.0xff.0xda]}, \texttt{[0x0]} and \texttt{[0x3.0xa0.0x0.0x0]}, are valid application addresses.
+For example \texttt{21.255.37}, \texttt{0} and \texttt{1.0.0.0.0}, are valid application addresses.
 
@@
 
-The root application or the root smart contract, with \texttt{applicationID=[0x0]}, is a privileged smart contract
+The root application or the root smart contract, with \texttt{applicationID = 0}, is a privileged smart contract
 responsible for installation/uninstallation of other smart contracts. The Argennon root smart contract
@@
 The ARG application or the ARG smart contract,
-with \texttt{applicationID=[0x1]}, controls the ARG token, the main
+with \texttt{applicationID = 1}, controls the ARG token, the main
 currency of the Argennon blockchain. This smart contract also manages a database of public keys and
@@
 \begin{verbatim}
-<account-id> ::= "["<decimal-prefix-code>"]"
-<decimal-prefix-code> ::= <dec-num>"."<decimal-prefix-code> | <dec-num>
+<account-id> ::= <hex-prefix-code>
+<hex-prefix-code> ::= "0x"<hex-num>"."<hex-prefix-code> | "0x"<hex-num>
 \end{verbatim}
-where \texttt{<dec-num>} is a normal decimal number between $0$ and $255$.
+where \texttt{<hex-num>} is a hexadecimal number between $0$ and $255$, using lower case
+letters \texttt{[a-f]} for showing digits greater than $9$.
 
-For example \texttt{[21.255.37]}, \texttt{[0]} and \texttt{[1.0.0.0.0]}, are valid standard symbolic representations
-of account addresses.
+For example \texttt{0x24.0xff.0xda}, \texttt{0x0} and \texttt{0x3.0xa0.0x0.0x0}, are valid standard symbolic
+representations of account addresses.
 
@@
     \item A list of heap locations for writing
-    \item A list of heap chunks it will deallocate
+    \item A list of heap chunks it will deallocate (if any)
     \item A list of methods it will delete (if any)
-    \item Number and size of heap chunks it will allocate
+    \item Number and size of heap chunks it will allocate (if any)
     \item Number and size of method bytecodes it will allocate (if any)
@@
 
+\begin{lstlisting}[language=python, frame=TB, float, title=An Argennon transaction in YAML format,label={lst:txn-example}]
+---
+fee: |
+    PUT /balances/0x73.0xa2?to=fee&amount=0.26&sig=5b73CbmwQNRC7fWUY15 HTTP/1.1
+call: |
+    POST httpb://bac.argennon.net/54.189.21/proposals HTTP/1.1
+    Content-Type: application/json; charset=utf-8
+    Content-Length: 69
+
+    {
+        "name": "Grant Proposal",
+        "recipient": "0x24.0x8f.0x29.0xa1",
+        "amount": 25000,
+        "sig"= "2a36Gtrw249wQCD70nWY49d"
+    }
+caps:
+    internal: 2500000 # maximum number of AVM execution clocks
+    external: 1000000
+    read: [(2654,3),(15642,0),(15642,1),(15642,3)]
+    write: [(15642,0),(20154,0),(20154,1)]
+\end{lstlisting}
+
 \subsection{Transaction Fee}\label{subsec:fee}

### 2021-07-13 tex/sections/elements.tex
Add RESTful properties to README.md
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 call: |
-    POST httpb://bac.argennon.net/54.189.21/proposals HTTP/1.1
+    POST bttp://bac.argennon.net/54.189.21/proposals HTTP/1.1
     Content-Type: application/json; charset=utf-8

### 2021-07-14 tex/sections/argon.tex
Add the virtual signature concept
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The Argon programming language is a class-based, object-oriented language designed for writing Argennon's smart
+The Argon programming language is a class-based, object-oriented language designed for writing Argennon smart
 contracts. The Argon programming language is inspired by Solidity and is similar to Java, with a number of aspects
@@
 to have exactly one \texttt{main} method and one \texttt{initialize} method. The \texttt{main} method is the
-only method of an Argon application which can be called by other smart contracts.
+only method of an Argon application which would be called by other smart contracts.
 
@@
 
-Access level modifiers determine whether other classes can use a particular field or invoke a particular method or
-if a method can be invoked externally by other smart contracts.
+Access level modifiers determine whether other classes can use a particular field or invoke a particular method.
 

### 2021-07-14 tex/sections/avm.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The Argennon\footnote{the classical pronunciation should be used:\textipa{/Ar"gen.non/}} Virtual Machine (AVM) is an
-abstract computing machine for executing Argennon's smart contracts. The
+abstract computing machine for executing Argennon smart contracts. The
 Argennon Virtual Machine knows nothing of the Argennon blockchain, transactions or any concept of gas usage, only of
-Argennon's identifier tries and the concept of execution sessions.
+Argennon identifier tries and the concept of execution sessions.
 
 Execution sessions are separate sessions of executing smart contract's code by the Argennon Virtual Machine.
-Every execution session starts with a single method invocation, but does not necessarily end when
-that method invocation completes. During an execution session additional method invocations could be scheduled,
-which will be executed after the initial method completion.
+Every execution session starts with a single method invocation, but does not necessarily end with
+that method invocation. During an execution session additional method invocations could be scheduled,
+which will be executed after the initial method completes.
 
 An execution session may complete normally or abruptly. When an execution session completes abruptly, it will
-not have any effects on the AVM state.
+not have any effects on the AVM's state.
 
@@
     \item \texttt{applicationID} is a prefix code built by \emph{applications} prefix tree.
-    \item \texttt{accountID} is a prefix code built by \emph{accounts} prefix tree.
+    \item \texttt{accountID} is a prefix code built by \emph{accounts} prefix tree. An \texttt{accountID} can
+    not be \texttt{0x0} or \texttt{0x1}.
     \item \texttt{methodID} is a composite prefix code built by concatenating an \texttt{applicationID} to
     a prefix code made by \emph{local} prefix tree: (\texttt{|} is the concatenating operator)
-
-    \texttt{methodID = (applicationID|<local-prefix-code>)}
+    \subitem \texttt{methodID = (applicationID|<local-prefix-code>)} .
     \item \texttt{chunkID} is a composite prefix code built by concatenating an \texttt{applicationID} to
     an \texttt{accountID} to a prefix code made by \emph{local} prefix tree:
-
-    \texttt{chunkID = (applicationID|accountID|<local-prefix-code>)}
+    \subitem \texttt{chunkID = (applicationID|accountID)} or
+    \subitem \texttt{chunkID = (applicationID|0x0|<local-prefix-code>)} or
+    \subitem \texttt{chunkID = (applicationID|0x1|accountID|<local-prefix-code>)} .
 \end{itemize}
@@
 Every AVM execution session has a queue of call stacks. A call stack contains all the information that is needed for
-restoring the AVM state, and continuing the execution after the method invocation completes.
+restoring the AVM's state and continuing the execution after the method invocation completes.
 This information is represented by a \texttt{CallInfo} struct:
@@
     operandStack
-}.
+} .
 \end{verbatim}
@@
 
-The Argennon Virtual Machine defines various run-time data areas that are used during an execution session
-(i.e.\ transaction), Some of these data areas are persistent and are stored on the blockchain, Other data areas
+The Argennon Virtual Machine defines various run-time data areas that are used during an execution session,
+Some of these data areas are persistent and are stored on the blockchain, Other data areas
 are per execution session or per method invocation. These data areas are created when their context starts and
@@
 except the \texttt{dispatcher} method, which is the only external method of any smart contract.
-This design enables the AVM to execute smart contracts with \emph{RESTful} APIs more efficiently.}
+This design enables the AVM to execute \emph{RESTful} smart contracts more efficiently.}
 
@@
 A smart contract can read any chunk stored in the heap by having its \texttt{chunkID}. However, it can only modify
-those chunks that was created by itself. In other words, every memory chunk in the AVM heap has an owner. Only
+those chunks that was created by itself. In other words, every memory chunk in the AVM's heap has an owner. Only
 the owner can modify a chunk while anyone can read that chunk. The owner of a chunk can be easily determined
@@
     internal method of another smart contract. Since the invoked method will always be executed in the context of
-    the invoking smart contract, a smart contract will not be able to modify another smart contract state by
+    the invoking smart contract, a smart contract will not be able to modify another smart contract's state by
     this instruction. This instruction facilitates code reuse and the usage of libraries.
-    If the invoked method completes abruptly, that will cause the invoker method to complete abruptly too.
+    If the invoked method completes abruptly, the invoker method completes abruptly too.
     \item \texttt{invoke\_dispatcher}: invokes the \texttt{dispatcher} method of a smart contract and changes the
@@
     an \texttt{i\_invoke} instruction completes abruptly, all changes made
-    by the invoked method to the AVM state will be reverted and the instruction will act like
+    by the invoked method to the AVM's state will be reverted and the instruction will act like
     a jump to the specified error handling code.
@@
     instruction, high performance native methods of the hosting machine, such as GPU accelerated operations,
-    could become available to AVM smart contracts. This instruction will not modify the AVM state.
+    could become available to AVM smart contracts. This instruction will not modify the AVM's state.
     \item \texttt{invoke\_later}: invokes the \texttt{dispatcher} method of another smart contract by creating a new
@@
 and operand stack is created. If the method was invoked by an \texttt{i\_invoke} instruction,
-the \texttt{Save()} function of the AVM heap is called before the execution of the method starts.
+the \texttt{Save()} function of the AVM's heap is called before the execution of the method starts.
 
@@
 \note{The AVM does not natively support polymorphism and virtual methods. A compiler could easily
-generates appropriate code for implementing these features.}
+generate appropriate code for implementing these features.}
 
@@
 \note{Any AVM instruction may throw an exception. This happens when the execution cost of a transaction has exceeded
-the declared execution cost of the transaction.}
+the declared maximum execution cost.}
 
@@
 the \texttt{pc} register is pointing to an \texttt{i\_invoke} instruction, the \texttt{Heap.Restore()} function is
-called to restore the state of the AVM heap. Then, the
+called to restore the state of the AVM's heap. Then, the
 local frame and operand stack of the invoker is restored, and execution continues after branching to the error handling
-code specified by the instruction operand.
+code specified by the instruction's operand.
 
@@
 execution of the first call stack is restored. This means
-that all changes made by the execution session to the AVM heap will be completely reverted.
+that all changes made by the execution session to the AVM's heap will be completely reverted.
 
@@
 can not execute that instruction again as long
-as the method executing that instruction completes (abruptly or normally). If a smart contract tries to execute
-the \texttt{enter} instruction, while the method executing that instruction for the first time has not yet completed,
+as the method executing the instruction completes (abruptly or normally). If a smart contract tries to execute
+the \texttt{enter} instruction while the method executing that instruction for the first time has not yet completed,
 an exception will be thrown.
 
-Smart contracts can completely prevent reentrancy by executing \texttt{enter} instruction as the first instruction
+Smart contracts can completely prevent reentrancy by executing the \texttt{enter} instruction as the first instruction
 of their \texttt{dispatcher} method.
@@
 \begin{itemize}
-    \item \texttt{alloc}: creates a new memory chunk of the specified size in the heap. the identifier
-    of the new chunk will be the concatenation of the \texttt{applicationID} of the creator of the chunk and
+    \item \texttt{alloc}: creates a new memory chunk of the specified size in the heap. The
+    new chunk's identifier will be the concatenation of the \texttt{applicationID} of the creator of the chunk and
     the \texttt{id} operand of the instruction. The \texttt{id} operand must be the concatenation of
@@
     generated by the \emph{local} prefix tree: \texttt{id = accountID|localID}. If the identifier is not valid or
-    if it already exists in the heap a catchable exception will be thrown.
+    if it already exists in the heap, a catchable exception will be thrown.
 
@@
     contract \texttt{applicationID} and the specified \texttt{id} operand. If the identifier is not valid or
-    if the chunk was not found a catchable exception will be thrown.
+    if the chunk was not found, a catchable exception will be thrown.
 \end{itemize}
@@
 When a smart contract allocates a new memory chunk, the identifier of the new chunk is not generated by
-the AVM, instead the smart contract can choose an identifier itself. This is a very important feature of
-the AVM heap, which allows smart contracts to use the AVM heap as a dictionary (map) data structure.
-Since the \texttt{chunkID} is a prefix code, any smart contract has its own identifier space and a smart contract
+the AVM. Instead, the smart contract can choose an identifier itself. This is a very important feature of
+the AVM's heap, which allows smart contracts to use the AVM's heap as a dictionary (map) data structure.
+Since the \texttt{chunkID} is a prefix code, any smart contract has its own identifier space, and a smart contract
 can easily generate unique identifiers for its chunks.
 
-The AVM requires the chosen identifier to be a prefix code made by concatenating a code compatible with
+The AVM requires the chosen identifier to be a prefix code compatible with
+the \emph{accounts} prefix tree, or a code made by concatenating a code compatible with
 the \emph{accounts} prefix tree to a code compatible with the \emph{local} prefix tree. This requirement
@@
 from a user to another user, first we need to make sure that the sender has authorized this operation. The
-Argennon virtual machine has no built-in mechanism for authorizing operations, but it provides a rich set of
+Argennon Virtual Machine has no built-in mechanism for authorizing operations, but it provides a rich set of
 cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
-passing cryptographic signatures as parameters to methods, a programmer, having user's public keys, can implement
+passing cryptographic signatures as parameters to methods, a programmer, having users' public keys, can implement
 the required logic for authorizing any operation.
@@
 
-In Argennon, the root smart contract is an updatable smart contract, which can be updated by the Argennon consensus
-protocol. This means that bugs or security vulnerabilities in the AVM standard library could be quickly patched and
+In Argennon, the root smart contract is an updatable smart contract, which can be updated by the Argennon governance
+system. This means that bugs or security vulnerabilities in the AVM standard library could be quickly patched and
 smart contracts could benefit from bugfixes and improvements of the AVM standard library even if they are

### 2021-07-14 tex/sections/elements.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 Argennon has two special smart contracts: the \emph{root smart contract}, also called the \emph{root application}, and
-the \emph{ARG smart contract}, which is also called the Argennon smart contract or the \emph{ARG application}.
+the \emph{ARG smart contract}, which is also called the \emph{Argennon smart contract} or the \emph{ARG application}.
 
@@
 The root application or the root smart contract, with \texttt{applicationID = 0}, is a privileged smart contract
-responsible for installation/uninstallation of other smart contracts. The Argennon root smart contract
+responsible for installation/uninstallation of other smart contracts. The Argennon's root smart contract
 performs three main operations:
@@
 
-As a result, every Argennon transaction contains two HTTP requests. In addition to these HTTP requests, the
+As a result, every Argennon transaction contains two HTTP requests. Besides these HTTP requests, the
 transaction is required to contain a resource declaration object, specifying the maximum amount of resources
@@
 instruction is always considered as a \emph{free} instruction and resources spent during its execution
-session will not be counted. As a result, only the second instruction could fail due to exceeding resource limits.
+session will not be counted. Thus, only the second instruction could fail due to exceeding resource limits.
 
@@
 \texttt{invoke\_dispatcher} and \texttt{invoke\_later} instructions,\footnote{By overall cost, we mean the execution
-cost needed for reaching the next instruction.} and the remaining execution cost will be considered as
+cost needed for reaching the next instruction.}. The remaining execution cost will be considered as
 the \emph{internal} cost. If a transaction reaches one of its maximum costs, executing any instruction which has
@@
 This way the execution cost of a smart contract is completely decoupled from the
-smart contract it calls, and a malicious contract can not make its invoker fail unconditionally
+smart contract it calls, and a malicious contract can not make its invoker fail
 by using infinite loops.}
@@
 Also, Argennon transactions are required to specify what heap or code area addresses they will access. This will
-enable validators to parallelize transaction validation as we will see in Section~\ref{sec:concurrency}. A transaction
-that tries to access a memory location that is not in its access list, will be rejected.
-Users could use off-chain \emph{smart contract oracles} to predict the list of memory locations their transactions need.
+enable validators to parallelize transaction validation as we will see in Section~\ref{sec:concurrency}. An instruction
+that tries to access a memory location that is not in the access list of the transaction, will throw an exception.
+Users could use off-chain \emph{execution oracles} to predict the list of memory locations their transactions need.
 
-A smart contract oracle is a full AVM emulator that keeps a full local copy of the AVM storage and can emulate AVM
-execution without accessing a ZK-EDB server. Smart contract oracles can be used for reporting useful information about
+An execution oracle is a full AVM emulator that keeps a full local copy of the AVM storage and can emulate AVM
+execution without accessing a ZK-EDB server. Execution oracles can be used for reporting useful information about
 Argennon transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
@@
 can receive the proposed fee of that transaction.
+
+\begin{lstlisting}[language=python, frame=TB, float, title=An Argennon transaction in YAML format,label={lst:txn-example}]
+---
+fee: |
+    PUT /balances/0x73.0xa2?to=fee&amount=0.26&sig=5b73CbmwQNRC7fWUY15 HTTP/1.1
+call: |
+    POST bttp://app.argennon.net/54.189.21/proposals HTTP/1.1
+    Content-Type: application/json; charset=utf-8
+    Content-Length: 77
+
+    {
+        "name": "Grant",
+        "recipient": "0x24.0x8f.0x29.0xa1",
+        "amount": 25000,
+        "sig"= "2a36Gtrw249wQCD70nWY49d"
+    }
+caps:
+    internal: 2500000 # maximum number of AVM execution clocks
+    external: 1000000
+    read: [(2654,3),(15642,0),(15642,1),(15642,3)]
+    write: [(15642,0),(20154,0),(20154,1)]
+\end{lstlisting}
 
@@
 
-\begin{lstlisting}[language=python, frame=TB, float, title=An Argennon transaction in YAML format,label={lst:txn-example}]
----
-fee: |
-    PUT /balances/0x73.0xa2?to=fee&amount=0.26&sig=5b73CbmwQNRC7fWUY15 HTTP/1.1
-call: |
-    POST bttp://bac.argennon.net/54.189.21/proposals HTTP/1.1
-    Content-Type: application/json; charset=utf-8
-    Content-Length: 69
-
-    {
-        "name": "Grant Proposal",
-        "recipient": "0x24.0x8f.0x29.0xa1",
-        "amount": 25000,
-        "sig"= "2a36Gtrw249wQCD70nWY49d"
-    }
-caps:
-    internal: 2500000 # maximum number of AVM execution clocks
-    external: 1000000
-    read: [(2654,3),(15642,0),(15642,1),(15642,3)]
-    write: [(15642,0),(20154,0),(20154,1)]
-\end{lstlisting}
-
 \subsection{Transaction Fee}\label{subsec:fee}
@@
 its fees by providing digital signatures of one or more accounts, authorizing the transfer of the amount of fee in
-ARG from one or more accounts to the fee sink accounts, and the fee is transferred by the first
+ARGs from one or more accounts to the fee sink accounts. This fee is transferred by the first
 \texttt{i\_invoke\_dispatcher} instruction of the transaction.
@@
 \note{An Argennon transaction always pays all of its proposed fee, no matter how much of its predefined resources
-were not used in the final emulation. This will incentivize users to report the resource usage of their
+were not used during the final emulation. This will incentivize users to report the resource usage of their
 transactions more accurately.}

### 2021-07-16 tex/sections/persistence.tex
Fix a readability issue
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-A page of the AVM storage should contain objects that have very similar access pattern. We expect that when a page
-is needed for validating a block, almost all of its objects are needed for either reading or writing. We also prefer
+A page of the AVM storage should contain objects that have very similar access pattern. Ideally, when a page is needed
+for validating a block, almost all of its objects should be needed for either reading or writing. We prefer
 that the objects are needed for the same access type. In other words, the objects of a page are chosen in a way that

### 2021-07-19 tex/sections/consensus.tex
Fix a readability issue
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 As we mentioned before, the block with height $n$ needs a certificate from the committee of
-validators with index $n$ modulo $m$. To decide about signing a block certificate, a validator
-checks the conditional validity\footnote{See Section~\ref{subsec:block-validation}} of the block, and
+validators with index $n$ modulo $m$. To decide about signing the certificate of a block which already has
+a certificate from the delegates, a validator checks the conditional
+validity\footnote{See Section~\ref{subsec:block-validation}} of the block, and
 if the block is valid he issues

### 2021-07-20 tex/sections/elements.tex
Fix some readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-For example \texttt{21.255.37}, \texttt{0} and \texttt{1.0.0.0.0}, are valid application addresses.
+For example \texttt{21.255.37}, \texttt{0}, \texttt{11.6} and \texttt{2.0.0.0.0}, are valid application addresses.
 
@@
     \item Installation of AVM smart contracts and determining the update policy of a smart
-    contract: if the contract is updatable or not, which accounts can update or uninstall the contract, and so
-    on.
-    \item Removing AVM smart contracts.
+    contract: if the contract is updatable or not, which accounts or smart contracts can update or uninstall
+    the contract, and so on.
+    \item Removing an AVM smart contract (if allowed).
     \item Updating an AVM smart contract (if allowed).
@@
 As a result, every Argennon transaction contains two HTTP requests. Besides these HTTP requests, the
-transaction is required to contain a resource declaration object, specifying the maximum amount of resources
+transaction is also required to contain a resource declaration object, specifying the maximum amount of resources
 it needs for execution.
@@
 Every Argennon transaction is required to specify a cap for all the resources it needs. This
-includes memory, network and processor related resources. When a transaction reaches one of its
+includes memory, network and processor related resources. If during emulation, a transaction reaches one of its
 pre-declared resource caps, executing any AVM instruction which uses that resource, will result in an AVM exception.
@@
 instruction is always considered as a \emph{free} instruction and resources spent during its execution
-session will not be counted. Thus, only the second instruction could fail due to exceeding resource limits.
+session will not be counted. Therefore, only the second instruction could fail due to exceeding resource limits.
 
@@
 cost needed for reaching the next instruction.}. The remaining execution cost will be considered as
-the \emph{internal} cost. If a transaction reaches one of its maximum costs, executing any instruction which has
-that type of cost, will throw an AVM exception.
+the \emph{internal} cost. If a transaction reaches one of its maximum execution costs, executing any instruction
+which has that type of cost, will throw an AVM exception.
 
-\note{When a transaction reaches its \texttt{maxExternalCost}, it can still keep executing
+\note{When a transaction reaches its \texttt{maxExternalCost}, it can still execute
 its own code, while it can not call other smart contracts.
 This way the execution cost of a smart contract is completely decoupled from the
-smart contract it calls, and a malicious contract can not make its invoker fail
+smart contract it calls, and a malicious contract can not make its invoker certainly fail
 by using infinite loops.}
@@
 call: |
-    POST bttp://app.argennon.net/54.189.21/proposals HTTP/1.1
+    POST bttp://dapp.argennon.net/54.189.21/proposals HTTP/1.1
     Content-Type: application/json; charset=utf-8

### 2021-07-26 tex/sections/avm.tex
Add identifier matching algorithm
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 that method invocation. During an execution session additional method invocations could be scheduled,
-which will be executed after the initial method completes.
+which will be executed after the initial method invocation completes.
 
@@
 
-The Argennon virtual machine does not have a fixed word size. Any instruction of the AVM has
+The Argennon Virtual Machine does not have a fixed word size. Any instruction of the AVM has
 its specific word size, and the addressable memory areas are byte addressable.
@@
 \emph{applications, accounts} and \emph{local}. Any identifier in Argennon is a
-prefix code built by using one or more of these prefix trees:
+prefix code built by one or more of these prefix trees:
 \begin{itemize}
@@
 uniquely extractable from any sequence. Assume that we have a string of digits in base $\beta$, we
-know that $k$ first digits belong to an Argennon's identifier, but we don't know the value of $k$.
+know that $k$ first digits belong to an Argennon identifier, but we don't know the value of $k$.
 Algorithm~\ref{alg:prefix_id} can be used to extract the prefixed identifier uniquely. Also, we can apply this algorithm
@@
 %##\includegraphics[width=17cm]{../img/Alg1s.png}
-\begin{algorithm}
+\begin{algorithm}[t]
     \DontPrintSemicolon
-    \SetKwData{Ready}{$R_e$}\SetKwData{V}{$v_f$}\SetKwData{Graph}{$G_e$}\SetKwData{Vertices}{$V$}\SetKwData
-    {Txns}{$T$}
     \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
     \Input{A sequence of $n$ digits in base $\beta$: $d_{1}d_{2}\dots d_{n}$ \newline
-    A prefix tree: $<A^{(1)},A^{(2)},A^{(3)},\dots>$}
+    A prefix tree: $<0,A^{(1)},A^{(2)},A^{(3)},\dots>$}
+    \BlankLine
     \Output{Valid identifier prefix of the sequence.}
@@
 
+When we have a prefixed identifier, and we want to know if a sequence of digits is marked by that identifier,
+we use Algorithm~\ref{alg:match_id} to match the prefixed identifier with the start of the sequence. The matching
+can be done with only three comparisons and an invalid prefixed identifier can also be detected.
+
 In Argennon the shorter prefix codes are assigned to more active accounts and smart contracts which tend to own more
@@
 of leaves in each level is chosen appropriately.
+
+\begin{algorithm}[h]
+    \DontPrintSemicolon
+    \SetKwData{Id}{$id$}
+    \SetKwInOut{Input}{input}\SetKwInOut{Output}{output}
+    \Input{A prefixed identifier in base $\beta$ with $n$ digits: $\Id=a_{1}a_{2}\dots a_{n}$ \newline
+    A sequence of digits in base $\beta$: $d_{1}d_{2}d_{3}\dots $ \newline
+    A prefix tree: $<0,A^{(1)},A^{(2)},A^{(3)},\dots>$
+    }
+    \BlankLine
+    \Output{$TURE$ if and only if the identifier is valid and the sequence starts with the identifier.}
+    \BlankLine
+    \If{$(0.a_{1}\dots a_{n})_\beta = (0.d_{1}\dots d_{n})_\beta$}
+    {
+        \If{$A^{(n-1)} \leq (0.a_{1}a_{2}\dots a_{n})_\beta < A^{(n)}$}
+        {
+            \KwRet{TRUE}\;
+        }
+    }
+    \KwRet{FALSE}\;
+    \caption{Matching a prefixed identifier}\label{alg:match_id}
+\end{algorithm}
 

### 2021-07-28 tex/sections/avm.tex
Fix some readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The Argennon Virtual Machine does not have a fixed word size. Any instruction of the AVM has
-its specific word size, and the addressable memory areas are byte addressable.
+The Argennon Virtual Machine does not have a fixed word size and uses \textbf{little-endian} byte order. Any
+instruction of the AVM has its specific word size, and the addressable memory areas are byte addressable.
 
@@
 uniquely extractable from any sequence. Assume that we have a string of digits in base $\beta$, we
-know that $k$ first digits belong to an Argennon identifier, but we don't know the value of $k$.
+know that the sequence starts with an Argennon identifier, but we do not know the length of that identifier.
 Algorithm~\ref{alg:prefix_id} can be used to extract the prefixed identifier uniquely. Also, we can apply this algorithm
@@
     \Input{A sequence of $n$ digits in base $\beta$: $d_{1}d_{2}\dots d_{n}$ \newline
-    A prefix tree: $<0,A^{(1)},A^{(2)},A^{(3)},\dots>$}
+    A prefix tree: $<A^{(1)},A^{(2)},A^{(3)},\dots>$}
     \BlankLine
@@
 we use Algorithm~\ref{alg:match_id} to match the prefixed identifier with the start of the sequence. The matching
-can be done with only three comparisons and an invalid prefixed identifier can also be detected.
+can be done with only three comparisons, and an invalid prefixed identifier can be detected and will not match
+any sequence.
 
@@
 
-The \texttt{methodID} field is the unique identifier of a method's bytecode that
+The \texttt{methodID} field is the unique identifier of a method's bytecode, which
 is a composite prefix code built by concatenating an \texttt{applicationID} to
@@
 \end{itemize}
-All data areas except operand stack, have their own address space. Operand stack is a
+All data areas except the operand stack, have their own address space. The operand stack is a
 last-in-first-out (LIFO) stack and is not addressable. Every AVM instruction operates on its specific

### 2022-03-28 tex/A.tex
Remove Argon chapter
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Re-add Argon chapter
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Save the draft of Ch1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-\lstset{
-    language=Java,
-    morekeywords={contract, initialize, constructor, external, virtual, map, account, signature, overrides, is, string},
-    basicstyle=\small\sffamily,
-    columns=fullflexible,
-%   keywordstyle=\bfseries,
-    breaklines=true,
-    showstringspaces=false,
-    mathescape
-}
 %\binoppenalty=99999
@@
 
-    \chapter{The Argennon Virtual Machine}\label{ch:AVM}
+    \chapter{The Argennon Smart Contract Execution Environment}\label{ch:AVM}
     \input{sections/avm}

### 2022-03-28 tex/sections/avm.tex
Save the draft of Ch1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The Argennon\footnote{the classical pronunciation should be used:\textipa{/Ar"gen.non/}} Virtual Machine (AVM) is an
-abstract computing machine for executing Argennon smart contracts. The
-Argennon Virtual Machine knows nothing of the Argennon blockchain, transactions or any concept of gas usage, only of
-Argennon identifier tries and the concept of execution sessions.
+The Argennon\footnote{The classical pronunciation should be used:\textipa{/Ar"gen.non/}} Smart Contract Execution
+Environment (AscEE) is an abstract execution environment for executing Argennon smarts contracts (a.k.a\ applications).
+An Argennon application essentially is a http server whose state is kept in the Argennon blockchain and
+its logic is described using an Argennon Standard Representation (ASR).
 
-Execution sessions are separate sessions of executing smart contract's code by the Argennon Virtual Machine.
-Every execution session starts with a single method invocation, but does not necessarily end with
-that method invocation. During an execution session additional method invocations could be scheduled,
-which will be executed after the initial method invocation completes.
+An Argennon Standard Representation (ASR) is a programming language for describing argennon applications, optimized
+for the architecture and requirements of the Argennon platform.
+Argennon supports two standard representations, one is a high level text based language which needs
+costly compilation before being executed on a hardware machine. The other one is a low level binary representation which
+usually can be executed, with minimal pre-processing, by a JIT compiler or an emulator. The high level
+language is intended for preserving high level information of applications logic to facilitate
+platform specific compiler optimization at host nodes. On the other hand, the low level language enables efficient
+execution of applications that are not frequently used.
 
-An execution session may complete normally or abruptly. When an execution session completes abruptly, it will
-not have any effects on the AVM's state.
+The state of an Argennon application is stored in byte addressable finite arrays of memory called
+\emph{heap chunks}. An application can have several heap chunks with different sizes, and can remove or
+resize its heap chunks or allocate new chunks. Every chunk belongs to exactly one application and can only be modified
+by its owner. In addition to heap chunk, every application also has an amount of non-persistent local memory for
+storing temporary data.
 
-\section{Data Types}\label{sec:data-types}
+The Argennon Smart Contract Execution Environment can be seen as a machine for executing http requests, producing
+their http response and updating related heap
+chunks. The AscEE executes requests sequentially \footnote{Actually requests are executed in parallel but by
+performing data dependency analysis the result is guaranteed to be identical with sequential execution of
+requests.} and each request has a separate \emph{execution session}. Execution sessions are separate sessions of
+executing smart contract's code by the Argennon Smart Contract Execution Environment. The state of an execution
+session will be
+destroyed at the end of the session and only the state of heap chunks is preserved. If a session fails and does not
+complete normally, it will not have any effect on the state of heap chunks.
 
-The Argennon Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
-and does not have to be done by the Argennon Virtual Machine itself.
+During an execution session an application can make http requests to other applications. Those requests will not start
+a new execution session and will be executed within the current session. In AscEE making a http request is similar to
+a function invocation, because of that, we sometimes refer to it as an application
+call.
 
-The Argennon Virtual Machine operates on two kinds of types: primitive types and identifier types. There are,
-correspondingly, two kinds of values that can be stored in memory locations, passed as arguments, returned by
-methods, and operated upon: primitive values and identifier values.
-
-Values of primitive or identifier types need not be tagged or otherwise be inspectable to determine their types
-at run time, or to be distinguished from values of other types. Instead, the instruction set of the Argennon
-Virtual Machine distinguishes its operand types using instructions intended to operate on values of specific
-types. For instance, \texttt{iadd64} assumes that its operands are two 64-bit signed integers or \texttt{delete}
-assumes that its operand is an identifier type.
-
-An identifier value in the Argennon Virtual Machine is a variable length array of bytes. Some instructions
-that work with identifiers are able to determine the length of their identifier operands, while some
-other instructions, for performance reasons, require the length to be specified.
-
-The Argennon Virtual Machine does not have a fixed word size and uses \textbf{little-endian} byte order. Any
-instruction of the AVM has its specific word size, and the addressable memory areas are byte addressable.
+The AscEE is designed based on \emph{optional
+decoupling principle}. When an application makes a request to another application, optionally it can request to be
+decoupled from the called application. That would mean the called application could not affect its caller's state
+by reentrancy, or could not fail the session by using excessive resources or performing illegal operations.
 
@@
 
-In the Argennon Virtual Machine four distinct identifier types exist: \texttt{applicationID}, \texttt{accountID},
-\texttt{methodID} and \texttt{chunkID}.
+In Argennon a unique identifier is assigned to every application, heap chunk and account. Therefore, three distinct
+identifier types exist: \texttt{appID}, \texttt{accountID}, and \texttt{chunkID}.
+All these identifiers are \emph{prefix codes}, and hence can be represented by
+\emph{prefix trees}\footnote{Also called tries.}.
 
-All these identifiers are \emph{prefix codes}, and hence can be represented by
-\emph{prefix trees}\footnote{Also called tries.}. The Argennon has three primitive prefix trees:
-\emph{applications, accounts} and \emph{local}. Any identifier in Argennon is a
-prefix code built by one or more of these prefix trees:
+Argennon has three primitive prefix trees:
+\emph{applications, accounts} and \emph{local}.
+All these trees are in base 256, with the maximum height
+of 8. An Argennon identifier may be simple or compound. A simple identifier is generated using a single trie, while a
+compound identifier is generated by concatenating prefix codes generated by two or more tries:
+
 \begin{itemize}
-    \item \texttt{applicationID} is a prefix code built by \emph{applications} prefix tree.
-    \item \texttt{accountID} is a prefix code built by \emph{accounts} prefix tree. An \texttt{accountID} can
-    not be \texttt{0x0} or \texttt{0x1}.
-    \item \texttt{methodID} is a composite prefix code built by concatenating an \texttt{applicationID} to
-    a prefix code made by \emph{local} prefix tree: (\texttt{|} is the concatenating operator)
-    \subitem \texttt{methodID = (applicationID|<local-prefix-code>)} .
+    \item \texttt{appID} is a prefix code built by \emph{applications} prefix tree. An application ID cannot
+    be \texttt{0x0}.
+
+    \item \texttt{accountID} is a prefix code built by \emph{accounts} prefix tree. An account ID cannot
+    be \texttt{0x0} or \texttt{0x1}.
+
     \item \texttt{chunkID} is a composite prefix code built by concatenating an \texttt{applicationID} to
     an \texttt{accountID} to a prefix code made by \emph{local} prefix tree:
-    \subitem \texttt{chunkID = (applicationID|accountID)} or
-    \subitem \texttt{chunkID = (applicationID|0x0|<local-prefix-code>)} or
-    \subitem \texttt{chunkID = (applicationID|0x1|accountID|<local-prefix-code>)} .
+    \subitem \texttt{chunkID = (applicationID|accountID|<local-prefix-code>)} .
 \end{itemize}
@@
 \]
-where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)}<A^{(i+1)}\). A typical choice for \(\beta\)
-could be \(2^8\).
+where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)} \leq A^{(i+1)}\). A typical choice for
+\(\beta\) could be \(2^8\).
 
@@
 
+\section{Heap Chunks}\label{heap}
+
+The persistent data of an Argennon application is stored in heap chunks. An application may have several heap chunks
+with different sizes, and can remove or
+resize its chunks or allocate new chunks. Every chunk belongs to exactly one application. Only the owner application can
+modify a chunk but there is no restrictions for reading a chunk.
+
+When an application allocates a new heap chunk, the identifier of the new chunk is not generated by
+the AscEE. Instead, the application can choose an identifier itself \footnote{provided it has a correct format}. This
+is an important feature of the AscEE's heap, which allows applications to use the AscEE's heap as a dictionary (map)
+data structure.
+Since the \texttt{chunkID} is a prefix code, any application has its own identifier space, and an application
+can easily generate unique identifiers for its chunks.
+
+A heap chunk can be considered as a continuous
+array of bytes. Every chunk has a size: \texttt{chunkSize} and a size upper bound: \texttt{sizeUpperBound}. The value of
+\texttt{chunkSize} can be determined uniquely at
+the start of
+every execution session, and it may be updated during the session like a normal memory location. On the other hand,
+the value of \texttt{sizeUpperBound} is constant for every block of the blockchain and is proposed by the block
+proposer.
+
+The address space of a chunk starts from zero and only offsets lower than \texttt{sizeUpperBound} are valid. Trying to
+access any offset higher than this value will result in a revert for the application.
+
+\subsection{Access Blocks}\label{subsec:access-blocks}
+
+Memory locations inside a chunk can only be accessed through access blocks. An access block is defined on a chunk
+and has an offset and a size and determines accessible memory locations inside a chunk. Multiple access blocks can
+be defined on a chunk, but they must be non overlapping. Access blocks are byte addressable and can have different
+access types:
+
+\begin{itemize}
+    \item \texttt{read\_only}: only allows read and check operations.
+    \item \texttt{writable} memory locations inside this access block can not be modified.
+    \item \texttt{check\_only}: only allows check operations. These operations query the persistence
+    status of memory locations.
+    \item \texttt{additive}: only allows addition operations without overflow checking. Note that the content of these
+    access blocks cannot be read.
+\end{itemize}
+
+Chunks are intended for simplifying proof checking of the data stored in the Argennon cloud and access blocks are
+required for better parallelization of request execution. An application should put the data it predicts is needed for
+validating a block in the same chunk and put the data it predicts will be needed in a single execution session in
+the same access block.
+
+\subsection{Chunk Resizing}\label{subsec:ch-resize}
+
+The value of \texttt{chunkSize} at the end of the execution session will determine if a memory location at an
+offset is persistent or not: Offsets lower than the chunk size are persistent, and higher offsets are not.
+Non-persistent locations will be re-initialized with zero, at the start of every execution session.
+
+The value of chunkSize can be modified during an execution session. However, the new size values can only be
+increasing or decreasing. More precisely, if a request declares that it wants to expand (shrink) a chunk, it can only
+increase (decrease) the value of chunkSize and any specified value for chunkSize during the execution session, needs
+to be greater (smaller) than the previous value of the chunk size. Any request that wants to expand (shrink) a chunk
+needs to specify a max size (min size). The value of chunkSize can not be set higher (lower) than this value.
+
+Usually an application should not have any assumption about the content of memory locations that are outside the chunk.
+While these locations are zero initialized at the start of every execution session, it should be noted that multiple
+invocations of an application may occur in a single execution session, and if one of them modifies a location outside
+the chunk, the changes can be seen by next invocations.
+
+There is no way for an application to query the chunk capacity. As a result, in the view of an application, accessing
+offsets higher than chunkSize results in undefined behaviour, while the behaviour is well-defined for validators.
+This enables validators to determine the validity of an offset at the start of the block validation in a parallelized
+preprocessing phase without executing requests.
+
+While an application can use \texttt{chunkSize} to determine if an offset is persistent or not, that is not
+considered a good practice. Reading \texttt{chunkSize} decreases transaction parallelization, and should be avoided.
+Instead, applications should use a built-in AscEE's function for checking the persistence status of memory addresses.
+
+An application can load a chunk with a valid prefix identifier even if that chunk does not exist. For a non-existent
+chunk the value of \texttt{chunkSize} is always zero.
+
+\section{Resource Management}\label{sec:res-man}
+
+Completing an execution session needs computational resources. The amount of resources used by an execution session
+needs to be monitored, otherwise a spammer would be able to easily exhaust the resources of the execution
+environment. Resource usage can be measured per session or per application call.
+
+The AscEE has two type of execution sessions: optimistic and monitored. Resource usage of an optimistic session
+is always measured per session and default pre-defined resource caps are used. An optimistic session must always
+complete normally. On the other hand in a monitored session some resources are measured per application and some
+caps are determined dynamically by the request. A monitored session is allowed to either complete normally or fail.
+
+Obviously resource monitoring is easier for optimistic sessions, and they can be executed more efficiently. However, if
+an optimistic session violates its resource caps, determining the point of failure requires precise measurement of
+used resources. Note that for implementing optional decoupling principle we need to determine the exact application
+which fails in the call chain. For example assume that an optimistic session containing an application call violates a
+1ms execution time cap. If the call happens exactly after 1ms a small fluctuation in measuring the execution time can
+change the point of failure between the called and the main application. On the other hand this is not an issue for
+monitored session because resource measurements and caps are defined per application.
+
+The following resources are defined and monitored for AscEE execution sessions:
+
+execution time:
+The amount of cpu time that is required for executing a session or an application. The execution time of a session
+is measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
+of cpu time needed for executing a predefined standard application which is used for benchmarking a host's performance.
+
+local memory: any memory usage of an application that is not part of the heap chunks and is not part of another resource
+will be considered as local memory usage. This resource is measured in bytes.
+
+heap access list: every session can only access heap locations that are declared in its access list. In addition,
+resizing
+heap chunks can only be done in the range of the pre-declared lower bound and upper bound. those predeclared
+locations of heap chunks that are
+
+applications list: a session can only make requests to applications that are declared in its application list.
+
+call depth: the number of nested application calls.
+
+number of deferred calls
+number of session locks
+number of virtual signatures
+
 \section{Arithmetics}\label{sec:arithmetics}
@@
 
-\section{Architecture}\label{sec:arch}
-
-\subsection{The \texttt{pc} Register}\label{subsec:the-pc-register}
-
-The Argennon Virtual Machine always has a single thread of execution and exactly has one \texttt{pc} register.
-When the Argennon Virtual Machine is executing a method, if that method is not native, the \texttt{pc} register
-contains the address of the AVM instruction currently being executed. If the method currently being executed
-is native, the value of the Argennon Virtual Machine's \texttt{pc} register is undefined.
-
-\subsection{Call Stack Queue}\label{subsec:call-stack}
-
-Every AVM execution session has a queue of call stacks. A call stack contains all the information that is needed for
-restoring the AVM's state and continuing the execution after the method invocation completes.
-This information is represented by a \texttt{CallInfo} struct:
-\begin{verbatim}
-CallInfo {
-    methodID,
-    pc,
-    localFrame,
-    operandStack
-} .
-\end{verbatim}
-
-The \texttt{methodID} field is the unique identifier of a method's bytecode, which
-is a composite prefix code built by concatenating an \texttt{applicationID} to
-a local identifier.
-
-Every method invocation has a corresponding \texttt{CallInfo} struct and when a method is invoked its
-\texttt{CallInfo} struct is pushed onto the call stack. Hence, the top of the call stack always contains
-the \texttt{CallInfo} struct of the current method. When a method invocation completes, its call
-information is popped from the call stack.
-
-An AVM execution session will continue as long as there is a non-empty call stack in the call stack queue. When the
-execution of the current call stack finishes, the execution of the next call stack in the call stack queue starts.
-It is possible that an AVM method invocation creates new call stacks. These newly created call stacks will be a part
-of the current execution session and are added at the end of the call stack queue. A newly created call stack always
-contains a single \texttt{CallInfo} struct. See Section~\ref{subsec:method-invocation} for more details.
-
-\subsection{Run-Time Data Areas}\label{subsec:data-areas}
-
-The Argennon Virtual Machine defines various run-time data areas that are used during an execution session,
-Some of these data areas are persistent and are stored on the blockchain, Other data areas
-are per execution session or per method invocation. These data areas are created when their context starts and
-destroyed when their context ends.
-
-The Argennon Virtual Machine has five run-time data areas:
-\begin{itemize}
-    \item Method Area
-    \item Constant Area
-    \item Local Frame
-    \item Operand Stack
-    \item Heap
-\end{itemize}
-All data areas except the operand stack, have their own address space. The operand stack is a
-last-in-first-out (LIFO) stack and is not addressable. Every AVM instruction operates on its specific
-data areas.
-
-\subsubsection{Method Area}
-
-The method area contains the byte-code of any method which can be called by a non-native method invocation
-instruction. In the Argennon Virtual Machine every method has a unique identifier:
-\texttt{methodID}, and the method area is a map from method identifiers to their byte codes.
-
-Every smart contract is required to have a special \texttt{dispatcher} method. The \texttt{dispatcher} method is the
-only method of a smart contract that can be called by other smart contracts and has a
-predefined \texttt{methodID}: \texttt{(ApplicationID|0)}.
-
-\note{The Argennon Virtual Machine does not have internal or external methods. All methods are considered internal
-except the \texttt{dispatcher} method, which is the only external method of any smart contract.
-This design enables the AVM to execute \emph{RESTful} smart contracts more efficiently.}
-
-Instructions that modify the code area can only be run in privileged mode, which means they can only be executed
-by the \emph{root} smart contract. As a result, the access of smart contracts to the code area is essentially
-read-only.
-
-\subsubsection{Constant Area}
-
-Every smart contract has a single constant area which is stored in the AVM code area as a special type of
-non-executable method. The constant area of a smart contract contains user
-defined constants.
-
-\note{Instructions that modify the constant area can only be run in privileged mode.}
-
-\subsubsection{Local Frame}
-
-A local frame is used to store methods parameters and local variables. A new frame is created each time a method
-is invoked, and it is destroyed when its method invocation completes, whether the completion is normal or abrupt.
-
-\subsubsection{Operand Stack}
-
-Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too. The
-Argennon Virtual Machine is a stack machine and its instructions take operands from the operand stack, operate on
-them, and push the result back onto the operand stack. An operand stack is destroyed when its owner method
-completes, whether that completion is normal or abrupt.
-
-\subsubsection{Heap}
-
-The heap of the Argennon Virtual Machine is a persistent memory area which stores \emph{memory chunks}. A Memory
-chunk is a byte addressable piece of continuous memory which has a separate address space starting from 0. Each chunk
-has a fixed size and different chunks need not be equally sized. Chunks are stored in the heap area and every chunk is
-assigned a unique identifier: \texttt{chunkID}. As a result, the address of every memory location inside
-the heap area can be considered as a pair: \texttt{(chunkID, offset)}.
-
-A smart contract can read any chunk stored in the heap by having its \texttt{chunkID}. However, it can only modify
-those chunks that was created by itself. In other words, every memory chunk in the AVM's heap has an owner. Only
-the owner can modify a chunk while anyone can read that chunk. The owner of a chunk can be easily determined
-by the \texttt{applicationID} part of the chunk identifier.
-
-\note{The reason behind this type of access control design is the fact that smart contract code is usually
-immutable. That means if a smart contract does not implement a getter mechanism for some parts of its internal
-data, this functionality can never be added later, and despite the internal data is publicly available, there will
-be no way for other smart contracts to use this data on-chain. This design eliminates the need for implementing
-\emph{trivial} getters.}
-
-The AVM Heap is able to save snapshots of its state and later restore them. This will enable the Argennon
-Virtual Machine to have state reversion capability. The snapshot management of the heap area is done by
-the following functions:
-\begin{itemize}
-    \item \texttt{Save()} saves a snapshot of the current state of the heap and pushes it onto the
-    snapshot stack.
-    \item \texttt{Restore()} pops the snapshot stored on the top of the snapshot stack and restores it. The
-    current state of the heap will be lost.
-    \item \texttt{Discard()} pops the snapshot stored on the top of the snapshot stack and discards it. The current
-    state of the heap will not change.
-\end{itemize}
-
-\note{These functions are internal functions of the AVM. They are not instructions, and can not be called by smart
-contracts.}
-
-\section{Instruction Set Overview}\label{sec:instruction-set-overview}
-
-An Argennon Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
-performed, followed by zero or more operands supplying arguments or data that are used by the operation.
-The number and size of the operands are determined solely by the opcode.
-
-\subsection{Method Invocation}\label{subsec:method-invocation}
-
-The Argennon Virtual Machine has four types of method invocation instruction:
-\begin{itemize}
-    \item \texttt{invoke\_internal}: invokes a method without changing the context of method execution. In other
-    words, the invoked method can only modify the state of the current smart contract.
-    A smart contract can invoke any method by \texttt{invoke\_internal} instruction, even if that method is an
-    internal method of another smart contract. Since the invoked method will always be executed in the context of
-    the invoking smart contract, a smart contract will not be able to modify another smart contract's state by
-    this instruction. This instruction facilitates code reuse and the usage of libraries.
-    If the invoked method completes abruptly, the invoker method completes abruptly too.
-    \item \texttt{invoke\_dispatcher}: invokes the \texttt{dispatcher} method of a smart contract and changes the
-    context of method execution to that
-    smart contract. If the invoked method completes abruptly, that will cause the invoker method to complete
-    abruptly too.
-    \item \texttt{i\_invoke\_dispatcher, i\_invoke\_internal}: independently invokes a method. These instructions
-    are similar to their \texttt{invoke} counterparts, with the difference that when
-    the invoked method completes abruptly, the invoker method can continue execution. When a method invoked by
-    an \texttt{i\_invoke} instruction completes abruptly, all changes made
-    by the invoked method to the AVM's state will be reverted and the instruction will act like
-    a jump to the specified error handling code.
-    \item \texttt{invoke\_native}: invokes a method that is not hosted by the Argennon Virtual Machine. By this
-    instruction, high performance native methods of the hosting machine, such as GPU accelerated operations,
-    could become available to AVM smart contracts. This instruction will not modify the AVM's state.
-    \item \texttt{invoke\_later}: invokes the \texttt{dispatcher} method of another smart contract by creating a new
-    call stack. The created call stack is added at the end of the call stack queue, and
-    the context of method execution will change after the execution of that call stack starts. This instruction does
-    not modify the \texttt{pc} register and the execution of the
-    invoker will continue after this instruction. As a result, \texttt{invoke\_later} can not return
-    any values to the caller and the return value of the called method (if any) will be discarded. If the invoked
-    method completes abruptly, it will cause \textbf{all the call stack queue to
-    complete abruptly.}
-\end{itemize}
-
-\note{The \texttt{invoke\_later} instruction can be used to call a smart contract while its reentracy lock is active.
-Because the invocation uses a separate call stack, it will not cause the lock to throw an exception.}
-
-Each time a method is invoked a new local frame
-and operand stack is created. If the method was invoked by an \texttt{i\_invoke} instruction,
-the \texttt{Save()} function of the AVM's heap is called before the execution of the method starts.
-
-The Argennon Virtual Machine uses local frames to pass parameters on
-method invocation. On method invocation, any parameters are passed in consecutive local variables stored in the
-method's local frame starting from address 0. The invoker of a method writes the parameters in the local frame
-of the invoked method using \texttt{arg} instructions.
-
-\note{The AVM does not natively support polymorphism and virtual methods. A compiler could easily
-generate appropriate code for implementing these features.}
-
-\subsection{Exceptions}\label{subsec:exceptions}
-
-An exception is thrown programmatically using the \texttt{athrow} instruction. Exceptions can also be thrown by
-various Argennon Virtual Machine instructions if they detect an abnormal condition.
-
-\note{Any AVM instruction may throw an exception. This happens when the execution cost of a transaction has exceeded
-the declared maximum execution cost.}
-
-\subsection{Method Invocation Completion}\label{subsec:method-invocation-completion}
-
-A method invocation completes normally if that invocation does not cause an exception to be thrown, either
-directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
-method completes normally and the invocation was not made by an \texttt{invoke\_later} instruction, then a value may be
-returned to the invoking method. This occurs when the invoked method executes one of the return instructions, the
-choice of which must be appropriate for the type of the value being returned (if any). Execution then continues
-normally in the invoking method's local frame with the returned
-value (if any) pushed onto the operand stack. If the method was invoked by an \texttt{invoke\_later} instruction,
-the execution of the next call stack (if any) begins and the returned value
-is discarded. When a method invocation that was made by
-an \texttt{i\_invoke} instruction completes normally,
-the \texttt{Heap.Discard()} function is called as a part of the return instruction.
-
-A method invocation completes abruptly when an exception is thrown by one its instructions.
-The \texttt{invoke\_internal} and \texttt{invoke\_dispatcher} instructions
-always throw an exception when their invoked method completes abruptly. A
-method invocation that completes abruptly never returns a value.
-
-When a method completes normally, the call stack is used to restore the state of the invoker,
-including its local frame and operand stack, with the \texttt{pc} register appropriately restored and incremented
-to skip past the method invocation instruction. If the current call stack is empty the next call stack in the call
-stack queue will be used for continuing the execution, and when there are no call stacks left in the queue the
-current execution session ends.
-
-When a method completes abruptly, the call stack is used to restore the \texttt{pc} register of the invoker. If the pc
-register points to an \texttt{invoke} instruction, the current method completes abruptly as well. Otherwise, if
-the \texttt{pc} register is pointing to an \texttt{i\_invoke} instruction, the \texttt{Heap.Restore()} function is
-called to restore the state of the AVM's heap. Then, the
-local frame and operand stack of the invoker is restored, and execution continues after branching to the error handling
-code specified by the instruction's operand.
-
-When an invocation which was made by an \texttt{invoke\_later} instruction completes abruptly, the
-heap snapshot that was saved before the start of the
-execution of the first call stack is restored. This means
-that all changes made by the execution session to the AVM's heap will be completely reverted.
-
-\subsection{Reentrancy Protection}\label{subsec:reentrancy}
+\section{Reentrancy Protection}\label{sec:reentrancy}
 
@@
 
-\subsection{Heap Allocation and Deallocation}\label{subsec:heap-allocation-instructions}
-
-Heap allocation and deallocation in the Argennon Virtual Machine is done by the following instructions:
-
-\begin{itemize}
-    \item \texttt{alloc}: creates a new memory chunk of the specified size in the heap. The
-    new chunk's identifier will be the concatenation of the \texttt{applicationID} of the creator of the chunk and
-    the \texttt{id} operand of the instruction. The \texttt{id} operand must be the concatenation of
-    an \texttt{accountID} and a prefix code
-    generated by the \emph{local} prefix tree: \texttt{id = accountID|localID}. If the identifier is not valid or
-    if it already exists in the heap, a catchable exception will be thrown.
-
-    \item \texttt{delete}: deletes the chunk that its identifier is the concatenation of the current smart
-    contract \texttt{applicationID} and the specified \texttt{id} operand. If the identifier is not valid or
-    if the chunk was not found, a catchable exception will be thrown.
-\end{itemize}
-
-When a smart contract allocates a new memory chunk, the identifier of the new chunk is not generated by
-the AVM. Instead, the smart contract can choose an identifier itself. This is a very important feature of
-the AVM's heap, which allows smart contracts to use the AVM's heap as a dictionary (map) data structure.
-Since the \texttt{chunkID} is a prefix code, any smart contract has its own identifier space, and a smart contract
-can easily generate unique identifiers for its chunks.
-
-The AVM requires the chosen identifier to be a prefix code compatible with
-the \emph{accounts} prefix tree, or a code made by concatenating a code compatible with
-the \emph{accounts} prefix tree to a code compatible with the \emph{local} prefix tree. This requirement
-enables AVM to detect invalid identifiers, while smart contracts can easily associate data with account identifiers.
-At the same time, A smart contract is able to create maps with costume keys by appending local identifiers
-to the zero \texttt{accountID}. Zero \texttt{accountID} does not correspond to a valid account in the Argennon
-blockchain.
+\section{Deferred Calls}\label{sec:deferred-calls}
 
@@
 
-\section{The AVM Standard Library}\label{sec:the-avm-standard-library}
+\section{The Argennon Standard Library}\label{sec:asl}
 

### 2022-03-28 tex/sections/avm.tex
Save the draft of Ch1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The AscEE has two type of execution sessions: optimistic and monitored. Resource usage of an optimistic session
-is always measured per session and default pre-defined resource caps are used. An optimistic session must always
-complete normally. On the other hand in a monitored session some resources are measured per application and some
-caps are determined dynamically by the request. A monitored session is allowed to either complete normally or fail.
+The AscEE has two type of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
+session is always measured per session and default pre-defined resource caps are used. On the other hand in a monitored
+session some resources are measured per application and some
+caps are determined dynamically by the request. An optimistic session must always
+complete normally but a monitored session is allowed to either complete normally or abruptly.
 
 Obviously resource monitoring is easier for optimistic sessions, and they can be executed more efficiently. However, if
-an optimistic session violates its resource caps, determining the point of failure requires precise measurement of
-used resources. Note that for implementing optional decoupling principle we need to determine the exact application
-which fails in the call chain. For example assume that an optimistic session containing an application call violates a
-1ms execution time cap. If the call happens exactly after 1ms a small fluctuation in measuring the execution time can
-change the point of failure between the called and the main application. On the other hand this is not an issue for
-monitored session because resource measurements and caps are defined per application.
+an optimistic session violates its resource caps, determining the point of failure requires precise resource
+measurement. Note that for implementing optional decoupling principle, we need to determine the exact application
+which has failed in the call chain. For example assume that an optimistic session containing an application call
+violates a 2 milliseconds execution time cap. If in the caller application the call happens exactly after 2
+milliseconds of execution, a small fluctuation in the execution time measurement can
+change the point of failure between the called and the main application. This will introduce nondeterministic
+behaviour which makes block validation impossible. On the other hand, for a monitored session this is not an issue
+because resource measurements and caps are defined per application, and we make sure that resource caps are not too
+small for any application call.
 
-The following resources are defined and monitored for AscEE execution sessions:
+Different computational resources are measured and monitored during an AscEE session:
+\begin{itemize}
+    \item \textbf{execution time}:
+    the amount of cpu time that is required for executing a session or an application. The execution time of a session
+    is measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
+    of cpu time needed for executing a predefined standard application which is used for benchmarking a host's
+    performance.
 
-execution time:
-The amount of cpu time that is required for executing a session or an application. The execution time of a session
-is measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
-of cpu time needed for executing a predefined standard application which is used for benchmarking a host's performance.
+    Optimistic sessions have a predefined \texttt{maxClocks} value which is determined by the Argennon protocol. This
+    value defines a bound on the total cpu time of the session and no per application measurement is done.
 
-local memory: any memory usage of an application that is not part of the heap chunks and is not part of another resource
+    Monitored sessions perform per application cpu-time measurement, and every application call during a monitored
+    session has a separate \texttt{maxClocks} value. This value determines the maximum amount of time that the cpu
+    can be used for executing that particular application call. It should be noted that the timer is paused when an
+    application makes a call to another application, until the control returns to the caller. An application call
+    needs at least 100 clocks and if the value of its \texttt{maxClock} is lower than this value the call will be
+    considered a failed call.
+
+    Each application call has some amount of \texttt{externalClocks}. When an application makes a request to another
+    application it has to \emph{forward} a portion of its external clocks to the called application. This amount
+    will determine the value of maxClocks for the called application.
+
+    The amount of external clocks for every application
+    call is defined to be \(2/3\) of its \texttt{maxClocks}. As a result the total number of clocks of a monitored
+    session is always less than \(3 \times \texttt{maxClocks}\) of the root application. The value of
+    \texttt{maxClocks} for
+    the root application call is determined by the transaction.
+
+    \item \textbf{local memory}:
+    any memory usage of an application that is not part of the heap chunks and is not part of another resource
     will be considered as local memory usage. This resource is measured in bytes.
+    \item \textbf{heap access list}:
+    every session can only access heap locations that are declared in its access list. In addition,
+    resizing heap chunks can only be done in the range of the pre-declared lower bound and upper bound.
+    \item \textbf{applications list}:
+    a session may only make requests to applications that are declared in its application list.
+    \item \textbf{call depth}:
+    the number of nested application calls can not be more than a threshold. This threshold is determined by the
+    Argennon protocol.
+    \item active differed calls:
+    \item virtual signatures:
+    \item number of locks
 
-heap access list: every session can only access heap locations that are declared in its access list. In addition,
-resizing
-heap chunks can only be done in the range of the pre-declared lower bound and upper bound. those predeclared
-locations of heap chunks that are
-
-applications list: a session can only make requests to applications that are declared in its application list.
-
-call depth: the number of nested application calls.
-
-number of deferred calls
-number of session locks
-number of virtual signatures
+\end{itemize}
 

### 2022-03-29 tex/sections/avm.tex
Save the draft of Ch1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 for the architecture and requirements of the Argennon platform.
-Argennon supports two standard representations, one is a high level text based language which needs
+Argennon supports two standard representations: one is a high level text based language which needs
 costly compilation before being executed on a hardware machine. The other one is a low level binary representation which
@@
 destroyed at the end of the session and only the state of heap chunks is preserved. If a session fails and does not
-complete normally, it will not have any effect on the state of heap chunks.
+complete normally, it will not have any effect on the heap chunks.
 
 During an execution session an application can make http requests to other applications. Those requests will not start
-a new execution session and will be executed within the current session. In AscEE making a http request is similar to
-a function invocation, because of that, we sometimes refer to it as an application
-call.
+a new execution session and will be executed within the current session. In AscEE making a http request to an
+application is similar to a function invocation therefore we also refer to it as an application call.
 
@@
 
-Argennon has three primitive prefix trees:
-\emph{applications, accounts} and \emph{local}.
+Argennon has four primitive prefix trees:
+\emph{applications, accounts, local} and \emph{varUint}.
 All these trees are in base 256, with the maximum height
@@
     \item \texttt{read\_only}: only allows read and check operations.
-    \item \texttt{writable} memory locations inside this access block can not be modified.
+    \item \texttt{writable}: memory locations inside this access block can not be modified.
     \item \texttt{check\_only}: only allows check operations. These operations query the persistence
@@
 Chunks are intended for simplifying proof checking of the data stored in the Argennon cloud and access blocks are
-required for better parallelization of request execution. An application should put the data it predicts is needed for
-validating a block in the same chunk and put the data it predicts will be needed in a single execution session in
-the same access block.
+required for better parallelization of the request execution. An application should put the data it predicts is
+needed for validating a block in the same chunk and the data it predicts is needed in a single execution
+session in the same access block.
 
@@
 Completing an execution session needs computational resources. The amount of resources used by an execution session
-needs to be monitored, otherwise a spammer would be able to easily exhaust the resources of the execution
+should be monitored, otherwise a malicious user would be able to easily spam and exhaust resources of the execution
 environment. Resource usage can be measured per session or per application call.
@@
 session some resources are measured per application and some
-caps are determined dynamically by the request. An optimistic session must always
-complete normally but a monitored session is allowed to either complete normally or abruptly.
+caps are determined dynamically by the transaction. An optimistic session must always
+complete normally, a monitored session is allowed to either complete normally or abruptly.
 
-Obviously resource monitoring is easier for optimistic sessions, and they can be executed more efficiently. However, if
+Obviously resource monitoring is easier and more efficient for optimistic sessions. However, if
 an optimistic session violates its resource caps, determining the point of failure requires precise resource
@@
 milliseconds of execution, a small fluctuation in the execution time measurement can
-change the point of failure between the called and the main application. This will introduce nondeterministic
+change the point of failure between the called and the main application. This can introduce nondeterministic
 behaviour which makes block validation impossible. On the other hand, for a monitored session this is not an issue
 because resource measurements and caps are defined per application, and we make sure that resource caps are not too
-small for any application call.
+small and can be measured easily.
 
@@
     Optimistic sessions have a predefined \texttt{maxClocks} value which is determined by the Argennon protocol. This
-    value defines a bound on the total cpu time of the session and no per application measurement is done.
+    value defines a bound on the \textbf{total} cpu time of the session and no per application measurement is done.
 
@@
     session has a separate \texttt{maxClocks} value. This value determines the maximum amount of time that the cpu
-    can be used for executing that particular application call. It should be noted that the timer is paused when an
+    can be used for executing that particular application call. It should be noted that this cpu timer is paused when an
     application makes a call to another application, until the control returns to the caller. An application call
@@
 
-    The amount of external clocks for every application
-    call is defined to be \(2/3\) of its \texttt{maxClocks}. As a result the total number of clocks of a monitored
+    The amount of external clocks of an application
+    call is defined to be \(2/3\) of its \texttt{maxClocks}. As a result, the total number of clocks of a monitored
     session is always less than \(3 \times \texttt{maxClocks}\) of the root application. The value of
@@
     \item \textbf{local memory}:
-    any memory usage of an application that is not part of the heap chunks and is not part of another resource
-    will be considered as local memory usage. This resource is measured in bytes.
+    any memory usage of an application that is not part of a heap chunk and is not part of another resource
+    will be considered as local memory usage. Local memory is not persistent and when an application finishes serving a
+    request and returns the http response (i.e.\ the application call completes) its local memory is deleted.
+    This resource is measured in bytes.
+
+    Optimistic sessions measure local memory usage per session and enforce a protocol-defined cap on the total amount of
+    used local memory in the session. Monitored sessions measure local memory usage per
+    application call and enforce a protocol defined-cap per application call. An application call which tries to use
+    more local memory than the cap will fail.
     \item \textbf{heap access list}:
@@
     \item \textbf{call depth}:
-    the number of nested application calls can not be more than a threshold. This threshold is determined by the
-    Argennon protocol.
+    during a session the number of nested application calls can not be more than a threshold. This threshold is
+    determined by the Argennon protocol.
     \item active differed calls:
@@
 \end{itemize}
+
+The Argennon protocol allows AscEE's implementations to have at max \%50 error in measuring execution time and at
+max \%25 error in measuring local memory usage. Other resources must be measured precisely.
+
+\section{Request Attachments}\label{sec:attachments}
+
+Attachments of a request is a list of request identifiers of the current block that are "attached" to the request.
+That means, for validating that request a validator first needs to "inject" the digest of attached requests into the
+httpRequest field of the request.
+
+The main usage of this feature is for fee payment. A request that wants to pay the fees for a number of requests
+declares those requests as its attachments. For paying fees the payer signs the digest of requests for which he
+wants to pay fees. After injecting the digest of those request by validators, that signature can be validated
+correctly and securely.
 

### 2022-03-30 tex/sections/avm.tex
Update Chapter 1
This is the first draft.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 Argennon supports two standard representations: one is a high level text based language which needs
-costly compilation before being executed on a hardware machine. The other one is a low level binary representation which
+costly compilation before being executed on a hardware machine. The other is a low level binary representation which
 usually can be executed, with minimal pre-processing, by a JIT compiler or an emulator. The high level
 language is intended for preserving high level information of applications logic to facilitate
-platform specific compiler optimization at host nodes. On the other hand, the low level language enables efficient
-execution of applications that are not frequently used.
+platform specific compiler optimization at host nodes. On the other hand, the low level language is designed for
+efficient execution of applications that are not frequently used.
 
@@
 resize its heap chunks or allocate new chunks. Every chunk belongs to exactly one application and can only be modified
-by its owner. In addition to heap chunk, every application also has an amount of non-persistent local memory for
+by its owner. In addition to heap chunks, every application also has an amount of non-persistent local memory for
 storing temporary data.
 
+\section{Execution Sessions}\label{sec:sessions}
+
 The Argennon Smart Contract Execution Environment can be seen as a machine for executing http requests, producing
-their http response and updating related heap
-chunks. The AscEE executes requests sequentially \footnote{Actually requests are executed in parallel but by
-performing data dependency analysis the result is guaranteed to be identical with sequential execution of
-requests.} and each request has a separate \emph{execution session}. Execution sessions are separate sessions of
-executing smart contract's code by the Argennon Smart Contract Execution Environment. The state of an execution
-session will be
+their http response and updating related heap chunks. The AscEE executes requests sequentially \footnote{Actually
+requests are executed in parallel but by performing data dependency analysis the result is guaranteed to be identical
+with sequential execution of requests.} and each request has a separate \emph{execution session}. Execution sessions
+are separate sessions of executing smart contract's code in order to fulfill \emph{external} http requests. An
+external request is a request that is not made by another Argennon application.
+
+The state and data of an execution session will be
 destroyed at the end of the session and only the state of heap chunks is preserved. If a session fails and does not
-complete normally, it will not have any effect on the heap chunks.
+complete normally, it will not have any effect on any heap chunks.
 
-During an execution session an application can make http requests to other applications. Those requests will not start
-a new execution session and will be executed within the current session. In AscEE making a http request to an
-application is similar to a function invocation therefore we also refer to it as an application call.
+During an execution session an application can make \emph{internal} http requests to other applications. Those
+requests will not start a new execution session and will be executed within the current session. In AscEE making a
+http request to an application is similar to a function invocation. For that reason, we also refer to internal requests
+as application calls.
 
-The AscEE is designed based on \emph{optional
-decoupling principle}. When an application makes a request to another application, optionally it can request to be
-decoupled from the called application. That would mean the called application could not affect its caller's state
-by reentrancy, or could not fail the session by using excessive resources or performing illegal operations.
+The AscEE is designed based on \emph{optional decoupling principle}. When an application makes a request to another
+application, optionally it can request to be decoupled from the called application. That would mean the called
+application could not affect its caller's state by reentrancy, or could not fail the session by using excessive
+resources or performing illegal operations.
 
@@
 All these trees are in base 256, with the maximum height
-of 8. An Argennon identifier may be simple or compound. A simple identifier is generated using a single trie, while a
+of 8.
+
+An Argennon identifier may be simple or compound. A simple identifier is generated using a single trie, while a
 compound identifier is generated by concatenating prefix codes generated by two or more tries:
@@
 
-In Argennon the shorter prefix codes are assigned to more active accounts and smart contracts which tend to own more
+In Argennon the shorter prefix codes are assigned to more active accounts and applications which tend to own more
 data objects in the system. The prefix trees are designed by analyzing empirical data to make sure the number
@@
 Since the \texttt{chunkID} is a prefix code, any application has its own identifier space, and an application
-can easily generate unique identifiers for its chunks.
+can easily find unique identifiers for its chunks.
 
@@
 the value of \texttt{sizeUpperBound} is constant for every block of the blockchain and is proposed by the block
-proposer.
+proposer. \texttt{sizeUpperBound} indicates the upper bound of the value of \texttt{chunkSize} for a chunk in a block.
+A validator can use this value to safely allocate memory for a chunk.
+
+The value of \texttt{chunkSize} at the end of the execution session will determine if a memory location at an
+offset is persistent or not: Offsets lower than the chunk size are persistent, and higher offsets are not.
+Non-persistent locations will be re-initialized with zero, at the start of every execution session.
 
@@
 and has an offset and a size and determines accessible memory locations inside a chunk. Multiple access blocks can
-be defined on a chunk, but they must be non overlapping. Access blocks are byte addressable and can have different
+be defined on a single chunk, but they must be non overlapping. Access blocks are byte addressable and can have
+different
 access types:
@@
     \item \texttt{read\_only}: only allows read and check operations.
-    \item \texttt{writable}: memory locations inside this access block can not be modified.
+    \item \texttt{writable}: allows reading and writing.
     \item \texttt{check\_only}: only allows check operations. These operations query the persistence
-    status of memory locations.
+    status of a memory location.
     \item \texttt{additive}: only allows addition operations without overflow checking. Note that the content of these
@@
 
-The value of \texttt{chunkSize} at the end of the execution session will determine if a memory location at an
-offset is persistent or not: Offsets lower than the chunk size are persistent, and higher offsets are not.
-Non-persistent locations will be re-initialized with zero, at the start of every execution session.
-
-The value of chunkSize can be modified during an execution session. However, the new size values can only be
+The value of \texttt{chunkSize} can be modified during an execution session. However, the new size values can only be
 increasing or decreasing. More precisely, if a request declares that it wants to expand (shrink) a chunk, it can only
-increase (decrease) the value of chunkSize and any specified value for chunkSize during the execution session, needs
-to be greater (smaller) than the previous value of the chunk size. Any request that wants to expand (shrink) a chunk
-needs to specify a max size (min size). The value of chunkSize can not be set higher (lower) than this value.
+increase (decrease) the value of \texttt{chunkSize} and any specified value during the execution
+session, needs to be greater (smaller) than the previous value of the chunk's size. Any request that wants to expand
+(shrink) a chunk needs to specify a max size (min size). The value of \texttt{chunkSize} can not be set higher
+(lower) than this value.
 
@@
 
-There is no way for an application to query the chunk capacity. As a result, in the view of an application, accessing
-offsets higher than chunkSize results in undefined behaviour, while the behaviour is well-defined for validators.
+There is no way for an application to query \texttt{sizeUpperBound} of a chunk. As a result, for an application,
+accessing offsets higher than \texttt{chunkSize} results in undefined behaviour, while the behaviour is well-defined
+in the view of validators.
 This enables validators to determine the validity of an offset at the start of the block validation in a parallelized
-preprocessing phase without executing requests.
+preprocessing phase without actually executing requests.
 
@@
 
+\section{Request Attachments}\label{sec:attachments}
+
+Attachment of a request is a list of request identifiers of the current block that are "attached" to the request.
+That means, for validating that request a validator first needs to "inject" the digest of attached requests into the
+http request.
+
+The main usage of this feature is for fee payment. A request that wants to pay the fees for a number of requests
+declares those requests as its attachments. For paying fees the payer signs the digest of requests for which he
+wants to pay fees. After injecting the digest of those request by validators, that signature can be validated
+correctly and securely.
+
+\section{Authorizing Operations}\label{sec:authorizing-operations}
+
+In blockchain applications, we usually need to authorize certain operations. For example, for sending an asset
+from a user to another user, first we need to make sure that the sender has authorized this operation.
+
+The AscEE uses \emph{authenticated message passing} for authorizing operations. In this method every execution
+session has a set of authenticated messages. Applications explicitly pass those messages for authorizing operations.
+Authentication of messages can be done using any type of cryptographic signature scheme in the preprocessing phase of
+request execution. For example, we could use BLS aggregate signature to authenticate all messages of a block in bulk.
+
+Moreover, application use built-in functions of AscEE to generate authenticated messages in run-time. This enables an
+application to authorize an operation even for an application that is not directly called.
+
+\note{Authorization by explicit messages and signatures eliminates the need for approval methods or call back patterns.}
+
+In addition to authenticated messages the AscEE provides a set of
+cryptographic functions for validating signatures and cryptographic entities. By using these functionalities and
+passing cryptographic signatures as parameters to methods, a programmer, having users' public keys, can implement
+the required logic for authorizing operations.
+
+\note{The Argennon Execution Environment has no instructions for issuing cryptographic signatures.}
+
+\section{Reentrancy Protection}\label{sec:reentrancy}
+
+The Argennon Smart Contract Execution Environment provides optional low level reentrancy protection by providing low
+level \emph{entrance locks}. When an application acquires an entrance lock it cannot acquire that lock again and trying
+to do so will result in a revert. The entrance lock of an application will be released when the application explicitly
+releases its entrance lock or when the call that had acquired the lock completes.
+
+The AscEE reentrancy protection mechanism is optional. An application can allow reentrancy, it can protect only certain
+areas of its code, or can completely disallow reentrancy.
+
+\section{Deferred Calls}\label{sec:deferred-calls}
+
+\ldots
+
 \section{Resource Management}\label{sec:res-man}
 
-Completing an execution session needs computational resources. The amount of resources used by an execution session
+Completing an execution session requires computational resources. The amount of resources used by an execution session
 should be monitored, otherwise a malicious user would be able to easily spam and exhaust resources of the execution
@@
 session is always measured per session and default pre-defined resource caps are used. On the other hand in a monitored
-session some resources are measured per application and some
-caps are determined dynamically by the transaction. An optimistic session must always
+session, some resources are measured per application call and some
+caps are determined dynamically by the external http request (i.e.\ transaction). An optimistic session must always
 complete normally, a monitored session is allowed to either complete normally or abruptly.
@@
 measurement. Note that for implementing optional decoupling principle, we need to determine the exact application
-which has failed in the call chain. For example assume that an optimistic session containing an application call
+which has failed in a call chain. For example assume that an optimistic session containing an application call
 violates a 2 milliseconds execution time cap. If in the caller application the call happens exactly after 2
@@
 change the point of failure between the called and the main application. This can introduce nondeterministic
-behaviour which makes block validation impossible. On the other hand, for a monitored session this is not an issue
-because resource measurements and caps are defined per application, and we make sure that resource caps are not too
+behaviour which could make block validation impossible. On the other hand, for a monitored session this is not an issue
+because resource measurements and caps are defined per application, and we can make sure that resource caps are not too
 small and can be measured easily.
@@
     \item \textbf{execution time}:
-    the amount of cpu time that is required for executing a session or an application. The execution time of a session
-    is measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
+    the amount of cpu time that is required for executing a session or an application call. The execution time of a
+    session is measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
     of cpu time needed for executing a predefined standard application which is used for benchmarking a host's
@@
 
-    Monitored sessions perform per application cpu-time measurement, and every application call during a monitored
+    Monitored sessions perform per application call cpu-time measurement, and every application call during a monitored
     session has a separate \texttt{maxClocks} value. This value determines the maximum amount of time that the cpu
-    can be used for executing that particular application call. It should be noted that this cpu timer is paused when an
-    application makes a call to another application, until the control returns to the caller. An application call
+    can be used for executing that particular application call. It should be noted that the cpu timer is paused when
+    the application makes a call to another application, and is resumed when the control returns. An application call
     needs at least 100 clocks and if the value of its \texttt{maxClock} is lower than this value the call will be
@@
     application it has to \emph{forward} a portion of its external clocks to the called application. This amount
-    will determine the value of maxClocks for the called application.
+    will determine the value of \texttt{maxClocks} for the called application.
 
@@
     call is defined to be \(2/3\) of its \texttt{maxClocks}. As a result, the total number of clocks of a monitored
-    session is always less than \(3 \times \texttt{maxClocks}\) of the root application. The value of
+    session is always less than \(3 \times \texttt{maxClocks}\) of the root application call. The value of
     \texttt{maxClocks} for
-    the root application call is determined by the transaction.
+    the root application call is determined by the external request (transaction).
 
@@
     determined by the Argennon protocol.
-    \item active differed calls:
-    \item virtual signatures:
-    \item number of locks
-
+    \item \textbf{active differed calls}:
+    \item \textbf{virtual signatures}:
+    \item \textbf{number of entrance locks}:
 \end{itemize}
@@
 
-\section{Request Attachments}\label{sec:attachments}
+\section{The ArgC Language}\label{sec:the-argc-language}
 
-Attachments of a request is a list of request identifiers of the current block that are "attached" to the request.
-That means, for validating that request a validator first needs to "inject" the digest of attached requests into the
-httpRequest field of the request.
+\subsection{The ArgC Standard Library}\label{sec:asl}
 
-The main usage of this feature is for fee payment. A request that wants to pay the fees for a number of requests
-declares those requests as its attachments. For paying fees the payer signs the digest of requests for which he
-wants to pay fees. After injecting the digest of those request by validators, that signature can be validated
-correctly and securely.
-
-\section{Arithmetics}\label{sec:arithmetics}
-
-The Argennon Virtual Machine supports signed integer and signed floating point operations. The Argennon Virtual
-Machine does not support any type of unsigned arithmetics. All arithmetic operations in the Argennon Virtual Machine
-are checked and any type of overflow or underflow will cause a catchable exception to be thrown.
-
-\section{Reentrancy Protection}\label{sec:reentrancy}
-
-The Argennon Virtual Machine provides low level reentrancy protection by defining a dedicated
-instruction: \texttt{enter}. When a smart contract executes the \texttt{enter} instruction, it
-can not execute that instruction again as long
-as the method executing the instruction completes (abruptly or normally). If a smart contract tries to execute
-the \texttt{enter} instruction while the method executing that instruction for the first time has not yet completed,
-an exception will be thrown.
-
-Smart contracts can completely prevent reentrancy by executing the \texttt{enter} instruction as the first instruction
-of their \texttt{dispatcher} method.
-
-\section{Deferred Calls}\label{sec:deferred-calls}
-
-\section{Authorizing Operations}\label{sec:authorizing-operations}
-
-In blockchain applications, we usually need to authorize certain operations. For example, for sending an asset
-from a user to another user, first we need to make sure that the sender has authorized this operation. The
-Argennon Virtual Machine has no built-in mechanism for authorizing operations, but it provides a rich set of
-cryptographic instructions for validating signatures and cryptographic entities. By using these instructions and
-passing cryptographic signatures as parameters to methods, a programmer, having users' public keys, can implement
-the required logic for authorizing any operation.
-
-\note{Authorization by explicit signatures eliminates the need for approval methods or call back patterns.}
-
-In addition to cryptographic signatures, the Argennon Virtual Machine provides instructions that enable
-smart contracts to issue \emph{virtual signatures}. A smart contract can virtually sign a message for a target
-smart contract by \texttt{v\_sign} instruction. When a smart contract
-virtually signs a message for a target, that message is stored in the internal memory of the AVM, associated with
-its signer and the target. The application that is the target of the signature can verify the message
-by \texttt{v\_verify} instruction. After verifying the signature, it will be removed from the AVM's internal memory
-to make sure a signature will not be used multiple times. The virtual signatures stored in the AVM's internal memory
-last for one execution session.
-
-\note{The Argennon virtual machine has no instructions for issuing cryptographic signatures.}
-
-\section{The Argennon Standard Library}\label{sec:asl}
-
-As we explained in Section~\ref{subsec:method-invocation} by using \texttt{invoke\_internal} instruction, a smart
-contract can invoke methods of another smart contract in its own context. AVM smart contracts use this instruction to
-invoke methods of the AVM standard library. Methods of the AVM standard library are stored
-with the \texttt{applicationID} zero in the AVM method area. This implies that the AVM standard library is actually
+An application can invoke methods of the ArgC Standard Library (ASL) in its own context. Methods of the ArgC standard
+library are stored as a special application in the Argennon blockchain.
 a part of the root smart contract.
-
-In Argennon, the root smart contract is an updatable smart contract, which can be updated by the Argennon governance
-system. This means that bugs or security vulnerabilities in the AVM standard library could be quickly patched and
-smart contracts could benefit from bugfixes and improvements of the AVM standard library even if they are
+In Argennon, some applications (smart contracts) are updatable. The ArgC Standard Library is an updatable smart
+contract which can be updated by the Argennon governance
+system. This means that bugs or security vulnerabilities in the ArgC Standard Library could be quickly patched and
+applications could benefit from bugfixes and improvements of the ArgC Standard Library even if they are
 non-updatable. Many important and useful functionalities,
 such as fungible and non-fungible assets, access control mechanisms,
-and general purpose DAOs are implemented in the Argennon standard library.
+and general purpose DAOs are implemented in the ArgC Standard Library.
 
@@
 of behaviour from a contract which complies with an Argennon standard.
+
+\section{Parallel Request Execution}\label{sec:concurrency}
+\input{sections/concurrency}

### 2022-03-31 tex/sections/avm.tex
Remove usages of "ZK-EDB"
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 Environment (AscEE) is an abstract execution environment for executing Argennon smarts contracts (a.k.a\ applications).
-An Argennon application essentially is a http server whose state is kept in the Argennon blockchain and
+An Argennon application essentially is an HTTP server whose state is kept in the Argennon blockchain and
 its logic is described using an Argennon Standard Representation (ASR).
@@
 usually can be executed, with minimal pre-processing, by a JIT compiler or an emulator. The high level
-language is intended for preserving high level information of applications logic to facilitate
+language is intended for preserving the high level information of applications logic to facilitate
 platform specific compiler optimization at host nodes. On the other hand, the low level language is designed for
@@
 The state of an Argennon application is stored in byte addressable finite arrays of memory called
-\emph{heap chunks}. An application can have several heap chunks with different sizes, and can remove or
+\emph{heap chunks}. An application may have several heap chunks with different sizes, and can remove or
 resize its heap chunks or allocate new chunks. Every chunk belongs to exactly one application and can only be modified
@@
 
-The Argennon Smart Contract Execution Environment can be seen as a machine for executing http requests, producing
-their http response and updating related heap chunks. The AscEE executes requests sequentially \footnote{Actually
+The Argennon Smart Contract Execution Environment can be seen as a machine for executing HTTP requests, producing
+their HTTP response and updating related heap chunks. The AscEE executes requests sequentially \footnote{Actually
 requests are executed in parallel but by performing data dependency analysis the result is guaranteed to be identical
 with sequential execution of requests.} and each request has a separate \emph{execution session}. Execution sessions
-are separate sessions of executing smart contract's code in order to fulfill \emph{external} http requests. An
+are separate sessions of executing smart contract's code in order to fulfill \emph{external} HTTP requests. An
 external request is a request that is not made by another Argennon application.
@@
 
-During an execution session an application can make \emph{internal} http requests to other applications. Those
+During an execution session an application can make \emph{internal} HTTP requests to other applications. Those
 requests will not start a new execution session and will be executed within the current session. In AscEE making a
-http request to an application is similar to a function invocation. For that reason, we also refer to internal requests
+HTTP request to an application is similar to a function invocation. For that reason, we also refer to internal requests
 as application calls.
@@
 That means, for validating that request a validator first needs to "inject" the digest of attached requests into the
-http request.
+HTTP request.
 
@@
 session, some resources are measured per application call and some
-caps are determined dynamically by the external http request (i.e.\ transaction). An optimistic session must always
+caps are determined dynamically by the external HTTP request (i.e.\ transaction). An optimistic session must always
 complete normally, a monitored session is allowed to either complete normally or abruptly.
@@
     will be considered as local memory usage. Local memory is not persistent and when an application finishes serving a
-    request and returns the http response (i.e.\ the application call completes) its local memory is deleted.
+    request and returns the HTTP response (i.e.\ the application call completes) its local memory is deleted.
     This resource is measured in bytes.

### 2022-03-31 tex/sections/persistence.tex
Remove usages of "ZK-EDB"
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The Argennon Virtual Machine has two persistent memory areas: \emph{method area}, and \emph{heap}. Method area stores
-bytecodes of methods\footnote{also it stores constant area blocks.}, and heap stores memory chunks. Both of these
-data elements, bytecodes and chunks, can be considered as continuous pieces of byte addressable memory. Throughout this
-chapter, we shall call these data elements \emph{objects}.
+The Argennon Smart Contract Execution Environment has two persistent memory areas: \emph{code area}, and \emph{heap}.
+Code area stores the Argennon Standard Representation of
+applications\footnote{also it stores applications' constants.}, and heap stores heap chunks.
+Both of these data elements, ASRs and chunks, can be considered as continuous arrays of bytes.
+Throughout this chapter, we shall call these data elements \emph{objects}.
 
@@
 
-In the AVM persistence layer, similar objects are clustered together and constitute a bigger data element which we call
-a \emph{page}.\footnote{we avoid calling them clusters, because usually a cluster refers to a \emph{set}. AVM object
+In the AscEE persistence layer, similar objects are clustered together and constitute a bigger data element which we
+call
+a \emph{page}.\footnote{we avoid calling them clusters, because usually a cluster refers to a \emph{set}. AscEE object
 clusters are not sets. They are ordered lists, like a page containing an ordered list of words or sentences.}
@@
 \[
-    P = (O_1,O_2,\dots,O_n),\quad i < j \; \Leftrightarrow \; \textrm{$O_i$ was added before $O_j$}\ .
+    P = (O_1,O_2,\dots,O_n),\quad i < j \; \Leftrightarrow \; \textrm{$O_i$ was added before $O_j$}\ \cdot
 \]
 
-A page of the AVM storage should contain objects that have very similar access pattern. Ideally, when a page is needed
+A page of the AscEE storage should contain objects that have very similar access pattern. Ideally, when a page is needed
 for validating a block, almost all of its objects should be needed for either reading or writing. We prefer
@@
 
-\section{Zero-knowledge Databases}\label{sec:zk-edb}
+\section{Publicly Verifiable Database Servers}\label{sec:zk-edb}
 
-Pages of the AVM storage are persisted using updatable zero-knowledge elementary databases (ZK-EDB). Argennon
-has three zero-knowledge databases: \emph{staking} database, which stores all the data that is associated with
-the Argennon consensus protocol. \emph{method} database, which stores the AVM method area, and \emph{heap} database,
-which stores the AVM heap. The commitment of these three ZK-EDBs are included in every block of the Argennon blockchain.
-
-We consider the following properties for a ZK-EDB:
+Pages of the AscEE storage are persisted using \emph{dynamic cryptographic accumulators}. Argennon
+has three dynamic cryptographic accumulators: \emph{staking} database, which stores all the data that is associated with
+the Argennon consensus protocol. \emph{code} database, which stores the AscEE code area, and \emph{heap} database,
+which stores the AscEE heap. The commitment of these three accumulators are included in every block of the Argennon
+blockchain. These accumulators, in the Argennon network, are hosted on special database nodes called Publicly
+Verifiable Data Base (PV-DB) servers.
+We consider the following properties for a PV-DB:
 \begin{itemize}
-    \item The ZK-EDB contains a mapping from a set of keys to a set of values.
+    \item The PV-DB contains a mapping from a set of keys to a set of values.
     \item Every state of the database has a commitment \(C\).
-    \item The ZK-EDB has a method \((D, \pi) = \text{get}(x)\), where \(x\) is a key and \(D\) is the associated data
+    \item The PV-DB has a method \((D, \pi) = \text{get}(x)\), where \(x\) is a key and \(D\) is the associated data
     with \(x\), and \(\pi\) is a proof.
     \item A user having \(C\) and \(\pi\) can verify that \(D\) is really associated with \(x\), and \(D\) is not
-    altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB\@.
+    altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the PV-DB\@.
     \item Having \(\pi\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
@@
 
-Pages of the AVM storage are stored in the ZK-EDBs, with an index: \texttt{pageIndex} as their key.
+Pages of the AscEE storage are stored in the PV-DBs, with an index: \texttt{pageIndex} as their key.
 The \texttt{pageIndex} is required to be smaller than a certain value, determined by the
-protocol, to facilitate the usage of ZK-EDBs that are based on vector commitments.
-For this reason, the AVM clustering algorithm always tries to reuse indices and keep the number of used indices
+protocol, to facilitate the usage of PV-DBs that are based on vector commitments.
+For this reason, the AscEE clustering algorithm always tries to reuse indices and keep the number of used indices
 as low as possible.
 
-The commitments of the AVM ZK-EDBs are affected by the way data objects are clustered. Therefore,
+The commitments of the AscEE cryptographic accumulators are affected by the way data objects are clustered. Therefore,
 the Argennon clustering algorithm has to be a part of the consensus protocol.
@@
 
-A block proposer could obtain clustering directives from any third party source\footnote{we can say the AVM clustering
+A block proposer could obtain clustering directives from any third party source\footnote{we can say the AscEE clustering
 algorithm is essentially off-chain.}. This will not

### 2022-04-01 tex/sections/avm.tex
Add introduction for three-step request execution
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 storing temporary data.
+
+The AscEE executes the requests contained in each block of the Argennon
+blockchain in a three-step procedure. The first step is the \emph{preprocessing step}. In this step, the required
+data structures for executing requests are constructed. This step is
+designed in a way that can be done fully in parallel for each request without any risks of data races. The second
+step is the \emph{Data Dependency Analysis (DDA) step}.
+In this step by analyzing data dependency between requests, the AscEE determines requests that can be run in parallel
+and requests that need to be run sequentially. This information is represented using an \emph{execution dag} data
+structure. In the final step, request are executed using this data structure and helper data structures built in the
+first step.
 
@@
 
-\section{Parallel Request Execution}\label{sec:concurrency}
+\section{Data Dependency Analysis}\label{sec:concurrency}
 \input{sections/concurrency}

### 2022-04-01 tex/sections/elements.tex
Update some parts of `elements.tex`
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-An Argennon application or a smart contract is a collection of method bytecodes and heap chunks that are stored in
-the AVM storage, identified by a unique application identifier. An application identifier, \texttt{applicationID}, is
+An Argennon application or smart contract is an HTTP server which is represented by an Argennon Standard
+Representation (ASR) and whose state is stored in the Argennon blockchain. Each Argennon application is identified by
+a unique application identifier.
+
+An application identifier, \texttt{applicationID}, is
 a unique prefix code generated by the \emph{applications} prefix tree. (See Section~\ref{sec:identifiers}.)
-
-An application identifier can
-be considered as the address of an application and has the following standard symbolic representation:
+An application identifier can be considered as the address of an application and has the following standard symbolic
+representation:
 \begin{verbatim}
@@
 \begin{itemize}
-    \item Installation of AVM smart contracts and determining the update policy of a smart
+    \item Installation of new Argennon applications and determining the update policy of a smart
     contract: if the contract is updatable or not, which accounts or smart contracts can update or uninstall
     the contract, and so on.
-    \item Removing an AVM smart contract (if allowed).
-    \item Updating an AVM smart contract (if allowed).
+    \item Removing an Argennon application (if allowed).
+    \item Updating an Argennon application (if allowed).
 \end{itemize}
@@
 
-Argennon accounts are entities defined inside the ARG smart contract.
+Argennon accounts are entities defined inside the ARG application.
 Every Argennon account is uniquely identified by a prefix code generated using \emph{accounts} prefix
@@
 \begin{verbatim}
-<account-id> ::= <hex-prefix-code>
-<hex-prefix-code> ::= "0x"<hex-num>"."<hex-prefix-code> | "0x"<hex-num>
+<account-id> ::= "0x"<hex-num>
 \end{verbatim}
-where \texttt{<hex-num>} is a hexadecimal number between $0$ and $255$, using lower case
+where \texttt{<hex-num>} is a hexadecimal number, using lower case
 letters \texttt{[a-f]} for showing digits greater than $9$.
 
-For example \texttt{0x24.0xff.0xda}, \texttt{0x0} and \texttt{0x3.0xa0.0x0.0x0}, are valid standard symbolic
+For example \texttt{0x24ffda}, \texttt{0x0} and \texttt{0x03a0000}, are valid standard symbolic
 representations of account addresses.
 
-A new account can be created by invoking \texttt{createAccount} method from the Argennon smart contract. For creating
+A new account can be created by sending a proper HTTP request to the ARG smart contract. For creating
 a new account two public keys need to be provided by the caller and registered in the Argennon smart contract.
@@
 \section{Transactions}\label{sec:transactions}
+\emph{subsection is outdated and needs updating!}
 
@@
 \subsection{Resource Declaration}\label{subsec:resource-declaration}
+\emph{subsection is outdated and needs updating!}
 
@@
 
-\begin{lstlisting}[language=python, frame=TB, float, title=An Argennon transaction in YAML format,label={lst:txn-example}]
+\begin{lstlisting}[language=python, frame=TB, float, title=An Argennon transaction in YAML format,
+    label={lst:txn-example}]
 ---

### 2022-04-01 tex/sections/persistence.tex
Update some parts of `elements.tex`
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 applications\footnote{also it stores applications' constants.}, and heap stores heap chunks.
-Both of these data elements, ASRs and chunks, can be considered as continuous arrays of bytes.
+Both of these data elements, ASRs and heap chunks, can be considered as continuous arrays of bytes.
 Throughout this chapter, we shall call these data elements \emph{objects}.
@@
 blockchain. These accumulators, in the Argennon network, are hosted on special database nodes called Publicly
-Verifiable Data Base (PV-DB) servers.
+Verifiable Database (PV-DB) servers.
 We consider the following properties for a PV-DB:

### 2022-04-02 tex/sections/avm.tex
Fix some readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 platform specific compiler optimization at host nodes. On the other hand, the low level language is designed for
-efficient execution of applications that are not frequently used.
+efficient direct execution of applications that are not frequently used.
 
@@
 In this step by analyzing data dependency between requests, the AscEE determines requests that can be run in parallel
-and requests that need to be run sequentially. This information is represented using an \emph{execution dag} data
-structure. In the final step, request are executed using this data structure and helper data structures built in the
-first step.
+and requests that need to be run sequentially. This information is represented using an \emph{execution dag} and in
+the final step, request are executed using this data structure.
 
@@
 
-The Argennon Smart Contract Execution Environment can be seen as a machine for executing HTTP requests, producing
-their HTTP response and updating related heap chunks. The AscEE executes requests sequentially \footnote{Actually
-requests are executed in parallel but by performing data dependency analysis the result is guaranteed to be identical
+The Argennon Smart Contract Execution Environment can be seen as a machine for executing Argennon applications to
+fulfill HTTP requests, produce their HTTP responses and update related heap chunks. The AscEE executes requests
+sequentially \footnote{Actually requests are executed in parallel but by performing data dependency analysis the
+result is guaranteed to be identical
 with sequential execution of requests.} and each request has a separate \emph{execution session}. Execution sessions
@@
 
-The state and data of an execution session will be
+The state of an execution session will be
 destroyed at the end of the session and only the state of heap chunks is preserved. If a session fails and does not
@@
 During an execution session an application can make \emph{internal} HTTP requests to other applications. Those
-requests will not start a new execution session and will be executed within the current session. In AscEE making a
+requests will not start a new execution session and will be executed within the current session. In AscEE making an
 HTTP request to an application is similar to a function invocation. For that reason, we also refer to internal requests
@@
 The AscEE is designed based on \emph{optional decoupling principle}. When an application makes a request to another
-application, optionally it can request to be decoupled from the called application. That would mean the called
+application, optionally it can choose to be decoupled from the called application. That would mean the called
 application could not affect its caller's state by reentrancy, or could not fail the session by using excessive
@@
 
-In Argennon a unique identifier is assigned to every application, heap chunk and account. Therefore, three distinct
+In Argennon a unique identifier is assigned to every application, heap chunk and account. Consequently, three distinct
 identifier types exist: \texttt{appID}, \texttt{accountID}, and \texttt{chunkID}.
@@
 \begin{itemize}
-    \item \texttt{appID} is a prefix code built by \emph{applications} prefix tree. An application ID cannot
+    \item \texttt{appID} is a prefix code built by \emph{applications} prefix tree. An \texttt{appID} cannot
     be \texttt{0x0}.
 
-    \item \texttt{accountID} is a prefix code built by \emph{accounts} prefix tree. An account ID cannot
+    \item \texttt{accountID} is a prefix code built by \emph{accounts} prefix tree. An \texttt{accountID} cannot
     be \texttt{0x0} or \texttt{0x1}.
@@
 resize its chunks or allocate new chunks. Every chunk belongs to exactly one application. Only the owner application can
-modify a chunk but there is no restrictions for reading a chunk.
+modify a chunk but there is no restrictions for reading a chunk \footnote{The reason behind this type of access
+control design is the fact that smart contract
+code is usually immutable. That means if a smart contract does not implement a
+getter mechanism for some parts of its internal data, this functionality can never
+be added later, and despite the internal data is publicly available, there will be no
+way for other smart contracts to use this data on-chain.}.
 
 When an application allocates a new heap chunk, the identifier of the new chunk is not generated by
-the AscEE. Instead, the application can choose an identifier itself \footnote{provided it has a correct format}. This
-is an important feature of the AscEE's heap, which allows applications to use the AscEE's heap as a dictionary (map)
-data structure.
+the AscEE. Instead, the application can choose an identifier itself, provided it has a correct format. This
+is an important feature of the AscEE heap, which allows applications to use the AscEE heap as a dictionary
+data structure \footnote{also called a map.}.
 Since the \texttt{chunkID} is a prefix code, any application has its own identifier space, and an application
@@
 the value of \texttt{sizeUpperBound} is constant for every block of the blockchain and is proposed by the block
-proposer. \texttt{sizeUpperBound} indicates the upper bound of the value of \texttt{chunkSize} for a chunk in a block.
+proposer. \texttt{sizeUpperBound} indicates the upper bound of \texttt{chunkSize} value for a chunk in a block.
 A validator can use this value to safely allocate memory for a chunk.
@@
 offset is persistent or not: Offsets lower than the chunk size are persistent, and higher offsets are not.
-Non-persistent locations will be re-initialized with zero, at the start of every execution session.
+Non-persistent locations will be re-initialized with zero at the start of every execution session.
 
@@
 Memory locations inside a chunk can only be accessed through access blocks. An access block is defined on a chunk
-and has an offset and a size and determines accessible memory locations inside a chunk. Multiple access blocks can
+and has an offset and a size and determines accessible memory locations inside the chunk. Multiple access blocks can
 be defined on a single chunk, but they must be non overlapping. Access blocks are byte addressable and can have
-different
-access types:
+different access types:
 
@@
     status of a memory location.
-    \item \texttt{additive}: only allows addition operations without overflow checking. Note that the content of these
-    access blocks cannot be read.
+    \item \texttt{additive}: only allows addition-like operations without overflow checking. Note that the content of
+    these access blocks cannot be read.
 \end{itemize}
 
-Chunks are intended for simplifying proof checking of the data stored in the Argennon cloud and access blocks are
+Chunks and access blocks have different purposes. Chunks are intended for simplifying proof checking of the data
+stored in the Argennon cloud \footnote{Note that
+the Argennon cloud uses cryptographic accumulators for preserving data integrity.} and access blocks are
 required for better parallelization of the request execution. An application should put the data it predicts is
@@
 
-An application can load a chunk with a valid prefix identifier even if that chunk does not exist. For a non-existent
+An application can load any chunk with a valid prefix identifier even if that chunk does not exist. For a non-existent
 chunk the value of \texttt{chunkSize} is always zero.
@@
 
-Attachment of a request is a list of request identifiers of the current block that are "attached" to the request.
-That means, for validating that request a validator first needs to "inject" the digest of attached requests into the
-HTTP request.
+The attachment of a request is a list of request identifiers of the current block that are ``attached'' to the request.
+That means, for validating that request a validator first needs to ``inject'' the digest of attached requests into the
+HTTP request. By doing so, the called application will have access to the digest of attachments in a secure way.
 
-The main usage of this feature is for fee payment. A request that wants to pay the fees for a number of requests
+The main usage of this feature is for fee payment. A request that wants to pay the fees for a number of requests,
 declares those requests as its attachments. For paying fees the payer signs the digest of requests for which he
 wants to pay fees. After injecting the digest of those request by validators, that signature can be validated
-correctly and securely.
+correctly and securely by the called application.
 
@@
 
-The AscEE uses \emph{authenticated message passing} for authorizing operations. In this method every execution
-session has a set of authenticated messages. Applications explicitly pass those messages for authorizing operations.
-Authentication of messages can be done using any type of cryptographic signature scheme in the preprocessing phase of
-request execution. For example, we could use BLS aggregate signature to authenticate all messages of a block in bulk.
+The AscEE uses \emph{authenticated message passing} for authorizing operations. In this method, every execution
+session has a set of authenticated messages, and applications explicitly pass those messages for authorizing operations.
+Authentication of messages can be efficiently done in parallel, using any type of cryptographic signature scheme in
+the preprocessing phase of request execution. For example, we could use BLS aggregate signature to authenticate all
+messages of a block in bulk.
 
-Moreover, application use built-in functions of AscEE to generate authenticated messages in run-time. This enables an
-application to authorize an operation even for an application that is not directly called.
-
-\note{Authorization by explicit messages and signatures eliminates the need for approval methods or call back patterns.}
+Moreover, applications can use built-in functions of AscEE to generate authenticated messages in run-time. This enables
+an application to authorize operations even for applications that are not directly called.
 
@@
 
-\note{The Argennon Execution Environment has no instructions for issuing cryptographic signatures.}
+Authorization by explicit messages and signatures eliminates the need for approval methods or call back patterns in
+Argennon.
+
+\note{The AscEE has no instructions for issuing cryptographic signatures.}
 
@@
 to do so will result in a revert. The entrance lock of an application will be released when the application explicitly
-releases its entrance lock or when the call that had acquired the lock completes.
+releases its lock or when the call that had acquired that lock completes.
 
@@
 measurement. Note that for implementing optional decoupling principle, we need to determine the exact application
-which has failed in a call chain. For example assume that an optimistic session containing an application call
-violates a 2 milliseconds execution time cap. If in the caller application the call happens exactly after 2
-milliseconds of execution, a small fluctuation in the execution time measurement can
+which has failed in a call chain. For example, assume that an optimistic session containing an application call
+violates a 2 milliseconds execution time cap. If the call happens exactly after 2 milliseconds
+in the caller application, a small fluctuation in the execution time measurement can
 change the point of failure between the called and the main application. This can introduce nondeterministic
-behaviour which could make block validation impossible. On the other hand, for a monitored session this is not an issue
-because resource measurements and caps are defined per application, and we can make sure that resource caps are not too
-small and can be measured easily.
+behaviour which could make block validation impossible. That's why monitored sessions exist. For a monitored session
+this is not an issue because resource measurements and caps are defined per application call. This way, we can make
+sure that resource caps are not too small and can be measured easily.
 
@@
     \item \textbf{execution time}:
-    the amount of cpu time that is required for executing a session or an application call. The execution time of a
-    session is measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
-    of cpu time needed for executing a predefined standard application which is used for benchmarking a host's
+    is the amount of cpu time that is required for executing a session or an application call. The execution time is
+    measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
+    of \textbf{cpu time} needed for executing a predefined standard application which is used for benchmarking a host's
     performance.
@@
     session has a separate \texttt{maxClocks} value. This value determines the maximum amount of time that the cpu
-    can be used for executing that particular application call. It should be noted that the cpu timer is paused when
-    the application makes a call to another application, and is resumed when the control returns. An application call
+    can be used for executing that particular application call. It should be noted that the cpu timer of the
+    application call is paused when the application makes a call to another application, and is resumed when the
+    control returns. An application call
     needs at least 100 clocks and if the value of its \texttt{maxClock} is lower than this value the call will be
@@
 
-    Optimistic sessions measure local memory usage per session and enforce a protocol-defined cap on the total amount of
-    used local memory in the session. Monitored sessions measure local memory usage per
-    application call and enforce a protocol defined-cap per application call. An application call which tries to use
-    more local memory than the cap will fail.
+    Optimistic sessions measure local memory usage per session and enforce a protocol-defined cap on the total amount
+    of local memory a session can use. Monitored sessions measure local memory usage per
+    application call and enforce a protocol-defined cap per application call. An application call which tries to use
+    more local memory than the cap, will fail.
     \item \textbf{heap access list}:
@@
 
-The Argennon protocol allows AscEE's implementations to have at max \%50 error in measuring execution time and at
+The Argennon protocol allows an AscEE's implementation to have at max \%50 error in measuring execution time and at
 max \%25 error in measuring local memory usage. Other resources must be measured precisely.

### 2022-04-03 tex/A.tex
Update Chapter 1
This is the first draft.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Restructure the paper
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
+\lstset{
+    language=Java,
+    morekeywords={main, initialize, constructor, external, virtual, Map, Account, signature, overrides, is, string},
+    basicstyle=\small\sffamily,
+    columns=fullflexible,
+%   keywordstyle=\bfseries,
+    breaklines=true,
+    showstringspaces=false,
+    mathescape
+}
+
 \title{Argennon: A Scalable Smart Contract Platform}
@@
 \begin{document}
-    %\maketitle
+    \maketitle
+    \begin{abstract}
+        Argennon is a cloud based blockchain and smart
+        contract platform which tries to solve shortcomings of existing platforms. The Argennon blockchain uses a hybrid
+        proof of stake (POS) consensus protocol, which is based on participation of every Argennon user.
+
+        Thanks to the
+        cloud based design of the Argennon blockchain, transaction validation does not require costly computational
+        resources, and normal personal computers with limited hardware capabilities can actively participate in the
+        Argennon consensus protocol. This property makes Argennon a truly decentralized and democratic blockchain and
+        one of the most secure existing platforms.
+
+        By using cryptographic accumulators, the integrity of data on the Argennon cloud is guaranteed, and there
+        is no need for trusting cloud servers. At the same time, by using a smart clustering algorithm the network usage
+        and the overhead of the commitment scheme is kept manageable.
+    \end{abstract}
     \tableofcontents
@@
 
-    \chapter{The Argon Language}\label{ch:argon-lang}
-    \input{sections/argon}
-
     \chapter{Persistence Layer}\label{ch:persistance}
@@
     \chapter{Networking Layer}\label{ch:networking}
+
     \section{Normal Mode}\label{sec:normal-mode}
@@
     \chapter{The Argennon Blockchain}\label{ch:argennon-blockchain}
-    \input{sections/elements}
 
@@
 
+    \input{sections/elements}
+
     \section{Incentive mechanism}\label{sec:incentive-mechanism}
@@
 
-    \section{Concurrent Transaction Validation}\label{sec:concurrency}
-    \input{sections/concurrency}
+    \chapter{Governance}\label{ch:governance}
 
-    \chapter{Governance}\label{ch:governance}
     \section{ADAGs}\label{sec:adags}
@@
     \note{not yet written...}
+
+    \chapter{The Argon Language}\label{ch:argon-lang}
+    \input{sections/argon}
 \end{document}

### 2022-04-03 tex/sections/avm.tex
Restructure the paper
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 and requests that need to be run sequentially. This information is represented using an \emph{execution dag} and in
-the final step, request are executed using this data structure.
+the final step, requests are executed using this data structure.
 
@@
 fulfill HTTP requests, produce their HTTP responses and update related heap chunks. The AscEE executes requests
-sequentially \footnote{Actually requests are executed in parallel but by performing data dependency analysis the
+sequentially\footnote{Actually requests are executed in parallel but by performing data dependency analysis the
 result is guaranteed to be identical
@@
 resize its chunks or allocate new chunks. Every chunk belongs to exactly one application. Only the owner application can
-modify a chunk but there is no restrictions for reading a chunk \footnote{The reason behind this type of access
+modify a chunk but there is no restrictions for reading a chunk\footnote{The reason behind this type of access
 control design is the fact that smart contract
@@
 is an important feature of the AscEE heap, which allows applications to use the AscEE heap as a dictionary
-data structure \footnote{also called a map.}.
+data structure\footnote{also called a map.}.
 Since the \texttt{chunkID} is a prefix code, any application has its own identifier space, and an application
@@
 Chunks and access blocks have different purposes. Chunks are intended for simplifying proof checking of the data
-stored in the Argennon cloud \footnote{Note that
+stored in the Argennon cloud\footnote{Note that
 the Argennon cloud uses cryptographic accumulators for preserving data integrity.} and access blocks are

### 2022-04-03 tex/sections/elements.tex
Restructure the paper
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 \end{verbatim}
-where \texttt{<dec-num>} is a normal decimal number between $0$ and $255$.
-
-For example \texttt{21.255.37}, \texttt{0}, \texttt{11.6} and \texttt{2.0.0.0.0}, are valid application addresses.
+where \texttt{<dec-num>} is a normal decimal number between $0$ and $255$. For example \texttt{21.255.37},
+\texttt{0}, \texttt{11.6} and \texttt{2.0.0.0.0}, are valid application addresses.
 
@@
 the \emph{ARG smart contract}, which is also called the \emph{Argennon smart contract} or the \emph{ARG application}.
+
+Argennon application use HTTP as the application protocol and they are advised to have a RESTful API design.
 
@@
 \section{Transactions}\label{sec:transactions}
-\emph{subsection is outdated and needs updating!}
 
-Every Argennon transaction consists of two \texttt{i\_invoke\_dispatcher} instructions, the first instruction always
-transfers the proposed fee of the transaction in ARGs from a sender account to the fee sink accounts, and
-the second performs the requested operation. If the first instruction fails, the transaction will not be added to
-the Argennon blockchain.
-
-Users interact with AVM smart contracts using
-the second \texttt{i\_invoke\_dispatcher} instruction of a transaction.
-Transferring all assets, including ARG, is done by that instruction.
-
-The \texttt{i\_invoke\_dispatcher} instruction invokes the \texttt{dispatcher} method of a smart contract. Every AVM
-smart contract has a \texttt{dispatcher} method, and the Argennon protocol requires
-the \texttt{dispatcher} method to accept an HTTP request as its argument and
-return an HTTP response.
-
-As a result, every Argennon transaction contains two HTTP requests. Besides these HTTP requests, the
-transaction is also required to contain a resource declaration object, specifying the maximum amount of resources
-it needs for execution.
-
-\note{Argennon smart contracts use HTTP as the application protocol and they are advised to have a RESTful API design.}
+An Argennon transaction consist of an HTTP request made by a user, to an Argennon application and a resource
+declaration object. Transactions can only be
+issued by users and applications can not create transactions. An Argennon transaction is also called
+an \emph{external request}.
 
 \subsection{Resource Declaration}\label{subsec:resource-declaration}
-\emph{subsection is outdated and needs updating!}
 
-Every Argennon transaction is required to specify a cap for all the resources it needs. This
-includes memory, network and processor related resources. If during emulation, a transaction reaches one of its
-pre-declared resource caps, executing any AVM instruction which uses that resource, will result in an AVM exception.
-
-As we know, every Argennon transaction consists of two \texttt{i\_invoke\_dipatcher} instructions. The first
-instruction is always considered as a \emph{free} instruction and resources spent during its execution
-session will not be counted. Therefore, only the second instruction could fail due to exceeding resource limits.
-
-The Argennon protocol defines an execution cost for
+Every Argennon transaction needs to specify a maximum for
+The Argennon protocol requires every  an execution cost for
 every AVM instruction, reflecting the amount of resources its emulation needs. Every

### 2022-04-14 tex/A.tex
Update abstract
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
         and the overhead of the commitment scheme is kept manageable.
+
+        The Argennon network relies on a permission-less network of Publicly Verifiable Database (PV-DB) servers,
+        which forms the Argennon cloud. A PV-DB server is a conventional data server which uses its computational and
+        storage resources to help the
+        Argennon network process transactions. A large portion of incentive rewards in the Argennon platform is
+        devoted to PV-DB servers. This will incentivize the development of conventional networking, storage and
+        computational hardware, which can benefit all areas of information technology.
     \end{abstract}

### 2022-04-14 tex/sections/avm.tex
Update abstract
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The Argennon\footnote{The classical pronunciation should be used:\textipa{/Ar"gen.non/}} Smart Contract Execution
-Environment (AscEE) is an abstract execution environment for executing Argennon smarts contracts (a.k.a\ applications).
-An Argennon application essentially is an HTTP server whose state is kept in the Argennon blockchain and
+Environment (AscEE) is an abstract execution environment for executing Argennon smarts contracts (a.k.a\ Argennon
+applications). An Argennon application essentially is an HTTP server whose state is kept in the Argennon blockchain and
 its logic is described using an Argennon Standard Representation (ASR).
@@
 language is intended for preserving the high level information of applications logic to facilitate
-platform specific compiler optimization at host nodes. On the other hand, the low level language is designed for
+platform specific compiler optimization at host nodes. The low level language, is designed for
 efficient direct execution of applications that are not frequently used.
@@
 blockchain in a three-step procedure. The first step is the \emph{preprocessing step}. In this step, the required
-data structures for executing requests are constructed. This step is
+data for executing requests are retrieved from the Argennon cloud and the helper data structures are constructed.
+This step is
 designed in a way that can be done fully in parallel for each request without any risks of data races. The second
@@
 In this step by analyzing data dependency between requests, the AscEE determines requests that can be run in parallel
-and requests that need to be run sequentially. This information is represented using an \emph{execution dag} and in
-the final step, requests are executed using this data structure.
+and requests that need to be run sequentially. This information is represented using an \emph{execution dag} data
+structure and in the final step, requests are executed using that data structure.
 

### 2022-04-15 tex/sections/avm.tex
Improve readability of "Resource Management" subsection
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 language is intended for preserving the high level information of applications logic to facilitate
-platform specific compiler optimization at host nodes. The low level language, is designed for
+platform specific compiler optimization at host nodes. The low level language, on the oder hand, is designed for
 efficient direct execution of applications that are not frequently used.
@@
 resize its heap chunks or allocate new chunks. Every chunk belongs to exactly one application and can only be modified
-by its owner. In addition to heap chunks, every application also has an amount of non-persistent local memory for
+by its owner. In addition to heap chunks, every application has an amount of non-persistent local memory for
 storing temporary data.
@@
 The Argennon Smart Contract Execution Environment can be seen as a machine for executing Argennon applications to
-fulfill HTTP requests, produce their HTTP responses and update related heap chunks. The AscEE executes requests
-sequentially\footnote{Actually requests are executed in parallel but by performing data dependency analysis the
+fulfill HTTP requests, produce their HTTP responses and update related heap chunks. The execution of requests can be
+considered sequential\footnote{Actually requests are executed in parallel but by performing data dependency analysis the
 result is guaranteed to be identical
-with sequential execution of requests.} and each request has a separate \emph{execution session}. Execution sessions
-are separate sessions of executing smart contract's code in order to fulfill \emph{external} HTTP requests. An
-external request is a request that is not made by another Argennon application.
+with sequential execution of requests.} and each request has a separate \emph{execution session}. An execution session
+is a separate session of executing smart contract's code in order to fulfill an \emph{external} HTTP request.
+External requests are requests that are not made by other Argennon applications.
 
@@
 requests will not start a new execution session and will be executed within the current session. In AscEE making an
-HTTP request to an application is similar to a function invocation. For that reason, we also refer to internal requests
-as application calls.
+HTTP request to an application is similar to a function invocation, and for that reason, we also refer to them as
+application calls.
 
@@
 All Argennon prefix trees have an equal branching factor \(\beta\). Therefore, we can represent an Argennon
-prefix tree as a sequence of fractional numbers\footnote{It's possible to have \(a_i=0\). For
-exmaple \(A^{(4)}=(0.2000)_{10}\) is correct.} in base \(\beta\):
+prefix tree as a sequence of fractional numbers\footnote{It's possible to have \(a_i=0\).
+For example, \(A^{(4)}=(0.2000)_{10}\) is correct.} in base \(\beta\):
 \[
@@
 \]
-where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)} \leq A^{(i+1)}\). A typical choice for
-\(\beta\) could be \(2^8\).
+where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)} \leq A^{(i+1)}\).\footnote{A typical choice for
+    \(\beta\) is \(2^8\).}
 
@@
 
-The AscEE uses \emph{authenticated message passing} for authorizing operations. In this method, every execution
-session has a set of authenticated messages, and applications explicitly pass those messages for authorizing operations.
+The AscEE uses \emph{Authenticated Message Passing} for authorizing operations. In this method, every execution
+session has a set of authenticated messages, and applications \textbf{explicitly} pass those messages for authorizing
+operations.
 Authentication of messages can be efficiently done in parallel, using any type of cryptographic signature scheme in
@@
 
-Moreover, applications can use built-in functions of AscEE to generate authenticated messages in run-time. This enables
-an application to authorize operations even for applications that are not directly called.
+Moreover, applications use built-in functions of AscEE to generate authenticated messages in run-time. This enables
+an application to authorize operations for another application even if that application is not directly called by
+itself.
 
-In addition to authenticated messages the AscEE provides a set of
-cryptographic functions for validating signatures and cryptographic entities. By using these functionalities and
+In addition to authenticated messages, the AscEE provides a set of
+cryptographic functions for validating signatures and cryptographic entities. By using these functions and
 passing cryptographic signatures as parameters to methods, a programmer, having users' public keys, can implement
@@
 
-Authorization by explicit messages and signatures eliminates the need for approval methods or call back patterns in
-Argennon.
+Authorization by explicit authenticated messages and signatures eliminates the need for approval mechanisms or call back
+patterns in Argennon.
 
@@
 Completing an execution session requires computational resources. The amount of resources used by an execution session
-should be monitored, otherwise a malicious user would be able to easily spam and exhaust resources of the execution
-environment. Resource usage can be measured per session or per application call.
+should be monitored and managed, otherwise a malicious user would be able to easily spam and exhaust resources of the
+execution environment.
+
+Resource usage can be measured per session or per application call. Obviously per session measurement is
+easier and more efficient. However, when we measure resources per session if
+a session violates its resource caps, determining the point of failure requires precise and error-free resource
+measurement. Note that for implementing optional decoupling principle, we need to determine the exact application call
+which has failed in a call chain. For example, assume that a session containing an application call
+violates a 2 milliseconds execution time cap. If the call happens exactly after 2 milliseconds
+in the caller's code, a small fluctuation in the execution time measurement can
+change the point of failure between the called and the main application. This can introduce nondeterministic
+behaviour which could make block validation impossible. That's why the AscEE sometimes uses per application call
+resource management. If we define the resource caps per application call and perform our measurements for each
+application call separately, by ensuring that caps are larger than the measurement error, errors in the
+measurements would not change the point of failure in the call chain anymore.
 
 The AscEE has two type of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
-session is always measured per session and default pre-defined resource caps are used. On the other hand in a monitored
-session, some resources are measured per application call and some
-caps are determined dynamically by the external HTTP request (i.e.\ transaction). An optimistic session must always
-complete normally, a monitored session is allowed to either complete normally or abruptly.
-
-Obviously resource monitoring is easier and more efficient for optimistic sessions. However, if
-an optimistic session violates its resource caps, determining the point of failure requires precise resource
-measurement. Note that for implementing optional decoupling principle, we need to determine the exact application
-which has failed in a call chain. For example, assume that an optimistic session containing an application call
-violates a 2 milliseconds execution time cap. If the call happens exactly after 2 milliseconds
-in the caller application, a small fluctuation in the execution time measurement can
-change the point of failure between the called and the main application. This can introduce nondeterministic
-behaviour which could make block validation impossible. That's why monitored sessions exist. For a monitored session
-this is not an issue because resource measurements and caps are defined per application call. This way, we can make
-sure that resource caps are not too small and can be measured easily.
+session is always measured per session and default pre-defined resource caps are used. On the other hand, for a
+monitored session, resources that can not be measured precisely are measured per application call and their caps are
+determined per application call by the external HTTP request (i.e.\ transaction). An optimistic session must
+always complete normally, a monitored session is allowed to either complete normally or abruptly.
 
@@
     of \textbf{cpu time} needed for executing a predefined standard application which is used for benchmarking a host's
-    performance.
+    performance. In other words, by definition the AscEE imaginary standard machine completes the standard benchmark
+    in 1000 clocks.
 
     Optimistic sessions have a predefined \texttt{maxClocks} value which is determined by the Argennon protocol. This
-    value defines a bound on the \textbf{total} cpu time of the session and no per application measurement is done.
+    value defines a bound on the \textbf{total} AscEE clocks of the session, and no per application measurement is done.
 
@@
     control returns. An application call
-    needs at least 100 clocks and if the value of its \texttt{maxClock} is lower than this value the call will be
+    needs at least 10 clocks and if the value of its \texttt{maxClock} is lower than this value the call will be
     considered a failed call.
@@
     Each application call has some amount of \texttt{externalClocks}. When an application makes a request to another
-    application it has to \emph{forward} a portion of its external clocks to the called application. This amount
-    will determine the value of \texttt{maxClocks} for the called application.
+    application it has to \emph{forward} a portion of its \textbf{external} clocks to the called application. This
+    amount will determine the value of \texttt{maxClocks} for the called application.
 
@@
     \texttt{maxClocks} for
-    the root application call is determined by the external request (transaction).
+    the root application call is determined by the external request (i.e.\ transaction).
 
     \item \textbf{local memory}:
-    any memory usage of an application that is not part of a heap chunk and is not part of another resource
+    any memory usage of an application that is not part of a heap chunk and is not part of another measured resource
     will be considered as local memory usage. Local memory is not persistent and when an application finishes serving a
@@
     of local memory a session can use. Monitored sessions measure local memory usage per
-    application call and enforce a protocol-defined cap per application call. An application call which tries to use
-    more local memory than the cap, will fail.
+    application call and enforce a protocol-defined cap for each application call separately. An application call
+    which tries to use more local memory than the cap, fails.
     \item \textbf{heap access list}:
@@
     resizing heap chunks can only be done in the range of the pre-declared lower bound and upper bound.
-    \item \textbf{applications list}:
-    a session may only make requests to applications that are declared in its application list.
+    \item \textbf{app access list}:
+    a session may only make requests to applications that are declared in its application access list.
     \item \textbf{call depth}:
     during a session the number of nested application calls can not be more than a threshold. This threshold is
-    determined by the Argennon protocol.
-    \item \textbf{active differed calls}:
+    determined by the Argennon protocol. It should be noted that a differed call is considered like a normal call and
+    increases the call depth by one level.
+    \item \textbf{differed calls}:
+    every application call can have a limited number of differed calls which is determined by the protocol.
+    To simplify the implementation, this limit is defined per application call. Since in Argennon the call depth is
+    limited, a per-application call limit will also define an implicit limit for the total number of active differed
+    calls.
     \item \textbf{virtual signatures}:

### 2022-04-16 tex/sections/avm.tex
Update Resource Management
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 easier and more efficient. However, when we measure resources per session if
-a session violates its resource caps, determining the point of failure requires precise and error-free resource
+a session violates its resource caps, determining the point of failure may require precise and error-free resource
 measurement. Note that for implementing optional decoupling principle, we need to determine the exact application call
 which has failed in a call chain. For example, assume that a session containing an application call
-violates a 2 milliseconds execution time cap. If the call happens exactly after 2 milliseconds
-in the caller's code, a small fluctuation in the execution time measurement can
-change the point of failure between the called and the main application. This can introduce nondeterministic
+violates a 2 milliseconds execution time cap. If in the caller's code, the call happens exactly after 2 milliseconds,
+a small fluctuation in the execution time measurement can
+change the point of failure between the called and the caller application. This can introduce nondeterministic
 behaviour which could make block validation impossible. That's why the AscEE sometimes uses per application call
 resource management. If we define the resource caps per application call and perform our measurements for each
-application call separately, by ensuring that caps are larger than the measurement error, errors in the
-measurements would not change the point of failure in the call chain anymore.
+application call separately, by using caps that are larger than the measurement error, errors in the
+measurements would not change the point of failure in the call chain.
 
@@
 determined per application call by the external HTTP request (i.e.\ transaction). An optimistic session must
-always complete normally, a monitored session is allowed to either complete normally or abruptly.
+always complete normally, a monitored session is allowed to either complete normally or abruptly. In Argennon, the
+type of execution session for each transaction is decided by the block proposer.
 
@@
 
-The Argennon protocol allows an AscEE's implementation to have at max \%50 error in measuring execution time and at
-max \%25 error in measuring local memory usage. Other resources must be measured precisely.
+In Argennon, execution time and local memory are considered \emph{nondeterministic} resources. A nondeterministic
+resource is a resource that can not be measured precisely and its measurement always contains a random error.
+
+When a transaction fails due to violation of a limit for a nondeterministic resource, the proposer is
+required to exactly specify the application call in the session that violates that limit. When validators are
+executing a monitored session, for each application call, they enforce considerably larger limits for
+nondeterministic resources, only when the proposer has declared that an application call violates a limit the
+validators will enforce the actual value of the limit. This simple mechanism ensures that, with a very high
+probability, validators agree with the proposer.
 

### 2022-04-16 tex/sections/elements.tex
Remove outdated parts
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-An Argennon transaction consist of an HTTP request made by a user, to an Argennon application and a resource
-declaration object. Transactions can only be
+An Argennon transaction consist of an HTTP request made by a user to an Argennon application, a resource
+declaration object and a list of signed messages. Transactions can only be
 issued by users and applications can not create transactions. An Argennon transaction is also called
@@
 
-Every Argennon transaction needs to specify a maximum for
-The Argennon protocol requires every  an execution cost for
-every AVM instruction, reflecting the amount of resources its emulation needs. Every
-transaction is required to specify two maximum execution costs: \texttt{maxInternalCost}
-and \texttt{maxExternalCost}. The \emph{external} execution cost of a transaction is the \textbf{overall} cost of its
-\texttt{invoke\_dispatcher} and \texttt{invoke\_later} instructions,\footnote{By overall cost, we mean the execution
-cost needed for reaching the next instruction.}. The remaining execution cost will be considered as
-the \emph{internal} cost. If a transaction reaches one of its maximum execution costs, executing any instruction
-which has that type of cost, will throw an AVM exception.
-
-\note{When a transaction reaches its \texttt{maxExternalCost}, it can still execute
-its own code, while it can not call other smart contracts.
-This way the execution cost of a smart contract is completely decoupled from the
-smart contract it calls, and a malicious contract can not make its invoker certainly fail
-by using infinite loops.}
-
-Also, Argennon transactions are required to specify what heap or code area addresses they will access. This will
-enable validators to parallelize transaction validation as we will see in Section~\ref{sec:concurrency}. An instruction
-that tries to access a memory location that is not in the access list of the transaction, will throw an exception.
-Users could use off-chain \emph{execution oracles} to predict the list of memory locations their transactions need.
-
-An execution oracle is a full AVM emulator that keeps a full local copy of the AVM storage and can emulate AVM
-execution without accessing a ZK-EDB server. Execution oracles can be used for reporting useful information about
-Argennon transactions such as accessed AVM heap or code area locations, exact amount of execution cost,
-and so on.
-
 Every Argennon transaction is required to provide the following information as an upper bound for the
@@
 \begin{itemize}
-    \item Maximum internal execution cost
-    \item Maximum external execution cost
-    \item A list of heap/code-area locations for reading
-    \item A list of heap locations for writing
-    \item A list of heap chunks it will deallocate (if any)
-    \item A list of methods it will delete (if any)
-    \item Number and size of heap chunks it will allocate (if any)
-    \item Number and size of method bytecodes it will allocate (if any)
+    \item Maximum AscEE clocks
+    \item The list of applications the request will call
+    \item The list of access blocks the request needs
+    \item \texttt{maxSize} for chunks it wants to expand
+    \item \texttt{minSize} for chunks it wants to shrink
+    \item A list of applications it will update (if any)
 \end{itemize}
@@
 ---
-fee: |
-    PUT /balances/0x73.0xa2?to=fee&amount=0.26&sig=5b73CbmwQNRC7fWUY15 HTTP/1.1
-call: |
-    POST bttp://dapp.argennon.net/54.189.21/proposals HTTP/1.1
+request: |
+    PATCH /balances/0x95ab HTTP/1.1
     Content-Type: application/json; charset=utf-8
-    Content-Length: 77
+    Content-Length: 46
 
-    {
-        "name": "Grant",
-        "recipient": "0x24.0x8f.0x29.0xa1",
-        "amount": 25000,
-        "sig"= "2a36Gtrw249wQCD70nWY49d"
-    }
+    {"to":0xaabc,"amount":1399,"sig":0}
+
+messages:
+   - issuer: 0x95ab000000000000
+     msg: {"to":0xaabc000000000000,"amount":1399,"forApp":0x100000000000000,"nonce":11}
+     sig: LNUC49Lhyz702uszzNcfaU3BhPIbdaSgzqDUKzbJzLPTlFS2J9GzHl-cDKb
+
 caps:
-    internal: 2500000 # maximum number of AVM execution clocks
-    external: 1000000
+    maxClocks: 150 # maximum number of AVM execution clocks
+    apps: [1,124.16]
     read: [(2654,3),(15642,0),(15642,1),(15642,3)]
@@
 \end{lstlisting}
-
-\subsection{Authorization}\label{subsec:txn-auth}
-
-Argennon transactions do not have a sender. The authorization of the requested operation is always done by checking the
-digital signatures that are provided as a part of the HTTP request to the \texttt{dispatcher} method.
-
-While every block of the Argennon blockchain stores the commitment of the transaction list, Argennon does not enforce
-storage of the transaction history. To be able to detect replay attacks, we require
-every signature that a user creates to have a nonce. This nonce consists of the issuance round of the signature
-and a sequence number: \texttt{(issuance,\ sequence)}. When a user creates more than one signature in a round, he
-must sequence his signatures starting from 0 (i.e.~the sequence number restarts from 0 in every round). We define
-a maximum lifetime for signatures, so a signature is invalid if \texttt{currentRound - issuance > maxLifeTime} or
-if a signature of the same user with a bigger or equal nonce is already used
-(i.e.~is recorded in the blockchain). A nonce is bigger than another nonce if it has an older issuance. If two
-nonces have an equal issuance, the nonce with the bigger sequence number will be considered bigger.
-
-To be able to detect invalid signatures, we keep the maximum nonce of used digital signatures per user. This information
-is stored in the ARG smart contract and when the difference between \texttt{issuance} component of the nonce and
-the current round becomes bigger than the maximum allowed lifetime of a signature, it can be safely
-deleted.\footnote{in some conventional
-blockchains, the nonce data can never be deleted, even if the account has zero balance and is no longer used.}
-
-\subsection{Transaction Fee}\label{subsec:fee}
-
-Every Argennon transaction is required to pay two types of fees: execution fee, which is paid for executing the
-transaction, and storage fee, which is paid for the amount of storage the transaction allocates.
-
-A transaction pays
-its fees by providing digital signatures of one or more accounts, authorizing the transfer of the amount of fee in
-ARGs from one or more accounts to the fee sink accounts. This fee is transferred by the first
-\texttt{i\_invoke\_dispatcher} instruction of the transaction.
-
-\note{An Argennon transaction always pays all of its proposed fee, no matter how much of its predefined resources
-were not used during the final emulation. This will incentivize users to report the resource usage of their
-transactions more accurately.}

### 2022-04-16 tex/sections/incentive.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Remove outdated parts
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 %! Author = aybehrouz
+
+\subsection{Fees}\label{subsec:fees}
+
+The Argennon protocol does not explicitly define any fees for normal transactions. Only for high priority
+transactions a fixed fee is determined by the governance system (See Section ...). Because the protection of the
+Argennon network against spams and DOS attacks is mostly done by the delegates, they are also responsible for
+collecting transaction fees. A good fee collection policy could considerably increase the chance of delegates for
+being reelected in the
+next round, therefore they are incentivized to use creative and effective methods\footnote{For example, they may
+allow a limited number of free transactions per month for every account.}.
+
+By the concept of request attachments, the Argennon protocol provides a method
+for trust-less fee payment. A user can simply define the transaction that he wants to pay fees for, as the attachment
+of the fee payment transaction. That way the fee payment transaction will be performed only if the attached
+transaction is also included in the same block. However, it should be noted that trust-less fee payment is not
+enforced by the protocol.
+
+While transaction fee is not enforced by the Argennon protocol another type of fee is mandatory: the
+\emph{storage fee}. The Argennon protocol requires that a storage fee is paid for every block of the Argennon
+blockchain. This fee must be paid by the delegates, and its amount depends on the storage overhead that the new block
+is imposing on the Argennon state. The exact amount of this fee is determined by the ADAGs.
 
@@
 
-\subsection{Incentives for ZK-EDB Servers}\label{subsec:zk-edb-servers}
+\subsection{Incentives for PV-DB Servers}\label{subsec:PV-DB-servers}
 
-The incentive mechanism for ZK-EDB servers should have the following properties:
+The incentive mechanism for PV-DB servers should have the following properties:
 
@@
     \item It incentivizes storing all storage pages and not only those pages that are used more frequently.
-    \item It incentivizes ZK-EDB servers to actively provide the required storage pages for validators.
-    \item Making more accounts will not provide any advantage for a ZK-EDB server.
+    \item It incentivizes PV-DB servers to actively provide the required storage pages for validators.
+    \item Making more accounts will not provide any advantage for a PV-DB server.
 \end{itemize}
 
-For our incentive mechanism, we require that every time a validator receives a storage page from a ZK-EDB, after
-validating the data, he give a receipt to the ZK-EDB server. In this receipt the validator signs the
+For our incentive mechanism, we require that every time a validator receives a storage page from a PV-DB, after
+validating the data, he give a receipt to the PV-DB server. In this receipt the validator signs the
 following information:
@@
 \begin{itemize}
-    \item \texttt{ownerAddr}: the account address of the ZK-EDB server.
+    \item \texttt{ownerAddr}: the account address of the PV-DB server.
     \item \texttt{receivedPageID}: the ID of the received page.
@@
 
-\note{In a round, an honest validator never gives a receipt for an identical page to two different ZK-EDB servers.}
+\note{In a round, an honest validator never gives a receipt for an identical page to two different PV-DB servers.}
 
-To incentivize ZK-EDB servers, a lottery will be held every round,\footnote{A round is the time interval between
+To incentivize PV-DB servers, a lottery will be held every round,\footnote{A round is the time interval between
 two consecutive blocks.} and a predefined amount of ARGs from
@@
 
-\note{One ZK-EDB server could own multiple winning tickets in a round.}
+\note{One PV-DB server could own multiple winning tickets in a round.}
 
@@
     \item The receipt round number is $r - 1$.
-    \item The signer did not sign a receipt for the same storage page for two different ZK-EDB servers in
+    \item The signer did not sign a receipt for the same storage page for two different PV-DB servers in
     the previous round.
@@
     \item \texttt{seed} is the current block seed.
-    \item \texttt{|} is a concatenation operator.
+    \item \texttt{|} is the concatenation operator.
 \end{itemize}
@@
 Also, based on the current block seed, a random storage page is
-selected as the challenge of the round. A ZK-EDB server that owns a winning receipt needs to broadcast a \emph{winning
+selected as the challenge of the round. A PV-DB server that owns a winning receipt needs to broadcast a \emph{winning
 ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
 challenge. Solving a round challenge requires the content of the storage page which was selected as the round
-challenge. This will encourage ZK-EDB servers to store all storage pages.
+challenge. This will encourage PV-DB servers to store all storage pages.
 
@@
 The winning tickets of the lottery of round $r$ need to be included in the block of the round $r$,
-otherwise they will be considered expired. Validation and prize distribution for the winning tickets of round
-$r$ will be done in the round $r + 1$. This way, \textbf{the content of the challenge page could be
+otherwise they will be considered expired. However, finalizing and prize distribution for the winning tickets
+should be done in a later round. This way, \textbf{the content of the challenge page could be
 kept secret during the lottery round.} Every winning ticket will get an equal share of the lottery prize.

### 2022-04-17 tex/sections/incentive.tex
Add block fee
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 Argennon network against spams and DOS attacks is mostly done by the delegates, they are also responsible for
-collecting transaction fees. A good fee collection policy could considerably increase the chance of delegates for
-being reelected in the
-next round, therefore they are incentivized to use creative and effective methods\footnote{For example, they may
+determining and collecting transaction fees. A good fee collection policy could considerably increase the chance of
+delegates for being reelected in the next round, therefore they are incentivized to use creative and effective
+methods\footnote{For example, they may
 allow a limited number of free transactions per month for every account.}.
@@
 By the concept of request attachments, the Argennon protocol provides a method
-for trust-less fee payment. A user can simply define the transaction that he wants to pay fees for, as the attachment
+for trust-less fee payment. When a user does not want to use off-chain fee payment methods, he can simply define his
+transaction as the attachment
 of the fee payment transaction. That way the fee payment transaction will be performed only if the attached
-transaction is also included in the same block. However, it should be noted that trust-less fee payment is not
-enforced by the protocol.
+transaction is also included in the same block.
 
-While transaction fee is not enforced by the Argennon protocol another type of fee is mandatory: the
-\emph{storage fee}. The Argennon protocol requires that a storage fee is paid for every block of the Argennon
-blockchain. This fee must be paid by the delegates, and its amount depends on the storage overhead that the new block
-is imposing on the Argennon state. The exact amount of this fee is determined by the ADAGs.
+While transaction fee is not enforced by the Argennon protocol, there are other types of fee that are mandatory: the
+\emph{database fee} and the \emph{block fee}. Both of these fees are required to be paid for every block of the Argennon
+blockchain and are paid by the delegates. The block fee is a constant fee that is paid for each new block of the
+blockchain and its amount is
+determined by the ADAGs. The database fee depends on the data access and storage overhead that a new block
+is imposing on the Argennon storage cloud. The amount of this fee is determined by the ADAGs, and is collected inside
+a special account: the \texttt{dbFeeSink}.
 
@@
 
-When the certificate of a committee of validators is included in a block, the signers of that certificate and the
-delegates will be rewarded. Every validator will be rewarded equally. The delegates will be rewarded
-proportional to the total stake of the signers of the certificate and their reward increases if the
-certificate has more signers.
+The validators who sign the certificate of a block will receive the block fee paid for that block. Every validator
+will be rewarded
+proportional to his stake (i.e voting power). As we mentioned before the block fee is a constant fee which the
+delegates pay for each block.
 
 Rewards will not be distributed instantly, instead they will be distributed at the end of the staking period.
-This will facilitate efficient implementations which avoid frequent updates in the AVM storage.
-Rewards of every staking period depends on the amount of fees that are collected during that period.
+This will facilitate efficient implementations which avoid frequent updates in the Argennon storage.
 

### 2022-04-18 tex/sections/incentive.tex
Add the delegates reward
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 determining and collecting transaction fees. A good fee collection policy could considerably increase the chance of
-delegates for being reelected in the next round, therefore they are incentivized to use creative and effective
+delegates for being reelected in the next terms, therefore they are incentivized to use creative and effective
 methods\footnote{For example, they may
@@
 
-By the concept of request attachments, the Argennon protocol provides a method
-for trust-less fee payment. When a user does not want to use off-chain fee payment methods, he can simply define his
-transaction as the attachment
-of the fee payment transaction. That way the fee payment transaction will be performed only if the attached
+In Argennon fee payment can be done off-chain or on-chain. Off-chain fee payment is more efficient and flexible but
+requires some level of trust in the delegates. For trust-less fee payment, the Argennon protocol provides the
+concept of request attachments (See Section~\ref{sec:attachments}).
+When a user does not want to use off-chain fee payment methods, he can simply define his transaction as the attachment
+of the fee payment transaction. That way, the fee payment transaction will be performed only if the attached
 transaction is also included in the same block.
@@
 determined by the ADAGs. The database fee depends on the data access and storage overhead that a new block
-is imposing on the Argennon storage cloud. The amount of this fee is determined by the ADAGs, and is collected inside
+is imposing on the Argennon storage cloud. The amount of this fee is determined by the ADAGs, and is collected in
 a special account: the \texttt{dbFeeSink}.
@@
 This will facilitate efficient implementations which avoid frequent updates in the Argennon storage.
+
+As long as ARG is allowed to be minted and its cap is not reached, the delegates will receive a reward at the
+\textbf{end} of their election term. This reward will consist of newly minted ARGs, and its amount will be
+determined by the ADAGs. In addition, for each block certificate that is added to the Argennon blockchain some amount
+of ARGs will be minted and added to the \texttt{dbFeeSink} account.
 

### 2022-04-20 tex/sections/avm.tex
Remove outdated parts
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Clarify resource management
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 a session violates its resource caps, determining the point of failure may require precise and error-free resource
-measurement. Note that for implementing optional decoupling principle, we need to determine the exact application call
-which has failed in a call chain. For example, assume that a session containing an application call
+measurement. Note that for implementing optional decoupling principle in Argennon, we need to determine the exact
+application call which has failed in a call chain. For example, assume that a session containing an application call
 violates a 2 milliseconds execution time cap. If in the caller's code, the call happens exactly after 2 milliseconds,
@@
 monitored session, resources that can not be measured precisely are measured per application call and their caps are
-determined per application call by the external HTTP request (i.e.\ transaction). An optimistic session must
-always complete normally, a monitored session is allowed to either complete normally or abruptly. In Argennon, the
+determined per application call by the external HTTP request (i.e.\ transaction). In Argennon, the
 type of execution session for each transaction is decided by the block proposer.
@@
     application it has to \emph{forward} a portion of its \textbf{external} clocks to the called application. This
-    amount will determine the value of \texttt{maxClocks} for the called application.
+    amount will determine the value of \texttt{maxClocks} for the called application, and is subtracted from the value
+    of \texttt{externalClocks} of the caller.
 
@@
 resource is a resource that can not be measured precisely and its measurement always contains a random error.
+Optimistic sessions are not allowed to fail because of violating nondeterministic resource limits. As a result, the
+block proposer must always choose a monitored session for a transaction that violates a nondeterministic resource cap
+and is included in his block.
 

### 2022-04-20 tex/A.tex
Clarify resource management
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
         contract platform which tries to solve shortcomings of existing platforms. The Argennon blockchain uses a hybrid
-        proof of stake (POS) consensus protocol, which is based on participation of every Argennon user.
+        proof of stake (HPoS) consensus protocol, which is based on participation of every Argennon user. This hybrid
+        protocol is able to combine the benefits of a centralized and a decentralized system.
 
-        Thanks to the
-        cloud based design of the Argennon blockchain, transaction validation does not require costly computational
-        resources, and normal personal computers with limited hardware capabilities can actively participate in the
-        Argennon consensus protocol. This property makes Argennon a truly decentralized and democratic blockchain and
-        one of the most secure existing platforms.
+        Thanks to the cloud based design of the Argennon blockchain, transaction validation does not require costly
+        computational resources, and normal personal computers with limited hardware capabilities can actively
+        participate in the Argennon consensus protocol. This property makes Argennon a truly decentralized and
+        democratic blockchain and one of the most secure existing platforms.
 
@@
         which forms the Argennon cloud. A PV-DB server is a conventional data server which uses its computational and
-        storage resources to help the
-        Argennon network process transactions. A large portion of incentive rewards in the Argennon platform is
-        devoted to PV-DB servers. This will incentivize the development of conventional networking, storage and
-        computational hardware, which can benefit all areas of information technology.
+        storage resources to help the Argennon network process transactions. A large portion of incentive rewards in
+        the Argennon platform is devoted to PV-DB servers. This will incentivize the development of conventional
+        networking, storage and computational hardware, which can benefit all areas of information technology.
     \end{abstract}

### 2022-04-20 tex/sections/avm.tex
Clarify resource management
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 monitored session, resources that can not be measured precisely are measured per application call and their caps are
-determined per application call by the external HTTP request (i.e.\ transaction). In Argennon, the
-type of execution session for each transaction is decided by the block proposer.
+determined per application call by the external HTTP request (i.e.\ transaction). The block proposer decides the
+type of the execution session for each transaction.
+
+In most consensus protocols, we can assume that the block proposer has enough incentive to filter out transactions
+that spam run-time resources. Here by a run-time resource, we refer to a resource that at run-time, a limited amount
+of it is available, but its surplus can not be stored for later use. Execution time and local memory are
+examples of a run-time resource but permanent storage is not a run-time resource.
+
+If a proposed block contains many transactions which need a lot of run-time resources, validators would not
+be able to validate all transactions in a timely manner. Consequently, they may decide to reject the block or if they
+spend enough resources, the confirmation of that block could take more time than usual. Longer block time is not
+favoured by block proposers, because it means less throughput of the system which usually means less overall rewards
+for them. In Argennon the management of run-time resources, is left to the block proposer.
 

### 2022-04-22 tex/A.tex
Fix some readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-        The Argennon network relies on a permission-less network of Publicly Verifiable Database (PV-DB) servers,
-        which forms the Argennon cloud. A PV-DB server is a conventional data server which uses its computational and
+        The Argennon cloud is formed by a permission-less network of Publicly Verifiable Database (PV-DB) servers. A
+        PV-DB server in Argennon, is a conventional data server which uses its computational and
         storage resources to help the Argennon network process transactions. A large portion of incentive rewards in
-        the Argennon platform is devoted to PV-DB servers. This will incentivize the development of conventional
+        the Argennon platform is devoted to PV-DB servers. This incentivizes the development of conventional
         networking, storage and computational hardware, which can benefit all areas of information technology.

### 2022-04-22 tex/sections/avm.tex
Fix some readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 for the architecture and requirements of the Argennon platform.
-Argennon supports two standard representations: one is a high level text based language which needs
-costly compilation before being executed on a hardware machine. The other is a low level binary representation which
+Argennon supports two standard representations. One representation is a high level text based language which needs
+costly compilation before being executed on a hardware machine. The other, is a low level binary representation which
 usually can be executed, with minimal pre-processing, by a JIT compiler or an emulator. The high level
@@
 This step is
-designed in a way that can be done fully in parallel for each request without any risks of data races. The second
+designed in a way that can be done fully in parallel for each request without any risk of data races. The second
 step is the \emph{Data Dependency Analysis (DDA) step}.
@@
 a session violates its resource caps, determining the point of failure may require precise and error-free resource
-measurement. Note that for implementing optional decoupling principle in Argennon, we need to determine the exact
-application call which has failed in a call chain. For example, assume that a session containing an application call
+measurement. For example, assume that a session containing an application call
 violates a 2 milliseconds execution time cap. If in the caller's code, the call happens exactly after 2 milliseconds,
 a small fluctuation in the execution time measurement can
-change the point of failure between the called and the caller application. This can introduce nondeterministic
+change the point of failure between the called and the caller application. Note that for implementing optional
+decoupling principle in Argennon, we need to determine the
+exact application call which has failed in a call chain, and this fluctuation introduces a nondeterministic
 behaviour which could make block validation impossible. That's why the AscEE sometimes uses per application call
-resource management. If we define the resource caps per application call and perform our measurements for each
-application call separately, by using caps that are larger than the measurement error, errors in the
-measurements would not change the point of failure in the call chain.
+resource management. If we define the resource caps per application call and
+perform our measurements for each application call separately, by using caps that are larger than the measurement
+error, errors in the measurements would not change the point of failure in the call chain.
 
@@
 be able to validate all transactions in a timely manner. Consequently, they may decide to reject the block or if they
-spend enough resources, the confirmation of that block could take more time than usual. Longer block time is not
+spend enough resources, the confirmation of that block could take more than usual time. Longer block time is not
 favoured by block proposers, because it means less throughput of the system which usually means less overall rewards
-for them. In Argennon the management of run-time resources, is left to the block proposer.
+for them. In the Argennon protocol the management of run-time resources, is left to the block proposer.
 
@@
 Optimistic sessions are not allowed to fail because of violating nondeterministic resource limits. As a result, the
-block proposer must always choose a monitored session for a transaction that violates a nondeterministic resource cap
-and is included in his block.
+block proposer must always choose a monitored session for a transaction that is included in his block and violates a
+nondeterministic resource cap.
 
@@
 executing a monitored session, for each application call, they enforce considerably larger limits for
-nondeterministic resources, only when the proposer has declared that an application call violates a limit the
+nondeterministic resources, only when the proposer has declared that an application call violates a limit, the
 validators will enforce the actual value of the limit. This simple mechanism ensures that, with a very high

### 2022-07-01 tex/sections/avm.tex
Fix some readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 Environment (AscEE) is an abstract execution environment for executing Argennon smarts contracts (a.k.a\ Argennon
-applications). An Argennon application essentially is an HTTP server whose state is kept in the Argennon blockchain and
-its logic is described using an Argennon Standard Representation (ASR).
+applications) in a safe and secure environment. An Argennon application essentially is an HTTP server whose state is
+kept in the Argennon blockchain and its logic is described using an Argennon Standard Representation (ASR).
 
 An Argennon Standard Representation (ASR) is a programming language for describing argennon applications, optimized
-for the architecture and requirements of the Argennon platform.
-Argennon supports two standard representations. One representation is a high level text based language which needs
+for the architecture and properties of the Argennon platform.
+Argennon supports two standard representations: One is a high level text based representation which needs
 costly compilation before being executed on a hardware machine. The other, is a low level binary representation which
@@
 platform specific compiler optimization at host nodes. The low level language, on the oder hand, is designed for
-efficient direct execution of applications that are not frequently used.
+efficient direct execution of applications that are not frequently used and do not benefit from compiler optimizations.
 
@@
 considered sequential\footnote{Actually requests are executed in parallel but by performing data dependency analysis the
-result is guaranteed to be identical
-with sequential execution of requests.} and each request has a separate \emph{execution session}. An execution session
-is a separate session of executing smart contract's code in order to fulfill an \emph{external} HTTP request.
+result is guaranteed to be identical with sequential execution of requests.} and each request has a separate
+\emph{execution session}. Therefore, an execution
+session is a separate session of executing smart contract's code in order to fulfill an \emph{external} HTTP request.
 External requests are requests that are not made by other Argennon applications.
@@
 application, optionally it can choose to be decoupled from the called application. That would mean the called
-application could not affect its caller's state by reentrancy, or could not fail the session by using excessive
+application could not affect its caller's state by reentrancy, or could not abort the session by using excessive
 resources or performing illegal operations.
@@
 
-An Argennon identifier may be simple or compound. A simple identifier is generated using a single trie, while a
-compound identifier is generated by concatenating prefix codes generated by two or more tries:
+An Argennon identifier may be simple or compound. A simple identifier is generated using a single tree, while a
+compound identifier is generated by concatenating prefix codes generated by two or more trees:
 
@@
 
-All Argennon prefix trees have an equal branching factor \(\beta\). Therefore, we can represent an Argennon
+All Argennon prefix trees have an equal branching factor \(\beta\)\footnote{A typical choice for \(\beta\) is \(2^8\)
+    .}, and
+we can represent an Argennon
 prefix tree as a sequence of fractional numbers\footnote{It's possible to have \(a_i=0\).
@@
 \]
-where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)} \leq A^{(i+1)}\).\footnote{A typical choice for
-    \(\beta\) is \(2^8\).}
+where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)} \leq A^{(i+1)}\).
 
@@
 
-The persistent data of an Argennon application is stored in heap chunks. An application may have several heap chunks
+The persistent data of an Argennon application is stored in continuous finite
+arrays of bytes, called heap chunks. An application may have several heap chunks
 with different sizes, and can remove or
@@
 
-A heap chunk can be considered as a continuous
-array of bytes. Every chunk has a size: \texttt{chunkSize} and a size upper bound: \texttt{sizeUpperBound}. The value of
+Every heap chunk has a size: \texttt{chunkSize} and a size upper bound: \texttt{sizeUpperBound}. The value of
 \texttt{chunkSize} can be determined uniquely at
 the start of
-every execution session, and it may be updated during the session like a normal memory location. On the other hand,
-the value of \texttt{sizeUpperBound} is constant for every block of the blockchain and is proposed by the block
-proposer. \texttt{sizeUpperBound} indicates the upper bound of \texttt{chunkSize} value for a chunk in a block.
-A validator can use this value to safely allocate memory for a chunk.
+every execution session, and it may be updated by the owner application during the session like a normal memory
+location. On the other hand,
+the value of \texttt{sizeUpperBound} is constant for every block of the Argennon blockchain and is proposed by the block
+proposer. This value indicates the upper bound of \texttt{chunkSize} for a specific chunk during the execution of all
+requests of a block, and it can be used by validators for memory allocation of chunks during block validation.
 
@@
 
-Memory locations inside a chunk can only be accessed through access blocks. An access block is defined on a chunk
-and has an offset and a size and determines accessible memory locations inside the chunk. Multiple access blocks can
-be defined on a single chunk, but they must be non overlapping. Access blocks are byte addressable and can have
-different access types:
+Memory locations inside a chunk can only be accessed through access blocks. An access block is an access interval
+defined on a chunk. It has an offset and a size and determines the accessible memory locations inside the chunk.
+Multiple access blocks can be defined on a single chunk, but they must be non overlapping.
+Access blocks are byte addressable and can have different access types:
 
 \begin{itemize}
+    \item \texttt{check\_only}: only allows check operations. These operations query the persistence
+    status of a memory location.
     \item \texttt{read\_only}: only allows read and check operations.
     \item \texttt{writable}: allows reading and writing.
-    \item \texttt{check\_only}: only allows check operations. These operations query the persistence
-    status of a memory location.
-    \item \texttt{additive}: only allows addition-like operations without overflow checking. Note that the content of
-    these access blocks cannot be read.
+    \item \texttt{additive}: only allows additive operations. These operations are different types of mathematical
+    addition without any overflow checking. Note that the content of these access blocks cannot be read.
 \end{itemize}
@@
 the Argennon cloud uses cryptographic accumulators for preserving data integrity.} and access blocks are
-required for better parallelization of the request execution. An application should put the data it predicts is
-needed for validating a block in the same chunk and the data it predicts is needed in a single execution
-session in the same access block.
+required for better parallelization of the request execution. An application should put the data needed for
+validating a block in the same chunk and the data needed in a single execution session in the same access block.
 
@@
 The AscEE uses \emph{Authenticated Message Passing} for authorizing operations. In this method, every execution
-session has a set of authenticated messages, and applications \textbf{explicitly} pass those messages for authorizing
-operations.
-Authentication of messages can be efficiently done in parallel, using any type of cryptographic signature scheme in
-the preprocessing phase of request execution. For example, we could use BLS aggregate signature to authenticate all
-messages of a block in bulk.
+session has a set of authenticated messages, and those messages are \textbf{explicitly} passed in requests to
+applications for authorizing operations. These messages act exactly like digital signatures and applications can
+ensure that they are issued by specific accounts. The only difference is that the process of message authentication
+is performed by the AscEE internally and is not performed by the application. For accounts, the AscEE uses digital
+signatures to authenticate messages. Signatures for each request can be validated in parallel during the
+preprocessing phase of the request execution, and any type of
+cryptographic signature scheme can be used. For example, we could use BLS aggregate signature to authenticate all
+messages of a block of the blockchain in bulk.
 
-Moreover, applications use built-in functions of AscEE to generate authenticated messages in run-time. This enables
-an application to authorize operations for another application even if that application is not directly called by
-itself.
+Moreover, applications can use built-in functions of the AscEE to generate authenticated messages in run-time.
+This enables an application to authorize operations for another application even if it is not calling that
+application directly.
 
@@
 
+In most consensus protocols, we can assume that the block proposer has enough incentive to filter out transactions
+that spam run-time resources. Here by a run-time resource, we mean a resource that at run-time, a limited amount
+of it is available, but its surplus can not be stored for later use. Execution time and local memory are
+examples of such a resource but permanent storage is not.
+
+If a proposed block contains many transactions which need a lot of run-time resources, validators would not
+be able to validate all transactions in a timely manner. Consequently, they may decide to reject the block or if they
+spend enough resources, the confirmation of that block could take more than usual. Longer block time is not
+favoured by block proposers, because it means less throughput of the system which usually means less overall rewards
+for them. In the Argennon protocol the management of run-time resources, is left to the block proposer.
+
 Resource usage can be measured per session or per application call. Obviously per session measurement is
@@
 exact application call which has failed in a call chain, and this fluctuation introduces a nondeterministic
-behaviour which could make block validation impossible. That's why the AscEE sometimes uses per application call
-resource management. If we define the resource caps per application call and
+behaviour which could make block validation impossible. That's why we need per application call
+resource management sometimes. If we define the resource caps per application call and
 perform our measurements for each application call separately, by using caps that are larger than the measurement
@@
 The AscEE has two type of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
-session is always measured per session and default pre-defined resource caps are used. On the other hand, for a
+session is always measured per session and \textbf{default} pre-defined resource caps are used. On the other hand, for a
 monitored session, resources that can not be measured precisely are measured per application call and their caps are
@@
 type of the execution session for each transaction.
-
-In most consensus protocols, we can assume that the block proposer has enough incentive to filter out transactions
-that spam run-time resources. Here by a run-time resource, we refer to a resource that at run-time, a limited amount
-of it is available, but its surplus can not be stored for later use. Execution time and local memory are
-examples of a run-time resource but permanent storage is not a run-time resource.
-
-If a proposed block contains many transactions which need a lot of run-time resources, validators would not
-be able to validate all transactions in a timely manner. Consequently, they may decide to reject the block or if they
-spend enough resources, the confirmation of that block could take more than usual time. Longer block time is not
-favoured by block proposers, because it means less throughput of the system which usually means less overall rewards
-for them. In the Argennon protocol the management of run-time resources, is left to the block proposer.
 
@@
     of \textbf{cpu time} needed for executing a predefined standard application which is used for benchmarking a host's
-    performance. In other words, by definition the AscEE imaginary standard machine completes the standard benchmark
+    performance. In other words, by definition, the AscEE imaginary standard machine completes the standard benchmark
     in 1000 clocks.
@@
 Optimistic sessions are not allowed to fail because of violating nondeterministic resource limits. As a result, the
-block proposer must always choose a monitored session for a transaction that is included in his block and violates a
+block proposer must always choose a monitored session for a transaction that is included in the block and violates a
 nondeterministic resource cap.
 
-When a transaction fails due to violation of a limit for a nondeterministic resource, the proposer is
-required to exactly specify the application call in the session that violates that limit. When validators are
+When a transaction fails due to the violation of a limit for a nondeterministic resource, the proposer is
+required to exactly specify the application call which violates that limit in the execution session. When validators are
 executing a monitored session, for each application call, they enforce considerably larger limits for
-nondeterministic resources, only when the proposer has declared that an application call violates a limit, the
+nondeterministic resources. Only in case the proposer has declared that an application call violates a limit, the
 validators will enforce the actual value of the limit. This simple mechanism ensures that, with a very high
-probability, validators agree with the proposer.
+probability, validators agree with the proposer, although their measurements could be different.
 
@@
 
-An application can invoke methods of the ArgC Standard Library (ASL) in its own context. Methods of the ArgC standard
-library are stored as a special application in the Argennon blockchain.
-a part of the root smart contract.
 In Argennon, some applications (smart contracts) are updatable. The ArgC Standard Library is an updatable smart

### 2022-07-06 tex/A.tex
Update abstract
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-\title{Argennon: A Scalable Smart Contract Platform}
+\title{Argennon: A Scalable Cloud Based Smart Contract Platform}
 \author{aybehrouz}
@@
     \begin{abstract}
-        Argennon is a cloud based blockchain and smart
-        contract platform which tries to solve shortcomings of existing platforms. The Argennon blockchain uses a hybrid
-        proof of stake (HPoS) consensus protocol, which is based on participation of every Argennon user. This hybrid
-        protocol is able to combine the benefits of a centralized and a decentralized system.
-
-        Thanks to the cloud based design of the Argennon blockchain, transaction validation does not require costly
-        computational resources, and normal personal computers with limited hardware capabilities can actively
+        Argennon is a next generation cloud based blockchain and smart
+        contract platform. The Argennon blockchain uses
+        a hybrid proof of stake (HPoS) consensus protocol, which is capable of combining the benefits of
+        a centralized and a decentralized system. Thanks to the cloud
+        based design of the Argennon blockchain, transaction validation does not require costly
+        computational resources, and normal personal computers or smartphones, with limited hardware capabilities, can
+        validate transactions and actively
         participate in the Argennon consensus protocol. This property makes Argennon a truly decentralized and
         democratic blockchain and one of the most secure existing platforms.
-
-        By using cryptographic accumulators, the integrity of data on the Argennon cloud is guaranteed, and there
-        is no need for trusting cloud servers. At the same time, by using a smart clustering algorithm the network usage
-        and the overhead of the commitment scheme is kept manageable.
 
@@
         networking, storage and computational hardware, which can benefit all areas of information technology.
+
+        By using cryptographic accumulators, the integrity of data on the Argennon cloud is guaranteed, and there
+        is no need for trusting cloud servers. At the same time, by using a smart clustering algorithm the network usage
+        and the overhead of the cryptographic commitment scheme is kept manageable.
     \end{abstract}

### 2022-07-06 tex/sections/avm.tex
Update abstract
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 Argennon supports two standard representations: One is a high level text based representation which needs
-costly compilation before being executed on a hardware machine. The other, is a low level binary representation which
+costly compilation before being executed on a hardware machine. The other is a low level binary representation which
 usually can be executed, with minimal pre-processing, by a JIT compiler or an emulator. The high level
-language is intended for preserving the high level information of applications logic to facilitate
-platform specific compiler optimization at host nodes. The low level language, on the oder hand, is designed for
+language is intended for preserving the high level information of the application logic to facilitate
+platform specific compiler optimization at a host machine. The low level language, on the oder hand, is designed for
 efficient direct execution of applications that are not frequently used and do not benefit from compiler optimizations.
@@
 blockchain in a three-step procedure. The first step is the \emph{preprocessing step}. In this step, the required
-data for executing requests are retrieved from the Argennon cloud and the helper data structures are constructed.
-This step is
+data for executing requests are retrieved from the Argennon cloud, the cryptographic proofs are verified and the
+helper data structures for next steps are constructed. This step is
 designed in a way that can be done fully in parallel for each request without any risk of data races. The second
@@
 and requests that need to be run sequentially. This information is represented using an \emph{execution dag} data
-structure and in the final step, requests are executed using that data structure.
+structure and in the final step, the requests are executed using this data structure.
 
@@
 we can represent an Argennon
-prefix tree as a sequence of fractional numbers\footnote{It's possible to have \(a_i=0\).
-For example, \(A^{(4)}=(0.2000)_{10}\) is correct.} in base \(\beta\):
+prefix tree as a sequence of fractional numbers in base \(\beta\):
 \[
@@
 \]
-where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)} \leq A^{(i+1)}\).
+where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)} \leq A^{(i+1)}\). \footnote{It's possible to
+have \(a_i=0\). For example, \(A^{(4)}=(0.2000)_{10}\) is correct.}
 

### 2022-07-06 tex/sections/concurrency.tex
Update abstract
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 Every block of the Argennon blockchain contains a list of transactions. This list is an ordered list and the
-effect of its contained transactions must be applied to the AVM state sequentially as they appear in the ordered
+effect of its contained transactions must be applied to the AscEE state sequentially as they appear in the ordered
 list. This ordering is solely chosen by the block proposer, and users should not have any assumptions about
@@
 The fact that block transactions constitute a sequential list, does not mean they can not be executed and applied
-to the AVM state concurrently. Many transactions are actually independent and the order of their execution does not
+to the AscEE state concurrently. Many transactions are actually independent and the order of their execution does not
 matter. These transactions can be safely validated in parallel by validators.
 
-A transaction can change the AVM state by modifying either the code area or the AVM heap. In Argennon, all
+A transaction can change the AscEE state by modifying either the code area or the AscEE heap. In Argennon, all
 transactions declare the list of memory locations they want to read or write. This will enable us to determine the
@@
     \item Executing the transactions of \(T\) in any order that \emph{respects} \(G_e\) will result in
-    the same AVM state.
+    the same AscEE state.
     \begin{itemize}
@@
 Having the execution DAG of a set of transactions, using Algorithm~\ref{alg:exec_dag}, we can apply the transaction
-set to the AVM state concurrently, using multiple processor, while we can be sure that the resulted AVM state will
+set to the AscEE state concurrently, using multiple processor, while we can be sure that the resulted AscEE state will
 always be the same no matter how many processor we have used.
@@
     \KwData{The execution dag $\Graph = (\Vertices,E)$ of transaction set \Txns}
-    \KwResult{The state of the AVM after applying \Txns with any ordering respecting \Graph}
+    \KwResult{The state after applying \Txns with any ordering respecting \Graph}
     \BlankLine
@@
 memory locations that \(u\) modifies, we can run \(u\) and \(v\) in parallel. We just need to make sure
-\(u\) does not see any changes \(v\) is making in AVM memory. This can be done by appropriate versioning
+\(u\) does not see any changes \(v\) is making in AscEE memory. This can be done by appropriate versioning
 of the memory locations which is shared between \(u\) and \(v\). We call this method \emph{memory spooling}.
@@
 
-Concurrent counter data structure is a part of the AVM standard library, and any smart contract can use this data
+Concurrent counter data structure is a part of the ArgC standard library, and any smart contract can use this data
 structure for storing the balance record of highly active accounts.

### 2022-07-07 tex/sections/avm.tex
Restructure `Heap Chunks` section
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-Every heap chunk has a size: \texttt{chunkSize} and a size upper bound: \texttt{sizeUpperBound}. The value of
-\texttt{chunkSize} can be determined uniquely at
-the start of
-every execution session, and it may be updated by the owner application during the session like a normal memory
-location. On the other hand,
-the value of \texttt{sizeUpperBound} is constant for every block of the Argennon blockchain and is proposed by the block
-proposer. This value indicates the upper bound of \texttt{chunkSize} for a specific chunk during the execution of all
-requests of a block, and it can be used by validators for memory allocation of chunks during block validation.
-
-The value of \texttt{chunkSize} at the end of the execution session will determine if a memory location at an
-offset is persistent or not: Offsets lower than the chunk size are persistent, and higher offsets are not.
-Non-persistent locations will be re-initialized with zero at the start of every execution session.
-
-The address space of a chunk starts from zero and only offsets lower than \texttt{sizeUpperBound} are valid. Trying to
-access any offset higher than this value will result in a revert for the application.
-
-\subsection{Access Blocks}\label{subsec:access-blocks}
-
-Memory locations inside a chunk can only be accessed through access blocks. An access block is an access interval
-defined on a chunk. It has an offset and a size and determines the accessible memory locations inside the chunk.
-Multiple access blocks can be defined on a single chunk, but they must be non overlapping.
-Access blocks are byte addressable and can have different access types:
-
-\begin{itemize}
-    \item \texttt{check\_only}: only allows check operations. These operations query the persistence
-    status of a memory location.
-    \item \texttt{read\_only}: only allows read and check operations.
-    \item \texttt{writable}: allows reading and writing.
-    \item \texttt{additive}: only allows additive operations. These operations are different types of mathematical
-    addition without any overflow checking. Note that the content of these access blocks cannot be read.
-\end{itemize}
-
-Chunks and access blocks have different purposes. Chunks are intended for simplifying proof checking of the data
-stored in the Argennon cloud\footnote{Note that
-the Argennon cloud uses cryptographic accumulators for preserving data integrity.} and access blocks are
-required for better parallelization of the request execution. An application should put the data needed for
-validating a block in the same chunk and the data needed in a single execution session in the same access block.
-
 \subsection{Chunk Resizing}\label{subsec:ch-resize}
 
-The value of \texttt{chunkSize} can be modified during an execution session. However, the new size values can only be
+At the start of executing requests of a block, a validator can consider two values for every
+heap chunk, the size: \texttt{chunkSize} and a size upper bound: \texttt{sizeUpperBound}. The value of
+\texttt{chunkSize} can be determined uniquely at the start of
+every execution session, and it may be updated during the session by the owner application. On the other hand,
+the value of \texttt{sizeUpperBound} is determined and fixed for each block and is proposed by the block
+proposer. This value indicates the upper bound of \texttt{chunkSize} during the execution of
+requests of the block, and it can be used by the validator for safe memory allocation.
+
+The value of \texttt{chunkSize}, can be modified during an execution session. However, the new size values can only be
 increasing or decreasing. More precisely, if a request declares that it wants to expand (shrink) a chunk, it can only
@@
 (lower) than this value.
+
+The value of \texttt{chunkSize} at the end of an execution session will determine if a memory location at an
+offset is persistent or not: Offsets lower than the chunk size are persistent, and higher offsets are not.
+Non-persistent locations will be re-initialized with zero at the start of every execution session.
 
@@
 considered a good practice. Reading \texttt{chunkSize} decreases transaction parallelization, and should be avoided.
-Instead, applications should use a built-in AscEE's function for checking the persistence status of memory addresses.
+Instead, applications should use a built-in AscEE function for checking the persistence status of memory addresses.
 
@@
 
+The address space of a chunk starts from zero and only offsets lower than \texttt{sizeUpperBound} are valid. Trying to
+access any offset higher than this value will always result in a revert for the application.
+
+\subsection{Access Blocks}\label{subsec:access-blocks}
+
+Memory locations inside a chunk can only be accessed through access blocks. An access block is an access interval
+defined on a chunk, which determines the accessible memory locations inside the chunk. Each access block has an offset
+and a size. Multiple access blocks can be defined on a single chunk, but they must be non overlapping.
+Access blocks are byte addressable and may have different access types:
+
+\begin{itemize}
+    \item \texttt{check\_only}: only allows check operations. These operations query the persistence
+    status of a memory location.
+    \item \texttt{read\_only}: only allows read and check operations.
+    \item \texttt{writable}: allows reading and writing.
+    \item \texttt{additive}: only allows additive operations. These operations are addition-like operations without
+    any overflow checking. These operations must be associative and commutative. Note that the content of
+    these access blocks cannot be read.
+\end{itemize}
+
+Chunks and access blocks have different purposes. Chunks are intended for simplifying proof checking of the data
+stored in the Argennon cloud\footnote{Note that
+the Argennon cloud uses cryptographic accumulators for preserving data integrity.} and access blocks are
+required for better parallelization of the request execution. An application should put the data needed for
+validating a single block in the same chunk and the data needed in a single execution session in the same access block.
+
 \section{Request Attachments}\label{sec:attachments}
@@
 The attachment of a request is a list of request identifiers of the current block that are ``attached'' to the request.
-That means, for validating that request a validator first needs to ``inject'' the digest of attached requests into the
-HTTP request. By doing so, the called application will have access to the digest of attachments in a secure way.
+That means, for validating the request a validator first needs to ``inject'' the digest of attached requests into the
+HTTP request text. By doing so, the called application will have access to the digest of attachments in a secure way.
 
@@
 wants to pay fees. After injecting the digest of those request by validators, that signature can be validated
-correctly and securely by the called application.
+correctly and securely by the application that handles fee payment.
 
@@
 signatures to authenticate messages. Signatures for each request can be validated in parallel during the
-preprocessing phase of the request execution, and any type of
+preprocessing step of the block validation, and any type of
 cryptographic signature scheme can be used. For example, we could use BLS aggregate signature to authenticate all

### 2022-07-24 tex/sections/avm.tex
Fix some whitepaper typos

@@
 
-An Argennon Standard Representation (ASR) is a programming language for describing argennon applications, optimized
+An Argennon Standard Representation (ASR) is a programming language for describing Argennon applications, optimized
 for the architecture and properties of the Argennon platform.
@@
 In this step by analyzing data dependency between requests, the AscEE determines requests that can be run in parallel
-and requests that need to be run sequentially. This information is represented using an \emph{execution dag} data
+and requests that need to be run sequentially. This information is represented using an \emph{execution DAG} data
 structure and in the final step, the requests are executed using this data structure.
@@
 preprocessing step of the block validation, and any type of
-cryptographic signature scheme can be used. For example, we could use BLS aggregate signature to authenticate all
+cryptographic signature scheme can be used. For example, we could use BLS aggregate signatures to authenticate all
 messages of a block of the blockchain in bulk.
@@
 
-The AscEE has two type of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
+The AscEE has two types of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
 session is always measured per session and \textbf{default} pre-defined resource caps are used. On the other hand, for a

### 2022-07-24 tex/sections/consensus.tex
Fix some whitepaper typos

@@
 The credibility of a block of the Argennon blockchain is determined by the certificates it has received
-from different sets of users, knows as committees. There are two primary type of certificate committees in
+from different sets of users, known as committees. There are two primary type of certificate committees in
 Argennon: the committee of \emph{delegates} and the committee
@@
 
-Besides the main delegates' committee, a reserve committee of delegates consists of three members is elected by users
+Besides the main delegates' committee, a reserve committee of delegates consisting of three members is elected by users
 either through the ADAGs or by \emph{emergency agreement} during the recovery protocol. In case the main committee
@@
 
-To prevent blockchain from halting in such situations, the protocol performs a predefined partial
+To prevent the blockchain from halting in such situations, the protocol performs a predefined partial
 reshuffling of committee members.
@@
 of the blockchain to be a special \emph{status block}. A status block is a special block which can only contain
-status change transactions. The status block need to be certified by the delegates and by 2/3 of the total stake of
+status change transactions. The status block needs to be certified by the delegates and by 2/3 of the total stake of
 validators. The \texttt{online}/\texttt{offline} status of validators will not be considered in the validity of
@@
 
-A validator always choose a valid fork block over a block of the main chain. However, as we mentioned before,
+A validator always chooses a valid fork block over a block of the main chain. However, as we mentioned before,
 a validator never signs a certificate for two different blocks with the same height. Consequently, if a validator
@@
 
-Unfortunately, there is a subtle problem with this approach. It is not clear that in a real world economic system
+Unfortunately, there is a subtle problem with this approach. It is not clear in a real world economic system
 how much of the main currency of the system can be locked and kept out of the circulation indefinitely. It seems that
 this amount for currencies like US dollar, is quite low comparing to the total market cap of the currency.
-This means that for a real world currency this type of staking mechanisms will result in putting the
+This means that for a real world currency this type of staking mechanism will result in putting the
 fate of the system in the hands of the owners of a small fraction of the total supply.

### 2022-07-24 tex/sections/incentive.tex

@@
 in a block, the account will be disabled forever in the ARG smart contract. Disabling an account in the
-ARG smart contract will prevent that account from signing any valid signature in the future.
+ARG smart contract will prevent that account from signing any valid signatures in the future.
 
@@
     a fork block and a normal block at the same height usually is a malicious behaviour. However, it will not be
-    penalized because there are circumstances that a honest user could mistakenly do that.}
+    penalized because there are circumstances that an honest user could mistakenly do that.}
 \end{itemize}
@@
 For our incentive mechanism, we require that every time a validator receives a storage page from a PV-DB, after
-validating the data, he give a receipt to the PV-DB server. In this receipt the validator signs the
+validating the data, he gives a receipt to the PV-DB server. In this receipt the validator signs the
 following information:
@@
 two consecutive blocks.} and a predefined amount of ARGs from
-\texttt{dbFeeSink} account will be distributed between winners as a prize. This prize will be divided equally
+\texttt{dbFeeSink} account will be distributed between the winners as a prize. This prize will be divided equally
 between all \emph{winning tickets} of the lottery.

### 2022-07-24 tex/sections/persistence.tex

@@
 
-A page of the AscEE storage should contain objects that have very similar access pattern. Ideally, when a page is needed
+A page of the AscEE storage should contain objects that have very similar access patterns. Ideally, when a page is needed
 for validating a block, almost all of its objects should be needed for either reading or writing. We prefer
@@
 has three dynamic cryptographic accumulators: \emph{staking} database, which stores all the data that is associated with
-the Argennon consensus protocol. \emph{code} database, which stores the AscEE code area, and \emph{heap} database,
+the Argennon consensus protocol, \emph{code} database, which stores the AscEE code area, and \emph{heap} database,
 which stores the AscEE heap. The commitment of these three accumulators are included in every block of the Argennon

### 2022-07-24 tex/A.tex
Fix a minor error
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Update abstract
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
         a hybrid proof of stake (HPoS) consensus protocol, which is capable of combining the benefits of
-        a centralized and a decentralized system. Thanks to the cloud
-        based design of the Argennon blockchain, transaction validation does not require costly
-        computational resources, and normal personal computers or smartphones, with limited hardware capabilities, can
-        validate transactions and actively
+        a centralized and a decentralized system. Argennon validators do not store the blockchain state and
+        transaction validation does not require costly
+        computational resources. In Argennon, normal personal computers or smartphones, with limited hardware
+        capabilities, can validate transactions and actively
         participate in the Argennon consensus protocol. This property makes Argennon a truly decentralized and
@@
         is no need for trusting cloud servers. At the same time, by using a smart clustering algorithm the network usage
-        and the overhead of the cryptographic commitment scheme is kept manageable.
+        and the overhead of the cryptographic proofs is kept manageable.
     \end{abstract}

### 2022-07-31 tex/A.tex
Update abstract
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
         a hybrid proof of stake (HPoS) consensus protocol, which is capable of combining the benefits of
-        a centralized and a decentralized system. Argennon validators do not store the blockchain state and
-        transaction validation does not require costly
-        computational resources. In Argennon, normal personal computers or smartphones, with limited hardware
+        a centralized and a decentralized system. In Argennon, ledger storage and transaction validation are
+        outsourced to the cloud and normal personal computers or smartphones, with limited hardware
         capabilities, can validate transactions and actively
@@
 
-        The Argennon cloud is formed by a permission-less network of Publicly Verifiable Database (PV-DB) servers. A
-        PV-DB server in Argennon, is a conventional data server which uses its computational and
-        storage resources to help the Argennon network process transactions. A large portion of incentive rewards in
-        the Argennon platform is devoted to PV-DB servers. This incentivizes the development of conventional
-        networking, storage and computational hardware, which can benefit all areas of information technology.
+        In Argennon, Computational Integrity (CI) is achieved by using succinct cryptographic proofs (STARK/SNARK)
+        and data integrity is guaranteed by cryptographic accumulators. There is no need to trust cloud servers.
+        At the same time, a smart clustering algorithm keeps
+        the network usage and the overhead of commitment schemes manageable.
 
-        By using cryptographic accumulators, the integrity of data on the Argennon cloud is guaranteed, and there
-        is no need for trusting cloud servers. At the same time, by using a smart clustering algorithm the network usage
-        and the overhead of the cryptographic proofs is kept manageable.
+        The Argennon protocol strongly incentivizes the formation of a permission-less network of Publicly Verifiable
+        Cloud (PVC) servers. A PVC server in Argennon, is a conventional data server which uses its computational and
+        storage resources to help the Argennon network process transactions. This encourages the development
+        of conventional networking, storage and compute hardware, which can benefit all areas of information technology.
+        This contrasts with the approach of some older blockchains that incentivizes the development of a totally
+        useless technology of hash calculation.
+
     \end{abstract}

### 2022-08-02 tex/A.tex
Add a brief introduction
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 \usepackage{amsmath}
+\usepackage{mathtools}
 
@@
         a hybrid proof of stake (HPoS) consensus protocol, which is capable of combining the benefits of
-        a centralized and a decentralized system. In Argennon, ledger storage and transaction validation are
+        a centralized and a decentralized system. In Argennon, ledger storage and transaction processing are
         outsourced to the cloud and normal personal computers or smartphones, with limited hardware
@@
 
-        In Argennon, Computational Integrity (CI) is achieved by using succinct cryptographic proofs (STARK/SNARK)
-        and data integrity is guaranteed by cryptographic accumulators. There is no need to trust cloud servers.
-        At the same time, a smart clustering algorithm keeps
-        the network usage and the overhead of commitment schemes manageable.
+        The Argennon cloud is trustless and publicly verifiable. Computational Integrity (CI) is
+        achieved by using succinct cryptographic proofs (STARK/SNARK)
+        and data integrity is guaranteed by cryptographic accumulators. At the same time, a smart
+        clustering algorithm keeps the bandwidth usage and the overhead of cryptography manageable
+        for validators.
 
@@
         of conventional networking, storage and compute hardware, which can benefit all areas of information technology.
-        This contrasts with the approach of some older blockchains that incentivizes the development of a totally
+        This contrasts with the approach of some older blockchains that incentivized the development of a totally
         useless technology of hash calculation.
@@
     \tableofcontents
+
+    \chapter{Introduction}\label{ch:intro}
+    \input{sections/intro}
 

### 2022-08-02 tex/sections/intro.tex
Add a brief introduction
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

The most conventional use case of a blockchain is financial application. This makes the security of the consensus
protocol of a blockchain crucially important. Unfortunately, many currently used blockchains are vulnerable to a
certain type
of consensus attack, know as the bribery attack. In a bribery attack, an adversary tries to corrupt participants of a
protocol and make them violate the protocol by offering them money.

At the time of writing of this document,
the total mining reward for a bitcoin block is around \$150,000. If we assume that mining rewards define the utility
function of bitcoin miners accurately, one could hire all hashing power of the bitcoin network for one hour by spending
only \$750,000. The situation is not much different for PoS blockchains, as long as the total stake\footnote{Here, by
stake, we mean a real number measuring the total interest of a user in the system. We are not referring, in
particular, to the locked amount of a user's money that is known as stake in some PoS protocols.}
of validators is a relatively small value. This problem is more severe in blockchains that use randomly selected small
sets of validators. Selecting these random sets by hidden random procedures would not help, as long as the
validator himself knows he has been selected, before casting his vote.

It appears that the only solution to this important vulnerability is to effectively participate all the stake of the
system in the consensus protocol. This large participation secures the protocol against bribery or
collusion because the adversary would have to spend unrealistic amount of money to bribe enough users.

However, for effective participation in the consensus protocol, a validator needs to be able to detect
illegal transactions. Detecting illegal transaction can be done by accessing the ledger
state and executing transactions according to the protocol predefined rules. The
ledger state, even for small blockchains, could be several
hundreds of gigabytes, and executing transactions could easily become costly when a blockchain is acting as a
smart contract platform. This computational and storage overhead could prevent most of
ordinary users from any type of participation in the process of securing a blockchain.

Although a fully decentralized blockchain based on the participation of every user looks appealing, it is not as perfect
as it might seem. A blockchain consensus protocol relies on a network of computers, not humans. Ordinary users use
simple and similar computer systems. That means, they all have similar vulnerabilities that could be used by an
adversary to attack the consensus protocol. For example, a malware might be
able to infect a lot of normal computers in the world, and then it might be used to hijack the majority of
participants in the consensus protocol and compromise the security of the blockchain. Securing a computer
system against cyberattacks needs planning ahead and access to engineering resources. Special software and hardware,
like custom-built operating systems and isolated specialized hardware would be required. This is not something a normal
user can do. Only powerful centralized entities that have enough financial and technical resources are able to build
systems that are resilient against sophisticated cyberattacks. So in this regard, a centralized system is arguably
superior to a decentralized \footnote{Note
that decentralized and distributed are two different concepts.} system.

To overcome these difficulties, Argennon uses a hybrid proof of stake consensus protocol. A small committee of
delegates is democratically elected
by the Argennon users and is responsible for minting new blocks of the Argennon blockchain. Each
minted block is required to get approved by its corresponding validator assembly. Assemblies are very large, and they
contain at least three percent of the total stake of the system. A block is approved if it gets approval votes of at
least \(2/3\) of the total stake of its validator assembly.

Each block of the Argennon blockchain contains a set of Computational Integrity (CI) statements and a commitment to
its ledger state. This set defines how
the ledger state of a block can be reached from the state of its previous block via a set of intermediate states.
Each CI statement defines an ordered list of transactions (called external requests in Argennon) and determines
the intermediate states before and after executing those transactions:
\[
    \tau_{(s_i,s_{i+1},\mathbb{R})} \coloneqq \text{``$s_{i+1}$ is the next state after executing external requests
        $\mathbb{R}$ on state $s_i$''}
\]
Validators can\footnote{Using the Argennon cloud is optional for a validator.} receive succinct cryptographic proofs
of these CI statements from the Argennon Publicly Verifiable Cloud. They can also receive the state of the previous
block (the required part of it) and proofs that show the received data is consistent with the previous block commitment.

Verifying a succinct proof can be exponentially faster than replaying the computation. Moreover, verifications of
different statements are independent and can be done in parallel. Therefore, a validator can use multiple cores for
verifying CI statements of a single block and different validator assemblies can simultaneously verify different blocks.
In addition, proof generation of different CI statements can be done in parallel similarly. However, for parallel
proof generation, the state transition needs to be known in advance. That's why in Argennon, delegates do not try
to generate proofs and focus all their computational power on executing requests and generating the state transition
as fast as possible.

Independence of CI statements is useful, but is not enough for having a truly scalable blockchain. To increase
parallelism, the Argennon protocol enforces all external requests to pre-declare their memory access locations. That
would enable a block proposer\footnote{In Argennon, delegates are the only block proposers.} to use Data Dependency
Analysis\footnote{See Section~\ref{sec:concurrency}} to indicate independent sets of transactions and use those sets
for parallel processing of transactions. More importantly, these sets can be used for generating CI statements that
are defined on the same initial state and their proof can be generated independently without the need to calculate
the state transition in advance.

The Argennon protocol strongly incentivizes the formation of a permission-less network of Publicly Verifiable
Cloud (PVC) servers. To do so, the Argennon protocol conducts repetitive automatic lotteries between PVC servers.
A PVC server can increase its chance of winning by (i) generating proofs for more CI statements, (ii) providing the
state data to more validators and, (iii) storing all parts of the state instead of more frequently used parts.

A PVC server in Argennon, is a conventional data server which uses its computational and
storage resources to help the Argennon network process transactions. This encourages the development
of conventional networking, storage and compute hardware, which can benefit all areas of information technology.

### 2022-08-02 tex/sections/intro.tex
Fix some readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
-The most conventional use case of a blockchain is financial application. This makes the security of the consensus
-protocol of a blockchain crucially important. Unfortunately, many currently used blockchains are vulnerable to a
-certain type
-of consensus attack, know as the bribery attack. In a bribery attack, an adversary tries to corrupt participants of a
-protocol and make them violate the protocol by offering them money.
+The most conventional use of blockchains is in financial applications. Therefore, the security level of the consensus
+protocol of a blockchain has a crucial importance. Unfortunately, many currently used blockchains are vulnerable to a
+certain type of consensus attack, know as the bribery attack. In a bribery attack, an adversary tries to corrupt
+participants of a protocol and make them violate the protocol by offering them money.
 
 At the time of writing of this document,
-the total mining reward for a bitcoin block is around \$150,000. If we assume that mining rewards define the utility
-function of bitcoin miners accurately, one could hire all hashing power of the bitcoin network for one hour by spending
+the total mining reward for a bitcoin block is around \$150,000. If we assume that mining rewards accurately define
+the utility function of a bitcoin miner, one could hire all hashing power of the bitcoin network for one hour by
+spending
 only \$750,000. The situation is not much different for PoS blockchains, as long as the total stake\footnote{Here, by
@@
 particular, to the locked amount of a user's money that is known as stake in some PoS protocols.}
-of validators is a relatively small value. This problem is more severe in blockchains that use randomly selected small
-sets of validators. Selecting these random sets by hidden random procedures would not help, as long as the
-validator himself knows he has been selected, before casting his vote.
+of the validator set of a block is a relatively small value. This problem is more severe in blockchains that use
+randomly selected small sets of validators, and selecting these random sets by hidden random procedures would not
+help, since the validator himself knows he has been selected, before casting his vote.
 
 It appears that the only solution to this important vulnerability is to effectively participate all the stake of the
-system in the consensus protocol. This large participation secures the protocol against bribery or
+system in the consensus protocol. This large participation makes the protocol resilient against bribery or
 collusion because the adversary would have to spend unrealistic amount of money to bribe enough users.
@@
 illegal transactions. Detecting illegal transaction can be done by accessing the ledger
-state and executing transactions according to the protocol predefined rules. The
+state and executing transactions according to the protocol defined rules. The
 ledger state, even for small blockchains, could be several
@@
 adversary to attack the consensus protocol. For example, a malware might be
-able to infect a lot of normal computers in the world, and then it might be used to hijack the majority of
+able to infect a lot of normal computers in the world, and then it could be used to hijack the majority of
 participants in the consensus protocol and compromise the security of the blockchain. Securing a computer
@@
 like custom-built operating systems and isolated specialized hardware would be required. This is not something a normal
-user can do. Only powerful centralized entities that have enough financial and technical resources are able to build
-systems that are resilient against sophisticated cyberattacks. So in this regard, a centralized system is arguably
-superior to a decentralized \footnote{Note
-that decentralized and distributed are two different concepts.} system.
+user can afford. Only powerful centralized entities, having large financial and technical resources, could build
+systems that are resilient against sophisticated cyberattacks. In this regard, we have to admit, a centralized system is
+arguably superior to a decentralized \footnote{Note that decentralized and distributed are two different concepts.}
+system.
 
-To overcome these difficulties, Argennon uses a hybrid proof of stake consensus protocol. A small committee of
+To overcome these difficulties, Argennon uses a hybrid Proof of Stake consensus protocol. A small committee of
 delegates is democratically elected
-by the Argennon users and is responsible for minting new blocks of the Argennon blockchain. Each
-minted block is required to get approved by its corresponding validator assembly. Assemblies are very large, and they
-contain at least three percent of the total stake of the system. A block is approved if it gets approval votes of at
-least \(2/3\) of the total stake of its validator assembly.
+by the Argennon users and is responsible for minting new blocks of the blockchain. Each
+minted block is required to get approved by its corresponding validator assembly. Assemblies are very large sets of
+validators containing at least three percent of the total stake of the system. A block is approved if it gets
+approval votes from at least \(2/3\) of the total stake of its validator assembly.
 
 Each block of the Argennon blockchain contains a set of Computational Integrity (CI) statements and a commitment to
-its ledger state. This set defines how
+the corresponding ledger state of the block. This set defines how
 the ledger state of a block can be reached from the state of its previous block via a set of intermediate states.
@@
 Verifying a succinct proof can be exponentially faster than replaying the computation. Moreover, verifications of
-different statements are independent and can be done in parallel. Therefore, a validator can use multiple cores for
+different CI statements are independent and can be done in parallel. Therefore, a validator can use multiple cores for
 verifying CI statements of a single block and different validator assemblies can simultaneously verify different blocks.

### 2022-08-03 tex/sections/intro.tex
Fix more readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
-The most conventional use of blockchains is in financial applications. Therefore, the security level of the consensus
-protocol of a blockchain has a crucial importance. Unfortunately, many currently used blockchains are vulnerable to a
-certain type of consensus attack, know as the bribery attack. In a bribery attack, an adversary tries to corrupt
+The most conventional use of blockchains is in financial applications. This gives a crucial importance to the
+security level of the consensus protocol used in a blockchain. Unfortunately, many currently used blockchains are
+vulnerable to a
+certain type of consensus attack, known as the bribery attack. In a bribery attack, an adversary tries to corrupt
 participants of a protocol and make them violate the protocol by offering them money.
@@
 particular, to the locked amount of a user's money that is known as stake in some PoS protocols.}
-of the validator set of a block is a relatively small value. This problem is more severe in blockchains that use
+of the validator set of a block is a relatively small value. This is more severe in blockchains that use
 randomly selected small sets of validators, and selecting these random sets by hidden random procedures would not
@@
 illegal transactions. Detecting illegal transaction can be done by accessing the ledger
-state and executing transactions according to the protocol defined rules. The
+state and executing transactions according to the protocol rules. The
 ledger state, even for small blockchains, could be several
 hundreds of gigabytes, and executing transactions could easily become costly when a blockchain is acting as a
-smart contract platform. This computational and storage overhead could prevent most of
+smart contract platform. This computational and storage overhead, in practice, could prevent most of
 ordinary users from any type of participation in the process of securing a blockchain.
@@
 as it might seem. A blockchain consensus protocol relies on a network of computers, not humans. Ordinary users use
-simple and similar computer systems. That means, they all have similar vulnerabilities that could be used by an
-adversary to attack the consensus protocol. For example, a malware might be
+simple and similar computer systems. That means, they all have similar vulnerabilities which could be used by an
+adversary to catastrophically attack the consensus protocol. For example, a malware might be
 able to infect a lot of normal computers in the world, and then it could be used to hijack the majority of
@@
 system against cyberattacks needs planning ahead and access to engineering resources. Special software and hardware,
-like custom-built operating systems and isolated specialized hardware would be required. This is not something a normal
-user can afford. Only powerful centralized entities, having large financial and technical resources, could build
+like custom-built operating systems and isolated specialized hardware is required. This is not something a normal
+user can afford. Only powerful centralized entities having large financial and technical resources, could build
 systems that are resilient against sophisticated cyberattacks. In this regard, we have to admit, a centralized system is
@@
 minted block is required to get approved by its corresponding validator assembly. Assemblies are very large sets of
-validators containing at least three percent of the total stake of the system. A block is approved if it gets
+validators including at least three percent of the total stake of the system. A block is approved if it gets
 approval votes from at least \(2/3\) of the total stake of its validator assembly.
@@
 Each block of the Argennon blockchain contains a set of Computational Integrity (CI) statements and a commitment to
-the corresponding ledger state of the block. This set defines how
-the ledger state of a block can be reached from the state of its previous block via a set of intermediate states.
-Each CI statement defines an ordered list of transactions (called external requests in Argennon) and determines
-the intermediate states before and after executing those transactions:
+the corresponding ledger state of the block. The CI set defines how
+the ledger state of the block can be reached from the state of its previous block via a set of intermediate states.
+Each individual CI statement defines an ordered list of transactions, called external requests in Argennon, and
+determines the intermediate states before and after executing those transactions:
 \[
@@
 verifying CI statements of a single block and different validator assemblies can simultaneously verify different blocks.
-In addition, proof generation of different CI statements can be done in parallel similarly. However, for parallel
+In addition, proof generation of different CI statements similarly can be done in parallel. However, for parallel
 proof generation, the state transition needs to be known in advance. That's why in Argennon, delegates do not try
@@
 
+%% Argennon execution layer is optimized for this ....
+
 Independence of CI statements is useful, but is not enough for having a truly scalable blockchain. To increase
@@
 would enable a block proposer\footnote{In Argennon, delegates are the only block proposers.} to use Data Dependency
-Analysis\footnote{See Section~\ref{sec:concurrency}} to indicate independent sets of transactions and use those sets
-for parallel processing of transactions. More importantly, these sets can be used for generating CI statements that
+Analysis\footnote{See Section~\ref{sec:concurrency}} to indicate independent sets of external requests (i.e.\
+transactions) and use those sets for parallel processing. More importantly, these sets can be used for generating CI
+statements that
 are defined on the same initial state and their proof can be generated independently without the need to calculate

### 2022-08-03 tex/sections/intro.tex
Fix more readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The most conventional use of blockchains is in financial applications. This gives a crucial importance to the
-security level of the consensus protocol used in a blockchain. Unfortunately, many currently used blockchains are
+security of the consensus protocol used in a blockchain. Unfortunately, many currently used blockchains are
 vulnerable to a
@@
 of the validator set of a block is a relatively small value. This is more severe in blockchains that use
-randomly selected small sets of validators, and selecting these random sets by hidden random procedures would not
-help, since the validator himself knows he has been selected, before casting his vote.
+randomly selected small sets of validators. These small sets could be easily bribed and selecting these random sets
+by hidden random procedures would not help, since the validator himself knows he has been selected, before casting
+his vote.
 
@@
 simple and similar computer systems. That means, they all have similar vulnerabilities which could be used by an
-adversary to catastrophically attack the consensus protocol. For example, a malware might be
-able to infect a lot of normal computers in the world, and then it could be used to hijack the majority of
-participants in the consensus protocol and compromise the security of the blockchain. Securing a computer
-system against cyberattacks needs planning ahead and access to engineering resources. Special software and hardware,
-like custom-built operating systems and isolated specialized hardware is required. This is not something a normal
-user can afford. Only powerful centralized entities having large financial and technical resources, could build
+adversary to catastrophically attack the consensus protocol. For example, a malware using a common zero-day
+vulnerability might be able to infect a large portion of normal personal computers. This malware could be used by an
+adversary to control the majority of participant computers in the consensus protocol and compromise the security of the
+blockchain.
+
+Securing a computer system against cyberattacks needs planning ahead and access to engineering resources.
+Special software and hardware, like custom-built operating systems and isolated specialized hardware is required.
+This is not something a normal user can afford. Only powerful centralized entities having large financial and
+technical resources, could build
 systems that are resilient against sophisticated cyberattacks. In this regard, we have to admit, a centralized system is
-arguably superior to a decentralized \footnote{Note that decentralized and distributed are two different concepts.}
+arguably superior to a decentralized\footnote{Note that decentralized and distributed are two different concepts.}
 system.
@@
 the ledger state of the block can be reached from the state of its previous block via a set of intermediate states.
-Each individual CI statement defines an ordered list of transactions, called external requests in Argennon, and
-determines the intermediate states before and after executing those transactions:
+Each individual CI statement defines an ordered list of external requests\footnote{External requests in Argennon are
+similar to transactions in older blockchains.} and determines the intermediate states before and after executing those
+requests:
 \[
@@
 Validators can\footnote{Using the Argennon cloud is optional for a validator.} receive succinct cryptographic proofs
-of these CI statements from the Argennon Publicly Verifiable Cloud. They can also receive the state of the previous
+of these CI statements from the Argennon cloud. They can also receive the state of the previous
 block (the required part of it) and proofs that show the received data is consistent with the previous block commitment.
@@
 Verifying a succinct proof can be exponentially faster than replaying the computation. Moreover, verifications of
-different CI statements are independent and can be done in parallel. Therefore, a validator can use multiple cores for
-verifying CI statements of a single block and different validator assemblies can simultaneously verify different blocks.
+different CI statements are independent and can be done in parallel. As a result, a validator can use multiple cores for
+verifying CI statements of a single block and different validator assemblies can simultaneously and independently
+verify different blocks.
 In addition, proof generation of different CI statements similarly can be done in parallel. However, for parallel
 proof generation, the state transition needs to be known in advance. That's why in Argennon, delegates do not try
-to generate proofs and focus all their computational power on executing requests and generating the state transition
-as fast as possible.
+to generate proofs and focus all their computational power on executing external requests and
+generating the state transition as fast as possible.
 
@@
 would enable a block proposer\footnote{In Argennon, delegates are the only block proposers.} to use Data Dependency
-Analysis\footnote{See Section~\ref{sec:concurrency}} to indicate independent sets of external requests (i.e.\
+Analysis\footnote{See Section~\ref{sec:concurrency}} (DDA) to indicate independent sets of external requests (i.e.\
 transactions) and use those sets for parallel processing. More importantly, these sets can be used for generating CI

### 2022-08-03 tex/sections/intro.tex
Fix more readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 simple and similar computer systems. That means, they all have similar vulnerabilities which could be used by an
-adversary to catastrophically attack the consensus protocol. For example, a malware using a common zero-day
-vulnerability might be able to infect a large portion of normal personal computers. This malware could be used by an
-adversary to control the majority of participant computers in the consensus protocol and compromise the security of the
+adversary to catastrophically attack the consensus protocol. For example, if a malware, probably using a common zero-day
+vulnerability, has the ability to infect a large portion of normal personal computers, it could be used by an
+adversary to control the majority of participant computers in the consensus and compromise the security of the
 blockchain.
@@
 
-To overcome these difficulties, Argennon uses a hybrid Proof of Stake consensus protocol. A small committee of
+To overcome these difficulties, Argennon uses a hybrid Proof of Stake consensus protocol, which is
+capable of combining the benefits of a centralized and a decentralized system. A small committee of
 delegates is democratically elected

### 2022-08-05 tex/sections/intro.tex
Update the introduction chapter
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2022-08-03 tex/sections/intro.tex
@@
 At the time of writing of this document,
-the total mining reward for a bitcoin block is around \$150,000. If we assume that mining rewards accurately define
-the utility function of a bitcoin miner, one could hire all hashing power of the bitcoin network for one hour by
-spending
-only \$750,000. The situation is not much different for PoS blockchains, as long as the total stake\footnote{Here, by
+the total mining reward for a bitcoin block is around \$150,000. If we assume, in decision theoretic terminology,
+that the mining reward accurately defines the utility function of a bitcoin miner, one could hire all hashing power
+of the bitcoin network for one hour by
+spending only \$750,000. The situation is not much different for PoS blockchains, as long as the total
+stake\footnote{Here, by
 stake, we mean a real number measuring the total interest of a user in the system. We are not referring, in
 particular, to the locked amount of a user's money that is known as stake in some PoS protocols.}
-of the validator set of a block is a relatively small value. This is more severe in blockchains that use
+of the validator set is a relatively small value. This is more severe in blockchains that use
 randomly selected small sets of validators. These small sets could be easily bribed and selecting these random sets
@@
 system in the consensus protocol. This large participation makes the protocol resilient against bribery or
-collusion because the adversary would have to spend unrealistic amount of money to bribe enough users.
+collusion because the adversary would need to spend unrealistic amount of money to bribe enough users.
 
@@
 simple and similar computer systems. That means, they all have similar vulnerabilities which could be used by an
-adversary to catastrophically attack the consensus protocol. For example, a malware using a common zero-day
-vulnerability might be able to infect a large portion of normal personal computers. This malware could be used by an
-adversary to control the majority of participant computers in the consensus protocol and compromise the security of the
+adversary to catastrophically attack the consensus protocol. For example, if a malware, probably using a common zero-day
+vulnerability, has the ability to infect a large portion of normal personal computers, it could be used by an
+adversary to control the majority of participants in the consensus protocol and compromise the security of the
 blockchain.
@@
 
-To overcome these difficulties, Argennon uses a hybrid Proof of Stake consensus protocol. A small committee of
+To overcome these difficulties, Argennon uses a Hybrid Proof of Stake consensus protocol, which is
+capable of combining the benefits of a centralized and a decentralized system. A small committee of
 delegates is democratically elected
 by the Argennon users and is responsible for minting new blocks of the blockchain. Each
-minted block is required to get approved by its corresponding validator assembly. Assemblies are very large sets of
-validators including at least three percent of the total stake of the system. A block is approved if it gets
-approval votes from at least \(2/3\) of the total stake of its validator assembly.
+minted block is required to get approved by its corresponding validator committee. Validator committees are very
+large sets of validators including at least three percent of the total stake of the system. A block is approved if it
+gets approval votes from at least \(2/3\) of the total stake of its validator committee.
+
+Every block of the Argennon blockchain needs a certificate from both the committee of delegates and
+the committee of validators. A block is considered final after its \textbf{next} block receives \textbf{both} of
+its certificates. The Argennon protocol ensures that as long as more than half of the total stake of validators is
+controlled by honest users, the probability of discarding a final block is near zero even if all the delegates are
+malicious.
+
+Block generation is centralized in Argennon. This brings some interesting features to the Argennon platform, such as off
+chain fee payment, optimistic instant transaction confirmation and runoff protection. However, centralized
+block generation also increases the possibility of transaction censorship. This problem is addressed by a special
+High Priority Transaction (HPT) protocol in Argennon.
 
@@
 Validators can\footnote{Using the Argennon cloud is optional for a validator.} receive succinct cryptographic proofs
-of these CI statements from the Argennon cloud. They can also receive the state of the previous
+of these CI statements (STARK/SNARK) from the Argennon cloud. They can also receive the state of the previous
 block (the required part of it) and proofs that show the received data is consistent with the previous block commitment.
@@
 different CI statements are independent and can be done in parallel. As a result, a validator can use multiple cores for
-verifying CI statements of a single block and different validator assemblies can simultaneously and independently
-verify different blocks.
+verifying CI statements of a \textbf{single} block and different validator committees can simultaneously and
+independently verify \textbf{different} blocks.
 In addition, proof generation of different CI statements similarly can be done in parallel. However, for parallel
@@
 
-%% Argennon execution layer is optimized for this ....
+Argennon applications (i.e smart contracts) are stored in a high level text based language on the Argennon
+blockchain. This language is intended for preserving the
+high level information of the application logic to facilitate platform specific compiler optimization at a host
+machine. This enables delegates to compile and optimize Argennon applications for their specific hardware platforms
+and execute applications efficiently.
 
@@
 transactions) and use those sets for parallel processing. More importantly, these sets can be used for generating CI
-statements that
-are defined on the same initial state and their proof can be generated independently without the need to calculate
-the state transition in advance.
+statements that are defined on the \textbf{same initial state} and their proof can be generated independently without
+the need to calculate the state transition in advance.
 
-The Argennon protocol strongly incentivizes the formation of a permission-less network of Publicly Verifiable
+The Argennon protocol strongly incentivizes the formation of a \textbf{permission-less} network of Publicly Verifiable
 Cloud (PVC) servers. To do so, the Argennon protocol conducts repetitive automatic lotteries between PVC servers.

### 2022-08-06 tex/sections/intro.tex
Fix some readability issues in Chapter 1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 large sets of validators including at least three percent of the total stake of the system. A block is approved if it
-gets approval votes from at least \(2/3\) of the total stake of its validator committee.
+takes approval votes from at least \(2/3\) of the total stake of its validator committee.
 
 Every block of the Argennon blockchain needs a certificate from both the committee of delegates and
-the committee of validators. A block is considered final after its \textbf{next} block receives \textbf{both} of
+the committee of validators. A block is considered final after its \textbf{next} block gets \textbf{both} of
 its certificates. The Argennon protocol ensures that as long as more than half of the total stake of validators is
@@
 
-Block generation is centralized in Argennon. This brings some interesting features to the Argennon platform, such as off
-chain fee payment, optimistic instant transaction confirmation and runoff protection. However, centralized
-block generation also increases the possibility of transaction censorship. This problem is addressed by a special
-High Priority Transaction (HPT) protocol in Argennon.
+Centralized block generation brings some interesting features to the Argennon platform, such as off
+chain fee payment, optimistic instant transaction confirmation and runoff protection. However, it also increases
+the possibility of transaction censorship. In Argennon, this problem is addressed by a special
+High Priority Transaction (HPT) protocol.
 
@@
 
-Argennon applications (i.e smart contracts) are stored in a high level text based language on the Argennon
-blockchain. This language is intended for preserving the
+Argennon applications (i.e smart contracts) are stored in a high level text based language on the blockchain. This
+language is intended for preserving the
 high level information of the application logic to facilitate platform specific compiler optimization at a host

### 2022-08-06 tex/sections/intro.tex
Fix some readability issues in Chapter 1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 as it might seem. A blockchain consensus protocol relies on a network of computers, not humans. Ordinary users use
-simple and similar computer systems. That means, they all have similar vulnerabilities which could be used by an
-adversary to catastrophically attack the consensus protocol. For example, if a malware, probably using a common zero-day
+simple and similar computer systems. That means, they all have similar vulnerabilities and weaknesses which could be
+used by an adversary to catastrophically attack the consensus protocol. For example, if a malware, probably using a
+common zero-day
 vulnerability, has the ability to infect a large portion of normal personal computers, it could be used by an
@@
 
-Centralized block generation brings some interesting features to the Argennon platform, such as off
-chain fee payment, optimistic instant transaction confirmation and runoff protection. However, it also increases
-the possibility of transaction censorship. In Argennon, this problem is addressed by a special
-High Priority Transaction (HPT) protocol.
+Centralized block generation brings some interesting features to the Argennon platform, such as flexible and lower
+fees, off chain fee payment, optimistic instant transaction confirmation, and runoff protection.
+However, it also increases the possibility of transaction censorship. In Argennon, this problem is addressed by a
+special High Priority Transaction (HPT) protocol.
 
 Each block of the Argennon blockchain contains a set of Computational Integrity (CI) statements and a commitment to
-the corresponding ledger state of the block. The CI set defines how
-the ledger state of the block can be reached from the state of its previous block via a set of intermediate states.
+the final ledger state of the block. The CI set defines how the final ledger state of the block can be reached from
+the state of its previous block via a set of intermediate states.
 Each individual CI statement defines an ordered list of external requests\footnote{External requests in Argennon are
@@
 machine. This enables delegates to compile and optimize Argennon applications for their specific hardware platforms
-and execute applications efficiently.
+and execute applications efficiently, ensuring the state transition can be found as fast as possible.
 

### 2022-08-07 tex/A.tex
changed chapter 1 of whitepaper

@@
         a centralized and a decentralized system. In Argennon, ledger storage and transaction processing are
-        outsourced to the cloud and normal personal computers or smartphones, with limited hardware
+        outsourced to the Argennon cloud and normal personal computers or smartphones, with limited hardware
         capabilities, can validate transactions and actively
-        participate in the Argennon consensus protocol. This property makes Argennon a truly decentralized and
+        participate in the consensus protocol. This property makes Argennon a truly decentralized and
         democratic blockchain and one of the most secure existing platforms.
 
-        The Argennon cloud is trustless and publicly verifiable. Computational Integrity (CI) is
+        The Argennon cloud consists of a Committee of Delegates (CoD) and Publicly Verifiable Cloud (PVC) servers. 
+        The CoD is elected per term by Proof of Stake and is responsible for processing transactions and creating blocks.
+        After validator assemblies have verified the validity of these blocks they are stored in the PVC servers.
+        PVC servers are publicly verifiable. Computational Integrity (CI) is
         achieved by using succinct cryptographic proofs (STARK/SNARK)
@@
 
-        The Argennon protocol strongly incentivizes the formation of a permission-less network of Publicly Verifiable
-        Cloud (PVC) servers. A PVC server in Argennon, is a conventional data server which uses its computational and
-        storage resources to help the Argennon network process transactions. This encourages the development
-        of conventional networking, storage and compute hardware, which can benefit all areas of information technology.
-        This contrasts with the approach of some older blockchains that incentivized the development of a totally
-        useless technology of hash calculation.
-
+        This set up allows for very high transaction throughput while at the same time remaining highly secure. The network 
+        stays safe as long as at least one delegate and at least 1/3 of the validators in all validator assemblies remain honest.
     \end{abstract}

### 2022-08-07 tex/A.tex
changed chapter 1 of whitepaper some more

@@
         a centralized and a decentralized system. In Argennon, ledger storage and transaction processing are
-        outsourced to the Argennon cloud and normal personal computers or smartphones, with limited hardware
-        capabilities, can validate transactions and actively
-        participate in the consensus protocol. This property makes Argennon a truly decentralized and
-        democratic blockchain and one of the most secure existing platforms.
+        outsourced to the Argennon cloud while normal personal computers or smartphones, with limited hardware
+        capabilities, can validate transactions and actively participate in the consensus protocol.
 
@@
         PVC servers are publicly verifiable. Computational Integrity (CI) is
-        achieved by using succinct cryptographic proofs (STARK/SNARK)
-        and data integrity is guaranteed by cryptographic accumulators. At the same time, a smart
-        clustering algorithm keeps the bandwidth usage and the overhead of cryptography manageable
+        achieved by using succinct cryptographic proofs (STARK/SNARK) and data integrity is guaranteed by cryptographic accumulators. 
+        At the same time, a smart clustering algorithm keeps the bandwidth usage and the overhead of cryptography manageable
         for validators.
 
-        This set up allows for very high transaction throughput while at the same time remaining highly secure. The network 
-        stays safe as long as at least one delegate and at least 1/3 of the validators in all validator assemblies remain honest.
+        These properties make it possible for Argennon to achieve very high transaction throughput while requiring no trust
+        in central parties.  
     \end{abstract}

### 2022-08-07 tex/sections/intro.tex
Explain more about delegates in Chapter 1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 capable of combining the benefits of a centralized and a decentralized system. A small committee of
-delegates is democratically elected
-by the Argennon users and is responsible for minting new blocks of the blockchain. Each
-minted block is required to get approved by its corresponding validator committee. Validator committees are very
+delegates is democratically elected by users via the Argennon Decentralized Autonomous Governance System
+(ADAGs). This committee usually\footnote{The election term and the number of committee members can be changed by the
+ADAGs.} is elected for a one-year term, has five members, and is responsible for minting new
+blocks of the blockchain.
+Each minted block is required to get approved by its corresponding validator committee. Validator committees are very
 large sets of validators including at least three percent of the total stake of the system. A block is approved if it

### 2022-08-07 tex/sections/intro.tex
typos, change PV-DB to PVC, and change confusing sentence

--- 2022-08-06 tex/sections/intro.tex
@@
 system in the consensus protocol. This large participation makes the protocol resilient against bribery or
-collusion because the adversary would need to spend unrealistic amount of money to bribe enough users.
+collusion because the adversary would need to spend unrealistic amounts of money to bribe enough users.
 

### 2022-08-07 tex/sections/intro.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2022-08-07 tex/sections/intro.tex
@@
 Each minted block is required to get approved by its corresponding validator committee. Validator committees are very
-large sets of validators including at least three percent of the total stake of the system. A block is approved if it
-takes approval votes from at least \(2/3\) of the total stake of its validator committee.
+large sets of validators including at least three percent of the maximum possible stake of the system. A block is
+approved if it takes approval votes from at least \(2/3\) of the total stake of its validator committee.
 

### 2022-08-08 tex/sections/intro.tex
Clarify some parts of Ch1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
-The most conventional use of blockchains is in financial applications. This gives a crucial importance to the
+The most common use for blockchains is in financial applications. This gives a crucial importance to the
 security of the consensus protocol used in a blockchain. Unfortunately, many currently used blockchains are
-vulnerable to a
-certain type of consensus attack, known as the bribery attack. In a bribery attack, an adversary tries to corrupt
-participants of a protocol and make them violate the protocol by offering them money.
+vulnerable to a certain type of consensus attack, known as the bribery attack. In a bribery attack, an adversary
+tries to corrupt participants of a protocol by offering them money to make them violate the protocol.
 
@@
 spending only \$750,000. The situation is not much different for PoS blockchains, as long as the total
-stake\footnote{Here, by
+stake of the validator set is a relatively small value. Here by
 stake, we mean a real number measuring the total interest of a user in the system. We are not referring, in
-particular, to the locked amount of a user's money that is known as stake in some PoS protocols.}
-of the validator set is a relatively small value. This is more severe in blockchains that use
-randomly selected small sets of validators. These small sets could be easily bribed and selecting these random sets
+particular, to some locked amount of a user's money that is known as stake in some PoS protocols.
+
+This problem is more severe in blockchains that use randomly selected small sets of validators. These small sets
+usually have low total stake and could be easily bribed and corrupted. Selecting these random sets
 by hidden random procedures would not help, since the validator himself knows he has been selected, before casting
@@
 
-It appears that the only solution to this important vulnerability is to effectively participate all the stake of the
-system in the consensus protocol. This large participation makes the protocol resilient against bribery or
+It appears that the only solution to this important vulnerability is to effectively participate all the stakeholders
+of the system in the consensus protocol. This large participation makes the protocol resilient against bribery or
 collusion because the adversary would need to spend unrealistic amounts of money to bribe enough users.
@@
 However, for effective participation in the consensus protocol, a validator needs to be able to detect
-illegal transactions. Detecting illegal transaction can be done by accessing the ledger
+illegal transactions. Detecting illegal transactions can be done by accessing the ledger
 state and executing transactions according to the protocol rules. The

### 2022-08-08 tex/sections/intro.tex
Add a notice for Chapter 2
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-To overcome these difficulties, Argennon uses a Hybrid Proof of Stake consensus protocol, which is
+To overcome these difficulties, Argennon\footnote{The classical pronunciation should be used:\textipa{/Ar"gen.non/}}
+uses a Hybrid Proof of Stake consensus protocol, which is
 capable of combining the benefits of a centralized and a decentralized system. A small committee of

### 2022-08-11 tex/A.tex
Add a notice for Chapter 2
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Add a notice for Chapter 5 and fix typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2022-08-02 tex/A.tex
@@
         outsourced to the cloud and normal personal computers or smartphones, with limited hardware
-        capabilities, can validate transactions and actively
+        capabilities, are able to validate transactions and actively
         participate in the Argennon consensus protocol. This property makes Argennon a truly decentralized and
@@
         Cloud (PVC) servers. A PVC server in Argennon, is a conventional data server which uses its computational and
-        storage resources to help the Argennon network process transactions. This encourages the development
-        of conventional networking, storage and compute hardware, which can benefit all areas of information technology.
-        This contrasts with the approach of some older blockchains that incentivized the development of a totally
-        useless technology of hash calculation.
+        storage resources to help the Argennon network process transactions.
 
@@
     \chapter{The Argennon Smart Contract Execution Environment}\label{ch:AVM}
+    \note{This chapter is outdated and does not include CI proofs. Currently a new version is being rewritten.}
     \input{sections/avm}
@@
     \chapter{The Argennon Blockchain}\label{ch:argennon-blockchain}
+    \note{Currently a new version is being rewritten to include CI proofs.}
 

### 2022-08-13 tex/sections/avm.tex

@@
 
-The Argennon\footnote{The classical pronunciation should be used:\textipa{/Ar"gen.non/}} Smart Contract Execution
+The Argennon Smart Contract Execution
 Environment (AscEE) is an abstract execution environment for executing Argennon smarts contracts (a.k.a\ Argennon

### 2022-08-13 tex/sections/intro.tex

@@
 vulnerable to a certain type of consensus attack, known as the bribery attack. In a bribery attack, an adversary
-tries to corrupt participants of a protocol by offering them money to make them violate the protocol.
+tries to corrupt participants of a protocol by offering them money and making them violate the protocol.
 
@@
 Although a fully decentralized blockchain based on the participation of every user looks appealing, it is not as perfect
-as it might seem. A blockchain consensus protocol relies on a network of computers, not humans. Ordinary users use
-simple and similar computer systems. That means, they all have similar vulnerabilities and weaknesses which could be
-used by an adversary to catastrophically attack the consensus protocol. For example, if a malware, probably using a
+as it might seem. The consensus protocol of a blockchain relies on a network of computers, not humans. Ordinary users
+use simple and similar computer systems. That means, they all have similar vulnerabilities and weaknesses which could be
+used by an adversary to catastrophically attack the consensus protocol. For instance, if a malware, probably using a
 common zero-day
@@
 capable of combining the benefits of a centralized and a decentralized system. A small committee of
-delegates is democratically elected by users via the Argennon Decentralized Autonomous Governance System
+delegates is democratically elected by users via the Argennon Decentralized Autonomous Governance system
 (ADAGs). This committee usually\footnote{The election term and the number of committee members can be changed by the
 ADAGs.} is elected for a one-year term, has five members, and is responsible for minting new
-blocks of the blockchain.
+blocks of the Argennon blockchain.
 Each minted block is required to get approved by its corresponding validator committee. Validator committees are very
@@
 However, it also increases the possibility of transaction censorship. In Argennon, this problem is addressed by a
-special High Priority Transaction (HPT) protocol.
+special High Priority Request (HPR) protocol.
 
@@
 Each individual CI statement defines an ordered list of external requests\footnote{External requests in Argennon are
-similar to transactions in older blockchains.} and determines the intermediate states before and after executing those
-requests:
+similar to transactions in older blockchains.} and determines the state before and after executing those requests:
 \[
@@
 Verifying a succinct proof can be exponentially faster than replaying the computation. Moreover, verifications of
-different CI statements are independent and can be done in parallel. As a result, a validator can use multiple cores for
-verifying CI statements of a \textbf{single} block and different validator committees can simultaneously and
+different CI statements are independent of each other and can be done in parallel. As a result, a validator can use
+multiple cores for verifying CI statements of a \textbf{single} block and different validator committees can
+simultaneously and
 independently verify \textbf{different} blocks.
@@
 of conventional networking, storage and compute hardware, which can benefit all areas of information technology.
+This contrasts with the approach of some older blockchains that incentivized the development of a totally
+useless technology of hash calculation.

### 2022-08-13 tex/sections/incentive.tex
Update abstract, fix typos and add draw.io diagram (#12)
* changed chapter 1 of whitepaper

* changed chapter 1 of whitepaper some more

* create folder for diagrams

* Added Argennon-fees.drawio

* remove empty file

* Update Argennon-fees.drawio

* typos, change PV-DB to PVC, and change confusing sentence

Co-authored-by: Behrouz <behrouz_ayati@yahoo.com>

@@
 
-Only behaviours will be punished that can not happen due to a mistake or an attack. These behaviours include:
+Punishable behaviours include:
 \begin{itemize}
@@
 
-\subsection{Incentives for PV-DB Servers}\label{subsec:PV-DB-servers}
+\subsection{Incentives for PVC Servers}\label{subsec:PVC-servers}
 
-The incentive mechanism for PV-DB servers should have the following properties:
+The incentive mechanism for PVC servers should have the following properties:
 
@@
     \item It incentivizes storing all storage pages and not only those pages that are used more frequently.
-    \item It incentivizes PV-DB servers to actively provide the required storage pages for validators.
-    \item Making more accounts will not provide any advantage for a PV-DB server.
+    \item It incentivizes PVC servers to actively provide the required storage pages for validators.
+    \item Making more accounts will not provide any advantage for a PVC server.
 \end{itemize}
 
-For our incentive mechanism, we require that every time a validator receives a storage page from a PV-DB, after
-validating the data, he gives a receipt to the PV-DB server. In this receipt the validator signs the
+For our incentive mechanism, we require that every time a validator receives a storage page from a PVC, after
+validating the data, he gives a receipt to the PVC server. In this receipt the validator signs the
 following information:
@@
 \begin{itemize}
-    \item \texttt{ownerAddr}: the account address of the PV-DB server.
+    \item \texttt{ownerAddr}: the account address of the PVC server.
     \item \texttt{receivedPageID}: the ID of the received page.
@@
 
-\note{In a round, an honest validator never gives a receipt for an identical page to two different PV-DB servers.}
+\note{In a round, an honest validator never gives a receipt for an identical page to two different PVC servers.}
 
-To incentivize PV-DB servers, a lottery will be held every round,\footnote{A round is the time interval between
+To incentivize PVC servers, a lottery will be held every round,\footnote{A round is the time interval between
 two consecutive blocks.} and a predefined amount of ARGs from
@@
 
-\note{One PV-DB server could own multiple winning tickets in a round.}
+\note{One PVC server could own multiple winning tickets in a round.}
 
@@
     \item The receipt round number is $r - 1$.
-    \item The signer did not sign a receipt for the same storage page for two different PV-DB servers in
+    \item The signer did not sign a receipt for the same storage page for two different PVC servers in
     the previous round.
@@
 Also, based on the current block seed, a random storage page is
-selected as the challenge of the round. A PV-DB server that owns a winning receipt needs to broadcast a \emph{winning
+selected as the challenge of the round. A PVC server that owns a winning receipt needs to broadcast a \emph{winning
 ticket} to claim his prize. The winning ticket consists of a winning receipt and a \emph{solution} to the round
 challenge. Solving a round challenge requires the content of the storage page which was selected as the round
-challenge. This will encourage PV-DB servers to store all storage pages.
+challenge. This will encourage PVC servers to store all storage pages.
 

### 2022-08-13 tex/A.tex
Update abstract, fix typos and add draw.io diagram (#12)
* changed chapter 1 of whitepaper

* changed chapter 1 of whitepaper some more

* create folder for diagrams

* Added Argennon-fees.drawio

* remove empty file

* Update Argennon-fees.drawio

* typos, change PV-DB to PVC, and change confusing sentence

Co-authored-by: Behrouz <behrouz_ayati@yahoo.com>
Add more explanation about validators in intro
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
         The Argennon cloud is trustless and publicly verifiable. Computational Integrity (CI) is
-        achieved by using succinct cryptographic proofs (STARK/SNARK)
+        achieved by using succinct argument of knowledge systems (STARK/SNARK)
         and data integrity is guaranteed by cryptographic accumulators. At the same time, a smart
@@
 
+        These properties make it possible for Argennon to achieve very high transaction throughput while requiring no trust
+        in central parties.  
     \end{abstract}

### 2022-08-13 tex/sections/intro.tex
Add more explanation about validators in intro
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 blocks of the Argennon blockchain.
-Each minted block is required to get approved by its corresponding validator committee. Validator committees are very
-large sets of validators including at least three percent of the maximum possible stake of the system. A block is
-approved if it takes approval votes from at least \(2/3\) of the total stake of its validator committee.
+Each minted block is required to get approved by its corresponding validator committee.
+Validator committees are very
+large sets of validators including at least three percent of the maximum possible stake of the system.
+Every committee of validators has an index between $0$ and $m - 1$, and is resposible for validating block number $n$,
+if $n$ modulo $m$ equals the committee index. A block is approved if it takes approval votes from at least $2/3$ of
+the total stake of its validator committee.
 

### 2022-08-14 tex/sections/intro.tex
Add explanation about safeguard chains in CH1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 large sets of validators including at least three percent of the maximum possible stake of the system.
-Every committee of validators has an index between $0$ and $m - 1$, and is resposible for validating block number $n$,
+Every committee of validators has an index between $0$ and $m - 1$, and is responsible for validating block number $n$,
 if $n$ modulo $m$ equals the committee index. A block is approved if it takes approval votes from at least $2/3$ of
@@
 Each block of the Argennon blockchain contains a set of Computational Integrity (CI) statements and a commitment to
-the final ledger state of the block. The CI set defines how the final ledger state of the block can be reached from
+the final ledger state of the block (i.e.\ the ledger state after executing all transactions of the block). The CI
+set defines how the final state of the block can be reached from
 the state of its previous block via a set of intermediate states.
@@
 
+Using succinct argument of knowledge systems makes an Argennon validator light enough to be implemented as a smart
+contract. By deploying an Argennon validator on another turing complete smart contract platform, Argennon could use
+more established blockchains as an extra layer of security, specially
+during the bootstrapping phase, when ARG is not well distributed yet.
+
 The Argennon protocol strongly incentivizes the formation of a \textbf{permission-less} network of Publicly Verifiable

### 2022-08-14 tex/sections/intro.tex
Improve readability
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-Using succinct argument of knowledge systems makes an Argennon validator light enough to be implemented as a smart
-contract. By deploying an Argennon validator on another turing complete smart contract platform, Argennon could use
-more established blockchains as an extra layer of security, specially
-during the bootstrapping phase, when ARG is not well distributed yet.
+Using succinct argument of knowledge systems makes the main functionalities of an Argennon validator light enough to
+be implemented as a smart contract. By deploying a validator contract on another platform, Argennon
+could use more established blockchains as an extra layer of security, specially
+during the bootstrapping phase, when ARG is not well distributed yet. To reduce the execution fee, this validator
+contract could be used for validating roll-ups of state transition.
 

### 2022-08-19 tex/A.tex
Add explanation about safeguard chains in CH1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Update intro and abstract
Several concepts have been explained more accurately now.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-\title{Argennon: A Scalable Cloud Based Smart Contract Platform}
+\title{Argennon: A Scalable Cloud Based Smart Contract Platform Using Argument of Knowledge Systems}
 \author{aybehrouz}
@@
         achieved by using succinct argument of knowledge systems (STARK/SNARK)
-        and data integrity is guaranteed by cryptographic accumulators. At the same time, a smart
-        clustering algorithm keeps the bandwidth usage and the overhead of cryptography manageable
-        for validators.
+        and data integrity is guaranteed by cryptographic accumulators.
 
@@
         storage resources to help the Argennon network process transactions.
-
-        These properties make it possible for Argennon to achieve very high transaction throughput while requiring no trust
-        in central parties.  
     \end{abstract}

### 2022-08-19 tex/sections/intro.tex
Update intro and abstract
Several concepts have been explained more accurately now.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 blocks of the Argennon blockchain.
-Each minted block is required to get approved by its corresponding validator committee.
+Each minted block gets certified by all members of the delegates committee and after that it required to get approved
+by its corresponding validator committee.
+
 Validator committees are very
-large sets of validators including at least three percent of the maximum possible stake of the system.
+large sets of validators including at least three percent of the maximum possible stake of the Argennon network.
 Every committee of validators has an index between $0$ and $m - 1$, and is responsible for validating block number $n$,
@@
 
-Every block of the Argennon blockchain needs a certificate from both the committee of delegates and
-the committee of validators. A block is considered final after its \textbf{next} block gets \textbf{both} of
-its certificates. The Argennon protocol ensures that as long as more than half of the total stake of validators is
+In case the main committee fails to generate new blocks or behaves maliciously, a special Recovery Protocol is
+initiated by validators.
+This recovery protocol can recover the functionality
+of the Argennon blockchain as long as more than $2/3$ of the total stake of the system is controlled by honest users
+and any network partition resolves after a finite amount of time. The Recovery Protocol uses two main emergency
+procedures to recover the functionality of the Argennon blockchain: \emph{emergency forking} and \emph{emergency
+agreement} protocol.
+
+In Argennon, a block is considered final after its \textbf{next} block gets certified by \textbf{both} delegates and
+validator committees.
+The Argennon protocol ensures that as long as more than half of the total stake of validators is
 controlled by honest users, the probability of discarding a final block is near zero even if all the delegates are
@@
 set defines how the final state of the block can be reached from
-the state of its previous block via a set of intermediate states.
-Each individual CI statement defines an ordered list of external requests\footnote{External requests in Argennon are
-similar to transactions in older blockchains.} and determines the state before and after executing those requests:
-\[
-    \tau_{(s_i,s_{i+1},\mathbb{R})} \coloneqq \text{``$s_{i+1}$ is the next state after executing external requests
-        $\mathbb{R}$ on state $s_i$''}
-\]
+the state of its previous block via a set of intermediate state transitions. Each individual CI statement defines
+an ordered list of external requests\footnote{External requests in Argennon are
+similar to transactions in older blockchains.} and determines the state before and after executing those requests.
+
+More formally, each CI statement has the form $\tau(\mathbb{S}, \mathbb{R}, \hat{\mathbb{S}})$, and states that
+$\hat{\mathbb{S}}$ is the commitment of the next state after executing an ordered list of external requests with commitment
+$\mathbb{R}$ on a state with commitment $\mathbb{S}$.
+Commitments to states are produced by a cryptographic accumulator.
 Validators can\footnote{Using the Argennon cloud is optional for a validator.} receive succinct cryptographic proofs
-of these CI statements (STARK/SNARK) from the Argennon cloud. They can also receive the state of the previous
-block (the required part of it) and proofs that show the received data is consistent with the previous block commitment.
+of these CI statements (STARK/SNARK) from the Argennon cloud and efficiently validate a block.
 
@@
 Cloud (PVC) servers. To do so, the Argennon protocol conducts repetitive automatic lotteries between PVC servers.
-A PVC server can increase its chance of winning by (i) generating proofs for more CI statements, (ii) providing the
-state data to more validators and, (iii) storing all parts of the state instead of more frequently used parts.
+A PVC server can increase its chance of winning by (i) generating proofs for more CI statements, (ii) storing all
+parts of the ledger state and providing \emph{proofs of storage}.
 

### 2022-08-19 tex/sections/intro.tex
Fix readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 blocks of the Argennon blockchain.
-Each minted block gets certified by all members of the delegates committee and after that it required to get approved
-by its corresponding validator committee.
+Each minted block gets certified by all members of the delegates committee and after that it must get approved
+by its corresponding validator assembly.
 
-Validator committees are very
+Validator assemblies are very
 large sets of validators including at least three percent of the maximum possible stake of the Argennon network.
-Every committee of validators has an index between $0$ and $m - 1$, and is responsible for validating block number $n$,
-if $n$ modulo $m$ equals the committee index. A block is approved if it takes approval votes from at least $2/3$ of
-the total stake of its validator committee.
+Every validator assembly has an index between $0$ and $m - 1$, and is responsible for validating block number $n$,
+if $n$ modulo $m$ equals the index of the assembly. A block is approved if it takes approval votes from at
+least $2/3$ of the total stake of its validator assembly.
 
@@
 
-In Argennon, a block is considered final after its \textbf{next} block gets certified by \textbf{both} delegates and
-validator committees.
+In Argennon, a block is considered final after its \textbf{next} block gets certified by \textbf{both} the
+delegates committee and its validator assembly.
 The Argennon protocol ensures that as long as more than half of the total stake of validators is
@@
 Centralized block generation brings some interesting features to the Argennon platform, such as flexible and lower
-fees, off chain fee payment, optimistic instant transaction confirmation, and runoff protection.
+fees, off chain fee payment, optimistic instant transaction confirmation, and front running protection.
 However, it also increases the possibility of transaction censorship. In Argennon, this problem is addressed by a
@@
 
-More formally, each CI statement has the form $\tau(\mathbb{S}, \mathbb{R}, \hat{\mathbb{S}})$, and states that
-$\hat{\mathbb{S}}$ is the commitment of the next state after executing an ordered list of external requests with commitment
-$\mathbb{R}$ on a state with commitment $\mathbb{S}$.
+More formally, each CI statement has the form $\tau_{\mathbb{S}, \hat{\mathbb{S}}, \mathbb{R}}$, and states that
+$\hat{\mathbb{S}}$ is the commitment of the next state after executing an ordered list of external requests with
+commitment $\mathbb{R}$ on a state which has commitment $\mathbb{S}$.
 Commitments to states are produced by a cryptographic accumulator.
 Validators can\footnote{Using the Argennon cloud is optional for a validator.} receive succinct cryptographic proofs
-of these CI statements (STARK/SNARK) from the Argennon cloud and efficiently validate a block.
+of these CI statements (STARK/SNARK) from the Argennon cloud and validate blocks without storing the ledger state
+or performing costly computation.
 
@@
 different CI statements are independent of each other and can be done in parallel. As a result, a validator can use
-multiple cores for verifying CI statements of a \textbf{single} block and different validator committees can
+multiple cores for verifying CI statements of a \textbf{single} block and different validator assemblies can
 simultaneously and
@@
 could use more established blockchains as an extra layer of security, specially
-during the bootstrapping phase, when ARG is not well distributed yet. To reduce the execution fee, this validator
-contract could be used for validating roll-ups of state transition.
+during the bootstrapping phase, when ARG is not well distributed yet. In addition, this contract will facilitate
+trustless bridging of assets to the Argennon blockchain. To reduce the execution fee, this validator
+contract could be used for validating only roll-ups of state transition.
 

### 2022-08-20 tex/sections/intro.tex
Fix readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 security of the consensus protocol used in a blockchain. Unfortunately, many currently used blockchains are
-vulnerable to a certain type of consensus attack, known as the bribery attack. In a bribery attack, an adversary
-tries to corrupt participants of a protocol by offering them money and making them violate the protocol.
+vulnerable to a certain type of consensus attack, known as the Bribery Attack. In a Bribery Attack, an adversary
+tries to corrupt participants of a protocol by offering them money and seducing them to violate the protocol.
 
 At the time of writing of this document,
-the total mining reward for a bitcoin block is around \$150,000. If we assume, in decision theoretic terminology,
-that the mining reward accurately defines the utility function of a bitcoin miner, one could hire all hashing power
-of the bitcoin network for one hour by
+the total mining reward for a Bitcoin block is around \$150,000. If we assume, in decision theoretic terminology,
+that the mining reward accurately defines the utility function of a Bitcoin miner, one could hire all hashing power
+of the Bitcoin network for one hour by
 spending only \$750,000. The situation is not much different for PoS blockchains, as long as the total
-stake of the validator set is a relatively small value. Here by
-stake, we mean a real number measuring the total interest of a user in the system. We are not referring, in
+stake of the validator set is a relatively small value. By
+stake, we mean a real number measuring the total interest of a user in the system, and we are not referring, in
 particular, to some locked amount of a user's money that is known as stake in some PoS protocols.
-
 This problem is more severe in blockchains that use randomly selected small sets of validators. These small sets
@@
 common zero-day
-vulnerability, has the ability to infect a large portion of normal personal computers, it could be used by an
-adversary to control the majority of participants in the consensus protocol and compromise the security of the
+vulnerability, has the ability to infect a large portion of normal personal computers, an adversary can use it to
+take control of the majority of participants in the consensus protocol and compromise the security of the
 blockchain.
@@
 Securing a computer system against cyberattacks needs planning ahead and access to engineering resources.
-Special software and hardware, like custom-built operating systems and isolated specialized hardware is required.
-This is not something a normal user can afford. Only powerful centralized entities having large financial and
-technical resources, could build
+Special software and hardware, like custom-built operating systems and isolated specialized hardware are required.
+This is not something a normal user can afford. But powerful centralized entities, having large financial and
+technical resources, could actually build
 systems that are resilient against sophisticated cyberattacks. In this regard, we have to admit, a centralized system is
@@
 and any network partition resolves after a finite amount of time. The Recovery Protocol uses two main emergency
-procedures to recover the functionality of the Argennon blockchain: \emph{emergency forking} and \emph{emergency
-agreement} protocol.
+procedures to recover the functionality of the Argennon blockchain: \emph{Emergency Forking} and \emph{Emergency
+Agreement} protocol.
 
@@
 malicious.
-
-Centralized block generation brings some interesting features to the Argennon platform, such as flexible and lower
-fees, off chain fee payment, optimistic instant transaction confirmation, and front running protection.
-However, it also increases the possibility of transaction censorship. In Argennon, this problem is addressed by a
-special High Priority Request (HPR) protocol.
 
@@
 different CI statements are independent of each other and can be done in parallel. As a result, a validator can use
-multiple cores for verifying CI statements of a \textbf{single} block and different validator assemblies can
+multiple cores for verifying CI statements of a \textbf{single block} and different validator assemblies can
 simultaneously and
-independently verify \textbf{different} blocks.
+independently verify \textbf{different blocks}.
 In addition, proof generation of different CI statements similarly can be done in parallel. However, for parallel
@@
 
-Using succinct argument of knowledge systems makes the main functionalities of an Argennon validator light enough to
+Centralized block generation brings some interesting features to the Argennon platform, such as flexible and lower
+fees, off chain fee payment, optimistic instant transaction confirmation, and front running protection.
+However, it also increases the possibility of transaction censorship. In Argennon, this problem is addressed by a
+special High Priority Request (HPR) protocol.
+
+Using Succinct Argument of Knowledge systems makes the main functionalities of an Argennon validator light enough to
 be implemented as a smart contract. By deploying a validator contract on another platform, Argennon
@@
 trustless bridging of assets to the Argennon blockchain. To reduce the execution fee, this validator
-contract could be used for validating only roll-ups of state transition.
+contract can be used for validating only roll-ups of the state transition.
 

### 2022-08-21 tex/sections/intro.tex
Mention arithmetiztion in the intro
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 machine. This enables delegates to compile and optimize Argennon applications for their specific hardware platforms
-and execute applications efficiently, ensuring the state transition can be found as fast as possible.
+and execute applications efficiently, ensuring the state transition can be found as fast as possible. Moreover,
+validators by using a special compiler, generate arithmetization of this language and use it for verifying CI proofs.
 
@@
 during the bootstrapping phase, when ARG is not well distributed yet. In addition, this contract will facilitate
-trustless bridging of assets to the Argennon blockchain. To reduce the execution fee, this validator
-contract can be used for validating only roll-ups of the state transition.
+trustless bridging of assets from that platform to the Argennon blockchain. To reduce the execution fee, only
+roll-ups of the state transition can be validated by the contract.
 

### 2022-08-24 tex/sections/intro.tex
Mention the ASAR
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 stake of the validator set is a relatively small value. By
-stake, we mean a real number measuring the total interest of a user in the system, and we are not referring, in
+stake, we mean a real number measuring the total interest of a user in the system, we are not referring, in
 particular, to some locked amount of a user's money that is known as stake in some PoS protocols.
@@
 Special software and hardware, like custom-built operating systems and isolated specialized hardware are required.
-This is not something a normal user can afford. But powerful centralized entities, having large financial and
-technical resources, could actually build
+This is not something a normal user can afford. Powerful centralized entities, having large financial and
+technical resources, could build
 systems that are resilient against sophisticated cyberattacks. In this regard, we have to admit, a centralized system is
@@
 
-More formally, each CI statement has the form $\tau_{\mathbb{S}, \hat{\mathbb{S}}, \mathbb{R}}$, and states that
+More formally, each CI statement has the form
+$\hat{\mathbb{S}} \coloneqq \tau(\mathbb{S}, \mathbb{R})$, and states that
 $\hat{\mathbb{S}}$ is the commitment of the next state after executing an ordered list of external requests with
 commitment $\mathbb{R}$ on a state which has commitment $\mathbb{S}$.
-Commitments to states are produced by a cryptographic accumulator.
+Commitments to states are produced by a cryptographic accumulator. $\tau$ is a transition function and is known to
+both the validator and the prover.
+
 Validators can\footnote{Using the Argennon cloud is optional for a validator.} receive succinct cryptographic proofs
-of these CI statements (STARK/SNARK) from the Argennon cloud and validate blocks without storing the ledger state
-or performing costly computation.
+(STARK/SNARK) of these CI statements from the Argennon cloud and validate blocks without storing the ledger
+state or performing costly computation.
 
@@
 
-Argennon applications (i.e smart contracts) are stored in a high level text based language on the blockchain. This
-language is intended for preserving the
+On the blockchain, Argennon applications (i.e smart contracts) are stored in a high level text based language, called
+the Argennon Standard Application Representation (ASAR). The ASAR is intended for preserving the
 high level information of the application logic to facilitate platform specific compiler optimization at a host
@@
 during the bootstrapping phase, when ARG is not well distributed yet. In addition, this contract will facilitate
-trustless bridging of assets from that platform to the Argennon blockchain. To reduce the execution fee, only
+trustless bridging of assets from that platform to the Argennon blockchain. To reduce execution fee, only
 roll-ups of the state transition can be validated by the contract.

### 2022-09-02 tex/A.tex
Add the APM
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 \documentclass[11pt, a4paper]{report}
-\usepackage[a4paper]{geometry}
 
@@
 \usepackage{mathtools}
+\usepackage[a4paper]{geometry}
+\usepackage{tikz}
 
@@
 
-    \chapter{The Argennon Smart Contract Execution Environment}\label{ch:AVM}
-    \note{This chapter is outdated and does not include CI proofs. Currently a new version is being rewritten.}
+    \chapter{The Argennon Smart Contract Execution Environment}\label{ch:AscEE}
     \input{sections/avm}
+
+    \chapter{The Argennon Prover Machine}\label{ch:APM}
+    \input{sections/apm}
 
@@
     Unlike conventional blockchains, Argennon does not use a P2P network architecture. Instead, it uses a
-    client-server topology, based on a permission-less list of ZK-EDB servers. ZK-EDB servers are a
+    client-server topology, based on a permission-less network of PVC servers. PVC servers are a
     crucial part of the Argennon ecosystem, and they form the backbone of the Argennon networking layer.

### 2022-09-02 tex/sections/apm.tex
Add the APM
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

%! Author = aybehrouz
%! Date = 9/2/22

The Argennon Prover Machine (APM) is a virtual machine tailored for efficient verification of AscEE computations by
argument systems. The APM has a minimal RISC architecture with a very compact instruction set. This ensures that
its transition function has an efficient circuit complexity. Here by circuit complexity we mean the size
of the smallest circuit that, given two adjacent states in the trace, verifies that the transition between the two
states indeed respects the APM specification. The APM is a stack machine and has a random access key-value memory.
The word size of the APM is 64-bit.

The Argennon Prover Machine gets as its primary input a vector $(\mathbf{C}_H,\mathbf{C}_P,\mathbf{C}_R)$, where
$\mathbf{C}_H$ is a commitment to the AscEE Heap, $\mathbf{C}_P$ is a commitment to the AscEE Program Area and
$\mathbf{C}_R$ is a commitment to an ordered list of requests. The APM then outputs a final
$\mathbf{accept}$ flag which indicates if the execution has ended successfully and also outputs the updated
commitments to the AscEE Heap and Program Area: $(\mathbf{C}_{H'},\mathbf{C}_{P'},\mathbf{accept})$.

Producing the required outputs from these inputs is not computationally feasible so the APM accepts an auxiliary
input vector $(H,\pi_H,P,\pi_P,R,\pi_R,h)$, where $\pi_X$ is a proof that proves $X$ can be opened w.r.t
$\mathbf{C}_X$. The auxiliary input $h$ is a hint that helps the APM make nondeterministic choices correctly.

The APM consists of four main modules:
\begin{itemize}
    \item \textbf{Preprocessor:} This module prepares the input data for the NEU and PRU. It verifies that $H,
    P,R$ are valid w.r.t the provided commitments. It also processes the input data and ensures it has the
    correct format and is valid. For example it verifies the signatures of authenticated messages or
    verifies that the proposed chunk bounds are valid.
    \item \textbf{Normal Execution Unit (NEU):} This module executes the requests whose execution completes
    normally. It outputs the updated heap chunks and an $accept$ flag. More formally it evaluates $(H',accept)=P(H,
    R_{good})$, where $H'$ is the updated heap chunks and the $accept$ flag will be set to false, If an
    application terminates abruptly.
    \item \textbf{Failure Repeater Unit (FRU):} This module evaluates $accept=P(H',R_{bad})$, where $R_{bad}$ is
    those requests whose execution terminates abnormally, $H'$ is the output of the NEU and the $accept$ flag is set
    to false if a request execution completes normally.
    \item \textbf{Postprocessor:} This module is responsible for calculating the updated commitments
    $(\mathbf{C}_{H'}, \mathbf{C}_{P'})$ and the final $\mathbf{accept}$ flag. Installing new applications or
    updating the code of existing application is performed by this module.
\end{itemize}

The configuration of the APM can be considered as a quadruple
$(\mathbb{S},\mathbb{L},\mathbb{T}_{NEU},\mathbb{T}_{FRU})$,
where $\mathbb{S}$ is the size of the internal stack, $\mathbb{L}$ is the amount of local memory and
$\mathbb{T}_{NEU}, \mathbb{T}_{FRU}$ are the number of cycles that the Normal Execution Unit (NEU)
and Failure Repeater Unit (FRU) runs for. It should be noted that the APM has one local memory which
is used for all application calls.

%
%\begin{tikzpicture}
%    \draw (0,0) -- (4,0) -- (4,4) -- (0,4) -- (0,0);
%\end{tikzpicture}

### 2022-09-02 tex/sections/avm.tex
Add the APM
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The Argennon Smart Contract Execution
-Environment (AscEE) is an abstract execution environment for executing Argennon smarts contracts (a.k.a\ Argennon
-applications) in a safe and secure environment. An Argennon application essentially is an HTTP server whose state is
-kept in the Argennon blockchain and its logic is described using an Argennon Standard Representation (ASR).
+The Argennon Smart Contract Execution Environment (AscEE) is an abstract execution environment for executing Argennon
+smarts contracts (a.k.a\ Argennon applications) in a secure and isolated environment. An Argennon application
+essentially is an HTTP server whose state is
+kept in the Argennon blockchain and its logic is described using an Argennon Standard Application Representation (ASAR).
 
-An Argennon Standard Representation (ASR) is a programming language for describing Argennon applications, optimized
-for the architecture and properties of the Argennon platform.
-Argennon supports two standard representations: One is a high level text based representation which needs
-costly compilation before being executed on a hardware machine. The other is a low level binary representation which
-usually can be executed, with minimal pre-processing, by a JIT compiler or an emulator. The high level
-language is intended for preserving the high level information of the application logic to facilitate
-platform specific compiler optimization at a host machine. The low level language, on the oder hand, is designed for
-efficient direct execution of applications that are not frequently used and do not benefit from compiler optimizations.
+An Argennon Standard Application Representation (ASAR) is a programming language for describing Argennon
+applications, optimized
+for the architecture and properties of the Argennon platform. This text based representation should be low level
+enough to enable easy compilation from any high level programming language and should be high level enough to preserve
+the high level information of the application logic and facilitate platform specific compiler optimization at a host
+machine. In this regard, the ASAR can be considered as an Intermediate Representation (IR).
 
@@
 resize its heap chunks or allocate new chunks. Every chunk belongs to exactly one application and can only be modified
-by its owner. In addition to heap chunks, every application has an amount of non-persistent local memory for
+by its owner. In addition to heap chunks, every application has a limited amount of non-persistent local memory for
 storing temporary data.
@@
 blockchain in a three-step procedure. The first step is the \emph{preprocessing step}. In this step, the required
-data for executing requests are retrieved from the Argennon cloud, the cryptographic proofs are verified and the
+data for executing requests are retrieved and verified and the
 helper data structures for next steps are constructed. This step is
@@
 The Argennon Smart Contract Execution Environment can be seen as a machine for executing Argennon applications to
-fulfill HTTP requests, produce their HTTP responses and update related heap chunks. The execution of requests can be
-considered sequential\footnote{Actually requests are executed in parallel but by performing data dependency analysis the
-result is guaranteed to be identical with sequential execution of requests.} and each request has a separate
-\emph{execution session}. Therefore, an execution
-session is a separate session of executing smart contract's code in order to fulfill an \emph{external} HTTP request.
-External requests are requests that are not made by other Argennon applications.
+fulfill \emph{external} HTTP requests\footnote{External requests are requests that are not made by other Argennon
+applications.}, produce their HTTP responses and update related heap chunks. The execution of requests can be
+considered sequential\footnote{Actually requests may be executed in parallel but by performing data dependency analysis
+the result is guaranteed to be identical with the sequential execution of requests.} and each request has a separate
+\emph{Execution Session}. Therefore, an Execution Session is a separate session of executing application's code in
+order to fulfill an External HTTP request. We call this external request the \emph{initiator} of the Execution Session.
 
-The state of an execution session will be
+The state of an Execution Session will be
 destroyed at the end of the session and only the state of heap chunks is preserved. If a session fails and does not
-complete normally, it will not have any effect on any heap chunks.
+complete normally, it will not have any effects on any heap chunks.
 
-During an execution session an application can make \emph{internal} HTTP requests to other applications. Those
-requests will not start a new execution session and will be executed within the current session. In AscEE making an
+During an Execution Session an application can make \emph{internal} HTTP requests to other applications. Those
+requests will not start a new Execution Session and will be executed within the current session. In AscEE making an
 HTTP request to an application is similar to a function invocation, and for that reason, we also refer to them as
@@
 
-The AscEE is designed based on \emph{optional decoupling principle}. When an application makes a request to another
-application, optionally it can choose to be decoupled from the called application. That would mean the called
-application could not affect its caller's state by reentrancy, or could not abort the session by using excessive
-resources or performing illegal operations.
+\section{Memory}\label{mem}
+
+Every Argennon application has two types of memory: local memory and heap. Local memory is not persistent and is
+destroyed when the application call ends. Local memory is used for storing local variables and is not directly
+addressable. Heap, on the other hand, is persistent and can be used for persisting data between application calls.
+Heap is addressable and provides low level direct access. Both local memory and heap are limited, but the
+AscEE does not specify any particular limits for them. When an application tries to use too much memory, that may
+cause the execution session to end abruptly. In that case, the Execution Session will not have any effects on the
+state of the heap.
+
+\subsection{Heap Chunks}\label{subsec:heap}
+
+The AscEE heap is split into chunks. Each heap chunk is a continuous finite array of bytes, has a unique identifier, and
+is byte addressable. An application may have several heap chunks with different sizes and can remove or
+resize its chunks or allocate new ones. Every chunk belongs to exactly one application. Only the owner application can
+modify a chunk but there is no restrictions for reading a chunk\footnote{The reason behind this type of access
+control design is the fact that smart contract
+code is usually immutable. That means if a smart contract does not implement a
+getter mechanism for some parts of its internal data, this functionality can never
+be added later, and despite the internal data is publicly available, there will be no
+way for other smart contracts to use this data on-chain.}.
+
+When an application allocates a new heap chunk, the identifier of the new chunk is not generated by
+the AscEE. Instead, the application can choose an identifier itself, provided it has a correct format. This
+is an important feature of the AscEE heap which allows applications to use the AscEE heap as a map
+data structure\footnote{also called a dictionary.}.
+Since the \texttt{chunkID} is a prefix code, any application has its own identifier space, and an application
+can easily find unique identifiers for its chunks.
+
+During an Execution Session every heap chunk has an access type which may disallow certain operations. This access
+type is declared by the initiator request of the Execution Session:
+
+\begin{itemize}
+    \item \texttt{check\_only}: only allows check operations. These operations query the persistence
+    status of a memory location.
+    \item \texttt{read\_only}: only allows read and check operations.
+    \item \texttt{writable}: allows reading and writing.
+    \item \texttt{additive}: only allows additive operations. By additive we mean an addition-like operator without
+    overflow checking which is associative and commutative. Note that the content of these access blocks cannot be read.
+\end{itemize}
+
+\subsubsection{Chunk Resizing}\label{subsubsec:ch-resize}
+
+At the start of executing requests of a block, a validator can consider two values for every
+heap chunk, the size: \texttt{chunkSize} and a size upper bound: \texttt{sizeUpperBound}. The value of
+\texttt{chunkSize} can be determined uniquely at the start of
+every Execution Session, and it may be updated during the session by the owner application. On the other hand,
+the value of \texttt{sizeUpperBound} proposed by the block proposer for each block and is calculated based on values
+declared by requests (i.e.\ transactions) declaring chunk resizing. This value needs to be an upper bound of all the
+declared values and indicates the upper bound of \texttt{chunkSize} during the execution of requests of a block.
+
+The value of \texttt{chunkSize}, can be modified during an Execution Session. However, the new size values can only be
+increasing or decreasing. More precisely, if a request declares that it wants to expand (shrink) a chunk, it can only
+increase (decrease) the value of \texttt{chunkSize} and any specified value during the execution
+session, needs to be greater (smaller) than the previous value of the chunk's size. Any request that wants to expand
+(shrink) a chunk needs to specify a max size (min size). The value of \texttt{chunkSize} can not be set higher
+(lower) than this value.
+
+The value of \texttt{chunkSize} at the end of an Execution Session will determine if a memory location at an
+offset is persistent or not: Offsets lower than the chunk size are persistent, and higher offsets are not.
+Non-persistent locations will be re-initialized with zero at the start of every Execution Session.
+
+Usually an application should not have any assumption about the content of memory locations that are outside the chunk.
+While these locations are zero initialized at the start of every Execution Session, it should be noted that multiple
+invocations of an application may occur in a single Execution Session, and if one of them modifies a location outside
+the chunk, the changes can be seen by next invocations.
+
+There is no way for an application to query \texttt{sizeUpperBound}. As a result, for an application,
+accessing offsets higher than \texttt{chunkSize} results in undefined behaviour, while the behaviour is well-defined
+in the view of validators.
+This enables validators to determine the validity of an offset at the start of the block validation in a parallelized
+preprocessing phase without actually executing requests.
+
+While an application can use \texttt{chunkSize} to determine if an offset is persistent or not, that is not
+considered a good practice. Reading \texttt{chunkSize} decreases transaction parallelization, and should be avoided.
+Instead, applications should use a built-in AscEE function for checking the persistence status of memory addresses.
+
+An application can load any chunk with a valid prefix identifier even if that chunk does not exist. For a non-existent
+chunk the value of \texttt{chunkSize} is always zero.
+
+The address space of a chunk starts from zero and only offsets lower than \texttt{sizeUpperBound} are valid. Trying to
+access any offset higher than this value will always result in a revert for the application.
 
@@
 
-\section{Heap Chunks}\label{heap}
-
-The persistent data of an Argennon application is stored in continuous finite
-arrays of bytes, called heap chunks. An application may have several heap chunks
-with different sizes, and can remove or
-resize its chunks or allocate new chunks. Every chunk belongs to exactly one application. Only the owner application can
-modify a chunk but there is no restrictions for reading a chunk\footnote{The reason behind this type of access
-control design is the fact that smart contract
-code is usually immutable. That means if a smart contract does not implement a
-getter mechanism for some parts of its internal data, this functionality can never
-be added later, and despite the internal data is publicly available, there will be no
-way for other smart contracts to use this data on-chain.}.
-
-When an application allocates a new heap chunk, the identifier of the new chunk is not generated by
-the AscEE. Instead, the application can choose an identifier itself, provided it has a correct format. This
-is an important feature of the AscEE heap, which allows applications to use the AscEE heap as a dictionary
-data structure\footnote{also called a map.}.
-Since the \texttt{chunkID} is a prefix code, any application has its own identifier space, and an application
-can easily find unique identifiers for its chunks.
-
-\subsection{Chunk Resizing}\label{subsec:ch-resize}
-
-At the start of executing requests of a block, a validator can consider two values for every
-heap chunk, the size: \texttt{chunkSize} and a size upper bound: \texttt{sizeUpperBound}. The value of
-\texttt{chunkSize} can be determined uniquely at the start of
-every execution session, and it may be updated during the session by the owner application. On the other hand,
-the value of \texttt{sizeUpperBound} is determined and fixed for each block and is proposed by the block
-proposer. This value indicates the upper bound of \texttt{chunkSize} during the execution of
-requests of the block, and it can be used by the validator for safe memory allocation.
-
-The value of \texttt{chunkSize}, can be modified during an execution session. However, the new size values can only be
-increasing or decreasing. More precisely, if a request declares that it wants to expand (shrink) a chunk, it can only
-increase (decrease) the value of \texttt{chunkSize} and any specified value during the execution
-session, needs to be greater (smaller) than the previous value of the chunk's size. Any request that wants to expand
-(shrink) a chunk needs to specify a max size (min size). The value of \texttt{chunkSize} can not be set higher
-(lower) than this value.
-
-The value of \texttt{chunkSize} at the end of an execution session will determine if a memory location at an
-offset is persistent or not: Offsets lower than the chunk size are persistent, and higher offsets are not.
-Non-persistent locations will be re-initialized with zero at the start of every execution session.
-
-Usually an application should not have any assumption about the content of memory locations that are outside the chunk.
-While these locations are zero initialized at the start of every execution session, it should be noted that multiple
-invocations of an application may occur in a single execution session, and if one of them modifies a location outside
-the chunk, the changes can be seen by next invocations.
-
-There is no way for an application to query \texttt{sizeUpperBound} of a chunk. As a result, for an application,
-accessing offsets higher than \texttt{chunkSize} results in undefined behaviour, while the behaviour is well-defined
-in the view of validators.
-This enables validators to determine the validity of an offset at the start of the block validation in a parallelized
-preprocessing phase without actually executing requests.
-
-While an application can use \texttt{chunkSize} to determine if an offset is persistent or not, that is not
-considered a good practice. Reading \texttt{chunkSize} decreases transaction parallelization, and should be avoided.
-Instead, applications should use a built-in AscEE function for checking the persistence status of memory addresses.
-
-An application can load any chunk with a valid prefix identifier even if that chunk does not exist. For a non-existent
-chunk the value of \texttt{chunkSize} is always zero.
-
-The address space of a chunk starts from zero and only offsets lower than \texttt{sizeUpperBound} are valid. Trying to
-access any offset higher than this value will always result in a revert for the application.
-
-\subsection{Access Blocks}\label{subsec:access-blocks}
-
-Memory locations inside a chunk can only be accessed through access blocks. An access block is an access interval
-defined on a chunk, which determines the accessible memory locations inside the chunk. Each access block has an offset
-and a size. Multiple access blocks can be defined on a single chunk, but they must be non overlapping.
-Access blocks are byte addressable and may have different access types:
-
-\begin{itemize}
-    \item \texttt{check\_only}: only allows check operations. These operations query the persistence
-    status of a memory location.
-    \item \texttt{read\_only}: only allows read and check operations.
-    \item \texttt{writable}: allows reading and writing.
-    \item \texttt{additive}: only allows additive operations. These operations are addition-like operations without
-    any overflow checking. These operations must be associative and commutative. Note that the content of
-    these access blocks cannot be read.
-\end{itemize}
-
-Chunks and access blocks have different purposes. Chunks are intended for simplifying proof checking of the data
-stored in the Argennon cloud\footnote{Note that
-the Argennon cloud uses cryptographic accumulators for preserving data integrity.} and access blocks are
-required for better parallelization of the request execution. An application should put the data needed for
-validating a single block in the same chunk and the data needed in a single execution session in the same access block.
-
 \section{Request Attachments}\label{sec:attachments}
@@
 
-\section{Authorizing Operations}\label{sec:authorizing-operations}
+\section{Authorization}\label{sec:auth}
 
 In blockchain applications, we usually need to authorize certain operations. For example, for sending an asset
-from a user to another user, first we need to make sure that the sender has authorized this operation.
+from a user to another user, we need to make sure that the sender has authorized that operation.
 
-The AscEE uses \emph{Authenticated Message Passing} for authorizing operations. In this method, every execution
-session has a set of authenticated messages, and those messages are \textbf{explicitly} passed in requests to
+The AscEE uses \emph{Authenticated Message Passing} for authorizing operations. In this method, every Execution
+Session has a set of authenticated messages, and those messages are \textbf{explicitly} passed in requests to
 applications for authorizing operations. These messages act exactly like digital signatures and applications can
-ensure that they are issued by specific accounts. The only difference is that the process of message authentication
-is performed by the AscEE internally and is not performed by the application. For accounts, the AscEE uses digital
-signatures to authenticate messages. Signatures for each request can be validated in parallel during the
-preprocessing step of the block validation, and any type of
-cryptographic signature scheme can be used. For example, we could use BLS aggregate signatures to authenticate all
-messages of a block of the blockchain in bulk.
+ensure that they are issued by their claimed issuer accounts. The only difference is that the process of
+message authentication is performed by the AscEE internally and an application does not explicitly verify cryptographic
+signatures.
 
-Moreover, applications can use built-in functions of the AscEE to generate authenticated messages in run-time.
+Each Execution Session has a list of authenticated messages. Each authenticated message has an index which will be
+used for passing the message to an application as a request parameter. The AscEE uses cryptographic signatures to
+authenticate messages for user accounts. The signatures are validated in parallel during the
+preprocessing step of the block validation, and any type of cryptographic signature scheme can be used.
+
+Also, applications can use built-in functions of the AscEE to generate authenticated messages in run-time.
 This enables an application to authorize operations for another application even if it is not calling that
@@
 In addition to authenticated messages, the AscEE provides a set of
-cryptographic functions for validating signatures and cryptographic entities. By using these functions and
+cryptographic functions for validating signatures and calculating cryptographic entities. By using these functions and
 passing cryptographic signatures as parameters to methods, a programmer, having users' public keys, can implement
@@
 
-Authorization by explicit authenticated messages and signatures eliminates the need for approval mechanisms or call back
-patterns in Argennon.
-
-\note{The AscEE has no instructions for issuing cryptographic signatures.}
+Authorizing operations by Authenticated Message Passing and explicit signatures eliminates the need for approval
+mechanisms or call back patterns in Argennon.\footnote{The AscEE has no instructions for issuing cryptographic
+signatures.}
 
@@
 
-The Argennon Smart Contract Execution Environment provides optional low level reentrancy protection by providing low
+The AscEE provides optional low level reentrancy protection by providing low
 level \emph{entrance locks}. When an application acquires an entrance lock it cannot acquire that lock again and trying
 to do so will result in a revert. The entrance lock of an application will be released when the application explicitly
-releases its lock or when the call that had acquired that lock completes.
+releases its lock or when the call that has acquired that lock completes.
 
@@
 \ldots
-
-\section{Resource Management}\label{sec:res-man}
-
-Completing an execution session requires computational resources. The amount of resources used by an execution session
-should be monitored and managed, otherwise a malicious user would be able to easily spam and exhaust resources of the
-execution environment.
-
-In most consensus protocols, we can assume that the block proposer has enough incentive to filter out transactions
-that spam run-time resources. Here by a run-time resource, we mean a resource that at run-time, a limited amount
-of it is available, but its surplus can not be stored for later use. Execution time and local memory are
-examples of such a resource but permanent storage is not.
-
-If a proposed block contains many transactions which need a lot of run-time resources, validators would not
-be able to validate all transactions in a timely manner. Consequently, they may decide to reject the block or if they
-spend enough resources, the confirmation of that block could take more than usual. Longer block time is not
-favoured by block proposers, because it means less throughput of the system which usually means less overall rewards
-for them. In the Argennon protocol the management of run-time resources, is left to the block proposer.
-
-Resource usage can be measured per session or per application call. Obviously per session measurement is
-easier and more efficient. However, when we measure resources per session if
-a session violates its resource caps, determining the point of failure may require precise and error-free resource
-measurement. For example, assume that a session containing an application call
-violates a 2 milliseconds execution time cap. If in the caller's code, the call happens exactly after 2 milliseconds,
-a small fluctuation in the execution time measurement can
-change the point of failure between the called and the caller application. Note that for implementing optional
-decoupling principle in Argennon, we need to determine the
-exact application call which has failed in a call chain, and this fluctuation introduces a nondeterministic
-behaviour which could make block validation impossible. That's why we need per application call
-resource management sometimes. If we define the resource caps per application call and
-perform our measurements for each application call separately, by using caps that are larger than the measurement
-error, errors in the measurements would not change the point of failure in the call chain.
-
-The AscEE has two types of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
-session is always measured per session and \textbf{default} pre-defined resource caps are used. On the other hand, for a
-monitored session, resources that can not be measured precisely are measured per application call and their caps are
-determined per application call by the external HTTP request (i.e.\ transaction). The block proposer decides the
-type of the execution session for each transaction.
-
-Different computational resources are measured and monitored during an AscEE session:
-\begin{itemize}
-    \item \textbf{execution time}:
-    is the amount of cpu time that is required for executing a session or an application call. The execution time is
-    measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
-    of \textbf{cpu time} needed for executing a predefined standard application which is used for benchmarking a host's
-    performance. In other words, by definition, the AscEE imaginary standard machine completes the standard benchmark
-    in 1000 clocks.
-
-    Optimistic sessions have a predefined \texttt{maxClocks} value which is determined by the Argennon protocol. This
-    value defines a bound on the \textbf{total} AscEE clocks of the session, and no per application measurement is done.
-
-    Monitored sessions perform per application call cpu-time measurement, and every application call during a monitored
-    session has a separate \texttt{maxClocks} value. This value determines the maximum amount of time that the cpu
-    can be used for executing that particular application call. It should be noted that the cpu timer of the
-    application call is paused when the application makes a call to another application, and is resumed when the
-    control returns. An application call
-    needs at least 10 clocks and if the value of its \texttt{maxClock} is lower than this value the call will be
-    considered a failed call.
-
-    Each application call has some amount of \texttt{externalClocks}. When an application makes a request to another
-    application it has to \emph{forward} a portion of its \textbf{external} clocks to the called application. This
-    amount will determine the value of \texttt{maxClocks} for the called application, and is subtracted from the value
-    of \texttt{externalClocks} of the caller.
-
-    The amount of external clocks of an application
-    call is defined to be \(2/3\) of its \texttt{maxClocks}. As a result, the total number of clocks of a monitored
-    session is always less than \(3 \times \texttt{maxClocks}\) of the root application call. The value of
-    \texttt{maxClocks} for
-    the root application call is determined by the external request (i.e.\ transaction).
-
-    \item \textbf{local memory}:
-    any memory usage of an application that is not part of a heap chunk and is not part of another measured resource
-    will be considered as local memory usage. Local memory is not persistent and when an application finishes serving a
-    request and returns the HTTP response (i.e.\ the application call completes) its local memory is deleted.
-    This resource is measured in bytes.
-
-    Optimistic sessions measure local memory usage per session and enforce a protocol-defined cap on the total amount
-    of local memory a session can use. Monitored sessions measure local memory usage per
-    application call and enforce a protocol-defined cap for each application call separately. An application call
-    which tries to use more local memory than the cap, fails.
-    \item \textbf{heap access list}:
-    every session can only access heap locations that are declared in its access list. In addition,
-    resizing heap chunks can only be done in the range of the pre-declared lower bound and upper bound.
-    \item \textbf{app access list}:
-    a session may only make requests to applications that are declared in its application access list.
-    \item \textbf{call depth}:
-    during a session the number of nested application calls can not be more than a threshold. This threshold is
-    determined by the Argennon protocol. It should be noted that a differed call is considered like a normal call and
-    increases the call depth by one level.
-    \item \textbf{differed calls}:
-    every application call can have a limited number of differed calls which is determined by the protocol.
-    To simplify the implementation, this limit is defined per application call. Since in Argennon the call depth is
-    limited, a per-application call limit will also define an implicit limit for the total number of active differed
-    calls.
-    \item \textbf{virtual signatures}:
-    \item \textbf{number of entrance locks}:
-\end{itemize}
-
-In Argennon, execution time and local memory are considered \emph{nondeterministic} resources. A nondeterministic
-resource is a resource that can not be measured precisely and its measurement always contains a random error.
-Optimistic sessions are not allowed to fail because of violating nondeterministic resource limits. As a result, the
-block proposer must always choose a monitored session for a transaction that is included in the block and violates a
-nondeterministic resource cap.
-
-When a transaction fails due to the violation of a limit for a nondeterministic resource, the proposer is
-required to exactly specify the application call which violates that limit in the execution session. When validators are
-executing a monitored session, for each application call, they enforce considerably larger limits for
-nondeterministic resources. Only in case the proposer has declared that an application call violates a limit, the
-validators will enforce the actual value of the limit. This simple mechanism ensures that, with a very high
-probability, validators agree with the proposer, although their measurements could be different.
 

### 2022-09-02 tex/sections/elements.tex
Add the APM
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 \end{lstlisting}
+
+\section{Resource Management}\label{sec:res-man}
+
+Completing an Execution Session requires computational resources. The amount of resources used by an Execution Session
+should be monitored and managed, otherwise a malicious user would be able to easily spam and exhaust resources of the
+execution environment.
+
+In most consensus protocols, we can assume that the block proposer has enough incentive to filter out transactions
+that spam run-time resources. Here by a run-time resource, we mean a resource that at run-time, a limited amount
+of it is available, but its surplus can not be stored for later use. Execution time and local memory are
+examples of such a resource but permanent storage is not.
+
+If a proposed block contains many transactions which need a lot of run-time resources, validators would not
+be able to validate all transactions in a timely manner. Consequently, they may decide to reject the block or if they
+spend enough resources, the confirmation of that block could take more than usual. Longer block time is not
+favoured by block proposers, because it means less throughput of the system which usually means less overall rewards
+for them. In the Argennon protocol the management of run-time resources, is left to the block proposer.
+
+Resource usage can be measured per session or per application call. Obviously per session measurement is
+easier and more efficient. However, when we measure resources per session if
+a session violates its resource caps, determining the point of failure may require precise and error-free resource
+measurement. For example, assume that a session containing an application call
+violates a 2 milliseconds execution time cap. If in the caller's code, the call happens exactly after 2 milliseconds,
+a small fluctuation in the execution time measurement can
+change the point of failure between the called and the caller application. Note that for implementing optional
+decoupling principle in Argennon, we need to determine the
+exact application call which has failed in a call chain, and this fluctuation introduces a nondeterministic
+behaviour which could make block validation impossible. That's why we need per application call
+resource management sometimes. If we define the resource caps per application call and
+perform our measurements for each application call separately, by using caps that are larger than the measurement
+error, errors in the measurements would not change the point of failure in the call chain.
+
+The AscEE has two types of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
+session is always measured per session and \textbf{default} pre-defined resource caps are used. On the other hand, for a
+monitored session, resources that can not be measured precisely are measured per application call and their caps are
+determined per application call by the external HTTP request (i.e.\ transaction). The block proposer decides the
+type of the Execution Session for each transaction.
+
+Different computational resources are measured and monitored during an AscEE session:
+\begin{itemize}
+    \item \textbf{execution time}:
+    is the amount of cpu time that is required for executing a session or an application call. The execution time is
+    measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
+    of \textbf{cpu time} needed for executing a predefined standard application which is used for benchmarking a host's
+    performance. In other words, by definition, the AscEE imaginary standard machine completes the standard benchmark
+    in 1000 clocks.
+
+    Optimistic sessions have a predefined \texttt{maxClocks} value which is determined by the Argennon protocol. This
+    value defines a bound on the \textbf{total} AscEE clocks of the session, and no per application measurement is done.
+
+    Monitored sessions perform per application call cpu-time measurement, and every application call during a monitored
+    session has a separate \texttt{maxClocks} value. This value determines the maximum amount of time that the cpu
+    can be used for executing that particular application call. It should be noted that the cpu timer of the
+    application call is paused when the application makes a call to another application, and is resumed when the
+    control returns. An application call
+    needs at least 10 clocks and if the value of its \texttt{maxClock} is lower than this value the call will be
+    considered a failed call.
+
+    Each application call has some amount of \texttt{externalClocks}. When an application makes a request to another
+    application it has to \emph{forward} a portion of its \textbf{external} clocks to the called application. This
+    amount will determine the value of \texttt{maxClocks} for the called application, and is subtracted from the value
+    of \texttt{externalClocks} of the caller.
+
+    The amount of external clocks of an application
+    call is defined to be \(2/3\) of its \texttt{maxClocks}. As a result, the total number of clocks of a monitored
+    session is always less than \(3 \times \texttt{maxClocks}\) of the root application call. The value of
+    \texttt{maxClocks} for
+    the root application call is determined by the external request (i.e.\ transaction).
+
+    \item \textbf{local memory}:
+    any memory usage of an application that is not part of a heap chunk and is not part of another measured resource
+    will be considered as local memory usage. Local memory is not persistent and when an application finishes serving a
+    request and returns the HTTP response (i.e.\ the application call completes) its local memory is deleted.
+    This resource is measured in bytes.
+
+    Optimistic sessions measure local memory usage per session and enforce a protocol-defined cap on the total amount
+    of local memory a session can use. Monitored sessions measure local memory usage per
+    application call and enforce a protocol-defined cap for each application call separately. An application call
+    which tries to use more local memory than the cap, fails.
+    \item \textbf{heap access list}:
+    every session can only access heap locations that are declared in its access list. In addition,
+    resizing heap chunks can only be done in the range of the pre-declared lower bound and upper bound.
+    \item \textbf{app access list}:
+    a session may only make requests to applications that are declared in its application access list.
+    \item \textbf{call depth}:
+    during a session the number of nested application calls can not be more than a threshold. This threshold is
+    determined by the Argennon protocol. It should be noted that a differed call is considered like a normal call and
+    increases the call depth by one level.
+    \item \textbf{differed calls}:
+    every application call can have a limited number of differed calls which is determined by the protocol.
+    To simplify the implementation, this limit is defined per application call. Since in Argennon the call depth is
+    limited, a per-application call limit will also define an implicit limit for the total number of active differed
+    calls.
+    \item \textbf{virtual signatures}:
+    \item \textbf{number of entrance locks}:
+\end{itemize}
+
+In Argennon, execution time and local memory are considered \emph{nondeterministic} resources. A nondeterministic
+resource is a resource that can not be measured precisely and its measurement always contains a random error.
+Optimistic sessions are not allowed to fail because of violating nondeterministic resource limits. As a result, the
+block proposer must always choose a monitored session for a transaction that is included in the block and violates a
+nondeterministic resource cap.
+
+When a transaction fails due to the violation of a limit for a nondeterministic resource, the proposer is
+required to exactly specify the application call which violates that limit in the Execution Session. When validators are
+executing a monitored session, for each application call, they enforce considerably larger limits for
+nondeterministic resources. Only in case the proposer has declared that an application call violates a limit, the
+validators will enforce the actual value of the limit. This simple mechanism ensures that, with a very high
+probability, validators agree with the proposer, although their measurements could be different.

### 2022-09-02 tex/sections/incentive.tex
Add the APM
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 For our incentive mechanism, we require that every time a validator receives a storage page from a PVC, after
-validating the data, he gives a receipt to the PVC server. In this receipt the validator signs the
+validating the data, he give a receipt to the PVC server. In this receipt the validator signs the
 following information:

### 2022-09-02 tex/sections/intro.tex
Add the APM
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 machine. This enables delegates to compile and optimize Argennon applications for their specific hardware platforms
-and execute applications efficiently, ensuring the state transition can be found as fast as possible. Moreover,
-validators by using a special compiler, generate arithmetization of this language and use it for verifying CI proofs.
+and execute applications efficiently, ensuring the state transition can be found as fast as possible.
+
+In addition to the ASAR, on the blockchain an Argennon application is stored in the Argennon Prover Machine (APM) code.
+This code is used by validators for CI proof verification. The Argennon Prover Machine is a virtual
+machine tailored for efficient verification of AscEE computations by
+argument systems. The APM has a minimal RISC architecture with a very compact instruction set. This ensures that
+its transition function has an efficient circuit complexity.
 

### 2022-09-03 tex/sections/apm.tex
Fix typos and readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The Argennon Prover Machine gets as its primary input a vector $(\mathbf{C}_H,\mathbf{C}_P,\mathbf{C}_R)$, where
-$\mathbf{C}_H$ is a commitment to the AscEE Heap, $\mathbf{C}_P$ is a commitment to the AscEE Program Area and
-$\mathbf{C}_R$ is a commitment to an ordered list of requests. The APM then outputs a final
-$\mathbf{accept}$ flag which indicates if the execution has ended successfully and also outputs the updated
-commitments to the AscEE Heap and Program Area: $(\mathbf{C}_{H'},\mathbf{C}_{P'},\mathbf{accept})$.
+$\mathbf{C}_H$ is a commitment to the AscEE Heap, $\mathbf{C}_P$ is a commitment to the AscEE program area and
+$\mathbf{C}_R$ is a commitment to an ordered list of requests. The APM then outputs the updated
+commitments to the AscEE heap and program area and a final $\mathbf{accept}$ flag which indicates if the execution has
+ended successfully or not: $(\mathbf{C}_{H'},\mathbf{C}_{P'},\mathbf{accept})$.
 
-Producing the required outputs from these inputs is not computationally feasible so the APM accepts an auxiliary
+Producing the required outputs from these inputs is not computationally feasible so the APM receives an auxiliary
 input vector $(H,\pi_H,P,\pi_P,R,\pi_R,h)$, where $\pi_X$ is a proof that proves $X$ can be opened w.r.t
-$\mathbf{C}_X$. The auxiliary input $h$ is a hint that helps the APM make nondeterministic choices correctly.
+$\mathbf{C}_X$ and $h$ is a hint that helps the APM make nondeterministic choices correctly.
 
@@
 \begin{itemize}
-    \item \textbf{Preprocessor:} This module prepares the input data for the NEU and PRU. It verifies that $H,
+    \item \textbf{Preprocessor:} This module prepares the input data for other modules. It verifies that $H,
     P,R$ are valid w.r.t the provided commitments. It also processes the input data and ensures it has the

### 2022-09-03 tex/sections/avm.tex
Fix typos and readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The Argennon Smart Contract Execution Environment (AscEE) is an abstract execution environment for executing Argennon
-smarts contracts (a.k.a\ Argennon applications) in a secure and isolated environment. An Argennon application
+The Argennon Smart Contract Execution Environment (AscEE) is an abstract high level execution environment for executing
+Argennon smarts contracts (a.k.a\ Argennon applications) in a secure and isolated environment. An Argennon application
 essentially is an HTTP server whose state is
@@
 
-An Argennon Standard Application Representation (ASAR) is a programming language for describing Argennon
+An Argennon Standard Application Representation is a programming language for describing Argennon
 applications, optimized
@@
 the result is guaranteed to be identical with the sequential execution of requests.} and each request has a separate
-\emph{Execution Session}. Therefore, an Execution Session is a separate session of executing application's code in
-order to fulfill an External HTTP request. We call this external request the \emph{initiator} of the Execution Session.
+\emph{execution session}. Therefore, an execution session is a separate session of executing application's code in
+order to fulfill an external HTTP request. We call this external request the \emph{initiator} of the execution session.
 
-The state of an Execution Session will be
+The state of an execution session will be
 destroyed at the end of the session and only the state of heap chunks is preserved. If a session fails and does not
@@
 
-During an Execution Session an application can make \emph{internal} HTTP requests to other applications. Those
-requests will not start a new Execution Session and will be executed within the current session. In AscEE making an
-HTTP request to an application is similar to a function invocation, and for that reason, we also refer to them as
-application calls.
+During an execution session an application can make \emph{internal} HTTP requests to other applications. Those
+requests will not start a new execution session and will be executed within the current session. In AscEE making an
+internal HTTP request to an application is similar to a function invocation, and for that reason, we also refer to
+them as \emph{application calls}.
 
@@
 AscEE does not specify any particular limits for them. When an application tries to use too much memory, that may
-cause the execution session to end abruptly. In that case, the Execution Session will not have any effects on the
+cause the execution session to end abruptly. In that case, the execution session will not have any effects on the
 state of the heap.
@@
 When an application allocates a new heap chunk, the identifier of the new chunk is not generated by
-the AscEE. Instead, the application can choose an identifier itself, provided it has a correct format. This
+the AscEE. Instead, the application can choose an identifier itself, provided the new id has a correct format. This
 is an important feature of the AscEE heap which allows applications to use the AscEE heap as a map
@@
 
-During an Execution Session every heap chunk has an access type which may disallow certain operations. This access
-type is declared by the initiator request of the Execution Session:
+During an execution session every heap chunk has an access type which may disallow certain operations. This access
+type is declared by the initiator request of the execution session:
 
@@
     \item \texttt{additive}: only allows additive operations. By additive we mean an addition-like operator without
-    overflow checking which is associative and commutative. Note that the content of these access blocks cannot be read.
+    overflow checking which is associative and commutative. Note that the content of these chunks cannot be read.
 \end{itemize}
@@
 At the start of executing requests of a block, a validator can consider two values for every
-heap chunk, the size: \texttt{chunkSize} and a size upper bound: \texttt{sizeUpperBound}. The value of
+heap chunk, its size: \texttt{chunkSize} and a size upper bound: \texttt{sizeUpperBound}. The value of
 \texttt{chunkSize} can be determined uniquely at the start of
-every Execution Session, and it may be updated during the session by the owner application. On the other hand,
-the value of \texttt{sizeUpperBound} proposed by the block proposer for each block and is calculated based on values
-declared by requests (i.e.\ transactions) declaring chunk resizing. This value needs to be an upper bound of all the
-declared values and indicates the upper bound of \texttt{chunkSize} during the execution of requests of a block.
+every execution session, and it may be updated during the session by the owner application. On the other hand,
+the value of \texttt{sizeUpperBound} is proposed by the block proposer for each block and is calculated based on values
+declared by external requests (i.e.\ transactions) which want to perform chunk resizing. This value needs to be
+an upper bound of all the declared resizing values and indicates the upper bound of \texttt{chunkSize} during the
+execution of a block.
 
-The value of \texttt{chunkSize}, can be modified during an Execution Session. However, the new size values can only be
-increasing or decreasing. More precisely, if a request declares that it wants to expand (shrink) a chunk, it can only
-increase (decrease) the value of \texttt{chunkSize} and any specified value during the execution
+The value of \texttt{chunkSize}, can be modified during an execution session. However, the new values of size can
+only be increasing or decreasing. More precisely, if a request declares that it wants to expand (shrink) a chunk, it
+can only increase (decrease) the value of \texttt{chunkSize} and any specified value during the execution
 session, needs to be greater (smaller) than the previous value of the chunk's size. Any request that wants to expand
@@
 
-The value of \texttt{chunkSize} at the end of an Execution Session will determine if a memory location at an
+The value of \texttt{chunkSize} at the end of an execution session will determine if a memory location at an
 offset is persistent or not: Offsets lower than the chunk size are persistent, and higher offsets are not.
-Non-persistent locations will be re-initialized with zero at the start of every Execution Session.
+Non-persistent locations will be re-initialized with zero at the start of every execution session.
 
 Usually an application should not have any assumption about the content of memory locations that are outside the chunk.
-While these locations are zero initialized at the start of every Execution Session, it should be noted that multiple
-invocations of an application may occur in a single Execution Session, and if one of them modifies a location outside
+While these locations are zero initialized at the start of every execution session, multiple
+invocations of an application may occur in a single execution session, and if one of them modifies a location outside
 the chunk, the changes can be seen by next invocations.
 
-There is no way for an application to query \texttt{sizeUpperBound}. As a result, for an application,
+There is no way for an application to query \texttt{sizeUpperBound}. As a result, for an application
 accessing offsets higher than \texttt{chunkSize} results in undefined behaviour, while the behaviour is well-defined
@@
 
-The AscEE uses \emph{Authenticated Message Passing} for authorizing operations. In this method, every Execution
-Session has a set of authenticated messages, and those messages are \textbf{explicitly} passed in requests to
+The AscEE uses \emph{Authenticated Message Passing} for authorizing operations. In this method, every execution
+session has a set of authenticated messages, and those messages are \textbf{explicitly} passed in requests to
 applications for authorizing operations. These messages act exactly like digital signatures and applications can
@@
 
-Each Execution Session has a list of authenticated messages. Each authenticated message has an index which will be
+Each execution session has a list of authenticated messages. Each authenticated message has an index which will be
 used for passing the message to an application as a request parameter. The AscEE uses cryptographic signatures to
 authenticate messages for user accounts. The signatures are validated in parallel during the
-preprocessing step of the block validation, and any type of cryptographic signature scheme can be used.
+preprocessing step, and any type of cryptographic signature scheme can be used.
 

### 2022-09-03 tex/sections/elements.tex
Fix typos and readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-Completing an Execution Session requires computational resources. The amount of resources used by an Execution Session
+Completing an execution session requires computational resources. The amount of resources used by an execution session
 should be monitored and managed, otherwise a malicious user would be able to easily spam and exhaust resources of the
@@
 determined per application call by the external HTTP request (i.e.\ transaction). The block proposer decides the
-type of the Execution Session for each transaction.
+type of the execution session for each transaction.
 
@@
 When a transaction fails due to the violation of a limit for a nondeterministic resource, the proposer is
-required to exactly specify the application call which violates that limit in the Execution Session. When validators are
+required to exactly specify the application call which violates that limit in the execution session. When validators are
 executing a monitored session, for each application call, they enforce considerably larger limits for

### 2022-09-03 tex/sections/intro.tex
Fix typos and readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-In addition to the ASAR, on the blockchain an Argennon application is stored in the Argennon Prover Machine (APM) code.
-This code is used by validators for CI proof verification. The Argennon Prover Machine is a virtual
+In addition to the ASAR, an Argennon application is stored on the blockchain in the Argennon Prover Machine (APM) code.
+This code is used by validators for CI proof verifications. The Argennon Prover Machine is a virtual
 machine tailored for efficient verification of AscEE computations by

### 2022-09-04 tex/sections/persistence.tex
Update persistence chapter
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2022-03-31 tex/sections/persistence.tex
@@
 
-The Argennon Smart Contract Execution Environment has two persistent memory areas: \emph{code area}, and \emph{heap}.
-Code area stores the Argennon Standard Representation of
+The Argennon Smart Contract Execution Environment has two persistent memory areas: \emph{program area}, and \emph{heap}.
+Program area stores the ASAR and the APM code of
 applications\footnote{also it stores applications' constants.}, and heap stores heap chunks.
-Both of these data elements, ASRs and chunks, can be considered as continuous arrays of bytes.
-Throughout this chapter, we shall call these data elements \emph{objects}.
+Both of these data elements can be considered as continuous arrays of bytes.
+Throughout this chapter, we shall call these data elements \emph{chunks}.
 
@@
 In the AscEE persistence layer, similar objects are clustered together and constitute a bigger data element which we
-call
-a \emph{page}.\footnote{we avoid calling them clusters, because usually a cluster refers to a \emph{set}. AscEE object
-clusters are not sets. They are ordered lists, like a page containing an ordered list of words or sentences.}
-A page is an ordered list of an arbitrary number of objects, which their order reflects the order they were added to
-the page:
-\[
-    P = (O_1,O_2,\dots,O_n),\quad i < j \; \Leftrightarrow \; \textrm{$O_i$ was added before $O_j$}\ \cdot
-\]
+call a \emph{page}.\footnote{we avoid calling them clusters, because usually a cluster refers to a \emph{set}. AscEE
+chunk clusters are not sets. They are ordered lists, like a page containing an ordered list of words or sentences.}
+A page is an ordered list of an arbitrary number of chunks. Every page has a \emph{native} chunk that has the same
+identifier as the page. In addition to the native chunk, a page can host any number of \emph{migrant} chunks.
+A page of the AscEE storage should consists of chunks that have very similar access patterns. Ideally, when a page is
+needed for validating a block, almost all of its chunks should be needed for either reading or writing. We prefer
+that the chunks are needed for the same access type. In other words, the chunks of a page should be chosen in a way that
+for validating a block, we need to either read all of them or modify\footnote{and probably read.} all of them.
 
-A page of the AscEE storage should contain objects that have very similar access pattern. Ideally, when a page is needed
-for validating a block, almost all of its objects should be needed for either reading or writing. We prefer
-that the objects are needed for the same access type. In other words, the objects of a page are chosen in a way that
-for validating a block, we usually need to either read all of them or modify\footnote{and probably read.} all of them.
+When a page contains migrants, its native chunk can not be migrated. If the page does not
+have any migrants, its native chunk can be migrated and after that the page will be converted into a special
+\texttt{<moved>} page. When a non-native chunk is migrated to another page, it will be simply removed from the page.
 
@@
 
-Pages of the AscEE storage are persisted using \emph{dynamic cryptographic accumulators}. Argennon
-has three dynamic cryptographic accumulators: \emph{staking} database, which stores all the data that is associated with
-the Argennon consensus protocol. \emph{code} database, which stores the AscEE code area, and \emph{heap} database,
-which stores the AscEE heap. The commitment of these three accumulators are included in every block of the Argennon
-blockchain. These accumulators, in the Argennon network, are hosted on special database nodes called Publicly
-Verifiable Data Base (PV-DB) servers.
+Pages of the AscEE storage are persisted using \emph{dynamic universal accumulators}. Argennon
+has two dynamic accumulators: \emph{program} database, which stores the AscEE program area, and \emph{heap} database,
+which stores the AscEE heap. The commitment of these accumulators are included in every block of the Argennon
+blockchain. In the Argennon cloud, nodes that store these accumulators are called Publicly Verifiable Database
+(PV-DB) servers.
+
 We consider the following properties for a PV-DB:
@@
 
-Pages of the AscEE storage are stored in the PV-DBs, with an index: \texttt{pageIndex} as their key.
-The \texttt{pageIndex} is required to be smaller than a certain value, determined by the
-protocol, to facilitate the usage of PV-DBs that are based on vector commitments.
-For this reason, the AscEE clustering algorithm always tries to reuse indices and keep the number of used indices
-as low as possible.
-
-The commitments of the AscEE cryptographic accumulators are affected by the way data objects are clustered. Therefore,
-the Argennon clustering algorithm has to be a part of the consensus protocol.
+The commitments of the AscEE cryptographic accumulators are affected by the way data chunks are clustered in pages.
+Therefore, the Argennon clustering algorithm has to be a part of the consensus protocol.
 
@@
 can only modify pages that were used for validating the block, and can
-include directives for moving an object from one page to another or directives specifying which pages will contain
-the newly created objects. These directives are always executed by nodes at the end of block validation.
+include directives for moving a chunk from one page to another or directives specifying which pages will contain
+newly created chunks. These directives are applied at the end of block validation, after executing requests.
 
-A block proposer could obtain clustering directives from any third party source\footnote{we can say the AscEE clustering
-algorithm is essentially off-chain.}. This will not
+A block proposer is allowed to obtain clustering directives from any third party source\footnote{we can say the AscEE
+clustering algorithm is essentially off-chain.}. This will not
 affect Argennon security, since the integrity of a database can not be altered by clustering directives.
@@
 
-\subsection{Vector Commitments}\label{subsec:impl-zk-edbs}
-
-Informally, vector commitments allow committing to an ordered sequence of $q$ values
-(i.e.\ a vector), rather than to single messages. This is done in a way such that it is later possible
-to open the commitment with respect to specific positions (e.g., to prove that $m_i$ is the $i$-th committed
-message). More precisely, vector commitments are required to satisfy what is called position binding.
-Position binding states that an adversary should not be able to open a commitment to two different
-values at the same position. While this property, by itself, would be trivial to realize using standard
-commitment schemes, what makes vector commitments interesting is that they are concise, i.e.,
-the size of the commitment string as well as the size of each opening \textbf{is independent of the
-vector length}.
-
-\note{not yet written...}
-
 \section{Object Clustering Algorithm}\label{sec:clustering}

### 2022-09-06 tex/sections/apm.tex
Update blocks description
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The Argennon Prover Machine gets as its primary input a vector $(\mathbf{C}_H,\mathbf{C}_P,\mathbf{C}_R)$, where
+The Argennon Prover Machine gets as its primary input a vector $(\mathbf{C}_H,\mathbf{C}_P,\mathbf{C}_R)$ where
 $\mathbf{C}_H$ is a commitment to the AscEE Heap, $\mathbf{C}_P$ is a commitment to the AscEE program area and
@@
 
-The configuration of the APM can be considered as a quadruple
-$(\mathbb{S},\mathbb{L},\mathbb{T}_{NEU},\mathbb{T}_{FRU})$,
-where $\mathbb{S}$ is the size of the internal stack, $\mathbb{L}$ is the amount of local memory and
-$\mathbb{T}_{NEU}, \mathbb{T}_{FRU}$ are the number of cycles that the Normal Execution Unit (NEU)
-and Failure Repeater Unit (FRU) runs for. It should be noted that the APM has one local memory which
-is used for all application calls.
+The configuration of the APM can be considered as a tuple
+\[
+(\mathcal{S},\mathcal{L}_{\text{NEU}},\mathcal{L}_{\text{FRU}},\mathcal{T}_{\text{NEU}},
+\mathcal{T}_{\text{FRU}})\ ,
+\]
+where $\mathcal{S}$ is the size of the internal stack, $\mathcal{L}_{\text{NEU}},\mathcal{L}_{\text{FRU}}$ are the
+amount of local memory of the normal execution and failure repeater units respectively, and
+$\mathcal{T}_{\text{NEU}},\mathcal{T}_{\text{FRU}}$ are the number of cycles that the NEU and FRU runs for. It should
+be noted that the APM does not use a different local memory for each application call.
 

### 2022-09-06 tex/sections/blocks.tex
Fix some whitepaper typos
Update blocks description
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The Argennon blockchain is a sequence of blocks. Every block represents an ordered list of transactions, intended to be
-executed by the Argennon Virtual Machine. The first block of the blockchain, the \emph{genesis} block, is a spacial
-block that fully describes the initial state of the AVM. Every block of the Argennon blockchain thus corresponds to a
-unique AVM state which can be calculated deterministically from the genesis block.
+The Argennon blockchain is a sequence of blocks. Every block represents an ordered list of external requests, intended
+to be executed by the Argennon Smart Contract Execution Environment (AscEE). The first block of the blockchain, the
+\emph{genesis} block, is a spacial block that fully describes the initial state of the AscEE. Every block of the
+Argennon blockchain thus corresponds to a unique AscEE state which can be calculated deterministically from the genesis
+block.
 
@@
         \hline
-        \textbf{Block} \\ [0.5ex]
+        \textbf{Block} \\ [0.6ex]
         \hline\hline
-        commitment to the staking database            \\ [1.2ex]
-        commitment to the method database             \\ [1.2ex]
-        commitment to the heap database               \\ [1.2ex]
-        commitment to the set of transactions         \\ [1.2ex]
-        a consecutive list of block certificates issued     \\
-        by validators' committee (if any)           \\ [1.2ex]
-        clustering directives                         \\ [1.2ex]
-        random seed                                   \\ [1.2ex]
+        height: $h$                          \\ [1.2ex]
+        commitment to the program database: $\mathbf{C}_{P}$             \\ [1.2ex]
+        commitment to the heap database: $\mathbf{C}_{H}$               \\ [1.2ex]
+        commitment to the ordered list of requests: $\mathbf{C}_{R}$        \\ [1.2ex]
+        clustering directives: $dir$                        \\ [1.2ex]
+        certificate of the validators assembly for \\
+        the block with height $h - k$: $v \mhyphen cert_{h-k}$         \\ [1.2ex]
+        certificate of the delegates for\\
+        the previous block: $d \mhyphen cert_{h-1}$        \\ [1.2ex]
         previous block hash                           \\ [1.2ex]
@@
 
-Having the previous AVM state, the transaction list and the clustering directives of a block, a node can calculate
-commitments to the staking, method and heap databases of the current block by emulating the AVM execution. If the
-node can obtain the previous block commitments from a trusted source, it does not need to have a trusted local
-copy of the AVM state,
-and it can reliably retrieve the required storage pages from a ZK-EDB server. We call this type of block verification
-\emph{conditional} block validation, since the validity of the current block is
+To validate a block three main conditions need to be verified: (i) commitments to the program and heap
+database result from applying the request list and clustering directives of the block to the previous AscEE
+state, (ii) $v \mhyphen cert_{h-k}$ is valid, (iii) previous block is valid and has height $h - 1$.
+
+We can denote condition (i) by a Computational Integrity (CI) statement:
+\begin{equation}
+    \label{eq:ci}
+    \mathbf{C}_{P_h},\mathbf{C}_{H_h} \coloneqq \tau(\mathbf{C}_{P_{h-1}},\mathbf{C}_{H_{h-1}},\mathbf{C}_{R_{h}},
+    dir_h)\ ,
+\end{equation}
+where $\tau$ is a transition function that encodes the AscEE computation logic and the necessary preprocessing and
+postprocessing steps\footnote{including opening and updating the state commitments}.
+
+Verifying statement~\ref{eq:ci} can be done by either replaying the AscEE computation and performing the required
+preprocessing and postprocessing steps, or alternatively by
+verifying a computation proof. To use computation proofs the verifier and prover need to share an arithmetized
+representation of $\tau$ and use it for both proof generation and verification.
+
+The Argennon Prover Machine provides a convenient and universal arithmetization for any Argennon
+application. Moreover, since a compiled version of every Argennon application to the APM code is stored in the AscEE
+program area, validators do not need to locally store the APM code of applications.
+
+The Argennon protocol does not enforce the usage of the APM. Validators and PVC servers can use different methods
+for arithmetization. This naturally would require that the ASAR of an application is used
+for generating the appropriate arithmetization.
+
+The certificate of validators, $v \mhyphen cert_{h-k}$, is an aggregate signature; validating it requires accessing the
+AscEE heap state at block $h-k-1$ to obtain public keys and stake values. Again, in addition to direct verification,
+a validator can use computation proofs, received from the Argennon cloud, to perform cheaper verification.
+
+For validating the previous block, instead of directly validating the content of the block, a validator only
+verifies the block certificates. Verifying $d \mhyphen cert_{h-1}$ is straightforward and verifying $v \mhyphen
+cert_{h-1}$ can be cheaply done by using computation proofs, obtained from the Argennon cloud. This type of block
+validation only validates the transition from block $h-1$ to block $h$, and the block is valid only if its previous
+block is valid too. We
+call this type of block verification \emph{conditional} block validation, since the validity of the current block is
 conditioned on the validity of the previous block.
 
-Interestingly, conditional block validation of multiple blocks can be done in parallel. If a node has enough bandwidth
-and computational resources, it can conditionally verify any number of blocks from a previously created blockchain
-simultaneously and in parallel. As we will see in Section~\ref{subsec:validators-committee}, this property plays an
-important role in the Argennon consensus protocol.
+In summary, a validator validates a block by verifying three computation proofs: $\pi_{\tau}$ which proves the
+transition is correct, $\pi_{h-k}$ which proves the validity of the included certificate of block $h-k$ and $\pi_{h-q}$
+which proves the previous block has a correct certificate of validators. It should be noted that these proofs are not
+a part of the block contents.
 
-To some extent, conditional validation of a single block could be parallelized as well. Many transactions
-in a block are actually independent and the order of their execution does not
-matter. These transactions can be safely validated in parallel. Section~\ref{sec:concurrency} further
-develops this concept.
+Interestingly, conditional block validation of multiple blocks can be done in parallel. Moreover, the required proofs
+can be generated independently and by different PVC servers. As we will see in
+Section~\ref{subsec:validators-committee}, this property plays an important role in the Argennon consensus protocol.
 
@@
 An Argennon block certificate is an aggregate signature of some predefined subset of accounts. This predefined subset
-is called the certificate committee and their signature ensures that the certified block is conditionally
+is called the certificate assembly and their signature ensures that the certified block is conditionally
 valid given the validity of some previous block.
 
-Argennon uses BLS aggregate signatures to represent block certificates. To better understand block certificates and
-the Argennon consensus protocol, we need to briefly review the BLS signature scheme and its aggregation mechanism.
+Argennon uses a parallel algorithm to produce block certificates and therefore the signature scheme needs to satisfy
+certain properties:
+\begin{itemize}
+    \item \textbf{Associative aggregation:} the signature aggregation operator is associative.
+    \item \textbf{Efficient cancellation:} if $S$ is the predefined set of users that we need their aggregate
+    signature, verifying an aggregate signature of $S-U$ can be done in time $O(T+|U|)$, where the aggregate
+    signature of $S$ can be verified in
+    $O(T)$.
+\end{itemize}
+
+An example for a signature scheme that supports all these properties is the BLS signature scheme.
+
+\subsubsection{BLS Signatures}
 

### 2022-09-07 tex/sections/apm.tex
Update resource management
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
     \item \textbf{Normal Execution Unit (NEU):} This module executes the requests whose execution completes
-    normally. It outputs the updated heap chunks and an $accept$ flag. More formally it evaluates $(H',accept)=P(H,
-    R_{good})$, where $H'$ is the updated heap chunks and the $accept$ flag will be set to false, If an
+    normally. It outputs the updated heap chunks and an $accept_N$ flag. More formally it evaluates $(H',accept_N)=P(H,
+    R_{good})$, where $H'$ is the updated heap chunks and the $accept_N$ flag will be set to false, If an
     application terminates abruptly.
-    \item \textbf{Failure Repeater Unit (FRU):} This module evaluates $accept=P(H',R_{bad})$, where $R_{bad}$ is
-    those requests whose execution terminates abnormally, $H'$ is the output of the NEU and the $accept$ flag is set
+    \item \textbf{Failure Repeater Unit (FRU):} This module evaluates $accept_F=P(H',R_{bad})$, where $R_{bad}$ is
+    those requests whose execution terminates abnormally, $H'$ is the output of the NEU and the $accept_F$ flag is set
     to false if a request execution completes normally.
+
+    The FRU has a higher circuit complexity than the NEU\@. Unlike the NEU, the FRU is able to restore the initial state
+    of the heap when an application fails. It also calculates the execution cost of every instruction and if the
+    application's execution cost exceeds its predefined cap, the FRU will terminate the application.
     \item \textbf{Postprocessor:} This module is responsible for calculating the updated commitments

### 2022-09-07 tex/sections/blocks.tex
Update resource management
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 In summary, a validator validates a block by verifying three computation proofs: $\pi_{\tau}$ which proves the
-transition is correct, $\pi_{h-k}$ which proves the validity of the included certificate of block $h-k$ and $\pi_{h-q}$
+transition is correct, $\pi_{h-k}$ which proves the validity of the included certificate of block $h-k$ and $\pi_{h-1}$
 which proves the previous block has a correct certificate of validators. It should be noted that these proofs are not

### 2022-09-07 tex/sections/consensus.tex
Update resource management
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 from different sets of users, known as committees. There are two primary type of certificate committees in
-Argennon: the committee of \emph{delegates} and the committee
-of \emph{validators}. Argennon has \emph{one} committee of delegates and $m$ committees of validators.
+Argennon: the committee of \emph{delegates} and the assembly of \emph{validators}. Argennon has \emph{one} committee
+of delegates and $m$ assemblies of validators.
 
 The committee of delegates issues a certificate for every block of the Argennon blockchain, and each
-committee of validators issues a certificate every $m$ blocks. A validators' committee will
-certify a block only if it has already been certified by the committee of delegates. Every committee of validators has
+assembly of validators issues a certificate every $m$ blocks. A validators assembly will
+certify a block only if it has already been certified by the committee of delegates. Every assembly of validators has
 an index between $0$ and $m - 1$, and it issues a certificate for block number $n$, if $n$ modulo $m$ equals
-the committee index.
+the assembly index.
 
 Every block of the Argennon blockchain needs a certificate from both the committee of delegates and
-the committee of validators. A block is considered final after its \textbf{next} block receives \textbf{both} of
+the assembly of validators. A block is considered final after its \textbf{next} block receives \textbf{both} of
 its certificates. In Argennon as long as more than half of the total stake of validators is controlled by honest users,
@@
 
-A block certificate issued by some members of a validators' committee is considered valid, if according to
+A block certificate issued by some members of a validators assembly is considered valid, if according to
 the staking database of the previous block \textbf{certified by the same committee}, we have:\footnote{If we calculate
@@
 
-As we mentioned before, the block with height $n$ needs a certificate from the committee of
+As we mentioned before, the block with height $n$ needs a certificate from the assembly of
 validators with index $n$ modulo $m$. To decide about signing the certificate of a block which already has
@@
 
-block certificates issued by committees of validators are included in the blocks of the Argennon blockchain.
+block certificates issued by validators assemblies are included in the blocks of the Argennon blockchain.
 A block can contain multiple certificates, provided that

### 2022-09-07 tex/sections/elements.tex
Update resource management
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2022-04-16 tex/sections/elements.tex
@@
 
-An Argennon application or smart contract is an HTTP server which is represented by an Argennon Standard
-Representation (ASR) and whose state is stored in the Argennon blockchain. Each Argennon application is identified by
+An Argennon application or smart contract is an HTTP server which is represented by an Argennon Standard Application
+Representation (ASAR) and whose state is stored in the Argennon blockchain. Each Argennon application is identified by
 a unique application identifier.
@@
 
-\section{Transactions}\label{sec:transactions}
+\section{External Requests}\label{sec:transactions}
 
-An Argennon transaction consist of an HTTP request made by a user to an Argennon application, a resource
-declaration object and a list of signed messages. Transactions can only be
-issued by users and applications can not create transactions. An Argennon transaction is also called
-an \emph{external request}.
+An Argennon \emph{external request} (i.e.\ transaction) consist of an HTTP request made by a user to an Argennon application, a
+resource declaration object and a list of signed messages. External requests can only be
+issued by users and requests created by applications are called \emph{internal request}.
 
@@
 \end{lstlisting}
+
+\section{Resource Management}\label{sec:res-man}
+
+Completing an execution session requires computational resources. The amount of resources used by an execution session
+should be monitored and managed, otherwise a malicious user would be able to easily spam and exhaust resources of the
+execution environment.
+
+In most consensus protocols, we can assume that the block proposer has enough incentive to filter out transactions
+that spam run-time resources. Here by a run-time resource, we mean a resource that at run-time, a limited amount
+of it is available, but its surplus can not be stored for later use. Execution time and local memory are
+examples of such a resource but permanent storage is not.
+
+If a proposed block contains many transactions which need a lot of run-time resources, validators would not
+be able to validate all transactions in a timely manner. Consequently, they may decide to reject the block or if they
+spend enough resources, the confirmation of that block could take more than usual. Longer block time is not
+favoured by block proposers, because it means less throughput of the system which usually means less overall rewards
+for them. In the Argennon protocol the management of run-time resources, is left to the block proposer.
+
+In Argennon the reference for resource usage is the Argennon Prover Machine. The APM has two different execution
+units: the NEU which is used for executing external requests that the block proposer claims will complete
+successfully and the FRU which executes requests that the proposer has classified as failed requests. The FRU has a
+more restricted resource management and less resources are available for requests that use the FRU.
+
+We recall from Section~\ref{ch:APM} that the APM configuration is a tuple:
+\[
+(\mathcal{S},\mathcal{L}_{\text{NEU}},\mathcal{L}_{\text{FRU}},\mathcal{T}_{\text{NEU}},\mathcal{T}_{\text{FRU}}).
+\]
+When requests are executed by the APM, the following run-time resources should be considered:
+
+\begin{itemize}
+    \item \textbf{execution cost}:
+    each APM instruction has a protocol defined cost and the execution cost of a program can be calculated
+    deterministically. Only the FRU performs these calculations and verifies that an external request does not exceed
+    its predeclared execution cost.
+
+    For executing all request of a block, at max, The NEU will run for
+    $\mathcal{T}_{\text{NEU}}$ steps and the FRU will run for
+    $\mathcal{T}_{\text{FRU}}$ steps.
+
+    \item \textbf{local memory}: $\mathcal{L}_{\text{NEU}}, \mathcal{L}_{\text{FRU}}$. The Argennon protocol requires
+    $\mathcal{L}_{\text{NEU}} = 3\mathcal{L}_{\text{FRU}}$, so the NEU has three times more local memory.
+
+    \item \textbf{stack size}: both the FPU and the NEU has the same stack size: $\mathcal{S}$
+
+    \item \textbf{heap access list}:
+    every session can only access heap locations that are declared in its access list. In addition,
+    resizing heap chunks can only be done in the range of the pre-declared lower bound and upper bound. Both the NEU
+    and the FRU enforce this restriction.
+    \item \textbf{app access list}:
+    a session may only make requests to applications that are declared in its application access list. Both the NEU
+    and the FRU enforce this restriction.
+    \item \textbf{call depth}:
+    during a session the number of nested application calls can not be more than a threshold. This threshold is
+    determined by the Argennon protocol. It should be noted that a differed call is considered like a normal call and
+    increases the call depth by one level. Only the FRU enforces this restriction.
+\end{itemize}
+
+In Argennon the proposer does not use the APM for executing requests and directly executes the ASAR of an
+application, which is much more efficient. Only when he wants to reject a transaction because of excessive resource
+usage, he will emulate the APM to make sure that the request will be rejected by the FRU. The differences between the
+NEU and FRU ensures that this policy is safe.

### 2022-09-08 tex/sections/blocks.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 An Argennon block certificate is an aggregate signature of some predefined subset of accounts. This predefined subset
-is called the certificate assembly and their signature ensures that the certified block is conditionally
+is called the certificate assembly or committee and their signature ensures that the certified block is conditionally
 valid given the validity of some previous block.
+
+Because it is not usually possible to collect the signatures of all members of a certificate committee, an Argennon
+block certificate essentially is an Accountable-Subgroup Multi-signature (ASM).
 
@@
     \item \textbf{Efficient cancellation:} if $S$ is the predefined set of users that we need their aggregate
-    signature, verifying an aggregate signature of $S-U$ can be done in time $O(T+|U|)$, where the aggregate
-    signature of $S$ can be verified in
-    $O(T)$.
+    signature, verifying an aggregate signature of $S-U$ can be done in time $O(T+|U|)$, if the aggregate
+    signature of $S$ can be verified in $O(T)$.
 \end{itemize}
@@
 
-Because it is not usually possible to collect the signatures of all members of a certificate committee, an Argennon
-block certificate essentially is an Accountable-Subgroup Multi-signature (ASM). Argennon uses a simple ASM scheme
-based on BLS aggregate signatures.
-
+Argennon uses a simple ASM scheme based on BLS aggregate signatures.
 Argennon block certificates constitute an ordered sequence based on the order of blocks they certify. If we show

### 2022-09-08 tex/sections/consensus.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-In Argennon, signature aggregation is mostly performed by ZK-EDB servers. To distribute the aggregation workload
-between different servers, Every committee of validators is divided into pre-determined groups, and each ZK-EDB
+In Argennon, signature aggregation is mostly performed by PVC servers. To distribute the aggregation workload
+between different servers, every validators assembly is divided into pre-determined groups, and each PVC
 server is responsible for signature aggregation of one group. To make sure that there is enough redundancy, the
-total number of groups should be less than the number of ZK-EDB servers and each group should be assigned to
-multiple ZK-EDB servers.
+total number of groups should be less than the number of PVC servers and each group should be assigned to
+multiple PVC servers.
 
 Any member of a group knows all the servers that are responsible for signature aggregation of his group. When a member
-signs a block certificate, he sends his signature to \textbf{all} the severs that aggregate the signatures of his group.
+signs a block certificate, he sends his signature to all of the severs that aggregate the signatures of his group.
 These servers aggregate the signatures they receive and then send the aggregated signature to the delegates.
@@
 
-The role of the delegates in the signature aggregation is limited. The important part of the work is done by ZK-EDB
-servers. As long as there are enough honest ZK-EDB servers, the network will be able to perform signature aggregation
+The role of the delegates in the signature aggregation is limited. The important part of the work is done by PVC
+servers. As long as there are enough honest PVC servers, the network will be able to perform signature aggregation
 even if the delegates are malicious.

### 2022-09-09 tex/sections/apm.tex
Fixe some typos and readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The Argennon Prover Machine gets as its primary input a vector $(\mathbf{C}_H,\mathbf{C}_P,\mathbf{C}_R)$ where
-$\mathbf{C}_H$ is a commitment to the AscEE Heap, $\mathbf{C}_P$ is a commitment to the AscEE program area and
+$\mathbf{C}_H$ is a commitment to the AscEE heap, $\mathbf{C}_P$ is a commitment to the AscEE program area and
 $\mathbf{C}_R$ is a commitment to an ordered list of requests. The APM then outputs the updated
@@
     correct format and is valid. For example it verifies the signatures of authenticated messages or
-    verifies that the proposed chunk bounds are valid.
+    checks that the proposed chunk bounds are valid.
     \item \textbf{Normal Execution Unit (NEU):} This module executes the requests whose execution completes
-    normally. It outputs the updated heap chunks and an $accept_N$ flag. More formally it evaluates $(H',accept_N)=P(H,
-    R_{good})$, where $H'$ is the updated heap chunks and the $accept_N$ flag will be set to false, If an
+    normally. It outputs the updated heap chunks and an $accept_N$ flag. More formally it evaluates
+    $H',accept_N=P(H,R_{good})$, where $H'$ is the updated heap chunks and $accept_N$ flag will be set to false, If an
     application terminates abruptly.
     \item \textbf{Failure Repeater Unit (FRU):} This module evaluates $accept_F=P(H',R_{bad})$, where $R_{bad}$ is
-    those requests whose execution terminates abnormally, $H'$ is the output of the NEU and the $accept_F$ flag is set
+    those requests whose execution terminates abnormally, $H'$ is the output of the NEU and $accept_F$ flag is set
     to false if a request execution completes normally.

### 2022-09-09 tex/sections/avm.tex
Fixe some typos and readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The Argennon Smart Contract Execution Environment (AscEE) is an abstract high level execution environment for executing
-Argennon smarts contracts (a.k.a\ Argennon applications) in a secure and isolated environment. An Argennon application
-essentially is an HTTP server whose state is
+Argennon smarts contracts (a.k.a\ Argennon applications) in an efficient and isolated environment. An Argennon
+application essentially is an HTTP server whose state is
 kept in the Argennon blockchain and its logic is described using an Argennon Standard Application Representation (ASAR).
@@
 applications, optimized
-for the architecture and properties of the Argennon platform. This text based representation should be low level
-enough to enable easy compilation from any high level programming language and should be high level enough to preserve
+for the architecture and properties of the Argennon platform. This text based representation is low level
+enough to enable easy compilation from any high level programming language and is high level enough to preserve
 the high level information of the application logic and facilitate platform specific compiler optimization at a host
@@
 the result is guaranteed to be identical with the sequential execution of requests.} and each request has a separate
-\emph{execution session}. Therefore, an execution session is a separate session of executing application's code in
+\emph{execution session}. In other words, an execution session is a separate session of executing application's code in
 order to fulfill an external HTTP request. We call this external request the \emph{initiator} of the execution session.
@@
 
-During an execution session an application can make \emph{internal} HTTP requests to other applications. Those
+During an execution session an application can make \emph{internal} HTTP requests to other applications. These
 requests will not start a new execution session and will be executed within the current session. In AscEE making an
@@
 cause the execution session to end abruptly. In that case, the execution session will not have any effects on the
-state of the heap.
+state of heap.
 
@@
 
-During an execution session every heap chunk has an access type which may disallow certain operations. This access
-type is declared by the initiator request of the execution session:
+During an execution session every heap chunk has an access-type which may disallow certain operations. The
+access-type of a chunk is declared by the initiator request of the execution session:
 
@@
 At the start of executing requests of a block, a validator can consider two values for every
-heap chunk, its size: \texttt{chunkSize} and a size upper bound: \texttt{sizeUpperBound}. The value of
+heap chunk, the size: \texttt{chunkSize} and a size upper bound: \texttt{sizeUpperBound}. The value of
 \texttt{chunkSize} can be determined uniquely at the start of
@@
 the value of \texttt{sizeUpperBound} is proposed by the block proposer for each block and is calculated based on values
-declared by external requests (i.e.\ transactions) which want to perform chunk resizing. This value needs to be
+declared by external requests (i.e.\ transactions) that want to perform chunk resizing. This value needs to be
 an upper bound of all the declared resizing values and indicates the upper bound of \texttt{chunkSize} during the
 execution of a block.
+
+The address space of a chunk starts from zero and only offsets lower than \texttt{sizeUpperBound} are valid. Trying to
+access any offset higher than \texttt{sizeUpperBound} will always cause the execution session to end abruptly. There
+is no way for an application to query \texttt{sizeUpperBound}. As a result, in the view of an application,
+accessing offsets higher than \texttt{chunkSize} will result in undefined behaviour, while the behaviour is
+well-defined for validators.
+This enables validators to determine the validity of an offset at the start of the block validation in a parallelized
+preprocessing phase without actually executing requests.
 
@@
 can only increase (decrease) the value of \texttt{chunkSize} and any specified value during the execution
-session, needs to be greater (smaller) than the previous value of the chunk's size. Any request that wants to expand
-(shrink) a chunk needs to specify a max size (min size). The value of \texttt{chunkSize} can not be set higher
+session, needs to be greater (smaller) than the previous value of the \texttt{chunkSize}. Any request that wants to
+expand (shrink) a chunk needs to specify a max size (min size). The value of \texttt{chunkSize} can not be set higher
 (lower) than this value.
@@
 
-Usually an application should not have any assumption about the content of memory locations that are outside the chunk.
+Usually an application should not have any assumptions about the content of memory locations that are outside the chunk.
 While these locations are zero initialized at the start of every execution session, multiple
@@
 the chunk, the changes can be seen by next invocations.
-
-There is no way for an application to query \texttt{sizeUpperBound}. As a result, for an application
-accessing offsets higher than \texttt{chunkSize} results in undefined behaviour, while the behaviour is well-defined
-in the view of validators.
-This enables validators to determine the validity of an offset at the start of the block validation in a parallelized
-preprocessing phase without actually executing requests.
 
@@
 
-The address space of a chunk starts from zero and only offsets lower than \texttt{sizeUpperBound} are valid. Trying to
-access any offset higher than this value will always result in a revert for the application.
-
 \section{Identifiers}\label{sec:identifiers}
@@
 
-Argennon has four primitive prefix trees:
-\emph{applications, accounts, local} and \emph{varUint}.
-All these trees are in base 256, with the maximum height
-of 8.
+Argennon has four primitive prefix trees: \emph{applications, accounts, local} and \emph{varUint}. All these trees
+are in base 256, with the maximum height of 8.
 
@@
 
-All Argennon prefix trees have an equal branching factor \(\beta\)\footnote{A typical choice for \(\beta\) is \(2^8\)
-    .}, and
-we can represent an Argennon
+All Argennon prefix trees have an equal branching
+factor \(\beta\)\footnote{A typical choice for \(\beta\) is \(2^8\).}, and we can represent an Argennon
 prefix tree as a sequence of fractional numbers in base \(\beta\):
@@
 
-The attachment of a request is a list of request identifiers of the current block that are ``attached'' to the request.
-That means, for validating the request a validator first needs to ``inject'' the digest of attached requests into the
-HTTP request text. By doing so, the called application will have access to the digest of attachments in a secure way.
+The attachment of a request is a list of request identifiers of the current block that are \emph{attached} to the
+request. That means, for validating the request a validator first needs to \emph{inject} the digest of attachments
+into the current HTTP request text. By doing so, the called application will have access to the digest of
+attachments in a secure and authenticated way.
 
@@
 used for passing the message to an application as a request parameter. The AscEE uses cryptographic signatures to
-authenticate messages for user accounts. The signatures are validated in parallel during the
-preprocessing step, and any type of cryptographic signature scheme can be used.
+authenticate messages for user accounts. The signatures are validated during the
+preprocessing step in parallel, and any type of cryptographic signature scheme can be used.
 
@@
 \section{The ArgC Language}\label{sec:the-argc-language}
-
+\note{This section is outdated...}
 \subsection{The ArgC Standard Library}\label{sec:asl}
@@
 All Argennon standards, for instance ARC standard series, which defines standards regarding transferable assets,
-are defined based on how a contract should use the AVM standard library. As a result, Argennon standards are
+are defined based on how a contract should use the ArgC standard library. As a result, Argennon standards are
 different from conventional blockchain standards. Argennon standards define some type of standard logic and

### 2022-09-09 tex/sections/blocks.tex
Fixe some typos and readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-To validate a block three main conditions need to be verified: (i) commitments to the program and heap
+To validate a block three main conditions need to be validated: (i) commitments to the program and heap
 database result from applying the request list and clustering directives of the block to the previous AscEE
@@
 
-The certificate of validators, $v \mhyphen cert_{h-k}$, is an aggregate signature; validating it requires accessing the
-AscEE heap state at block $h-k-1$ to obtain public keys and stake values. Again, in addition to direct verification,
-a validator can use computation proofs, received from the Argennon cloud, to perform cheaper verification.
+The certificate of validators, $v \mhyphen cert_{h-k}$, is an aggregate signature. Validating it requires accessing the
+AscEE heap state at block $h-k-m$ to obtain public keys and stake values. Here $m$ is the total number of validator
+assemblies and $k$ is the maximum allowed length of
+the unvalidated chain. (See Section\ref{subsec:validators-committee})
+Again, in addition to direct verification, a validator can use computation proofs, received from the Argennon cloud,
+to perform cheaper verification of this certificate.
 
 For validating the previous block, instead of directly validating the content of the block, a validator only
-verifies the block certificates. Verifying $d \mhyphen cert_{h-1}$ is straightforward and verifying $v \mhyphen
-cert_{h-1}$ can be cheaply done by using computation proofs, obtained from the Argennon cloud. This type of block
-validation only validates the transition from block $h-1$ to block $h$, and the block is valid only if its previous
-block is valid too. We
-call this type of block verification \emph{conditional} block validation, since the validity of the current block is
+verifies the block certificates. Verifying $d \mhyphen cert_{h-1}$ is straightforward. Verifying $v \mhyphen
+cert_{h-1}$, which should be performed based on stakes at block $h-m-1$, can be cheaply done by using computation
+proofs, obtained from
+the
+Argennon cloud.
+
+This type of block validation only validates the transition from block $h-1$ to block $h$, and the block is valid
+only if its previous block is valid. We
+call this type of block validation \emph{conditional block validation}, since the validity of the current block is
 conditioned on the validity of the previous block.
+
+Interestingly, conditional block validation of multiple blocks can be done in parallel. Moreover, the required proofs
+can be generated independently and by different PVC servers. As we will see in
+Section~\ref{subsec:validators-committee}, this property plays an important role in the Argennon consensus protocol.
 
@@
 a part of the block contents.
-
-Interestingly, conditional block validation of multiple blocks can be done in parallel. Moreover, the required proofs
-can be generated independently and by different PVC servers. As we will see in
-Section~\ref{subsec:validators-committee}, this property plays an important role in the Argennon consensus protocol.
 

### 2022-09-09 tex/sections/consensus.tex
Fixe some typos and readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The \texttt{minValidatorsStake}
-threshold is determined by the ADAGs, but it can never be higher than $500$ ARGs.
+threshold is determined by the ADAGs, but it can never be higher than $1000$ ARGs.
 
-Every \texttt{committeeLifeTime} number of blocks, randomly $m$ committees are selected from
-validators, in a way that the total stakes of committees are almost equal, and every
-account is a member of \textbf{at least} one committee.
+Every \texttt{AssemblyLifeTime} number of blocks, randomly $m$ assemblies are selected from
+validators, in a way that the total stakes of assemblies are almost equal, and every
+account is a member of \textbf{at least} one assembly.
 
@@
 \note{When a validator changes his status, the change has no effect until the block containing the status
-change transaction gets certified by his committee.}
+change transaction gets certified by his assembly.}
 
 A block certificate issued by some members of a validators assembly is considered valid, if according to
-the staking database of the previous block \textbf{certified by the same committee}, we have:\footnote{If we calculate
-the stake values based on the previous block a malicious committee can select the validators of the next block.}
+the staking database of the previous block \textbf{certified by the same assembly}, we have:\footnote{If we calculate
+the stake values based on the previous block a malicious assembly can select the validators of the next block.}
 \begin{itemize}
-    \item The total stake of \texttt{online} members of the committee is higher than \texttt{minOnlineStake} fraction
-    of the total stake of the committee. This threshold can be changed by the ADAGs, but it can never be lower
+    \item The total stake of \texttt{online} members of the assembly is higher than \texttt{minOnlineStake} fraction
+    of the total stake of the assembly. This threshold can be changed by the ADAGs, but it can never be lower
     than $2/3$.
@@
     \item The sum of stake values of the certificate signers is higher than $3/4$ of the total stake
-    of the committee members that have \texttt{online} status.
+    of the assembly members that have \texttt{online} status.
 \end{itemize}
@@
 The delegates can generate blocks very fast. Therefore, the Argennon blockchain always has an
-unvalidated part, which contains blocks that have a certificate from the committee of delegates but have not
+unvalidated part, which contains blocks that have a certificate from the assembly of delegates but have not
 yet received a certificate from the validators.
@@
 Some validators may also require seeing a certificate from
-some community driven committee. An honest validator never signs a certificate for two different blocks with the
+some community driven committees. An honest validator never signs a certificate for two different blocks with the
 same height.
 
-So in Argennon, the block validation by committees is performed in parallel, and validators
-do not wait for seeing the certificate of the previous block validators to start transaction validation. On the
+So in Argennon, the block validation by assemblies is performed in parallel, and validators
+do not wait for seeing the certificate of the previous block assembly to start block validation. On the
 other hand, the block certificates are published and broadcast sequentially. A validator does not publish its
-certificate if the certificate of the previous block has not been published yet. This ensures that an invalid
-fork made by malicious delegates will not receive any certificates from validators.
+certificate if the certificate of the previous block validator assembly has not been published yet. This ensures that
+an invalid fork made by malicious delegates will not receive any certificates from any validator assemblies.
 

### 2022-09-09 tex/sections/elements.tex
Remove the update notice
We remove the update notice, but the chapter is not finished yet.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Fixe some typos and readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-An application identifier, \texttt{applicationID}, is
-a unique prefix code generated by the \emph{applications} prefix tree. (See Section~\ref{sec:identifiers}.)
-An application identifier can be considered as the address of an application and has the following standard symbolic
-representation:
+An application identifier, \texttt{applicationID}, is a unique prefix code generated by the \emph{applications}
+prefix tree. (See Section~\ref{sec:identifiers}) An application identifier can be considered as the address of an
+application and has the following standard symbolic representation:
 \begin{verbatim}
@@
 \end{verbatim}
-where \texttt{<dec-num>} is a normal decimal number between $0$ and $255$. For example \texttt{21.255.37},
-\texttt{0}, \texttt{11.6} and \texttt{2.0.0.0.0}, are valid application addresses.
+where \texttt{<dec-num>} is a normal decimal number between $0$ and $255$.
+
+For example \texttt{21.255.37}, \texttt{0}, \texttt{11.6} and \texttt{2.0.0.0.0}, are valid symbolic
+representations of application addresses.
 
@@
 
-Argennon application use HTTP as the application protocol and they are advised to have a RESTful API design.
+Argennon applications use HTTP as the application protocol, and they are advised to have a RESTful API design.
 
@@
 with \texttt{applicationID = 1}, controls the ARG token, the main
-currency of the Argennon blockchain. This smart contract also manages a database of public keys and
-handles signature verification.
+currency of the Argennon blockchain. This smart contract also manages a database of public keys and stake values.
 
@@
 One public key will be used for issuing digital signatures, and the other one will be used for voting. The
-provided public keys need to meet certain cryptographic requirements,\footnote{Argennon uses Prove
-Knowledge of the Secret Key (KOSK) scheme.} and can not be already registered in the system.
+provided public keys need to meet certain cryptographic requirements.\footnote{Argennon uses Prove
+Knowledge of the Secret Key (KOSK) scheme.}
 
@@
 
-\subsection{Resource Declaration}\label{subsec:resource-declaration}
+\subsection{Resource Declaration Object}\label{subsec:resource-declaration}
 
@@
 \begin{itemize}
-    \item Maximum AscEE clocks
+    \item Maximum execution cost
     \item The list of applications the request will call
-    \item The list of access blocks the request needs
+    \item The list of chunks the request needs
     \item \texttt{maxSize} for chunks it wants to expand
@@
 caps:
-    maxClocks: 150 # maximum number of AVM execution clocks
+    maxCost: 150      # the cost of execution by the APM
     apps: [1,124.16]
@@
 favoured by block proposers, because it means less throughput of the system which usually means less overall rewards
-for them. In the Argennon protocol the management of run-time resources, is left to the block proposer.
+for them.
 
+In the Argennon protocol the management of run-time resources, is left to the block proposer.
 In Argennon the reference for resource usage is the Argennon Prover Machine. The APM has two different execution
@@
 successfully and the FRU which executes requests that the proposer has classified as failed requests. The FRU has a
-more restricted resource management and less resources are available for requests that use the FRU.
+more restricted resource management and less resources are available for requests that use the FRU\@.
 
-We recall from Section~\ref{ch:APM} that the APM configuration is a tuple:
+We recall from Chapter~\ref{ch:APM} that the APM configuration is a tuple:
 \[
@@
 
-    For executing all request of a block, at max, The NEU will run for
+    For executing the whole requests of a block, at max, The NEU will run for
     $\mathcal{T}_{\text{NEU}}$ steps and the FRU will run for
@@
 
-    \item \textbf{local memory}: $\mathcal{L}_{\text{NEU}}, \mathcal{L}_{\text{FRU}}$. The Argennon protocol requires
-    $\mathcal{L}_{\text{NEU}} = 3\mathcal{L}_{\text{FRU}}$, so the NEU has three times more local memory.
+    \item \textbf{local memory}: denoted by $\mathcal{L}_{\text{NEU}}, \mathcal{L}_{\text{FRU}}$. The Argennon protocol
+    requires: $\mathcal{L}_{\text{NEU}} = 3\mathcal{L}_{\text{FRU}}$, so the Normal Execution Unit has three
+    times more local memory.
 
@@
 In Argennon the proposer does not use the APM for executing requests and directly executes the ASAR of an
-application, which is much more efficient. Only when he wants to reject a transaction because of excessive resource
-usage, he will emulate the APM to make sure that the request will be rejected by the FRU. The differences between the
-NEU and FRU ensures that this policy is safe.
+application. That is much more efficient. Only when the proposer wants to reject a transaction due to excessive
+resource usage, he will emulate the APM to make sure the request will be rejected by the FRU. The differences
+between the NEU and FRU ensures that this policy is safe.

### 2022-09-09 tex/sections/persistence.tex
Fixe some typos and readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 In the AscEE persistence layer, similar objects are clustered together and constitute a bigger data element which we
-call a \emph{page}.\footnote{we avoid calling them clusters, because usually a cluster refers to a \emph{set}. AscEE
+call a \emph{page}.\footnote{We avoid calling them clusters, because usually a cluster refers to a \emph{set}. AscEE
 chunk clusters are not sets. They are ordered lists, like a page containing an ordered list of words or sentences.}
@@
 identifier as the page. In addition to the native chunk, a page can host any number of \emph{migrant} chunks.
-A page of the AscEE storage should consists of chunks that have very similar access patterns. Ideally, when a page is
+A page of the AscEE storage should consist of chunks that have very similar access patterns. Ideally, when a page is
 needed for validating a block, almost all of its chunks should be needed for either reading or writing. We prefer
@@
 blockchain. In the Argennon cloud, nodes that store these accumulators are called Publicly Verifiable Database
-(PV-DB) servers.
+(PV-DB) servers.\footnote{Usually PVC servers are also PV-DB servers.}
 
@@
 
-The commitments of the AscEE cryptographic accumulators are affected by the way data chunks are clustered in pages.
+The commitments of the AscEE cryptographic accumulators are affected by the way chunks are clustered in pages.
 Therefore, the Argennon clustering algorithm has to be a part of the consensus protocol.
@@
 
-A block proposer is allowed to obtain clustering directives from any third party source\footnote{we can say the AscEE
-clustering algorithm is essentially off-chain.}. This will not
+A block proposer is allowed to obtain clustering directives from any third party source.\footnote{We can say the AscEE
+clustering algorithm is essentially off-chain.} This will not
 affect Argennon security, since the integrity of a database can not be altered by clustering directives.

### 2022-09-17 tex/sections/consensus.tex
Update the normal consensus
We removed status blocks and an analysis is added for the effect of online/offline status in the protocol.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The credibility of a block of the Argennon blockchain is determined by the certificates it has received
-from different sets of users, known as committees. There are two primary type of certificate committees in
+The credibility of a block of the Argennon blockchain is determined by the certificates it receives
+from different sets of users, known as committees. There are two primary types of certificate committee in
 Argennon: the committee of \emph{delegates} and the assembly of \emph{validators}. Argennon has \emph{one} committee
@@
 The committee of delegates issues a certificate for every block of the Argennon blockchain, and each
-assembly of validators issues a certificate every $m$ blocks. A validators assembly will
+assembly of validators issues a certificate every $m$ blocks. A validator assembly will
 certify a block only if it has already been certified by the committee of delegates. Every assembly of validators has
-an index between $0$ and $m - 1$, and it issues a certificate for block number $n$, if $n$ modulo $m$ equals
+an index between $0$ and $m - 1$, and it issues a certificate for block with height $h$, if $h$ modulo $m$ equals
 the assembly index.
@@
 In addition to primary committees, Argennon has several community driven committees. Certificates of these
-committees are not required for block finality, but they could be used by members of the
-validators' committee to better decide about the validity of a block.
+committees are not required for block finality, but they could be used by members of validator assemblies to better
+decide about the validity of a block.
 
@@
 Argennon Decentralized Autonomous Governance system (ADAGs\footnote{pronounced \textipa{/eI-dagz/}.}).
-At the start of the Argennon mainnet, this committee will have
-five members, and later its size could be changed by the ADAGs in a procedure described
-in Section~\ref{sec:adags}.
+At the start of the Argennon main-net, this committee will be elected for one-year terms and will have five members.
+Later, this could be changed by the ADAGs in a procedure described in Section~\ref{sec:adags}.
 
@@
 
-Besides the main delegates' committee, a reserve committee of delegates consisting of three members is elected by users
+Besides the main committee, a reserve committee of delegates consisting of three members is elected by users
 either through the ADAGs or by \emph{emergency agreement} during the recovery protocol. In case the main committee
 fails to generate new blocks or behaves maliciously, the task of
-block generation will be assigned to the reserve committee until a new main delegates' committee is elected
-through the ADAGs.
+block generation will be assigned to the reserve committee until the main committee comes back online or a new
+committee is elected through the ADAGs.
 
-Usually, the delegates are large organizations, and they have enough computational resources to generate blocks
-very fast. However, a block is not completely final if it does not have the certificate of its validators.
-A certified block by the delegates will not be accepted by the network, if the last block certified by
-the validators is behind it more than a certain number of blocks.
+A block certified only by the delegates committee is credible, but is not considered completely final until it
+receives the certificate of its validator assembly. Since a block at height $h$ contains the validators certificate
+of the block at height $h-k$, the unvalidated part of the blockchain can not be longer than $k$ blocks.
 
 The committee of delegates may use any type of agreement protocol to reach consensus on the
-next block. Usually a very simple and fast protocol can do the job: one of the members
-is randomly chosen as the proposer, and other members vote "yes" or "no" on the proposed block.
+next block. Usually, the delegates are large organizations, and they can communicate with each
+other efficiently using their reliable networking infrastructure. This mostly eliminates the complications of their
+consensus protocol and any protocol will have a good performance in practice. Usually a very simple and fast protocol
+can do the job: one of the members is randomly chosen as the proposer, and other members vote "yes" or "no" on the
+proposed block. For better performance, they should run their agreement protocol for reaching consensus about small
+batches of transactions in their mem-pools, instead of the whole next block.
 
-If one of the delegates lose its network connectivity, no new blocks can be generated. For this reason,
+If one of the delegates loses its network connectivity, no new blocks can be generated. For this reason,
 the delegates should invest on different types of communication infrastructure, to make sure they never lose
@@
 
-\subsection{Validators}\label{subsec:validators-committee}
+\subsection{The Assemblies of Validators}\label{subsec:validators-committee}
 
@@
 system, and is measured in ARGs. Any account whose stake value is higher than
-\texttt{minValidatorsStake} threshold is considered a \emph{validator}.
-The \texttt{minValidatorsStake}
+\texttt{minValidatorStake} threshold is considered a \emph{validator}.
+The \texttt{minValidatorStake}
 threshold is determined by the ADAGs, but it can never be higher than $1000$ ARGs.
@@
 validators, in a way that the total stakes of assemblies are almost equal, and every
-account is a member of \textbf{at least} one assembly.
+account is a member of \textbf{at least} one assembly.\footnote{An account can be the member of multiple assemblies.}
+
+The value of $m$ is determined by the ADAGs, but it can never be higher than $32$. This way, it is guaranteed
+that on average, any block of the Argennon blockchain is validated by at least $3\%$ of the total ARG supply.
+
+\subsubsection{Activity Status}
 
 Every validator has a status which can be either \texttt{online} or \texttt{offline}.
-This status is stored in the ARG smart contract and is a part of the staking database. A validator can change
-his status through a method invocation
-from the ARG smart contract. When an account sets its status to \texttt{offline}, it receives a small reward, and
-it can not change it back to \texttt{online} for \texttt{statusCoolDown} number of blocks.
+This status is stored in the ARG smart contract and is part of the staking database. A validator can change
+his status to \texttt{offline} through an external request (transaction) to the ARG application. In this request he
+exactly specifies for how long he wants to be offline and after this period his status will automatically be considered
+\texttt{online} again. When a validator sets his status to \texttt{offline} for some period of time, he
+will receive a small portion of the maximum possible reward that a validator can receive in that period of time by
+actively participating in the consensus protocol. This ensures that a validator has an incentive for changing his
+status to offline rather than simply becoming inactive.
 
-\note{When a validator changes his status, the change has no effect until the block containing the status
-change transaction gets certified by his assembly.}
+The staking database of a validator assembly can only be updated by the assembly itself. That means, an external
+request which changes the status of a validator can only be included in a block that is being validated by the
+assembly of that validator.
+
+There is no transaction type for changing the status of a validator to \texttt{online}. A malicious delegates committee
+would be able to censor this type of transactions and prevent honest validators from coming back online. For this reason
+the status of a validator is considered online automatically when the specified period of time for being offline ends.
 
@@
 the staking database of the previous block \textbf{certified by the same assembly}, we have:\footnote{If we calculate
-the stake values based on the previous block a malicious assembly can select the validators of the next block.}
+the stake values based on the previous block, a malicious assembly can select the validators of the next block.}
 \begin{itemize}
@@
 
+If according to the staking database of block $h$, the total online stake of the assembly with index $h$ modulo $m$ is
+lower than \texttt{minOnlineStake} threshold, the block $h + m$ can never be certified by validators. To prevent the
+blockchain from halting in such situations, the validator assembly with index $h$ modulo $m$ will get merged into the
+assembly that has the most online stake at block $h$. This will decrease the number of assemblies to $m-1$, and the
+indices of assemblies will be updated appropriately.
+
+\subsubsection{Certificate Signing}
+
 The delegates can generate blocks very fast. Therefore, the Argennon blockchain always has an
-unvalidated part, which contains blocks that have a certificate from the assembly of delegates but have not
+unvalidated part which contains the blocks that have a certificate from the committee of delegates but have not
 yet received a certificate from the validators.
 
-As we mentioned before, the block with height $n$ needs a certificate from the assembly of
-validators with index $n$ modulo $m$. To decide about signing the certificate of a block which already has
+As we mentioned before, the block with height $h$ needs a certificate from the assembly of
+validators with index $h$ modulo $m$. To decide about signing the certificate of a block which already has
 a certificate from the delegates, a validator checks the conditional
-validity\footnote{See Section~\ref{subsec:block-validation}} of the block, and
-if the block is valid he issues
+validity of the block (See Section~\ref{subsec:block-validation}), and if the block is valid, he issues
 an "accept" signature. If the block is invalid, he initiates the recovery protocol. The validator will broadcast the
-certificate \textbf{only after} he sees the certificate of the validators of the previous block.
+certificate \textbf{only after} he sees the certificate of the validator assembly of the previous block.
 Some validators may also require seeing a certificate from
-some community driven committees. An honest validator never signs a certificate for two different blocks with the
+some community based committees. An honest validator never signs a certificate for two different blocks with the
 same height.
 
-So in Argennon, the block validation by assemblies is performed in parallel, and validators
-do not wait for seeing the certificate of the previous block assembly to start block validation. On the
-other hand, the block certificates are published and broadcast sequentially. A validator does not publish its
-certificate if the certificate of the previous block validator assembly has not been published yet. This ensures that
+Consequently, in Argennon the block validation by assemblies is performed in parallel, and validators
+do not wait for seeing the validators certificate of the previous block to start block validation. On the
+other hand, the block certificates are published and broadcast sequentially. A validator does not publish his
+signature, if the certificate of the previous block validator assembly has not been published yet. This ensures that
 an invalid fork made by malicious delegates will not receive any certificates from any validator assemblies.
 
-The value of $m$ is determined by the ADAGs. but it can never be higher than $25$. This way, it is guaranteed
-that on average, any block of the Argennon blockchain is validated by at least $0.02$ of the total ARG supply.
+\subsubsection{Analysis}
 
-block certificates issued by validators assemblies are included in the blocks of the Argennon blockchain.
-A block can contain multiple certificates, provided that
-those certificates belong to consecutive blocks.
+We analyze the minimum amount of stake that is required for conducting different types of attacks against the Argennon
+blockchain. In these attack scenarios, we assume that a single validator assembly is corrupted, all the delegates
+are malicious and the adversary is able to partition the network arbitrarily.
 
-\subsection{Status Blocks}\label{subsec:status-blocks}
+We denote the total stake of the corrupted validator assembly with $s$ and the total stake of malicious users of the
+assembly with $m$. We use $d$ to denote the stake of the assembly's users with \texttt{offline} status and $h$ to denote
+the total stake of the assembly's users with \texttt{online} status that do not participate in the
+protocol.\footnote{$d$ stands for \emph{deactivated} and $h$ stands for \emph{hidden}.} We assume
+that a certificate is accepted if it is signed by more than $r$ fraction of the total online stake of the assembly.
+We obtain the minimum required malicious stake for three types of attacks:
+\begin{itemize}
+    \item Confirming an invalid block:
+    \[ m > r(s-d) \]
+    \item Forking the blockchain by double voting and network partitioning:
+    \[m > (2r-1)(s-d)+h \]
+    \item Halting the blockchain by refusing to vote:
+    \[m > (1-r)(s-d)-h \]
+\end{itemize}
 
-If according to the staking database of block $n$, the total online stake of the committee with index $n$ modulo $m$ is
-lower than \texttt{minOnlineStake} threshold, the block $n + m$ can never be certified by validators.
+In Argennon we have $r=\frac{3}{4}$ and $d < \frac{1}{3}s$.
+Consequently, in Argennon to confirm an invalid block, the adversary needs at least $\frac{1}{2}$ of the total stake
+of the assembly. For forking the blockchain, interestingly $m$ is
+minimized when $h=0$, and $\frac{1}{3}s$ is the minimum required stake. For halting the blockchain,
+an adversary requires a stake higher than $\frac{1}{6}s-h$.
 
-To prevent the blockchain from halting in such situations, the protocol performs a predefined partial
-reshuffling of committee members.
-In this reshuffling which is based on the block random seed, some online members from other committees will
-be moved to the committee without enough online stake to make it active again.
+In particular we are interested in comparing the Argennon protocol with a simple protocol that accepts a certificate
+if it is signed by more than $\frac{2}{3}$ of the total stake of the assembly and there is no \texttt{online/offline}
+status for users.
 
-If the reshuffling can not solve the problem due to low total online stake, the protocol requires the next block
-of the blockchain to be a special \emph{status block}. A status block is a special block which can only contain
-status change transactions. The status block needs to be certified by the delegates and by 2/3 of the total stake of
-validators. The \texttt{online}/\texttt{offline} status of validators will not be considered in the validity of
-the status block certificate.\footnote{Theoretically at the status block, the total online stake of the system could
-be very low. Therefore, the status block should not be certified only by online stake.} After applying
-the transactions of the status block, the total online stake of all committees of validators
-must go higher than \texttt{minOnlineStake} threshold.
+To compare the minimum stake required for halting the blockchain in two protocols, we can solve the following
+inequality:
+\[
+    \frac{1}{4}(s-d)-h > \frac{1}{3}s-h\ \cdot
+\]
+We observe that as long as $d>\frac{1}{9}s$, the minimum required stake for halting the blockchain is higher in the
+Argennon protocol.
 
-\subsection{Signature Aggregation}\label{subsec:sig-agg}
+\subsubsection{Signature Aggregation}\label{subsubsec:sig-agg}
+
+The validators certificate of a block is an aggregate signature of members of the corresponding assembly of
+validators. Validator assemblies could include millions of users and calculating their aggregate
+signature requires an efficient distributed algorithm.
 
 In Argennon, signature aggregation is mostly performed by PVC servers. To distribute the aggregation workload
-between different servers, every validators assembly is divided into pre-determined groups, and each PVC
+between different servers, every validator assembly is divided into pre-determined groups, and each PVC
 server is responsible for signature aggregation of one group. To make sure that there is enough redundancy, the
@@
 Any member of a group knows all the servers that are responsible for signature aggregation of his group. When a member
-signs a block certificate, he sends his signature to all of the severs that aggregate the signatures of his group.
-These servers aggregate the signatures they receive and then send the aggregated signature to the delegates.
+signs a block certificate, he sends his signature to all the severs that aggregate the signatures of his group.
+These PVC servers aggregate the signatures they receive and then send the aggregated signature to the delegates.
 Furthermore, the delegates aggregate these signatures to produce the final block certificate
-and then include it in the next block.
+and then broadcast it to the PVC servers network.
 
-The role of the delegates in the signature aggregation is limited. The important part of the work is done by PVC
-servers. As long as there are enough honest PVC servers, the network will be able to perform signature aggregation
-even if the delegates are malicious.
+The role of the delegates in the signature aggregation algorithm is limited. The important part of the work is done by
+PVC servers and slightly modified versions of this algorithm can perform signature aggregation
+even if all the delegates are malicious, as long as there are enough honest PVC servers.
 

### 2022-09-17 tex/sections/consensus.tex
Add a new version of emergency forking
The algorithm still is a draft.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The recovery protocol can recover the functionality
-of the Argennon blockchain as long as more than 2/3 of the total stake of the system is controlled by honest users
+of the Argennon blockchain as long as more than $2/3$ of the total stake of the system is controlled by honest users
 and any network partition resolves after a finite amount of time. The recovery protocol uses two main emergency
@@
 by validators.
-A valid fork request is an unexpired request signed by more than half of the
-total \texttt{online} stake of the validators.
+A valid fork request is an unexpired request signed by more than half of the total stake of validators.
 
-For forking at block $b$, the reserve committee of delegates
-makes a special \emph{fork block} which only contains a valid fork request, and its parent is the block $b$.
-The height of the fork block is
-$b + 1$ so it needs a valid certificate from the committee of validator with the index $b+1$ modulo $m$.
+For forking at block $h$, the reserve committee of delegates
+makes a special \emph{fork block} which only contains a valid fork request, and its parent is the block $h$.
+The height of the fork block therefore is
+$h + 1$, so it needs a valid certificate from the assembly of validators with the index $h+1$ modulo $m$.
 
-For signing a fork block, a validator ensures that the block is signed by the reserve committee and contains
-a valid fork request. The parent of the fork block does not necessarily need a validators' certificate. If the
-parent does not have a certificate, the validator checks the certificate of the block before the parent and
-the conditional validity of the parent instead. This enables the reserve committee
-to recover the liveliness of
-the blockchain in a situation where a malicious committee has generated multiple blocks at the same height.
+For signing a fork block, a validator ensures that the block is signed by the reserve committee, contains
+a valid fork request and its parent is not issued by the same reserve committee. The parent of the fork block does not
+necessarily need a validators certificate. If the parent does not have a certificate, the validator checks the
+certificate of the block before the parent and the conditional validity of the parent instead. This enables the
+reserve committee to recover the liveliness of the blockchain in a situation where a malicious committee has
+generated multiple blocks at the same height. Notice that the block before the parent always need a validators
+certificate.
 
 A validator always chooses a valid fork block over a block of the main chain. However, as we mentioned before,
-a validator never signs a certificate for two different blocks with the same height. Consequently, if a validator
-has already signed the block $b$ of the main chain, he will not sign the fork block and vice versa.
+a honest validator \textbf{never signs a certificate for two different blocks with the same height}. Consequently, if a
+validator has already signed the block $h+1$ of the main chain, he will not sign the fork block at height $h+1$ and
+vice versa.
 
-When the fork block is broadcast, it is possible that the validators of the committee with index $b+1$ modulo $m$
-get divided between the fork block and the block $b+1$ of the main chain, in a way that no block gets enough validators.
-This will cause the blockchain to halt. To prevent this, the protocol allows the reserve delegates to revoke a
-fork block. After a fork block
-is revoked, the validators who voted for it are allowed to vote for another block with the same
-height.
-
-To revoke a fork block with height $b+1$, the delegates need to seal the fork by adding a special
-\emph{seal block} after the fork block.\footnote{The parent of the seal block must be the fork block.} The seal block
-has the height $b+2$ and needs to be certified by the validators' committee with the
-index $b+2$ modulo $m$. The fork block is considered revoked only after the seal block is certified by the validators.
-
-The seal block is not a normal block and validators who signed a certificate for a seal block are allowed
-to sign a certificate for a block with the same height and vice versa. However, it should be noted that
-generating a block with the same parent as the seal block is considered a malicious behaviour of the reserve
-committee and validators will not sign such a block.
-
-As long as more than half of the total online stake of every committee of validators is controlled
-by honest users, a malicious committee of delegates can not use the emergency forking procedure to discard blocks
-that have a certificate from validators.
+The reserve committee of delegates is not allowed to generate multiple fork blocks at the same height. However, they
+can generate multiple fork blocks with different heights, as long as the parent block is generated by the previous
+committee. When the reserve committee generates a fork block at height $h+1$, but the block fails to receive a
+certificate from validators, the committee will generate a new fork block at height $h+2$.
 
@@
 valid blocks do not get discarded, including blocks that are not certified by the validators yet.
+
+\paragraph{Analysis}
+
+When the reserve committee gets activated, the main committee might have been malicious, so any number of blocks could
+exist at each height. However, at each height at max one block can have a validators certificate. Moreover, if at some
+height there are not any blocks with a validators certificate, then no blocks at higher heights can have a
+validators certificate, because validators do not sign the certificate of a normal block if its parent does
+not have a certificate.
+
+If $h_{\max}$ denotes the height of the highest block with a validators certificate and a fork block is generated at
+height $h$, as long as more than $2/3$ stake of all the assemblies of validators are honest, we have:
+\begin{itemize}
+    \item if $h > h_{\max} + 2$, the fork block can not receive a certificate from validators.
+    \item if $h = h_{\max} + 2$, the fork block can always receive a certificate from validators if network
+    partitions do not last forever.
+    \item if $h = h_{\max} + 1$, the fork block may receive a certificate from validators or not. It is possible that
+    the validators of the assembly with index $h+1$ modulo $m$ get divided between the fork block and a block at
+    height $h+1$ of the main chain.
+    \item if $h < h_{\max} + 1$, the fork block can not receive a certificate from validators.
+\end{itemize}
+
+Obviously, if the reserve committee generates two fork blocks at height $h_{\max} + 2$ and $h_{\max} + 1$, both blocks
+could receive a validators certificate. However, this will not be problematic, since when the fork block at height
+$h_{\max} + 2$ has a validators certificate, any block added after the fork block at height $h_{\max} + 1$ can not
+receive the certificate of validators and the fork block at this height is effectively sealed.
+
+During the emergency forking, at max one block with a validators certificate may be discarded. This happens when
+the malicious main committee of delegates has forked the main chain by producing blocks $b_1$ and $b_2$ at height
+$h$; the block $b_1$ has received a validators certificate and validators have not certified any blocks at height $h+1$.
+The reserve committee adds a fork block after block $b_2$ and that will discard $b_1$. Notice that if a block at
+height $h+1$ had been certified by validators the fork block chould not have received a certificate from validators.
 
@@
 and the \emph{confirmation phase} which confirms the selected proposal. The voting phase is done in rounds.
-Each round lasts for approximately $\lambda$ units of time, and after $k$ rounds, the current agreement session ends and a
+Each round lasts for approximately $\lambda$ units of time, and after $k$ rounds, the current agreement session ends
+and a
 new session starts. All votes and messages are tagged in a way that the messages of one session can
@@
 in round $r$ of the voting phase:
+
 \paragraph{Voting Phase:}
@@
 \end{itemize}
+
 \paragraph{Confirmation Phase for Proposal $p$:}

### 2022-09-18 tex/sections/consensus.tex
Clarify the emergency agreement protocol and fix a typo
We had forgotten to stress that users who final-voted a value should also r-vote it.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
+The emergency agreement protocol is essentially an election procedure and involves human interaction. Users need to
+determine who they want to vote for by interacting with the software. As long as users can not agree upon electing a
+candidate, the voting process has to continue.
+
 The agreement protocol consists of two phases: the \emph{voting phase}, which selects a single proposal
 and the \emph{confirmation phase} which confirms the selected proposal. The voting phase is done in rounds.
-Each round lasts for approximately $\lambda$ units of time, and after $k$ rounds, the current agreement session ends
-and a
-new session starts. All votes and messages are tagged in a way that the messages of one session can
+Each round lasts for approximately $\lambda$ units of time, and after $k$ rounds, the current \emph{agreement session}
+ends and a new session starts. All votes and messages are tagged in a way that the messages of one session can
 not be used in another session.
 
-Users cast three type of votes: \emph{$i$-votes}, which are votes that are valid only in
-round $i$, \emph{final-votes}, which are votes that are valid in any round, and \emph{c-votes}, which are votes
-used only in the confirmation phase.
+Users cast three types of vote: \emph{$r$-votes}, which are votes that are valid only in
+round $r$, \emph{final-votes}, which are votes that are valid in any round, and \emph{c-votes}, which are votes
+used only in the confirmation phase. Votes are weighted based on users' stakes and \texttt{online/offline} status of
+users is not considered. When we say 2/3 votes, we mean the sum of the stake of voters is 2/3 of the total stake.
 
-A user executes the following procedure
-in round $r$ of the voting phase:
-
-\paragraph{Voting Phase:}
+\newpage
+A user executes the following procedure in each agreement session:
+\paragraph{Voting Phase in Round $r$:}
 \begin{itemize}
-    \item if the user has not yet final-voted any value, he $r$-votes a single desired proposal.
+    \item if the user has final-voted any value, he $r$-votes his final-vote, otherwise he $r$-votes a single desired
+    proposal.
     \item if he sees more than 2/3 $r$-votes for a proposal $p$, he final-votes $p$.
@@
     \item user c-votes $p$.
-    \item if he sees more than 2/3 c-votes for $p$ he selects $p$ and ends the agreement protocol.
     \item if $clock > k \cdot \lambda$ user starts a new agreement session.
+\end{itemize}
+
+\paragraph{Termination:}
+\begin{itemize}
+    \item as soon as the user sees more than 2/3 c-votes for $p$ he selects $p$ and ends the agreement protocol.
 \end{itemize}
@@
 
+A honest user during a session only final-votes a single proposal. This ensures that as long as $2/3$ of the total
+stake of the system is controlled by honest users, two different proposal can not get $2/3$ final-votes. However, when
+a user sees $2/3$ final-votes for a proposal, he can not be sure that other users have also seen those votes and have
+ended the voting phase. If enough stake is still in the voting phase the agreement protocol may confirm a
+different proposal.
+
+The final confirmation phase ensures that when a proposal gets more than $2/3$ c-votes, more than $1/3$ of
+\textbf{honest} stake has ended the voting phase and hence no other proposal can be confirmed by the agreement protocol.
+
+In unlikely circumstances, an adversary can prevent honest users from reaching agreement in a session. As an example
+assume that $s$ is the total stake of users and a malicious user controls $m$ stake. We have two proposals $p_1$ and
+$p_2$. In the first round all users vote for $p_1$, but only $s/3 - m$ stake sees more than $2/3$ $1$-votes. At the
+next round all the remaining stake votes for $p_2$. Again, only $s/3 - m$ stake sees more than $2/3$
+$2$-votes. In round $3$, the malicious user stops voting. Since $s/3 - m$ stake final-voted for $p_1$ and another
+$s/3 - m$ stake final-voted for $p_2$, the remaining $s/3$ honest stake is not enough for reaching an agreement.
+Therefore, after $k$ rounds the session will end without an agreement and a new session will be started.
+
+In later agreement sessions, if honest users do not change their choices drastically or if messages are transmitted
+more reliably between users, an agreement can be reached eventually.
+
 When the emergency agreement protocol is used for electing a new reserve committee to fork the blockchain, the
 confirmation phase could be skipped. In that case, the confirmation of the fork block by the appropriate committee
-of validators acts like the confirmation phase.
+of validators could act like the confirmation phase.
 

### 2022-09-18 tex/sections/consensus.tex
Update subsection "Initiating the Recovery Protocol"
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 of the validators and its start time has not passed. The validator calculates
-the stake values based on the staking database of the last final block in his blockchain without
+the stake values based on the staking database of the last \textbf{final} block in his blockchain without
 considering the \texttt{online}/\texttt{offline} status of validators.
@@
 A honest user during a session only final-votes a single proposal. This ensures that as long as $2/3$ of the total
-stake of the system is controlled by honest users, two different proposal can not get $2/3$ final-votes. However, when
+stake of the system is controlled by honest users, two different proposals can not get $2/3$ final-votes. However, when
 a user sees $2/3$ final-votes for a proposal, he can not be sure that other users have also seen those votes and have
@@
 
-When a validator does not receive any blocks for \texttt{blockTimeOut} amount of time, or when he sees an
-evidence which proves the delegates are malicious, he initiates the recovery protocol.
+When the validator software does not receive any blocks for \texttt{blockTimeOut} amount of time, or when it observes an
+evidence which proves the delegates are malicious, after prompting the user and after his confirmation, it will
+initiate the recovery protocol.
 
-To do so, first the validator activates the censorship resilient mode of his networking module, then he checks the
-validity of the blocks that do not have a validators' certificate and determines the
-last valid block of his blockchain.
+To do so, first the validator software activates the censorship resilient mode of the networking module, then it checks
+the validity of the blocks that do not have a validators certificate and determines the last valid block of its
+version of the blockchain.
 
-In the next step, he will sign and broadcast an \textbf{emergency fork request} message, alongside some useful metadata
-such as the last valid block of his blockchain and the evidence of delegates' misbehaviour.\footnote{this metadata is
-not a part of the fork request.}
+In the next step, it will sign and broadcast an \textbf{emergency fork request} message, alongside some useful metadata
+such as the last valid block of its blockchain and the evidence of delegates' misbehaviour.\footnote{this metadata is
+not a part of the fork request.} Before starting the recovery protocol, validators try to synchronize their blockchains
+as much as possible.
 
-If the reserve committee of delegates is already active, or if a validator sees a valid fork request signed
+If the reserve committee of delegates is already active, or if the validator software sees a valid fork request signed
 by more than half of the total online stake of the validators, but does not receive the fork
-block after a certain amount of time, he will sign and broadcast a request for
+block after a certain amount of time, after user confirmation, it will sign and broadcast a request for
 \textbf{emergency agreement} on a new reserve committee. The agreement on new delegates usually needs user
@@
     \item two different blocks with the same parent
-    \item a block that has an invalid format
 \end{itemize}
@@
 of stake the user has in the system. Conventionally in these systems, a user's stake is considered to be equal with the
-amount of native system tokens, he has "staked" in the system. A user stakes his tokens by locking them in
+amount of native system tokens, he has ``staked'' in the system. A user stakes his tokens by locking them in
 his account or a separate staking account for some period of time. During this time, he will not be able to transfer

### 2022-09-19 tex/sections/consensus.tex
Fix an error
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The committee of delegates is responsible for creating new blocks of the Argennon blockchain, and it issues a
-certificate for every block of the Argennon blockchain. A certificate needs to be signed
-by \textbf{all} of the committee members in order to be considered valid.
+The committee of delegates is responsible for creating new blocks of the Argennon blockchain, and issues a
+certificate for every block of the Argennon blockchain. The certificate needs to be signed
+by \textbf{all} the committee members in order to be considered valid.
 
@@
 
-A block certified only by the delegates committee is credible, but is not considered completely final until it
-receives the certificate of its validator assembly. Since a block at height $h$ contains the validators certificate
-of the block at height $h-k$, the unvalidated part of the blockchain can not be longer than $k$ blocks.
+A block certified only by the committee of delegates is relatively credible, but it is not considered final
+until its next block receives the certificate of its validator assembly. Since a block at height $h$ contains the
+validators certificate of the block at height $h-k$, the unvalidated part of the Argennon blockchain can not be longer
+than $k$ blocks.
 
 The committee of delegates may use any type of agreement protocol to reach consensus on the
-next block. Usually, the delegates are large organizations, and they can communicate with each
+next block. Usually the delegates are large organizations, and they can communicate with each
 other efficiently using their reliable networking infrastructure. This mostly eliminates the complications of their
-consensus protocol and any protocol will have a good performance in practice. Usually a very simple and fast protocol
-can do the job: one of the members is randomly chosen as the proposer, and other members vote "yes" or "no" on the
-proposed block. For better performance, they should run their agreement protocol for reaching consensus about small
-batches of transactions in their mem-pools, instead of the whole next block.
+consensus protocol and any protocol could have a good performance in practice. Usually a very simple and fast protocol
+can do the job: one of the members is randomly chosen as the proposer, and other members vote ``yes'' or ``no'' on the
+proposed block. For better performance, the delegates should run their agreement protocol for reaching consensus about
+small batches of transactions in their mem-pools, instead of the whole next block.
 
 If one of the delegates loses its network connectivity, no new blocks can be generated. For this reason,
-the delegates should invest on different types of communication infrastructure, to make sure they never lose
+the delegates should invest on different types of communication infrastructure, to make sure they will never lose
 connectivity to each other and to the Argennon network.
@@
 Every \texttt{AssemblyLifeTime} number of blocks, randomly $m$ assemblies are selected from
-validators, in a way that the total stakes of assemblies are almost equal, and every
-account is a member of \textbf{at least} one assembly.\footnote{An account can be the member of multiple assemblies.}
+validators, in a way that the total stakes of different assemblies are almost equal, and every
+account is a member of \textbf{at least} one assembly.\footnote{An account can be a member of multiple assemblies.}
 
 The value of $m$ is determined by the ADAGs, but it can never be higher than $32$. This way, it is guaranteed
-that on average, any block of the Argennon blockchain is validated by at least $3\%$ of the total ARG supply.
+that on average, any block of the Argennon blockchain is validated by at least $2\%$ of the total ARG supply.
 
@@
 his status to \texttt{offline} through an external request (transaction) to the ARG application. In this request he
-exactly specifies for how long he wants to be offline and after this period his status will automatically be considered
+exactly specifies for how long he wants to be offline and after this period his status will be automatically considered
 \texttt{online} again. When a validator sets his status to \texttt{offline} for some period of time, he
@@
 
-The staking database of a validator assembly can only be updated by the assembly itself. That means, an external
-request which changes the status of a validator can only be included in a block that is being validated by the
-assembly of that validator.
+The staking database of a validator assembly can be updated only by the assembly itself. That means, an external
+request which changes the status of a validator can be included only in a block that is validated by the assembly of
+that validator.
 
@@
 the staking database of the previous block \textbf{certified by the same assembly}, we have:\footnote{If we calculate
-the stake values based on the previous block, a malicious assembly can select the validators of the next block.}
+the stake values based on the previous block, a malicious assembly can select the validators of the next block by
+manipulating the staking database.}
 \begin{itemize}
@@
 
-\subsubsection{Certificate Signing}
+The merging will continue recursively until the online stake of all remaining assemblies is higher than
+\texttt{minOnlineStake}. If eventually all assemblies get merged together and only one assembly remains, the condition
+for validity of block certificates changes: A certificate of validators will be considered valid if the sum of stakes
+of the certificate signers is higher than $2/3$ of the total stake of validators and \texttt{online/offline}
+status of validators becomes ineffective.
 
-The delegates can generate blocks very fast. Therefore, the Argennon blockchain always has an
+\subsubsection{Signing the Block Certificate}
+
+The delegates can generate blocks very fast. Consequently, the Argennon blockchain always has an
 unvalidated part which contains the blocks that have a certificate from the committee of delegates but have not
@@
 validity of the block (See Section~\ref{subsec:block-validation}), and if the block is valid, he issues
-an "accept" signature. If the block is invalid, he initiates the recovery protocol. The validator will broadcast the
+an ``accept'' signature. If the block is invalid, he initiates the recovery protocol. The validator will broadcast the
 certificate \textbf{only after} he sees the certificate of the validator assembly of the previous block.
@@
 blockchain. In these attack scenarios, we assume that a single validator assembly is corrupted, all the delegates
-are malicious and the adversary is able to partition the network arbitrarily.
+are malicious and the adversary is able to fully control message transmission and partition the network arbitrarily.
 
 We denote the total stake of the corrupted validator assembly with $s$ and the total stake of malicious users of the
-assembly with $m$. We use $d$ to denote the stake of the assembly's users with \texttt{offline} status and $h$ to denote
-the total stake of the assembly's users with \texttt{online} status that do not participate in the
+assembly with $m$. We use $d$ to denote the stake of users of the assembly who have \texttt{offline} status and $h$
+to denote the stake of users of the assembly who have \texttt{online} status and do not participate in the
 protocol.\footnote{$d$ stands for \emph{deactivated} and $h$ stands for \emph{hidden}.} We assume
@@
 Consequently, in Argennon to confirm an invalid block, the adversary needs at least $\frac{1}{2}$ of the total stake
-of the assembly. For forking the blockchain, interestingly $m$ is
+of an assembly. For forking the blockchain, interestingly $m$ is
 minimized when $h=0$, and $\frac{1}{3}s$ is the minimum required stake. For halting the blockchain,
@@
 \[
-    \frac{1}{4}(s-d)-h > \frac{1}{3}s-h\ \cdot
+    \frac{1}{4}(s-d)-h > \frac{1}{3}s-(d+h)\ \cdot
 \]
 We observe that as long as $d>\frac{1}{9}s$, the minimum required stake for halting the blockchain is higher in the
-Argennon protocol.
+Argennon protocol and the value of $h$ does not matter.
 

### 2022-09-20 tex/sections/consensus.tex
Complete the new emergency forking protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The merging will continue recursively until the online stake of all remaining assemblies is higher than
-\texttt{minOnlineStake}. If eventually all assemblies get merged together and only one assembly remains, the condition
+\texttt{minOnlineStake} fraction.
+If eventually all assemblies get merged together and only one assembly remains, the condition
 for validity of block certificates changes: A certificate of validators will be considered valid if the sum of stakes
@@
 
-In particular we are interested in comparing the Argennon protocol with a simple protocol that accepts a certificate
+In particular, we are interested in comparing the Argennon protocol with a simple protocol that accepts a certificate
 if it is signed by more than $\frac{2}{3}$ of the total stake of the assembly and there is no \texttt{online/offline}
@@
 
-The recovery protocol can recover the functionality
-of the Argennon blockchain as long as more than $2/3$ of the total stake of the system is controlled by honest users
-and any network partition resolves after a finite amount of time. The recovery protocol uses two main emergency
-procedures to recover the functionality of the Argennon blockchain: \emph{emergency forking} and \emph{emergency
+We assume that an adversary is able to fully control message transmission between users and is able to partition the
+network arbitrarily for finite periods of time. Under these circumstances, the recovery protocol can recover the
+functionality of the Argennon blockchain as long as more than $2/3$ of the total stake of every validator assembly is
+controlled by honest users. The recovery protocol uses two main emergency procedures to
+recover the functionality of the Argennon blockchain: \emph{emergency forking} and \emph{emergency
 agreement} protocol.
@@
 from the validators.
-This fork needs to be confirmed by validators and can not discard any blocks that has been already certified
-by validators.
+This fork needs to be confirmed by validators and can not discard more than one block that has been certified by
+validators.
 A valid fork request is an unexpired request signed by more than half of the total stake of validators.
@@
 makes a special \emph{fork block} which only contains a valid fork request, and its parent is the block $h$.
-The height of the fork block therefore is
-$h + 1$, so it needs a valid certificate from the assembly of validators with the index $h+1$ modulo $m$.
+The height of the fork block therefore is $h + 1$, and the fork block needs a valid certificate from the assembly of
+validators with the index $h+1$ modulo $m$. When a
+fork block gets certified by validators, its parent is also confirmed and will become a part of the blockchain, even if
+it does not have a validators certificate.
 
-For signing a fork block, a validator ensures that the block is signed by the reserve committee, contains
-a valid fork request and its parent is not issued by the same reserve committee. The parent of the fork block does not
-necessarily need a validators certificate. If the parent does not have a certificate, the validator checks the
-certificate of the block before the parent and the conditional validity of the parent instead. This enables the
+For signing a fork block at height $h+1$, a validator ensures that the following conditions hold:
+\begin{itemize}
+    \item the fork block is signed by the reserve committee.
+    \item the fork block contains a valid fork request.
+    \item the parent block of the fork block is issued by the previous committee.
+    \item the parent block of the fork block is certified by validators, or the parent block is conditionally
+    valid and there is a fork block with height $h$ which is certified by validators, or the parent block is
+    conditionally valid and the parent block of the parent has a validators certificate.
+    \item the validator has not already signed a certificate for a fork or normal block at height $h+1$.
+\end{itemize}
+
+The parent of the fork block does not necessarily need a validators certificate. This enables the
 reserve committee to recover the liveliness of the blockchain in a situation where a malicious committee has
-generated multiple blocks at the same height. Notice that the block before the parent always need a validators
+generated multiple blocks at the same height. Notice that the block before the parent always needs a validators
 certificate.
 
-A validator always chooses a valid fork block over a block of the main chain. However, as we mentioned before,
-a honest validator \textbf{never signs a certificate for two different blocks with the same height}. Consequently, if a
-validator has already signed the block $h+1$ of the main chain, he will not sign the fork block at height $h+1$ and
-vice versa.
+A validator always chooses a valid fork block over a block of the main chain and may sign different fork
+blocks with different heights. However, as we mentioned before, an honest validator
+\textbf{never signs a certificate for two different blocks with the same height}. Consequently, a validator never
+signs two fork blocks at the same height, and if he has already signed the fork block at height $h+1$, he will not
+sign the block $h+1$ of the main chain and vice versa.
 
-The reserve committee of delegates is not allowed to generate multiple fork blocks at the same height. However, they
-can generate multiple fork blocks with different heights, as long as the parent block is generated by the previous
-committee. When the reserve committee generates a fork block at height $h+1$, but the block fails to receive a
-certificate from validators, the committee will generate a new fork block at height $h+2$.
+The reserve committee of delegates is allowed to generate multiple fork blocks with different heights, as long as the
+parent block is generated by the previous committee. When the reserve committee generates multiple fork blocks
+at different heights, the next normal block must be always added after the fork block with the highest height.
 
-Moreover, an honest committee of delegates will always try to perform the emergency forking in such a way that
-valid blocks do not get discarded, including blocks that are not certified by the validators yet.
+The reserve committee of delegates should try to perform the emergency forking in such a way that
+valid blocks do not get discarded, including blocks that have not been certified by the validators yet.
+
+For forking the blockchain, the reserve committee uses a straightforward algorithm. Let $h_v$ be the height of the last
+block with a validator certificate and $h_v+k$ be the height of the last valid block that the reserve committee has
+seen. For forking the main chain, the reserve committee generates all fork blocks with heights $h_v+1,h_v+2,\dots,
+h_v+k+1$. The parent of the fork block with height $h_v+i$ will be the block $h_v+i-1$ of the main chain. Then the
+reserve committee will wait until the fork block with height $h_v+k+1$ receives a certificate form validators. The
+reserve committee then will continue the normal chain after that fork block and the fork block with height $h_v+k+1$
+will be the parent of the first normal block generated by the reserve committee.
 
@@
 height there are not any blocks with a validators certificate, then no blocks at higher heights can have a
-validators certificate, because validators do not sign the certificate of a normal block if its parent does
+validators certificate either, because validators do not sign the certificate of a normal block if its parent does
 not have a certificate.
 
-If $h_{\max}$ denotes the height of the highest block with a validators certificate and a fork block is generated at
-height $h$, as long as more than $2/3$ stake of all the assemblies of validators are honest, we have:
+If $h_{\max}$ denotes the height of the highest block with a validators certificate, as long as more than $2/3$ stake
+of all the assemblies of validators are honest, for a fork block with height $h_f$ we have:
 \begin{itemize}
-    \item if $h > h_{\max} + 2$, the fork block can not receive a certificate from validators.
-    \item if $h = h_{\max} + 2$, the fork block can always receive a certificate from validators if network
+    \item if $h_f \leq h_{\max}$, the fork block can not receive a certificate from validators.
+    \item if $h_f = h_{\max} + 1$, the fork block may receive a certificate from validators or not. It is possible that
+    the validators of the assembly with index $h_f$ modulo $m$ get divided between the fork block and a block at
+    height $h_f$ of the main chain.
+    \item if $h_f = h_{\max} + 2$, the fork block can always receive a certificate from validators, if network
     partitions do not last forever.
-    \item if $h = h_{\max} + 1$, the fork block may receive a certificate from validators or not. It is possible that
-    the validators of the assembly with index $h+1$ modulo $m$ get divided between the fork block and a block at
-    height $h+1$ of the main chain.
-    \item if $h < h_{\max} + 1$, the fork block can not receive a certificate from validators.
+    \item if $h_f \geq h_{\max} + 3$, the fork block can always receive a certificate from validators \textbf{only if} a
+    fork block at height $h_f-1$ gets certified by validators.
 \end{itemize}
 
-Obviously, if the reserve committee generates two fork blocks at height $h_{\max} + 2$ and $h_{\max} + 1$, both blocks
-could receive a validators certificate. However, this will not be problematic, since when the fork block at height
-$h_{\max} + 2$ has a validators certificate, any block added after the fork block at height $h_{\max} + 1$ can not
-receive the certificate of validators and the fork block at this height is effectively sealed.
+Consequently, if two fork blocks at heights $h_0$ and $h_0+k$ are generated by the reserve committee, if both blocks
+receive a certified from validators, we must have $h_0 > h_{\max}$, and there must exist fork blocks with heights
+$h_0+1,\dots,h_0+k-1$ which are certified by validators. That means if a malicious reserve committee generates a normal
+block after any of the fork blocks with heights less than $h_0+k$, that normal block can not receive a certificate from
+validators.
 
@@
 
-\newpage
 A user executes the following procedure in each agreement session:

### 2022-09-20 tex/sections/consensus.tex
Fix an important typo in the emergency voting
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
     proposal.
-    \item if he sees more than 2/3 $r$-votes for a proposal $p$, he final-votes $p$.
+    \item if the user has \textbf{not} final-voted any value, if he sees more than 2/3 $r$-votes for a proposal $p$, he
+    final-votes $p$.
     \item if he sees more than 2/3 final-votes for a proposal, he goes to the confirmation phase for $p$.
-    \item if $clock > r \cdot \lambda$ and $r < k$ user goes to the round $r + 1$ and if $r = k$ user starts a
+    \item if $clock > r \cdot \lambda$ and $r < k$ the user goes to the round $r + 1$ and if $r = k$ user starts a
     new agreement session.
@@
 \begin{itemize}
-    \item user c-votes $p$.
+    \item the user c-votes $p$.
     \item if $clock > k \cdot \lambda$ user starts a new agreement session.
@@
 a user sees $2/3$ final-votes for a proposal, he can not be sure that other users have also seen those votes and have
-ended the voting phase. If enough stake is still in the voting phase the agreement protocol may confirm a
+ended the voting phase. If enough stake is still in the voting phase, the agreement protocol may confirm a
 different proposal.
-
 The final confirmation phase ensures that when a proposal gets more than $2/3$ c-votes, more than $1/3$ of
@@
 assume that $s$ is the total stake of users and a malicious user controls $m$ stake. We have two proposals $p_1$ and
-$p_2$. In the first round all users vote for $p_1$, but only $s/3 - m$ stake sees more than $2/3$ $1$-votes. At the
-next round all the remaining stake votes for $p_2$. Again, only $s/3 - m$ stake sees more than $2/3$
+$p_2$. In the first round all users $1$-vote for $p_1$, but only $s/3 - m$ stake sees more than $2/3$ $1$-votes. At the
+next round all the remaining stake $2$-votes for $p_2$. Again, only $s/3 - m$ stake sees more than $2/3$
 $2$-votes. In round $3$, the malicious user stops voting. Since $s/3 - m$ stake final-voted for $p_1$ and another
 $s/3 - m$ stake final-voted for $p_2$, the remaining $s/3$ honest stake is not enough for reaching an agreement.
-Therefore, after $k$ rounds the session will end without an agreement and a new session will be started.
+Consequently, after $k$ rounds the session will end without an agreement and a new session will be started.
 

### 2022-09-20 tex/sections/consensus.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The height of the fork block therefore is $h + 1$, and the fork block needs a valid certificate from the assembly of
-validators with the index $h+1$ modulo $m$. When a
+validators with index $h+1$ modulo $m$. When a
 fork block gets certified by validators, its parent is also confirmed and will become a part of the blockchain, even if
@@
 
-For forking the blockchain, the reserve committee uses a straightforward algorithm. Let $h_v$ be the height of the last
+For forking the blockchain, the reserve committee uses a straightforward algorithm: let $h_v$ be the height of the last
 block with a validator certificate and $h_v+k$ be the height of the last valid block that the reserve committee has
 seen. For forking the main chain, the reserve committee generates all fork blocks with heights $h_v+1,h_v+2,\dots,
-h_v+k+1$. The parent of the fork block with height $h_v+i$ will be the block $h_v+i-1$ of the main chain. Then the
-reserve committee will wait until the fork block with height $h_v+k+1$ receives a certificate form validators. The
-reserve committee then will continue the normal chain after that fork block and the fork block with height $h_v+k+1$
+h_v+k+1$. The parent of the fork block with height $h_v+i$ will be the block $h_v+i-1$ of the main chain. The
+reserve committee will wait until the fork block with height $h_v+k+1$ receives a certificate from validators and
+then will continue the normal chain after that fork block. Hence, the fork block with height $h_v+k+1$
 will be the parent of the first normal block generated by the reserve committee.
@@
 If $h_{\max}$ denotes the height of the highest block with a validators certificate, as long as more than $2/3$ stake
-of all the assemblies of validators are honest, for a fork block with height $h_f$ we have:
+of every assembly of validators are honest, for a fork block with height $h_f$ we have:\footnote{This fork block
+forks the blockchain at height $h_f-1$.}
 \begin{itemize}
@@
 $h_0+1,\dots,h_0+k-1$ which are certified by validators. That means if a malicious reserve committee generates a normal
-block after any of the fork blocks with heights less than $h_0+k$, that normal block can not receive a certificate from
+block after any fork block with height less than $h_0+k$, that normal block can not receive a certificate from
 validators.
@@
 $h$; the block $b_1$ has received a validators certificate and validators have not certified any blocks at height $h+1$.
-The reserve committee adds a fork block after block $b_2$ and that will discard $b_1$. Notice that if a block at
-height $h+1$ had been certified by validators the fork block chould not have received a certificate from validators.
+The reserve committee adds a fork block whose parent is block $b_2$ and that will essentially discard $b_1$. Notice
+that if a block at height $h+1$ had been certified by validators the fork block could not have received a certificate
+from validators.
 

### 2022-09-23 tex/sections/consensus.tex
Add new emergency agreement protocol
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-The agreement protocol consists of two phases: the \emph{voting phase}, which selects a single proposal
-and the \emph{confirmation phase} which confirms the selected proposal. The voting phase is done in rounds.
-Each round lasts for approximately $\lambda$ units of time, and after $k$ rounds, the current \emph{agreement session}
-ends and a new session starts. All votes and messages are tagged in a way that the messages of one session can
-not be used in another session.
-
-Users cast three types of vote: \emph{$r$-votes}, which are votes that are valid only in
-round $r$, \emph{final-votes}, which are votes that are valid in any round, and \emph{c-votes}, which are votes
-used only in the confirmation phase. Votes are weighted based on users' stakes and \texttt{online/offline} status of
-users is not considered. When we say 2/3 votes, we mean the sum of the stake of voters is 2/3 of the total stake.
+The voting process is done in rounds and each round usually lasts for approximately $\lambda$ units of time.
+$\lambda$ is selected by th ADAGs and could be several hours. All votes and messages \textbf{are tagged} in a way
+that a vote cast in a round can not be used in another round. Votes are weighted based on users' stakes and
+\texttt{online/offline} status of users is not considered. When we say 2/3 votes, we mean the sum of the stake of
+voters is 2/3 of the total stake.
 
@@
 \begin{itemize}
-    \item if the user has final-voted any value, he $r$-votes his final-vote, otherwise he $r$-votes a single desired
+    \item if the user has locked his vote on a proposal $p$, he votes $p$, otherwise he votes a single desired
     proposal.
-    \item if the user has \textbf{not} final-voted any value, if he sees more than 2/3 $r$-votes for a proposal $p$, he
-    final-votes $p$.
-    \item if he sees more than 2/3 final-votes for a proposal, he goes to the confirmation phase for $p$.
-    \item if $clock > r \cdot \lambda$ and $r < k$ the user goes to the round $r + 1$ and if $r = k$ user starts a
-    new agreement session.
-\end{itemize}
-
-\paragraph{Confirmation Phase for Proposal $p$:}
-\begin{itemize}
-    \item the user c-votes $p$.
-    \item if $clock > k \cdot \lambda$ user starts a new agreement session.
+    \item when $clock = \lambda$, if the users has seen more than $2/3$ votes for a proposal $p$, he votes $p$-win,
+    otherwise he votes $draw$. A user votes either $p$-win or $draw$, not both.
+    \item when $clock = k \lambda$ for $k=2,3,\dots$, if the user has seen more than $2/3$ votes for a proposal
+    $p$ he votes $p$-lock.
+    \item if the users sees more than 2/3 $draw$ votes, he goes to the round $r+1$ and sets $clock=0$.
+    \item if the user sees more than 2/3 $p$-lock votes, he goes to the round $r+1$, sets $clock=0$ and locks his
+    vote on $p$.
 \end{itemize}
@@
 \begin{itemize}
-    \item as soon as the user sees more than 2/3 c-votes for $p$ he selects $p$ and ends the agreement protocol.
+    \item as soon as the user sees more than 2/3 $p$-win votes for $p$, he selects $p$ and ends
+    the agreement protocol. p-win votes can be for any round but all must belong to the same round.
 \end{itemize}
@@
 
-A honest user during a session only final-votes a single proposal. This ensures that as long as $2/3$ of the total
-stake of the system is controlled by honest users, two different proposals can not get $2/3$ final-votes. However, when
-a user sees $2/3$ final-votes for a proposal, he can not be sure that other users have also seen those votes and have
-ended the voting phase. If enough stake is still in the voting phase, the agreement protocol may confirm a
-different proposal.
-The final confirmation phase ensures that when a proposal gets more than $2/3$ c-votes, more than $1/3$ of
-\textbf{honest} stake has ended the voting phase and hence no other proposal can be confirmed by the agreement protocol.
+A honest user during a round only votes a single proposal and either votes $p$-win or $draw$. This ensures that
+as long as more than $2/3$ of the total stake of the system is controlled by honest users, no two different proposals
+can get more than $2/3$ votes. In addition, only one of $p$-win or $draw$ can get more than $2/3$ votes. When a
+user sees more than $2/3$ $p$-win votes, he can be sure that more than $1/3$ of the \textbf{honest}
+stake has voted $p$-win, so $draw$ has less than $2/3$ votes. Therefore, for going to the next round we
+will need 2/3 $p$-lock votes and all honest users will lock their vote on $p$. As a result only $p$ can be confirmed by
+the agreement protocol.
 
-In unlikely circumstances, an adversary can prevent honest users from reaching agreement in a session. As an example
-assume that $s$ is the total stake of users and a malicious user controls $m$ stake. We have two proposals $p_1$ and
-$p_2$. In the first round all users $1$-vote for $p_1$, but only $s/3 - m$ stake sees more than $2/3$ $1$-votes. At the
-next round all the remaining stake $2$-votes for $p_2$. Again, only $s/3 - m$ stake sees more than $2/3$
-$2$-votes. In round $3$, the malicious user stops voting. Since $s/3 - m$ stake final-voted for $p_1$ and another
-$s/3 - m$ stake final-voted for $p_2$, the remaining $s/3$ honest stake is not enough for reaching an agreement.
-Consequently, after $k$ rounds the session will end without an agreement and a new session will be started.
-
-In later agreement sessions, if honest users do not change their choices drastically or if messages are transmitted
-more reliably between users, an agreement can be reached eventually.
-
-When the emergency agreement protocol is used for electing a new reserve committee to fork the blockchain, the
-confirmation phase could be skipped. In that case, the confirmation of the fork block by the appropriate committee
-of validators could act like the confirmation phase.
+If honest users can agree upon some proposal value, the agreement protocol will converge to that value. When a
+single hones user terminates the protocol, he can convince all other honest users to terminate their protocol by
+sending those 2/3 $p$-win votes that he has seen. When we go to the next round all honest users will lock their vote on
+the same proposal and an agreement could be reached in next rounds. We will never get stuck in a round. If at some
+round $draw$ gets less than 2/3 votes, that means at least $\epsilon$ honest stake has voted
+$p$-win.\footnote{It's possible that the $\epsilon$ stake has not voted yet. However based on the finite time
+partitioning assumption, at some point that honest stake should get connected to the network and vote.}
+Consequently, there must be more than 2/3 votes for some proposal $p$ which convinced the $\epsilon$ honest stake to
+vote $p$-win. Therefore, after waiting long enough, all the honest stake will see those votes and will eventually
+vote for $p$-lock, so $p$-lock will get more than 2/3 votes.
 
@@
 \[
-    B = B_{k+1} = B_{k+2} = \dots = B_{t}\ .
+    B = B_{k+1} = B_{k+2} = \dots = B_{t}\ \cdot
 \]
@@
 \[
-    M_{t} = M_{k} + (t - k) \alpha (B - M_{k})\ .
+    M_{t} = M_{k} + (t - k) \alpha (B - M_{k})\ \cdot
 \]

### 2022-09-23 tex/sections/consensus.tex
Fix typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The voting process is done in rounds and each round usually lasts for approximately $\lambda$ units of time.
-$\lambda$ is selected by th ADAGs and could be several hours. All votes and messages \textbf{are tagged} in a way
+$\lambda$ is selected by the ADAGs and could be several hours. All votes and messages \textbf{are tagged} in a way
 that a vote cast in a round can not be used in another round. Votes are weighted based on users' stakes and
@@
     otherwise he votes $draw$. A user votes either $p$-win or $draw$, not both.
+    \item if the user sees more than 2/3 $draw$ votes, he goes to the round $r+1$ and sets $clock=0$.
+    \item if the user sees more than 2/3 $p$-lock votes, he goes to the round $r+1$, sets $clock=0$ and locks his
+    vote on $p$.
     \item when $clock = k \lambda$ for $k=2,3,\dots$, if the user has seen more than $2/3$ votes for a proposal
     $p$ he votes $p$-lock.
-    \item if the users sees more than 2/3 $draw$ votes, he goes to the round $r+1$ and sets $clock=0$.
-    \item if the user sees more than 2/3 $p$-lock votes, he goes to the round $r+1$, sets $clock=0$ and locks his
-    vote on $p$.
 \end{itemize}
@@
     \item as soon as the user sees more than 2/3 $p$-win votes for $p$, he selects $p$ and ends
-    the agreement protocol. p-win votes can be for any round but all must belong to the same round.
+    the agreement protocol. The $p$-win votes can be for any round, but all must belong to the same round.
 \end{itemize}
 
-We assume that users have clocks with the same speed, and $\lambda \gg \epsilon$, where $\epsilon$ is the maximum
+We assume that users have clocks with the same speed, and $\lambda \gg \delta$, where $\delta$ is the maximum
 clock difference between users. We also assume that more than 2/3 of the total stake of the system is controlled
@@
 
-A honest user during a round only votes a single proposal and either votes $p$-win or $draw$. This ensures that
-as long as more than $2/3$ of the total stake of the system is controlled by honest users, no two different proposals
-can get more than $2/3$ votes. In addition, only one of $p$-win or $draw$ can get more than $2/3$ votes. When a
-user sees more than $2/3$ $p$-win votes, he can be sure that more than $1/3$ of the \textbf{honest}
-stake has voted $p$-win, so $draw$ has less than $2/3$ votes. Therefore, for going to the next round we
-will need 2/3 $p$-lock votes and all honest users will lock their vote on $p$. As a result only $p$ can be confirmed by
-the agreement protocol.
+A honest user during a round only votes a single proposal. This ensures that as long as more than $2/3$ of the total
+stake of the system is controlled by honest users, no two different proposals can get more than $2/3$ votes. Also, he
+either votes $p$-win or $draw$, so only one of $p$-win or $draw$ can get more than $2/3$ votes. Consequently, when a
+user sees more than $2/3$ $p$-win votes, he can be sure that $draw$ has less than $2/3$ votes. Therefore, for going
+to the next round we will need 2/3 $p$-lock
+votes and all honest users will lock their vote on $p$ when they start the next round. As a result only $p$ can be
+confirmed by the agreement protocol.
 
-If honest users can agree upon some proposal value, the agreement protocol will converge to that value. When a
-single hones user terminates the protocol, he can convince all other honest users to terminate their protocol by
-sending those 2/3 $p$-win votes that he has seen. When we go to the next round all honest users will lock their vote on
-the same proposal and an agreement could be reached in next rounds. We will never get stuck in a round. If at some
+When a single honest user terminates the protocol, he can convince all other honest users to terminate their protocol
+by sending those 2/3 $p$-win votes that he has seen.
+
+If honest users can agree upon some proposal value, the agreement protocol will converge to that value. When a round
+ends and we go to the next round, all honest users will lock their vote on
+the same proposal or no one will lock his vote, so an agreement could be reached in next rounds. We will never get
+stuck in a round. If at some
 round $draw$ gets less than 2/3 votes, that means at least $\epsilon$ honest stake has voted
-$p$-win.\footnote{It's possible that the $\epsilon$ stake has not voted yet. However based on the finite time
+$p$-win.\footnote{It is possible that the $\epsilon$ stake has not voted yet. However based on the finite time
 partitioning assumption, at some point that honest stake should get connected to the network and vote.}
-Consequently, there must be more than 2/3 votes for some proposal $p$ which convinced the $\epsilon$ honest stake to
+That means there must be more than 2/3 votes for some proposal $p$ which convinced the $\epsilon$ honest stake to
 vote $p$-win. Therefore, after waiting long enough, all the honest stake will see those votes and will eventually
-vote for $p$-lock, so $p$-lock will get more than 2/3 votes.
+vote for $p$-lock, and $p$-lock will get more than 2/3 votes.
 

### 2022-09-24 tex/sections/consensus.tex
Fix typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
     proposal.
-    \item when $clock = \lambda$, if the users has seen more than $2/3$ votes for a proposal $p$, he votes $p$-win,
+    \item when $clock = \lambda$, if the user has seen more than $2/3$ votes for a proposal $p$, he votes $p$-win,
     otherwise he votes $draw$. A user votes either $p$-win or $draw$, not both.
@@
     \item if honest users can agree upon some proposal value, the agreement protocol will converge to that value
-    after a finite number of sessions.
+    after a finite number of rounds.
 \end{itemize}

### 2022-09-24 tex/sections/intro.tex
Fix typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 security of the consensus protocol used in a blockchain. Unfortunately, many currently used blockchains are
-vulnerable to a certain type of consensus attack, known as the Bribery Attack. In a Bribery Attack, an adversary
+vulnerable to a certain type of consensus attack, known as the \emph{bribery attack}. In a bribery attack, an adversary
 tries to corrupt participants of a protocol by offering them money and seducing them to violate the protocol.
@@
 
-In case the main committee fails to generate new blocks or behaves maliciously, a special Recovery Protocol is
+In case the main committee fails to generate new blocks or behaves maliciously, a special \emph{recovery protocol} is
 initiated by validators.
@@
 of the Argennon blockchain as long as more than $2/3$ of the total stake of the system is controlled by honest users
-and any network partition resolves after a finite amount of time. The Recovery Protocol uses two main emergency
+and any network partition resolves after a finite amount of time. The recovery protocol uses two main emergency
 procedures to recover the functionality of the Argennon blockchain: \emph{Emergency Forking} and \emph{Emergency

### 2022-09-26 tex/sections/apm.tex
Fix readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-Producing the required outputs from these inputs is not computationally feasible so the APM receives an auxiliary
+Producing the required outputs from these inputs is not computationally feasible, so the APM receives an auxiliary
 input vector $(H,\pi_H,P,\pi_P,R,\pi_R,h)$, where $\pi_X$ is a proof that proves $X$ can be opened w.r.t
@@
 
-The configuration of the APM can be considered as a tuple
+The configuration of the APM can be considered as a tuple:
 \[

### 2022-09-26 tex/sections/blocks.tex
Update the normal consensus
We removed status blocks and an analysis is added for the effect of online/offline status in the protocol.

Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Fix readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
         clustering directives: $dir$                        \\ [1.2ex]
-        certificate of the validators assembly for \\
+        certificate of the validator assembly for \\
         the block with height $h - k$: $v \mhyphen cert_{h-k}$         \\ [1.2ex]
@@
 To validate a block three main conditions need to be validated: (i) commitments to the program and heap
-database result from applying the request list and clustering directives of the block to the previous AscEE
+database are resulted from applying the request list and clustering directives of the block to the previous AscEE
 state, (ii) $v \mhyphen cert_{h-k}$ is valid, (iii) previous block is valid and has height $h - 1$.
@@
 preprocessing and postprocessing steps, or alternatively by
-verifying a computation proof. To use computation proofs the verifier and prover need to share an arithmetized
-representation of $\tau$ and use it for both proof generation and verification.
+verifying a computational integrity proof\footnote{More accurately we should call it an argument instead of a proof.
+However, using the word argument can confuse a reader who is not familiar with the subject, so we avoid it here.}.
+To use computational integrity proofs the verifier and prover need to
+share an arithmetized representation of $\tau$ and use it for both proof generation and verification.
 
-The Argennon Prover Machine provides a convenient and universal arithmetization for any Argennon
+The Argennon Prover Machine provides a convenient and universal way for arithmetization of any Argennon
 application. Moreover, since a compiled version of every Argennon application to the APM code is stored in the AscEE
@@
 
-The Argennon protocol does not enforce the usage of the APM. Validators and PVC servers can use different methods
-for arithmetization. This naturally would require that the ASAR of an application is used
-for generating the appropriate arithmetization.
+The Argennon protocol does not enforce the usage of the APM. Validators and PVC servers can use any argument system
+with any arithmetization, and if required, the ASAR of an application can be used
+for generating the appropriate arithmetization instead of the APM\@.
 
@@
 the unvalidated chain. (See Section\ref{subsec:validators-committee})
-Again, in addition to direct verification, a validator can use computation proofs, received from the Argennon cloud,
-to perform cheaper verification of this certificate.
+Again, in addition to direct verification, a validator can use computational integrity proofs, received from the
+Argennon cloud, to perform cheaper verification of this certificate.
 
@@
 verifies the block certificates. Verifying $d \mhyphen cert_{h-1}$ is straightforward. Verifying $v \mhyphen
-cert_{h-1}$, which should be performed based on stakes at block $h-m-1$, can be cheaply done by using computation
-proofs, obtained from
-the
-Argennon cloud.
+cert_{h-1}$, which should be performed based on stakes at block $h-m-1$, can be cheaply done by using computational
+proofs, obtained from the Argennon cloud.
 
@@
 
-In summary, a validator validates a block by verifying three computation proofs: $\pi_{\tau}$ which proves the
-transition is correct, $\pi_{h-k}$ which proves the validity of the included certificate of block $h-k$ and $\pi_{h-1}$
-which proves the previous block has a correct certificate of validators. It should be noted that these proofs are not
-a part of the block contents.
+In summary, a validator validates a block by verifying three computational integrity proofs: $\pi_{\tau}$ which proves
+the transition is correct, $\pi_{h-k}$ which proves the validity of the included certificate of block $h-k$ and
+$\pi_{h-1}$ which proves the previous block has a correct certificate of validators. It should be noted that these
+proofs are not a part of the block contents and therefore the exact argument system used for generating them is not
+specified by the Argennon protocol.
 
@@
 
-Argennon uses a parallel algorithm to produce block certificates and therefore the signature scheme needs to satisfy
-certain properties:
+The Argennon network uses a parallel algorithm to produce block certificates and therefore the signature scheme needs
+to satisfy certain properties:
 \begin{itemize}

### 2022-09-26 tex/sections/consensus.tex
Fix readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 the assembly of validators. A block is considered final after its \textbf{next} block receives \textbf{both} of
-its certificates. In Argennon as long as more than half of the total stake of validators is controlled by honest users,
+its certificates. In Argennon as long as more than $2/3$ of the total stake of validators is controlled by honest users,
 the probability of discarding a final block is near zero even if all the delegates are malicious.

### 2022-09-26 tex/sections/intro.tex
Fix readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 This recovery protocol can recover the functionality
-of the Argennon blockchain as long as more than $2/3$ of the total stake of the system is controlled by honest users
-and any network partition resolves after a finite amount of time. The recovery protocol uses two main emergency
-procedures to recover the functionality of the Argennon blockchain: \emph{Emergency Forking} and \emph{Emergency
-Agreement} protocol.
+of the Argennon blockchain as long as more than $2/3$ of the total stake of every validator assembly is controlled by
+honest users and any network partition resolves after a finite amount of time. The recovery protocol uses two main
+emergency procedures to recover the functionality of the Argennon blockchain: \emph{Emergency Forking} and
+\emph{Emergency Agreement} protocol.
 
@@
 delegates committee and its validator assembly.
-The Argennon protocol ensures that as long as more than half of the total stake of validators is
+The Argennon protocol ensures that as long as more than $2/3$ of the total stake of validators is
 controlled by honest users, the probability of discarding a final block is near zero even if all the delegates are
@@
 argument systems. The APM has a minimal RISC architecture with a very compact instruction set. This ensures that
-its transition function has an efficient circuit complexity.
+its transition function has an efficient circuit complexity. However, the Argennon protocol does not enforce the usage
+of the APM. Validators and PVC servers can use any argument system with any arithmetization, and if required, the ASAR
+of an application can be used for generating the appropriate arithmetization instead of the APM\@.
 

### 2023-01-11 tex/sections/avm.tex
Improve readability of Chapter 2
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The Argennon Smart Contract Execution Environment (AscEE) is an abstract high level execution environment for executing
-Argennon smarts contracts (a.k.a\ Argennon applications) in an efficient and isolated environment. An Argennon
+Argennon smarts contracts a.k.a\ Argennon applications in an efficient and isolated environment. An Argennon
 application essentially is an HTTP server whose state is
@@
 resize its heap chunks or allocate new chunks. Every chunk belongs to exactly one application and can only be modified
-by its owner. In addition to heap chunks, every application has a limited amount of non-persistent local memory for
+by its owner. Besides heap chunks, every application has a limited amount of non-persistent local memory for
 storing temporary data.
@@
 In this step by analyzing data dependency between requests, the AscEE determines requests that can be run in parallel
-and requests that need to be run sequentially. This information is represented using an \emph{execution DAG} data
-structure and in the final step, the requests are executed using this data structure.
+and requests that need to be run sequentially. This information is represented using an \emph{execution DAG} and in
+the final step, the requests are executed using this data structure.
 
@@
 the result is guaranteed to be identical with the sequential execution of requests.} and each request has a separate
-\emph{execution session}. In other words, an execution session is a separate session of executing application's code in
-order to fulfill an external HTTP request. We call this external request the \emph{initiator} of the execution session.
+\emph{execution session}. An execution session is a separate session of executing application's code in
+order to fulfill an external HTTP request. This external request is the \emph{initiator} of the execution session.
 
@@
 Every Argennon application has two types of memory: local memory and heap. Local memory is not persistent and is
-destroyed when the application call ends. Local memory is used for storing local variables and is not directly
-addressable. Heap, on the other hand, is persistent and can be used for persisting data between application calls.
-Heap is addressable and provides low level direct access. Both local memory and heap are limited, but the
-AscEE does not specify any particular limits for them. When an application tries to use too much memory, that may
-cause the execution session to end abruptly. In that case, the execution session will not have any effects on the
-state of heap.
+destroyed when the application call ends. Heap, on the other hand, is persistent and can be used for persisting data
+between application calls. Local memory is used for storing local variables and is not directly
+addressable. Heap is addressable and provides low level direct access. Both local memory and heap are limited, but
+the limit is not specified by the AscEE. If an application tries to use too much
+memory, that may cause the execution session to end abruptly. In that case, the execution session will not have any
+effects on the state of heap.
 
@@
 When an application allocates a new heap chunk, the identifier of the new chunk is not generated by
-the AscEE. Instead, the application can choose an identifier itself, provided the new id has a correct format. This
-is an important feature of the AscEE heap which allows applications to use the AscEE heap as a map
+the AscEE. Instead, the application can choose an identifier itself, provided the new identifier has a correct format.
+This is an important feature of the AscEE heap which enables applications to use the AscEE heap as a map
 data structure\footnote{also called a dictionary.}.
 Since the \texttt{chunkID} is a prefix code, any application has its own identifier space, and an application
-can easily find unique identifiers for its chunks.
+can easily find unique identifiers for its chunks. (See Section~\ref{sec:identifiers})
 
-During an execution session every heap chunk has an access-type which may disallow certain operations. The
-access-type of a chunk is declared by the initiator request of the execution session:
+During an execution session every heap chunk has an access-type which may disallow certain operations during that
+session. This access-type is declared by the initiator request of the execution session for every chunk:
 
@@
     \item \texttt{writable}: allows reading and writing.
-    \item \texttt{additive}: only allows additive operations. By additive we mean an addition-like operator without
-    overflow checking which is associative and commutative. Note that the content of these chunks cannot be read.
+    \item \texttt{additive}: only allows additive operations. By additive, we mean an addition-like operator without
+    overflow checking which is associative and commutative. Note that the content of the chunk cannot be read.
 \end{itemize}
@@
 every execution session, and it may be updated during the session by the owner application. On the other hand,
-the value of \texttt{sizeUpperBound} is proposed by the block proposer for each block and is calculated based on values
-declared by external requests (i.e.\ transactions) that want to perform chunk resizing. This value needs to be
-an upper bound of all the declared resizing values and indicates the upper bound of \texttt{chunkSize} during the
-execution of a block.
+the value of \texttt{sizeUpperBound} can be determined uniquely at the start of block validation and is proposed by the
+block proposer for each block. This value is calculated based on resizing values
+declared by external requests (i.e.\ transactions) that want to perform chunk resizing and needs to be
+an upper bound of all the declared resizing values, indicating an upper bound of \texttt{chunkSize} during the
+execution of the requests of a block.
 
 The address space of a chunk starts from zero and only offsets lower than \texttt{sizeUpperBound} are valid. Trying to
-access any offset higher than \texttt{sizeUpperBound} will always cause the execution session to end abruptly. There
-is no way for an application to query \texttt{sizeUpperBound}. As a result, in the view of an application,
-accessing offsets higher than \texttt{chunkSize} will result in undefined behaviour, while the behaviour is
-well-defined for validators.
-This enables validators to determine the validity of an offset at the start of the block validation in a parallelized
-preprocessing phase without actually executing requests.
+access any offset higher than \texttt{sizeUpperBound} will always cause the execution session to end abruptly. The
+value of \texttt{sizeUpperBound} is always greater than \texttt{chunkSize} and there is no way for an application to
+query \texttt{sizeUpperBound}. As a result, in the view of an application,
+accessing offsets higher than \texttt{chunkSize} results in undefined behaviour, while the behaviour is
+well-defined in the view of a validator.
+This enables validators to determine the validity of an offset at the start of the block validation in a
+preprocessing phase without actually executing requests, while \texttt{sizeUpperBound} can be determined in a simple
+parallelized algorithm.
 
@@
 can only increase (decrease) the value of \texttt{chunkSize} and any specified value during the execution
-session, needs to be greater (smaller) than the previous value of the \texttt{chunkSize}. Any request that wants to
+session, needs to be greater (smaller) than the previous value of \texttt{chunkSize}. Any request that wants to
 expand (shrink) a chunk needs to specify a max size (min size). The value of \texttt{chunkSize} can not be set higher
@@
 
-An application can load any chunk with a valid prefix identifier even if that chunk does not exist. For a non-existent
+An application may load any chunk with a valid prefix identifier even if that chunk does not exist. For a non-existent
 chunk the value of \texttt{chunkSize} is always zero.
@@
 \begin{itemize}
-    \item \texttt{appID} is a prefix code built by \emph{applications} prefix tree. An \texttt{appID} cannot
+    \item \texttt{appID} is a prefix code built by the \emph{applications} prefix tree. An \texttt{appID} cannot
     be \texttt{0x0}.
 
-    \item \texttt{accountID} is a prefix code built by \emph{accounts} prefix tree. An \texttt{accountID} cannot
+    \item \texttt{accountID} is a prefix code built by the \emph{accounts} prefix tree. An \texttt{accountID} cannot
     be \texttt{0x0} or \texttt{0x1}.
@@
     \item \texttt{chunkID} is a composite prefix code built by concatenating an \texttt{applicationID} to
-    an \texttt{accountID} to a prefix code made by \emph{local} prefix tree:
+    an \texttt{accountID} to a prefix code made by the \emph{local} prefix tree:
     \subitem \texttt{chunkID = (applicationID|accountID|<local-prefix-code>)} .
@@
 we use Algorithm~\ref{alg:match_id} to match the prefixed identifier with the start of the sequence. The matching
-can be done with only three comparisons, and an invalid prefixed identifier can be detected and will not match
+can be done with only three comparisons, while invalid identifiers can be detected and will not match
 any sequence.
@@
 The attachment of a request is a list of request identifiers of the current block that are \emph{attached} to the
-request. That means, for validating the request a validator first needs to \emph{inject} the digest of attachments
-into the current HTTP request text. By doing so, the called application will have access to the digest of
-attachments in a secure and authenticated way.
+request. That means, for validating this request a validator first needs to \emph{inject} the digest of attachments
+into its HTTP request text. By doing so, the called application will have access to the digest of
+attachments in a secure way, while it is ensured that the attached requests are included in the
+current block.
 
@@
 declares those requests as its attachments. For paying fees the payer signs the digest of requests for which he
-wants to pay fees. After injecting the digest of those request by validators, that signature can be validated
-correctly and securely by the application that handles fee payment.
+wants to pay fees. After injecting the digest of those request by validators, that signature can be validated by the
+application that handles fee payment, and it is guaranteed that the attached requests are actually included in the
+current block.
 

### 2023-01-11 tex/sections/blocks.tex
Improve readability of "Block Validation" SubSection
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 to be executed by the Argennon Smart Contract Execution Environment (AscEE). The first block of the blockchain, the
-\emph{genesis} block, is a spacial block that fully describes the initial state of the AscEE. Every block of the
+\emph{genesis} block, is a spacial block that fully describes the initial state of the AscEE and every block of the
 Argennon blockchain thus corresponds to a unique AscEE state which can be calculated deterministically from the genesis
@@
         the block with height $h - k$: $v \mhyphen cert_{h-k}$         \\ [1.2ex]
-        certificate of the delegates for\\
-        the previous block: $d \mhyphen cert_{h-1}$        \\ [1.2ex]
         previous block hash                           \\ [1.2ex]
@@
 \end{center}
-
-\subsection{Block Validation}\label{subsec:block-validation}
-
-To validate a block three main conditions need to be validated: (i) commitments to the program and heap
-database are resulted from applying the request list and clustering directives of the block to the previous AscEE
-state, (ii) $v \mhyphen cert_{h-k}$ is valid, (iii) previous block is valid and has height $h - 1$.
-
-We can denote condition (i) by a Computational Integrity (CI) statement:
-\begin{equation}
-    \label{eq:ci}
-    \mathbf{C}_{P_h},\mathbf{C}_{H_h} \coloneqq \tau(\mathbf{C}_{P_{h-1}},\mathbf{C}_{H_{h-1}},\mathbf{C}_{R_{h}},
-    dir_h)\ ,
-\end{equation}
-where $\tau$ is a transition function that encodes the AscEE computation logic and the necessary preprocessing and
-postprocessing steps\footnote{including opening and updating the state commitments}.
-
-Verifying statement~\ref{eq:ci} can be done by either replaying the AscEE computation and performing the required
-preprocessing and postprocessing steps, or alternatively by
-verifying a computational integrity proof\footnote{More accurately we should call it an argument instead of a proof.
-However, using the word argument can confuse a reader who is not familiar with the subject, so we avoid it here.}.
-To use computational integrity proofs the verifier and prover need to
-share an arithmetized representation of $\tau$ and use it for both proof generation and verification.
-
-The Argennon Prover Machine provides a convenient and universal way for arithmetization of any Argennon
-application. Moreover, since a compiled version of every Argennon application to the APM code is stored in the AscEE
-program area, validators do not need to locally store the APM code of applications.
-
-The Argennon protocol does not enforce the usage of the APM. Validators and PVC servers can use any argument system
-with any arithmetization, and if required, the ASAR of an application can be used
-for generating the appropriate arithmetization instead of the APM\@.
-
-The certificate of validators, $v \mhyphen cert_{h-k}$, is an aggregate signature. Validating it requires accessing the
-AscEE heap state at block $h-k-m$ to obtain public keys and stake values. Here $m$ is the total number of validator
-assemblies and $k$ is the maximum allowed length of
-the unvalidated chain. (See Section\ref{subsec:validators-committee})
-Again, in addition to direct verification, a validator can use computational integrity proofs, received from the
-Argennon cloud, to perform cheaper verification of this certificate.
-
-For validating the previous block, instead of directly validating the content of the block, a validator only
-verifies the block certificates. Verifying $d \mhyphen cert_{h-1}$ is straightforward. Verifying $v \mhyphen
-cert_{h-1}$, which should be performed based on stakes at block $h-m-1$, can be cheaply done by using computational
-proofs, obtained from the Argennon cloud.
-
-This type of block validation only validates the transition from block $h-1$ to block $h$, and the block is valid
-only if its previous block is valid. We
-call this type of block validation \emph{conditional block validation}, since the validity of the current block is
-conditioned on the validity of the previous block.
-
-Interestingly, conditional block validation of multiple blocks can be done in parallel. Moreover, the required proofs
-can be generated independently and by different PVC servers. As we will see in
-Section~\ref{subsec:validators-committee}, this property plays an important role in the Argennon consensus protocol.
-
-In summary, a validator validates a block by verifying three computational integrity proofs: $\pi_{\tau}$ which proves
-the transition is correct, $\pi_{h-k}$ which proves the validity of the included certificate of block $h-k$ and
-$\pi_{h-1}$ which proves the previous block has a correct certificate of validators. It should be noted that these
-proofs are not a part of the block contents and therefore the exact argument system used for generating them is not
-specified by the Argennon protocol.
 
@@
     The signature $\sigma$ is a \emph{single} group element.
-    \item $\textbf{Verify}(\PK,m,\sigma)$: if $e(g_1, \sigma) = e\big(\PK,\ \Hm(m)\big)$  then output "accept",
-    otherwise output "reject".
+    \item $\textbf{Verify}(\PK,m,\sigma)$: if $e(g_1, \sigma) = e\big(\PK,\ \Hm(m)\big)$  then output ``accept'',
+    otherwise output ``reject''.
 \end{itemize}
@@
     \label{eq:agg}
-    \sigma \gets \sigma_1 \cdots \sigma_n \in \G_0\ .
+    \sigma \gets \sigma_1 \cdots \sigma_n \in \G_0\ \cdot
 \end{equation}
@@
     \label{eq:aggdiff}
-    e(g_1, \sigma) = e\big(\PK_1,\ \Hm(m_1)\big) \cdots e\big(\PK_n,\ \Hm(m_n)\big)\ .
+    e(g_1, \sigma) = e\big(\PK_1,\ \Hm(m_1)\big) \cdots e\big(\PK_n,\ \Hm(m_n)\big)\ \cdot
 \end{equation}
@@
     \label{eq:aggsame}
-    e(g_1, \sigma) = e\Big(\PK_1 \cdots \PK_n,\ \Hm(m)\Big)\ .
+    e(g_1, \sigma) = e\Big(\PK_1 \cdots \PK_n,\ \Hm(m)\Big)\ \cdot
 \end{equation}
@@
 group is commutative, we have $(ab)^{-1}=a^{-1}b^{-1}$.}
+
+\subsection{Block Validation}\label{subsec:block-validation}
+
+To validate a block three main conditions need to be validated: (i) commitments to the program and heap
+database are resulted from applying the request list and clustering directives of the block to the previous AscEE
+state, (ii) previous block is valid and has height $h - 1$, (iii) $v \mhyphen cert_{h-k}$ is valid.
+
+We can denote condition (i) by a Computational Integrity (CI) statement:
+\begin{equation}
+    \label{eq:ci}
+    \mathbf{C}_{P_h},\mathbf{C}_{H_h} \coloneqq \tau(\mathbf{C}_{P_{h-1}},\mathbf{C}_{H_{h-1}},\mathbf{C}_{R_{h}},
+    dir_h)\ ,
+\end{equation}
+where $\tau$ is a transition function that encodes the AscEE computation logic and the necessary preprocessing and
+postprocessing steps\footnote{including opening and updating the state commitments}.
+
+Verifying statement~\ref{eq:ci} can be done by either replaying the AscEE computation and performing the required
+preprocessing and postprocessing steps, or alternatively by
+verifying a computational integrity proof\footnote{More accurately we should call it an argument instead of a proof.
+However, using the word argument can confuse a reader who is not familiar with the subject, so we avoid it here.}.
+To use computational integrity proofs the verifier and prover need to
+share an arithmetized representation of $\tau$ and use it for both proof generation and verification.
+
+The Argennon Prover Machine provides a convenient and universal way for arithmetization of any Argennon
+application. Moreover, since a compiled version of every Argennon application to the APM code is stored in the AscEE
+program area, validators that use the APM do not need to locally store the APM code of applications.
+
+The Argennon protocol does not enforce the usage of the APM. Validators and PVC servers can use any argument system
+with any arithmetization, and if required, the ASAR of an application can be used
+for generating the appropriate arithmetization instead of the APM arithmetization.
+
+For validating the previous block, instead of directly validating the contents of the block, a validator only
+verifies the block certificates. Each block of the Argennon blockchain has two certificates: the certificate of
+validators, $v \mhyphen cert$, and the certificate of the delegates, $d \mhyphen cert$. Verifying $d \mhyphen
+cert_{h-1}$ is straightforward but verifying $v \mhyphen cert_{h-1}$ is more challenging. $v \mhyphen cert_{h-1}$ is an
+aggregate signature and validating it requires accessing the AscEE heap state at block $h-1-m$, where $m$ is the
+total number of validator assemblies, to obtain public keys and stake values. Again, in addition to direct
+verification, a validator can use computational integrity proofs, received from the Argennon cloud, to perform
+cheaper verification of this certificate.
+
+The block at height $h$ includes a certificate of validators for block $h-k$ which is used to record the participation
+of validators and facilitate reward calculation. This certificate needs to be validated based on stake and public key
+database at block $h-k-m$, and can be cheaply done by using computational proofs, obtained from the Argennon cloud.
+Here $k$ is the maximum allowed length of the unvalidated part of the Argennon blockchain. (See
+Section~\ref{subsec:validators-committee})
+
+This type of block validation only validates the transition from block $h-1$ to block $h$, and the block is valid
+only if its previous block is valid. We
+call this type of block validation \emph{conditional block validation}, since the validity of the current block is
+conditioned on the validity of the previous block.
+
+Interestingly, conditional block validation of multiple blocks can be done in parallel. Moreover, the required proofs
+can be generated independently and by different PVC servers. As we will see in
+Section~\ref{subsec:validators-committee}, this property plays an important role in the Argennon consensus protocol.
+
+In summary, a validator validates a block by verifying three computational integrity proofs: $\pi_{\tau}$ which proves
+the transition is correct, $\pi_{h-1}$ which proves the previous block has a correct certificate of validators and
+$\pi_{h-k}$ which proves the validity of the included certificate of validators of block $h-k$. It should be noted that
+these proofs are not part of the block contents and therefore the exact argument system used for generating them is
+not specified by the Argennon protocol.

### 2023-01-11 tex/sections/concurrency.tex
Improve readability of "Block Validation" SubSection
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 effect of its contained transactions must be applied to the AscEE state sequentially as they appear in the ordered
-list. This ordering is solely chosen by the block proposer, and users should not have any assumptions about
-the ordering of transactions in a block.
+list.\footnote{This ordering is solely chosen by the block proposer, and users should not have any assumptions about
+the ordering of transactions in a block.}
 
-The fact that block transactions constitute a sequential list, does not mean they can not be executed and applied
+The fact that block transactions constitute a sequential list, does not imply they can not be executed and applied
 to the AscEE state concurrently. Many transactions are actually independent and the order of their execution does not

### 2023-01-12 tex/sections/blocks.tex
Fix readability issues of Subsection "Consensus"
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 verifies the block certificates. Each block of the Argennon blockchain has two certificates: the certificate of
-validators, $v \mhyphen cert$, and the certificate of the delegates, $d \mhyphen cert$. Verifying $d \mhyphen
+validators, $v \mhyphen cert$, and the certificate of delegates, $d \mhyphen cert$. Verifying $d \mhyphen
 cert_{h-1}$ is straightforward but verifying $v \mhyphen cert_{h-1}$ is more challenging. $v \mhyphen cert_{h-1}$ is an
-aggregate signature and validating it requires accessing the AscEE heap state at block $h-1-m$, where $m$ is the
+aggregate signature and validating it requires accessing the AscEE heap state at block~$h-1-m$, where $m$ is the
 total number of validator assemblies, to obtain public keys and stake values. Again, in addition to direct
@@
 
-The block at height $h$ includes a certificate of validators for block $h-k$ which is used to record the participation
+The block at height~$h$ includes a certificate of validators for block~$h-k$ which is used to record the participation
 of validators and facilitate reward calculation. This certificate needs to be validated based on stake and public key
-database at block $h-k-m$, and can be cheaply done by using computational proofs, obtained from the Argennon cloud.
+database at block~$h-k-m$, and can be cheaply done by using computational proofs, obtained from the Argennon cloud.
 Here $k$ is the maximum allowed length of the unvalidated part of the Argennon blockchain. (See
@@
 
-This type of block validation only validates the transition from block $h-1$ to block $h$, and the block is valid
+This type of block validation only validates the transition from block~$h-1$ to block~$h$, and the block is valid
 only if its previous block is valid. We
@@
 the transition is correct, $\pi_{h-1}$ which proves the previous block has a correct certificate of validators and
-$\pi_{h-k}$ which proves the validity of the included certificate of validators of block $h-k$. It should be noted that
+$\pi_{h-k}$ which proves the validity of the included certificate of validators of block~$h-k$. It should be noted that
 these proofs are not part of the block contents and therefore the exact argument system used for generating them is

### 2023-01-12 tex/sections/consensus.tex
Fix readability issues of Subsection "Consensus"
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

--- 2022-09-24 tex/sections/consensus.tex
@@
 the assembly of validators. A block is considered final after its \textbf{next} block receives \textbf{both} of
-its certificates. In Argennon as long as more than half of the total stake of validators is controlled by honest users,
-the probability of discarding a final block is near zero even if all the delegates are malicious.
+its certificates. In Argennon as long as more than $2/3$ of the total stake of validators is controlled by honest users,
+the probability of discarding a final block is zero even if all the delegates are malicious.
 
-In addition to primary committees, Argennon has several community driven committees. Certificates of these
+In addition to primary committees, Argennon could have several community driven committees. Certificates of these
 committees are not required for block finality, but they could be used by members of validator assemblies to better
@@
 At the start of the Argennon main-net, this committee will be elected for one-year terms and will have five members.
-Later, this could be changed by the ADAGs in a procedure described in Section~\ref{sec:adags}.
+Later, this can be changed by the ADAGs in a procedure described in Section~\ref{sec:adags}.
 
@@
 
-Besides the main committee, a reserve committee of delegates consisting of three members is elected by users
-either through the ADAGs or by \emph{emergency agreement} during the recovery protocol. In case the main committee
+Besides the main committee, a reserve committee of delegates consisting of three members is elected by validators
+either through the ADAGs or by the \emph{emergency agreement} during the recovery protocol. In case the main committee
 fails to generate new blocks or behaves maliciously, the task of
@@
 proposed block. For better performance, the delegates should run their agreement protocol for reaching consensus about
-small batches of transactions in their mem-pools, instead of the whole next block.
+small batches of transactions in their mem-pools, instead of the whole block.
 
-If one of the delegates loses its network connectivity, no new blocks can be generated. For this reason,
-the delegates should invest on different types of communication infrastructure, to make sure they will never lose
-connectivity to each other and to the Argennon network.
+If one of the delegates loses its network connectivity, no new blocks can be generated until the reserve committee
+gets activated. For this reason, the delegates should invest on different types of communication infrastructure, to
+make sure they will never lose connectivity to each other and to the Argennon network.
 
@@
 The \texttt{minValidatorStake}
-threshold is determined by the ADAGs, but it can never be higher than $1000$ ARGs.
+threshold is determined by the ADAGs, but it can never be higher than $1000$ ARG\@.
 
@@
 Every validator has a status which can be either \texttt{online} or \texttt{offline}.
-This status is stored in the ARG smart contract and is part of the staking database. A validator can change
+This status is stored in the ARG application and is part of the staking database. A validator can change
 his status to \texttt{offline} through an external request (transaction) to the ARG application. In this request he
@@
 
-There is no transaction type for changing the status of a validator to \texttt{online}. A malicious delegates committee
-would be able to censor this type of transactions and prevent honest validators from coming back online. For this reason
-the status of a validator is considered online automatically when the specified period of time for being offline ends.
+There is no transaction type for changing the status of a validator to \texttt{online}. A malicious committee
+of delegates would be able to censor this type of transactions and prevent honest validators from coming back online.
+For this reason the status of a validator is considered online automatically when the specified period of time for
+being offline ends.
 
@@
 of the certificate signers is higher than $2/3$ of the total stake of validators and \texttt{online/offline}
-status of validators becomes ineffective.
+status of validators becomes ineffective. This prevents the system from going into temporary deadlocks and the
+community will always be able to preserve the liveliness without waiting for the expiration of offline status of
+some accounts.
 
@@
 do not wait for seeing the validators certificate of the previous block to start block validation. On the
-other hand, the block certificates are published and broadcast sequentially. A validator does not publish his
-signature, if the certificate of the previous block validator assembly has not been published yet. This ensures that
-an invalid fork made by malicious delegates will not receive any certificates from any validator assemblies.
+other hand, the block certificates are published and broadcast sequentially. A validator does not publish his vote,
+if the certificate of the validator assembly of the previous block has not been published yet. This ensures
+that an invalid fork made by malicious delegates can not receive any valid certificates from any validator assemblies.
 
@@
 blockchain. In these attack scenarios, we assume that a single validator assembly is corrupted, all the delegates
-are malicious and the adversary is able to fully control message transmission and partition the network arbitrarily.
+are malicious and the adversary is able to fully control message transmission between nodes and partition the network
+arbitrarily.
 
@@
 \begin{itemize}
-    \item Confirming an invalid block:
+    \item Certifying an invalid block:
     \[ m > r(s-d) \]
@@
 of an assembly. For forking the blockchain, interestingly $m$ is
-minimized when $h=0$, and $\frac{1}{3}s$ is the minimum required stake. For halting the blockchain,
+minimized when $h=0$, thus the minimum required stake is $\frac{1}{3}s$. For halting the blockchain,
 an adversary requires a stake higher than $\frac{1}{6}s-h$.
@@
 
-To compare the minimum stake required for halting the blockchain in two protocols, we can solve the following
-inequality:
+We observe that the minimum required stake for halting the blockchain in the Argennon protocol will be higher, as
+long as the
+following inequality holds:
 \[
@@
 \]
-We observe that as long as $d>\frac{1}{9}s$, the minimum required stake for halting the blockchain is higher in the
+So if $d>\frac{1}{9}s$, the minimum required stake for halting the blockchain is higher in the
 Argennon protocol and the value of $h$ does not matter.

### 2023-01-13 tex/sections/blocks.tex
Add more explanation for "Emergency Forking"
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
     \item \textbf{Associative aggregation:} the signature aggregation operator is associative.
-    \item \textbf{Efficient cancellation:} if $S$ is the predefined set of users that we need their aggregate
-    signature, verifying an aggregate signature of $S-U$ can be done in time $O(T+|U|)$, if the aggregate
+    \item \textbf{Efficient cancellation:} if $S$ is a predefined and fixed set of users and $U$ is an arbitrary subset
+    of~$S$, verifying an aggregate signature of $S-U$ can be done in time $O(T+|U|)$, if the aggregate
     signature of $S$ can be verified in $O(T)$.
@@
 cert_{h-1}$ is straightforward but verifying $v \mhyphen cert_{h-1}$ is more challenging. $v \mhyphen cert_{h-1}$ is an
-aggregate signature and validating it requires accessing the AscEE heap state at block~$h-1-m$, where $m$ is the
+aggregate signature and validating it requires accessing the staking database at block~$h-1-m$, where $m$ is the
 total number of validator assemblies, to obtain public keys and stake values. Again, in addition to direct

### 2023-01-13 tex/sections/consensus.tex
Add more explanation for "Emergency Forking"
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 the probability of discarding a final block is zero even if all the delegates are malicious.
+
+Having multiple assemblies gives validators some ``resting'' time and allows a validator to be out of sync with the
+network for some time, without losing the opportunity to vote for any blocks.
 
@@
 validity of the block (See Section~\ref{subsec:block-validation}), and if the block is valid, he issues
-an ``accept'' signature. If the block is invalid, he initiates the recovery protocol. The validator will broadcast the
-certificate \textbf{only after} he sees the certificate of the validator assembly of the previous block.
+an ``accept'' signature. If the block is invalid, he initiates the recovery protocol. The validator will broadcast
+the signature \textbf{only after} he sees the certificate of the validator assembly of the previous block.
 Some validators may also require seeing a certificate from
@@
 controlled by honest users. The recovery protocol uses two main emergency procedures to
-recover the functionality of the Argennon blockchain: \emph{emergency forking} and \emph{emergency
+recover the functionality of the Argennon blockchain: the \emph{emergency forking} and \emph{emergency
 agreement} protocol.
@@
 fork block gets certified by validators, its parent is also confirmed and will become a part of the blockchain, even if
-it does not have a validators certificate.
+it does not have a certificate of validators.
 
@@
 block with a validator certificate and $h_v+k$ be the height of the last valid block that the reserve committee has
-seen. For forking the main chain, the reserve committee generates all fork blocks with heights $h_v+1,h_v+2,\dots,
-h_v+k+1$. The parent of the fork block with height $h_v+i$ will be the block $h_v+i-1$ of the main chain. The
-reserve committee will wait until the fork block with height $h_v+k+1$ receives a certificate from validators and
-then will continue the normal chain after that fork block. Hence, the fork block with height $h_v+k+1$
+seen. For forking the main chain, the reserve committee generates all fork blocks with heights~$h_v+1,h_v+2,\dots,
+h_v+k+1$. The parent of the fork block with height $h_v+i$ will be the block~$h_v+i-1$ of the main chain. The
+reserve committee will wait until the fork block with height~$h_v+k+1$ receives a certificate from validators and
+then will continue the normal chain after that fork block. Hence, the fork block with height~$h_v+k+1$
 will be the parent of the first normal block generated by the reserve committee.
@@
 If $h_{\max}$ denotes the height of the highest block with a validators certificate, as long as more than $2/3$ stake
-of every assembly of validators are honest, for a fork block with height $h_f$ we have:\footnote{This fork block
+of every assembly of validators is honest, for a fork block with height $h_f$ we have:\footnote{This fork block
 forks the blockchain at height $h_f-1$.}
@@
     \item if $h_f \leq h_{\max}$, the fork block can not receive a certificate from validators.
-    \item if $h_f = h_{\max} + 1$, the fork block may receive a certificate from validators or not. It is possible that
+    \item if $h_f = h_{\max} + 1$, when there is no main block with height~$h_f$, the fork block can always receive a
+    certificate from validators, otherwise the fork block may receive a certificate or not. It is possible that
     the validators of the assembly with index $h_f$ modulo $m$ get divided between the fork block and a block at
@@
 
-Consequently, if two fork blocks at heights $h_0$ and $h_0+k$ are generated by the reserve committee, if both blocks
-receive a certified from validators, we must have $h_0 > h_{\max}$, and there must exist fork blocks with heights
+The reserve committee does not know the value of $h_{\max}$. However, if the reserve committee creates all fork blocks
+with heights~$h_0,h_0+1,h_0+2,\dots$ every fork block with height $h \geq h_{\max} + 2$ will surely get a certificate.
+If there is no block in the main chain with height~$h_{\max} + 1$, then the fork block with height~$h_{\max} +
+1$ will also get a certificate, otherwise the reserve committee should be able to eventually find the main block
+at $h_{\max}+1$ and generate a fork block with height~$h_{\max} + 2$ as its successor. This way, the reserve
+committee will always be able to create at least one fork block which gets certified and continue the chain after the
+fork block with the highest height.
+
+If two fork blocks at heights $h_0$ and $h_0+k$ are generated by the reserve committee and both blocks receive a
+certified from validators, we must have $h_0 > h_{\max}$, and there must exist fork blocks with heights
 $h_0+1,\dots,h_0+k-1$ which are certified by validators. That means if a malicious reserve committee generates a normal
@@
 
+If a malicious reserve committee creates two fork blocks with the same height, either only one of
+them can get a certificate, which makes the other one ineffective or validators get split between blocks
+and the chain will halt. In this case the emergency agreement protocol will start.
+
 During the emergency forking, at max one block with a validators certificate may be discarded. This happens when
@@
 $h$; the block $b_1$ has received a validators certificate and validators have not certified any blocks at height $h+1$.
-The reserve committee adds a fork block whose parent is block $b_2$ and that will essentially discard $b_1$. Notice
-that if a block at height $h+1$ had been certified by validators the fork block could not have received a certificate
-from validators.
+The reserve committee adds a fork block whose parent is block $b_2$, that will essentially discard $b_1$. Notice
+that if a block at height $h+1$ had been certified by validators the fork block after $b_2$ could not have received a
+certificate from validators.
 

### 2023-01-13 tex/sections/consensus.tex
Improve readability of "Emergency Agreement"
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 A honest user during a round only votes a single proposal. This ensures that as long as more than $2/3$ of the total
-stake of the system is controlled by honest users, no two different proposals can get more than $2/3$ votes. Also, he
-either votes $p$-win or $draw$, so only one of $p$-win or $draw$ can get more than $2/3$ votes. Consequently, when a
+stake of the system is controlled by honest users, no two different proposals can get more than $2/3$ votes.
+As a result, we can not have 2/3 votes for both $p$-lock and $p'$-lock if $p \neq p'$. Also, a honest user
+either votes $p$-win or $draw$, so only one of $p$-win or $draw$ can get more than $2/3$ votes and when a
 user sees more than $2/3$ $p$-win votes, he can be sure that $draw$ has less than $2/3$ votes. Therefore, for going
@@
 
-When a single honest user terminates the protocol, he can convince all other honest users to terminate their protocol
-by sending those 2/3 $p$-win votes that he has seen.
+Note that when a single honest user terminates the protocol, he can convince all other honest users to terminate
+their protocol by sending those 2/3 $p$-win votes that he has seen.
 

### 2023-01-13 tex/sections/intro.tex
Improve readability of the first part of "intro"
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 At the time of writing of this document,
-the total mining reward for a Bitcoin block is around \$150,000. If we assume, in decision theoretic terminology,
-that the mining reward accurately defines the utility function of a Bitcoin miner, one could hire all hashing power
+the total mining reward for a Bitcoin block is around \$150,000, and the Bitcoin network approximately produces 5 blocks
+per hour. If we assume, in decision theoretic terminology, that the mining reward solely defines the utility
+function\footnote{In decision theory, the utility function measures the preference, worth or value of different
+alternatives for a decision maker.} of a Bitcoin miner, one should be able to hire all hashing power
 of the Bitcoin network for one hour by
 spending only \$750,000. The situation is not much different for PoS blockchains, as long as the total
-stake of the validator set is a relatively small value. By
-stake, we mean a real number measuring the total interest of a user in the system, we are not referring, in
-particular, to some locked amount of a user's money that is known as stake in some PoS protocols.
-This problem is more severe in blockchains that use randomly selected small sets of validators. These small sets
+stake\footnote{Here by stake, we mean a real number measuring the total interest of a user in the system, and we are
+not referring, in particular, to some locked amount of a user's money that is known as stake in some PoS protocols.}
+of the validator set is a relatively small value.
+This problem is even more severe in blockchains that use randomly selected small sets of validators. These small sets
 usually have low total stake and could be easily bribed and corrupted. Selecting these random sets
-by hidden random procedures would not help, since the validator himself knows he has been selected, before casting
+by hidden random procedures would not help, since the validator himself knows he has been selected before casting
 his vote.
@@
 smart contract platform. This computational and storage overhead, in practice, could prevent most of
-ordinary users from any type of participation in the process of securing a blockchain.
+ordinary users from any type of participation in the consensus protocol of a blockchain.
 
-Although a fully decentralized blockchain based on the participation of every user looks appealing, it is not as perfect
+A fully decentralized blockchain based on the participation of every user looks appealing, though it is not as perfect
 as it might seem. The consensus protocol of a blockchain relies on a network of computers, not humans. Ordinary users

### 2023-01-16 tex/sections/consensus.tex
Fix a readability issue
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 The reserve committee of delegates is able to fork the Argennon blockchain, if it receives a valid fork request
-from the validators.
-This fork needs to be confirmed by validators and can not discard more than one block that has been certified by
-validators.
+from validators.
 A valid fork request is an unexpired request signed by more than half of the total stake of validators.
+
+A fork created by the reserve committee needs to be confirmed by validators and can never discard more than one
+block which has received a certificate from validators.
 
@@
 
-The emergency agreement protocol is a resilient protocol for deciding between a set of proposals when no committee of
-delegates can be trusted. For initiating the protocol, a validator signs a message containing the subject
-of the agreement and a start time.
+The emergency agreement protocol is a resilient protocol for deciding between a set of proposals when the committee of
+delegates is not available or can not be trusted. For initiating the protocol, a validator signs a message containing
+the subject of the agreement and a start time.
 

### 2023-01-19 tex/sections/consensus.tex
Clarify the analysis of emergency forking
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
+\subsubsection{Signing the Block Certificate}
+
+The delegates can generate blocks very fast. Consequently, the Argennon blockchain always has an
+unvalidated part which contains the blocks that have a certificate from the committee of delegates but have not
+yet received a certificate from the validators.
+
+As we mentioned before, the block with height $h$ needs a certificate from the assembly of
+validators with index $h$ modulo $m$. To decide about signing the certificate of a block which already has
+a certificate from the delegates, a validator checks the conditional
+validity of the block (See Section~\ref{subsec:block-validation}), and if the block is valid, he issues
+an ``accept'' signature. If the block is invalid, he initiates the recovery protocol. The validator will broadcast
+the signature \textbf{only after} he sees the certificate of the validator assembly of the previous block.
+Some validators may also require seeing a certificate from
+some community based committees. An honest validator never signs a certificate for two different blocks with the
+same height.
+
+Consequently, in Argennon the block validation by assemblies is performed in parallel, and validators
+do not wait for seeing the validators certificate of the previous block to start block validation. On the
+other hand, the block certificates are published and broadcast sequentially. A validator does not publish his vote,
+if the certificate of the validator assembly of the previous block has not been published yet. This ensures
+that an invalid fork made by malicious delegates can not receive any valid certificates from any validator assemblies.
+
+\subsubsection{Signature Aggregation}\label{subsubsec:sig-agg}
+
+The validators certificate of a block is an aggregate signature of members of the corresponding assembly of
+validators. Validator assemblies could include millions of users and calculating their aggregate
+signature requires an efficient distributed algorithm.
+
+In Argennon, signature aggregation is mostly performed by PVC servers. To distribute the aggregation workload
+between different servers, every validator assembly is divided into pre-determined groups, and each PVC
+server is responsible for signature aggregation of one group. To make sure that there is enough redundancy, the
+total number of groups should be less than the number of PVC servers and each group should be assigned to
+multiple PVC servers.
+
+Any member of a group knows all the servers that are responsible for signature aggregation of his group. When a member
+signs a block certificate, he sends his signature to all the severs that aggregate the signatures of his group.
+These PVC servers aggregate the signatures they receive and then send the aggregated signature to the delegates.
+Furthermore, the delegates aggregate these signatures to produce the final block certificate
+and then broadcast it to the PVC servers network.
+
+The role of the delegates in the signature aggregation algorithm is limited. The important part of the work is done by
+PVC servers and slightly modified versions of this algorithm can perform signature aggregation
+even if all the delegates are malicious, as long as there are enough honest PVC servers.
+
 \subsubsection{Activity Status}
@@
 
-\subsubsection{Signing the Block Certificate}
-
-The delegates can generate blocks very fast. Consequently, the Argennon blockchain always has an
-unvalidated part which contains the blocks that have a certificate from the committee of delegates but have not
-yet received a certificate from the validators.
-
-As we mentioned before, the block with height $h$ needs a certificate from the assembly of
-validators with index $h$ modulo $m$. To decide about signing the certificate of a block which already has
-a certificate from the delegates, a validator checks the conditional
-validity of the block (See Section~\ref{subsec:block-validation}), and if the block is valid, he issues
-an ``accept'' signature. If the block is invalid, he initiates the recovery protocol. The validator will broadcast
-the signature \textbf{only after} he sees the certificate of the validator assembly of the previous block.
-Some validators may also require seeing a certificate from
-some community based committees. An honest validator never signs a certificate for two different blocks with the
-same height.
-
-Consequently, in Argennon the block validation by assemblies is performed in parallel, and validators
-do not wait for seeing the validators certificate of the previous block to start block validation. On the
-other hand, the block certificates are published and broadcast sequentially. A validator does not publish his vote,
-if the certificate of the validator assembly of the previous block has not been published yet. This ensures
-that an invalid fork made by malicious delegates can not receive any valid certificates from any validator assemblies.
-
 \subsubsection{Analysis}
@@
 Argennon protocol and the value of $h$ does not matter.
-
-\subsubsection{Signature Aggregation}\label{subsubsec:sig-agg}
-
-The validators certificate of a block is an aggregate signature of members of the corresponding assembly of
-validators. Validator assemblies could include millions of users and calculating their aggregate
-signature requires an efficient distributed algorithm.
-
-In Argennon, signature aggregation is mostly performed by PVC servers. To distribute the aggregation workload
-between different servers, every validator assembly is divided into pre-determined groups, and each PVC
-server is responsible for signature aggregation of one group. To make sure that there is enough redundancy, the
-total number of groups should be less than the number of PVC servers and each group should be assigned to
-multiple PVC servers.
-
-Any member of a group knows all the servers that are responsible for signature aggregation of his group. When a member
-signs a block certificate, he sends his signature to all the severs that aggregate the signatures of his group.
-These PVC servers aggregate the signatures they receive and then send the aggregated signature to the delegates.
-Furthermore, the delegates aggregate these signatures to produce the final block certificate
-and then broadcast it to the PVC servers network.
-
-The role of the delegates in the signature aggregation algorithm is limited. The important part of the work is done by
-PVC servers and slightly modified versions of this algorithm can perform signature aggregation
-even if all the delegates are malicious, as long as there are enough honest PVC servers.
 
@@
 
-The reserve committee does not know the value of $h_{\max}$. However, if the reserve committee creates all fork blocks
-with heights~$h_0,h_0+1,h_0+2,\dots$ every fork block with height $h \geq h_{\max} + 2$ will surely get a certificate.
-If there is no block in the main chain with height~$h_{\max} + 1$, then the fork block with height~$h_{\max} +
-1$ will also get a certificate, otherwise the reserve committee should be able to eventually find the main block
-at $h_{\max}+1$ and generate a fork block with height~$h_{\max} + 2$ as its successor. This way, the reserve
-committee will always be able to create at least one fork block which gets certified and continue the chain after the
-fork block with the highest height.
+The reserve committee must be able to create at least one fork block which gets a certificate from validators.
+The reserve committee may not know the value of $h_{\max}$.\footnote{This could happen when the network is
+partitioned.} However, if the reserve committee creates all fork blocks
+with heights~$h_0,h_0+1,h_0+2,\dots$ for some $h_0 \leq h_{\max} + 2$,
+then every fork block with height~$h \geq h_{\max} +
+2$ will surely get a certificate. Obviously $h_v \leq h_{\max}$, so if the main chain contains a block with height
+$h_{\max} + 1$ the reserve committee should be able to eventually find this block and generate the fork block with
+height $h_{\max} + 2$, which will surely get a certificate.
+If there is no block in the main chain with height~$h_{\max} + 1$, then the fork block with
+height~$h_{\max} + 1$ can get a certificate. This way, the reserve committee will always be able to continue the chain
+after the certified fork block with the highest height.
 
@@
 
-During the emergency forking, at max one block with a validators certificate may be discarded. This happens when
-the malicious main committee of delegates has forked the main chain by producing blocks $b_1$ and $b_2$ at height
+During the emergency forking, no more than one block with a validators certificate can be discarded. A fork block
+with height~$h+1$ can not receive a certificate without a certified normal parent block with
+height~$h$ or $h-1$, or another certified fork block with height~$h$.
+
+One certified block may be discarded when
+the malicious main committee of delegates has forked the main chain by producing blocks $b_1$ and $b_2$ with heights
 $h$; the block $b_1$ has received a validators certificate and validators have not certified any blocks at height $h+1$.
-The reserve committee adds a fork block whose parent is block $b_2$, that will essentially discard $b_1$. Notice
+The reserve committee adds a fork block whose parent is block $b_2$, and that will essentially discard $b_1$. Notice
 that if a block at height $h+1$ had been certified by validators the fork block after $b_2$ could not have received a

### 2023-02-05 tex/sections/apm.tex
Add a section about simplifying zk components
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 %! Date = 9/2/22
+
+\section{Introduction}\label{sec:introduction3}
 
@@
 
+\section{Simplifying Complex Components}\label{sec:simplifying-complex-components}
+
+Assume that we have a complex component which has a considerable arithmetic circuit complexity. If this component
+is needed in some arbitrary computation steps, we will need to repeat its circuit in every step of the computation. This
+will increase the complexity of proof generation considerably. Fortunately, there is a workaround and instead of
+repeating the circuit of the component, we can use a simpler cryptographic accumulator circuit during the
+computation, and use a final verification circuit at the end to verify the functionality of the component.
+
+Every component accepts an input $in$ and produces an output $out$. As a result we can denote the component by a
+deterministic function that maps an input sequence with any arbitrary length $k$, to an output sequence with the same
+length:
+\[
+    f((in_1,in_2,\dots,in_k)) \coloneqq (out_1,out_2,\dots,out_k)
+\]
+In every step of the computation, we replace this component with a cryptographic accumulator which receives $in$ and
+$out$ as its inputs and gets activated in the same computation steps that the component must be active. When the
+accumulator is active, it accumulates its inputs, so at the end of the computation it has computed a digest:
+\[
+    h(in_1,out_1,in_2,out_2,\dots,in_k,out_k) \coloneqq digest_f
+\]
+where $k$ is the number of steps that our component must have been active.
+
+Now the prover needs to prove that he knows values $in_1,out_1,\dots,in_k,out_k$ such that they are correctly
+produced by the component:
+\[
+    f((in_1,\dots,in_k)) = (out_1,\dots,out_k)
+\]
+and their digest is also correct:
+\[
+    h(in_1,out_1,\dots,in_k,out_k) = digest_f
+\]
+where $f$ and $digest_f$ are known to the prover and verifier.
+
+The circuit for verifying this assessment usually is straight forward. When the computation involves a large
+number of steps and the component is complex, this approach can reduce the cost of proof generation considerably. Memory
+components are good candidates for being simplified by this method.
+
+Interestingly, this approach can reduce the number of computation steps. When we use an accumulator circuit instead
+of the component, $out$ can be available in the same computation step that $in$ is available. This will
+eliminate the computation step that is needed for generating $out$ by the component's circuit.
+
 %

### 2023-02-05 tex/sections/apm.tex
Add `Selected.tex`
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
     P,R$ are valid w.r.t the provided commitments. It also processes the input data and ensures it has the
-    correct format and is valid. For example it verifies the signatures of authenticated messages or
+    correct format and is valid. For example, it verifies the signatures of authenticated messages or
     checks that the proposed chunk bounds are valid.
@@
 
+\pagebreak
+
 \section{Simplifying Complex Components}\label{sec:simplifying-complex-components}
@@
 is needed in some arbitrary computation steps, we will need to repeat its circuit in every step of the computation. This
-will increase the complexity of proof generation considerably. Fortunately, there is a workaround and instead of
+will increase the complexity of proof generation considerably. Fortunately, there is a workaround. Instead of
 repeating the circuit of the component, we can use a simpler cryptographic accumulator circuit during the
@@
 
-Interestingly, this approach can reduce the number of computation steps. When we use an accumulator circuit instead
-of the component, $out$ can be available in the same computation step that $in$ is available. This will
+Interestingly, this approach can reduce the number of computation steps as well. When we use an accumulator circuit
+instead of the component, $out$ can be available in the same computation step that $in$ is available. This will
 eliminate the computation step that is needed for generating $out$ by the component's circuit.

### 2023-02-06 tex/sections/apm.tex
Fix some typos
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-Assume that we have a complex component which has a considerable arithmetic circuit complexity. If this component
+Assume that we have a component which has a considerable arithmetic circuit complexity. If this complex component
 is needed in some arbitrary computation steps, we will need to repeat its circuit in every step of the computation. This
@@
 \]
-where $f$ and $digest_f$ are known to the prover and verifier.
+where $f$ is known to both the prover and verifier.
 
@@
 Interestingly, this approach can reduce the number of computation steps as well. When we use an accumulator circuit
-instead of the component, $out$ can be available in the same computation step that $in$ is available. This will
-eliminate the computation step that is needed for generating $out$ by the component's circuit.
+instead of the component, $out$ will be available in the same computation step that $in$ is available. This will
+eliminate the computation step that is needed for generating $out$ by our component's circuit.
 

### 2023-02-06 tex/sections/apm.tex
Fix a few issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Fix an error
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 will increase the complexity of proof generation considerably. Fortunately, there is a workaround. Instead of
-repeating the circuit of the component, we can use a simpler cryptographic accumulator circuit during the
+repeating the circuit of the component, we can use a simpler cryptographic hash calculator circuit during the
 computation, and use a final verification circuit at the end to verify the functionality of the component.
@@
 \]
-In every step of the computation, we replace this component with a cryptographic accumulator which receives $in$ and
+In every step of the computation, we replace this component with a cryptographic hash calculator which receives $in$ and
 $out$ as its inputs and gets activated in the same computation steps that the component must be active. When the
-accumulator is active, it accumulates its inputs, so at the end of the computation it has computed a digest:
+hash calculator is active, it hashes its inputs, so at the end of the computation it has computed a digest:
 \[
@@
 \]
-where $f$ is known to both the prover and verifier.
+where functions $f$ and $h$, are both known to the prover and verifier.
 
@@
 
-Interestingly, this approach can reduce the number of computation steps as well. When we use an accumulator circuit
+Interestingly, this approach can reduce the number of computation steps as well. When we use a hash calculator circuit
 instead of the component, $out$ will be available in the same computation step that $in$ is available. This will
 eliminate the computation step that is needed for generating $out$ by our component's circuit.
-
 %

### 2023-02-06 tex/sections/apm.tex
Fix an error
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-Interestingly, this approach can reduce the number of computation steps as well. When we use a hash calculator circuit
-instead of the component, $out$ will be available in the same computation step that $in$ is available. This will
-eliminate the computation step that is needed for generating $out$ by our component's circuit.
+Interestingly, this approach can also reduce the number of computation steps. When we use a hash calculator circuit
+instead of our component, $out$ will be available in the same computation step that $in$ is available. This will
+eliminate the computation step that is needed for generating $out$ by the component's circuit.
 %

### 2023-02-07 tex/sections/apm.tex
Clarify section 3.2
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 
-Assume that we have a component which has a considerable arithmetic circuit complexity. If this complex component
-is needed in some arbitrary computation steps, we will need to repeat its circuit in every step of the computation. This
+Assume that we have a computing machine $M$, that accepts an input $in$ and produces an output $out$. We want to
+calculate the output that results from applying $M$ on an initial input for $n$ times. Note that
+$M$ may not be state-less:
+\[
+    M_n(in_1, out_1, out_2,\dots,out_{n-1}) \coloneqq out_n
+\]
+
+We assume that $M$ is composed of a number of subcomponents. Every subcomponent is like a computing machine and accepts
+an input and produces an output. Like $M$, subcomponents are not necessarily state-less. When $M$ receives an input,
+based on the input, a number of subcomponents gets activated, they generate their output from the input, and then
+their outputs are combined to produce the output of $M$.
+
+To build a verifying circuit for this computation, we need to construct a circuit for $M$ and then repeat this circuit
+$n$ times to compute $out_n$ at the final step of the computation.\footnote{Since $M$ is not state-less, we may have
+to feed inputs from steps, say, $i-1,i-2,\dots$ to step $i$.}
+
+Assume that we have a subcomponent that has a considerable arithmetic circuit complexity. Even if this subcomponent is
+active in only a few computation steps, We will still have to repeat its circuit in every step of the computation. This
 will increase the complexity of proof generation considerably. Fortunately, there is a workaround. Instead of
 repeating the circuit of the component, we can use a simpler cryptographic hash calculator circuit during the
-computation, and use a final verification circuit at the end to verify the functionality of the component.
+computation. At the end, to verify the functionality of the component, we use a final verification circuit.
 
-Every component accepts an input $in$ and produces an output $out$. As a result we can denote the component by a
-deterministic function that maps an input sequence with any arbitrary length $k$, to an output sequence with the same
-length:
+Every subcomponent accepts an input $in$ and produces an output $out$. If the component gets activated for $k$ steps, we
+can denote it by a deterministic function that maps an input sequence with length $k$, to an output sequence with the
+same length:\footnote{Note that subcomponents are not active in every step of the computation, so $in_i$ is \textbf{not}
+indicating the input of the component at the $i$th step of the computation.}
 \[
@@
 instead of our component, $out$ will be available in the same computation step that $in$ is available. This will
-eliminate the computation step that is needed for generating $out$ by the component's circuit.
+eliminate one computation step that is needed for generating $out$ by the component's circuit.
 %

### 2023-02-08 tex/sections/apm.tex
Fix minor readability issues
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

@@
 calculate the output that results from applying $M$ on an initial input for $n$ times. Note that
-$M$ may not be state-less:
+$M$ may or may not be state-less:
 \[
@@
 
-Assume that we have a subcomponent that has a considerable arithmetic circuit complexity. Even if this subcomponent is
+Assume that $M$ contains a subcomponent that has a considerable arithmetic circuit complexity.
+Even if this subcomponent is
 active in only a few computation steps, We will still have to repeat its circuit in every step of the computation. This
 will increase the complexity of proof generation considerably. Fortunately, there is a workaround. Instead of
-repeating the circuit of the component, we can use a simpler cryptographic hash calculator circuit during the
-computation. At the end, to verify the functionality of the component, we use a final verification circuit.
+repeating the circuit of the subcomponent, we can use a simpler cryptographic hash calculator circuit during the
+computation. At the end, we use a final verification circuit to verify the functionality of the component.
 
@@
 
-Now the prover needs to prove that he knows values $in_1,out_1,\dots,in_k,out_k$ such that they are correctly
+Now the prover needs to prove that he knows values $in_1,out_1,in_2,out_2,\dots,in_k,out_k$ such that they are correctly
 produced by the component:
